### 2018-10-15 Journal.tex
,

\documentclass[a4paper,12pt]{report}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{cancel}
\usepackage{gensymb}
\usepackage{graphicx}
\usepackage{esint}
\usepackage{mdsymbol}
\usepackage{esvect} 
\usepackage{lipsum}
\usepackage{hhline}
\usepackage{eurosym}
\usepackage{listings}
\usepackage{booktabs}
\usepackage{amssymb}
\usepackage{mathrsfs}
\usepackage{commath}
\usepackage{adjustbox}
\usepackage{booktabs}
\usepackage{array}
\usepackage{lscape}
\usepackage[xspace]{ellipsis}
\usepackage{color}
\usepackage{float}
\usepackage{caption}
\usepackage{afterpage}

\definecolor{mygreen}{rgb}{0,0.6,0}
\definecolor{mygray}{rgb}{0.5,0.5,0.5}
\definecolor{mymauve}{rgb}{0.58,0,0.82}
\begin{document}

\title{Master's Project Journal}
\author{Conor Dooley}
\maketitle
\section*{Initial Research}
Mostly consisted of reading the theses sent to me by Elena/Brian to get an idea of what the "parent" projects involved and why my project was required.
\section*{D/NCO Research}
After 2nd meeting with supervisors the goal was to find out what options there were in terms of implementing an oscillator on an FPGA. Only two real options were phase accumulator \& ring oscillator. Both of these I knew of in advance of this research. Xilinx proprietary option with IODelays. Goal for next week to implement an oscillator of both main options.
\section*{Initial DCO Implementation}
Wrote Verilog in order to implement both a wave/inverter chain oscillator \& a counter based oscillator. Then made the test benches for these, only used post-synth simulation. Then implemented and tested. Counter based worked fine, runs perfectly @ generates divided clock. Wave seems to "toggle" an LED, but frequency is so high I cannot tell. Need to use a massive number of inverters, takes ages to synth \& therefore simulate. <- this seems to be vivado doing silly things. Turns out vivado is optimising away even with comb. loops allowed.

\end{document}

### 2018-10-19 matlab/importScopeData.m
Adding matlab & importing correct RingOsc

@@ -1,0 +1,1 @@
+[scope_x_data,scope_y_data] = importAgilentBin('scope_3.bin',1);

### 2018-10-19 matlab/process_scope_data.m

@@ -1,0 +1,8 @@
+y_fft = fft(scope_y_data);
+f_samp = 4E9;
+num_samp = length(scope_y_data);
+y_fft_scaled = abs(y_fft)/num_samp;
+freqs = f_samp*(1:1:num_samp/2)/(num_samp);
+y_fft_plottable = real(y_fft_scaled(1:floor(num_samp/2)));
+
+plot(freqs,y_fft_plottable)

### 2018-10-20 Journal.tex

@@
 Goal for next week to implement an oscillator of both main options.
-\section*{Initial DCO Implementation}
-Wrote Verilog in order to implement both a wave/inverter chain oscillator \& a counter based oscillator.
-Then made the test benches for these, only used post-synth simulation.
-Then implemented and tested.
-Counter based worked fine, runs perfectly @ generates divided clock.
-Wave seems to "toggle" an LED, but frequency is so high I cannot tell.
-Need to use a massive number of inverters, takes ages to synth \& therefore simulate.
-<- this seems to be vivado doing silly things.
-Turns out vivado is optimising away even with comb.
-loops allowed.
 

### 2018-10-21 matlab/generateSquareWave.m
process_scope_data.m

@@ -1,0 +1,12 @@
+function [t,square_wave] = generateSquareWave(period_mean,num_samp,samp_rate,phase_shift)
+    freq=1/period_mean;
+    offset=3.375/2;
+    amp=3.375/2;
+    duty=50;
+    t_max = num_samp*(1/samp_rate); 
+    t=linspace(0,t_max,num_samp);%100 seconds
+    square_wave=offset+amp*square(2*pi*freq.*t-duty+phase_shift*pi);
+%     plot(t,sq_wav)
+%     hold on;
+%     plot(t,scope_y_data)
+end

### 2018-10-21 matlab/getPeriodDrift.m

@@ -1,0 +1,10 @@
+function [period_drift] = plotPeriodDrift(periods)
+
+    period_drift = zeros(1,length(periods));
+    period_drift(1) = 0;
+    for i = 2:length(periods)
+        period_drift(i) = periods(i) - periods(1);
+    end
+
+end
+

### 2018-10-21 matlab/getPeriods.m

@@ -1,0 +1,8 @@
+function [periods] = getPeriods(init_cross)
+
+    periods = zeros(1,length(init_cross)-1);
+    for i = 2:length(init_cross)
+        periods(i-1) = init_cross(i) - init_cross(i-1);
+    end
+end
+

### 2018-10-21 matlab/process_scope_data.m

@@ -1,8 +1,59 @@
-y_fft = fft(scope_y_data);
+%% get data
+filename = 'scope_3.bin';
+[scope_x_data,scope_y_data] = importAgilentBin(filename,1);
+
+%% fft
 f_samp = 4E9;
 num_samp = length(scope_y_data);
-y_fft_scaled = abs(y_fft)/num_samp;
-freqs = f_samp*(1:1:num_samp/2)/(num_samp);
+freqs = f_samp*(1:1:num_samp/2)/(num_samp)/1E6;
+
+y_fft = fft(scope_y_data);
+y_fft_scaled = abs(y_fft);
 y_fft_plottable = real(y_fft_scaled(1:floor(num_samp/2)));
 
-plot(freqs,y_fft_plottable)
+
+num_datasets = 1;
+for loop_inc = 1:num_datasets
+    %% compute periods
+    [w,init_cross,final_cross,mid_level] = pulsewidth(scope_y_data,scope_x_data, 'Polarity', 'Positive');
+
+    [periods] = getPeriods(init_cross);
+    
+    period_mean(loop_inc) = mean(periods)*1E9; %in nanoseconds
+    period_jitter_std(loop_inc) = std(periods)*1E9; %in nanoseconds
+
+    % @~1000 samples 3.090 for 1 outside
+    % https://www.sitime.com/api/gated/AN10007-Jitter-and-measurement.pdf
+    period_jitter_p2p = 2*3.090*period_jitter_std*1E9; %in nanoseconds
+    
+    if(length(init_cross) >= 10000)
+        long_term_jitter(loop_inc) = periods(10000)-periods(1);
+    end   
+       
+    phase_shift = 0.10;
+    [t,square_wave] = generateSquareWave(period_mean,num_samp,f_samp,phase_shift);
+    
+    [~,init_cross_ideal,~,~] = pulsewidth(square_wave,scope_x_data, 'Polarity', 'Positive');
+
+    
+    [periods_ideal] = getPeriods(init_cross_ideal);
+    
+    time_interval_error = (periods - periods_ideal)*1E9; %in nanoseconds
+end
+if (num_datasets == 1)
+    figure
+    plot(t,scope_y_data)
+    title('Time Domain Data')
+    ylabel('Amplitude')
+    xlabel('Time (s)')
+    figure
+    plot(freqs,y_fft_plottable)
+    title('FFT')
+    ylabel('')
+    xlabel('Frequency (MHz)')
+    figure
[…]

### 2018-10-21 tex_files/Journal.tex
hournal update

--- 2018-10-15 Journal.tex
@@
 loops allowed.
+Needed to faff around the .xdc file.
+
+Eventually went for 1000 -> gives 1.58 MHz w/ inv chain. Works as expected. ~315 ps gate delay.
+\section*{Jitter detection}
+Started the week attempting to implement a system using the FPGA to measure jitter of the clock.
+Prior to this I implemented the ability to tune the frequencies of the oscillators.
+The RO was given 4 bit control, removing an extra pair of inverters each time.
+The PAccum was given 9 bit control.
+When I tried to do jitter detection I ran into a pair of problems. The first of these is my resolution was going to suffer.
+At 5 MHz RO clock Tnom is 200 nsec. Running the FPGA at 500 MHz gives 2 ns resolution. Period change through two oscillators is 1.2 ns. Cannot tell 3! jumps apart.
+Switched to using the agilent 4 GS/s scope, it has 250 psec resolution.
+Collected data friday, ran it through matlab, got some basic jitter calculations done.
+Wrote a bash script that implements a GPL c library that interacts with the scope over ethernet.
+Hopefully it works and can be used to automate datacapture process
+Will test Monday.
 

### 2018-10-22 matlab/process_scope_data.m
.

@@ -24,21 +24,21 @@ for loop_inc = 1:num_datasets
 
     % @~1000 samples 3.090 for 1 outside
     % https://www.sitime.com/api/gated/AN10007-Jitter-and-measurement.pdf
-    period_jitter_p2p = 2*3.090*period_jitter_std*1E9; %in nanoseconds
+    period_jitter_p2p = 2*3.090*period_jitter_std; %in nanoseconds
     
     if(length(init_cross) >= 10000)
         long_term_jitter(loop_inc) = periods(10000)-periods(1);
     end   
        
     phase_shift = 0.10;
-    [t,square_wave] = generateSquareWave(period_mean,num_samp,f_samp,phase_shift);
+    [t,square_wave] = generateSquareWave(period_mean/1E9,num_samp,f_samp,phase_shift);
     
     [~,init_cross_ideal,~,~] = pulsewidth(square_wave,scope_x_data, 'Polarity', 'Positive');
 
     
     [periods_ideal] = getPeriods(init_cross_ideal);
     
-    time_interval_error = (periods - periods_ideal)*1E9; %in nanoseconds
+    time_interval_error = (periods - periods_ideal);%*1E9; %in nanoseconds
 end
 if (num_datasets == 1)
     figure

### 2018-10-23 tex_files/Journal.tex
journal updates

@@
 \usepackage{cancel}
-\usepackage{gensymb}
+%\usepackage{gensymb}
 \usepackage{graphicx}
@@
 \usepackage{esvect} 
-\usepackage{lipsum}
+%\usepackage{lipsum}
 \usepackage{hhline}
@@
 Mostly consisted of reading the theses sent to me by Elena/Brian to get an idea of what the "parent" projects involved and why my project was required.
+\subsection*{Paper 1}
 \section*{D/NCO Research}
-After 2nd meeting with supervisors the goal was to find out what options there were in terms of implementing an oscillator on an FPGA.
-Only two real options were phase accumulator \& ring oscillator.
-Both of these I knew of in advance of this research.
-Xilinx proprietary option with IODelays.
-Goal for next week to implement an oscillator of both main options.
+After 2nd meeting with supervisors the goal was to find out what options there were in terms of implementing an oscillator on an FPGA. Only two real options were phase accumulator \& ring oscillator.
+Both of these I knew of in advance of this research. Xilinx proprietary option with IODelays. Goal for next week to implement an oscillator of both main options. %todo Link papers
 \section*{Initial DCO Implementation}
@@
 Will test Monday.
+\section*{MMCM Block}
+Set up the MMCM block such that I was able to generate both the 5 MHz signal \"required\" for the 7 segment display and a high freq.
+signal for the phase acccumultor as otherwise using the 100 MHz external signal as a input to the phase accumulator block resulted in Vivado errors.
+Now likely working at 160 Meg, just need to check w/ freq.
+meter.
 

### 2018-10-23 tex_files/Journal.tex
journal

--- 2018-10-20 Journal.tex
@@
 Mostly consisted of reading the theses sent to me by Elena/Brian to get an idea of what the "parent" projects involved and why my project was required.
+\subsection*{Paper 1}
 \section*{D/NCO Research}
-After 2nd meeting with supervisors the goal was to find out what options there were in terms of implementing an oscillator on an FPGA.
-Only two real options were phase accumulator \& ring oscillator.
-Both of these I knew of in advance of this research.
-Xilinx proprietary option with IODelays.
-Goal for next week to implement an oscillator of both main options.
+After 2nd meeting with supervisors the goal was to find out what options there were in terms of implementing an oscillator on an FPGA. Only two real options were phase accumulator \& ring oscillator.
+Both of these I knew of in advance of this research. Xilinx proprietary option with IODelays. Goal for next week to implement an oscillator of both main options. %TODO Link papers & this section
+\section*{Initial DCO Implementation}
+Having analysed the potential options for the oscillator in this network we decided to implement both of the two viable options so my task for the week was to do this.
+The phase accumulator was easy to implement, just requiring a counter with an adjustable addition value.\\
+The ring oscillator was more complicated although not because of the code itself being especially difficult, just requiring the use of a generate block to avoid having to type out the same line hundreds of times for each of the inverters required.
+The tricky part of this however was forcing Vivado to implement my verilog as written, as it wished to optimise away my inverter chains as it saw them as being redundant.
+The fact I had a combinatorial loop in my design presented it's own challenges in overcoming Vivado's complaints.
+
+%Wrote Verilog in order to implement both a wave/inverter chain oscillator \& a counter based oscillator.
+Then made the test benches for these, only used post-synth simulation.
+Then implemented and tested.
+Counter based worked fine, runs perfectly @ generates divided clock.
+Wave seems to "toggle" an LED, but frequency is so high I cannot tell.
+Need to use a massive number of inverters, takes ages to synth \& therefore simulate.
+<- this seems to be vivado doing silly things.
+Turns out vivado is optimising away even with comb.
+loops allowed.
+Needed to faff around the .xdc file.
+
+%Eventually went for 1000 -> gives 1.58 MHz w/ inv chain. Works as expected. ~315 ps gate delay.
+\section*{Jitter detection}
+Started the week attempting to implement a system using the FPGA to measure clock jitter.
+Prior to this I implemented the ability to tune the frequencies of the oscillators using the switches on the board.
+The ring oscillator was given 4 bit control, removing an extra pair of inverters each time, and an enable.
+As the phase accumulator has a naturally wide tuning range of to up half the clock frequency and the accumulator itself being 10 bits wide it made sense to give it 9 bit wide control as well as an enable.\\
+
+I then attempted to implement the jitter detection on the FPGA itself.
+I ran into some issues with the crossing of clock domains and went to speak to Brian about it.
+While talking to him we came to the conclusion that the resolution offered by a detector on the FPGA would not offer a resolution high enough to accurately measure the jitter.
+For example at a 500 MHz FPGA clock, which is at the higher end of what is attainable, the period is 2 nano seconds.
+As things stand the delay through one inverter is approximately 315 picoseconds each one adds 0.63 nanoseconds to the period of the signal.
+At 5 MHz then:
+$$ T_{nom} = 200\times10^{-9}\text{ ns ,}T_{next} = 202\times10^{-9}\text{ ns} $$
+$$ F_{nom} = 5\times10^{6}\text{ Hz ,}F_{next} = 4.950495\times10^{-9}\text{ Hz} $$
+$$ \Delta F_{FPGA} = 49.504\text{ kHz} $$
+While the ring oscillators adjustment resolution is 1.26 nanoseconds given each step removes an inverter pair this is a change of 
+$$ T_{nom} = 200\times10^{-9}\text{ ns ,}T_{next} = 201.26\times10^{-9}\text{ ns} $$
+$$ F_{nom} = 5\times10^{6}\text{ Hz ,}F_{next} = 4.968697\times10^{-9}\text{ Hz} $$
+$$ \Delta F_{RO} = 31.302\text{ kHz} $$
+As the supposed jitter detector cannot even detect a change in the period by a step it is not a useful method for the detection of jitter.\\
+In response to this we decided to do the jitter measurements used a 4 GS/s oscilloscope which has a resolution of 250 picoseconds, and following the same procedure this gives:
+%...
+$$ \Delta F_{Scope} = 6.24\text{ kHz} $$
+which is a significant improvement and allows for a number of measurement points for each oscillator resolution step.
+
+I then took a sample measurement of the clock output of the ring oscillator running at 5 MHz and began to process it in Matlab to get an idea of the jitter.
+To do this I wrote a basic script which calculated the mean and standard deviation of the periods in the sample as well and the Time Interval Error.
+\section*{OScope Control}
+Over the weekend I attempted to establish whether the oscilloscope could be controlled from my computer to automate the data collection process, and having initially hit a dead end with the Agilent provided Matlab functionality I found some an implementation in \texttt{C} which I have yet to test.
+
+\section*{MMCM Block}
+After running into issues with the use of the reference 100 MHz clock via the on chip clock manager at the same time as using it for the Phase Accumulator I had made.
+Accordingly I resetup the MMCM block in order to generate both the buffered high frequency clock for the Phase Accumulator and the lower frequency 5 MHz signal \"required\" for the 7 segment display.
 

### 2018-10-24 tex_files/Journal.tex
working more on tidying the journal

@@
 \section*{Initial Research}
-Mostly consisted of reading the theses sent to me by Elena/Brian to get an idea of what the "parent" projects involved and why my project was required.
-\subsection*{Paper 1}
+The first week after I received my project was spent researching the concepts involved, mostly through the reading of papers and theses sent to me by my supervisors.
+These allowed me to gain an understanding as to why my project was needed, the technology that was involved and the goal of the project.
+% I should probably make a thing in here regarding what these papers entailed?
+% Non trivial however, reviewing those might be a better goal for a less busy period
+
+%\subsection*{Paper 1}
 \section*{D/NCO Research}
-After 2nd meeting with supervisors the goal was to find out what options there were in terms of implementing an oscillator on an FPGA. Only two real options were phase accumulator \& ring oscillator.
-Both of these I knew of in advance of this research. Xilinx proprietary option with IODelays. Goal for next week to implement an oscillator of both main options. %TODO Link papers & this section
+After my second supervisor meeting a goal was set for the following week of researching the potential designs for the Digitally/Numerically Controller Oscillator for this network.
+Going into this I was aware of two potential designs, one of these being a chain of inverters, which provided there was an odd number would oscillate with a period given by:
+$$ 2\times\text{delay}\times n $$
+Where $\text{delay}$ is the time taken for a signal pro propagate through an inverter and $n$ is the number of inverters in the chain.\\
+The second option that I was aware of in advance of the research was a phase accumulator using a counter which would overflow at $2^n-1$ where $n$ is the bit width of the counter.
+By controlling the number that is added on each clock cycle of the main FPGA clock driving the counter the period of this oscillator can be adjusted.\\
+The only other option I managed to come across for an oscillator on an FPGA was using Xilinx proprietary \texttt{IODELAY} blocks, which have a permanently programmable delay at runtime \ref{IODELAY}.
+This could be coupled with an odd number of inverters in order to implement a quasi ring oscillator.
+However as these are IO blocks their usage is not particularly conducive to the creation of a network of ADPLLs.
+%\ref{IODELAY} https://www.xilinx.com/support/documentation/application_notes/xapp872.pdf
+%After 2nd meeting with supervisors the goal was to find out what options there were in terms of implementing an oscillator on an FPGA. Only two real options were phase accumulator \& ring oscillator.
+
+%Both of these I knew of in advance of this research. Xilinx proprietary option with IODelays. Goal for next week to implement an oscillator of both main options. %TODO Link papers & this section
 \section*{Initial DCO Implementation}
@@
 The tricky part of this however was forcing Vivado to implement my verilog as written, as it wished to optimise away my inverter chains as it saw them as being redundant.
-The fact I had a combinatorial loop in my design presented it's own challenges in overcoming Vivado's complaints.
+The fact I had a combinatorial loop in my design presented it's own challenges in overcoming Vivado's complaints, especially as the net/cell names kept changing.\\
+One of these problems came down to the usage of the incorrect Verilog keyword in order to prevent the optimisation of my inverters.
+I had taken this segment of the code from Brian's DSD notes and around the web many others had used the same directive in their code, \texttt{( * KEEP = true * )}, so some google-fu was required here.
+Eventually I discovered that \texttt{KEEP} was limited to the synthesis and implementation stages hence my design working in timing simulations yet failing to be pass the generation of a bitstream.\\
+I figured out that the correct directive was in fact \texttt{( * DONT\_TOUCH = true * )}, which while performing correctly and was able to carry on with fixing the combinatorial loop issue.\\
+It turns out there is a \textit{simple} solution to the combinatorial loops problem as Vivado informs the user while reading the logs. However the suggested remedy of \newline
+\texttt{set\_property ALLOW\_COMBINATORIAL\_LOOPS true [get\_nets \\<module\_hierarchy/net\_name>]}\newline
+is not actually a functioning solution to this issue.
+As it transpires the correct method is to select one of the cells that Vivado reports as being part of this loop and use this to locate the appropriate cell, shown in the following command.\newline
+\texttt{set\_property ALLOW\_COMBINATORIAL\_LOOPS true [get\_nets -of\_objects [get\_cells <module\_hierarchy/cell\_name>]]}\newline
+The other drawback of this method is that due to Vivado renaming cells and/or nets when the structure of the design changes the constraints are not automatically reapplied, however there may be a more generic way to implement this.
 
@@
 $$ \Delta F_{Scope} = 6.24\text{ kHz} $$
-which is a significant improvement and allows for a number of measurement points for each oscillator resolution step.
+which is a significant improvement and allows for a number of measurement points for each oscillator resolution step.\\
 
@@
 To do this I wrote a basic script which calculated the mean and standard deviation of the periods in the sample as well and the Time Interval Error.
+This week I also met with Eugene and demonstrated my code for him as well as giving him binary files containing 8 MSamp of data for both the ring oscillator and the phase accumulator.
+He will be able to analyse their jitter characteristics and from there begin to determine the optimum control coefficients.\\
+%TODO pierre or filip?
+Worked also with Philipe getting his oscillator to work, using a similar design to mine however implemented in VHDL.\\
+
 \section*{OScope Control}
-Over the weekend I attempted to establish whether the oscilloscope could be controlled from my computer to automate the data collection process, and having initially hit a dead end with the Agilent provided Matlab functionality I found some an implementation in \texttt{C} which I have yet to test.
+Over the weekend I attempted to establish whether the oscilloscope could be controlled from my computer to automate the data collection process, and having initially hit a dead end with the Agilent provided Matlab functionality I found some an implementation in \texttt{C} from a Professor in Nottingham.\\
+I worked on this on Wednesday and hit roadblocks with direct wiring. I got from Brian a router and was, after some initial setup issues, to get basic communication going with the scope.
+However when I attempted to extract the data from the oscilloscope I hit problems, a call to \texttt{new char[LEN]} was throwing a \texttt{std::Alloc} error which prevented the data actually being read back from the scope.
+I suspect this may possibly be due to the buffer I was attempting to create being 8 million elements or more in length. 
+I intend trying this again tomorrow with reduced sizes and hopefully this will work.
+There is also the potential that rather than extracting the data I could use the scope's inbuilt save to USB in order to circumvent this problem.
 
@@
 Accordingly I resetup the MMCM block in order to generate both the buffered high frequency clock for the Phase Accumulator and the lower frequency 5 MHz signal \"required\" for the 7 segment display.
+Testing again proved that both could be run simultaneously, the benefit of in phase clocks being that crossing of clock domains does not need to be considered.
 

### 2018-11-05 tex_files/Journal.tex
journal entry

@@
 
+\section*{Phase Detection}
+The next step then was to work on a phase detection mechanism for the ADPLL.
+First off I went down the simplest route and implemented a 1 bit bang-bang detector using just a single D flip flop.
+After meeting with Brian on Wednesday we discussed some of the challenges of implementing the phase detector on an FPGA.
+Foremost among these was the issue of metastability and the likelihood of both signals occurring at the same time.
+For an implementation on an ASIC where an arbitration circuit can be constructed using gates I do not have this luxury, as I am restricted to simple look-up table elements.
+As a temporary workaround for this we decided I would implement a phase detector with the assumption that the signals are synchronised to the FPGA clock.\\
+Working under this assumption I created a phase detector which counts the number of FPGA clock cycles between edges of the two signals in question.
+In order to avoid issues where the rising edge of the second signal detected occurs very shortly before the next rising edge of the first signal detected the maximum phase difference that can be detected is the high time of the first signal.\\
+This phase detector outputs signed 2s complement integers that can be fed directly into the loop filter which succeeds this block.
+
+The problem that this introduces then is the minimum phase difference detectable is 2.5 nano seconds which equates to a 65 KHz frequency or a 4.5$^\circ$ phase deviation.
+
 \end{document}

### 2018-11-13 matlab/plotVivadoSim.m
prior to on bus journal updates

@@ -1,0 +1,10 @@
+function plotVivadoSim(filepath)
+%UNTITLED Summary of this function goes here
+%   Detailed explanation goes here
+data = csvread(filepath);
+plot(data(:,1),data(:,2))
+title('Step response to ? frequency difference');
+xlabel('Time steps');
+ylabel('Error code');
+end
+

### 2018-11-13 tex_files/Journal.tex

@@
 He will be able to analyse their jitter characteristics and from there begin to determine the optimum control coefficients.\\
-%TODO pierre or filip?
-Worked also with Philipe getting his oscillator to work, using a similar design to mine however implemented in VHDL.\\
+Worked also with Pierre getting his oscillator to work, using a similar design to mine however implemented in VHDL.\\
 
@@
 
-\section*{Phase Detection}
+\section*{Phase Detection I}
 The next step then was to work on a phase detection mechanism for the ADPLL.
@@
 
+\section*{Phase Detection II}
+The above phase detector was created using a state machine with the following diagram describing its behaviour:\\
+%TODO insert image of the state transitions
+It is preceded by two synchronisers which attempt to avoid metastable conditions, which will be used for the time being so that I could advance until we found a method that worked more generally.\\
+I was able to test this phase detector and confirm that it was measuring a frequency difference in line with what I expected given my simulation results, despite spending too much time trying to figure out why the display was not giving me a reasonable output when it was just being updated far too quickly to be human readable.\\
+%TODO insert image of the RTL diagram
+I then realised I was also operating at a 160 MHz clock in both simulation and hardware so I attempted to push the clock up to the 400 MHz mark then but started to run into so timing issues on both inter and intra clock paths so I had to dial back the FPGA clock frequency to 258 MHz which was enough to remove all but what I think is an allowable timing violation on a synchroniser.\\
+Working with the 1:0.1 rule for proportional/integral gains I then implemented a PI loop filter as in Eldar's PhD on Friday in order to complete an entire multi-bit ADPLL.
+After ironing out one issue caused by a lack of bit extension causing strange behaviour on the control code if it was negative which had gone undetected in simulation I was able to get my PLL locking to an on-chip reference signal late on Friday.
+In more detail the 8 bit signed output of the error loop filter was being added to a 12 bit bias signal and due to a lack of sign extension -1 became a +511.
+This error would correct itself on the next cycle leading to signal which looked like: $||--|||||-----||--|||||$\\%TODO do I have a screenshot of this error?
+Pierre told me about his method for the arbiter so maybe I can implement that method myself rather than relying on my limited use case version.
+%TODO insert a good simulation
+
 \end{document}

### 2018-11-14 tex_files/Journal.tex
14-11-28

@@
 I was able to test this phase detector and confirm that it was measuring a frequency difference in line with what I expected given my simulation results, despite spending too much time trying to figure out why the display was not giving me a reasonable output when it was just being updated far too quickly to be human readable.\\
-%TODO insert image of the RTL diagram
+\begin{figure}[h]
+    \centering
+    \includegraphics[scale=0.33]{../ADPLL_RTL_VIVADO}
+    \caption{RTL Diagram of the ADPLL}
+    \label{fig:adpll_rtl}
+\end{figure}
 I then realised I was also operating at a 160 MHz clock in both simulation and hardware so I attempted to push the clock up to the 400 MHz mark then but started to run into so timing issues on both inter and intra clock paths so I had to dial back the FPGA clock frequency to 258 MHz which was enough to remove all but what I think is an allowable timing violation on a synchroniser.\\
@@
 In more detail the 8 bit signed output of the error loop filter was being added to a 12 bit bias signal and due to a lack of sign extension -1 became a +511.
-This error would correct itself on the next cycle leading to signal which looked like: $||--|||||-----||--|||||$\\%TODO do I have a screenshot of this error?
+This error would correct itself on the next cycle leading to signal which looked like: $||--|||||-----||--|||||$\\r
 Pierre told me about his method for the arbiter so maybe I can implement that method myself rather than relying on my limited use case version.
-%TODO insert a good simulation
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[scale=0.33]{../good_sim}
+    \caption{Example simulation}
+    \label{fig:example_sim}
+\end{figure}
+
+\section*{Basic Measurements}
+Lock range at bias of 82:
+$$4.19\rightarrow12.34\textrm{ MHz} $$ 
+Capture range at bias of 82:
+$$4.25\rightarrow12.76\textrm{ MHz}  $$
+Jitter at bias of 82:\\
+%TODO
+Jitter locked to 5~ MHz:\\
+%TODO
+Frequency sensitivity(control code to output):
+$$ 258\times10^6\frac{1}{2^{12}}=62.988\textrm{ MHz} $$
+Overall proportional gain (time $\rightarrow$ TDC $\rightarrow$ frequency):
+$$ 8.998 \textrm{ KHz/degree} $$
+LPF cutoff:\\
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[scale=0.33]{../HMO_locking_4M7_to5M0}
+    \caption{Example of locking at startup}
+    \label{fig:hmo_locking}
+\end{figure}%redo on good scope
 

### 2018-11-14 matlab/process_scope_data.m

@@ -1,6 +1,19 @@
 %% get data
-filename = 'scope_3.bin';
-[scope_x_data,scope_y_data] = importAgilentBin(filename,1);
+clear variables
+filename = 'scope_18.bin';
+% [scope_x_data,scope_y_data] = importAgilentBin(filename,2);
+num = 1;
+
+periods = [];
+for inc1 = 1:10
+    filename = char("scope_"+num2str(18+inc1)+".bin");
+    [scope_x_data(inc1,:),scope_y_data(inc1,:)] = importAgilentBin(filename,1);
+    
+    [w,init_cross,final_cross,mid_level] = pulsewidth(scope_y_data(inc1,:),scope_x_data(inc1,:), 'Polarity', 'Positive');
+    periods = [periods;getPeriods(init_cross)];
+end
+
+periods = reshape(periods, [1 size(periods,1)*size(periods,2)]);
 
 %% fft
 f_samp = 4E9;
@@ -15,25 +28,26 @@ y_fft_plottable = real(y_fft_scaled(1:floor(num_samp/2)));
 num_datasets = 1;
 for loop_inc = 1:num_datasets
     %% compute periods
-    [w,init_cross,final_cross,mid_level] = pulsewidth(scope_y_data,scope_x_data, 'Polarity', 'Positive');
-
-    [periods] = getPeriods(init_cross);
+%     [w,init_cross,final_cross,mid_level] = pulsewidth(scope_y_data,scope_x_data, 'Polarity', 'Positive');
+% 
+%     [periods] = getPeriods(init_cross);
     
     period_mean(loop_inc) = mean(periods)*1E9; %in nanoseconds
     period_jitter_std(loop_inc) = std(periods)*1E9; %in nanoseconds
 
     % @~1000 samples 3.090 for 1 outside
     % https://www.sitime.com/api/gated/AN10007-Jitter-and-measurement.pdf
-    period_jitter_p2p = 2*3.090*period_jitter_std; %in nanoseconds
+    period_jitter_p2p = 2*3.719*period_jitter_std; %in nanoseconds
     
-    if(length(init_cross) >= 10000)
-        long_term_jitter(loop_inc) = periods(10000)-periods(1);
-    end   
+%     if(length(periods) >= 10000)
+%         long_term_jitter(loop_inc) = periods(10000)-periods(1);
+%     end   
        
     phase_shift = 0.10;
[…]

### 2018-11-14 tex_files/Journal.tex
same date part 2

@@
 Jitter at bias of 82:\\
-%TODO
+$$ 200.1665\pm1.9689\textrm{ nsec} $$
 Jitter locked to 5~ MHz:\\
-%TODO
+$$ 200.009\pm2.7378\textrm{ nsec} $$
 Frequency sensitivity(control code to output):

### 2018-11-14 matlab/process_scope_data.m

@@ -6,7 +6,7 @@ num = 1;
 
 periods = [];
 for inc1 = 1:10
-    filename = char("scope_"+num2str(18+inc1)+".bin");
+    filename = char("scope_"+num2str(4+inc1)+".bin");
     [scope_x_data(inc1,:),scope_y_data(inc1,:)] = importAgilentBin(filename,1);
     
     [w,init_cross,final_cross,mid_level] = pulsewidth(scope_y_data(inc1,:),scope_x_data(inc1,:), 'Polarity', 'Positive');

### 2018-11-21 matlab/process_scope_data.m
2AM lul

@@ -1,7 +1,12 @@
 %% get data
 clear variables
 filename = 'scope_18.bin';
-% [scope_x_data,scope_y_data] = importAgilentBin(filename,2);
+[scope_x_data,scope_y_data] = importAgilentBin(filename,1);
+[scope_x_data,scope_y_data2] = importAgilentBin(filename,2);
+[w,init_cross,final_cross,mid_level] = pulsewidth(scope_y_data,scope_x_data, 'Polarity', 'Positive');
+fsamp = 4E9;
+num_samp = length(scope_y_data);
+periods = getPeriods(init_cross);
 num = 1;
 
 periods = [];
@@ -60,19 +65,32 @@ if (num_datasets == 1)
     title('Time Domain Data')
     ylabel('Amplitude')
     xlabel('Time (s)')
+    
     figure
     plot(freqs,y_fft_plottable)
     title('FFT')
     ylabel('')
     xlabel('Frequency (MHz)')
+    
     figure
     plot(time_interval_error)
     title('Time Interval Error')
     ylabel('Error (ns)')
     xlabel('Samples')
+    
     figure
     histfit(periods)
     title()
     xlabel()
     ylabel()
+    
+    figure
+    plot(t*1e6-7.95,scope_y_data,'b')
+    hold on;    
+    plot(t*1e6-7.95,scope_y_data2+4,'r')
+    ylabel('Amplitude')
+    xlabel('Time (s)')
+    xlim([0 9])
+    xlabel("Time (\mu sec)")
+    set(gca,'ytick',[])
 end

### 2018-11-21 matlab/importScopeData.m
1902

@@ -1,1 +1,1 @@
-[scope_x_data,scope_y_data] = importAgilentBin('scope_3.bin',1);
+[scope_x_data,scope_y_data] = importAgilentBin('scope_18.bin',1);

### 2018-11-21 matlab/plot_error_over_time.m

@@ -1,0 +1,28 @@
+%plot_error_over_time
+
+filename = 'scope_18.bin';
+[~,scope_y_data] = importAgilentBin(filename,1);
+[scope_x_data,scope_y_data2] = importAgilentBin(filename,2);
+[w,init_cross,final_cross,mid_level] = pulsewidth(scope_y_data,scope_x_data, 'Polarity', 'Positive');
+[w2,init_cross2,final_cross2,mid_level2] = pulsewidth(scope_y_data2,scope_x_data, 'Polarity', 'Positive');
+
+fsamp = 4E9;
+num_samp = size(scope_y_data,2);
+
+periods = getPeriods(init_cross);
+periods2 = getPeriods(init_cross2);
+
+time_interval_error = periods - periods2(size(periods2,2)-size(periods,2)+1:end);
+
+times = (1:1:size(time_interval_error,2))*mean(periods2)*1E6;
+
+figure
+plot(times,time_interval_error*1E9,'b')
+hold on;
+linex = [24.6 24.6];
+liney = [-100 300];
+plot(linex, liney,'r');
+ylim([-50 200])
+xlim([0 40])
+ylabel('Time Interval Error (ns)')
+xlabel('Time (\mus)')

### 2018-11-21 matlab/process_scope_data.m

@@ -1,17 +1,9 @@
 %% get data
 clear variables
-filename = 'scope_18.bin';
-[scope_x_data,scope_y_data] = importAgilentBin(filename,1);
-[scope_x_data,scope_y_data2] = importAgilentBin(filename,2);
-[w,init_cross,final_cross,mid_level] = pulsewidth(scope_y_data,scope_x_data, 'Polarity', 'Positive');
-fsamp = 4E9;
-num_samp = length(scope_y_data);
-periods = getPeriods(init_cross);
-num = 1;
 
 periods = [];
 for inc1 = 1:10
-    filename = char("scope_"+num2str(4+inc1)+".bin");
+    filename = char("scope_"+num2str(18+inc1)+".bin");
     [scope_x_data(inc1,:),scope_y_data(inc1,:)] = importAgilentBin(filename,1);
     
     [w,init_cross,final_cross,mid_level] = pulsewidth(scope_y_data(inc1,:),scope_x_data(inc1,:), 'Polarity', 'Positive');
@@ -58,6 +50,10 @@ for loop_inc = 1:num_datasets
     [periods_ideal] = getPeriods(init_cross_ideal);
     
     time_interval_error = (periods - periods_ideal);%*1E9; %in nanoseconds
+end
+
+if (false)
+    time_interval_error = periods - periods2(size(periods2,2)-size(periods,2)+1:end);
 end
 if (num_datasets == 1)
     figure

### 2018-11-24 tex_files/test.tex
final(ish) presentation

\documentclass[12pt]{article}
\usepackage{smartdiagram}
\usesmartdiagramlibrary{additions}
\begin{document}
	\begin{figure}
		\centering
		\smartdiagramset{
			uniform color list=white!60!green for 3 items,
			back arrow disabled=true,
			module minimum width=2cm,
			module minimum height=2cm,
			module x sep=4cm,
			text width=3cm,
			additions={
				additional item offset=20mm,
				additional item width=2cm,
				additional item height=2cm,
				additional item text width=3cm,
				additional item shadow=drop shadow,
				additional item bottom color=white!60!yellow,
				additional item border color=gray,
				additional arrow color=white!60!yellow,
		}}
		\smartdiagramadd[flow diagram:horizontal]{
			Background Reading, Design ADPLL, Test ADPLL
		}{
			below of module1/External Reference,below of module2/Inverter Chain PLL,below of module3/Interlinked PLLs
		}
		\smartdiagramconnect{->}{additional-module1/additional-module2,additional-module2/additional-module3}
		\begin{tikzpicture}[remember picture,overlay]% modified from p. 47 of manual
		\draw[additional item arrow type] (module3) |- ([yshift=-10mm]module1.south) -- (additional-module1);
		\end{tikzpicture}

		\vspace{40mm}\vspace{20mm}
		\smartdiagramset{
			uniform color list=white!60!orange for 6 items,
			back arrow disabled=true,
			module minimum width=2cm,
			module minimum height=2cm,
			module x sep=4cm,
			text width=3cm,
			additions={
				additional item offset=20mm,
				additional item width=2cm,
				additional item height=2cm,
				additional item text width=3cm,
				additional item shadow=drop shadow,
				additional item bottom color=white!60!black,
				additional item border color=gray,
				additional arrow color=orange,
		}}
		\begin{tikzpicture}[remember picture,overlay]% modified from p. 47 of manual
		\draw[additional item arrow type] (additional-module3) |- ([yshift=-10mm]additional-module1.south) -- (additional-module1);
		\end{tikzpicture}
		\smartdiagramadd[flow diagram:horizontal]{
			Acquire Better FPGA, Build ADPLL Network,  Test/Analyse Network
		}{
		below of module1/External Reference,below of module2/Inverter Chain PLL,below of module3/Interlinked PLLs
	}\par
	\end{figure}
\end{document}

### 2018-11-24 matlab/plot_error_over_time.m

@@ -27,3 +27,16 @@ xlim([0 40])
 ylabel('Time Interval Error (ns)')
 xlabel('Time (\mus)')
 text(30, 150, "Locked", 'Color', 'r');
+
+figure
+plot(times,1./periods*1E-6,'b')
+hold on;
+linex = [25.2 25.2];
+liney = [-100 300];
+plot(linex, liney,'r');
+ylim([2 7])
+xlim([0 40])
+%xlabel("Time (\mu sec)")
+%ylabel("Frequency (MHz)")
+title("Frequency (MHz) over Time (\mus)")
+text(30, 6.5, "Locked", 'Color', 'r');

### 2018-11-26 matlab/process_scope_data.m
...

@@ -31,7 +31,8 @@ for loop_inc = 1:num_datasets
     
     period_mean(loop_inc) = mean(periods)*1E9; %in nanoseconds
     period_jitter_std(loop_inc) = std(periods)*1E9; %in nanoseconds
-
+    period_jitter_rms(loop_inc) = 
+    
     % @~1000 samples 3.090 for 1 outside
     % https://www.sitime.com/api/gated/AN10007-Jitter-and-measurement.pdf
     period_jitter_p2p = 2*3.719*period_jitter_std; %in nanoseconds
@@ -49,8 +50,12 @@ for loop_inc = 1:num_datasets
     
     [periods_ideal] = getPeriods(init_cross_ideal);
     
-    time_interval_error = (periods - periods_ideal);%*1E9; %in nanoseconds
+    %time_interval_error = (periods - periods_ideal);%*1E9; %in nanoseconds
 end
+
+tmp = abs(periods-period_mean*1E-9);
+max(tmp)
+
 
 if (false)
     time_interval_error = periods - periods2(size(periods2,2)-size(periods,2)+1:end);

### 2019-01-01 tex_files/Journal.tex
...

@@
 
+\section*{RingADPLL}
+Had to change the number of inverters significantly as it the new layout since my last test of a significantly more simple system had a different layout, especially only 8 steps of 2 inverters rather than 32 in this design which \texttt{OscTestTopLevel} used with a raw Bang Bang Phase Detector.
+Unfortuantely all simulations broke and I was not actually able to simulate this re calibration of the ``bias''(really the number of inverters) and had to do it manually by generating bitstreams and seeing how high low I could pull the frequency of operation.
+
+\section*{External Reference}
+I was able to simply repurpose an Arduino PWM script I had previously made as the reference.
+I then was able to run the ADPLL at 10 MHz from a reference at 1.25 MHz.
+The choice to run at 10 rather than 5 Meg was due to instability that was introduce with a division by eight to use the reference and a resulting inability to actually lock without reducing the clock frequency of the phase detection circuit (4x required for 8x reference scaling).
+I also made some large changes to the Loop Filter to allow for gain variation both in width and value.
+
 \end{document}

### 2019-01-04 tex_files/Journal.tex
starting interim report

@@
 Had to change the number of inverters significantly as it the new layout since my last test of a significantly more simple system had a different layout, especially only 8 steps of 2 inverters rather than 32 in this design which \texttt{OscTestTopLevel} used with a raw Bang Bang Phase Detector.
-Unfortuantely all simulations broke and I was not actually able to simulate this re calibration of the ``bias''(really the number of inverters) and had to do it manually by generating bitstreams and seeing how high low I could pull the frequency of operation.
+Unfortunately all simulations broke and I was not actually able to simulate this re calibration of the ``bias''(really the number of inverters) and had to do it manually by generating bit streams and seeing how high low I could pull the frequency of operation.
 
 \section*{External Reference}
-I was able to simply repurpose an Arduino PWM script I had previously made as the reference.
+I was able to simply re purpose an Arduino PWM script I had previously made as the reference.
 I then was able to run the ADPLL at 10 MHz from a reference at 1.25 MHz.
@@
 
+\section*{Truncation in 2s Complement (LF)}
+This is an issue that needs addressing. I am truncating and that may corrupt the number. For now it doesn't matter however it may with certain configurations.
+
+\section*{RingOsc Sim Issue}
+For some reason I cannot run simulations of anything other than the RingOsc module itself. Any high level module fails to simulate properly, seemingly zero delay oscillation once the chain is enabled.
+
 \end{document}

### 2019-01-04 tex_files/placeholder.tex
starting interim report

\documentclass[12pt,english,british]{report}
\usepackage[a4paper]{geometry}
%\usepackage[T1]{fontenc}
\usepackage[latin9]{inputenc}
\geometry{verbose,tmargin=3cm,bmargin=3cm,lmargin=3.5cm,rmargin=2.4cm}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{cancel}
%\usepackage{gensymb}
\usepackage{graphicx}
\usepackage{esint}
\usepackage{mdsymbol}
\usepackage{esvect} 
%\usepackage{lipsum}
\usepackage{hhline}
\usepackage{eurosym}
\usepackage{listings}
\usepackage{booktabs}
\usepackage{amssymb}
\usepackage{mathrsfs}
\usepackage{commath}
\usepackage{adjustbox}
\usepackage{booktabs}
\usepackage{array}
\usepackage{lscape}
\usepackage[xspace]{ellipsis}
\usepackage{color}
\usepackage{float}
\usepackage{caption}
\usepackage{afterpage}

\definecolor{mygreen}{rgb}{0,0.6,0}
\definecolor{mygray}{rgb}{0.5,0.5,0.5}
\definecolor{mymauve}{rgb}{0.58,0,0.82}
\begin{document}

\title{Placeholder Tex Doc}
\author{Conor Dooley}
\maketitle
\chapter{Background Review}
\section{Brief Overview}

In a world where the demand for high performance hand-held computing devices continues to grow and the prevalence of ``smart'' devices is increasing the problem of maintaining the steady gain in performance of Systems-On-Chip remains at the forefront. %TODO cite?
The main drivers of performance in Systems-On-Chip are the number of transistors on a chip, the clocking frequency and parallelism. However the latter will not be addressed in this assessment, partly as the performance of the paralellised operation is largely governed by the number of transistors and clock frequency to begin with. %TODO Excessive?
As the number of transistors on a chip has increased roughly following Moore's Law the increase in clock frequency has not been able to follow a similar linear trajectory, which clock frequency having remained roughly equivalent for the last 10 years \cite{ross2008cpu}.
This plateauing of clock frequency has been caused by the high power consumption due to the demands placed by the global distribution of a high frequency clock, which is often the single biggest consumer of power on the chip. %TODO cite (can use tiwari1998reducing)
In a world where low power devices are desirable, especially with the growing demand of Internet-Of-Things where high power consumption goes directly against one of the key pillars of the technology. %TODO Pillars of what?
In digital systems two main approaches are used when designing the clocking system. In both cases the chip is broken down into small areas in which all transistors are clocked synchronously, with the size constrained by the ability to deliver a quality clock signal to all transistors. The first of these methods is Globally Synchronous Locally Synchronous, or GSLS, where all of these sub areas on the chip operate synchronously with each other. In practice however this is very difficult to achieve as extremely high precision is required across the entire area of the chip, and doing so leads to high power consumption.
In contrast a Globally Asynchronous Locally Synchronous clock delivery system ensures that in each sub division of the chip clocking is synchronous however the ``local'' areas are independent from each other. This reduces the clocking system's complexity and thus the power consumption and chip area used at the expense of communication speed between blocks. %TODO citation?
A GSLS system however has the advantages of deterministic behaviour and greater rates of communication between clocking areas and as such remains a desirable system design. A number of methods which deliver GSLS clocking exist at present such as clock trees as well as emerging technologies such as ADPLL networks.

\section{The Impact of Clocking Errors}

In Figure \ref{fig:eldar_why_precise_clocking} the data path between two synchronously clocked registers is shown, with the circuit's function being carried out by the combinatorial network between the registers.
Each register has a setup time, which represents the amount of time that the input value to a register must remain constant before the clock edge, and a hold time, the time for which the input must remain constant after a clock edge.
\begin{figure}[h]
	\centering
	\includegraphics[scale=0.5]{../eldar_why_precise_clocking}
	\caption{Data Flow in a Clocked System \cite{zianbetov2013distributed}.}
	\label{fig:eldar_why_precise_clocking}
\end{figure}\newline
A lack of synchronisation between the clock edges will manifest itself as a time difference between the clocking events at both registers, $\Delta T = t^i_k - t^f_k$. $\Delta T$ is considered to be ergodic and can be described by an average deviation called skew and random process that is normally modelled as a Gaussian random variable. If $\Delta T$ is negative this reduces the time available for the intervening combinatorial network thereby having the same effect as a reduction in clocking frequency.
The most common sources of clock error are caused by mismatches which usually stem from production such as differences in the length of clocking paths, buffer delays or in the parameters of either active or passive components in the clock distribution network. These all manifest themselves as skew, while the noise in active components will appear as jitter in the clock signal.

\section{Traditional Solutions}

A number of traditional solutions exist which provide GSLS clocking systems, using a variety of techniques. The most simple of these use the clock distribution network's symmetry in order to deliver the clock in phase at all locations in the chip and are named according to their geometry with the main variants being branch-trees, X-trees or H-trees. %TODO insert figure
\begin{figure}[h]
	\centering
	\includegraphics[scale=0.5]{../eldar_why_precise_clocking}
	\caption{Data Flow in a Clocked System \cite{zianbetov2013distributed}.}%TODO insert figure
	\label{fig:trees}
\end{figure}
While on the surface these appear simple the task of obtaining an exact matching is, in practise, the limiting factor in this design. Even if the clock distribution system is geometrically symmetrical production mismatches in either active or passive components will lead to a skew that varies from part to part. In order to minimise the impact of production tolerances the dimensions of components in the distribution network can be increased, thus reducing the relative variation possible. However this has the impact of increasing the power consumption of the distribution network \cite{tiwari1998reducing}.

A mesh clock distribution network is an alternate design where the clock is delivered using a Cartesian grid of distribution lines and variation in skew is inversely proportional to the density of the grid. According to Abdelhadi \textit{et al} (2010) clock meshes ``\textit{achieve low and deterministic skew, low skew variations, and low jitter}'', all desirable characteristics for a clock distribution system. However they dissipate more power due to extra capacitive loading due to vast number of lines required to form the grid and similarly to tree distribution networks suffer from potential mismatch in production and alleviation through increasing dimensions will again lead to higher power consumption \cite{abdelhadi2010timing}. 
\begin{figure}[h]
	\centering
	\includegraphics[scale=0.5]{../eldar_why_precise_clocking}
	\caption{Data Flow in a Clocked System \cite{zianbetov2013distributed}.}%TODO insert figure
	\label{fig:mesh}
\end{figure}

Tree distribution systems are able to correct for some of the skew created in the fabrication process by means of a skew compensation system, of which there are two main variants. Firstly a centralised skew compensation system has a single 
In a clocking distribution system using a tree type network the skew due to the manufacturing process can be partly accounted for by means of a skew compensation system, of which there are two main types. These types are defined by the location of the control mechanism with those having one controller located at the clock source being known as ``centralised'' methods and those with multiple controllers in the individual clocking areas known as ``decentralised''.

\chapter{Progress to Date}

\newpage
\bibliography{bib} 
\bibliographystyle{IEEEtran}

\end{document}

### 2019-01-04 tex_files/placeholder.tex
6 pgs

@@
-\documentclass[12pt,english,british]{report}
+\documentclass[11pt,english,british]{report}
 \usepackage[a4paper]{geometry}
@@
 \usepackage[latin9]{inputenc}
+\usepackage{setspace}
+
 \geometry{verbose,tmargin=3cm,bmargin=3cm,lmargin=3.5cm,rmargin=2.4cm}
@@
 \section{Brief Overview}
-
+\doublespacing
 In a world where the demand for high performance hand-held computing devices continues to grow and the prevalence of ``smart'' devices is increasing the problem of maintaining the steady gain in performance of Systems-On-Chip remains at the forefront.
@@
 As the number of transistors on a chip has increased roughly following Moore's Law the increase in clock frequency has not been able to follow a similar linear trajectory, which clock frequency having remained roughly equivalent for the last 10 years \cite{ross2008cpu}.
+
 This plateauing of clock frequency has been caused by the high power consumption due to the demands placed by the global distribution of a high frequency clock, which is often the single biggest consumer of power on the chip.
@@
 %TODO Pillars of what?
+
 In digital systems two main approaches are used when designing the clocking system.
@@
 
+Designs do exist in which the electrical lines used in the tree network are replaced by waveguides for optical signals, which negates the electrical losses however this technique suffers from being a newer technology and as such is expensive to realise \cite{haurylau2006chip}.
+
 A mesh clock distribution network is an alternate design where the clock is delivered using a Cartesian grid of distribution lines and variation in skew is inversely proportional to the density of the grid.
@@
 
-Tree distribution systems are able to correct for some of the skew created in the fabrication process by means of a skew compensation system, of which there are two main variants.
-Firstly a centralised skew compensation system has a single
 In a clocking distribution system using a tree type network the skew due to the manufacturing process can be partly accounted for by means of a skew compensation system, of which there are two main types.
 These types are defined by the location of the control mechanism with those having one controller located at the clock source being known as ``centralised'' methods and those with multiple controllers in the individual clocking areas known as ``decentralised''.
+In a centralised skew compensation circuit the skew across the chip is calculated by the central controller which then manipulates the distribution network in order to deliver a more in-phase clock around the chip.
+This however adds complication to a circuit which already is a large consumer of power and as such increases the power usage.
+As the name suggest a decentralised skew compensation technique delegates this responsibility to the individual clock regions.
+For example Yamashita \textit{et al} designed a system in which each clocking area or ``leaf node'' contains a partial clock tree.
+Each of these ``leaves'' is able to compare its clock phase to the neighbouring node and based on the result tune an adjustable delay buffer \cite{yamashita2005dynamic}.
+While this method can compensate for process, voltage and temperature variation it does not address the power consumption due to the delivery of a high frequency clock across the entire chip area.
+
+\section{Multi-oscillator Designs}
+The designs described previously are all similar in that they have a single central oscillator that provides the clock for all areas of the chip, whereas the following methods attempt to synchronise multiple oscillators, each of which provides the clock for a single clocking area.
+The main advantage of a multi-oscillator design is that as each clocking area has its clock created locally there is no problem with signal degradation.
+Similar to a decentralised skew compensation scheme comparisons are only made to neighbouring blocks and as such there is reduced power loss due to transmission.
+\begin{figure}[h]
+	\centering
+	\includegraphics[scale=0.5]{../mizuno1998noise}
+	\caption{Coupled Oscillator Clock Delivery Circuit \cite{mizuno1998noise}.}
+	\label{fig:mizuno1998noise}
+\end{figure}
+One such method is a network of oscillators as in Figure \ref{fig:mizuno1998noise} which uses coupled PLLs to generate local clocks.
+The advantage of this method is that only the control voltage $v_c$ needs delivery to all areas of the chip thus alleviating the need for a power hungry distribution circuit while also being more noise immune than the transmission of a high frequency clock.
+However this design still suffers from clock variation as all VCOs are fed the same control voltage, and this is acknowledged by the authors:
+\begin{quotation}
+	\singlespacing
+	\textit{Unfortunately, as with the conventional ...
+method, distributing the VCOs over the entire chip causes the problem that jitter and skew are increased by variations in the fabrication process (static), temperature, and power supply (dynamic)} \cite{mizuno1998noise}.
+	\doublespacing
+\end{quotation}
+This design also only compares the phase of a single branch of the distribution network with the reference in order calculate the control voltage which again may contribute to a degradation in clock synchronisation.
+%TODO why
+
+Another potential multi-oscillator design synchronises the oscillators using the phase relationship with the oscillators in the neighbouring clocking area.
+Once again this negates the requirement for a global distribution structure and comparisons need only be made with neighbouring areas, and as a phase lock loop is being used only a divided down version of the clock is required.
+This in turn means the hardware transporting the divided clock signal to the phase comparator has significantly lower requirements placed on it thus lowering the power consumption.
+Pratt and Nguyen initially proposed this design in their paper published in 1995 entitled ``\textit{Distributed Synchronous Clocking}'' in which they propose a Cartesian grid of clocking areas, each with their own PLL, which has become known as a PLL Network \cite{pratt1995distributed}.
+In this design any given node is synchronised with its neighbours and one of the corner nodes is additionally synchronised with the reference.
+According to the authors this is ``a simple, effective way to achieve low cost, high quality, low skew clock generation in a synchronous parallel processor''.
+This methodology was implemented by Gutnik \textit{et al} (2000) who found:
+\begin{quote}
+	\singlespacing
+	\textit{Design and measurements on this chip confirm that generating and synchronizing multiple clocks on chip is feasible.
+Neither the power nor the area overhead of multiple PLLs is substantial compared to the cost of distributing the clock by conventional means} \cite{gutnik2000active}.
+	\doublespacing
+\end{quote}
+\begin{figure}[h]
+	\centering
+	\includegraphics[scale=0.5]{../gutnik2000active}
+	\caption{PLL Network Topology \cite{gutnik2000active}.}
+	\label{fig:gutnik2000active}
+\end{figure}
+
+\section{ADPLL Networks}
 

### 2019-01-05 tex_files/placeholder.tex
10 pages

@@
 	\includegraphics[scale=0.5]{../eldar_why_precise_clocking}
-	\caption{Data Flow in a Clocked System \cite{zianbetov2013distributed}.}
+	\caption{Data Flow in a Clocked System \cite{zianbetov2013phd}.}
 	\label{fig:eldar_why_precise_clocking}
@@
 However this has the impact of increasing the power consumption of the distribution network \cite{tiwari1998reducing}.
-
 Designs do exist in which the electrical lines used in the tree network are replaced by waveguides for optical signals, which negates the electrical losses however this technique suffers from being a newer technology and as such is expensive to realise \cite{haurylau2006chip}.
@@
 	\includegraphics[scale=0.5]{../eldar_why_precise_clocking}
-	\caption{Data Flow in a Clocked System \cite{zianbetov2013distributed}.}%TODO insert figure
+	\caption{Data Flow in a Clocked System \cite{zianbetov2013phd}.}%TODO insert figure
 	\label{fig:mesh}
@@
 
+As a PLL network is an analog circuit its integration in a modern IC is a barrier to usage and it has not been used in any commercial designs \cite{zianbetov2013distributed}.
+An alternative design that is more suitable for current fabrication techniques eschews from using analog components and instead implements the network of controlled oscillators using only digital circuitry, hence the name All-Digital PLL.
+A 4x4 ADPLL network was designed and prototyped in 65 nm CMOS by Zianbetov and Shan in order to test the suitability of the technique as a clock distributor \cite{zianbetov2013phd,shan2014phd}.
+%TODO fix phd references
+In this design the oscillators are coupled using digital phase comparators which attempt to measure the phase error between two oscillators with each non edge node being connected to four neighbours.
+Figure \ref{fig:eldar_node} shows high level detail of the architecture of both the entire clocking system and that of an individual node in the design.
+A digital PLL network has the additional advantages over its analog counterparts that it can benefit from advancements in digital circuit design suites, be reconfigurable and programmable and has a significantly greater immunity to perturbations inherent to its digital nature \cite{zianbetov2013phd}.
+This last advantage is of particular use in a digital environment, as otherwise there is potential for clock degradation resulting from switching of transistors.
+\begin{figure}[h]
+	\centering
+	\includegraphics[scale=0.35]{../ccirc_2013_arch}
+	\caption{Architecture of the ADPLL network and of a single node \cite{zianbetov2013distributed}.}
+	\label{fig:eldar_node}
+\end{figure}
+Looking at the design of a given node it is notable that the function carried out by the Error Combiner  is akin to a average therefore as mentioned by Pratt and Nyugen there is potential to settle in a stable ``mode'' in which the oscillators are not synchronised.
+In their paper they present a method where initial start-up is performed uni-directionally and once all nodes are close full connectivity can be restored \cite{pratt1995distributed}.
+In creating an entirely digital system Zianbetov and Shan could exploit the reconfigurability and and implement this system at start-up.
+
+\section{ADPLL Architecture}
+As indicated in Figure \ref{fig:mulkeen_pll} the three main building blocks of a conventional Phase Lock Loop are the Phase Detector, Loop Filter and Controllable Oscillator.
+In the All-Digital counterpart each of these traditionally analog components are replaced by their digital counterparts, necessitating quantisation in order to remain physically realisable.
+\begin{figure}[h]
+	\centering
+	\includegraphics[scale=0.7]{mulkeen_pll}
+	\caption{Block Diagram of a Phase Lock Loop, \textit{Wireless Systems Notes}, Mulkeen (2017).}
+	\label{fig:mulkeen_pll}
+\end{figure}
+
+\subsection{Digitally Controllable Oscillator}
+In a digital system there are a very limited number of voltages representable, and frequently there are only two levels, so using this to control the oscillator is not a viable strategy.
+Instead a fixed bit width signal is used to control the oscillator's period, selecting the number of inverters in a ring oscillator or the varactor configuration of a travelling wave oscillator \cite{chen2011rotary}.
+The decisions made in the design of the digitally controllable oscillator, or DCO, determine many of the other ADPLL parameters.
+While tuning range and centre frequency as well as linearity carry over from the analog counterpart a DCO also has a frequency step which in combination with the bit width of the control signal determines the range over which the oscillator can be tuned.
+Figure \ref{fig:my_ring} illustrates a basic ring oscillator design.
+In a ring oscillator is an inherently unstable circuit composed of an odd number of inveters.
+where the period of oscillation is given by twice entire chain's delay.
+The frequency of operation can then be set by modulating the length of the chain by means of \texttt{f\_select}.
+The main impact of output frequency quantisation is that only frequencies which are integer multiples of the frequency step away from the centre can be easily reproduced, with intermediate values only obtainable in a manner akin to Fractional-N synthesis with the control code toggling back and forth.
+This acts as a source of jitter in the system.
+\begin{figure}[h]
+	\centering
+	\includegraphics[scale=0.35]{../inverter_chain}
+	\caption{Basic Ring/Inverter Chain Oscillator.}
+	\label{fig:my_ring}
+\end{figure}
+
+\subsection{Digital Phase Detector}
+Once again quantisation impacts the Phase Detector, as rather than a continuous output the phase detector in an ADPLL has a finite number of output values thus limiting the accuracy of the phase detector.
+A second form of quantisation is also present as unlike an analog system a digital phase detector does not provide continuous data in the time domain either, instead relying on sampling.
+At its most basic a digital phase comparator may only output an indication of which signal is leading, a design known as a Bang-Bang Detector, which can be constructed using a single D Flip Flop.
+As the output only has two levels the resultant word is only 1 bit wide and as such limits the range over which the output frequency can be controlled.
+More complex designs such as that in Figure \ref{fig:shan_bb_pd}, implemented by Shan, build on this by measuring the time difference between edges of the signals using a Time-to-Digital Converter, or TDC, in his case using a tapped delay line \cite{shan2014phd}.
+This essentially mimics a counter and allows for a non binary form of phase measurement.
+
+\begin{figure}[h] %TODO top
+	\centering
+	\includegraphics[scale=0.35]{../shan_bb_pd}
+	\caption{Bang-bang phase/frequency detector architecture. \cite{shan2014phd}.}
+	\label{fig:shan_bb_pd}
+\end{figure}
+\subsection{Digital Loop Filter}
+\begin{figure}[h]
+	\centering
+	\includegraphics[scale=0.35]{../shan_bb_pd} %TODO this diagram
+	\caption{Basic PI Controller Architecture.}
+	\label{fig:my_simple_pi}
+\end{figure}
+The loop filter in an ADPLL can be implemented as a PI controller as it is just a low pass filter, such as that in Figure \ref{fig:my_simple_pi}.
+In the case of an ADPLL node in a network the input of this filter is a weighted sum of two to four phase errors, each corresponding to a neighbouring node.
+In a digital system the proportional section is implemented by a simple multiplier whereas the integral path is a combination of a multiplier and a delayed summation stage.
+The value of these gains determine the response and stability of the oscillator system.
+The transfer function of such a circuit is given by:
+\begin{equation*}
+	H(z) = \alpha + \beta\frac{1}{1-z^{-1}} = \frac{(\alpha + \beta) - \alpha z^{-1}}{1-z^{-1}}~~~\cite{shan2014phd}
+\end{equation*}
+It has been found by Koskin \textit{et al} (2018) that stable operation can be achieved when the integral gain, $k_i$, is less than half the proportional gain, $k_p$ \cite{koskin2018generation}.
+In the same study a range of values were found which would produce low jitter operation of the network.
+As these value are all less than one the filter must implement fixed point arithmetic in an effort to maintain the simplicity of the clock distribution network, rather than incurring the complexity penalty of floating point calculations.
+
 \chapter{Progress to Date}

### 2019-01-06 tex_files/placeholder.tex
...

@@
 \maketitle
+\doublespacing
 \chapter{Background Review}
 \section{Brief Overview}
-\doublespacing
 In a world where the demand for high performance hand-held computing devices continues to grow and the prevalence of ``smart'' devices is increasing the problem of maintaining the steady gain in performance of Systems-On-Chip remains at the forefront.
@@
 %TODO Excessive?
-As the number of transistors on a chip has increased roughly following Moore's Law the increase in clock frequency has not been able to follow a similar linear trajectory, which clock frequency having remained roughly equivalent for the last 10 years \cite{ross2008cpu}.
+As the number of transistors on a chip has increased roughly following Moore's Law the increase in clock frequency has not been able to follow a similar linear trajectory, which clock frequency having remained roughly equivalent for the last number of years \cite{ross2008cpu}.
 
@@
 \section{The Impact of Clocking Errors}
-
 In Figure \ref{fig:eldar_why_precise_clocking} the data path between two synchronously clocked registers is shown, with the circuit's function being carried out by the combinatorial network between the registers.
@@
 \section{Traditional Solutions}
-
 A number of traditional solutions exist which provide GSLS clocking systems, using a variety of techniques.
@@
 \section{ADPLL Networks}
-
 As a PLL network is an analog circuit its integration in a modern IC is a barrier to usage and it has not been used in any commercial designs \cite{zianbetov2013distributed}.
@@
 
+\section{My Contribution}
+The planned contribution of this project is to create an FPGA based prototyping platform for the testing of network parameters and architectural variations.
+An FPGA as a prototyping tool for networks was previously utilised by Zianbetov and Shan as the final stage before the creation of a custom chip as it allowed for ``a partial validation of the ASIC design, through a verification of the essential functional features of the designed clocking network \cite{zianbetov2013phd,shan2014phd}.
+They however note that there are some fundamental differences between the two environments in particular that the FPGA based prototype is unable to create the mixed signal circuits used in an ASIC resulting in alternative designs being mandated for certain components such as the Time-to-Digital Converter.
+Similarly most commericially available FPGA boards were noted to have clock frequencies well below the gigahertz levels possible in ASIC designs.
+In spite of these problems the FPGA as a prototyping solution is valuable as it allows for hardware based testing of architectures and the simulation of non trivial circuit behaviours.
+
+The novelty of this particular implementation is that in the case of Zianbetov and Shan only a small network was simulated on an FPGA and at a frequency close to 50 kHz, however this project will attempt to significantly increase the frequency of operation to the MHz range as well as expanding the size of the network, from 16 oscillators to potentially 100.
+This is done with the aim of assisting work being undertaken by a number of researchers in UCD who are working on aspects of ADPLL networks who would stand to benefit from such a platform.
+
 \chapter{Progress to Date}
+The first stage of this project was the creation of an ADPLL design that would be used for the network.
+Due to limitations of the connections available on the FPGA evaluation board used to this point, a Digilent/Xilinx Nexys 4, signals at frequencies significantly beyond 5 MHz suffered from major degradation as the connections were not designed for such frequencies.
+Thus a target frequency was set at 5 MHz while this board was in use.
+After initial research into potential oscillator designs and what would be feasible given the constraints of the particular platform design began with the aforementioned oscillator.
+\section{Controllable Oscillator}
+In my research into potential designs for numerically controllable oscillators that were viable on an FPGA I came across two main options.
+The first of these was the ring oscillator, constructed from an odd number of inverters strung together in a chain \cite{predraig}.
+In this case the frequency of operation is determined by the number of oscillators in the chain as the delay through the individual inverters determines the half period of the signal.
+This period can then be modified by adding or removing inverters in a manner that maintains the odd number of total inverters.
+The propagation time through an individual inverter is dependant on the layout of the ring and thus this design must be used with some care.\\
+The second option is one that I was aware of in advance this project's beginning.
+A phase accumulator makes use of the overflow property of a counter, and has a frequency determined by the width of the counter and the frequency at which it is clocked.
+The oscillator operates as follows: Once per clock cycle the counter will increment by a number $k$ until overflow is reached at $2^n-1$ where $n$ is the bit width of the counter.
+By controlling the magnitude of $k$ added on each cycle of the driving clock the frequency of this oscillator can be adjusted.
+The most significant bit of this counter can form the clock signal as it has an approximately 50\% duty cycle.\\
+As it is not reliant on the clock of FPGA the ring oscillator design is a more interesting design in the context of simulation, but the phase accumulator is significantly easier to manage due to being more easily simulatable.
+The decision was made to create both designs in order to keep options open.\\
+The only other option I managed to come across for an oscillator on an FPGA was using Xilinx proprietary \texttt{IODELAY} blocks, which have a permanently programmable delay at runtime \ref{IODELAY}.
+This could be coupled with an odd number of inverters in order to implement a quasi ring oscillator.
+However as these are IO blocks their usage is not particularly conducive to the creation of a network of ADPLLs.\\
+The phase accumulator was designed to have a bit width of 12, all of which could be set as the control code.
+As the maximum frequency of operation that had no timing violations was 258 MHz the frequency step of this oscillator design was 62.988 kHz and before integration into an ADPLL had a frequency range from 62.988 kHz to just under half the clock frequency at 128.94 MHz.
+Figure \ref{fig:my_pa} details the simple construction of this oscillator.\\
+\begin{figure}[h]
+	\centering
+	\includegraphics[scale=0.5]{../phase_accum}
+	\caption{Phase Accumulator Block Diagram.}
+	\label{fig:my_pa}
+\end{figure}
+As originally implemented for testing purposes the inverter ring oscillator had a propagation time per inverter of 315 picoseconds, thus each inverter would add 0.63 nano seconds to the total period, and had 4 control bits.
+In order to obtain a 5 MHz output from this type of ring oscillator a ring of of 317 inverters is required, and to set this as the centre point of the frequency range the total number of inverters was 325.
+At 5 MHz the frequency step of this design was calculated as:
+\begin{align*}
+T_{nom} = 200\times10^{-9}\text{ ns ,}T_{next} = 201.26\times10^{-9}\text{ ns} \\
+F_{nom} = 5\times10^{6}\text{ Hz ,}F_{next} = 4.968697\times10^{-9}\text{ Hz}
+\end{align*} 
+\begin{equation*}
+\Delta F_{RO} = 31.302\text{ kHz}
+\end{equation*}
+%TODO look into more detail on yoke.
+
+\begin{figure}[h]
+	\centering
+	\includegraphics[scale=0.35]{../ring_osc}
+	\caption{Simplified Ring Oscillator Block Diagram.}
+	\label{fig:my_ring_osc}
+\end{figure}
+\chapter{Future Work}
 

### 2019-01-07 tex_files/placeholder.tex
...

@@
 \end{equation*}
-%TODO look into more detail on yoke.
-
+%TODO look into more detail on yoke with later work.
 \begin{figure}[h]
@@
 	\label{fig:my_ring_osc}
+\end{figure}
+\section{Phase Detector}
+Moving along the chain the next block then is the phase detector, or more accurately phase comparator.
+At this stage in development this block just compares the oscillator's output with that of just one reference.
+In the future this block will be followed up with one which performs a weighted sum of all neighbouring nodes.
+The phase detector used in Shan's design has an arbitration circuit that is not synthesisable and an alternate design is required for use on an FPGA.
+The eventual design I chose has two main parts, firstly an up/down counter performs the role of the Time-to-Digital converter while also indicating the sign of the phase difference.
+In order to control the count, and indicate the direction in which to do so, a state machine determines which edge occurs first.
+In order to circumvent the possibility of a metastability condition being reached the incoming signals are synchronised to the higher frequency FPGA clock prior to their entry into the phase detector.
+From there the state machine can operate as designed, in accordance with the state transition diagram in Figure \ref{fig:my_state_machine}.
+\begin{figure}[h]
+	\centering
+	\includegraphics[scale=0.35]{../state_trans_diagram}
+	\caption{Phase Detection State Machine State Transition Diagram.}
+	\label{fig:my_state_machine}
+\end{figure}
+The state machine behaves as follows: Starting in the "Waiting" state, if the reference has a rising edge first it is said to be leading the oscillators output and the sign of the phase error is positive.
+Alternatively if the opposite occurs and the first rising edge is seen on the oscillator output the phase error is given a negative sign and the TDC counts downwards.
+In either case this continues until a rising edge occurs on the other signal which stops the count, updates the phase error output of the block and finally clears the counter.
+Between "Copy \& Clear" events the output of the block is held constant.
+There is a second condition that can stop the count and that is the occurrence of a falling edge of the signal that began the count.
+This does however limit the rate of lock acquisition in the case where the phase difference is greater than 180$^\circ$ or the frequency difference is significant.
+The operation however is the same as a conventional method when either locked or within a normal operating tolerance of being locked.\\
+In order to represent both the count value and sign in one value the counter operates in twos complement which allows the output be used directly in subsequent blocks.
+The counter itself has a bit width of 8 thereby having a range of -128 to 127 which at 5 MHz gives a phase resolution of:
+\begin{align*}
+	T_{res} &= \frac{1}{258\times 10^6} = 3.8760~ns \\
+	T_{period} &= \frac{1}{5\times 10^6} = 200~ns \\
+	\textrm{Ratio} &=\frac{3.8760}{200} = 0.0194 \\
+	\textrm{Resolution} & = 0.0194\times 360^\circ = 6.9768^\circ
+\end{align*}
+
+\section{Loop Filter}
+As with Shan's design I used a PI controller as the loop filter in order to align with the typical technology and also benefit from the work of member of the research team in UCD \cite{shan2014phd,koskin2018generation}.
+%TODO eugene
+The bit with of the filter's input is once again 8 bits to match up with the input from the phase detector and outputs a 9 bit value in order to accommodate the summation of an 8 bit number from each of the paths.
+I selected $k_p = 1$ and $k_i = 0.125$ in order to align with the region of stable coefficients discovered by Eugene \cite{koskin2018generation}.
+%TODO redo image of LF
+\begin{figure}[h]
+	\centering
+	\includegraphics[scale=0.35]{../loop_filter}
+	\caption{PI Loop Filter RTL diagram.}
+	\label{fig:my_pi}
+\end{figure}\\
+Figure \ref{fig:my_pi} showing the RTL diagram of the filter shows the two distinct paths through the filter, the more simple of these being the proportional one.
+The multiplication necessitates the increased width of the path in order to preserve the accuracy of the fixed point calculations, a width which is then cut down to an 8 bit value before being added to the output of the integral branch.
+The proportional gain value has a fixed point width of 2.1 allowing for values from -2 to 1.5, however in testing to this point only 1 and 0.5 have been used.
+The integral path is similar with the gain having a fixed width of 1.3, a range of -1 to 0.750, and for my testing set to 0.125.
+After the gain is applied via an accumulator with a delayed input clocked by the 258 MHz FPGA clock.
+Once again the output of this path is cut back down to an 8 bit value.\\
+The addition of two 8 bit values results in a 9 bit signed output from the phase detection block which is then combined with the bias to form the control code for the oscillator.
+As a higher control code leads to a larger frequency with the phase accumulator this combination is done by additions, whereas for the ring oscillator the relationship is inverse as we are controlling the period, so instead subtraction is used.
+
+\section{ADPLL}
+Combining these blocks, in this case using the Phase Accumulator style oscillator, gives the ADPLL shown in Figure \ref{fig:my_adpll} on which I performed my initial testing and simulations.
+\begin{figure}[h]
+	\centering
+	\includegraphics[scale=0.35]{../rtl}
+	\caption{Phase Accumulator ADPLL RTL diagram.}
+	\label{fig:my_adpll}
 \end{figure}

### 2019-01-08 tex_files/placeholder.tex
...

@@
 \end{figure}
+
+\chapter{Test Results}
+
 \chapter{Future Work}
+So far all work has been done using a single phase lock loop in order to have a suitable basis for the network and going forward this PLL will be used in that role.
+The first stage of network implementation will interlink two PLLs, with one of them also receiving input from the external reference.
+This test will serve as both validation of the phase coupling ability of the network as well as a test of the error summation block and the ability to lock from startup.
+For the sake of more realistic performance the inverter chain ADPLL will be used ahead of the Phase Accumulator based design.\\
+A shortcoming of the current FPGA evaluation board is the inability to input or extract signals in the multiple MHz range without significant distortion, and this will be resolved with the arrival of an FPGA more suited to extracting and importing higher frequency signals.
+With an ability to extract higher frequency signals for measurement and analysis purposes the target frequency of 5 MHz can potentially be increased into the tens of MHz range.\\
+With simple coupling implemented the logical next stage is the implementation of a small network, possibly 2x2, before scaling up to a size suitable to the FPGA in use.
+At this point analysis of measurement results can be undertaken and any refinements to the network can be made.
+Additionally, if time permits, the Phase Accumulator based design could be implemented and tested and comparisons made between both designs.
 

### 2019-01-09 tex_files/placeholder.tex
...

@@
 	\centering
-	\includegraphics[scale=0.5]{../eldar_why_precise_clocking}
-	\caption{Data Flow in a Clocked System \cite{zianbetov2013distributed}.}%TODO insert figure
+	\includegraphics[scale=0.3]{../trees}
+	\caption{H and Branch Tree Clock Distribution Systems.}%TODO insert figure
 	\label{fig:trees}
@@
 	\centering
-	\includegraphics[scale=0.5]{../eldar_why_precise_clocking}
-	\caption{Data Flow in a Clocked System \cite{zianbetov2013phd}.}%TODO insert figure
+	\includegraphics[scale=0.6]{eldar_mesh}
+	\caption{Mesh Clock Distribution System \cite{zianbetov2013phd}.}%TODO insert figure
 	\label{fig:mesh}
@@
 	\centering
-	\includegraphics[scale=0.35]{../shan_bb_pd} %TODO this diagram
+	\includegraphics[scale=0.45]{../pi_simple} %TODO this diagram
 	\caption{Basic PI Controller Architecture.}
@@
 Combining these blocks, in this case using the Phase Accumulator style oscillator, gives the ADPLL shown in Figure \ref{fig:my_adpll} on which I performed my initial testing and simulations.
+For ease of testing the PLL replicated the reference exactly, as this would facilitate easier testing of the lock and capture ranges of the design.
 \begin{figure}[h]
@@
 	\label{fig:my_adpll}
-\end{figure}
+\end{figure}\\
+Once this testing had been completed with the phase accumulator based ADPLL testing progressed to the design implementing an inverter chain and scaling from the reference using an eight times divider.
+This necessitated a change in the bit width of the control signal to match the 5 bit wide signal desired, thus leading to the design in Figure \ref{fig:my_ring_adpll}.
+\begin{figure}[h]
+	\centering
+	\includegraphics[scale=0.35]{../ro_rtl}
+	\caption{Ring Oscillator ADPLL RTL diagram.}
+	\label{fig:my_ring_adpll}
+\end{figure}\\
+ The Loop Filter is clocked at the generated output frequency as this ensures it will be synchronised with the oscillator.
+This is important when using the inverter chain oscillator in order to avoid metastability as the generated output clock it may not be synchronised with the high frequency clock of the FPGA.
 
+\chapter{Test Results}
 \chapter{Test Results}
@@
 \chapter{Future Work}
+
 So far all work has been done using a single phase lock loop in order to have a suitable basis for the network and going forward this PLL will be used in that role.
@@
 At this point analysis of measurement results can be undertaken and any refinements to the network can be made.
-Additionally, if time permits, the Phase Accumulator based design could be implemented and tested and comparisons made between both designs.
+Additionally, if time permits, the Phase Accumulator based design could be implemented and tested and comparisons made between both designs.\\
+As with the single ADPLL testing the important characteristics to look out for will be the lock and capture ranges of the network which should mirror that of the ADPLLs used in it's construction.
+Similarly the jitter of a number of clocks across the network will be examined in order, and any skew between the PLL outputs investigated.
+Also of interest will be the initial locking behaviour, in particular how quickly the network can be brought up from ``cold''.
 

### 2019-01-10 tex_files/placeholder.tex
draft 1

@@
 \usepackage{setspace}
+\geometry{verbose,tmargin=3cm,bmargin=3cm,lmargin=3.5cm,rmargin=2.4cm}
+%\usepackage{fancyhdr}
+%\pagestyle{fancy}
+%\setcounter{secnumdepth}{3}
+\usepackage[british]{babel}
 
-\geometry{verbose,tmargin=3cm,bmargin=3cm,lmargin=3.5cm,rmargin=2.4cm}
+%%% Add this %%%
+\usepackage{titlesec}
+\titleformat{\chapter}[display]
+{\normalfont\huge\bfseries}{\chaptertitlename~\thechapter}{18pt}{\Huge}
+\titlespacing*{\chapter}{0pt}{50pt}{40pt}
+\titlespacing*{name=\chapter,numberless}{0pt}{-30pt}{10pt}
+%%%   End   %%%
+
 \usepackage{amsmath}
@@
 \definecolor{mymauve}{rgb}{0.58,0,0.82}
+
 \begin{document}
@@
 \author{Conor Dooley}
-\maketitle
+\begin{titlepage}
+
+	\pagenumbering{gobble}
+
+	\setstretch{1.25}
+
+	\begin{spacing}{3}
+
+		\noindent \begin{center}
+
+			{\huge{}ADPLL Network on an FPGA}
+			\par\end{center}{\huge \par}
+	\end{spacing}
+
+	\bigskip{}
+
+	\begin{center}
+		{\Large{}Conor Dooley}
+	\end{center}
+
+	\vspace{1.25cm}
+
+	\begin{center}
+		%\includegraphics[width=0.15\paperwidth]{ucd_crest.pdf}\vspace{0.75cm}
+		\includegraphics[width=0.2\paperwidth]{UCD_crest.pdf}\vspace{0.25cm}
+	\end{center}
+
+	\bigskip{}
+
+	\begin{center}
+		{\huge{}Masters Project Interim Report }
+	\end{center}
+
+	\vspace{1.cm}
+
+	{\Large{}\bigskip{}
+		\bigskip{}
+	}{\Large \par}
+
+	\noindent \begin{center}
+		\begin{tabular}{lll}
+			\textbf{\large{}Research Supervisors:} & \qquad{}\qquad{} & {\large{}Dr Elena Blokhina \& Brian Mulkeen}\\
+		\end{tabular}
+
+		\par\end{center}
+
+	\vspace{1cm}
+	\noindent \begin{center}
+		{\large{}January  2019}
+		\par\end{center}{\large \par}
+
+\end{titlepage}
 \doublespacing
@@
 
-\chapter{Test Results}
-\chapter{Test Results}
+\chapter{Results}
+
+Initially simulations were utilised to validate the behaviour of the individual blocks and of the system as a whole while using the Phase Accumulator based design.
+From these simulations I was able to test the performance of the ADPLL while locked to a signal at 5 MHz and its behaviour during the acquisition period.
+A signal at 5 MHz lies close to midway between two increments of the phase accumulator and as such the oscillator must switch between two frequencies leading to the introduction of jitter.
+From simulations performed in Vivado locked to a signal at 5 MHz I observed jitter of 3.491 ns, based on the differences between sequential periods, and a peak to peak worst case value of 9.299 ns.\\
+Similarly I performed simulations of the locking performance of the design starting with a deviation of 213 KHz and observing the time taken to achieve a lock.
+213 KHz represents an offset of three frequency steps plus the difference between 5 MHz and the nearest integer frequency step.
+Figure \ref{fig:vivado_sim} contains the visual results of the tests, in which a repetitive pattern can be seen to begin after 20.91 $\mu$s indicating a lock as been acquired.
+\begin{figure}[h]
+	\centering
+	%\includegraphics[scale=0.5]{../vivado_sim}
+	\includegraphics[scale=0.45]{../sim_locking_rect}
+	\caption{Locking Dynamics Simulation.}
+	\label{fig:vivado_sim}
+\end{figure}\\
+Identical tests were performed post implementation using a Digilent Nexys4 development board in addition to measuring the lock and capture ranges of the design.
+Firstly in order to determine the baseline for jitter in the system the oscillator was biased to 4.9761 MHz and sequential period jitter was observed to be 1.9689 ns.
+Connecting then the remainder of the ADPLL and by varying the reference the lock and capture ranges were found to be 4.19-12.76 MHz and 4.25-12.34 MHz respectively when the bias point was 5 MHz.
+Somewhat surprisingly this is not centred on the bias point, suggesting a potential flaw in the system that needs to be addressed.
+A identical locking test was performed in the implemented design, Figure \ref{fig:impl}, where using the same method a time to lock of 25.2 $\mu$s was observed.
+While locked jitter was better in both average period to period variation and peak to peak, which can be attributed to the FPGA clock frequency not exactly matching that designed and as such the distance between 5 MHz and the nearest frequency step being reduced, thereby leading to less switching required to generate a 5 MHz signal.
+\begin{figure}[h]
+	\centering
+	\includegraphics[scale=0.45]{../sim_locking_rect}
+	\caption{Locking Dynamics Measurement.}
+	\label{fig:impl}
+\end{figure}\\
+Similarly tests were carried out with the Ring Oscillator based design at a bias point of 5 MHz.
+425 inverters, with 5 bit control, were used to obtain an output in the region of 5 MHz with the bias set to the midpoint which is close to 5 MHz.
+The expected lock range given the width the use of an otherwise identical design would be the entire range of 32 steps or 64 inverters which can be easily estimated:
+\begin{align*}
+	T_{inv} &= \frac{200\times 10^{-6}}{425-2\times 16} = 0.50891\textrm{~ns} \\
+	F_{min} &= \frac{1}{(425)\times 0.50891\textrm{~ns}} = 4.623\textrm{~MHz} \\
+	F_{max} &= \frac{1}{(425-2\times 32)\times 0.50891\textrm{~ns}} = 5.443\textrm{~MHz}
+\end{align*}
+However in testing the lock and capture ranges both came out to be 4.86-5.15 MHz, so significantly below the estimates, which highlights an issue that requires investigation.\\
+
+Testing was also carried out of the ADPLL containing a divider in both divide by 4 and 8 modes to examine their impact on locking behaviour.
+The results of this testing are displayed in Table \ref{table:1}.
+The jitter in this case is rather low as the 10 MHz target frequency lies close to an integer increment of the phase accumulator design used for this test.
+Lock and capture ranges are significantly more symmetrical in these configurations than at a target of 5 MHz and as expected there is a decrease in range with the greater division values.
+\begin{figure}[h]
+	\centering
+	\includegraphics[scale=0.5]{../matlab/painintheass}
+	\caption{Testing of ADPLLs employing div4 \& div8.}
+	\label{fig:div}
+\end{figure}
+\begin{table}[!ht]
+	\begin{center} 
+		\begin{tabular}{c|c|c|c|c}           
+			& Period Jitter & Peak-Peak Jitter & Lock Range & Capture Range \\
+			div4 & 1.9858 ns & 6.4854 ns & 5.000-17.200 MHz & 5.620-16.700 MHz \\
+			div8 & 1.5702 ns & 6.1594 ns & 5.768-15.262 MHz & 5.928-14.216 MHz \\
+		\end{tabular}
+		\caption{Divider Comparison.}
+		\label{table:b_c_perf}
+	\end{center}
+\end{table}
 
 \chapter{Future Work}
-
+Firstly I will be addressing the issues regarding the lock and capture ranges, however the first of those may be fixed by a transition to a higher frequency as evidenced by the 10 MHz testing.
 So far all work has been done using a single phase lock loop in order to have a suitable basis for the network and going forward this PLL will be used in that role.
@@
 Similarly the jitter of a number of clocks across the network will be examined in order, and any skew between the PLL outputs investigated.
-Also of interest will be the initial locking behaviour, in particular how quickly the network can be brought up from ``cold''.
+Also of interest will be the initial locking behaviour, in particular how quickly the network can be brought up from ``cold''.\\
 

### 2019-01-10 matlab/a10M_quad_plot.m

@@ -1,0 +1,38 @@
+filename = 'scope_32.bin';
+[~,scope_y_data] = importAgilentBin(filename,1);
+[scope_x_data,scope_y_data2] = importAgilentBin(filename,2);
+[w,init_cross,final_cross,mid_level] = pulsewidth(scope_y_data,scope_x_data, 'Polarity', 'Positive');
+[w2,init_cross2,final_cross2,mid_level2] = pulsewidth(scope_y_data2,scope_x_data, 'Polarity', 'Positive');
+
+f_samp = 4E9;
+num_samp = length(scope_y_data);
+t_max = num_samp*(1/f_samp); 
+t=linspace(0,t_max,num_samp);
+
+figure
+hold on; 
+plot(t*1e6-7.95-0.09,scope_y_data+4,'b');
+plot(t*1e6-7.95-0.09,scope_y_data2,'r');
+
+filename = 'scope_31.bin';
+[~,scope_y_data] = importAgilentBin(filename,1);
+[scope_x_data,scope_y_data2] = importAgilentBin(filename,2);
+[w,init_cross,final_cross,mid_level] = pulsewidth(scope_y_data,scope_x_data, 'Polarity', 'Positive');
+[w2,init_cross2,final_cross2,mid_level2] = pulsewidth(scope_y_data2,scope_x_data, 'Polarity', 'Positive');
+
+move_dist = 295;
+plot((t(move_dist:end)-move_dist*1/4E9)*1e6-7.95,scope_y_data(move_dist:end)+8,'g');
+
+num_samp = length(scope_y_data);
+t_max = num_samp*(1/f_samp); 
+t=linspace(0,t_max/2,num_samp);
+
+move_dist = 178;
+plot((t(move_dist:end)-move_dist*1/4E9)*1e6-7.95,scope_y_data2(move_dist:end)+12,'k');
+
+ylabel('Amplitude')
+xlabel('Time (s)')
+xlim([80 81])
+xlabel("Time (\mu sec)")
+set(gca,'ytick',[])
+legend(["10M div 8", "ref div 8", "10M div 4", "ref div 4"])

### 2019-01-22 tex_files/placeholder.tex

--- 2019-01-08 tex_files/placeholder.tex
@@
 \usepackage{setspace}
+\geometry{verbose,tmargin=3cm,bmargin=3cm,lmargin=3.5cm,rmargin=2.4cm}
+%\usepackage{fancyhdr}
+%\pagestyle{fancy}
+%\setcounter{secnumdepth}{3}
+\usepackage[british]{babel}
 
-\geometry{verbose,tmargin=3cm,bmargin=3cm,lmargin=3.5cm,rmargin=2.4cm}
 \usepackage{amsmath}
@@
 
+			\usepackage{titlesec}
+
+			\titleformat{\chapter}[display]
+			{\normalfont\huge\bfseries}{\chaptertitlename\ \thechapter}{20pt}{\Huge}
+
+			% this alters "before" spacing (the second length argument) to 0
+			\titlespacing*{\chapter}{0pt}{0pt}{32pt}
+
+			\usepackage{etoolbox}
+			\makeatletter
+			\patchcmd{\chapter}{\if@openright\cleardoublepage\else\clearpage\fi}{}{}{}
+			\makeatother
+
+			\usepackage[unicode=true,pdfusetitle,
+			bookmarks=true,bookmarksnumbered=false,bookmarksopen=false,
+			breaklinks=true,pdfborder={0 0 0},backref=page,colorlinks=false]
+			{hyperref}
+
 \definecolor{mygreen}{rgb}{0,0.6,0}
@@
 \author{Conor Dooley}
-\maketitle
+\begin{titlepage}
+
+%	\pagenumbering{gobble}
+
+	\setstretch{1.25}
+
+	\begin{spacing}{3}
+
+		\noindent \begin{center}
+
+			{\huge{}ADPLL Network on an FPGA}
+			\par\end{center}{\huge \par}
+	\end{spacing}
+
+	\bigskip{}
+
+	\begin{center}
+		{\Large{}Conor Dooley}
+	\end{center}
+
+	\vspace{1.25cm}
+
+	\begin{center}
+		%\includegraphics[width=0.15\paperwidth]{ucd_crest.pdf}\vspace{0.75cm}
+		\includegraphics[width=0.2\paperwidth]{UCD_crest.pdf}\vspace{0.25cm}
+	\end{center}
+
+	\bigskip{}
+
+	\begin{center}
+		{\huge{}Masters Project Interim Report }
+	\end{center}
+
+	\vspace{1.cm}
+
+	{\Large{}\bigskip{}
+		\bigskip{}
+	}{\Large \par}
+
+	\noindent \begin{center}
+		\begin{tabular}{lll}
+			\textbf{\large{}Research Supervisors:} & \qquad{}\qquad{} & {\large{}Dr Elena Blokhina \& Brian Mulkeen}\\
+		\end{tabular}
+
+		\par\end{center}
+
+	\vspace{1cm}
+	\noindent \begin{center}
+		{\large{}January  2019}
+		\par\end{center}{\large \par}
+
+\end{titlepage}
+\setcounter{page}{1} 
+\pagenumbering{roman}
+\setstretch{1.05}
+
+\addcontentsline{toc}{chapter}{\contentsname}
+
+\tableofcontents{}
+
+\pagebreak{}
+
+%\begin{dedication}
+%Dedicated to google and wikipedia  
+%\end{dedication}
+
+%\setcounter{page}{1} 
+%\pagenumbering{roman}
+
+%\addtocontents{toc}{\protect{\pdfbookmark[0]{Abstract}{toc}}}
+%\pagestyle{empty}
+%\addcontentsline{toc}{chapter}{Abstract}
+%\phantomsection
+
+\pagestyle{plain}
+
+\setstretch{1.3}
 \doublespacing
+\pagenumbering{arabic}
+\setcounter{page}{1} 
 \chapter{Background Review}
 \section{Brief Overview}
-In a world where the demand for high performance hand-held computing devices continues to grow and the prevalence of ``smart'' devices is increasing the problem of maintaining the steady gain in performance of Systems-On-Chip remains at the forefront.
+In a world where the demand for high performance hand-held computing devices continues to grow and the prevalence of ``smart'' devices is increasing, the problem of maintaining the steady gain in performance of Systems-On-Chip remains at the fore.
 %TODO cite?
-The main drivers of performance in Systems-On-Chip are the number of transistors on a chip, the clocking frequency and parallelism.
-However the latter will not be addressed in this assessment, partly as the performance of the paralellised operation is largely governed by the number of transistors and clock frequency to begin with.
-%TODO Excessive?
-As the number of transistors on a chip has increased roughly following Moore's Law the increase in clock frequency has not been able to follow a similar linear trajectory, which clock frequency having remained roughly equivalent for the last number of years \cite{ross2008cpu}.
+The main drivers of performance in Systems-On-Chip are the number of transistors on a chip and the frequency at which it is clocked.
+As the number of transistors on a chip has increased roughly following Moore's Law, the increase in clock frequency has not been able to follow a similar linear trajectory, having remained roughly equivalent for the last number of years \cite{ross2008cpu}.
 
-This plateauing of clock frequency has been caused by the high power consumption due to the demands placed by the global distribution of a high frequency clock, which is often the single biggest consumer of power on the chip.
-%TODO cite (can use tiwari1998reducing)
-In a world where low power devices are desirable, especially with the growing demand of Internet-Of-Things where high power consumption goes directly against one of the key pillars of the technology.
+This plateauing of clock frequency has been caused by high power consumption due to the demands placed by the global distribution of a high frequency clock, often the single biggest consumer of power on the chip \cite{tiwari1998reducing}.
+In a world where low power devices are desirable, especially with the growing demand of the Internet-Of-Things where high power consumption goes directly against one of the key pillars of the technology.
 %TODO Pillars of what?
 
-In digital systems two main approaches are used when designing the clocking system.
-In both cases the chip is broken down into small areas in which all transistors are clocked synchronously, with the size constrained by the ability to deliver a quality clock signal to all transistors.
-The first of these methods is Globally Synchronous Locally Synchronous, or GSLS, where all of these sub areas on the chip operate synchronously with each other.
-In practice however this is very difficult to achieve as extremely high precision is required across the entire area of the chip, and doing so leads to high power consumption.
-In contrast a Globally Asynchronous Locally Synchronous clock delivery system ensures that in each sub division of the chip clocking is synchronous however the ``local'' areas are independent from each other.
-This reduces the clocking system's complexity and thus the power consumption and chip area used at the expense of communication speed between blocks.
+In digital systems, two main approaches are used when designing the clocking system.
+In both cases, the chip is broken down into small areas in which all transistors are clocked synchronously, with the size constrained by the ability to deliver a quality clock signal to all transistors.
+The first of these methods is Globally Synchronous Locally Synchronous, or GSLS, where all of these sub areas on the chip are synchronised with each other.
+In practice, however, this is very difficult to achieve, as extremely high precision is required across the entire area of the chip, and doing so leads to high power consumption.
+In contrast in a Globally Asynchronous Locally Synchronous clock delivery system the ``local'' areas are not synchronised with other.
+This reduces the clocking system's complexity and thus the power consumption and chip area used, at the expense of communication speed between blocks.
 %TODO citation?
-A GSLS system however has the advantages of deterministic behaviour and greater rates of communication between clocking areas and as such remains a desirable system design.
+A GSLS, system, however has the advantages of deterministic behaviour and greater rates of communication between clocking areas and, as such, remains a desirable system design.
 A number of methods which deliver GSLS clocking exist at present such as clock trees as well as emerging technologies such as ADPLL networks.
@@
 	\label{fig:eldar_why_precise_clocking}
-\end{figure}\newline
+\end{figure}\\
+
 A lack of synchronisation between the clock edges will manifest itself as a time difference between the clocking events at both registers, $\Delta T = t^i_k - t^f_k$.
-$\Delta T$ is considered to be ergodic and can be described by an average deviation called skew and random process that is normally modelled as a Gaussian random variable.
-If $\Delta T$ is negative this reduces the time available for the intervening combinatorial network thereby having the same effect as a reduction in clocking frequency.
-The most common sources of clock error are caused by mismatches which usually stem from production such as differences in the length of clocking paths, buffer delays or in the parameters of either active or passive components in the clock distribution network.
+$\Delta T$ is considered to be ergodic and can be described by an average deviation called skew and random process, normally modelled as a Gaussian random variable.
+If $\Delta T$ is negative this reduces the time available for the intervening combinatorial network thereby, having the same effect as a reduction in clocking frequency.
+The most common sources of clock error are caused by mismatches which usually stem from production, such as differences in the length of clocking paths, buffer delays or in the parameters of either active or passive components in the clock distribution network.
 These all manifest themselves as skew, while the noise in active components will appear as jitter in the clock signal.
@@
 A number of traditional solutions exist which provide GSLS clocking systems, using a variety of techniques.
-The most simple of these use the clock distribution network's symmetry in order to deliver the clock in phase at all locations in the chip and are named according to their geometry with the main variants being branch-trees, X-trees or H-trees.
-%TODO insert figure
+The most simple of these use the clock distribution network's symmetry in order to deliver the clock in phase, at all locations in the chip and are named according to their geometry with the main variants being branch-trees, X-trees or H-trees.
 \begin{figure}[h]
 	\centering
-	\includegraphics[scale=0.5]{../eldar_why_precise_clocking}
-	\caption{Data Flow in a Clocked System \cite{zianbetov2013distributed}.}%TODO insert figure
+	\includegraphics[scale=0.2]{../trees}
+	\caption{H and Branch Tree Clock Distribution Systems.}
 	\label{fig:trees}
 \end{figure}
-While on the surface these appear simple the task of obtaining an exact matching is, in practise, the limiting factor in this design.
-Even if the clock distribution system is geometrically symmetrical production mismatches in either active or passive components will lead to a skew that varies from part to part.
-In order to minimise the impact of production tolerances the dimensions of components in the distribution network can be increased, thus reducing the relative variation possible.
+
+While on the surface these appear simple, the task of obtaining an exact matching is, in practice, the limiting factor in this design.
+Even if the clock distribution system is geometrically symmetrical, production mismatches in either active or passive components will lead to a skew that varies from part to part.
+In order to minimise the impact of production tolerances, the dimensions of components in the distribution network can be increased, thus reducing the relative variation possible.
 However this has the impact of increasing the power consumption of the distribution network \cite{tiwari1998reducing}.
-Designs do exist in which the electrical lines used in the tree network are replaced by waveguides for optical signals, which negates the electrical losses however this technique suffers from being a newer technology and as such is expensive to realise \cite{haurylau2006chip}.
+Designs do exist in which the electrical lines used in the tree network are replaced by waveguides for optical signals, which negates the electrical losses, however, this technique suffers from being a newer technology and as such is expensive to realise \cite{haurylau2006chip}.
 
@@
 According to Abdelhadi \textit{et al} (2010) clock meshes ``\textit{achieve low and deterministic skew, low skew variations, and low jitter}'', all desirable characteristics for a clock distribution system.
-However they dissipate more power due to extra capacitive loading due to vast number of lines required to form the grid and similarly to tree distribution networks suffer from potential mismatch in production and alleviation through increasing dimensions will again lead to higher power consumption \cite{abdelhadi2010timing}.
+However they dissipate more power due to extra capacitive loading attributable to vast number of lines required to form the grid Similarly tree distribution networks suffer from potential mismatch in production and alleviation through increasing of distribution system dimensions will again lead to higher power consumption \cite{abdelhadi2010timing}.
 \begin{figure}[h]
 	\centering
-	\includegraphics[scale=0.5]{../eldar_why_precise_clocking}
-	\caption{Data Flow in a Clocked System \cite{zianbetov2013phd}.}%TODO insert figure
+	\includegraphics[scale=0.6]{eldar_mesh}
+	\caption{Mesh Clock Distribution System \cite{zianbetov2013phd}.}
 	\label{fig:mesh}
@@
 
-In a clocking distribution system using a tree type network the skew due to the manufacturing process can be partly accounted for by means of a skew compensation system, of which there are two main types.
-These types are defined by the location of the control mechanism with those having one controller located at the clock source being known as ``centralised'' methods and those with multiple controllers in the individual clocking areas known as ``decentralised''.
-In a centralised skew compensation circuit the skew across the chip is calculated by the central controller which then manipulates the distribution network in order to deliver a more in-phase clock around the chip.
-This however adds complication to a circuit which already is a large consumer of power and as such increases the power usage.
+In a tree type distribution system, skew due to the manufacturing process can be partly accounted for by means of a skew compensation system, of which there are two main types.
+These types are defined by the location of the control mechanism with those having one controller located at the clock source, known as ``centralised'' methods, and those with multiple controllers in the individual clocking areas known as ``decentralised''.
+In a centralised skew compensation circuit, the skew across the chip is calculated by the central controller which then manipulates the distribution network in order to deliver a more in-phase clock around the chip.
+This, however, adds complication to a circuit which already is a large consumer of power and as such increases the power usage.
+
 As the name suggest a decentralised skew compensation technique delegates this responsibility to the individual clock regions.
-For example Yamashita \textit{et al} designed a system in which each clocking area or ``leaf node'' contains a partial clock tree.
-Each of these ``leaves'' is able to compare its clock phase to the neighbouring node and based on the result tune an adjustable delay buffer \cite{yamashita2005dynamic}.
-While this method can compensate for process, voltage and temperature variation it does not address the power consumption due to the delivery of a high frequency clock across the entire chip area.
+For example, Yamashita \textit{et al} (2005) designed a system in which each clocking area or ``leaf node'' contains a partial clock tree.
+Each of these ``leaves'' is able to compare its clock phase to the neighbouring node, and based on the result, tune an adjustable delay buffer \cite{yamashita2005dynamic}.
+While this method can compensate for process, voltage and temperature variation, it does not address the power consumption due to the delivery of a high frequency clock across the entire chip area.
 
 \section{Multi-oscillator Designs}
-The designs described previously are all similar in that they have a single central oscillator that provides the clock for all areas of the chip, whereas the following methods attempt to synchronise multiple oscillators, each of which provides the clock for a single clocking area.
-The main advantage of a multi-oscillator design is that as each clocking area has its clock created locally there is no problem with signal degradation.
+The designs described previously, are all similar in that they have a single central oscillator that provides the clock for all areas of the chip, whereas the following methods attempt to synchronise multiple oscillators, each of which provides the clock for a single clocking area.
+The main advantage of a multi-oscillator design, is that as each clocking area has its clock created locally, there is no problem with signal degradation.
 Similar to a decentralised skew compensation scheme comparisons are only made to neighbouring blocks and as such there is reduced power loss due to transmission.
@@
 One such method is a network of oscillators as in Figure \ref{fig:mizuno1998noise} which uses coupled PLLs to generate local clocks.
-The advantage of this method is that only the control voltage $v_c$ needs delivery to all areas of the chip thus alleviating the need for a power hungry distribution circuit while also being more noise immune than the transmission of a high frequency clock.
+The advantage of this method is that only the control voltage $v_c$ needs delivery to all areas of the chip, thus alleviating the need for a power hungry distribution circuit, while also being more noise-immune than the transmission of a high frequency clock.
 However this design still suffers from clock variation as all VCOs are fed the same control voltage, and this is acknowledged by the authors:
@@
 \end{quotation}
-This design also only compares the phase of a single branch of the distribution network with the reference in order calculate the control voltage which again may contribute to a degradation in clock synchronisation.
+This design also only compares the phase of a single branch of the distribution network with the reference, in order calculate the control voltage which again may contribute to a degradation in clock synchronisation.
 %TODO why
@@
 Another potential multi-oscillator design synchronises the oscillators using the phase relationship with the oscillators in the neighbouring clocking area.
-Once again this negates the requirement for a global distribution structure and comparisons need only be made with neighbouring areas, and as a phase lock loop is being used only a divided down version of the clock is required.
-This in turn means the hardware transporting the divided clock signal to the phase comparator has significantly lower requirements placed on it thus lowering the power consumption.
+Once again, this negates the requirement for a global distribution structure and comparisons need only be made with neighbouring areas.
+As a phase lock loop is being used only a divided down version of the clock is required.
+This in turn means the hardware transporting the divided clock signal to the phase comparator, has significantly lower requirements placed on it, thus lowering the power consumption.
 Pratt and Nguyen initially proposed this design in their paper published in 1995 entitled ``\textit{Distributed Synchronous Clocking}'' in which they propose a Cartesian grid of clocking areas, each with their own PLL, which has become known as a PLL Network \cite{pratt1995distributed}.
@@
 \section{ADPLL Networks}
-As a PLL network is an analog circuit its integration in a modern IC is a barrier to usage and it has not been used in any commercial designs \cite{zianbetov2013distributed}.
+As a PLL network is an analog circuit, its integration in a modern IC is a barrier to usage and it has not been used in any commercial designs \cite{zianbetov2013distributed}.
 An alternative design that is more suitable for current fabrication techniques eschews from using analog components and instead implements the network of controlled oscillators using only digital circuitry, hence the name All-Digital PLL.
@@
 %TODO fix phd references
+
 In this design the oscillators are coupled using digital phase comparators which attempt to measure the phase error between two oscillators with each non edge node being connected to four neighbours.
@@
 \end{figure}
-Looking at the design of a given node it is notable that the function carried out by the Error Combiner  is akin to a average therefore as mentioned by Pratt and Nyugen there is potential to settle in a stable ``mode'' in which the oscillators are not synchronised.
-In their paper they present a method where initial start-up is performed uni-directionally and once all nodes are close full connectivity can be restored \cite{pratt1995distributed}.
-In creating an entirely digital system Zianbetov and Shan could exploit the reconfigurability and and implement this system at start-up.
+Looking at the design of a given node it is notable that the function carried out by the Error Combiner is akin to a average, therefore as mentioned by Pratt and Nyugen, there is potential to settle in a stable ``mode'' in which the oscillators are not synchronised.
+In their paper, they present a method where initial start-up is performed uni-directionally and, once all nodes are close to alignment, full connectivity can be restored, however this was not viable in an analog system \cite{pratt1995distributed}.
+In creating an entirely digital system, Zianbetov and Shan could exploit the reconfigurability and and implement this system at start-up.
 
@@
 	\centering
-	\includegraphics[scale=0.7]{mulkeen_pll}
-	\caption{Block Diagram of a Phase Lock Loop, \textit{Wireless Systems Notes}, Mulkeen (2017).}
+	\includegraphics[scale=0.5]{mulkeen_pll}
+	\caption{Block Diagram of a Phase Lock Loop, \textit{Wireless Systems Notes}, B. Mulkeen (2017).}
 	\label{fig:mulkeen_pll}
@@
 The decisions made in the design of the digitally controllable oscillator, or DCO, determine many of the other ADPLL parameters.
-While tuning range and centre frequency as well as linearity carry over from the analog counterpart a DCO also has a frequency step which in combination with the bit width of the control signal determines the range over which the oscillator can be tuned.
+While tuning range and centre frequency as well as linearity carry over from the analog counterpart, a DCO also has a frequency step which in combination with the bit width of the control signal determines the range over which the oscillator can be tuned.
 Figure \ref{fig:my_ring} illustrates a basic ring oscillator design.
-In a ring oscillator is an inherently unstable circuit composed of an odd number of inveters.
-where the period of oscillation is given by twice entire chain's delay.
+A ring oscillator is an inherently unstable circuit composed of an odd number of inverters where the period of oscillation is given by twice entire chain's delay.
 The frequency of operation can then be set by modulating the length of the chain by means of \texttt{f\_select}.
-The main impact of output frequency quantisation is that only frequencies which are integer multiples of the frequency step away from the centre can be easily reproduced, with intermediate values only obtainable in a manner akin to Fractional-N synthesis with the control code toggling back and forth.
+The main impact of output frequency quantisation is that only frequencies which are integer multiples of the frequency step away from the centre frequency can be easily reproduced, with intermediate values only obtainable in a manner akin to Fractional-N synthesis with the control code toggling back and forth.
 This acts as a source of jitter in the system.
@@
 	\centering
-	\includegraphics[scale=0.35]{../inverter_chain}
+	\includegraphics[scale=0.275]{../inverter_chain}
 	\caption{Basic Ring/Inverter Chain Oscillator.}
@@
 \subsection{Digital Phase Detector}
-Once again quantisation impacts the Phase Detector, as rather than a continuous output the phase detector in an ADPLL has a finite number of output values thus limiting the accuracy of the phase detector.
-A second form of quantisation is also present as unlike an analog system a digital phase detector does not provide continuous data in the time domain either, instead relying on sampling.
-At its most basic a digital phase comparator may only output an indication of which signal is leading, a design known as a Bang-Bang Detector, which can be constructed using a single D Flip Flop.
-As the output only has two levels the resultant word is only 1 bit wide and as such limits the range over which the output frequency can be controlled.
+Once again quantisation impacts the Phase Detector, as rather than a continuous output the phase detector in an ADPLL has a finite number of output values, thus limiting the accuracy of the phase detector.
+A second form of quantisation is also present, as unlike an analog system, a digital phase detector does not provide continuous data in the time domain either, instead relying on sampling.
+At its most basic, a digital phase comparator may only output an indication of which signal is leading, a design known as a Bang-Bang Detector, which can be constructed using a single D Flip Flop.
+As the output only has two levels the resultant word is only 1 bit wide and as such, limits the range over which the output frequency can be controlled.
 More complex designs such as that in Figure \ref{fig:shan_bb_pd}, implemented by Shan, build on this by measuring the time difference between edges of the signals using a Time-to-Digital Converter, or TDC, in his case using a tapped delay line \cite{shan2014phd}.
@@
 In the case of an ADPLL node in a network the input of this filter is a weighted sum of two to four phase errors, each corresponding to a neighbouring node.
-In a digital system the proportional section is implemented by a simple multiplier whereas the integral path is a combination of a multiplier and a delayed summation stage.
+In a digital system the proportional section is implemented by a simple multiplier, whereas the integral path is a combination of a multiplier and a delayed summation stage.
 The value of these gains determine the response and stability of the oscillator system.
@@
 It has been found by Koskin \textit{et al} (2018) that stable operation can be achieved when the integral gain, $k_i$, is less than half the proportional gain, $k_p$ \cite{koskin2018generation}.
-In the same study a range of values were found which would produce low jitter operation of the network.
-As these value are all less than one the filter must implement fixed point arithmetic in an effort to maintain the simplicity of the clock distribution network, rather than incurring the complexity penalty of floating point calculations.
+In the same study a range of values was found which would produce low jitter operation of the network.
+As these values are all less than one, the filter must implement fixed point arithmetic in an effort to maintain the simplicity of the clock distribution network, rather than incurring the complexity penalty of floating point calculations.
 
@@
 The planned contribution of this project is to create an FPGA based prototyping platform for the testing of network parameters and architectural variations.
-An FPGA as a prototyping tool for networks was previously utilised by Zianbetov and Shan as the final stage before the creation of a custom chip as it allowed for ``a partial validation of the ASIC design, through a verification of the essential functional features of the designed clocking network \cite{zianbetov2013phd,shan2014phd}.
-They however note that there are some fundamental differences between the two environments in particular that the FPGA based prototype is unable to create the mixed signal circuits used in an ASIC resulting in alternative designs being mandated for certain components such as the Time-to-Digital Converter.
-Similarly most commericially available FPGA boards were noted to have clock frequencies well below the gigahertz levels possible in ASIC designs.
-In spite of these problems the FPGA as a prototyping solution is valuable as it allows for hardware based testing of architectures and the simulation of non trivial circuit behaviours.
+An FPGA as a prototyping tool for networks was previously utilised by Zianbetov and Shan as the final stage before the creation of a custom chip, as it allowed for ``a partial validation of the ASIC design, through a verification of the essential functional features of the designed clocking network'' \cite{zianbetov2013phd,shan2014phd}.
+They note however, that there are some fundamental differences between the two environments, in particular that the FPGA based prototype is unable to create the mixed signal circuits used in an ASIC, resulting in alternative designs being mandated for certain components such as the Time-to-Digital Converter.
+Similarly most commercially available FPGA boards were noted to have clock frequencies well below the gigahertz levels possible in ASIC designs.
+In spite of these problems, the FPGA as a prototyping solution is valuable as it allows for hardware-based testing of architectures and the simulation of non trivial circuit behaviours.
 
-The novelty of this particular implementation is that in the case of Zianbetov and Shan only a small network was simulated on an FPGA and at a frequency close to 50 kHz, however this project will attempt to significantly increase the frequency of operation to the MHz range as well as expanding the size of the network, from 16 oscillators to potentially 100.
-This is done with the aim of assisting work being undertaken by a number of researchers in UCD who are working on aspects of ADPLL networks who would stand to benefit from such a platform.
+The novelty of this particular implementation is that in the case of Zianbetov and Shan only a small network was simulated on an FPGA and at a frequency close to 50 kHz, however this project will attempt to significantly increase the frequency of operation to the MHz range, as well as expanding the size of the network, from 16 oscillators to potentially 100.
+This is done with the aim of assisting work being undertaken by a number of researchers in UCD working on aspects of ADPLL networks who would stand to benefit from such a platform.
 
@@
 The first stage of this project was the creation of an ADPLL design that would be used for the network.
-Due to limitations of the connections available on the FPGA evaluation board used to this point, a Digilent/Xilinx Nexys 4, signals at frequencies significantly beyond 5 MHz suffered from major degradation as the connections were not designed for such frequencies.
+Due to limitations of the connections available on the FPGA evaluation board used to this point, a Digilent/Xilinx Nexys 4, outputting signals at frequencies significantly beyond 5 MHz, suffered from major degradation as the connections were not designed for such frequencies.
 Thus a target frequency was set at 5 MHz while this board was in use.
-After initial research into potential oscillator designs and what would be feasible given the constraints of the particular platform design began with the aforementioned oscillator.
+After initial research into potential oscillator designs and what would be feasible given the constraints of the particular platform, design began with the aforementioned oscillator.
+
 \section{Controllable Oscillator}
@@
 The first of these was the ring oscillator, constructed from an odd number of inverters strung together in a chain \cite{predraig}.
-In this case the frequency of operation is determined by the number of oscillators in the chain as the delay through the individual inverters determines the half period of the signal.
+In this case the frequency of operation is determined by the number of oscillators in the chain, as the delay through the individual inverters determines the half period of the signal.
 This period can then be modified by adding or removing inverters in a manner that maintains the odd number of total inverters.
-The propagation time through an individual inverter is dependant on the layout of the ring and thus this design must be used with some care.\\
-The second option is one that I was aware of in advance this project's beginning.
+The propagation time through an individual inverter is dependant on the layout of the ring and thus this design must be used with some care.
+
+The second option is one that I was aware of in advance of the beginning of this project.
 A phase accumulator makes use of the overflow property of a counter, and has a frequency determined by the width of the counter and the frequency at which it is clocked.
-The oscillator operates as follows: Once per clock cycle the counter will increment by a number $k$ until overflow is reached at $2^n-1$ where $n$ is the bit width of the counter.
+The oscillator operates as follows: once per clock cycle the counter will increment by a number $k$ until overflow is reached at $2^n-1$ where $n$ is the bit width of the counter.
 By controlling the magnitude of $k$ added on each cycle of the driving clock the frequency of this oscillator can be adjusted.
-The most significant bit of this counter can form the clock signal as it has an approximately 50\% duty cycle.\\
-As it is not reliant on the clock of FPGA the ring oscillator design is a more interesting design in the context of simulation, but the phase accumulator is significantly easier to manage due to being more easily simulatable.
+The most significant bit of this counter can form the clock signal, as it has an approximately 50\% duty cycle.
+
+As it is not reliant on the clock of FPGA, the ring oscillator design is a more interesting design in the context of simulation, but the phase accumulator is significantly easier to manage due to being more easily simulatable.
 The decision was made to create both designs in order to keep options open.\\
-The only other option I managed to come across for an oscillator on an FPGA was using Xilinx proprietary \texttt{IODELAY} blocks, which have a permanently programmable delay at runtime \ref{IODELAY}.
+The only other option I managed to come across for an oscillator on an FPGA was using Xilinx proprietary \texttt{IODELAY} blocks, which have a permanently programmable delay at runtime \cite{iodelay}.
 This could be coupled with an odd number of inverters in order to implement a quasi ring oscillator.
-However as these are IO blocks their usage is not particularly conducive to the creation of a network of ADPLLs.\\
+However as these are IO blocks, their usage is not particularly conducive to the creation of a network of ADPLLs.
+
 The phase accumulator was designed to have a bit width of 12, all of which could be set as the control code.
-As the maximum frequency of operation that had no timing violations was 258 MHz the frequency step of this oscillator design was 62.988 kHz and before integration into an ADPLL had a frequency range from 62.988 kHz to just under half the clock frequency at 128.94 MHz.
-Figure \ref{fig:my_pa} details the simple construction of this oscillator.\\
+As the maximum frequency of operation that had no timing violations was 258 MHz, the frequency step of this oscillator design was 62.988 kHz and before integration into an ADPLL had a frequency range from 62.988 kHz to just under half the clock frequency at 128.94 MHz.
+Figure \ref{fig:my_pa} details the simple construction of this oscillator.
 \begin{figure}[h]
 	\centering
-	\includegraphics[scale=0.5]{../phase_accum}
+	\includegraphics[scale=0.4]{../phase_accum}
 	\caption{Phase Accumulator Block Diagram.}
@@
 \end{figure}
-As originally implemented for testing purposes the inverter ring oscillator had a propagation time per inverter of 315 picoseconds, thus each inverter would add 0.63 nano seconds to the total period, and had 4 control bits.
-In order to obtain a 5 MHz output from this type of ring oscillator a ring of of 317 inverters is required, and to set this as the centre point of the frequency range the total number of inverters was 325.
+
+As originally implemented for testing purposes, the inverter ring oscillator had a propagation time per inverter of 315 picoseconds, thus each inverter would add 0.63 nano seconds to the total period, and had 4 control bits.
+In order to obtain a 5 MHz output from this type of ring oscillator a ring of of 317 inverters is required, and to set this as the centre point of the frequency range, the total number of inverters was 325.
 At 5 MHz the frequency step of this design was calculated as:
@@
 \end{equation*}
-%TODO look into more detail on yoke with later work.
 \begin{figure}[h]
 	\centering
-	\includegraphics[scale=0.35]{../ring_osc}
+	\includegraphics[scale=0.275]{../ring_osc}
 	\caption{Simplified Ring Oscillator Block Diagram.}
@@
 \section{Phase Detector}
-Moving along the chain the next block then is the phase detector, or more accurately phase comparator.
-At this stage in development this block just compares the oscillator's output with that of just one reference.
+Moving along the chain the next block is the phase detector, or more accurately, phase comparator.
+At this stage in development this block compares the oscillator's output with that of just one reference.
 In the future this block will be followed up with one which performs a weighted sum of all neighbouring nodes.
@@
 The eventual design I chose has two main parts, firstly an up/down counter performs the role of the Time-to-Digital converter while also indicating the sign of the phase difference.
-In order to control the count, and indicate the direction in which to do so, a state machine determines which edge occurs first.
-In order to circumvent the possibility of a metastability condition being reached the incoming signals are synchronised to the higher frequency FPGA clock prior to their entry into the phase detector.
-From there the state machine can operate as designed, in accordance with the state transition diagram in Figure \ref{fig:my_state_machine}.
+In order to control the count and indicate the direction in which to do so, a state machine determines which edge occurs first.
+In order to circumvent the possibility of a metastability condition being reached, the incoming signals are synchronised to the higher frequency FPGA clock, prior to their entry into the phase detector.
+From there, the state machine can operate as designed, in accordance with the state transition diagram in Figure \ref{fig:my_state_machine}.
 \begin{figure}[h]
@@
 \end{figure}
+
 The state machine behaves as follows: Starting in the "Waiting" state, if the reference has a rising edge first it is said to be leading the oscillators output and the sign of the phase error is positive.
-Alternatively if the opposite occurs and the first rising edge is seen on the oscillator output the phase error is given a negative sign and the TDC counts downwards.
-In either case this continues until a rising edge occurs on the other signal which stops the count, updates the phase error output of the block and finally clears the counter.
-Between "Copy \& Clear" events the output of the block is held constant.
+Alternatively if the opposite occurs and the first rising edge is seen on the oscillator output, the phase error is given a negative sign and the TDC counts downwards.
+In either case, this continues until a rising edge occurs on the other signal which stops the count, updates the phase error output of the block and finally clears the counter.
+Between "Copy \& Clear" events, the output of the block is held constant.
 There is a second condition that can stop the count and that is the occurrence of a falling edge of the signal that began the count.
-This does however limit the rate of lock acquisition in the case where the phase difference is greater than 180$^\circ$ or the frequency difference is significant.
-The operation however is the same as a conventional method when either locked or within a normal operating tolerance of being locked.\\
-In order to represent both the count value and sign in one value the counter operates in twos complement which allows the output be used directly in subsequent blocks.
-The counter itself has a bit width of 8 thereby having a range of -128 to 127 which at 5 MHz gives a phase resolution of:
+This does however limit the rate of lock acquisition in the case where the phase difference is greater than 180$^\circ$ or where the frequency difference is significant.
+The operation, however, is the same as a conventional method when either locked or within a normal operating tolerance of being locked.
+
+In order to represent both the count value and sign in one value, the counter operates in twos complement which allows the output be used directly in subsequent blocks.
+The counter itself has a bit width of 8, thereby having a range of -128 to 127 which at 5 MHz gives a phase resolution of:
 \begin{align*}
@@
 \section{Loop Filter}
-As with Shan's design I used a PI controller as the loop filter in order to align with the typical technology and also benefit from the work of member of the research team in UCD \cite{shan2014phd,koskin2018generation}.
-%TODO eugene
-The bit with of the filter's input is once again 8 bits to match up with the input from the phase detector and outputs a 9 bit value in order to accommodate the summation of an 8 bit number from each of the paths.
+As with Shan's design, I used a PI controller as the loop filter, in order to align with the typical technology and to benefit from the work of members of the research team in UCD \cite{shan2014phd,koskin2018generation}.
+The bit with of the filter's input is once again 8 bits, to match up with the input from the phase detector and outputs a 9 bit value in order to accommodate the summation of an 8 bit number from each of the paths.
 I selected $k_p = 1$ and $k_i = 0.125$ in order to align with the region of stable coefficients discovered by Eugene \cite{koskin2018generation}.
-%TODO redo image of LF
 \begin{figure}[h]
@@
 	\label{fig:my_pi}
-\end{figure}\\
-Figure \ref{fig:my_pi} showing the RTL diagram of the filter shows the two distinct paths through the filter, the more simple of these being the proportional one.
+\end{figure}
+
+Figure \ref{fig:my_pi} showing the RTL diagram of the filter shows the two distinct paths through the filter, the simpler of these being the proportional one.
 The multiplication necessitates the increased width of the path in order to preserve the accuracy of the fixed point calculations, a width which is then cut down to an 8 bit value before being added to the output of the integral branch.
-The proportional gain value has a fixed point width of 2.1 allowing for values from -2 to 1.5, however in testing to this point only 1 and 0.5 have been used.
-The integral path is similar with the gain having a fixed width of 1.3, a range of -1 to 0.750, and for my testing set to 0.125.
-After the gain is applied via an accumulator with a delayed input clocked by the 258 MHz FPGA clock.
+The proportional gain value has a fixed point width of 2.1 allowing for values from -2 to 1.5, however, in testing to this point only 1 and 0.5 have been used.
+The integral path is similar, with the gain having a fixed width of 1.3, a range of -1 to 0.750, and for my testing set to 0.125.
+After the gain is applied an accumulator with a delayed input clocked by the 258 MHz FPGA clock implements the integration.
 Once again the output of this path is cut back down to an 8 bit value.\\
 The addition of two 8 bit values results in a 9 bit signed output from the phase detection block which is then combined with the bias to form the control code for the oscillator.
-As a higher control code leads to a larger frequency with the phase accumulator this combination is done by additions, whereas for the ring oscillator the relationship is inverse as we are controlling the period, so instead subtraction is used.
+As a higher control code leads to a larger frequency with the phase accumulator, this combination is done by additions, whereas for the ring oscillator the relationship is inverse as we are controlling the period, so instead subtraction is used.
 
@@
 Combining these blocks, in this case using the Phase Accumulator style oscillator, gives the ADPLL shown in Figure \ref{fig:my_adpll} on which I performed my initial testing and simulations.
+For ease of testing the PLL replicated the reference exactly, as this would facilitate easier testing of the lock and capture ranges of the design.
 \begin{figure}[h]
 	\centering
-	\includegraphics[scale=0.35]{../rtl}
+	\includegraphics[scale=0.275]{../rtl}
 	\caption{Phase Accumulator ADPLL RTL diagram.}
 	\label{fig:my_adpll}
+\end{figure}\\
+Once this testing had been completed with the phase accumulator based ADPLL testing progressed to the design which implemented an inverter chain and scaling from the reference using an eight times divider.
+This necessitated a change in the bit width of the control signal to match the 5 bit wide signal desired, thus leading to the design in Figure \ref{fig:my_ring_adpll}.
+\begin{figure}[h]
+	\centering
+	\includegraphics[scale=0.275]{../ro_rtl}
+	\caption{Ring Oscillator ADPLL RTL diagram.}
+	\label{fig:my_ring_adpll}
+\end{figure}\\
+ The Loop Filter is clocked at the generated output frequency, as this ensures it will be synchronised with the oscillator.
+This is important when using the inverter chain oscillator, in order to avoid metastability as the generated output clock may not be synchronised with the high frequency clock of the FPGA.
+
+\chapter{Results}
+Initially simulations were utilised to validate the behaviour of the individual blocks and of the system as a whole, while using the Phase Accumulator based design.
+From these simulations, I was able to test the performance of the ADPLL, while locked to a signal at 5 MHz and its behaviour during the acquisition period.
+A signal at 5 MHz lies close to midway between two increments of the phase accumulator and as such the oscillator must switch between two frequencies leading to the introduction of jitter.
+From simulations performed in Vivado locked to a signal at 5 MHz, I observed jitter of 3.491 ns, based on the differences between sequential periods, and a peak-to-peak worst case value of 9.299 ns.
+
+Similarly, I performed simulations of the locking performance of the design, starting with a deviation of 213 KHz and observing the time taken to achieve a lock.
+213 KHz represents an offset of three frequency steps, plus the difference between 5 MHz and the nearest integer frequency step.
+Figure \ref{fig:vivado_sim} contains the visual results of the tests, in which a repetitive pattern can be seen to begin after 20.91 $\mu$s indicating a lock as been acquired.
+\begin{figure}[h]
+	\centering
+	\includegraphics[scale=0.45]{../sim_locking_rect}
+	\caption{Locking Dynamics Simulation.}
+	\label{fig:vivado_sim}
 \end{figure}
 
-\chapter{Test Results}
+Identical tests were performed post-implementation, using a Digilent Nexys4 development board, in addition to measuring the lock and capture ranges of the design.
+Firstly in order to determine the baseline for jitter in the system, the oscillator was biased to 4.9761 MHz and sequential period jitter was observed to be 1.9689 ns.
+Connecting the remainder of the ADPLL and by varying the reference, the lock and capture ranges were found to be 4.19-12.76 MHz and 4.25-12.34 MHz respectively, when the bias point was 5 MHz.
+Somewhat surprisingly this is not centred on the bias point, suggesting a potential flaw in the system that needs to be addressed.
+A identical locking test was performed in the implemented design, Figure \ref{fig:impl}, where using the same method, a time to lock of 25.2 $\mu$s was observed.
+While locked jitter was better in both average period-to-period variation and in peak-to-peak, which can be attributed to the FPGA clock frequency not exactly matching that designed and as such the distance between 5 MHz and the nearest frequency step being reduced, thereby leading to less switching required to generate a 5 MHz signal.
+\begin{figure}[h]
+	\centering
+	\includegraphics[scale=0.45]{../sim_locking_rect}
+	\caption{Locking Dynamics Measurement.}
+	\label{fig:impl}
+\end{figure}
+
+Similarly tests were carried out with the Ring Oscillator based design at a bias point of 5 MHz.
+425 inverters, with 5 bit control, were used to obtain an output in the region of 5 MHz with the bias set to the midpoint which is close to 5 MHz.
+The expected lock range given the use of an otherwise identical design would be the entire range of 32 steps or 64 inverters which can be easily estimated:
+\begin{align*}
+	T_{inv} &= \frac{200\times 10^{-6}}{425-2\times 16} = 0.50891\textrm{~ns} \\
+	F_{min} &= \frac{1}{(425)\times 0.50891\textrm{~ns}} = 4.623\textrm{~MHz} \\
+	F_{max} &= \frac{1}{(425-2\times 32)\times 0.50891\textrm{~ns}} = 5.443\textrm{~MHz}
+\end{align*}
+However in testing the lock and capture ranges, both came out to be 4.86-5.15 MHz, so significantly below the estimates, which highlights an issue that requires investigation.
+
+Testing was also carried out of the ADPLL containing a divider, in both divide by 4 and 8 modes, to examine their impact on locking behaviour.
+The results of this testing are displayed in Table \ref{table:b_c_perf}.
+The jitter in this case is rather low, as the 10 MHz target frequency lies close to an integer increment of the phase accumulator design used for this test.
+Lock and capture ranges are significantly more symmetrical in these configurations than at a target of 5 MHz and, as expected, there is a decrease in range with the greater division values.
+\begin{figure}[h]
+	\centering
+	\includegraphics[scale=0.5]{../matlab/painintheass}
+	\caption{Testing of ADPLLs employing div4 \& div8.}
+	\label{fig:div}
+\end{figure}
+\begin{table}[!ht]
+	\begin{center} 
+		\begin{tabular}{c|c|c|c|c}           
+			& Period Jitter & Peak-Peak Jitter & Lock Range & Capture Range \\
+			div4 & 1.9858 ns & 6.4854 ns & 5.000-17.200 MHz & 5.620-16.700 MHz \\
+			div8 & 1.5702 ns & 6.1594 ns & 5.768-15.262 MHz & 5.928-14.216 MHz \\
+		\end{tabular}
+		\caption{Divider Comparison.}
+		\label{table:b_c_perf}
+	\end{center}
+\end{table}
 
 \chapter{Future Work}
-So far all work has been done using a single phase lock loop in order to have a suitable basis for the network and going forward this PLL will be used in that role.
-The first stage of network implementation will interlink two PLLs, with one of them also receiving input from the external reference.
-This test will serve as both validation of the phase coupling ability of the network as well as a test of the error summation block and the ability to lock from startup.
-For the sake of more realistic performance the inverter chain ADPLL will be used ahead of the Phase Accumulator based design.\\
+So far, all work has been done using a single phase lock loop in order to have a suitable basis for the network, and the first stage of network implementation will interlink two  of these PLLs, with one of them also receiving input from the external reference.
+This test will serve as both validation of the phase-coupling ability of the network as well as a test of the error summation block and the ability to lock from startup.
+For the sake of more realistic performance, the inverter chain ADPLL will be used ahead of the Phase Accumulator based design.
+With simple coupling implemented, the logical next stage is the implementation of a small network, possibly 2x2, before scaling up to a size suitable to the FPGA in use.
+At this point, analysis of measurement results can be undertaken and any refinements to the network can be made.
+Additionally, if time permits, the Phase Accumulator based design could be implemented and tested and comparisons made between both designs.
+
 A shortcoming of the current FPGA evaluation board is the inability to input or extract signals in the multiple MHz range without significant distortion, and this will be resolved with the arrival of an FPGA more suited to extracting and importing higher frequency signals.
-With an ability to extract higher frequency signals for measurement and analysis purposes the target frequency of 5 MHz can potentially be increased into the tens of MHz range.\\
-With simple coupling implemented the logical next stage is the implementation of a small network, possibly 2x2, before scaling up to a size suitable to the FPGA in use.
-At this point analysis of measurement results can be undertaken and any refinements to the network can be made.
-Additionally, if time permits, the Phase Accumulator based design could be implemented and tested and comparisons made between both designs.
+With an ability to extract higher frequency signals for measurement and analysis purposes, the target frequency of 5 MHz can potentially be increased into the tens of MHz range.
+
+As with the single ADPLL testing the important characteristics to look out for will be the lock and capture ranges of the network which should mirror that of the ADPLLs used in it's construction.
+Similarly the jitter of a number of clocks across the network will be examined in order, and any skew between the PLL outputs investigated.
+Also of interest will be the initial locking behaviour, in particular how quickly the network can be brought up from ``cold''.
 

### 2019-01-25 tex_files/Journal.tex
dsf

@@
 
+\section*{External Reference II}
+Problem was with the offset of the reference, once fixed things worked as expected. I previously had it zero centred rather than zero-min
+
+\section*{Gain Testing}
+Varied gains to find lower jitter operation. Required fixing of the loop filter truncation. Optimal gain changed significantly based on the configuration in use.
+
+\section*{Chained PLLs}
+Chained together 2 PLLs and the reference as so:
+\begin{equation*}
+	ref -> pll1 <=> pll2
+\end{equation*}
+required the creation of a weighted error combiner. Moved the PFD out of the PLL block for reuse between neighbours.
+
+\section*{Baby's First Floorplanning}
+Did things with floorplanning in order to ensure the ring oscillator doesnt change its frequency all the time.
+
 \end{document}

### 2019-02-04 tex_files/Journal.tex

@@
 
+\section*{Ring Osc}
+Tried to get this into a PLL and lock to the reference. Struggled for ~6 hours with it and never got close to locking.
+
+\section*{2x2 Network}
+Tried this w/ PA PLL,  couldnt get locking, got something that was "close" to locked with huge amounts of jitter.
+Will attempt to locate the issue tomorrow.
+Potentially there is an error in the PLL, or a missing error negation.
+
 \end{document}

### 2019-02-12 tex_files/Journal.tex

@@
 Will attempt to locate the issue tomorrow.
-Potentially there is an error in the PLL, or a missing error negation.
+Potentially there is an error in the PLL, or a missing error negation.\\
+Got the 2x2 PLL to lock at 10 MHz w/ div 8.
+Also designed in unidirectional startup.
+First got that to lock, in doing so I found 3!
+separate errors in the NetworkADPLL (or sub module) code.
+From there I joined toggled off and got locking.
+
+\section*{Ring ADPLL}
+Got this to lock today.
+Spent all day yesterday debugging, realised that having an 8 bit Phase Detector with a 5 bit Control Code is basically impossible.
+Switched to 4 bit Phase Detector, and this lead to locking quite quickly.
+As an aside, 401 not 301 inverters are required.
+
+\section*{Gains}
+Table \ref{table:sdf} contains the gains required to achieve locking, and in the case of the PA ADPLL, the most visually optimal jitter.
+I have not yet checked the ring on the better scope so the values may not in fact be optimal.
+\begin{table}[!ht]
+	\begin{center} 
+		\begin{tabular}{c|c|c|c|c}           
+			Configuration & $K_p$ Width & $K_p$ & $K_i$ Width & $K_i$ \\
+			PA ADPLL (5 MHz) & 2.1 & 001  & 1.4 & 0001 \\
+			PA ADPLL div8 & 1.4 & 1001 & 1.7 & 0001 \\
+			PA Chained & 1.4 & 0001 & 1.8 & 0001 \\
+			PA 2x2 & 1.5 & 1000 & 1.7 & 0001 \\
+			RO ADPLL & 1.5 & 0111 & 1.7 & 0001 \\
+			RO ADPLL div2/4 & 1001 & 1000 & 1.7 & 0001 \\
+		\end{tabular}
+		\caption{PA @ 10, RO @ 5}
+		\label{table:sdf}
+	\end{center}
+\end{table}
 

### 2019-02-18 tex_files/Journal.tex
ring osc network lock journal update

@@
 			PA Chained & 1.4 & 0001 & 1.8 & 0001 \\
-			PA 2x2 & 1.5 & 1000 & 1.7 & 0001 \\
+			PA 2x2 & 1.5 & 1110 & 1.7 & 0001 \\
 			RO ADPLL & 1.5 & 0111 & 1.7 & 0001 \\
 			RO ADPLL div2/4 & 1001 & 1000 & 1.7 & 0001 \\
+            RO 2x2 & 1.4 & 0100 & 1.7 & 0001 \\
 		\end{tabular}
@@
 
+\section*{Ring Osc Locking}
+After many trials and tribulations, including a situation where the ring adpll was locking to the inverse of the reference applied to it I finally managed to get the ring locked at 5 MHz with a 4 times divider.
+Turns out the shifting of $K_p$ and $K_i$ in the same direction did not help matters as I thought it might.
+In fact increasing $K_p$ while decreasing $k_i$ turned out to be the solution to my problems.
+Along the way I located an error where I was using 5'd16 as the bias which doesnt work with a subtraction of 15 to -16, as it would give overflow.
+Athough I had actually  made the counter the same in both directions so this didnt actually manifest itself as an issue.
+The lock range is now heavily impacted by the gains as I discovered when attempting to copy gains from the PA ADPLL.
+The key to solving it was going through all the fixed point arith.
+and checking that the widths made sense, which made my realise I had made a false assummtion that going from an 8 to a 5 bit detector/filter resulted in a div by eight times more at the exit of the filter, but in fact had the same div by 4.
+After discovering this I had it locking within 3 hours of work, when prior I had spent ~ 9 days on the calendar and ~30 hours working on the problem.
+Locking perofmance looks potentially better than the 10 MHz perf of the ADPLL (magnitude not \% for the sake of comp.
+However div is by 4 now not 8, and res is twice as good.) In the coming week I will be characterising the perf of these different designs to compare effectively.
 \end{document}

### 2019-02-21 matlab/a10M_quad_plot.m
~~~

@@ -1,4 +1,4 @@
-filename = 'scope_32.bin';
+filename = 'PA_10M_div4.bin';
 [~,scope_y_data] = importAgilentBin(filename,1);
 [scope_x_data,scope_y_data2] = importAgilentBin(filename,2);
 [w,init_cross,final_cross,mid_level] = pulsewidth(scope_y_data,scope_x_data, 'Polarity', 'Positive');
@@ -16,7 +16,7 @@ plot(t*1e6-7.95-0.09,scope_y_data2,'r');
 
 c = scope_y_data2;
 
-filename = 'scope_31.bin';
+filename = 'PA_10M_div4_1.bin';
 [~,scope_y_data] = importAgilentBin(filename,1);
 [scope_x_data,scope_y_data2] = importAgilentBin(filename,2);
 [w,init_cross,final_cross,mid_level] = pulsewidth(scope_y_data,scope_x_data, 'Polarity', 'Positive');

### 2019-02-21 matlab/process_scope_data.m

@@ -3,8 +3,8 @@ clear variables
 
 periods = [];
 for inc1 = 1:10
-    filename = char("scope_"+num2str(18+inc1)+".bin");
-    [scope_x_data(inc1,:),scope_y_data(inc1,:)] = importAgilentBin(filename,2);
+    filename = char("PA_10M_div4.bin");
+    [scope_x_data(inc1,:),scope_y_data(inc1,:)] = importAgilentBin(filename,1);
     
     [w,init_cross,final_cross,mid_level] = pulsewidth(scope_y_data(inc1,:),scope_x_data(inc1,:), 'Polarity', 'Positive');
     periods = [periods;getPeriods(init_cross)];
@@ -23,13 +23,14 @@ y_fft_plottable = real(y_fft_scaled(1:floor(num_samp/2)));
 
 num_datasets = 1;
 for loop_inc = 1:num_datasets
-    %% compute periods
+%% compute periods
 %     [w,init_cross,final_cross,mid_level] = pulsewidth(scope_y_data,scope_x_data, 'Polarity', 'Positive');
 % 
 %     [periods] = getPeriods(init_cross);
     
     period_mean(loop_inc) = mean(periods)*1E9; %in nanoseconds
     period_jitter_std(loop_inc) = std(periods)*1E9; %in nanoseconds
+    period_max_vs_min(loop_inc) = (max(periods)-min(periods))*1E9;
    % period_jitter_rms(loop_inc) = 
     
     % @~1000 samples 3.090 for 1 outside
@@ -40,7 +41,7 @@ for loop_inc = 1:num_datasets
 %         long_term_jitter(loop_inc) = periods(10000)-periods(1);
 %     end   
        
-    phase_shift = 0.10;
+    phase_shift = 0;
     [t,square_wave] = generateSquareWave(period_mean/1E9,num_samp,f_samp,phase_shift);
     
 %     [~,init_cross_ideal,~,~] = pulsewidth(square_wave,scope_x_data, 'Polarity', 'Positive');
@@ -72,11 +73,11 @@ if (num_datasets == 1)
     ylabel('')
     xlabel('Frequency (MHz)')
     
-    figure
-    plot(time_interval_error)
-    title('Time Interval Error')
-    ylabel('Error (ns)')
-    xlabel('Samples')
+    %figure
+    %plot(time_interval_error)
+    %title('Time Interval Error')
+    %label('Error (ns)')
+    %xlabel('Samples')
     
     figure
     histfit(periods)

### 2019-03-02 tex_files/Journal.tex
journal update

@@
 However div is by 4 now not 8, and res is twice as good.) In the coming week I will be characterising the perf of these different designs to compare effectively.
+
+\section*{Issues}
+This week I took some measurements of my 2x2 networks in order to justify the transition over to the ring based design.
+However I was unable to get satisfactory performance out of the ring design, with total jitter being approximately 50\% of the period and a very narrow band of gains which actually did anything.
+Speaking to Elena she suggested that the problem might be the 4 ns resolution on the detector but 1 ns period increments.
+However 4 ns is the lowest increment I could possibly get so an alternate design was required.
+I.E.
+returning to TDC design.
+
+\section*{Tapped Delay Line}
+I spent this week creating this design, starting with the logic that would determine the sign was the simple side of things, but much more challenging was the delay line.
+As an FPGA is not intended to implement systems like this liberal use of KEEP\_TRUE was required alongside many variations of logic to find one which had suitable timing to implement this system.
+I eventually found one which worked in simulation, so I intend to test it in the coming week.
+
 \end{document}

### 2019-03-07 matlab/a1_3.m
kdsjf

@@ -1,4 +1,4 @@
-filename = 'cdcd.bin';
+filename = 'scope_35.bin';
 [~,scope_y_data] = importAgilentBin(filename,1);
 [~,scope_y_data2] = importAgilentBin(filename,2);
 [~,scope_y_data3] = importAgilentBin(filename,3);

### 2019-03-07 matlab/plot_4_waves.m

@@ -1,4 +1,4 @@
-filename = 'tcope_2.bin';
+filename = 'scope_36.bin';
 [~,scope_y_data] = importAgilentBin(filename,1);
 [~,scope_y_data2] = importAgilentBin(filename,2);
 [~,scope_y_data3] = importAgilentBin(filename,3);

### 2019-03-09 tex_files/Journal.tex
jurnal

@@
 
+\section*{Twas the week before the midterm}
+This week was eventful but frustrating.
+I did not manage to make progress on any goals per say, but I discovered what was the cause behind the inverse locking that I had been seeing sometimes.
+Turns out it was being caused by the phase detector design which I was using being stable for very small gains and a small number of bit phase detection and if the falling edge of signal 1 toggled back and forth between rising on signal two I would get oscillation between positive and negative, or lead and lag to use the correct terms, and with the tiny gains making it actually stable in this state.
+Changing the detector to a more conventional style that did not terminate a measurement until the next rising edge on the opposite signal fixed this probem.
+The new issue now is that I cannot get low jitter locking, seeingly down to the behaviour of the loop filter.\\
+The loop filter appears to work correctly when the two paths are taken in isolation but when combining the paths it breaks down.
+Either through being too aggressive and doesnt lock or from the ki value being too small, and when it is too small the output of the loop filter toggles between a value of min and maxval thus creating a huge amount of jitter.
 \end{document}

### 2019-03-14 matlab/a1_3.m
issues with pll_22 for whatever reason. Others lock but offset in freq so overlap is poor.

@@ -1,4 +1,4 @@
-filename = 'scope_35.bin';
+filename = 'cdcd.bin';
 [~,scope_y_data] = importAgilentBin(filename,1);
 [~,scope_y_data2] = importAgilentBin(filename,2);
 [~,scope_y_data3] = importAgilentBin(filename,3);

### 2019-03-14 matlab/cd___anal.m

@@ -1,0 +1,30 @@
+filename = 'c_____0.bin';
+[scope_x_data,scope_y_data] = importAgilentBin(filename,3);
+
+f_samp = 4E9;
+num_samp = length(scope_y_data);
+t_max = num_samp*(1/f_samp); 
+t=linspace(0,t_max,num_samp);
+
+[~,init_cross4,~,~] = pulsewidth(scope_y_data,scope_x_data, 'Polarity', 'Positive');
+
+periods4 = getPeriods(init_cross4);
+
+std_period_4 = std(periods4);
+
+max_4 = max(periods4-mean(periods4));
+min_4 = min(periods4-mean(periods4));
+p2p_4 = max_4-min_4;
+
+clc
+string = num2str(std_period_4) + ", " + num2str(max_4) + ", " + num2str(min_4) + ", " + num2str(p2p_4)
+
+
+
+
+% figure
+% hold on;
+% histfit(periods4)
+% title('')
+% xlabel('')
+% ylabel('')

### 2019-03-14 tex_files/Journal.tex
journal

@@
 Along the way I located an error where I was using 5'd16 as the bias which doesnt work with a subtraction of 15 to -16, as it would give overflow.
-Athough I had actually  made the counter the same in both directions so this didnt actually manifest itself as an issue.
+Although I had actually  made the counter the same in both directions so this didnt actually manifest itself as an issue.
 The lock range is now heavily impacted by the gains as I discovered when attempting to copy gains from the PA ADPLL.
@@
 Either through being too aggressive and doesnt lock or from the ki value being too small, and when it is too small the output of the loop filter toggles between a value of min and maxval thus creating a huge amount of jitter.
+\section{pm 10 nano}
+I finally have got good performance but I am not sure if I solved the problem.
+Four main changes were made, which in isolation had little impact.
+Firstly in order to obtain a full output range I removed the extra bits being added in the loop filter after multiplied by ki or kp.
+This meant that a division by four of a number where the upper bit was empty became a div two with the whole number thus doubling the output range.
+The second change was to move the cycle delay from the input to output of the loop filter thus ensuring that the output would change at a more appropriate time.
+I also chose to clock this on a signal from earlier in the inverter chain thus ensuring it would always be finished changing before the clock edge.
+I also swapped the size of the control code and detector width to match the design of the PA, so I went from 5 bit cc and 6 bit error to 6 bit cc and 5 bit error.
+
+As a result of these changes I was able to get a scope image where there was plus minus 10 nanoseconds on the edge visible in screen.
+Down from 40 nanoseconds that I had had on Tuesday AM and down from the 40\% ish of the period with the divider in place.
+I did remove the divider for this testing to make things easier and as of Thursday 14/03 it had not been replaced.
+The plan is that next week I will reimplement the divider and once I have better aligned the 4 PLLs I have, go up to a 3x3 network.
+
 \end{document}

### 2019-03-19 conf_paper/paper.tex
conf paper motivation

\section*{Motivation}
The technical motivation underlying this paper is the creation of an Field Programmable Gate Arary (FPGA) based prototyping platform for All Digital Phase Lock Loop (ADPLL) networks. As the creation of Application Specific Integrated Circuits (ASICs) is an expensive and time consuming process, any mistake has the potential to be rather costly. As such prior to the manufacture of an IC it is important to ensure that any errors made in the design have been weeded out prior to this stage.
Simulations at either at theoretical or gate/transistor levels have global usage in minimising such errors due to the ubiquity of simulators and their ease of use.
However simulations are only as good as the model used to describe the dynamics of the system, and emulating real jitter and other behaviours of a complex system is a rather difficult task. %TODO wording of this last sentence
An FPGA based prototype allows the system designers to validate the performance of both design and simulation, particularly the response to key noise sources such as power supply noise. An FPGA is ideal for this task as it leverages the existing skillset of a digital designer, permits the re-use of certain blocks, and most importantly enables cost effective and rapid reworking of the design.\\
This paper will examine the process of creating such a system, highlight the differences between potential designs and address some of the challenges and pitfalls that may be encountered along the way. This paper will also demonstrate that FPGA based prototyping can play a central role on the pathway to the implementation of an ADPLL network on a custom chip, or indeed any number of similar applications. \

### 2019-03-19 conf_paper/paper.tex
paper page 1

@@
-\section*{Motivation}
-The technical motivation underlying this paper is the creation of an Field Programmable Gate Arary (FPGA) based prototyping platform for All Digital Phase Lock Loop (ADPLL) networks.
+\documentclass[conference]{IEEEtran}
+%\IEEEoverridecommandlockouts
+% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
+\usepackage{cite}
+\usepackage{amsmath,amssymb,amsfonts}
+\usepackage{algorithmic}
+\usepackage{graphicx}
+\usepackage{textcomp}
+\usepackage{xcolor}
+\usepackage[british]{babel}
+\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
+		T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
+\begin{document}
+
+\title{FPGA Based Prototyping of an ADPLL Network.\\}
+
+\author{\IEEEauthorblockN{C. Dooley, E. Blokhina, B. Mulkeen}
+	\IEEEauthorblockA{\textit{School of Electronic Engineering} \\
+		\textit{University College Dublin}\\
+		Belfield, Dublin 4, Ireland. \\
+		email address} %TODO
+}
+
+\maketitle
+
+\begin{abstract}
+	This document is a model and instructions for \LaTeX.
+	This and the IEEEtran.cls file define the components of your paper [title, text, heads, etc.]. *CRITICAL: Do Not Use Symbols, Special Characters, Footnotes, 
+	or Math in Paper Title or Abstract.
+\end{abstract}
+
+\begin{IEEEkeywords}
+	FPGA, ADPLL, Simulation, Prototype
+\end{IEEEkeywords}
+\section{Motivation}
+The technical motivation underlying this paper is the creation of an Field Programmable Gate Array (FPGA) based prototyping platform for All Digital Phase Lock Loop (ADPLL) networks.
 As the creation of Application Specific Integrated Circuits (ASICs) is an expensive and time consuming process, any mistake has the potential to be rather costly.
 As such prior to the manufacture of an IC it is important to ensure that any errors made in the design have been weeded out prior to this stage.
-Simulations at either at theoretical or gate/transistor levels have global usage in minimising such errors due to the ubiquity of simulators and their ease of use.
+Simulations at either at theoretical or gate/transistor levels have global usage in minimising such errors due to the ubiquity of simulators and their ease of use.\\
 However simulations are only as good as the model used to describe the dynamics of the system, and emulating real jitter and other behaviours of a complex system is a rather difficult task.
@@
 An FPGA is ideal for this task as it leverages the existing skillset of a digital designer, permits the re-use of certain blocks, and most importantly enables cost effective and rapid reworking of the design.\\
+An FPGA as a prototyping tool has seen use in the work of Dimitri Galayko and his research group in UPMC Sorbonne and in Elena Blokhina's team in University College Dublin, and has been used in the testing of established designs prior to their implementation in custom silicon\cite{zianbetov2013phd,shan2014phd}, the creation of realistic models for use in high level simulations\cite{theboys2019} and the exploration of new designs for digital blocks before progressing on to true digital design on a gate level.
+%TODO references
+In the case of Shan et.
+al \cite{shan2014phd} direct testing of control parameters was performed using an FPGA based design where all parameters were scaled down proportionally and thus the ideal gains for the loop filter found in simulation could be tested in a more realistic setting.\\
+Similarly Koskin et.
+al \cite{theboys2019} used the analysis of a number of ADPLLs implemented on an FPGA both independently and in a network to validate previously obtained theoretical results while avoiding the costs associated with the development of a custom IC.\\
+Prototyping on an FPGA does have its drawbacks, most notably the mixed-signal circuits central to the operation of an ADPLL are not implementable on an FPGA and while most fundamental digital circuit elements are present these elements are not true transistor level implementations of their functionality but rather implemented by lookup tables and other multi-role structures on the FPGA.
+These drawbacks present a challenge to a designer as they restrict the potential architectures of key blocks.
+This is of particular concern when trying to mimic the behaviour of an established design on an FPGA.\\
+The particular system that this paper will discuss is an ADPLL network.
+Based on an idea first proposed in a 1995 paper by Pratt and Nguyen \cite{pratt1995distributed}, in which they proposed an alternative clock distribution network for ICs using a Cartesian grid of clocking areas, each with their own Phase Lock Loop, which has become known as a PLL Network .
+This methodology was implemented by Gutnik et.
+al \cite{gutnik2000active} who found it to be feasible.
+Subsequently Javiden et.
+al
+\cite{javidan2011all} proposed the implementation of such a system using All Digital PLLs which better suited the use cases of the technology and avoided some of the flaws pointed out in the original paper.\\
 This paper will examine the process of creating such a system, highlight the differences between potential designs and address some of the challenges and pitfalls that may be encountered along the way.
 This paper will also demonstrate that FPGA based prototyping can play a central role on the pathway to the implementation of an ADPLL network on a custom chip, or indeed any number of similar applications.
-\
+The paper is organised as follows: Section II describes the system that will be used on the FPGA and Section III the implementation thereof and the challenges encountered in the process.
+Finally Section IV will present example measurements made on the platform and discuss their merit.
+
+\section{System Architecture}
+
+\newpage
+\bibliography{conf.bib} 
+\bibliographystyle{IEEEtran}
+
+\end{document}

### 2019-03-19 conf_paper/paper.tex
page 1 done

@@
 \usepackage{graphicx}
+\usepackage{float}
 \usepackage{textcomp}
@@
 \section{System Architecture}
+An ADPLL network as the name would suggest is created from a number of ADPLLs that coupled using digital phase comparators which attempt to measure the phase error between two oscillators, with each non edge node being connected to four neighbours.
+Figure \ref{fig:adpll_network} illustrates such a system on a small scale.
+The individual ADPLLs are made up of a digitally controllable oscillator, an error combination block and a digital loop filter with the aforementioned digital phase detectors lying between each pair of oscillators.
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=0.25\textwidth]{example}
+	\caption{ADPLL Architecture.}
+	\label{fig:adpll_network}
+	\includegraphics[width=0.25\textwidth]{example}
+	\caption{ADPLL Architecture.}
+	\label{fig:adpll_base}
+\end{figure}
+ Figure \ref{fig:adpll_base} demonstrates the ADPLL design used in this paper.
 

### 2019-03-19 conf_paper/paper.tex
adding images

@@
 Figure \ref{fig:adpll_network} illustrates such a system on a small scale.
-The individual ADPLLs are made up of a digitally controllable oscillator, an error combination block and a digital loop filter with the aforementioned digital phase detectors lying between each pair of oscillators.
+The individual ADPLLs, often called distributed ADPLLs as the phase detectors are shared between nodes, are made up of a digitally controllable oscillator, an error combination block and a digital loop filter with the aforementioned digital phase detectors lying between each pair of oscillators.
 \begin{figure}[h]
 	\centering
-	\includegraphics[width=0.25\textwidth]{example}
-	\caption{ADPLL Architecture.}
+	\includegraphics[width=0.25\textwidth]{adpll_network}
+	\caption{ADPLL Network Architecture.}
 	\label{fig:adpll_network}
-	\includegraphics[width=0.25\textwidth]{example}
-	\caption{ADPLL Architecture.}
+	\vspace{0.5cm}
+	\includegraphics[width=0.4\textwidth]{dist_adpll}
+	\caption{Distributed ADPLL Design.}
 	\label{fig:adpll_base}

### 2019-03-20 conf_paper/paper.tex
second page of the conference paper has been written, onto the third ;)

@@
 An FPGA based prototype allows the system designers to validate the performance of both design and simulation, particularly the response to key noise sources such as power supply noise.
-An FPGA is ideal for this task as it leverages the existing skillset of a digital designer, permits the re-use of certain blocks, and most importantly enables cost effective and rapid reworking of the design.\\
+An FPGA is ideal for this task as it leverages the existing skill set of a digital designer, permits the re-use of certain blocks, and most importantly enables cost effective and rapid reworking of the design.\\
 An FPGA as a prototyping tool has seen use in the work of Dimitri Galayko and his research group in UPMC Sorbonne and in Elena Blokhina's team in University College Dublin, and has been used in the testing of established designs prior to their implementation in custom silicon\cite{zianbetov2013phd,shan2014phd}, the creation of realistic models for use in high level simulations\cite{theboys2019} and the exploration of new designs for digital blocks before progressing on to true digital design on a gate level.
@@
 An ADPLL network as the name would suggest is created from a number of ADPLLs that coupled using digital phase comparators which attempt to measure the phase error between two oscillators, with each non edge node being connected to four neighbours.
-Figure \ref{fig:adpll_network} illustrates such a system on a small scale.
+Figure \ref{fig:adpll_network} illustrates such a system on a small scale, with a number of ADPLLs laid out in a Cartesian grid.
 The individual ADPLLs, often called distributed ADPLLs as the phase detectors are shared between nodes, are made up of a digitally controllable oscillator, an error combination block and a digital loop filter with the aforementioned digital phase detectors lying between each pair of oscillators.
@@
 \end{figure}
- Figure \ref{fig:adpll_base} demonstrates the ADPLL design used in this paper.
+\subsection{Digitally Controlled Oscillators}
+While regular Voltage Controlled Oscillator (VCO) is tuned by the input voltage over a continuous range of values in a digital system there are a very limited number of voltages representable, most commonly just two levels.
+As such using the voltage itself to control an oscillator is not a viable strategy and instead Digitally, or as they are sometimes known as Numerically, Controlled Oscillators (DCO) accept an \textit{n} bit wide control signal that alters the period of the output waveform.
+While tuning range, centre frequency and linearity apply as in a VCO, a DCO also has a minimum frequency step as a result of the non continuous control values.
+%TODO WORDING
+In combination with the bit width of the control signal the minimum frequency step limits the range over which the oscillator can be tuned.
+Frequently the output of the DCO is divided down to allow the control blocks in the feedback path that are clocked on this signal to run at a lower and thus easier frequency.
+\subsection{Loop Filter}
+The loop filter is a key aspect of a phase lock loop as without it only the current output of the phase detector could be used to compute the control voltage of the VCO.
+In an ADPLL it performs an identical role, however the digital environment results the traditional loop filter being replaced by a discrete time Proportional Integral (PI) controller, with the integral path of the controller implemented by an accumulator.
+\subsection{Error Combiner}
+The Error Combiner, as the name suggests, performs a summation of the error signals from the phase detectors lying between any neighbouring oscillator and its neighbours.
+It is often convenient for this to be performed as a weighted sum, thus avoiding different gain requirements for the loop filter of oscillators in the corners of the grid that only have a pair of neighbours.
+The ability to change the weights during operation enables ``uni-directional start-up'' which Javiden et.
+al
+\cite{javidan2011all} found prevented the system from entering unwanted stable equilibriums warned of in Pratt' and Nguyen's seminal paper.
+\subsection{Phase Detector}
+As in a conventional PLL the phase detector outputs a value proportional to the difference in phase between a pair of signals, which is then output as a continuous voltage.
+A digital phase detector then attempts to do the same by measuring the difference between fixed reference points in the waveform, which as a digital system deals with square waves this is typically the rising edge of each input to the block.
+Similarly to the oscillator quantisation also plays a role in the behaviour of this block, with the time difference between each rising edge having a minimum possible measurement step.
+In distributed ADPLLs the phase detector has dual outputs, each the inverse of the other, in order to feed the correct value to both of the connected error combiners.
+
+\section{FPGA Prototyping}
+\subsection{FPGA Restrictions}
+The use of an FPGA comes with many restrictions as to how a complex system such as an ADPLL can be designed with the mixed signal blocks, the DCO and phase detectors, in particular being curtailed by the lack of transistor level control of the design and layout.\\
+Two potential solutions exist which alleviate this problem, and are suited to different frequency ranges and/or goals of the designer.
+Firstly the clock signal generated by the FPGA's clock distribution block can be used to drive accumulator based oscillators and phase detectors a technique suitable in two main use cases: The emulation of the performance of an already designed system operating at a proportionally cut down frequency.
+At frequencies that are orders of magnitude lower than the maximum obtainable clock frequency of the FPGA's clock manager.
+The downside of this approach is that the minimum steps are dictated by the clock frequency of the FPGA, which at frequencies that are not orders of magnitude lower than the FPGA clock will result in coarse resolutions for detector or oscillator, and thus greater jitter seen at the oscillator output.\\
+The second solution to the problem is more akin to the real world system, and uses the delay through primitive elements such as inverters, which can be deduced from measurements on the FPGA of choice but will vary depending on the layout selected.
+In the case of the FPGA used for this paper, a Xilinx Artix-7 (XC7A100T-1CSG324C) this delay is in the region of 300 picoseconds but varies depending on the routing between elements.
+This design is better suited to scenarios where the desired operating frequency is a significant fraction of the maximum usable clock of the FPGA or it is particularly desirable to have asynchronous behaviour in the system as in an ASIC.
+It is not suited to lower frequency designs as the number of inverters required will run into spacial constraints.
+\subsection*{Digitally Controlled Oscillators on an FPGA}
 

### 2019-03-20 conf_paper/paper.tex
page 3 started

@@
 It is not suited to lower frequency designs as the number of inverters required will run into spacial constraints.
-\subsection*{Digitally Controlled Oscillators on an FPGA}
+\subsection{Digitally Controlled Oscillators on an FPGA}
+There are three key designs that be used to implement a DCO on an FPGA, two of which rely on the FPGA clock and one using inverter primitives. 
+\subsubsection*{Design 1}
+The first of these is a linear in period design using an accumulator architecture, with a variable reload value on overflow.
+Here the control code modifies the reload value around a bias point and thus adjusts the time taken for overflow to be reached with the oscillator's period given as $T_{osc} = (2^n - (\textrm{bias}+\textrm{code}))\times\textrm{clock period}$.
+Period linearity is a desirable characteristic as it mimics the behaviour of many mixed-signal implementations, and is the key use case of this method.
+The choice of output signal however highlights the drawbacks of this method.
+With the accumulator's MSB acting as the output less than half of the control codes are feasible.
+As the minimum period is set by the number of unused bits and the maximum period fixed at twice the minimum the frequency and tuning range combinations are very limited.
+In addition the wave output by this oscillator is not infact square, but skewed by the bias point.
+\subsubsection*{Design 2}
+A linear in frequency design using an accumulator is also viable
 

### 2019-03-21 conf_paper/paper.tex
running out of space lol

@@
 There are three key designs that be used to implement a DCO on an FPGA, two of which rely on the FPGA clock and one using inverter primitives.
+Each of these designs fulfil a different role in mimicking the behaviour of an ASIC based mixed signal DCO.
+
 \subsubsection*{Design 1}
 The first of these is a linear in period design using an accumulator architecture, with a variable reload value on overflow.
-Here the control code modifies the reload value around a bias point and thus adjusts the time taken for overflow to be reached with the oscillator's period given as $T_{osc} = (2^n - (\textrm{bias}+\textrm{code}))\times\textrm{clock period}$.
+Here the control code modifies the reload value around a bias point and thus adjusts the time taken for overflow to be reached with the oscillator's period given as $T_{osc} = (2^{\textrm{width}} - (\textrm{bias}+\textrm{code}))\times\textrm{clock period}$.
 Period linearity is a desirable characteristic as it mimics the behaviour of many mixed-signal implementations, and is the key use case of this method.
-The choice of output signal however highlights the drawbacks of this method.
-With the accumulator's MSB acting as the output less than half of the control codes are feasible.
+The choice of the accumulator's Most Significant Bit (MSB) as the output signal however highlights the drawbacks of this method as less than half of the control codes are feasible.
 As the minimum period is set by the number of unused bits and the maximum period fixed at twice the minimum the frequency and tuning range combinations are very limited.
-In addition the wave output by this oscillator is not infact square, but skewed by the bias point.
+In addition the wave output by this oscillator is only square at the maximum period and control code changes vary the duty cycle of the signal.
+
 \subsubsection*{Design 2}
-A linear in frequency design using an accumulator is also viable
+If frequency linearity is a desired characteristic of the oscillator an accumulator based, FPGA clock driven design exists for this use case also. 
+Here the control code $k$ is added to the value stored in the accumulator at each rising edge of the FPGA clock until overflow is reached at $2^n-1$ where $n$ is the bit width of the counter thus valuing each control code increment at $\frac{\textrm{FPGA clock}}{2^{\textrm{width}}}$ Hz.
+Once again the MSB of the accumulator is used as the output signal, however with this design the MSB will spend an equivalent duration at both logic high and low.
+Again this design is better suited to lower frequencies, where the output of the DCO is orders of magnitude lower than the clock signal driving it.
+
+\subsubsection*{Design 3}
+The third and final potential DCO design is an inverter ring/chain which makes use of the delay through inverter primitives, thus bearing significant resemblance to DCOs used on ASICs.
+On an FPGA the lack of precise control over the layout of blocks means the exact value of the delay is uncertain and may vary noticeably from one implementation to another, however this misalignment and the asynchronous behaviour stemming from the lack of an FPGA driven clock increase the similarity between this design and an ASIC.
+The period is set by the number of inverters at $2\times(\textrm{num inverters})\times(\textrm{delay})$, and is varied by adding or removing a pair of inverters for every control code increment.
+This design works best when the desired period is within a couple orders of magnitude of the delay through each pair of inverters, beyond this point the number of look-up tables and routing resources required to implement the design becomes unsustainable.
+
+\subsection{Phase Detectors on an FPGA}
+The two options for replicating the behaviour of a phase detector also employ the same two design principles, with the option better suited to lower frequency signals once again being driven by the clock generated by the FPGA and an inverter based design better suiting higher frequency systems where the greater resolution offered is beneficial.\\
+A  Sign-Number (SigNum) detector is a common method of performing phase-frequency detection in an ADPLL and is the method that will be discussed in this paper.
+As the name suggests this design constitutes two sections, with one determining the sign of the phase difference and the other its magnitude.
+The first issue with implementing a phase detector on an FPGA is the issue of metastability, where the output of a circuit element is at an undetermined level, which can occur if both signs were to have edges within a short time of another or the set-up/hold times of a register are violated.
+This issue is easily addressed in an ASIC with an arbiter, however an arbiter is a mixed signal circuit that is not synthesisable on an FPGA.
+
+\subsubsection*{Design A}
+The first potential design uses a Finite State Machine (FSM) to determine which edge which edge has occurred first and thus the sign, and a counter that computes the time difference between each rising edge.
+This counter is clocked by the FPGA clock, thus giving the system a time resolution of $\frac{1}{F_{\textrm{FPGA}}}$.
+As before this results in a design that is only suitable when the frequency of operation is well below that of the FPGA clock due to resolution constraints.
+In order to avoid the problem of metastability the incoming waveforms can be synchronised to the clock of the FPGA for no penalty beyond the extra clock cycle taken as this synchronisation performs the same quantisation that would otherwise have occurred in the counter measuring the time difference.
+This design is better suited for use with the accumulator based oscillators due to the overlapping frequency ranges, however it can also be used in conjunction with an inverter ring.
+Important to note is that in the case of the Artix-7 used in this paper this results in a four times worse phase difference resolution than period resolution which made low jitter operation difficult, but this combination could be made possible by removing eight rather than two inverters for each control code step.
+%TODO plot this later
+
+\subsubsection*{Design B}
+Better suited for use with the inverter based oscillator is a system that mimics the behaviour of a tapped delay line working asynchronously to the FPGA clock.
+As the FPGA clock is not used, and that of the ADPLL cannot be used to measure itself an unclocked sign detection circuit is required, an example of which can be seen in Figure \ref{fig:segn_pdetdl}.
+This is nearly identical to sign detection methods used on an ASIC with the arbitration circuit replaced by a second SR Latch which alleviates the issue of metastability.\\
+A tapped delay line can be emulated on an FPGA through the use of inverter pairs due to their short and approximately known delay placed between each tap.
+The resolution then is the propagation time through the pair, although this will once again depend on the layout specifics, which was approximately 600 picoseconds on an Artix-7.
+This is significantly lower than the 4 nanoseconds achieved using the previous design and thus more suitable for use with an inverter ring oscillator as this resolution is better than the time resolution of the oscillator.
+Figure \ref{fig:num_pdetdl} illustrates the configuration of the circuit used to emulate the function of a delay line that will be used in this paper.
+\begin{figure}[h]
+    \centering
+    %\includegraphics[width=0.25\textwidth]{../new_pdet1}
+    \includegraphics[width=0.25\textwidth]{../bad_waveform}
+    \caption{Design B Sign Detection RTL Diagram.}
+    \label{fig:segn_pdetdl}
+\end{figure}
+\begin{figure}[h]
+    \centering
+    %\includegraphics[width=0.4\textwidth]{../new_pdet2}
+    \includegraphics[width=0.25\textwidth]{../bad_waveform}
+    \caption{Tapped Delay Line Emulation RTL Diagram.}
+    \label{fig:num_pdetdl}
+\end{figure}
 

### 2019-03-22 matlab/cd___anal.m
doctors

@@ -1,4 +1,4 @@
-filename = 'c_____0.bin';
+filename = 'scope_41.bin';
 [scope_x_data,scope_y_data] = importAgilentBin(filename,3);
 
 f_samp = 4E9;

### 2019-03-22 matlab/plot_4_waves.m

@@ -1,4 +1,4 @@
-filename = 'scope_36.bin';
+filename = 'scope_37.bin';
 [~,scope_y_data] = importAgilentBin(filename,1);
 [~,scope_y_data2] = importAgilentBin(filename,2);
 [~,scope_y_data3] = importAgilentBin(filename,3);

### 2019-03-22 conf_paper/paper.tex
images
doctors
draft 1 done

@@
 \usepackage[british]{babel}
+\newcommand\T{\rule{0pt}{2.6ex}}       % Top strut
+\newcommand\B{\rule[-1.2ex]{0pt}{0pt}} % Bottom strut
 \def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
@@
     \centering
-    %\includegraphics[width=0.25\textwidth]{../new_pdet1}
-    \includegraphics[width=0.25\textwidth]{../bad_waveform}
+    \includegraphics[width=0.4\textwidth]{../new_pdet1}
     \caption{Design B Sign Detection RTL Diagram.}
@@
     \centering
-    %\includegraphics[width=0.4\textwidth]{../new_pdet2}
-    \includegraphics[width=0.25\textwidth]{../bad_waveform}
+    \includegraphics[width=0.4\textwidth]{../new_pdet2}
     \caption{Tapped Delay Line Emulation RTL Diagram.}
@@
 \end{figure}
+\section{Comparison of Designs}
+This paper will examine the performance of a number of combinations of these blocks and the performance thereof, specifically oscillator 2 with phase detector A and oscillator 3 with detectors A \& B.
+For each test there were four ADPLLs on the FPGA arranged in a Cartesian Grid and as such each ADPLL will be referred to using matrix indexing corresponding to their grid location.
+For each configuration three different modes of operation were examined: Each ADPLL locking to the same 5 MHz external reference, each ADPLL connected to the left and above neighbours in a ``uni-directional'' network and finally a true ADPLL network where all neighbours are connected.
+For comparing each configuration gains which had previously been determined to provide the best jitter performance were chosen.
+The data was captured using an Agilent MSO7054 A at 4 GSa/s.
+Each point of data is the mean of five 200 microsecond long captures which were then analysed in Matlab, and Table \ref{table:b_c_perf} displays the results of this analysis.\\
+The first columns of the table relate to the cycle-to-cycle jitter of the system, the difference between sequential distances, and the second portion gives the Time Interval Error, or difference between the actual timing of the edge and the ``ideal'' time as set by the reference.
+This combination of measurements will show how well each design performs the task of staying consistent from cycle to cycle both in relation to itself and the system at large.
 
-\newpage
+\begin{table}[!ht]
+	\begin{center}
+		\begin{small}
+			\begin{tabular}{cc|c|c|c|c|c|c|}           
+				\cline{3-8}
+				&& \multicolumn{3}{c|}{Jitter Standard Deviation (ns)} & \multicolumn{3}{c|}{Time Interval Error. (ns)} \T\\
+				\cline{3-8}	
+				&&PLL 11&PLL 12&PLL 22&PLL 11&PLL 12&PLL 22\T\\
+				\hline
+				\multicolumn{2}{|l|}{Osc 2, PDET A}&-&-&-&-&-&-\T\\
+				\multicolumn{2}{|r|}{Free PLLs}&1.5988&1.6178&1.6220&&& \T\\
+				\multicolumn{2}{|r|}{Uni-dir.}&1.9804&2.0316&2.1317&&& \T\\
+				\multicolumn{2}{|r|}{Bi-dir.}&2.0124&2.0681&2.0732&&& \T\\
+				\hline
+				\multicolumn{2}{|l|}{Osc 3, PDET A}&-&-&-&-&-&-\T\\
+				\multicolumn{2}{|r|}{Free PLLs}&0.62499 $\mu$s&0.75919 $\mu$s&0.79049  $\mu$s&&& \T\\
+				\multicolumn{2}{|r|}{Uni-dir.}&0.76822 $\mu$s&0.9009 $\mu$s&1.1407 $\mu$s&&& \T\\
+				\multicolumn{2}{|r|}{Bi-dir.}&0.70393 $\mu$s&0.70201 $\mu$s&1.241 $\mu$s&&& \T\\
+				\hline
+				\multicolumn{2}{|l|}{Osc 3, PDET B}&-&-&-&-&-&-\T\\
+				\multicolumn{2}{|r|}{Free PLLs}&4.1680&3.9803&5.0741&&& \T\\
+				\multicolumn{2}{|r|}{Uni-dir.}&4.3820&5.7315&6.2336 &&& \T\\
+				\multicolumn{2}{|r|}{Bi-dir.}&5.6137&7.1831&7.1612 &&& \T\\
+				\hline
+				\B
+			\end{tabular}
+		\end{small}
+		\caption{}
+		\label{table:b_c_perf}
+	\end{center}
+\end{table}
+
+From the table it is apparent that the combination of the inverter chain DCO and the FPGA clock driven phase detector is an ill advised choice, as the jitter present in the system is several orders of magnitude worse than either other situation.
+In this system the FPGA clock is at 258 MHz, the maximum possible, thus giving a detection resolution of 4 nanoseconds.
+This is 4 times greater than the period resolution of approximately 1.2 nano seconds of the oscillator, and requires the gains of the PI filter to be increased fourfold in order to respond appropriately to the error detected.
+The problems of relatively inaccurate detection and increased gains combine to produce a system that is only able to achieve poor locking performance.
+Both other combinations work well however, delivering a level of jitter that is both low yet allows an indication how impactful variations in architecture or control parameters to be seen visually on an oscilloscope.\\
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=0.25\textwidth]{distrib_ring}
+	\caption{Design B Sign Detection RTL Diagram.}
+	\label{fig:dist1}
+\end{figure}
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=0.25\textwidth]{distrib_pa}
+	\caption{Tapped Delay Line Emulation RTL Diagram.}
+	\label{fig:dist2}
+\end{figure}
+Figures \ref{fig:dist1} \& \ref{fig:dist2} paint a picture of the unrealistic nature of the FPGA clocked designs as in comparison to the ring based design each oscillator will have identical intrinsic period steps thus eliminating any jitter that maybe be seen in the system due to misalignment.
+This effects the suitability of the FPGA clocked designs for the improvement of simulation models or the examination of how a novel block maybe impact the behaviour of the network.
+\section*{Conclusion}
+An FPGA is a suitable environment for the prototyping of ADPLL networks and can fill a number of roles depending on the designer's requirements.
+However the designer should bear in mind the platform's limitations especially when attempting to use a frequency that is not orders of magnitude lower than the maximum clock available from the FPGA's clock manager.
+Inverter based and FPGA clocked designs both have a niche in which they are the preferred option, with the clocked design better suiting replication of an existing design at scaled down frequencies and the inverter based designs better in the modelling and analysis of ADPLL network behaviour.
 \bibliography{conf.bib} 

### 2019-03-22 matlab/a10M_quad_plot.m

@@ -1,4 +1,4 @@
-filename = 'PA_10M_div4.bin';
+filename = 'c_____1.bin';
 [~,scope_y_data] = importAgilentBin(filename,1);
 [scope_x_data,scope_y_data2] = importAgilentBin(filename,2);
 [w,init_cross,final_cross,mid_level] = pulsewidth(scope_y_data,scope_x_data, 'Polarity', 'Positive');

### 2019-03-22 matlab/a1_3.m

@@ -1,4 +1,4 @@
-filename = 'cdcd.bin';
+filename = 'scope_41.bin';
 [~,scope_y_data] = importAgilentBin(filename,1);
 [~,scope_y_data2] = importAgilentBin(filename,2);
 [~,scope_y_data3] = importAgilentBin(filename,3);
@@ -9,12 +9,12 @@ num_samp = length(scope_y_data);
 t_max = num_samp*(1/f_samp); 
 t=linspace(0,t_max,num_samp);
 
-figure(1)
-hold on; 
-plot(t*1e6-7.95-0.09,scope_y_data+15,'b');
-plot(t*1e6-7.95-0.09,scope_y_data2+10,'r');
-plot(t*1e6-7.95-0.09,scope_y_data3+5,'g');
-plot(t*1e6-7.95-0.09,scope_y_data4+0,'k');
+% figure(1)
+% hold on; 
+% plot(t*1e6-7.95-0.09,scope_y_data+15,'b');
+% plot(t*1e6-7.95-0.09,scope_y_data2+10,'r');
+% plot(t*1e6-7.95-0.09,scope_y_data3+5,'g');
+% plot(t*1e6-7.95-0.09,scope_y_data4+0,'k');
 
 [~,init_cross2,~,~] = pulsewidth(scope_y_data2,scope_x_data, 'Polarity', 'Positive');
 [~,init_cross3,~,~] = pulsewidth(scope_y_data3,scope_x_data, 'Polarity', 'Positive');
@@ -23,6 +23,10 @@ plot(t*1e6-7.95-0.09,scope_y_data4+0,'k');
 periods2 = getPeriods(init_cross2);
 periods3 = getPeriods(init_cross3);
 periods4 = getPeriods(init_cross4);
+
+mean_period_2 = mean(periods2);
+mean_period_3 = mean(periods3);
+mean_period_4 = mean(periods4);
 
 std_period_2 = std(periods2);
 std_period_3 = std(periods3);
@@ -38,23 +42,40 @@ max_4 = max(periods4-mean(periods4));
 min_4 = min(periods4-mean(periods4));
 p2p_4 = max_4-min_4;
 
-diff2_3 = init_cross2-init_cross3;
-diff3_4 = init_cross3-init_cross4;
-diff4_2 = init_cross4-init_cross2;
+% diff2_3 = init_cross2-init_cross3;
+% diff3_4 = init_cross3-init_cross4;
[…]

### 2019-03-22 matlab/cd___anal.m

@@ -1,4 +1,4 @@
-filename = 'scope_41.bin';
+filename = 'c_____0.bin';
 [scope_x_data,scope_y_data] = importAgilentBin(filename,3);
 
 f_samp = 4E9;

### 2019-03-22 matlab/cheating.m

@@ -1,0 +1,34 @@
+% two_a_plls = [1.6051e-09, 1.5924e-09, 1.6146e-09, 1.6251e-09, 1.619e-09, 1.621e-09,];
+% two_a_uni = [1.9819e-09, 2.0627e-09, 2.2021e-09, 1.9789e-09, 2.0005e-09, 2.0612e-09,];
+% two_a_bi = [2.0303e-09, 2.1097e-09, 2.1253e-09, 1.9946e-09, 2.0264e-09, 2.0212e-09,];
+% 
+% three_a_plls = [6.2499e-07, 7.5919e-07, 7.9049e-07,];
+% three_a_uni = [7.6822e-07, 9.009e-07,  1.1407e-06,];
+% three_a_bi = [7.0393e-07, 7.0201e-07, 1.0241e-06,];
+% 
+% three_b_plls = [4.1561e-09, 4.0044e-10, 4.1883e-09, 3.9882e-10, 4.1628e-09, 3.9888e-10, 4.1648e-09, 3.9396e-10,];
+% three_b_uni = [6.1918e-09, 5.2058e-09, 1.9428e-08, 6.2755e-09, 5.2531e-09, 1.9096e-08, 2.2232e-08,];
+% three_b_bi = [5.5243e-09, 7.1567e-09, 6.9324e-09, 5.4574e-09, 7.2364e-09, 7.4538e-09, 5.3828e-09, 7.1966e-09, 7.4565e-09, 5.8469e-09, 7.1654e-09, 6.9995e-09, 5.8569e-09, 7.1604e-09, 6.9636e-09, ];
+% 
+% mean(two_a_plls)
+% mean(two_a_uni)
+% mean(two_a_bi)
+% mean(three_a_plls)
+% mean(three_a_uni)
+% mean(three_a_bi)
+% mean(three_b_plls)
+% mean(three_b_uni)
+% mean(three_b_bi)
+
+mean([6.9324,
+7.4538,
+7.4565,
+6.9995,
+6.9636,
+
+
+
+
+  
+])
+

### 2019-03-23 conf_paper/paper.tex
new LF test module without fractions & 4 page + ref draft of paper

@@
 \begin{abstract}
-    This document is a model and instructions for \LaTeX.
-    This and the IEEEtran.cls file define the components of your paper [title, text, heads, etc.]. *CRITICAL: Do Not Use Symbols, Special Characters, Footnotes, 
-    or Math in Paper Title or Abstract.
+    This paper introduces and compares the implementation of a number of FPGA based ADPLL network prototyping architectures.
+Comparison is then made between the expected performance and role of each method as a development tool.
 \end{abstract}
@@
 \begin{IEEEkeywords}
-    FPGA, ADPLL, Simulation, Prototype
+    FPGA, ADPLL, Simulation, Prototyping
 \end{IEEEkeywords}
@@
 An FPGA is ideal for this task as it leverages the existing skill set of a digital designer, permits the re-use of certain blocks, and most importantly enables cost effective and rapid reworking of the design.\\
-An FPGA as a prototyping tool has seen use in the work of Dimitri Galayko and his research group in UPMC Sorbonne and in Elena Blokhina's team in University College Dublin, and has been used in the testing of established designs prior to their implementation in custom silicon\cite{zianbetov2013phd,shan2014phd}, the creation of realistic models for use in high level simulations\cite{theboys2019} and the exploration of new designs for digital blocks before progressing on to true digital design on a gate level.
+An FPGA as a prototyping tool has seen use in UPMC Sorbonne and University College Dublin, and has been used in the testing of established designs prior to their implementation in custom silicon\cite{zianbetov2013phd,shan2014phd}, the validation of realistic models for use in high level simulations\cite{theboys2019} and the exploration of new designs for digital blocks before progressing on to true digital design on a gate level.
 %TODO references
@@
 
-\subsubsection*{Design 1}
-The first of these is a linear in period design using an accumulator architecture, with a variable reload value on overflow.
+%\subsubsection*{Design 1}
+%The first of these is a linear in period design using an accumulator architecture, with a variable reload value on overflow.
 Here the control code modifies the reload value around a bias point and thus adjusts the time taken for overflow to be reached with the oscillator's period given as $T_{osc} = (2^{\textrm{width}} - (\textrm{bias}+\textrm{code}))\times\textrm{clock period}$.
@@
 As the minimum period is set by the number of unused bits and the maximum period fixed at twice the minimum the frequency and tuning range combinations are very limited.
-In addition the wave output by this oscillator is only square at the maximum period and control code changes vary the duty cycle of the signal.
+%In addition the wave output by this oscillator is only square at the maximum period and control code changes vary the duty cycle of the signal.
 
-\subsubsection*{Design 2}
+\subsubsection*{Design 1}
 If frequency linearity is a desired characteristic of the oscillator an accumulator based, FPGA clock driven design exists for this use case also. 
@@
 
-\subsubsection*{Design 3}
+\subsubsection*{Design 2}
 The third and final potential DCO design is an inverter ring/chain which makes use of the delay through inverter primitives, thus bearing significant resemblance to DCOs used on ASICs.
@@
 The period is set by the number of inverters at $2\times(\textrm{num inverters})\times(\textrm{delay})$, and is varied by adding or removing a pair of inverters for every control code increment.
-This design works best when the desired period is within a couple orders of magnitude of the delay through each pair of inverters, beyond this point the number of look-up tables and routing resources required to implement the design becomes unsustainable.
+This design works best when the desired period is within a couple orders of magnitude of the delay through each pair of inverters, beyond this point the number of look-up tables and routing resources required to implement the design becomes unsustainable.\\
+A third potential design exists, similar to Oscillator 1 in which the control code varies the reload value of the accumulator at overflow, however this design will not be discussed in this paper.
 
@@
 \section{Comparison of Designs}
-This paper will examine the performance of a number of combinations of these blocks and the performance thereof, specifically oscillator 2 with phase detector A and oscillator 3 with detectors A \& B.
+This paper will examine the performance of a number of combinations of these blocks and the performance thereof, specifically oscillator 1 with phase detector A and oscillator 2 with detectors A \& B.
 For each test there were four ADPLLs on the FPGA arranged in a Cartesian Grid and as such each ADPLL will be referred to using matrix indexing corresponding to their grid location.
@@
 	\begin{center}
-		\begin{small}
+		\begin{tiny}
 			\begin{tabular}{cc|c|c|c|c|c|c|}           
@@
 				\hline
-				\multicolumn{2}{|l|}{Osc 2, PDET A}&-&-&-&-&-&-\T\\
+				\multicolumn{2}{|l|}{Osc 1, PDET A}&-&-&-&-&-&-\T\\
 				\multicolumn{2}{|r|}{Free PLLs}&1.5988&1.6178&1.6220&&& \T\\
@@
 				\hline
-				\multicolumn{2}{|l|}{Osc 3, PDET A}&-&-&-&-&-&-\T\\
+				\multicolumn{2}{|l|}{Osc 2, PDET A}&-&-&-&-&-&-\T\\
 				\multicolumn{2}{|r|}{Free PLLs}&0.62499 $\mu$s&0.75919 $\mu$s&0.79049  $\mu$s&&& \T\\
@@
 				\hline
-				\multicolumn{2}{|l|}{Osc 3, PDET B}&-&-&-&-&-&-\T\\
+				\multicolumn{2}{|l|}{Osc 2, PDET B}&-&-&-&-&-&-\T\\
 				\multicolumn{2}{|r|}{Free PLLs}&4.1680&3.9803&5.0741&&& \T\\
@@
 			\end{tabular}
-		\end{small}
+		\end{tiny}
 		\caption{}
@@
 	\includegraphics[width=0.25\textwidth]{distrib_ring}
-	\caption{Design B Sign Detection RTL Diagram.}
+	\caption{Distribution of Periods for Inverter Based Design.}
 	\label{fig:dist1}
@@
 	\includegraphics[width=0.25\textwidth]{distrib_pa}
-	\caption{Tapped Delay Line Emulation RTL Diagram.}
+	\caption{Distribution of Periods for FPGA Clocked Design.}
 	\label{fig:dist2}
@@
 Inverter based and FPGA clocked designs both have a niche in which they are the preferred option, with the clocked design better suiting replication of an existing design at scaled down frequencies and the inverter based designs better in the modelling and analysis of ADPLL network behaviour.
-\bibliography{conf.bib} 
-\bibliographystyle{IEEEtran}
+
+\bibliographystyle{ieeetran}
+\bibliography{conf}
 

### 2019-03-23 tex_files/Journal.tex
journal

@@
 After discovering this I had it locking within 3 hours of work, when prior I had spent ~ 9 days on the calendar and ~30 hours working on the problem.
-Locking perofmance looks potentially better than the 10 MHz perf of the ADPLL (magnitude not \% for the sake of comp.
+Locking performance looks potentially better than the 10 MHz perf of the ADPLL (magnitude not \% for the sake of comp.
 However div is by 4 now not 8, and res is twice as good.) In the coming week I will be characterising the perf of these different designs to compare effectively.
@@
 Turns out it was being caused by the phase detector design which I was using being stable for very small gains and a small number of bit phase detection and if the falling edge of signal 1 toggled back and forth between rising on signal two I would get oscillation between positive and negative, or lead and lag to use the correct terms, and with the tiny gains making it actually stable in this state.
-Changing the detector to a more conventional style that did not terminate a measurement until the next rising edge on the opposite signal fixed this probem.
-The new issue now is that I cannot get low jitter locking, seeingly down to the behaviour of the loop filter.\\
+Changing the detector to a more conventional style that did not terminate a measurement until the next rising edge on the opposite signal fixed this problem.
+The new issue now is that I cannot get low jitter locking, seemingly down to the behaviour of the loop filter.\\
 The loop filter appears to work correctly when the two paths are taken in isolation but when combining the paths it breaks down.
-Either through being too aggressive and doesnt lock or from the ki value being too small, and when it is too small the output of the loop filter toggles between a value of min and maxval thus creating a huge amount of jitter.
-\section{pm 10 nano}
+Either through being too aggressive and doesn't lock or from the ki value being too small, and when it is too small the output of the loop filter toggles between a value of min and maxval thus creating a huge amount of jitter.
+\section*{pm 10 nano}
 I finally have got good performance but I am not sure if I solved the problem.
@@
 The plan is that next week I will reimplement the divider and once I have better aligned the 4 PLLs I have, go up to a 3x3 network.
+\section*{Alignment}
+Found another lingering PF issue this week that was narrowing the locking range.
+Other than that I spent the week aligning oscillators in order to have a network.
+2x2 was fine, 3x3 was a struggle.
+The LF issue that I suspect exists maybe the source of the narrow locking range I have seen.
+It remains to be seen however as this week was conference paper writing week and as such things have been slow in other areas (also 2 assignments).
+Made a new loop filter this weekend without fractions which may be easier to debug (Pierre's suggestion).
 

### 2019-03-25 conf_paper/paper.tex
paper spellcheck as of Anne, threebythreeringtest, twobytwo back to being 2x2 and new constraint files added as appropriate

@@
 \usepackage{float}
+\usepackage{subfig}
 \usepackage{textcomp}
@@
         Belfield, Dublin 4, Ireland. \\
-        email address} %TODO
+        conor.dooley@ucdconnect.ie} %TODO
 }
@@
 As the creation of Application Specific Integrated Circuits (ASICs) is an expensive and time consuming process, any mistake has the potential to be rather costly.
-As such prior to the manufacture of an IC it is important to ensure that any errors made in the design have been weeded out prior to this stage.
-Simulations at either at theoretical or gate/transistor levels have global usage in minimising such errors due to the ubiquity of simulators and their ease of use.\\
+As such, prior to the manufacture of an IC, it is important to ensure that any errors made in the design have been weeded out.
+Simulations either at theoretical or gate/transistor levels have global usage in minimising such errors due to the ubiquity of simulators and their ease of use.\\
 However simulations are only as good as the model used to describe the dynamics of the system, and emulating real jitter and other behaviours of a complex system is a rather difficult task.
@@
 In the case of Shan et.
-al \cite{shan2014phd} direct testing of control parameters was performed using an FPGA based design where all parameters were scaled down proportionally and thus the ideal gains for the loop filter found in simulation could be tested in a more realistic setting.\\
+al \cite{shan2014phd}, direct testing of control parameters was performed using an FPGA based design, where all parameters were scaled down proportionally and thus the ideal gains for the loop filter found in simulation, could be tested in a more realistic setting.\\
 Similarly Koskin et.
-al \cite{theboys2019} used the analysis of a number of ADPLLs implemented on an FPGA both independently and in a network to validate previously obtained theoretical results while avoiding the costs associated with the development of a custom IC.\\
+al \cite{theboys2019} used the analysis of a number of ADPLLs implemented on an FPGA, both independently and in a network, to validate previously obtained theoretical results while avoiding the costs associated with the development of a custom IC.\\
 Prototyping on an FPGA does have its drawbacks, most notably the mixed-signal circuits central to the operation of an ADPLL are not implementable on an FPGA and while most fundamental digital circuit elements are present these elements are not true transistor level implementations of their functionality but rather implemented by lookup tables and other multi-role structures on the FPGA.
-These drawbacks present a challenge to a designer as they restrict the potential architectures of key blocks.
+These drawbacks present a challenge to a designer as they restrict the potential architecture of key blocks.
 This is of particular concern when trying to mimic the behaviour of an established design on an FPGA.\\
 The particular system that this paper will discuss is an ADPLL network.
-Based on an idea first proposed in a 1995 paper by Pratt and Nguyen \cite{pratt1995distributed}, in which they proposed an alternative clock distribution network for ICs using a Cartesian grid of clocking areas, each with their own Phase Lock Loop, which has become known as a PLL Network .
+This is based on an idea first proposed in a 1995 paper by Pratt and Nguyen \cite{pratt1995distributed}, in which they proposed an alternative clock distribution network for ICs using a Cartesian grid of clocking areas, each with their own Phase Lock Loop, which has become known as a PLL Network .
 This methodology was implemented by Gutnik et.
@@
 al
-\cite{javidan2011all} proposed the implementation of such a system using All Digital PLLs which better suited the use cases of the technology and avoided some of the flaws pointed out in the original paper.\\
+\cite{javidan2011all} proposed the implementation of such a system using All Digital PLLs, which better suited the use cases of the technology and avoided some of the flaws pointed out in the original paper.\\
 This paper will examine the process of creating such a system, highlight the differences between potential designs and address some of the challenges and pitfalls that may be encountered along the way.
@@
 The paper is organised as follows: Section II describes the system that will be used on the FPGA and Section III the implementation thereof and the challenges encountered in the process.
-Finally Section IV will present example measurements made on the platform and discuss their merit.
+Finally, Section IV will present example measurements made on the platform and discuss their merit.
 
 \section{System Architecture}
-An ADPLL network as the name would suggest is created from a number of ADPLLs that coupled using digital phase comparators which attempt to measure the phase error between two oscillators, with each non edge node being connected to four neighbours.
+An ADPLL network, as the name would suggest, is created from a number of ADPLLs that are coupled using digital phase comparators which attempt to measure the phase error between two oscillators, with each non edge node being connected to four neighbours.
 Figure \ref{fig:adpll_network} illustrates such a system on a small scale, with a number of ADPLLs laid out in a Cartesian grid.
@@
 \subsection{Digitally Controlled Oscillators}
-While regular Voltage Controlled Oscillator (VCO) is tuned by the input voltage over a continuous range of values in a digital system there are a very limited number of voltages representable, most commonly just two levels.
-As such using the voltage itself to control an oscillator is not a viable strategy and instead Digitally, or as they are sometimes known as Numerically, Controlled Oscillators (DCO) accept an \textit{n} bit wide control signal that alters the period of the output waveform.
+While regular Voltage Controlled Oscillator (VCO) is tuned by the input voltage over a continuous range of values, in a digital system there are a very limited number of voltages representable, most commonly just two levels.
+As such, using the voltage itself to control an oscillator is not a viable strategy and instead Digitally, or as they are sometimes known as Numerically, Controlled Oscillators (DCO) accept an \textit{n} bit wide control signal that alters the period of the output waveform.
 While tuning range, centre frequency and linearity apply as in a VCO, a DCO also has a minimum frequency step as a result of the non continuous control values.
 %TODO WORDING
-In combination with the bit width of the control signal the minimum frequency step limits the range over which the oscillator can be tuned.
-Frequently the output of the DCO is divided down to allow the control blocks in the feedback path that are clocked on this signal to run at a lower and thus easier frequency.
+In combination with the bit width of the control signal, the minimum frequency step limits the range over which the oscillator can be tuned.
+Frequently the output of the DCO is divided down to allow the control blocks in the feedback path that are clocked on this signal, to run at a lower and thus easier frequency.
 \subsection{Loop Filter}
 The loop filter is a key aspect of a phase lock loop as without it only the current output of the phase detector could be used to compute the control voltage of the VCO.
-In an ADPLL it performs an identical role, however the digital environment results the traditional loop filter being replaced by a discrete time Proportional Integral (PI) controller, with the integral path of the controller implemented by an accumulator.
+In an ADPLL it performs an identical role however, the digital environment results the traditional loop filter being replaced by a discrete time Proportional Integral (PI) controller, with the integral path of the controller implemented by an accumulator.
 \subsection{Error Combiner}
@@
 It is often convenient for this to be performed as a weighted sum, thus avoiding different gain requirements for the loop filter of oscillators in the corners of the grid that only have a pair of neighbours.
-The ability to change the weights during operation enables ``uni-directional start-up'' which Javiden et.
-al
-\cite{javidan2011all} found prevented the system from entering unwanted stable equilibriums warned of in Pratt' and Nguyen's seminal paper.
+The ability to change the weights during operation enables ``uni-directional start-up'', which Javiden et.
+al \cite{javidan2011all} found prevented the system from entering unwanted stable equilibriums warned of in Pratt and Nguyen's seminal paper.
 \subsection{Phase Detector}
 As in a conventional PLL the phase detector outputs a value proportional to the difference in phase between a pair of signals, which is then output as a continuous voltage.
-A digital phase detector then attempts to do the same by measuring the difference between fixed reference points in the waveform, which as a digital system deals with square waves this is typically the rising edge of each input to the block.
-Similarly to the oscillator quantisation also plays a role in the behaviour of this block, with the time difference between each rising edge having a minimum possible measurement step.
+A digital phase detector then attempts to do the same by measuring the difference between fixed reference points in the waveform, which as a digital system deals with square waves, this is typically the rising edge of each input to the block.
+Similarly to the oscillator, quantisation also plays a role in the behaviour of this block, with the time difference between each rising edge having a minimum possible measurement step.
 In distributed ADPLLs the phase detector has dual outputs, each the inverse of the other, in order to feed the correct value to both of the connected error combiners.
@@
 \subsection{FPGA Restrictions}
-The use of an FPGA comes with many restrictions as to how a complex system such as an ADPLL can be designed with the mixed signal blocks, the DCO and phase detectors, in particular being curtailed by the lack of transistor level control of the design and layout.\\
+The use of an FPGA comes with many restrictions as to how a complex system such as an ADPLL can be designed with the mixed signal blocks, the DCO and phase detectors in particular being curtailed by the lack of transistor level control of the design and layout.\\
 Two potential solutions exist which alleviate this problem, and are suited to different frequency ranges and/or goals of the designer.
-Firstly the clock signal generated by the FPGA's clock distribution block can be used to drive accumulator based oscillators and phase detectors a technique suitable in two main use cases: The emulation of the performance of an already designed system operating at a proportionally cut down frequency.
+Firstly the clock signal generated by the FPGA's clock distribution block can be used to drive accumulator based oscillators and phase detectors, a technique suitable in two main use cases: The emulation of the performance of an already designed system operating at a proportionally cut down frequency.
 At frequencies that are orders of magnitude lower than the maximum obtainable clock frequency of the FPGA's clock manager.
@@
 The second solution to the problem is more akin to the real world system, and uses the delay through primitive elements such as inverters, which can be deduced from measurements on the FPGA of choice but will vary depending on the layout selected.
-In the case of the FPGA used for this paper, a Xilinx Artix-7 (XC7A100T-1CSG324C) this delay is in the region of 300 picoseconds but varies depending on the routing between elements.
+In the case of the FPGA used for this paper, a Xilinx Artix-7 (XC7A100T-1CSG324C), this delay is in the region of 300 picoseconds but varies depending on the routing between elements.
 This design is better suited to scenarios where the desired operating frequency is a significant fraction of the maximum usable clock of the FPGA or it is particularly desirable to have asynchronous behaviour in the system as in an ASIC.
@@
 \subsubsection*{Design 1}
-If frequency linearity is a desired characteristic of the oscillator an accumulator based, FPGA clock driven design exists for this use case also. 
-Here the control code $k$ is added to the value stored in the accumulator at each rising edge of the FPGA clock until overflow is reached at $2^n-1$ where $n$ is the bit width of the counter thus valuing each control code increment at $\frac{\textrm{FPGA clock}}{2^{\textrm{width}}}$ Hz.
-Once again the MSB of the accumulator is used as the output signal, however with this design the MSB will spend an equivalent duration at both logic high and low.
+If frequency linearity is a desired characteristic of the oscillator, an accumulator based, FPGA clock driven design exists for this use case also. 
+Here the control code $k$ is added to the value stored in the accumulator at each rising edge of the FPGA clock until overflow is reached at $2^n-1$, where $n$ is the bit width of the counter, thus valuing each control code increment at $\frac{\textrm{FPGA clock}}{2^{\textrm{width}}}$ Hz.
+Once again the MSB of the accumulator is used as the output signal, however, with this design, the MSB will spend an equivalent duration at both logic high and low.
 Again this design is better suited to lower frequencies, where the output of the DCO is orders of magnitude lower than the clock signal driving it.
@@
 The third and final potential DCO design is an inverter ring/chain which makes use of the delay through inverter primitives, thus bearing significant resemblance to DCOs used on ASICs.
-On an FPGA the lack of precise control over the layout of blocks means the exact value of the delay is uncertain and may vary noticeably from one implementation to another, however this misalignment and the asynchronous behaviour stemming from the lack of an FPGA driven clock increase the similarity between this design and an ASIC.
+On an FPGA, the lack of precise control over the layout of blocks, means the exact value of the delay is uncertain and may vary noticeably from one implementation to another, however, this misalignment and the asynchronous behaviour stemming from the lack of an FPGA driven clock, increase the similarity between this design and an ASIC.
 The period is set by the number of inverters at $2\times(\textrm{num inverters})\times(\textrm{delay})$, and is varied by adding or removing a pair of inverters for every control code increment.
-This design works best when the desired period is within a couple orders of magnitude of the delay through each pair of inverters, beyond this point the number of look-up tables and routing resources required to implement the design becomes unsustainable.\\
-A third potential design exists, similar to Oscillator 1 in which the control code varies the reload value of the accumulator at overflow, however this design will not be discussed in this paper.
+This design works best when the desired period is within a couple orders of magnitude of the delay through each pair of inverters, beyond this point, the number of look-up tables and routing resources required to implement the design becomes unsustainable.\\
+
+A third potential design exists, similar to Oscillator 1 in which the control code varies the reload value of the accumulator at overflow, however, this design will not be discussed in this paper.
 
 \subsection{Phase Detectors on an FPGA}
-The two options for replicating the behaviour of a phase detector also employ the same two design principles, with the option better suited to lower frequency signals once again being driven by the clock generated by the FPGA and an inverter based design better suiting higher frequency systems where the greater resolution offered is beneficial.\\
+The two options for replicating the behaviour of a phase detector also employ the same two design principles, with the option better suited to lower frequency signals once again, being driven by the clock generated by the FPGA and an inverter based design better suiting higher frequency systems, where the greater resolution offered is beneficial.\\
 A  Sign-Number (SigNum) detector is a common method of performing phase-frequency detection in an ADPLL and is the method that will be discussed in this paper.
-As the name suggests this design constitutes two sections, with one determining the sign of the phase difference and the other its magnitude.
+As the name suggests, this design constitutes two sections, with one determining the sign of the phase difference and the other, its magnitude.
 The first issue with implementing a phase detector on an FPGA is the issue of metastability, where the output of a circuit element is at an undetermined level, which can occur if both signs were to have edges within a short time of another or the set-up/hold times of a register are violated.
-This issue is easily addressed in an ASIC with an arbiter, however an arbiter is a mixed signal circuit that is not synthesisable on an FPGA.
+This issue is easily addressed in an ASIC with an arbiter, however, an arbiter is a mixed signal circuit that is not synthesisable on an FPGA.
 
 \subsubsection*{Design A}
-The first potential design uses a Finite State Machine (FSM) to determine which edge which edge has occurred first and thus the sign, and a counter that computes the time difference between each rising edge.
+The first potential design uses a Finite State Machine (FSM) to determine which edge has occurred first and thus the sign, and a counter that computes the time difference between each rising edge.
 This counter is clocked by the FPGA clock, thus giving the system a time resolution of $\frac{1}{F_{\textrm{FPGA}}}$.
-As before this results in a design that is only suitable when the frequency of operation is well below that of the FPGA clock due to resolution constraints.
-In order to avoid the problem of metastability the incoming waveforms can be synchronised to the clock of the FPGA for no penalty beyond the extra clock cycle taken as this synchronisation performs the same quantisation that would otherwise have occurred in the counter measuring the time difference.
-This design is better suited for use with the accumulator based oscillators due to the overlapping frequency ranges, however it can also be used in conjunction with an inverter ring.
-Important to note is that in the case of the Artix-7 used in this paper this results in a four times worse phase difference resolution than period resolution which made low jitter operation difficult, but this combination could be made possible by removing eight rather than two inverters for each control code step.
+As with the FPGA driven oscillator, this results in a design that is only suitable when the frequency of operation is well below that of the FPGA clock due to resolution constraints.
+In order to avoid the problem of metastability, the incoming waveforms can be synchronised to the clock of the FPGA for no penalty beyond the extra clock cycle taken, as this synchronisation performs the same quantisation that would otherwise have occurred in the counter measuring the time difference.
+This design is better suited for use with the accumulator based oscillators due to the overlapping suitable frequency ranges, however, it can also be used in conjunction with an inverter ring.
+Important to note is that in the case of the Artix-7 used in this paper, this results in a four times worse phase difference resolution than period resolution which made low jitter operation difficult, but this combination could be made possible by removing eight rather than two inverters for each control code step.
 %TODO plot this later
@@
 \subsubsection*{Design B}
-Better suited for use with the inverter based oscillator is a system that mimics the behaviour of a tapped delay line working asynchronously to the FPGA clock.
-As the FPGA clock is not used, and that of the ADPLL cannot be used to measure itself an unclocked sign detection circuit is required, an example of which can be seen in Figure \ref{fig:segn_pdetdl}.
-This is nearly identical to sign detection methods used on an ASIC with the arbitration circuit replaced by a second SR Latch which alleviates the issue of metastability.\\
-A tapped delay line can be emulated on an FPGA through the use of inverter pairs due to their short and approximately known delay placed between each tap.
+Better suited for use with the inverter based oscillator, is a system that mimics the behaviour of a tapped delay line, working asynchronously to the FPGA clock.
+As the FPGA clock is not used, and that of the ADPLL cannot be used to measure itself, an unclocked sign detection circuit is required, an example of which can be seen in Figure \ref{fig:segn_pdetdl}.
+This is nearly identical to sign detection methods used on an ASIC, with the arbitration circuit replaced by a second SR Latch, which alleviates the issue of metastability.\\
+A tapped delay line can be emulated on an FPGA through the use of inverter pairs, due to their short and approximately known delay placed between each tap.
 The resolution then is the propagation time through the pair, although this will once again depend on the layout specifics, which was approximately 600 picoseconds on an Artix-7.
-This is significantly lower than the 4 nanoseconds achieved using the previous design and thus more suitable for use with an inverter ring oscillator as this resolution is better than the time resolution of the oscillator.
+This is significantly lower than the 4 nanoseconds achieved using the previous design and thus more suitable for use with an inverter ring oscillator, as this resolution is better than the time resolution of the oscillator.
 Figure \ref{fig:num_pdetdl} illustrates the configuration of the circuit used to emulate the function of a delay line that will be used in this paper.
@@
 \section{Comparison of Designs}
-This paper will examine the performance of a number of combinations of these blocks and the performance thereof, specifically oscillator 1 with phase detector A and oscillator 2 with detectors A \& B.
+This paper will now examine the performance of a number of combinations of these blocks and the performance thereof, specifically oscillator 1 with phase detector A and oscillator 2 with detectors A \& B.
 For each test there were four ADPLLs on the FPGA arranged in a Cartesian Grid and as such each ADPLL will be referred to using matrix indexing corresponding to their grid location.
 For each configuration three different modes of operation were examined: Each ADPLL locking to the same 5 MHz external reference, each ADPLL connected to the left and above neighbours in a ``uni-directional'' network and finally a true ADPLL network where all neighbours are connected.
-For comparing each configuration gains which had previously been determined to provide the best jitter performance were chosen.
+For comparison, each configuration gains which had previously been determined to provide the best jitter performance were chosen.
 The data was captured using an Agilent MSO7054 A at 4 GSa/s.
-Each point of data is the mean of five 200 microsecond long captures which were then analysed in Matlab, and Table \ref{table:b_c_perf} displays the results of this analysis.\\
-The first columns of the table relate to the cycle-to-cycle jitter of the system, the difference between sequential distances, and the second portion gives the Time Interval Error, or difference between the actual timing of the edge and the ``ideal'' time as set by the reference.
+Each point of data is the mean of five 200 microsecond long captures which were then analysed in Matlab, Table \ref{table:b_c_perf} displays the results of this analysis.\\
+The first columns of the table relate to the cycle-to-cycle jitter of the system, the difference between sequential rising edges, and the second portion gives the Time Interval Error, the difference between the actual timing of the edge and the ``ideal'' time as set by the reference.
 This combination of measurements will show how well each design performs the task of staying consistent from cycle to cycle both in relation to itself and the system at large.
@@
 
-From the table it is apparent that the combination of the inverter chain DCO and the FPGA clock driven phase detector is an ill advised choice, as the jitter present in the system is several orders of magnitude worse than either other situation.
-In this system the FPGA clock is at 258 MHz, the maximum possible, thus giving a detection resolution of 4 nanoseconds.
-This is 4 times greater than the period resolution of approximately 1.2 nano seconds of the oscillator, and requires the gains of the PI filter to be increased fourfold in order to respond appropriately to the error detected.
+From the table it is apparent that the combination of the inverter chain DCO and the FPGA clock driven phase detector is an ill-advised choice, as the jitter present in the system is several orders of magnitude worse than either other situation.
+In this system, the FPGA clock is at 258 MHz, the maximum possible, thus giving a detection resolution of 3.875 nanoseconds.
+This is 3.5 times greater than the period resolution of approximately 1.2 nano second period resolution of the oscillator, and requires the gains of the PI filter to be increased fourfold in order to respond appropriately to the error detected.
 The problems of relatively inaccurate detection and increased gains combine to produce a system that is only able to achieve poor locking performance.
-Both other combinations work well however, delivering a level of jitter that is both low yet allows an indication how impactful variations in architecture or control parameters to be seen visually on an oscilloscope.\\
-\begin{figure}[h]
+Both other combinations work well however, delivering a level of jitter that is both low yet allows an indication of how impactful variations in architecture or control parameters are to be seen visually on an oscilloscope.\\
+
+\begin{figure}[h]%
     \centering
-	\includegraphics[width=0.25\textwidth]{distrib_ring}
-	\caption{Distribution of Periods for Inverter Based Design.}
-	\label{fig:dist1}
+    \subfloat[Inverter Based Design.]{{\includegraphics[width=0.5\linewidth]{distrib_ring} }}%
+    \subfloat[FPGA Clocked Design.]{{\includegraphics[width=0.5\linewidth]{distrib_pa} }}%
+    \caption{Example Distribution of Periods.\label{fig:dists}}    
 \end{figure}
-\begin{figure}[h]
-	\centering
-	\includegraphics[width=0.25\textwidth]{distrib_pa}
-	\caption{Distribution of Periods for FPGA Clocked Design.}
-	\label{fig:dist2}
-\end{figure}
-Figures \ref{fig:dist1} \& \ref{fig:dist2} paint a picture of the unrealistic nature of the FPGA clocked designs as in comparison to the ring based design each oscillator will have identical intrinsic period steps thus eliminating any jitter that maybe be seen in the system due to misalignment.
-This effects the suitability of the FPGA clocked designs for the improvement of simulation models or the examination of how a novel block maybe impact the behaviour of the network.
+%\begin{figure}[h]
+%   \centering
+%   \subfloat[FPGA Clocked Design][width=0.25\textwidth]{distrib_pa}
+%   \caption{Distribution of Periods.}
+%   \label{fig:dist2}
+%\end{figure}
+Figure \ref{fig:dists} paints a picture of the unrealistic nature of the FPGA clocked designs as in comparison to the ring based design, each oscillator will have identical intrinsic period steps thus eliminating any jitter that maybe be seen in the system due to misalignment.
+This effects the suitability of the FPGA clocked designs for the improvement of simulation models, or the examination of how a novel block may impact the behaviour of the network.
 \section*{Conclusion}
 An FPGA is a suitable environment for the prototyping of ADPLL networks and can fill a number of roles depending on the designer's requirements.
-However the designer should bear in mind the platform's limitations especially when attempting to use a frequency that is not orders of magnitude lower than the maximum clock available from the FPGA's clock manager.
-Inverter based and FPGA clocked designs both have a niche in which they are the preferred option, with the clocked design better suiting replication of an existing design at scaled down frequencies and the inverter based designs better in the modelling and analysis of ADPLL network behaviour.
+However, the designer should bear in mind the platform's limitations, especially when attempting to use a frequency that is not orders of magnitude lower than the maximum clock available from the FPGA's clock manager.
+Inverter based and FPGA clocked designs, both have a niche in which they are the preferred option, with the clocked design better suiting replication of an existing design at scaled-down frequencies and the inverter based designs better in the modelling and analysis of ADPLL network behaviour.
 

### 2019-03-25 conf_paper/paper.tex
meas1 & 3

@@
 \author{\IEEEauthorblockN{C. Dooley, E. Blokhina, B. Mulkeen}
-    \IEEEauthorblockA{\textit{School of Electronic Engineering} \\
-        \textit{University College Dublin}\\
-        Belfield, Dublin 4, Ireland. \\
-        conor.dooley@ucdconnect.ie} %TODO
+    \IEEEauthorblockA{\textit{School of Electrical \& Electronic Engineering, University College Dublin}
 }
@@
 \maketitle
-
 \begin{abstract}
@@
 \end{abstract}
-
-\begin{IEEEkeywords}
-    FPGA, ADPLL, Simulation, Prototyping
-\end{IEEEkeywords}
 \section{Motivation}

### 2019-03-25 matlab/a1_3.m

@@ -1,13 +1,81 @@
-filename = 'scope_41.bin';
-[~,scope_y_data] = importAgilentBin(filename,1);
-[~,scope_y_data2] = importAgilentBin(filename,2);
-[~,scope_y_data3] = importAgilentBin(filename,3);
-[scope_x_data,scope_y_data4] = importAgilentBin(filename,4);
+for inc = 1:15
+    filename = char("new/z1a_" + num2str(inc) + ".bin");
+    [~,scope_y_data1] = importAgilentBin(filename,1);
+    [~,scope_y_data2] = importAgilentBin(filename,2);
+    [~,scope_y_data3] = importAgilentBin(filename,3);
+    [scope_x_data,scope_y_data4] = importAgilentBin(filename,4);
 
-f_samp = 4E9;
-num_samp = length(scope_y_data);
-t_max = num_samp*(1/f_samp); 
-t=linspace(0,t_max,num_samp);
+    f_samp = 4E9;
+    num_samp = length(scope_y_data1);
+    t_max = num_samp*(1/f_samp); 
+    t=linspace(0,t_max,num_samp);
+
+    [~,init_cross1,~,~] = pulsewidth(scope_y_data1,scope_x_data, 'Polarity', 'Positive');
+    [~,init_cross2,~,~] = pulsewidth(scope_y_data2,scope_x_data, 'Polarity', 'Positive');
+    [~,init_cross3,~,~] = pulsewidth(scope_y_data3,scope_x_data, 'Polarity', 'Positive');
+    [~,init_cross4,~,~] = pulsewidth(scope_y_data4,scope_x_data, 'Polarity', 'Positive');
+
+    periods1 = getPeriods(init_cross1); 
+    periods2 = getPeriods(init_cross2);
+    periods3 = getPeriods(init_cross3);
+    periods4 = getPeriods(init_cross4);
+
+    mean_period_1 = mean(periods1);
+    mean_period_2 = mean(periods2);
+    mean_period_3 = mean(periods3);
+    mean_period_4 = mean(periods4);
+
+    std_period_2(inc) = std(periods2);
+    std_period_3(inc) = std(periods3);
+    std_period_4(inc) = std(periods4);
+
+%     max_2 = max(periods2-mean(periods2));
[…]

### 2019-03-25 matlab/cd___anal.m

@@ -1,24 +1,25 @@
-filename = 'c_____0.bin';
-[scope_x_data,scope_y_data] = importAgilentBin(filename,3);
+for inc = 1:15
+    filename = char("new/des2a_" + num2str(inc) + ".bin");
+    [scope_x_data,scope_y_data] = importAgilentBin(filename,3);
 
-f_samp = 4E9;
-num_samp = length(scope_y_data);
-t_max = num_samp*(1/f_samp); 
-t=linspace(0,t_max,num_samp);
+    f_samp = 4E9;
+    num_samp = length(scope_y_data);
+    t_max = num_samp*(1/f_samp); 
+    t=linspace(0,t_max,num_samp);
 
-[~,init_cross4,~,~] = pulsewidth(scope_y_data,scope_x_data, 'Polarity', 'Positive');
+    [~,init_cross4,~,~] = pulsewidth(scope_y_data,scope_x_data, 'Polarity', 'Positive');
 
-periods4 = getPeriods(init_cross4);
+    periods4 = getPeriods(init_cross4);
 
-std_period_4 = std(periods4);
+    std_period_4 = std(periods4);
 
-max_4 = max(periods4-mean(periods4));
-min_4 = min(periods4-mean(periods4));
-p2p_4 = max_4-min_4;
+    max_4 = max(periods4-mean(periods4));
+    min_4 = min(periods4-mean(periods4));
+    p2p_4 = max_4-min_4;
 
-clc
-string = num2str(std_period_4) + ", " + num2str(max_4) + ", " + num2str(min_4) + ", " + num2str(p2p_4)
-
+    clc
+    string = num2str(std_period_4) + ", " + num2str(max_4) + ", " + num2str(min_4) + ", " + num2str(p2p_4)
+end
 
 
 

### 2019-03-25 conf_paper/paper.tex
tweaks and measurements

@@
 
-\title{FPGA Based Prototyping of an ADPLL Network.\\}
+\title{FPGA Based Prototyping of an ADPLL Network\\}
 
@@
 	}
-
+}
 \maketitle
+\vspace{-4.5cm}
 \begin{abstract}
@@
     \label{fig:adpll_base}
+    \vspace{-0.5cm}
 \end{figure}
@@
     \label{fig:segn_pdetdl}
+    \vspace{-0.5cm}
 \end{figure}
@@
     \label{fig:num_pdetdl}
+    \vspace{-0.5cm}
 \end{figure}
@@
         \begin{tiny}
+        	\setlength{\tabcolsep}{.9\tabcolsep}
             \begin{tabular}{cc|c|c|c|c|c|c|}           
                 \cline{3-8}
-                && \multicolumn{3}{c|}{Jitter Standard Deviation (ns)} & \multicolumn{3}{c|}{Time Interval Error. (ns)} \T\\
+                && \multicolumn{3}{c|}{Jitter Standard Deviation (ns)} & \multicolumn{3}{c|}{Max. Time Interval Error. (ns)} \T\\
                 \cline{3-8} 
@@
                 \multicolumn{2}{|l|}{Osc 1, PDET A}&-&-&-&-&-&-\T\\
-                \multicolumn{2}{|r|}{Free PLLs}&1.5988&1.6178&1.6220&&& \T\\
-                \multicolumn{2}{|r|}{Uni-dir.}&1.9804&2.0316&2.1317&&& \T\\
-                \multicolumn{2}{|r|}{Bi-dir.}&2.0124&2.0681&2.0732&&& \T\\
+                \multicolumn{2}{|r|}{Free PLLs} &1.4377 &1.4714&1.4217   &6.1910&6.3667&6.8052 \T\\
+                \multicolumn{2}{|r|}{Uni-dir.}  &0.91589&1.0993&1.0875   &8.3619&8.9768&10.974 \T\\
+                \multicolumn{2}{|r|}{Bi-dir.}   &1.0470 &1.0979&1.0306   &24.998&33.767&33.204 \T\\
                 \hline
                 \multicolumn{2}{|l|}{Osc 2, PDET A}&-&-&-&-&-&-\T\\
-                \multicolumn{2}{|r|}{Free PLLs}&0.62499 $\mu$s&0.75919 $\mu$s&0.79049  $\mu$s&&& \T\\
-                \multicolumn{2}{|r|}{Uni-dir.}&0.76822 $\mu$s&0.9009 $\mu$s&1.1407 $\mu$s&&& \T\\
-                \multicolumn{2}{|r|}{Bi-dir.}&0.70393 $\mu$s&0.70201 $\mu$s&1.241 $\mu$s&&& \T\\
+                \multicolumn{2}{|r|}{Free PLLs}&&& 0&&& \T\\
+                \multicolumn{2}{|r|}{Uni-dir.}&&& 0&&& \T\\
+                \multicolumn{2}{|r|}{Bi-dir.}&&& 0&&& \T\\
                 \hline
                 \multicolumn{2}{|l|}{Osc 2, PDET B}&-&-&-&-&-&-\T\\
-                \multicolumn{2}{|r|}{Free PLLs}&4.1680&3.9803&5.0741&&& \T\\
-                \multicolumn{2}{|r|}{Uni-dir.}&4.3820&5.7315&6.2336 &&& \T\\
-                \multicolumn{2}{|r|}{Bi-dir.}&5.6137&7.1831&7.1612 &&& \T\\
+                \multicolumn{2}{|r|}{Free PLLs} &7.8090&7.8827&6.6272   &17.336&21.151&24.302 \T\\
+                \multicolumn{2}{|r|}{Uni-dir.}  &7.7866&8.1145&7.3598   &17.649&29.962&37.051 \T\\
+                \multicolumn{2}{|r|}{Bi-dir.}   &7.7575&7.7729&6.6962   &30.019&40.577&45.617 \T\\
                 \hline
@@
     \end{center}
+\vspace{-0.5cm}
 \end{table}
-
 From the table it is apparent that the combination of the inverter chain DCO and the FPGA clock driven phase detector is an ill-advised choice, as the jitter present in the system is several orders of magnitude worse than either other situation.
@@
     \caption{Example Distribution of Periods.\label{fig:dists}}    
+    \vspace{-0.5cm}
 \end{figure}

### 2019-03-25 matlab/a1_3.m

@@ -1,5 +1,6 @@
-for inc = 1:15
-    filename = char("new/z1a_" + num2str(inc) + ".bin");
+for inc = 1:1
+    %filename = char("new/z1a_" + num2str(inc) + ".bin");
+    filename = 'cdcd5.bin';
     [~,scope_y_data1] = importAgilentBin(filename,1);
     [~,scope_y_data2] = importAgilentBin(filename,2);
     [~,scope_y_data3] = importAgilentBin(filename,3);
@@ -15,6 +16,7 @@ for inc = 1:15
     [~,init_cross3,~,~] = pulsewidth(scope_y_data3,scope_x_data, 'Polarity', 'Positive');
     [~,init_cross4,~,~] = pulsewidth(scope_y_data4,scope_x_data, 'Polarity', 'Positive');
 
+    len = length(init_cross1);
     periods1 = getPeriods(init_cross1); 
     periods2 = getPeriods(init_cross2);
     periods3 = getPeriods(init_cross3);
@@ -38,11 +40,25 @@ for inc = 1:15
 %     max_4 = max(periods4-mean(periods4));
 %     min_4 = min(periods4-mean(periods4));
 %     p2p_4 = max_4-min_4;
-
-    TIE_2(inc) = max(abs(periods2-mean_period_1));
-    TIE_3(inc) = max(abs(periods3-mean_period_1));
-    TIE_4(inc) = max(abs(periods4-mean_period_1));
-
+%     length(init_cross2) - length(init_cross1)
+%     if length(init_cross2) == length(init_cross1)
+%         TIE_2(inc) = max(abs(init_cross2(1:len)-init_cross1));
+%     elseif length(init_cross2) > length(init_cross1)
+%         TIE_2(inc) = max(abs(init_cross2(2:end)-init_cross1));
+%         inc
+%     end
+%     if length(init_cross3) == length(init_cross1)
+%         TIE_3(inc) = max(abs(init_cross3(1:len)-init_cross1));
+%     elseif length(init_cross3) > length(init_cross1)
+%         TIE_3(inc) = max(abs(init_cross3(2:end)-init_cross1));
+%         inc
+%         end
+%     if length(init_cross4) == length(init_cross1)
+%         TIE_4(inc) = max(abs(init_cross4(1:len)-init_cross1));
[…]

### 2019-03-26 conf_paper/critique.tex
critique done, 1.5 pages w/ stupid textwidth, more measurements added

\documentclass[a4paper,12pt]{article}

\title{Conference Paper Critique}
\author{Conor Dooley}

\begin{document}

\maketitle

The intention is to submit this paper to the 2019 edition of the International Conference on Synthesis, Modelling, Analysis and Simulation Methods and Applications to Circuit Design (SMACD), a conference dedicated to the modelling and simulation of, among other areas, mixed-signal circuits which aligns with the paper's goal of modelling an ADPLL network, itself a mixed-signal circuit. The final deadline for submission to this conference is the 1st of April 2019, the Monday after this submission is due internally, which is quite convenient. The paper is most suited to this conference as it describes the challenges and realities of implementing a platform for the hardware modelling and simulation of a complex system in a cost and time inexpensive manner. The paper could also possibly be submitted to the 2019 3rd International Conference on Circuits, System and Simulation (ICCSS 2019) with a deadline of the 12th of April, however, this conference is in China so if this paper was to be accepted this would present a logistical difficulty.\\

This paper was inspired by the usage of FPGAs as a prototyping tool by Eldar Zianbetov and Chuan Shan in the course of their development of an ADPLL network on an ASIC, however this was carried out at a significantly reduced frequency of operation. Here in UCD the technique has been used by Eugene Koskin and Pierre Bisleaux in their as of yet unpublished paper to validate theoretical models developed by Eugene during the course of his PhD. The initial goal of the project was to investigate a number of possible architectures for this prototyping platform, and this paper explains the designs analysed and the benefits of each technique \\

The two most glaring omissions in the paper are the third oscillator design mentioned in the course of the paper yet is not analysed and the rather poor performance of Design ``1a'' (ring oscillator and clocked phase detector) which was unable to acquire lock at 5 MHz and had to be analysed at the reduced frequency of 1.25 MHz (the output of the in-circuit divider). %TODO
The third oscillator, briefly mentioned in the paper, is of the type most similar to that used by Zianbetlov/Shan in their paper, which like Oscillator 1 is made up of a counter driven by the FPGA clock. However, unlike Oscillator 1 it is linear in period rather than frequency thus making it more comparable to the ring oscillator design I used. It would make for a more appropriate comparison of designs and has greater correlation with real world usage had it been used but was not as Oscillator 1 was chosen as the clocked design early in the project.\\

There is also a lack of detail regarding the implementation of each module used in the design, beyond images of the detector and high level descriptions due to the limited space available. Similarly the technical motivation does not contain a particularly large number of references beyond those demonstrating the usage of FPGAs in this role and the explanation of where the idea of ADPLL networks came from. This is due to the lack of publications addressing hardware simulation of such networks and the assumption the reader will be aware of the blocks involved in a Phase Lock Loop.

\end{document}

### 2019-03-26 conf_paper/paper.tex
critique done, 1.5 pages w/ stupid textwidth, more measurements added

@@
 
-\title{FPGA Based Prototyping of an ADPLL Network\\}
+\title{FPGA Based Modelling of an ADPLL Network\\}
 
@@
 An FPGA is ideal for this task as it leverages the existing skill set of a digital designer, permits the re-use of certain blocks, and most importantly enables cost effective and rapid reworking of the design.\\
-An FPGA as a prototyping tool has seen use in UPMC Sorbonne and University College Dublin, and has been used in the testing of established designs prior to their implementation in custom silicon\cite{zianbetov2013phd,shan2014phd}, the validation of realistic models for use in high level simulations\cite{theboys2019} and the exploration of new designs for digital blocks before progressing on to true digital design on a gate level.
+An FPGA as a prototyping tool has seen use in UPMC Sorbonne and University College Dublin, and has been used in the testing of established designs prior to their implementation in custom silicon \cite{zianbetov2013phd,shan2014phd}, the validation of realistic models for use in high level simulations\cite{theboys2019} and the exploration of new designs for digital blocks before progressing on to true digital design on a gate level.
 %TODO references
@@
 al
-\cite{javidan2011all} proposed the implementation of such a system using All Digital PLLs, which better suited the use cases of the technology and avoided some of the flaws pointed out in the original paper.\\
+\cite{javidan2011all} proposed the implementation of such a system using All Digital PLLs, which better suited the use cases of the technology and avoided some of the flaws pointed out in the original paper.
+Zianbetov and Shan followed on from this proposition by implementing such a network on a custom IC \cite{zianbetov2013phd,shan2014phd}.\\
 This paper will examine the process of creating such a system, highlight the differences between potential designs and address some of the challenges and pitfalls that may be encountered along the way.

### 2019-03-26 conf_paper/paper.tex
paper measurements and reword finished

@@
     \IEEEauthorblockA{\textit{School of Electrical \& Electronic Engineering, University College Dublin}
+    	\vspace{-1.2cm}
 	}
@@
 \maketitle
-\vspace{-4.5cm}
 \begin{abstract}
@@
                 \multicolumn{2}{|l|}{Osc 2, PDET A}&-&-&-&-&-&-\T\\
-                \multicolumn{2}{|r|}{Free PLLs}&&& 0&&& \T\\
-                \multicolumn{2}{|r|}{Uni-dir.}&&& 0&&& \T\\
-                \multicolumn{2}{|r|}{Bi-dir.}&&& 0&&& \T\\
+                \multicolumn{2}{|r|}{Free PLLs} &78.458&74.025&67.779  &146.09&143.70&142.03 \T\\
+                \multicolumn{2}{|r|}{Uni-dir.}  &78.428&72.646&65.665  &146.23&298.49&200.10 \T\\
+                \multicolumn{2}{|r|}{Bi-dir.}   &78.436&73.827&67.648  &235.73&497.60&311.28 \T\\
                 \hline
@@
 \end{table}
-From the table it is apparent that the combination of the inverter chain DCO and the FPGA clock driven phase detector is an ill-advised choice, as the jitter present in the system is several orders of magnitude worse than either other situation.
+It is apparently that the combination of the inverter chain DCO and the FPGA clocked phase detector is an ill advised choice, with the system's jitter being an order of magnitude worse than the design using inverter delay based detector.
+However when the resolution of the detector is then considered, a potential cause for this issue is found.
+The detector in case two has a resolution of 3.875 nanoseconds in comparison to the 1.2 nanosecond period step of the oscillator, requiring an increase in the loop filter gains.
+The problems of relatively inaccurate detection and increased gains combine to produce a system that is only able to achieve poor locking performance.
+In case three then the detector's resolution improves to 0.6 nanoseconds and produces relatively good system performance.
+This combination does not perform as well as that of case one, but that is attributable to the unrealistic nature of the FPGA clocked design.
+Both case one and three work well, delivering a level of jitter that is both low yet allows an indication of how impactful variations in architecture or control parameters are to be seen visually on an oscilloscope with appropriate persistence.\\
+
+%From the table it is apparent that the combination of the inverter chain DCO and the FPGA clock driven phase detector is an ill-advised choice, as the jitter present in the system is several orders of magnitude worse than either other situation.
+In order to obtain these measurements a four times divider was inserted into the feedback path, and hence the presence of jitter greater than the period of 200 ns.
 In this system, the FPGA clock is at 258 MHz, the maximum possible, thus giving a detection resolution of 3.875 nanoseconds.
-This is 3.5 times greater than the period resolution of approximately 1.2 nano second period resolution of the oscillator, and requires the gains of the PI filter to be increased fourfold in order to respond appropriately to the error detected.
-The problems of relatively inaccurate detection and increased gains combine to produce a system that is only able to achieve poor locking performance.
-Both other combinations work well however, delivering a level of jitter that is both low yet allows an indication of how impactful variations in architecture or control parameters are to be seen visually on an oscilloscope.\\
+%This is 3.5 times greater than the period resolution of approximately 1.2 nano second period resolution of the oscillator, and requires the gains of the PI filter to be increased fourfold in order to respond appropriately to the error detected.
+The
+%Both other combinations work well however, delivering a level of jitter that is both low yet allows an indication of how impactful variations in architecture or control parameters are to be seen visually on an oscilloscope.\\
 
@@
 %\end{figure}
-Figure \ref{fig:dists} paints a picture of the unrealistic nature of the FPGA clocked designs as in comparison to the ring based design, each oscillator will have identical intrinsic period steps thus eliminating any jitter that maybe be seen in the system due to misalignment.
+Figure \ref{fig:dists} paints a picture of the unrealistic nature of the FPGA clocked design in comparison to its inverter based counterpart, each oscillator will have identical intrinsic period steps thus eliminating any jitter that maybe be seen in the system due to misalignment.
 This effects the suitability of the FPGA clocked designs for the improvement of simulation models, or the examination of how a novel block may impact the behaviour of the network.

### 2019-03-26 matlab/a1_3.m

@@ -1,6 +1,6 @@
 for inc = 1:1
-    %filename = char("new/z1a_" + num2str(inc) + ".bin");
-    filename = 'cdcd5.bin';
+    %filename = char("new/des2a_" + num2str(inc) + ".bin");
+    filename = char("scope_41.bin");
     [~,scope_y_data1] = importAgilentBin(filename,1);
     [~,scope_y_data2] = importAgilentBin(filename,2);
     [~,scope_y_data3] = importAgilentBin(filename,3);
@@ -40,7 +40,7 @@ for inc = 1:1
 %     max_4 = max(periods4-mean(periods4));
 %     min_4 = min(periods4-mean(periods4));
 %     p2p_4 = max_4-min_4;
-%     length(init_cross2) - length(init_cross1)
+% 
 %     if length(init_cross2) == length(init_cross1)
 %         TIE_2(inc) = max(abs(init_cross2(1:len)-init_cross1));
 %     elseif length(init_cross2) > length(init_cross1)
@@ -102,9 +102,9 @@ mean_jitter_4_s3 = mean(std_period_4(11:15))
 % 
 figure
 hold on;
-histfit((periods2)*1e9)
-histfit((periods3)*1e9)
-histfit((periods4)*1e9)
+histfit((periods2)*1e9+6)
+histfit((periods3)*1e9+6)
+histfit((periods4)*1e9+6)
 title('')
 xlabel('Period (ns)','fontsize',16);
 ylabel('Occurance Rate','fontsize',16);

### 2019-03-27 conf_paper/paper.tex
fdsjkl

@@
 The technical motivation underlying this paper is the creation of an Field Programmable Gate Array (FPGA) based prototyping platform for All Digital Phase Lock Loop (ADPLL) networks.
-As the creation of Application Specific Integrated Circuits (ASICs) is an expensive and time consuming process, any mistake has the potential to be rather costly.
-As such, prior to the manufacture of an IC, it is important to ensure that any errors made in the design have been weeded out.
+As the creation of Application Specific Integrated Circuits (ASICs) is an expensive and time consuming process, and mistakes made in the design may require the manufacture of a new chip, it is important to ensure that any errors made in the design have been located and addressed before manufacture..
 Simulations either at theoretical or gate/transistor levels have global usage in minimising such errors due to the ubiquity of simulators and their ease of use.\\
@@
 While regular Voltage Controlled Oscillator (VCO) is tuned by the input voltage over a continuous range of values, in a digital system there are a very limited number of voltages representable, most commonly just two levels.
-As such, using the voltage itself to control an oscillator is not a viable strategy and instead Digitally, or as they are sometimes known as Numerically, Controlled Oscillators (DCO) accept an \textit{n} bit wide control signal that alters the period of the output waveform.
+As such, using the voltage itself to control an oscillator is not a viable strategy and instead Digitally (or Numerically) Controlled Oscillators (DCO) accept an \textit{n} bit wide control signal that alters the period of the output waveform.
 While tuning range, centre frequency and linearity apply as in a VCO, a DCO also has a minimum frequency step as a result of the non continuous control values.
@@
 The loop filter is a key aspect of a phase lock loop as without it only the current output of the phase detector could be used to compute the control voltage of the VCO.
-In an ADPLL it performs an identical role however, the digital environment results the traditional loop filter being replaced by a discrete time Proportional Integral (PI) controller, with the integral path of the controller implemented by an accumulator.
+In an ADPLL it performs an identical role however, the digital environment results in the traditional loop filter being replaced by a discrete time Proportional Integral (PI) controller, with the integral path of the controller implemented by an accumulator.
 \subsection{Error Combiner}
@@
 \subsection{Phase Detector}
-As in a conventional PLL the phase detector outputs a value proportional to the difference in phase between a pair of signals, which is then output as a continuous voltage.
-A digital phase detector then attempts to do the same by measuring the difference between fixed reference points in the waveform, which as a digital system deals with square waves, this is typically the rising edge of each input to the block.
-Similarly to the oscillator, quantisation also plays a role in the behaviour of this block, with the time difference between each rising edge having a minimum possible measurement step.
+In a conventional PLL the phase detector measures a value proportional to the difference in phase between a pair of signals, which is then output as a continuous voltage.
+A digital phase detector then attempts to do the same by measuring the time difference between fixed reference points in the waveform, which as a digital system deals with square waves, is typically the rising edge of each input to the block.
+As in the oscillator, quantisation also plays a role in the behaviour of this block, with the time difference between each rising edge having a minimum possible measurement step.
 In distributed ADPLLs the phase detector has dual outputs, each the inverse of the other, in order to feed the correct value to both of the connected error combiners.
@@
 Two potential solutions exist which alleviate this problem, and are suited to different frequency ranges and/or goals of the designer.
-Firstly the clock signal generated by the FPGA's clock distribution block can be used to drive accumulator based oscillators and phase detectors, a technique suitable in two main use cases: The emulation of the performance of an already designed system operating at a proportionally cut down frequency.
-At frequencies that are orders of magnitude lower than the maximum obtainable clock frequency of the FPGA's clock manager.
-The downside of this approach is that the minimum steps are dictated by the clock frequency of the FPGA, which at frequencies that are not orders of magnitude lower than the FPGA clock will result in coarse resolutions for detector or oscillator, and thus greater jitter seen at the oscillator output.\\
+Firstly the clock signal generated by the FPGA's clock distribution block can be used to drive accumulator based oscillators and phase detectors, a technique suitable in two main use cases: The emulation of the performance of an already designed system operating at a proportionally cut down frequency or if the target frequency is low.
+The downside of this approach is that the minimum steps are dictated by the clock frequency of the FPGA, which at frequencies that are not orders of magnitude lower than the FPGA clock will result in coarse resolutions for detector or oscillator, and thus greater jitter seen at the oscillator output.
+Accordingly this method only suitable for use at frequencies that are orders of magnitude lower than the maximum obtainable clock frequency of the FPGA's clock manager.\\
 The second solution to the problem is more akin to the real world system, and uses the delay through primitive elements such as inverters, which can be deduced from measurements on the FPGA of choice but will vary depending on the layout selected.
@@
 \subsection{Digitally Controlled Oscillators on an FPGA}
-There are three key designs that be used to implement a DCO on an FPGA, two of which rely on the FPGA clock and one using inverter primitives.
-Each of these designs fulfil a different role in mimicking the behaviour of an ASIC based mixed signal DCO.
+There are three key designs that be used to implement a DCO on an FPGA, two relying on the FPGA clock and one using inverter primitives.
+Each of these designs fulfils a different role in mimicking the behaviour of an ASIC based mixed signal DCO.
 
@@
 As the name suggests, this design constitutes two sections, with one determining the sign of the phase difference and the other, its magnitude.
-The first issue with implementing a phase detector on an FPGA is the issue of metastability, where the output of a circuit element is at an undetermined level, which can occur if both signs were to have edges within a short time of another or the set-up/hold times of a register are violated.
+The first issue with implementing a phase detector on an FPGA is the issue of metastability, where the output of a circuit element is at an undetermined level, which can occur if both signals were to have edges within a short time of one another or the set-up/hold times of a register are violated.
 This issue is easily addressed in an ASIC with an arbiter, however, an arbiter is a mixed signal circuit that is not synthesisable on an FPGA.
@@
 \subsubsection*{Design A}
-The first potential design uses a Finite State Machine (FSM) to determine which edge has occurred first and thus the sign, and a counter that computes the time difference between each rising edge.
+The first potential design uses a Finite State Machine (FSM) to determine which edge has occurred first and thus the sign, and a counter that computes the time difference between each rising edge of each signal.
 This counter is clocked by the FPGA clock, thus giving the system a time resolution of $\frac{1}{F_{\textrm{FPGA}}}$.
@@
 This design is better suited for use with the accumulator based oscillators due to the overlapping suitable frequency ranges, however, it can also be used in conjunction with an inverter ring.
-Important to note is that in the case of the Artix-7 used in this paper, this results in a four times worse phase difference resolution than period resolution which made low jitter operation difficult, but this combination could be made possible by removing eight rather than two inverters for each control code step.
+%Important to note is that in the case of the Artix-7 used in this paper, this results in a four times worse phase difference resolution than period resolution which made low jitter operation difficult, but this combination could be made possible by removing eight rather than two inverters for each control code step.
 %TODO plot this later
@@
 For each configuration three different modes of operation were examined: Each ADPLL locking to the same 5 MHz external reference, each ADPLL connected to the left and above neighbours in a ``uni-directional'' network and finally a true ADPLL network where all neighbours are connected.
-For comparison, each configuration gains which had previously been determined to provide the best jitter performance were chosen.
-The data was captured using an Agilent MSO7054 A at 4 GSa/s.
+For comparison, each configuration uses loop filter gains which had previously been determined to provide the best jitter performance.
+The data was captured using an Agilent MSO7054-A oscilloscope at 4 GSa/s.
 Each point of data is the mean of five 200 microsecond long captures which were then analysed in Matlab, Table \ref{table:b_c_perf} displays the results of this analysis.\\
@@
 However when the resolution of the detector is then considered, a potential cause for this issue is found.
-The detector in case two has a resolution of 3.875 nanoseconds in comparison to the 1.2 nanosecond period step of the oscillator, requiring an increase in the loop filter gains.
+The detector in case two has a resolution of 3.875 nanoseconds, as 258 MHz is the maximum FPGA clock frequency that can be used with this design, in comparison to the 1.2 nanosecond period step of the oscillator, requiring an increase in the loop filter gains.
 The problems of relatively inaccurate detection and increased gains combine to produce a system that is only able to achieve poor locking performance.

### 2019-03-28 matlab/a1_3.m
another shite day lol. VHDL LF start

@@ -1,6 +1,6 @@
-for inc = 1:1
-    %filename = char("new/des2a_" + num2str(inc) + ".bin");
-    filename = char("scope_41.bin");
+for inc = 1:10
+    filename = char("new/uhh" + num2str(inc) + ".bin");
+    %filename = char("scope_41.bin");
     [~,scope_y_data1] = importAgilentBin(filename,1);
     [~,scope_y_data2] = importAgilentBin(filename,2);
     [~,scope_y_data3] = importAgilentBin(filename,3);
@@ -31,34 +31,34 @@ for inc = 1:1
     std_period_3(inc) = std(periods3);
     std_period_4(inc) = std(periods4);
 
-%     max_2 = max(periods2-mean(periods2));
-%     min_2 = min(periods2-mean(periods2));
-%     p2p_2 = max_2-min_2;
-%     max_3 = max(periods3-mean(periods3));
-%     min_3 = min(periods3-mean(periods3));
-%     p2p_3 = max_3-min_3;
-%     max_4 = max(periods4-mean(periods4));
-%     min_4 = min(periods4-mean(periods4));
-%     p2p_4 = max_4-min_4;
-% 
-%     if length(init_cross2) == length(init_cross1)
-%         TIE_2(inc) = max(abs(init_cross2(1:len)-init_cross1));
-%     elseif length(init_cross2) > length(init_cross1)
-%         TIE_2(inc) = max(abs(init_cross2(2:end)-init_cross1));
-%         inc
-%     end
-%     if length(init_cross3) == length(init_cross1)
-%         TIE_3(inc) = max(abs(init_cross3(1:len)-init_cross1));
-%     elseif length(init_cross3) > length(init_cross1)
-%         TIE_3(inc) = max(abs(init_cross3(2:end)-init_cross1));
-%         inc
-%         end
-%     if length(init_cross4) == length(init_cross1)
-%         TIE_4(inc) = max(abs(init_cross4(1:len)-init_cross1));
-%     elseif length(init_cross4) > length(init_cross1)
-%         TIE_4(inc) = max(abs(init_cross4(2:end)-init_cross1));
-%         inc
-%     end
+    max_2 = max(periods2-mean(periods2));
[…]

### 2019-03-28 conf_paper/paper.tex

@@
 \subsubsection*{Design 1}
-If frequency linearity is a desired characteristic of the oscillator, an accumulator based, FPGA clock driven design exists for this use case also. 
+If frequency linearity is a desired characteristic of the oscillator, an accumulator based, FPGA clock driven design is possible. 
 Here the control code $k$ is added to the value stored in the accumulator at each rising edge of the FPGA clock until overflow is reached at $2^n-1$, where $n$ is the bit width of the counter, thus valuing each control code increment at $\frac{\textrm{FPGA clock}}{2^{\textrm{width}}}$ Hz.
-Once again the MSB of the accumulator is used as the output signal, however, with this design, the MSB will spend an equivalent duration at both logic high and low.
-Again this design is better suited to lower frequencies, where the output of the DCO is orders of magnitude lower than the clock signal driving it.
+The accumulator's Most Significant Bit (MSB) is used as the output of this block and forms a square wave.
+This design is better suited to lower frequencies, where the output of the DCO is orders of magnitude lower than the clock signal driving it, as this ensures that the incremental change due to the control code remains a small fraction of the period.
 
 \subsubsection*{Design 2}
-The third and final potential DCO design is an inverter ring/chain which makes use of the delay through inverter primitives, thus bearing significant resemblance to DCOs used on ASICs.
+The second potential DCO design is an inverter ring/chain which makes use of the delay through inverter primitives, thus bearing significant resemblance to DCOs used on ASICs.
 On an FPGA, the lack of precise control over the layout of blocks, means the exact value of the delay is uncertain and may vary noticeably from one implementation to another, however, this misalignment and the asynchronous behaviour stemming from the lack of an FPGA driven clock, increase the similarity between this design and an ASIC.
@@
 This design works best when the desired period is within a couple orders of magnitude of the delay through each pair of inverters, beyond this point, the number of look-up tables and routing resources required to implement the design becomes unsustainable.\\
-
 A third potential design exists, similar to Oscillator 1 in which the control code varies the reload value of the accumulator at overflow, however, this design will not be discussed in this paper.
@@
 For each test there were four ADPLLs on the FPGA arranged in a Cartesian Grid and as such each ADPLL will be referred to using matrix indexing corresponding to their grid location.
-For each configuration three different modes of operation were examined: Each ADPLL locking to the same 5 MHz external reference, each ADPLL connected to the left and above neighbours in a ``uni-directional'' network and finally a true ADPLL network where all neighbours are connected.
+For each configuration three different modes of operation were examined: Each ADPLL locked to the same 5 MHz external reference, each ADPLL connected to the left and above neighbours in a ``uni-directional'' network and finally a true ADPLL network where all neighbours are connected.
 For comparison, each configuration uses loop filter gains which had previously been determined to provide the best jitter performance.
 The data was captured using an Agilent MSO7054-A oscilloscope at 4 GSa/s.
-Each point of data is the mean of five 200 microsecond long captures which were then analysed in Matlab, Table \ref{table:b_c_perf} displays the results of this analysis.\\
+Each point of data is the mean of five 200 microsecond long captures which were then analysed in Matlab, Table \ref{table:b_c_perf} displays the results of this analysis.
 The first columns of the table relate to the cycle-to-cycle jitter of the system, the difference between sequential rising edges, and the second portion gives the Time Interval Error, the difference between the actual timing of the edge and the ``ideal'' time as set by the reference.
@@
 \end{table}
-It is apparently that the combination of the inverter chain DCO and the FPGA clocked phase detector is an ill advised choice, with the system's jitter being an order of magnitude worse than the design using inverter delay based detector.
-However when the resolution of the detector is then considered, a potential cause for this issue is found.
-The detector in case two has a resolution of 3.875 nanoseconds, as 258 MHz is the maximum FPGA clock frequency that can be used with this design, in comparison to the 1.2 nanosecond period step of the oscillator, requiring an increase in the loop filter gains.
+It is apparent that the combination of the inverter chain DCO and the FPGA clocked phase detector is an ill advised choice, with the system's jitter being an order of magnitude worse than the design using inverter delay based detector.
+In order to obtain measurements for this combination, due to the poor performance, a four times divider was inserted into the feedback path and the comparison made between this divided output and a reference signal at 1.25 MHz, and hence the presence of jitter greater than the period of 200 ns.
+In this system, the FPGA clock is at 258 MHz, the maximum possible to satisfy timing requirements, thus giving a detection resolution of 3.875 nanoseconds.
+This is significantly greater than the 1.2 nanosecond period step of the oscillator, requiring an increase in the loop filter gains when compared to the first combination in the table.
 The problems of relatively inaccurate detection and increased gains combine to produce a system that is only able to achieve poor locking performance.
@@
 %From the table it is apparent that the combination of the inverter chain DCO and the FPGA clock driven phase detector is an ill-advised choice, as the jitter present in the system is several orders of magnitude worse than either other situation.
-In order to obtain these measurements a four times divider was inserted into the feedback path, and hence the presence of jitter greater than the period of 200 ns.
-In this system, the FPGA clock is at 258 MHz, the maximum possible, thus giving a detection resolution of 3.875 nanoseconds.
 %This is 3.5 times greater than the period resolution of approximately 1.2 nano second period resolution of the oscillator, and requires the gains of the PI filter to be increased fourfold in order to respond appropriately to the error detected.

### 2019-03-29 conf_paper/critique.tex
fixed the lad

@@
 
-The two most glaring omissions in the paper are the third oscillator design mentioned in the course of the paper yet is not analysed and the rather poor performance of Design ``1a'' (ring oscillator and clocked phase detector) which was unable to acquire lock at 5 MHz and had to be analysed at the reduced frequency of 1.25 MHz (the output of the in-circuit divider).
+The two most glaring omissions in the paper are the third oscillator design mentioned in the course of the paper yet is not analysed and the rather poor performance of Design ``2a'' (ring oscillator and clocked phase detector) which was unable to acquire lock at 5 MHz and had to be analysed at the reduced frequency of 1.25 MHz (the output of the in-circuit divider).
 %TODO

### 2019-03-30 conf_paper/paper.tex
new measurements and paper update

@@
                 \multicolumn{2}{|l|}{Osc 2, PDET A}&-&-&-&-&-&-\T\\
-                \multicolumn{2}{|r|}{Free PLLs} &78.458&74.025&67.779  &146.09&143.70&142.03 \T\\
-                \multicolumn{2}{|r|}{Uni-dir.}  &78.428&72.646&65.665  &146.23&298.49&200.10 \T\\
-                \multicolumn{2}{|r|}{Bi-dir.}   &78.436&73.827&67.648  &235.73&497.60&311.28 \T\\
+                \multicolumn{2}{|r|}{Free PLLs} &0.56728&0& 0&5.3797&0&0 \T\\
+                \multicolumn{2}{|r|}{Uni-dir.}  &0.57365&0& 0&6.0632&0&0 \T\\
+                \multicolumn{2}{|r|}{Bi-dir.}   &0.58760&0& 0&15.466&0&0 \T\\
                 \hline
                 \multicolumn{2}{|l|}{Osc 2, PDET B}&-&-&-&-&-&-\T\\
-                \multicolumn{2}{|r|}{Free PLLs} &7.8090&7.8827&6.6272   &17.336&21.151&24.302 \T\\
-                \multicolumn{2}{|r|}{Uni-dir.}  &7.7866&8.1145&7.3598   &17.649&29.962&37.051 \T\\
-                \multicolumn{2}{|r|}{Bi-dir.}   &7.7575&7.7729&6.6962   &30.019&40.577&45.617 \T\\
+                \multicolumn{2}{|r|}{Free PLLs} &0.71635&0&0  &4.8293&0&0 \T\\
+                \multicolumn{2}{|r|}{Uni-dir.}  &0.76883&0&0  &4.7922&0&0 \T\\
+                \multicolumn{2}{|r|}{Bi-dir.}   &0.79292&0&0  &6.5277&0&0 \T\\
                 \hline
                 \B
+                %% old 2b
+                %&7.8090&7.8827&6.6272   &17.336&21.151&24.302
+                %&7.7866&8.1145&7.3598   &17.649&29.962&37.051
+                %&7.7575&7.7729&6.6962   &30.019&40.577&45.617
+                %% old 2a
+                %&78.458&74.025&67.779  &146.09&143.70&142.03
+                %&78.428&72.646&65.665  &146.23&298.49&200.10
+                %&78.436&73.827&67.648  &235.73&497.60&311.28
+
+                %&5.6728e-10&0&0&5.3797e-09&0&0
+                %&5.7365e-10&0&0&6.0632e-09&0&0
+                %&5.8760e-10&0&0&1.5466e-08&0&0
+
             \end{tabular}
@@
 \end{table}
-It is apparent that the combination of the inverter chain DCO and the FPGA clocked phase detector is an ill advised choice, with the system's jitter being an order of magnitude worse than the design using inverter delay based detector.
+From the table it is apparent that the coarser period and detection steps of Design ``1A''  (3.875 nanoseconds as at greater clock frequencies timing violations would occur) cannot deliver the same jitter performance as the inverter ring driven designs.
+Looking at the two subsequent designs their lesser flexibility is made up for by their more realistic jitter behaviour.
+As would be expected best performance occurs when the network is in PLL mode and as the quality of the reference degrades as the network is connected up performance becomes worse.
+In terms of cycle-to-cycle jitter the best performing design is that of ``2A'', however, its time interval error is significantly worse especially in bilateral network mode.
+This more aggressive degradation as the network is linked up can be attributed to the worse detection resolution of the FPGA clocked phase detector, four times worse than the 1.2 second step of the entirely inverter based Design ``2B''.\\
+The further a PLL is away from the reference feeding ADPLL 11 the worse performance is, peaking furthest from the reference in ADPLL 22.
+%TODO insert comment here when I have more oscillators for 1 lines or so.
+
+The lesser cycle-to-cycle of Design ``2A'' follows from the implementation independent design of the phase detector.
+Unlike the inverter based design where the implementation may result in the propagation delay between taps on the delay line changing, from both tap to tap in the same detector and the same tap in different detectors, the clocked design has identical detection resolutions as each register is driven by the FPGA clock.
+This can be highlighted in the more easily demonstrable case of the clocked and inverter based oscillators.\\
+\begin{figure}[h]%
+    \centering
+    \subfloat[Inverter Based Design.]{{\includegraphics[width=0.5\linewidth]{distrib_ring} }}%
+    \subfloat[FPGA Clocked Design.]{{\includegraphics[width=0.5\linewidth]{distrib_pa} }}%
+    \caption{Example Distribution of Periods.\label{fig:dists}}    
+    \vspace{-0.5cm}
+\end{figure}
+Figure \ref{fig:dists} paints a picture of the unrealistic nature of the FPGA clocked design in comparison to its inverter based counterpart, each oscillator will have identical intrinsic period steps thus eliminating any jitter that maybe be seen in the system due to misalignment.
+This effects the suitability of the FPGA clocked designs for the improvement of simulation models, or the examination of how a novel block may impact the behaviour of the network.
+\section*{Conclusion}
+An FPGA is a suitable environment for the prototyping of ADPLL networks and can fill a number of roles depending on the designer's requirements.
+However, the designer should bear in mind the platform's limitations, especially when attempting to use a frequency that is not orders of magnitude lower than the maximum clock available from the FPGA's clock manager.
+Inverter based and FPGA clocked designs, both have a niche in which they are the preferred option, with the clocked design better suiting replication of an existing design at scaled-down frequencies and the inverter based designs better in the modelling and analysis of ADPLL network behaviour.
+
+\bibliographystyle{ieeetran}
+\bibliography{conf}
+
+\end{document}
+%It is apparent that the combination of the inverter chain DCO and the FPGA clocked phase detector is an ill advised choice, with the system's jitter being an order of magnitude worse than the design using inverter delay based detector.
 In order to obtain measurements for this combination, due to the poor performance, a four times divider was inserted into the feedback path and the comparison made between this divided output and a reference signal at 1.25 MHz, and hence the presence of jitter greater than the period of 200 ns.
@@
 %Both other combinations work well however, delivering a level of jitter that is both low yet allows an indication of how impactful variations in architecture or control parameters are to be seen visually on an oscilloscope.\\
-
-\begin{figure}[h]%
-    \centering
-    \subfloat[Inverter Based Design.]{{\includegraphics[width=0.5\linewidth]{distrib_ring} }}%
-    \subfloat[FPGA Clocked Design.]{{\includegraphics[width=0.5\linewidth]{distrib_pa} }}%
-    \caption{Example Distribution of Periods.\label{fig:dists}}    
-    \vspace{-0.5cm}
-\end{figure}
-%\begin{figure}[h]
-%   \centering
-%   \subfloat[FPGA Clocked Design][width=0.25\textwidth]{distrib_pa}
-%   \caption{Distribution of Periods.}
-%   \label{fig:dist2}
-%\end{figure}
-Figure \ref{fig:dists} paints a picture of the unrealistic nature of the FPGA clocked design in comparison to its inverter based counterpart, each oscillator will have identical intrinsic period steps thus eliminating any jitter that maybe be seen in the system due to misalignment.
-This effects the suitability of the FPGA clocked designs for the improvement of simulation models, or the examination of how a novel block may impact the behaviour of the network.
-\section*{Conclusion}
-An FPGA is a suitable environment for the prototyping of ADPLL networks and can fill a number of roles depending on the designer's requirements.
-However, the designer should bear in mind the platform's limitations, especially when attempting to use a frequency that is not orders of magnitude lower than the maximum clock available from the FPGA's clock manager.
-Inverter based and FPGA clocked designs, both have a niche in which they are the preferred option, with the clocked design better suiting replication of an existing design at scaled-down frequencies and the inverter based designs better in the modelling and analysis of ADPLL network behaviour.
-
-\bibliographystyle{ieeetran}
-\bibliography{conf}
-
-\end{document}

### 2019-03-30 matlab/a1_3.m

@@ -1,10 +1,43 @@
-for inc = 1:1
-    %filename = char("new/des2a_" + num2str(inc) + ".bin");
-    filename = char("scope_41.bin");
-    [~,scope_y_data1] = importAgilentBin(filename,1);
-    [~,scope_y_data2] = importAgilentBin(filename,2);
-    [~,scope_y_data3] = importAgilentBin(filename,3);
-    [scope_x_data,scope_y_data4] = importAgilentBin(filename,4);
+for inc = 1:15
+    %filename = char("conor_test/settings1_" + num2str(inc) + ".bin");
+    %filename = char("scope_41.bin");
+    
+    path3       = "test_conor/settings3_" + num2str(inc) + ".bin";
+    fileID      = fopen(path3);
+    A           = fread(fileID, 'double');
+    fclose(fileID);
+    N        = 2; 
+    step        = 2e6;
+    width_vect  = 4e6;
+    ss          = [];
+    for chan=0:N-1 %start at 0 for chan1
+        deb = chan*width_vect+1;
+        ss = [ss A(deb:deb+step-1, 1) A(deb+step:deb+2*step-1, 1)];
+    end
+
+%     Ni = N-1;
+%     for jkl = 0:Ni
+%          [w,init_cross,final_cross,mid_level] = pulsewidth(ss(:,2+2*jkl),ss(:,1+2*jkl), 'Polarity', 'Positive');
+%          switch jkl
+%              case 0
+%                 PP = init_cross;
+%              otherwise
+%                  if length(PP)>length(init_cross)
+%                     PP = [PP(1:length(init_cross),:) init_cross];
+%                  elseif length(PP)<length(init_cross)
+%                     PP = [PP init_cross(1:length(PP),:)];  
+%                  else
+%                     PP = [PP init_cross]; 
+%                  end
+%          end
+%     end
+    scope_x_data = ss(:,1);
+    scope_y_data1 = ss(:,2);
+    scope_y_data2 = ss(:,4);
+%     [~,scope_y_data1] = importAgilentBin(filename,1);
+%     [~,scope_y_data2] = importAgilentBin(filename,2);
[…]

### 2019-03-30 matlab/test_conor/MATLAB_data_processing_pll_network_Pb.m

@@ -1,0 +1,936 @@
+clear all;
+clc;
+fclose all;
+format long
+
+disp('**********  OPENING FILES   ***********');
+
+file_type = 'binary';       % choice : 'binary', 'csv', 'dat' : Default Value = 'binary'
+       % is there a majority gate measurement in the file? 'on' or 'off' : Default Value = 'off'
+serie_or_append = 'append';  % as the other file, 'serie' or 'append'
+pll_order = [11];           % Default Value = [11]
+% pll_order = [11 12 13 23]; % which pll is opened
+
+for kp=1:1
+    kp
+    for stuff=1:length(pll_order);
+        clear R_EK_2
+        clear R_EK_coef
+        clear R_PB_2
+        clear R_PB_coef 
+        clear ss
+        clear PP
+        clear T_sig
+        stuff
+        if strcmp(file_type,'dat')
+            switch  N
+                case 3
+                    path = './ADPLL_network/measurement/ADPLL_3x1_network/kp=05_ki=1..13_64/';
+                    s1 = importdata(strcat(path, 'sig_div1.dat' ));
+                    s2 = importdata(strcat(path, 'sig_div2.dat'));
+                    s3 = importdata(strcat(path, 'sig_div3.dat'));
+                    s4 = importdata(strcat(path, 'sig_mg.dat'));
+                    s5 = importdata(strcat(path, 'sig_ref.dat'));
+                    ss = [s1 s2 s3 s4 s5];
+                case 5
+                    path = './measure/PLL_x5/serie33/';
+                    s1 = importdata(strcat(path, 'sig_div1.dat' ));
+                    s2 = importdata(strcat(path, 'sig_div2.dat'));
+                    s3 = importdata(strcat(path, 'sig_div3.dat'));
+                    s4 = importdata(strcat(path, 'sig_div4.dat'));
+                    s5 = importdata(strcat(path, 'sig_div5.dat'));
+                    s6 = importdata(strcat(path, 'sig_mg.dat'));
+                    s7 = importdata(strcat(path, 'sig_ref.dat'));
+                    ss = [s1 s2 s3 s4 s5 s6 s7];
+                case 7
+                    path = './measure/vacances7PLL/';
+                    s1 = importdata(strcat(path, 'sig_div1.dat' ));
+                    s2 = importdata(strcat(path, 'sig_div2.dat'));
+                    s3 = importdata(strcat(path, 'sig_div3.dat'));
[…]

### 2019-03-30 matlab/test_conor/fft_Bisiaux.m

@@ -1,0 +1,15 @@
+function [Xfft, Yfft] = fft_Bisiaux(Fs,fft_pts,x_input)
+%FFT_BISIAUX Summary of this function goes here
+%   Detailed explanation goes here
+% Nfft        = pow2(14); %length(v_modul);%   
+% Xfft_moi    = Fs*(0:Nfft/2-1)/Nfft;
+% Yfft        = fft(v_modul, Nfft);
+% Yfft        = Yfft(1:Nfft/2);
+% Yfft        = abs(Yfft)/length(Yfft);
+Nfft        = fft_pts; %length(v_modul);%   
+Xfft        = Fs*(0:Nfft/2-1)/Nfft;
+Yfft        = fft(x_input, Nfft);
+Yfft        = Yfft(1:Nfft/2);
+Yfft        = abs(Yfft)/length(Yfft);
+end
+

### 2019-03-30 matlab/test_conor/hdlcoder_board_customization.m

@@ -1,0 +1,12 @@
+function r = hdlcoder_board_customization
+% Board plugin registration file
+% 1. Any registration file with this name on MATLAB path will be picked up
+% 2. Registration file returns a cell array pointing to the location of
+%    the board plugins
+% 3. Board plugin must be a package folder accessible from MATLAB path,
+%    and contains a board definition file
+r = { ...
+      'ZyboRegistration.plugin_board', ...
+     };
+ boardName = 'ZYBO';
+end

### 2019-03-30 matlab/test_conor/plugin_board.m

@@ -1,0 +1,34 @@
+function hB = plugin_board()
+% Board definition
+% Construct board object
+hB = hdlcoder.Board;
+hB.BoardName    = 'ZYBO';
+% FPGA device information
+hB.FPGAVendor   = 'Xilinx';
+hB.FPGAFamily   = 'Zynq';
+hB.FPGADevice   = 'xc7z010';
+hB.FPGAPackage  = 'clg400';
+hB.FPGASpeed    = '-1';
+% Tool information
+hB.SupportedTool = {'Xilinx Vivado'};
+% FPGA JTAG chain position
+hB.JTAGChainPosition = 2;
+%% Add interfaces
+% Standard "External Port" interface
+hB.addExternalPortInterface( ...
+    'IOPadConstraint', {'IOSTANDARD = LVCMOS33'});
+% Custom board external I/O interface
+hB.addExternalIOInterface( ...
+  'InterfaceID',    'LEDs General Purpose', ...
+  'InterfaceType',  'OUT', ...
+  'PortName',       'LEDs', ...
+  'PortWidth',      4, ...
+  'FPGAPin',        {'M14', 'M15', 'G14', 'D18'}, ...
+  'IOPadConstraint', {'IOSTANDARD = LVCMOS33'});
+hB.addExternalIOInterface( ...
+   'InterfaceID',    'Push Buttons', ...
+   'InterfaceType',  'IN', ...
+   'PortName',       'PushButtons', ...
+   'PortWidth',      4, ...
+   'FPGAPin',        {'R18', 'P16', 'V16', 'Y16'}, ...
+   'IOPadConstraint', {'IOSTANDARD = LVCMOS33'});

### 2019-03-30 matlab/test_conor/plugin_rd.m

@@ -1,0 +1,22 @@
+function hRD = plugin_rd()
+% Reference design definition
+% Construct reference design object
+hRD = hdlcoder.ReferenceDesign('SynthesisTool', 'Xilinx Vivado');
+hRD.ReferenceDesignName = 'Demo system';
+hRD.BoardName = 'ZYBO';
+% Tool information
+hRD.SupportedToolVersion = {'2017.2','2017.4'};
+% add custom Vivado design
+hRD.addCustomVivadoDesign( ...
+    'CustomBlockDesignTcl', 'design_led.tcl', ...
+    'VivadoBoardPart','digilentinc.com:zybo:part0:1.0');
+%% Add interfaces
+% add clock interface
+hRD.addClockInterface( ...
+  'ClockConnection',   'clk_wiz_0/clk_out1', ...
+  'ResetConnection',   'proc_sys_reset_0/peripheral_aresetn');
+% add AXI4 and AXI4-Lite slave interfaces
+hRD.addAXI4SlaveInterface( ...
+    'InterfaceConnection', 'axi_interconnect_0/M00_AXI', ...
+    'BaseAddress',         '0x40010000', ...
+    'MasterAddressSpace',  'processing_system7_0/Data');

### 2019-03-31 tex_files/Journal.tex

@@
 Made a new loop filter this weekend without fractions which may be easier to debug (Pierre's suggestion).
+\section*{2x2 Locking Well, Paper submitted}
+Turns out that the problem with my ring oscillators was that I was driving things in the wrong direction due to double negation.
+Did eventually solve this on Friday AFTER the paper was submitted.
+Performance is great now, 3x3 should work well (fingers crossed) on Monday.
+Unfortunately I wanted to submit to a conference on Monday but a combination of issues likely means that the project will not compile the references so when I collect the 30 missing datapoints and go home to do that it will likely be too late.
 

### 2019-04-01 matlab/a1_3.m
as

@@ -2,11 +2,11 @@ for inc = 1:15
     %filename = char("conor_test/settings1_" + num2str(inc) + ".bin");
     %filename = char("scope_41.bin");
     
-    path3       = "test_conor/settings3_" + num2str(inc) + ".bin";
+    path3       = "test_conor/monday/design2b_" + num2str(inc) + ".bin";
     fileID      = fopen(path3);
     A           = fread(fileID, 'double');
     fclose(fileID);
-    N        = 2; 
+    N           = 4; 
     step        = 2e6;
     width_vect  = 4e6;
     ss          = [];
@@ -34,6 +34,8 @@ for inc = 1:15
     scope_x_data = ss(:,1);
     scope_y_data1 = ss(:,2);
     scope_y_data2 = ss(:,4);
+    scope_y_data3 = ss(:,6);
+    scope_y_data4 = ss(:,8);
 %     [~,scope_y_data1] = importAgilentBin(filename,1);
 %     [~,scope_y_data2] = importAgilentBin(filename,2);
 %     [~,scope_y_data3] = importAgilentBin(filename,3);
@@ -45,25 +47,24 @@ for inc = 1:15
 
     [~,init_cross1,~,~] = pulsewidth(scope_y_data1,scope_x_data, 'Polarity', 'Positive');
     [~,init_cross2,~,~] = pulsewidth(scope_y_data2,scope_x_data, 'Polarity', 'Positive');
-%     [~,init_cross3,~,~] = pulsewidth(scope_y_data3,scope_x_data, 'Polarity', 'Positive');
-%     [~,init_cross4,~,~] = pulsewidth(scope_y_data4,scope_x_data, 'Polarity', 'Positive');
+    [~,init_cross3,~,~] = pulsewidth(scope_y_data3,scope_x_data, 'Polarity', 'Positive');
+    [~,init_cross4,~,~] = pulsewidth(scope_y_data4,scope_x_data, 'Polarity', 'Positive');
 
     len = length(init_cross1);
     periods(1,:) = getPeriods(init_cross1); 
     periods(2,:) = getPeriods(init_cross2);
-%     periods(3,:) = getPeriods(init_cross3);
-%     periods(4,:) = getPeriods(init_cross4);
+    periods(3,:) = getPeriods(init_cross3);
+    periods(4,:) = getPeriods(init_cross4);
 
[…]

### 2019-04-05 thesis/Abstract.tex
abstracts

\documentclass[11pt,english,british]{report}
\usepackage[a4paper]{geometry}
\geometry{verbose,tmargin=3cm,bmargin=3cm,lmargin=3.5cm,rmargin=2.4cm}
\begin{document}
Low power, high frequency clock distribution systems will be in ever increasing demand in the near future as the need for high performance digital circuitry grows. At these frequencies, however, the conventional clock distribution systems are unable provide a clock signal of adequate quality without compromising on either of these problems. Many devices have turned away from using Globally Synchronous clock distribution systems in favour of those that divide the area of the chip into Globally Asynchronous but Locally Synchronous areas. However, new technologies seek to enable the use of Globally Synchronous methods at high frequencies, one of which being the use of oscillators coupled in phase, each responsible for delivering the clock to a subregion of the chip.\\
Each oscillator forms part of a Phase Lock Loop (PLL), and in order to enable the synchronisation of the network each PLL is linked those controlling the adjacent clock regions. For a digital system it is expedient to implement these PLLs digitally as an All-Digital Phase Lock Loop (ADPLL) as this provides a number of advantages. It has been shown in theory and experimentally that this method can produce the high quality clock desired with low power consumption.\\
As the production of test chips is expensive and time consuming through simulation and validation of a design is vital, traditionally carried out for mixed signal circuitry using complex behavioural and theoretical models. For a ADPLL Network the consistency of a signal both with respect to itself, and to the other clock signals on the chip is a key performance attribute and much of the variation is due to random processes that may be difficult to simulate effectively.\\
This thesis will demonstrate that the Field Programmable Gate Array (FPGA) can be used in order to simulate, model or validate ADPLL network architectures in a cost and time effective manner, as a complement to conventional methods. The key benefit is that many system dynamics that will be seen when a network is implemented on an Application Specific Integrated Circuit (ASIC), but be overlooked in software simulations can be examined in the hardware simulation that an FPGA provides.\\
This thesis implements networks using three designs of ADPLL, each using a different architecture, and highlights the use cases to which each is best suited. The performance of each design is then analysed and this compared to the suggested use cases. Additionally the impact of more minor architectural modifications is tested and documented.
\end{document}

### 2019-04-05 thesis/LayAbstract.tex
abstracts

\documentclass[11pt,english,british]{report}
\usepackage[a4paper]{geometry}
\geometry{verbose,tmargin=3cm,bmargin=3cm,lmargin=3.5cm,rmargin=2.4cm}
\begin{document}
With the increasing proliferation of ``smart'' devices the need for low power yet high speed devices has never been greater. Each smart device contains a processor to control the device, each containing complex circuitry, requiring extremely exact synchronisation. The task of this synchronisation falls to the ``clock'' signal which must occur at the same instant in all areas of the processor. Conventional solutions to this problem cannot satisfy both power and speed requirements simultaneously, so designers have proposed the ADPLL Network which approaches the problem from the other side. Rather than generate the clock once and send it around the chip, which consumes a large amount of power, an ADPLL network divides the chip into a grid and generates many signals that each serve a local area, only requiring non time critical control signals to be sent over large distances.\\
The production of chips to test designs is both expensive and time consuming so designers must ensure that mistakes have not been made, accomplished by simulating the behaviour of the designs using complex behavioural and theoretical models. This thesis will discuss the use of Field Programmable Gate Arrays (FPGA) as a hardware testing, simulation and validation platform, to be used prior to test chip production, that is both cost and time effective. The hardware nature of an FPGA enables the analysis of behaviours that would not be possible in software, without the cost and time penalties of a custom chip. This is made possible by the ability to reconfigure the chip at will, albeit with comparatively lesser capabilities.\\
This thesis implements networks using three designs of ADPLL, each using a different architecture, and highlights the use cases to which each is best suited. The performance of each design is then analysed and this compared to the suggested use cases. Additionally the impact of more minor architectural modifications is tested and documented.
%what do I do?
\end{document}

### 2019-04-06 conf_paper/paper.tex
ieeetran fixed
as
initial thesis background

@@
 
-\author{\IEEEauthorblockN{C. Dooley, E. Blokhina, B. Mulkeen}
-    \IEEEauthorblockA{\textit{School of Electrical \& Electronic Engineering, University College Dublin}
-    	\vspace{-1.2cm}
+\author{\IEEEauthorblockN{C. Dooley$^1$, E. Blokhina$^1$, B. Mulkeen$^1$, D. Galayko$^2$}
+    \IEEEauthorblockA{\textit{$^1$School of Electrical \& Electronic Engineering, University College Dublin}\\
+    \textit{$^2$LIP6 Lab, Sorbonne Université}
+    	\vspace{-0.8cm}
 	}
@@
                 \multicolumn{2}{|l|}{Osc 2, PDET A}&-&-&-&-&-&-\T\\
-                \multicolumn{2}{|r|}{Free PLLs} &0.56728&0& 0&5.3797&0&0 \T\\
-                \multicolumn{2}{|r|}{Uni-dir.}  &0.57365&0& 0&6.0632&0&0 \T\\
-                \multicolumn{2}{|r|}{Bi-dir.}   &0.58760&0& 0&15.466&0&0 \T\\
+                \multicolumn{2}{|r|}{Free PLLs} &0.56222&0.49224&0.64124 &5.3039&7.8626&5.4634 \T\\
+                \multicolumn{2}{|r|}{Uni-dir.}  &0.55040&0.51284&0.62796 &5.5312&7.5359&7.9697 \T\\
+                \multicolumn{2}{|r|}{Bi-dir.}   &0.57056&0.50403&0.61644 &11.515&15.943&17.428 \T\\
                 \hline
                 \multicolumn{2}{|l|}{Osc 2, PDET B}&-&-&-&-&-&-\T\\
-                \multicolumn{2}{|r|}{Free PLLs} &0.71635&0&0  &4.8293&0&0 \T\\
-                \multicolumn{2}{|r|}{Uni-dir.}  &0.76883&0&0  &4.7922&0&0 \T\\
-                \multicolumn{2}{|r|}{Bi-dir.}   &0.79292&0&0  &6.5277&0&0 \T\\
+                \multicolumn{2}{|r|}{Free PLLs} &0.66757&0.84477&0.71308 &5.9142&6.1353&5.3939 \T\\
+                \multicolumn{2}{|r|}{Uni-dir.}  &0.66934&0.83894&0.71271 &5.8779&7.1746&6.8914 \T\\
+                \multicolumn{2}{|r|}{Bi-dir.}   &0.68372&0.83193&0.68570 &7.5705&9.6292&9.7715 \T\\
                 \hline
@@
 In terms of cycle-to-cycle jitter the best performing design is that of ``2A'', however, its time interval error is significantly worse especially in bilateral network mode.
-This more aggressive degradation as the network is linked up can be attributed to the worse detection resolution of the FPGA clocked phase detector, four times worse than the 1.2 second step of the entirely inverter based Design ``2B''.\\
+This more aggressive degradation as the network is linked up can be attributed to the worse detection resolution of the FPGA clocked phase detector, four times worse than the 1.2 nanosecond step of the entirely inverter based Design ``2B''.\\
 The further a PLL is away from the reference feeding ADPLL 11 the worse performance is, peaking furthest from the reference in ADPLL 22.
-%TODO insert comment here when I have more oscillators for 1 lines or so.
-
 The lesser cycle-to-cycle of Design ``2A'' follows from the implementation independent design of the phase detector.
@@
 
-\bibliographystyle{ieeetran}
+\bibliographystyle{IEEEtran}
 \bibliography{conf}

### 2019-04-06 thesis/Abstract.tex
initial thesis background

@@
 As the production of test chips is expensive and time consuming through simulation and validation of a design is vital, traditionally carried out for mixed signal circuitry using complex behavioural and theoretical models.
-For a ADPLL Network the consistency of a signal both with respect to itself, and to the other clock signals on the chip is a key performance attribute and much of the variation is due to random processes that may be difficult to simulate effectively.\\
+For an ADPLL Network the consistency of a signal both with respect to itself, and to the other clock signals on the chip is a key performance attribute and much of the variation is due to random processes that may be difficult to simulate effectively.\\
 This thesis will demonstrate that the Field Programmable Gate Array (FPGA) can be used in order to simulate, model or validate ADPLL network architectures in a cost and time effective manner, as a complement to conventional methods.

### 2019-04-06 thesis/Background.tex
initial thesis background

--- 2019-01-22 tex_files/placeholder.tex
@@
-\documentclass[11pt,english,british]{report}
-\usepackage[a4paper]{geometry}
-%\usepackage[T1]{fontenc}
-\usepackage[latin9]{inputenc}
-\usepackage{setspace}
-\geometry{verbose,tmargin=3cm,bmargin=3cm,lmargin=3.5cm,rmargin=2.4cm}
-%\usepackage{fancyhdr}
-%\pagestyle{fancy}
-%\setcounter{secnumdepth}{3}
-\usepackage[british]{babel}
+\section{Brief Overview}
+In a world where the demand for high performance hand-held computing devices continues to grow and the prevalence of ``smart'' devices is increasing, there is unprecedented demand for \ac{SOC} devices to control systems as varied as medical devices and entertainment systems.
+As these applications become more and more demanding, with ever increasing amounts of data to process and the expectation that today's devices will outperform those of yesterday, the problem of maintaining the steady gain in performance of \acp{SOC} remains at the fore.
 
-\usepackage{amsmath}
-\usepackage{amssymb}
-\usepackage{cancel}
-%\usepackage{gensymb}
-\usepackage{graphicx}
-\usepackage{esint}
-\usepackage{mdsymbol}
-\usepackage{esvect} 
-%\usepackage{lipsum}
-\usepackage{hhline}
-\usepackage{eurosym}
-\usepackage{listings}
-\usepackage{booktabs}
-\usepackage{amssymb}
-\usepackage{mathrsfs}
-\usepackage{commath}
-\usepackage{adjustbox}
-\usepackage{booktabs}
-\usepackage{array}
-\usepackage{lscape}
-\usepackage[xspace]{ellipsis}
-\usepackage{color}
-\usepackage{float}
-\usepackage{caption}
-\usepackage{afterpage}
-
-			\usepackage{titlesec}
-
-			\titleformat{\chapter}[display]
-			{\normalfont\huge\bfseries}{\chaptertitlename\ \thechapter}{20pt}{\Huge}
-
-			% this alters "before" spacing (the second length argument) to 0
-			\titlespacing*{\chapter}{0pt}{0pt}{32pt}
-
-			\usepackage{etoolbox}
-			\makeatletter
-			\patchcmd{\chapter}{\if@openright\cleardoublepage\else\clearpage\fi}{}{}{}
-			\makeatother
-
-			\usepackage[unicode=true,pdfusetitle,
-			bookmarks=true,bookmarksnumbered=false,bookmarksopen=false,
-			breaklinks=true,pdfborder={0 0 0},backref=page,colorlinks=false]
-			{hyperref}
-
-\definecolor{mygreen}{rgb}{0,0.6,0}
-\definecolor{mygray}{rgb}{0.5,0.5,0.5}
-\definecolor{mymauve}{rgb}{0.58,0,0.82}
-
-\begin{document}
-
-\title{Placeholder Tex Doc}
-\author{Conor Dooley}
-\begin{titlepage}
-
-%	\pagenumbering{gobble}
-
-	\setstretch{1.25}
-
-	\begin{spacing}{3}
-
-		\noindent \begin{center}
-
-			{\huge{}ADPLL Network on an FPGA}
-			\par\end{center}{\huge \par}
-	\end{spacing}
-
-	\bigskip{}
-
-	\begin{center}
-		{\Large{}Conor Dooley}
-	\end{center}
-
-	\vspace{1.25cm}
-
-	\begin{center}
-		%\includegraphics[width=0.15\paperwidth]{ucd_crest.pdf}\vspace{0.75cm}
-		\includegraphics[width=0.2\paperwidth]{UCD_crest.pdf}\vspace{0.25cm}
-	\end{center}
-
-	\bigskip{}
-
-	\begin{center}
-		{\huge{}Masters Project Interim Report }
-	\end{center}
-
-	\vspace{1.cm}
-
-	{\Large{}\bigskip{}
-		\bigskip{}
-	}{\Large \par}
-
-	\noindent \begin{center}
-		\begin{tabular}{lll}
-			\textbf{\large{}Research Supervisors:} & \qquad{}\qquad{} & {\large{}Dr Elena Blokhina \& Brian Mulkeen}\\
-		\end{tabular}
-
-		\par\end{center}
-
-	\vspace{1cm}
-	\noindent \begin{center}
-		{\large{}January  2019}
-		\par\end{center}{\large \par}
-
-\end{titlepage}
-\setcounter{page}{1} 
-\pagenumbering{roman}
-\setstretch{1.05}
-
-\addcontentsline{toc}{chapter}{\contentsname}
-
-\tableofcontents{}
-
-\pagebreak{}
-
-%\begin{dedication}
-%Dedicated to google and wikipedia  
-%\end{dedication}
-
-%\setcounter{page}{1} 
-%\pagenumbering{roman}
-
-%\addtocontents{toc}{\protect{\pdfbookmark[0]{Abstract}{toc}}}
-%\pagestyle{empty}
-%\addcontentsline{toc}{chapter}{Abstract}
-%\phantomsection
-
-\pagestyle{plain}
-
-\setstretch{1.3}
-\doublespacing
-\pagenumbering{arabic}
-\setcounter{page}{1} 
-\chapter{Background Review}
-\section{Brief Overview}
-In a world where the demand for high performance hand-held computing devices continues to grow and the prevalence of ``smart'' devices is increasing, the problem of maintaining the steady gain in performance of Systems-On-Chip remains at the fore.
-%TODO cite?
-The main drivers of performance in Systems-On-Chip are the number of transistors on a chip and the frequency at which it is clocked.
-As the number of transistors on a chip has increased roughly following Moore's Law, the increase in clock frequency has not been able to follow a similar linear trajectory, having remained roughly equivalent for the last number of years \cite{ross2008cpu}.
-
+The main drivers of performance in \acp{SOC} are the number of transistors on a chip, which is correlated with the number of calculations that can be carried out simultaneously, and the frequency at which the device operates, which determines the number of calculations performed per second.
+Moore's Law, based on the famous observation by Gordon Moore in 1965\cite{moore1965cramming}, predicted a doubling in the transistor count of \ac{IC}s per year for the forthcoming decade.
+This behaviour has carried on to this day as a result of the ever decreasing size of transistors, and has only begun to slow down in recent years.
+However, as the number of transistors on a chip has increased roughly following Moore's Law, the increase in clock frequency has not been able to follow a similar linear trajectory, having remained roughly equivalent for the last number of years \cite{ross2008cpu}, indeed the clock speed of the Intel Core family of \ac{CPU}s has not changed since their introduction in 2009 \cite{intelark}.
+\begin{figure}[h]
+	\centering
+	\includegraphics[scale=0.4]{../eldar_last_30_yrs}
+	\caption{Frequency of the Intel microprocessors over past 30 years \cite{zianbetov2013phd}.}
+	\label{fig:eldar_last_30_yrs}
+\end{figure}\\
 This plateauing of clock frequency has been caused by high power consumption due to the demands placed by the global distribution of a high frequency clock, often the single biggest consumer of power on the chip \cite{tiwari1998reducing}.
-In a world where low power devices are desirable, especially with the growing demand of the Internet-Of-Things where high power consumption goes directly against one of the key pillars of the technology.
-%TODO Pillars of what?
+With the growth of the \ac{IOT} market where low power devices are desirable, with many of the emerging uses of \ac{SOC}s being portable and thus without a permanent power source, high power consumption goes directly against one of the key pillars of the technology.
+This forces many of these devices to use lower performance hardware in order to reduce the power consumption, and increase the battery life, of their devices.
 
@@
 In both cases, the chip is broken down into small areas in which all transistors are clocked synchronously, with the size constrained by the ability to deliver a quality clock signal to all transistors.
-The first of these methods is Globally Synchronous Locally Synchronous, or GSLS, where all of these sub areas on the chip are synchronised with each other.
-In practice, however, this is very difficult to achieve, as extremely high precision is required across the entire area of the chip, and doing so leads to high power consumption.
-In contrast in a Globally Asynchronous Locally Synchronous clock delivery system the ``local'' areas are not synchronised with other.
+The first of these methods is \ac{GSLS}, where the clock signals in each of these subregions of the chip are synchronised with one other.
+In practice, however, this is very difficult to achieve, as extremely high precision is required across the ever increasing number of transistors and the entire area of the chip, and doing so leads to high power consumption.
+
+In contrast in a \ac{GALS} clock delivery system the ``local'' areas are not synchronised with other.
 This reduces the clocking system's complexity and thus the power consumption and chip area used, at the expense of communication speed between blocks.
-%TODO citation?
-A GSLS, system, however has the advantages of deterministic behaviour and greater rates of communication between clocking areas and, as such, remains a desirable system design.
-A number of methods which deliver GSLS clocking exist at present such as clock trees as well as emerging technologies such as ADPLL networks.
+This disadvantage comes from the need to then somehow synchronise the messages being sent from one area to another to avoid the corruption of any messages.
+A \ac{GSLS}, system, however has the advantages of deterministic behaviour and greater rates of communication between clocking areas and, as such, remains a desirable system design.
+A number of methods which deliver \ac{GSLS} clocking exist at present such as clock trees as well as emerging technologies such as ADPLL networks.
 
@@
 	\centering
-	\includegraphics[scale=0.5]{../eldar_why_precise_clocking}
+	\includegraphics[scale=0.6]{../eldar_why_precise_clocking}
 	\caption{Data Flow in a Clocked System \cite{zianbetov2013phd}.}
 	\label{fig:eldar_why_precise_clocking}
-\end{figure}\\
+\end{figure}
 
@@
 If $\Delta T$ is negative this reduces the time available for the intervening combinatorial network thereby, having the same effect as a reduction in clocking frequency.
-The most common sources of clock error are caused by mismatches which usually stem from production, such as differences in the length of clocking paths, buffer delays or in the parameters of either active or passive components in the clock distribution network.
-These all manifest themselves as skew, while the noise in active components will appear as jitter in the clock signal.
+Correspondingly a positive $\Delta T$ for depicted registers implies a negative $\Delta T$ for $R_f$ and the subsequent register.
+The most common sources of clock error are caused by mismatches which usually stem from production, such as differences in the length of clocking paths, buffer delays or in the parameters of either active or passive components in the clock distribution network, which as the size of components on an \ac{IC} reduces becomes more difficult to avoid.
+All sources of mismatch will manifest themselves in the clock distribution system as skew between transistors, while the noise in active components or the power supply system will appear as jitter in the clock signal.
 
 \section{Traditional Solutions}
-A number of traditional solutions exist which provide GSLS clocking systems, using a variety of techniques.
-The most simple of these use the clock distribution network's symmetry in order to deliver the clock in phase, at all locations in the chip and are named according to their geometry with the main variants being branch-trees, X-trees or H-trees.
+A number of traditional solutions exist which provide \ac{GSLS} clocking systems, using a variety of techniques.
+The most simple of these implement clock distribution systems that are symmetrical in order to distribute a centrally generated clock signal to all areas of the chip at the same phase.
+These systems are named in accordance with their geometry, with the most common variants being branch, X or H trees.
 \begin{figure}[h]
 	\centering
-	\includegraphics[scale=0.2]{../trees}
+	\includegraphics[scale=0.33]{../trees}
 	\caption{H and Branch Tree Clock Distribution Systems.}
@@
 While on the surface these appear simple, the task of obtaining an exact matching is, in practice, the limiting factor in this design.
-Even if the clock distribution system is geometrically symmetrical, production mismatches in either active or passive components will lead to a skew that varies from part to part.
+Even if the clock distribution system is geometrically symmetrical by design, production mismatches in either active or passive components will lead to a skew that varies from part to part.
 In order to minimise the impact of production tolerances, the dimensions of components in the distribution network can be increased, thus reducing the relative variation possible.
 However this has the impact of increasing the power consumption of the distribution network \cite{tiwari1998reducing}.
-Designs do exist in which the electrical lines used in the tree network are replaced by waveguides for optical signals, which negates the electrical losses, however, this technique suffers from being a newer technology and as such is expensive to realise \cite{haurylau2006chip}.
 
-A mesh clock distribution network is an alternate design where the clock is delivered using a Cartesian grid of distribution lines and variation in skew is inversely proportional to the density of the grid.
+A mesh clock distribution network is an alternate design where the clock is delivered using a Cartesian grid of distribution lines.
+Compared to a tree type system, the variation in skew seen with a clock mesh is inversely proportional to the density of the grid while the sources of jitter remain identical.
 According to Abdelhadi \textit{et al} (2010) clock meshes ``\textit{achieve low and deterministic skew, low skew variations, and low jitter}'', all desirable characteristics for a clock distribution system.
-However they dissipate more power due to extra capacitive loading attributable to vast number of lines required to form the grid Similarly tree distribution networks suffer from potential mismatch in production and alleviation through increasing of distribution system dimensions will again lead to higher power consumption \cite{abdelhadi2010timing}.
+However they dissipate more power due to extra capacitive loading, attributable to vast number of lines required to form the grid.
+Similarly mesh distribution networks suffer from potential mismatch in production and alleviation through increasing of the dimensions of interconnects will, as with a tree type system, lead to higher power consumption \cite{abdelhadi2010timing}.
 \begin{figure}[h]
 	\centering
-	\includegraphics[scale=0.6]{eldar_mesh}
+	\includegraphics[scale=0.7]{../tex_files/eldar_mesh}
 	\caption{Mesh Clock Distribution System \cite{zianbetov2013phd}.}
@@
 \end{figure}
+Alternative designs replace the electrical lines used in the tree networks with waveguides for optical signals, with only the distribution in the local area carried out using regular wires.
+This technique presents many advantages \cite{chen2006chip}: optical clock delivery is immune to the noise sources that affect electrical clock distribution systems, consume less power and do not suffer from the electrical losses present in a regular tree system.
 
-In a tree type distribution system, skew due to the manufacturing process can be partly accounted for by means of a skew compensation system, of which there are two main types.
-These types are defined by the location of the control mechanism with those having one controller located at the clock source, known as ``centralised'' methods, and those with multiple controllers in the individual clocking areas known as ``decentralised''.
+\section{Skew Compensation}
+In a tree type distribution system, skew is the main issue affecting clock accuracy and as such some effort has gone into addressing the problem.
+Skew due to the manufacturing process can be, at least, partly accounted for by means of active control through a skew compensator.
+This is a circuit, or controller, that compares the skew of each local clocking area on the chip and attempts to ensure in-phase clock delivery.
+Two main strategies exist to provide skew compensation, each named according to the location of the control mechanism.
+Designs featuring the controller located at the clock source, are known as ``centralised'' methods, and those with multiple controllers in the individual clocking areas known as ``decentralised''.
+Regardless of the controller placement these techniques allow for the tuning of the propagation delay between the centralised clock source and the local clocking areas.
+
 In a centralised skew compensation circuit, the skew across the chip is calculated by the central controller which then manipulates the distribution network in order to deliver a more in-phase clock around the chip.
-This, however, adds complication to a circuit which already is a large consumer of power and as such increases the power usage.
+This calculation is done by measuring the round trip time from the clock source to both the root of local clock tree, and to the individual ``leaves'' of the tree.
+The controller then has a limited ability to tune the propagation path.
+The downsides of this technique are the resolution of both the measurement and compensation are poor, allowing for the correction of just skew and not of any jitter that may be present in the system, and that the extra circuitry required for both the tunability of the forward path and the two extra return paths contribute to an increased footprint and power consumption.
 
-As the name suggest a decentralised skew compensation technique delegates this responsibility to the individual clock regions.
+As the name suggest a decentralised skew compensation technique delegates the responsibility of tuning the propagation path to the individual clock regions.
+This strategy has the advantage of not requiring the return paths present in a ``centralised'' design.
+Instead comparison is made betweeen the leaves of different clocking areas and on this basis the propagation delay is varied.
 For example, Yamashita \textit{et al} (2005) designed a system in which each clocking area or ``leaf node'' contains a partial clock tree.
 Each of these ``leaves'' is able to compare its clock phase to the neighbouring node, and based on the result, tune an adjustable delay buffer \cite{yamashita2005dynamic}.
-While this method can compensate for process, voltage and temperature variation, it does not address the power consumption due to the delivery of a high frequency clock across the entire chip area.
+While this method can compensate for process, voltage and temperature variation, it does not address the power consumption due to the delivery of a high frequency clock across the entire chip area nor does it have any impact on clock jitter.
 
+%DOWN TO HERE DONE FOR ACRO
 \section{Multi-oscillator Designs}
@@
 	\centering
-	\includegraphics[scale=0.5]{../mizuno1998noise}
+	\includegraphics[scale=0.7]{../mizuno1998noise}
 	\caption{Coupled Oscillator Clock Delivery Circuit \cite{mizuno1998noise}.}
@@
 \begin{quotation}
-	\singlespacing
 	\textit{Unfortunately, as with the conventional ...
 method, distributing the VCOs over the entire chip causes the problem that jitter and skew are increased by variations in the fabrication process (static), temperature, and power supply (dynamic)} \cite{mizuno1998noise}.
-	\doublespacing
 \end{quotation}
@@
 \begin{quote}
-	\singlespacing
 	\textit{Design and measurements on this chip confirm that generating and synchronizing multiple clocks on chip is feasible.
 Neither the power nor the area overhead of multiple PLLs is substantial compared to the cost of distributing the clock by conventional means} \cite{gutnik2000active}.
-	\doublespacing
 \end{quote}
@@
 	\centering
-	\includegraphics[scale=0.5]{../gutnik2000active}
+	\includegraphics[scale=0.7]{../gutnik2000active}
 	\caption{PLL Network Topology \cite{gutnik2000active}.}
@@
 	\centering
-	\includegraphics[scale=0.35]{../ccirc_2013_arch}
+	\includegraphics[scale=0.4]{../ccirc_2013_arch}
 	\caption{Architecture of the ADPLL network and of a single node \cite{zianbetov2013distributed}.}
@@
 	\centering
-	\includegraphics[scale=0.5]{mulkeen_pll}
+	\includegraphics[scale=0.5]{../tex_files/mulkeen_pll}
 	\caption{Block Diagram of a Phase Lock Loop, \textit{Wireless Systems Notes}, B. Mulkeen (2017).}
@@
 \end{figure}
-
-\subsection{Digitally Controllable Oscillator}
-In a digital system there are a very limited number of voltages representable, and frequently there are only two levels, so using this to control the oscillator is not a viable strategy.
-Instead a fixed bit width signal is used to control the oscillator's period, selecting the number of inverters in a ring oscillator or the varactor configuration of a travelling wave oscillator \cite{chen2011rotary}.
-The decisions made in the design of the digitally controllable oscillator, or DCO, determine many of the other ADPLL parameters.
-While tuning range and centre frequency as well as linearity carry over from the analog counterpart, a DCO also has a frequency step which in combination with the bit width of the control signal determines the range over which the oscillator can be tuned.
-Figure \ref{fig:my_ring} illustrates a basic ring oscillator design.
-A ring oscillator is an inherently unstable circuit composed of an odd number of inverters where the period of oscillation is given by twice entire chain's delay.
-The frequency of operation can then be set by modulating the length of the chain by means of \texttt{f\_select}.
-The main impact of output frequency quantisation is that only frequencies which are integer multiples of the frequency step away from the centre frequency can be easily reproduced, with intermediate values only obtainable in a manner akin to Fractional-N synthesis with the control code toggling back and forth.
-This acts as a source of jitter in the system.
-\begin{figure}[h]
-	\centering
-	\includegraphics[scale=0.275]{../inverter_chain}
-	\caption{Basic Ring/Inverter Chain Oscillator.}
-	\label{fig:my_ring}
-\end{figure}
-
-\subsection{Digital Phase Detector}
-Once again quantisation impacts the Phase Detector, as rather than a continuous output the phase detector in an ADPLL has a finite number of output values, thus limiting the accuracy of the phase detector.
-A second form of quantisation is also present, as unlike an analog system, a digital phase detector does not provide continuous data in the time domain either, instead relying on sampling.
-At its most basic, a digital phase comparator may only output an indication of which signal is leading, a design known as a Bang-Bang Detector, which can be constructed using a single D Flip Flop.
-As the output only has two levels the resultant word is only 1 bit wide and as such, limits the range over which the output frequency can be controlled.
-More complex designs such as that in Figure \ref{fig:shan_bb_pd}, implemented by Shan, build on this by measuring the time difference between edges of the signals using a Time-to-Digital Converter, or TDC, in his case using a tapped delay line \cite{shan2014phd}.
-This essentially mimics a counter and allows for a non binary form of phase measurement.
-
-\begin{figure}[h] %TODO top
-	\centering
-	\includegraphics[scale=0.35]{../shan_bb_pd}
-	\caption{Bang-bang phase/frequency detector architecture. \cite{shan2014phd}.}
-	\label{fig:shan_bb_pd}
-\end{figure}
-\subsection{Digital Loop Filter}
-\begin{figure}[h]
-	\centering
-	\includegraphics[scale=0.45]{../pi_simple} %TODO this diagram
-	\caption{Basic PI Controller Architecture.}
-	\label{fig:my_simple_pi}
-\end{figure}
-The loop filter in an ADPLL can be implemented as a PI controller as it is just a low pass filter, such as that in Figure \ref{fig:my_simple_pi}.
-In the case of an ADPLL node in a network the input of this filter is a weighted sum of two to four phase errors, each corresponding to a neighbouring node.
-In a digital system the proportional section is implemented by a simple multiplier, whereas the integral path is a combination of a multiplier and a delayed summation stage.
-The value of these gains determine the response and stability of the oscillator system.
-The transfer function of such a circuit is given by:
-\begin{equation*}
-	H(z) = \alpha + \beta\frac{1}{1-z^{-1}} = \frac{(\alpha + \beta) - \alpha z^{-1}}{1-z^{-1}}~~~\cite{shan2014phd}
-\end{equation*}
-It has been found by Koskin \textit{et al} (2018) that stable operation can be achieved when the integral gain, $k_i$, is less than half the proportional gain, $k_p$ \cite{koskin2018generation}.
-In the same study a range of values was found which would produce low jitter operation of the network.
-As these values are all less than one, the filter must implement fixed point arithmetic in an effort to maintain the simplicity of the clock distribution network, rather than incurring the complexity penalty of floating point calculations.
-
-\section{My Contribution}
-The planned contribution of this project is to create an FPGA based prototyping platform for the testing of network parameters and architectural variations.
-An FPGA as a prototyping tool for networks was previously utilised by Zianbetov and Shan as the final stage before the creation of a custom chip, as it allowed for ``a partial validation of the ASIC design, through a verification of the essential functional features of the designed clocking network'' \cite{zianbetov2013phd,shan2014phd}.
-They note however, that there are some fundamental differences between the two environments, in particular that the FPGA based prototype is unable to create the mixed signal circuits used in an ASIC, resulting in alternative designs being mandated for certain components such as the Time-to-Digital Converter.
-Similarly most commercially available FPGA boards were noted to have clock frequencies well below the gigahertz levels possible in ASIC designs.
-In spite of these problems, the FPGA as a prototyping solution is valuable as it allows for hardware-based testing of architectures and the simulation of non trivial circuit behaviours.
-
-The novelty of this particular implementation is that in the case of Zianbetov and Shan only a small network was simulated on an FPGA and at a frequency close to 50 kHz, however this project will attempt to significantly increase the frequency of operation to the MHz range, as well as expanding the size of the network, from 16 oscillators to potentially 100.
-This is done with the aim of assisting work being undertaken by a number of researchers in UCD working on aspects of ADPLL networks who would stand to benefit from such a platform.
-
-\chapter{Progress to Date}
-The first stage of this project was the creation of an ADPLL design that would be used for the network.
-Due to limitations of the connections available on the FPGA evaluation board used to this point, a Digilent/Xilinx Nexys 4, outputting signals at frequencies significantly beyond 5 MHz, suffered from major degradation as the connections were not designed for such frequencies.
-Thus a target frequency was set at 5 MHz while this board was in use.
-After initial research into potential oscillator designs and what would be feasible given the constraints of the particular platform, design began with the aforementioned oscillator.
-
-\section{Controllable Oscillator}
-In my research into potential designs for numerically controllable oscillators that were viable on an FPGA I came across two main options.
-The first of these was the ring oscillator, constructed from an odd number of inverters strung together in a chain \cite{predraig}.
-In this case the frequency of operation is determined by the number of oscillators in the chain, as the delay through the individual inverters determines the half period of the signal.
-This period can then be modified by adding or removing inverters in a manner that maintains the odd number of total inverters.
-The propagation time through an individual inverter is dependant on the layout of the ring and thus this design must be used with some care.
-
-The second option is one that I was aware of in advance of the beginning of this project.
-A phase accumulator makes use of the overflow property of a counter, and has a frequency determined by the width of the counter and the frequency at which it is clocked.
-The oscillator operates as follows: once per clock cycle the counter will increment by a number $k$ until overflow is reached at $2^n-1$ where $n$ is the bit width of the counter.
-By controlling the magnitude of $k$ added on each cycle of the driving clock the frequency of this oscillator can be adjusted.
-The most significant bit of this counter can form the clock signal, as it has an approximately 50\% duty cycle.
-
-As it is not reliant on the clock of FPGA, the ring oscillator design is a more interesting design in the context of simulation, but the phase accumulator is significantly easier to manage due to being more easily simulatable.
-The decision was made to create both designs in order to keep options open.\\
-The only other option I managed to come across for an oscillator on an FPGA was using Xilinx proprietary \texttt{IODELAY} blocks, which have a permanently programmable delay at runtime \cite{iodelay}.
-This could be coupled with an odd number of inverters in order to implement a quasi ring oscillator.
-However as these are IO blocks, their usage is not particularly conducive to the creation of a network of ADPLLs.
-
-The phase accumulator was designed to have a bit width of 12, all of which could be set as the control code.
-As the maximum frequency of operation that had no timing violations was 258 MHz, the frequency step of this oscillator design was 62.988 kHz and before integration into an ADPLL had a frequency range from 62.988 kHz to just under half the clock frequency at 128.94 MHz.
-Figure \ref{fig:my_pa} details the simple construction of this oscillator.
-\begin{figure}[h]
-	\centering
-	\includegraphics[scale=0.4]{../phase_accum}
-	\caption{Phase Accumulator Block Diagram.}
-	\label{fig:my_pa}
-\end{figure}
-
-As originally implemented for testing purposes, the inverter ring oscillator had a propagation time per inverter of 315 picoseconds, thus each inverter would add 0.63 nano seconds to the total period, and had 4 control bits.
-In order to obtain a 5 MHz output from this type of ring oscillator a ring of of 317 inverters is required, and to set this as the centre point of the frequency range, the total number of inverters was 325.
-At 5 MHz the frequency step of this design was calculated as:
-\begin{align*}
-T_{nom} = 200\times10^{-9}\text{ ns ,}T_{next} = 201.26\times10^{-9}\text{ ns} \\
-F_{nom} = 5\times10^{6}\text{ Hz ,}F_{next} = 4.968697\times10^{-9}\text{ Hz}
-\end{align*} 
-\begin{equation*}
-\Delta F_{RO} = 31.302\text{ kHz}
-\end{equation*}
-\begin{figure}[h]
-	\centering
-	\includegraphics[scale=0.275]{../ring_osc}
-	\caption{Simplified Ring Oscillator Block Diagram.}
-	\label{fig:my_ring_osc}
-\end{figure}
-\section{Phase Detector}
-Moving along the chain the next block is the phase detector, or more accurately, phase comparator.
-At this stage in development this block compares the oscillator's output with that of just one reference.
-In the future this block will be followed up with one which performs a weighted sum of all neighbouring nodes.
-The phase detector used in Shan's design has an arbitration circuit that is not synthesisable and an alternate design is required for use on an FPGA.
-The eventual design I chose has two main parts, firstly an up/down counter performs the role of the Time-to-Digital converter while also indicating the sign of the phase difference.
-In order to control the count and indicate the direction in which to do so, a state machine determines which edge occurs first.
-In order to circumvent the possibility of a metastability condition being reached, the incoming signals are synchronised to the higher frequency FPGA clock, prior to their entry into the phase detector.
-From there, the state machine can operate as designed, in accordance with the state transition diagram in Figure \ref{fig:my_state_machine}.
-\begin{figure}[h]
-	\centering
-	\includegraphics[scale=0.35]{../state_trans_diagram}
-	\caption{Phase Detection State Machine State Transition Diagram.}
-	\label{fig:my_state_machine}
-\end{figure}
-
-The state machine behaves as follows: Starting in the "Waiting" state, if the reference has a rising edge first it is said to be leading the oscillators output and the sign of the phase error is positive.
-Alternatively if the opposite occurs and the first rising edge is seen on the oscillator output, the phase error is given a negative sign and the TDC counts downwards.
-In either case, this continues until a rising edge occurs on the other signal which stops the count, updates the phase error output of the block and finally clears the counter.
-Between "Copy \& Clear" events, the output of the block is held constant.
-There is a second condition that can stop the count and that is the occurrence of a falling edge of the signal that began the count.
-This does however limit the rate of lock acquisition in the case where the phase difference is greater than 180$^\circ$ or where the frequency difference is significant.
-The operation, however, is the same as a conventional method when either locked or within a normal operating tolerance of being locked.
-
-In order to represent both the count value and sign in one value, the counter operates in twos complement which allows the output be used directly in subsequent blocks.
-The counter itself has a bit width of 8, thereby having a range of -128 to 127 which at 5 MHz gives a phase resolution of:
-\begin{align*}
-	T_{res} &= \frac{1}{258\times 10^6} = 3.8760~ns \\
-	T_{period} &= \frac{1}{5\times 10^6} = 200~ns \\
-	\textrm{Ratio} &=\frac{3.8760}{200} = 0.0194 \\
-	\textrm{Resolution} & = 0.0194\times 360^\circ = 6.9768^\circ
-\end{align*}
-
-\section{Loop Filter}
-As with Shan's design, I used a PI controller as the loop filter, in order to align with the typical technology and to benefit from the work of members of the research team in UCD \cite{shan2014phd,koskin2018generation}.
-The bit with of the filter's input is once again 8 bits, to match up with the input from the phase detector and outputs a 9 bit value in order to accommodate the summation of an 8 bit number from each of the paths.
-I selected $k_p = 1$ and $k_i = 0.125$ in order to align with the region of stable coefficients discovered by Eugene \cite{koskin2018generation}.
-\begin{figure}[h]
-	\centering
-	\includegraphics[scale=0.35]{../loop_filter}
-	\caption{PI Loop Filter RTL diagram.}
-	\label{fig:my_pi}
-\end{figure}
-
-Figure \ref{fig:my_pi} showing the RTL diagram of the filter shows the two distinct paths through the filter, the simpler of these being the proportional one.
-The multiplication necessitates the increased width of the path in order to preserve the accuracy of the fixed point calculations, a width which is then cut down to an 8 bit value before being added to the output of the integral branch.
-The proportional gain value has a fixed point width of 2.1 allowing for values from -2 to 1.5, however, in testing to this point only 1 and 0.5 have been used.
-The integral path is similar, with the gain having a fixed width of 1.3, a range of -1 to 0.750, and for my testing set to 0.125.
-After the gain is applied an accumulator with a delayed input clocked by the 258 MHz FPGA clock implements the integration.
-Once again the output of this path is cut back down to an 8 bit value.\\
-The addition of two 8 bit values results in a 9 bit signed output from the phase detection block which is then combined with the bias to form the control code for the oscillator.
-As a higher control code leads to a larger frequency with the phase accumulator, this combination is done by additions, whereas for the ring oscillator the relationship is inverse as we are controlling the period, so instead subtraction is used.
-
-\section{ADPLL}
-Combining these blocks, in this case using the Phase Accumulator style oscillator, gives the ADPLL shown in Figure \ref{fig:my_adpll} on which I performed my initial testing and simulations.
-For ease of testing the PLL replicated the reference exactly, as this would facilitate easier testing of the lock and capture ranges of the design.
-\begin{figure}[h]
-	\centering
-	\includegraphics[scale=0.275]{../rtl}
-	\caption{Phase Accumulator ADPLL RTL diagram.}
-	\label{fig:my_adpll}
-\end{figure}\\
-Once this testing had been completed with the phase accumulator based ADPLL testing progressed to the design which implemented an inverter chain and scaling from the reference using an eight times divider.
-This necessitated a change in the bit width of the control signal to match the 5 bit wide signal desired, thus leading to the design in Figure \ref{fig:my_ring_adpll}.
-\begin{figure}[h]
-	\centering
-	\includegraphics[scale=0.275]{../ro_rtl}
-	\caption{Ring Oscillator ADPLL RTL diagram.}
-	\label{fig:my_ring_adpll}
-\end{figure}\\
- The Loop Filter is clocked at the generated output frequency, as this ensures it will be synchronised with the oscillator.
-This is important when using the inverter chain oscillator, in order to avoid metastability as the generated output clock may not be synchronised with the high frequency clock of the FPGA.
-
-\chapter{Results}
-Initially simulations were utilised to validate the behaviour of the individual blocks and of the system as a whole, while using the Phase Accumulator based design.
-From these simulations, I was able to test the performance of the ADPLL, while locked to a signal at 5 MHz and its behaviour during the acquisition period.
-A signal at 5 MHz lies close to midway between two increments of the phase accumulator and as such the oscillator must switch between two frequencies leading to the introduction of jitter.
-From simulations performed in Vivado locked to a signal at 5 MHz, I observed jitter of 3.491 ns, based on the differences between sequential periods, and a peak-to-peak worst case value of 9.299 ns.
-
-Similarly, I performed simulations of the locking performance of the design, starting with a deviation of 213 KHz and observing the time taken to achieve a lock.
-213 KHz represents an offset of three frequency steps, plus the difference between 5 MHz and the nearest integer frequency step.
-Figure \ref{fig:vivado_sim} contains the visual results of the tests, in which a repetitive pattern can be seen to begin after 20.91 $\mu$s indicating a lock as been acquired.
-\begin{figure}[h]
-	\centering
-	\includegraphics[scale=0.45]{../sim_locking_rect}
-	\caption{Locking Dynamics Simulation.}
-	\label{fig:vivado_sim}
-\end{figure}
-
-Identical tests were performed post-implementation, using a Digilent Nexys4 development board, in addition to measuring the lock and capture ranges of the design.
-Firstly in order to determine the baseline for jitter in the system, the oscillator was biased to 4.9761 MHz and sequential period jitter was observed to be 1.9689 ns.
-Connecting the remainder of the ADPLL and by varying the reference, the lock and capture ranges were found to be 4.19-12.76 MHz and 4.25-12.34 MHz respectively, when the bias point was 5 MHz.
-Somewhat surprisingly this is not centred on the bias point, suggesting a potential flaw in the system that needs to be addressed.
-A identical locking test was performed in the implemented design, Figure \ref{fig:impl}, where using the same method, a time to lock of 25.2 $\mu$s was observed.
-While locked jitter was better in both average period-to-period variation and in peak-to-peak, which can be attributed to the FPGA clock frequency not exactly matching that designed and as such the distance between 5 MHz and the nearest frequency step being reduced, thereby leading to less switching required to generate a 5 MHz signal.
-\begin{figure}[h]
-	\centering
-	\includegraphics[scale=0.45]{../sim_locking_rect}
-	\caption{Locking Dynamics Measurement.}
-	\label{fig:impl}
-\end{figure}
-
-Similarly tests were carried out with the Ring Oscillator based design at a bias point of 5 MHz.
-425 inverters, with 5 bit control, were used to obtain an output in the region of 5 MHz with the bias set to the midpoint which is close to 5 MHz.
-The expected lock range given the use of an otherwise identical design would be the entire range of 32 steps or 64 inverters which can be easily estimated:
-\begin{align*}
-	T_{inv} &= \frac{200\times 10^{-6}}{425-2\times 16} = 0.50891\textrm{~ns} \\
-	F_{min} &= \frac{1}{(425)\times 0.50891\textrm{~ns}} = 4.623\textrm{~MHz} \\
-	F_{max} &= \frac{1}{(425-2\times 32)\times 0.50891\textrm{~ns}} = 5.443\textrm{~MHz}
-\end{align*}
-However in testing the lock and capture ranges, both came out to be 4.86-5.15 MHz, so significantly below the estimates, which highlights an issue that requires investigation.
-
-Testing was also carried out of the ADPLL containing a divider, in both divide by 4 and 8 modes, to examine their impact on locking behaviour.
-The results of this testing are displayed in Table \ref{table:b_c_perf}.
-The jitter in this case is rather low, as the 10 MHz target frequency lies close to an integer increment of the phase accumulator design used for this test.
-Lock and capture ranges are significantly more symmetrical in these configurations than at a target of 5 MHz and, as expected, there is a decrease in range with the greater division values.
-\begin{figure}[h]
-	\centering
-	\includegraphics[scale=0.5]{../matlab/painintheass}
-	\caption{Testing of ADPLLs employing div4 \& div8.}
-	\label{fig:div}
-\end{figure}
-\begin{table}[!ht]
-	\begin{center} 
-		\begin{tabular}{c|c|c|c|c}           
-			& Period Jitter & Peak-Peak Jitter & Lock Range & Capture Range \\
-			div4 & 1.9858 ns & 6.4854 ns & 5.000-17.200 MHz & 5.620-16.700 MHz \\
-			div8 & 1.5702 ns & 6.1594 ns & 5.768-15.262 MHz & 5.928-14.216 MHz \\
-		\end{tabular}
-		\caption{Divider Comparison.}
-		\label{table:b_c_perf}
-	\end{center}
-\end{table}
-
-\chapter{Future Work}
-So far, all work has been done using a single phase lock loop in order to have a suitable basis for the network, and the first stage of network implementation will interlink two  of these PLLs, with one of them also receiving input from the external reference.
-This test will serve as both validation of the phase-coupling ability of the network as well as a test of the error summation block and the ability to lock from startup.
-For the sake of more realistic performance, the inverter chain ADPLL will be used ahead of the Phase Accumulator based design.
-With simple coupling implemented, the logical next stage is the implementation of a small network, possibly 2x2, before scaling up to a size suitable to the FPGA in use.
-At this point, analysis of measurement results can be undertaken and any refinements to the network can be made.
-Additionally, if time permits, the Phase Accumulator based design could be implemented and tested and comparisons made between both designs.
-
-A shortcoming of the current FPGA evaluation board is the inability to input or extract signals in the multiple MHz range without significant distortion, and this will be resolved with the arrival of an FPGA more suited to extracting and importing higher frequency signals.
-With an ability to extract higher frequency signals for measurement and analysis purposes, the target frequency of 5 MHz can potentially be increased into the tens of MHz range.
-
-As with the single ADPLL testing the important characteristics to look out for will be the lock and capture ranges of the network which should mirror that of the ADPLLs used in it's construction.
-Similarly the jitter of a number of clocks across the network will be examined in order, and any skew between the PLL outputs investigated.
-Also of interest will be the initial locking behaviour, in particular how quickly the network can be brought up from ``cold''.
-
-\newpage
-\bibliography{bib} 
-\bibliographystyle{IEEEtran}
-
-\end{document}

### 2019-04-06 thesis/Introduction.tex
initial thesis background

This thesis will put forward the \ac{FPGA} as a tool in the design of \ac{ADPLL} networks to bridge the gap between software simulations and implementation in custom silicon by providing a hardware-based simulation, modelling and validation platform. While an FPGA lacks the direct control over the schematic and layout that an \ac{ASIC} provides, the hardware nature of this platform enables the analysis of system dynamics that are not easily modelled in simulation and the testing.

The goal of this project is to design and implement an extensible platform that can be used by the research team in \ac{UCD} going forward as they seek to understand the behaviour of \ac{ADPLL} networks at a higher level and to serve as a hardware test-bed for proposed new architectures or system components. In order to accomplish this goal a number of potential \ac{ADPLL} architectures will be investigated, implemented and tested to ensure they are function correctly. Individual \acsp{ADPLL}, however, will not give sufficient insight into the behaviour of a network, so once the \ac{ADPLL} designs have been established, each will be implemented as part of a network of increasing sizes. Each contrasting design will be analysed based on the results of measurements and tests, and these results will be used to corroborate claims made regarding which \ac{FPGA} based \ac{ADPLL} design or \ac{ADPLL} network architecture is best suited for particular use cases.

The design of each block, or component part, used in the network will be discussed, starting with the reasons for their selection and an explanation of the design methodology, along with any major pitfalls encountered their creation. The impact on performance caused by modifying the design of these blocks will again be assessed on the basis of measurement results, before comparison is made to both theoretical expectations and their use case.

An \ac{FPGA} based test platform is ideal for those who wish to examine system dynamics without the time delays, financial burden or expertise required to develop a complete mixed-signal system on an \ac{ASIC} but retain the ability to realistically simulate the behaviour of an \ac{ADPLL} network. The result of this project is such a platform, designed to be extensible, with flexibility built into each component/module used.

### 2019-04-07 thesis/Background.tex
sunday funday pt1

@@
 The designs described previously, are all similar in that they have a single central oscillator that provides the clock for all areas of the chip, whereas the following methods attempt to synchronise multiple oscillators, each of which provides the clock for a single clocking area.
-The main advantage of a multi-oscillator design, is that as each clocking area has its clock created locally, there is no problem with signal degradation.
-Similar to a decentralised skew compensation scheme comparisons are only made to neighbouring blocks and as such there is reduced power loss due to transmission.
+The main advantages of a multi-oscillator design, are that as each clocking area has its clock created locally, there is no degradation in the quality of the signal as it is distributed around the chip and the number of potential noise sources is reduced.
+In order to obtain global synchronisation some method of comparison between local clocking areas is required, and how this is done depends on the architecture.
+Regardless of the comparison is made, it is carried out between neighbouring clocking areas and as such the feedback network need not have a large footprint or power overhead.
+
+One such method is a network of oscillators as in Figure \ref{fig:mizuno1998noise} which uses coupled \ac{PLL} to generate local clocks.
+Here the output of a leaf node is compared with an external reference and the operating frequency of each \ac{VCO} tuned based on the result.
 \begin{figure}[h]
@@
 \end{figure}
-One such method is a network of oscillators as in Figure \ref{fig:mizuno1998noise} which uses coupled PLLs to generate local clocks.
-The advantage of this method is that only the control voltage $v_c$ needs delivery to all areas of the chip, thus alleviating the need for a power hungry distribution circuit, while also being more noise-immune than the transmission of a high frequency clock.
-However this design still suffers from clock variation as all VCOs are fed the same control voltage, and this is acknowledged by the authors:
+The advantage of this method is the simplicity of the feedback network, requiring just the divided clock output from a single leaf node.
+The \ac{VCO}s then adjusted by the control voltage, $v_c$, which needs delivery to all areas of the chip.
+However, this is a regular signal and as such does not suffer from skew or jitter.
+This alleviates the need for a power hungry distribution circuit, while also being more noise-immune than the transmission of a high frequency clock.
+However this design still suffers from clock variation as all \ac{VCO}s are fed the same control voltage, and thus the manufacturing tolerance issues present in conventional designs persists here also.
+This is acknowledged by the authors:
 \begin{quotation}
 	\textit{Unfortunately, as with the conventional ...
-method, distributing the VCOs over the entire chip causes the problem that jitter and skew are increased by variations in the fabrication process (static), temperature, and power supply (dynamic)} \cite{mizuno1998noise}.
+method, distributing the \ac{VCO}s over the entire chip causes the problem that jitter and skew are increased by variations in the fabrication process (static), temperature, and power supply (dynamic)} \cite{mizuno1998noise}.
 \end{quotation}
-This design also only compares the phase of a single branch of the distribution network with the reference, in order calculate the control voltage which again may contribute to a degradation in clock synchronisation.
-%TODO why
+This type of multi-oscillator design is implemented by analogue circuits, and as a result not only are the clock signals, but also the control signals are liable to variation due to noise, fabrication mismatch and power supply dynamics.
 
-Another potential multi-oscillator design synchronises the oscillators using the phase relationship with the oscillators in the neighbouring clocking area.
-Once again, this negates the requirement for a global distribution structure and comparisons need only be made with neighbouring areas.
-As a phase lock loop is being used only a divided down version of the clock is required.
-This in turn means the hardware transporting the divided clock signal to the phase comparator, has significantly lower requirements placed on it, thus lowering the power consumption.
-Pratt and Nguyen initially proposed this design in their paper published in 1995 entitled ``\textit{Distributed Synchronous Clocking}'' in which they propose a Cartesian grid of clocking areas, each with their own PLL, which has become known as a PLL Network \cite{pratt1995distributed}.
+Another potential multi-oscillator clock distribution system uses the phase relationship between the oscillators driving neighbouring clock areas in order to obtain synchronisation.
+Once again, this negates the requirement for a global distribution structure and the signals used for comparisons need only be sent between neighbouring clocking areas.
+As a \ac{PLL} is being used it is again possible to perform the phase comparisons using a divided version of the generated clock.
+This in turn means the hardware transporting the divided clock signal to the phase comparator, has significantly lower requirements placed on it, thus lowering the power consumption due to electrical losses.
+Pratt and Nguyen initially proposed method of clock distribution in their 1995 paper entitled ``\textit{Distributed Synchronous Clocking}'' in which they propose a Cartesian grid of clocking areas, each with their own \ac{PLL}, which has become known as a \ac{PLL} Network \cite{pratt1995distributed}.
 In this design any given node is synchronised with its neighbours and one of the corner nodes is additionally synchronised with the reference.
 According to the authors this is ``a simple, effective way to achieve low cost, high quality, low skew clock generation in a synchronous parallel processor''.
-This methodology was implemented by Gutnik \textit{et al} (2000) who found:
-\begin{quote}
-	\textit{Design and measurements on this chip confirm that generating and synchronizing multiple clocks on chip is feasible.
-Neither the power nor the area overhead of multiple PLLs is substantial compared to the cost of distributing the clock by conventional means} \cite{gutnik2000active}.
-\end{quote}
+They did, however, note the presence of a phenomenon called ``mode locking'', which is setting of the network into a stable equilibrium where there are non-zero relative phases.
+
 \begin{figure}[h]
@@
 
+This architecture of clock distribution network was then implemented by Gutnik \textit{et al} (2000) who fabricated a 4x4 array of oscillators, operating at a centre frequency of 1.2 MHz.
+The oscillator was implemented as a voltage controlled ``nMOS-loaded differential ring oscillator'', and in order to mode locking the phase detector was implemented as a highly non-linear circuit.
+The design was a success and the authors concluded:
+\begin{quote}
+	\textit{Design and measurements on this chip confirm that generating and synchronizing multiple clocks on chip is feasible.
+Neither the power nor the area overhead of multiple \acs{PLL}s is substantial compared to the cost of distributing the clock by conventional means} \cite{gutnik2000active}.
+\end{quote}
 \section{ADPLL Networks}
-As a PLL network is an analog circuit, its integration in a modern IC is a barrier to usage and it has not been used in any commercial designs \cite{zianbetov2013distributed}.
-An alternative design that is more suitable for current fabrication techniques eschews from using analog components and instead implements the network of controlled oscillators using only digital circuitry, hence the name All-Digital PLL.
-A 4x4 ADPLL network was designed and prototyped in 65 nm CMOS by Zianbetov and Shan in order to test the suitability of the technique as a clock distributor \cite{zianbetov2013phd,shan2014phd}.
-%TODO fix phd references
+As a \ac{PLL} network is an analog circuit, its integration in a modern \ac{IC} is a barrier to usage, and as such it has not been used in any commercial designs \cite{zianbetov2013distributed}.
+An alternative design that is more suitable for current fabrication techniques eschews from using analogue components and instead implements the network of controlled oscillators using only digital circuitry, hence the name All-Digital \ac{PLL}.
+A 4x4 \ac{ADPLL} network was designed and prototyped in 65 nm CMOS by Zianbetov and Shan in order to test the suitability of the technique as a clock distributor \cite{zianbetov2013phd,shan2014phd}.
 
-In this design the oscillators are coupled using digital phase comparators which attempt to measure the phase error between two oscillators with each non edge node being connected to four neighbours.
+In this design the oscillators are once again laid out in a Cartesian grid, with each node coupled to their neighbours in phase.
+As this is now a digital system the coupling is carried out using digital phase comparators, which attempt to measure the phase difference between two oscillators.
 Figure \ref{fig:eldar_node} shows high level detail of the architecture of both the entire clocking system and that of an individual node in the design.
-A digital PLL network has the additional advantages over its analog counterparts that it can benefit from advancements in digital circuit design suites, be reconfigurable and programmable and has a significantly greater immunity to perturbations inherent to its digital nature \cite{zianbetov2013phd}.
+The digital nature of this architecture brings with it a number of advantages over traditional analogue implementations, as it can benefit from advancements in digital circuit design suites, be reconfigurable and programmable and has a significantly greater immunity to perturbations inherent to its digital nature, as the exact voltage of signals is of no importance \cite{zianbetov2013phd}.
 This last advantage is of particular use in a digital environment, as otherwise there is potential for clock degradation resulting from switching of transistors.
+The drawback of the switch to a digital architecture however is the presence of quantisation.
+Analogue designs both deliver continuous control signals to the oscillators and have a continuous phase detection capability, unlike a digital system where these actions are carried out with fixed resolution.
 \begin{figure}[h]
@@
 \end{figure}
-Looking at the design of a given node it is notable that the function carried out by the Error Combiner is akin to a average, therefore as mentioned by Pratt and Nyugen, there is potential to settle in a stable ``mode'' in which the oscillators are not synchronised.
-In their paper, they present a method where initial start-up is performed uni-directionally and, once all nodes are close to alignment, full connectivity can be restored, however this was not viable in an analog system \cite{pratt1995distributed}.
-In creating an entirely digital system, Zianbetov and Shan could exploit the reconfigurability and and implement this system at start-up.
+
+Looking at the design of a given node it is notable that the function carried out by the Error Combiner is akin to a average, therefore as mentioned by Pratt and Nyugen, there is potential to a mode locked equilibrium in which the oscillators are not synchronised.
+In their paper, they presented a method where initial start-up was performed uni-directionally and, once all nodes are close to alignment, full connectivity could be restored, however this was not viable in an analogue system as reconfigurability was not an option \cite{pratt1995distributed}.
+In creating an entirely digital system, Zianbetov and Shan could exploit reconfigurability and implement a uni-directional start-up and thus avoid the problem of convergence into a mode locked state, without having to design a non-linear phase detector.
 
 \section{ADPLL Architecture}
-As indicated in Figure \ref{fig:mulkeen_pll} the three main building blocks of a conventional Phase Lock Loop are the Phase Detector, Loop Filter and Controllable Oscillator.
-In the All-Digital counterpart each of these traditionally analog components are replaced by their digital counterparts, necessitating quantisation in order to remain physically realisable.
+As indicated in Figure \ref{fig:mulkeen_pll} the three main building blocks of a conventional \ac{PLL} are the \ac{PD}, \ac{LF} and \ac{VCO}.
+In an \ac{ADPLL} these blocks are then replaced by their digital counterparts, necessitating quantisation in order to remain physically realisable.
 \begin{figure}[h]
@@
 	\includegraphics[scale=0.5]{../tex_files/mulkeen_pll}
-	\caption{Block Diagram of a Phase Lock Loop, \textit{Wireless Systems Notes}, B. Mulkeen (2017).}
+	\caption{Block Diagram of a \acl{PLL}, \textit{Wireless Systems Notes}, B. Mulkeen (2017).}
 	\label{fig:mulkeen_pll}
 \end{figure}
+
+\subsection{Digitally Controllable Oscillator}
+In a digital system there are a very limited number of voltages representable, most commonly just two, so using a voltage to control the oscillator is not a viable strategy.
+Instead a fixed bit width signal is used to control the oscillator's period, selecting the number of inverters in a ring oscillator or the varactor configuration of a travelling wave oscillator \cite{chen2011rotary}.
+The decisions made in the design of the \ac{DCO}, or \ac{NCO}, determine many of the other \ac{ADPLL} parameters.
+While tuning range and centre frequency, as well as linearity, carry over from the analogue counterpart, a \ac{DCO} also has a frequency step which in combination with the bit width of the control signal determines the range over which the oscillator can be tuned.
+Figure \ref{fig:my_ring} illustrates a basic ring oscillator design.
+A ring oscillator is an inherently unstable circuit composed of an odd number of inverters connected in a circle, which allows a signal to propagate infinitely, with the signal at any point in the circuit appearing as a square wave.
+The half-period of this oscillator is the time taken for the signal to propagate once through the chain, $n$ times the propagation delay through one inverter.
+The frequency of operation can then be set by modulating the length of the chain, in steps of two inverters to maintain an odd number, by means of \texttt{f\_select}.
+The main impact of output frequency quantisation is that only frequencies which are integer multiples of the frequency step away from the centre frequency can be easily reproduced, with intermediate values only obtainable in a manner akin to Fractional-N synthesis with the control code toggling back and forth.
+This acts as a source of jitter in the system.
+\begin{figure}[h]
+	\centering
+	\includegraphics[scale=0.275]{../inverter_chain}
+	\caption{Basic Ring/Inverter Chain Oscillator.}
+	\label{fig:my_ring}
+\end{figure}
+
+It is also possible to implement a \ac{NCO} by means of a counter, in a manner that will produce either linear period or linear frequency steps.
+The former is achieved by setting the reload point of the counter
+
+\subsection{Digital Phase Detector}
+Once again quantisation impacts the Phase Detector, as rather than a continuous output the phase detector in an ADPLL has a finite number of output values, thus limiting the accuracy of the phase detector.
+A second form of quantisation is also present, as unlike an analog system, a digital phase detector does not provide continuous data in the time domain either, instead relying on sampling.
+At its most basic, a digital phase comparator may only output an indication of which signal is leading, a design known as a Bang-Bang Detector, which can be constructed using a single D Flip Flop.
+As the output only has two levels the resultant word is only 1 bit wide and as such, limits the range over which the output frequency can be controlled.
+More complex designs such as that in Figure \ref{fig:shan_bb_pd}, implemented by Shan, build on this by measuring the time difference between edges of the signals using a Time-to-Digital Converter, or TDC, in his case using a tapped delay line \cite{shan2014phd}.
+This essentially mimics a counter and allows for a non binary form of phase measurement.
+
+\begin{figure}[h] %TODO top
+	\centering
+	\includegraphics[scale=0.35]{../shan_bb_pd}
+	\caption{Bang-bang phase/frequency detector architecture. \cite{shan2014phd}.}
+	\label{fig:shan_bb_pd}
+\end{figure}
+\subsection{Digital Loop Filter}
+\begin{figure}[h]
+	\centering
+	\includegraphics[scale=0.45]{../pi_simple} %TODO this diagram
+	\caption{Basic PI Controller Architecture.}
+	\label{fig:my_simple_pi}
+\end{figure}
+The loop filter in an ADPLL can be implemented as a PI controller as it is just a low pass filter, such as that in Figure \ref{fig:my_simple_pi}.
+In the case of an ADPLL node in a network the input of this filter is a weighted sum of two to four phase errors, each corresponding to a neighbouring node.
+In a digital system the proportional section is implemented by a simple multiplier, whereas the integral path is a combination of a multiplier and a delayed summation stage.
+The value of these gains determine the response and stability of the oscillator system.
+The transfer function of such a circuit is given by \cite{shan2014phd}:
+\begin{equation*}
+	H(z) = \alpha + \beta\frac{1}{1-z^{-1}} = \frac{(\alpha + \beta) - \alpha z^{-1}}{1-z^{-1}}
+\end{equation*}
+It has been found by Koskin \textit{et al} (2018) that stable operation can be achieved when the integral gain, $k_i$, is less than half the proportional gain, $k_p$ \cite{koskin2018generation}.
+In the same study a range of values was found which would produce low jitter operation of the network.
+As these values are all less than one, the filter must implement fixed point arithmetic in an effort to maintain the simplicity of the clock distribution network, rather than incurring the complexity penalty of floating point calculations.
+
+\section{The Role of FPGAs}

### 2019-04-07 thesis/Background.tex
end of sunday

@@
 \end{quote}
+
+The remaining benefits of such a clock distribution system are: As the individual oscillators have their own control signal mismatch between different oscillators is not a factor as they will also receive different control signals.
+Secondly, and unlike the conventional methods, sources of jitter in the system such as power supply dynamics can be accounted for.
+Finally symmetry between the different oscillators is not required, once again attributable to the individual control signals in use.
 \section{ADPLL Networks}
@@
 
-\section{ADPLL Architecture}
+\section{\acl{ADPLL} Architecture}
 As indicated in Figure \ref{fig:mulkeen_pll} the three main building blocks of a conventional \ac{PLL} are the \ac{PD}, \ac{LF} and \ac{VCO}.
 In an \ac{ADPLL} these blocks are then replaced by their digital counterparts, necessitating quantisation in order to remain physically realisable.
+The ``All-Digital'' moniker is a misnomer as the oscillator and \acl{PD} are usually both implemented by mixed signal circuits.
 \begin{figure}[h]
@@
 
-\subsection{Digitally Controllable Oscillator}
+\subsection{Digitally Controlled Oscillator}
 In a digital system there are a very limited number of voltages representable, most commonly just two, so using a voltage to control the oscillator is not a viable strategy.
@@
 
-It is also possible to implement a \ac{NCO} by means of a counter, in a manner that will produce either linear period or linear frequency steps.
-The former is achieved by setting the reload point of the counter
+It is also possible to implement an \ac{NCO} by means of a counter, in a manner that will produce either linear period or linear frequency steps.
+Both methods use the most significant bit of the counter's value to form the output signal.
+Period linearity is achieved by varying the reload value of the counter after overflow depending on the control code, thereby changing the period by a multiple of a fixed step.
+Alternatively frequency linearity can be achieved if the reload value is left constant, but instead the amount added to the counter every clock cycle is changed according to the control code, once again a fixed step size is used.
 
 \subsection{Digital Phase Detector}
-Once again quantisation impacts the Phase Detector, as rather than a continuous output the phase detector in an ADPLL has a finite number of output values, thus limiting the accuracy of the phase detector.
+Once again quantisation impacts the \acl{PD}, as rather than a continuous output the phase detector in an \ac{ADPLL} has a finite number of output values, thus limiting the accuracy of the phase detector.
 A second form of quantisation is also present, as unlike an analog system, a digital phase detector does not provide continuous data in the time domain either, instead relying on sampling.
-At its most basic, a digital phase comparator may only output an indication of which signal is leading, a design known as a Bang-Bang Detector, which can be constructed using a single D Flip Flop.
+At its most basic, a digital phase comparator may only output an indication of which signal is leading, a design known as a Bang-Bang Detector, which can be constructed using a single D Flip Flop with one the generated signal connected to the ``D'' input and the reference signal acting as the clock.
 As the output only has two levels the resultant word is only 1 bit wide and as such, limits the range over which the output frequency can be controlled.
-More complex designs such as that in Figure \ref{fig:shan_bb_pd}, implemented by Shan, build on this by measuring the time difference between edges of the signals using a Time-to-Digital Converter, or TDC, in his case using a tapped delay line \cite{shan2014phd}.
-This essentially mimics a counter and allows for a non binary form of phase measurement.
+More complex designs such as that in Figure \ref{fig:shan_bb_pd}, implemented by Shan, build on this by measuring the time difference between edges of the signals using a \ac{TDC} in his case using a \ac{TDL} \cite{shan2014phd}.
+A \ac{TDL} is constructed by a chain of elements of a fixed delay, and the signal to be timed is applied to this start of this chain.
+After the timing interval elapsed the values at each point in the chain are examined, and using temperature coding, these are converted to a digital signal, the width of which is the bit width of the \ac{PD}'s error signal.
+This mimics a time measurement and allows for the phase difference to be recorded in a non binary manner.
 
-\begin{figure}[h] %TODO top
+\begin{figure}[h]
 	\centering
@@
 \end{figure}
+
 \subsection{Digital Loop Filter}
@@
 	\centering
-	\includegraphics[scale=0.45]{../pi_simple} %TODO this diagram
+	\includegraphics[scale=0.45]{../pi_simple}
 	\caption{Basic PI Controller Architecture.}
 	\label{fig:my_simple_pi}
-\end{figure}
-The loop filter in an ADPLL can be implemented as a PI controller as it is just a low pass filter, such as that in Figure \ref{fig:my_simple_pi}.
-In the case of an ADPLL node in a network the input of this filter is a weighted sum of two to four phase errors, each corresponding to a neighbouring node.
-In a digital system the proportional section is implemented by a simple multiplier, whereas the integral path is a combination of a multiplier and a delayed summation stage.
-The value of these gains determine the response and stability of the oscillator system.
-The transfer function of such a circuit is given by \cite{shan2014phd}:
+\end{figure}%TODO ask mulkeen if this is IIR
+The \acl{LF} in an \ac{ADPLL} can be implemented as a \ac{PI} controller, as only a low-pass filter is required, such as that in Figure \ref{fig:my_simple_pi}.
+In the case of a node in an \ac{ADPLL} network the input of this filter is a weighted of the phase difference relative to the neighbouring local clocking areas.
+In one example topology a digital system the proportional section can be implemented by a simple multiplier, whereas the integral path is constructed by adding the result of a multiplication by the proportional gain to an accumulator.
+This delayed summation can be easily implemented by an accumulator to which the current value of the multiplication is added each cycle, and as such the system has an infinite impulse response.
+The value of these gains determine the response and stability of the \ac{ADPLL} network.
+The transfer function of such a controller is given by \cite{shan2014phd}:
 \begin{equation*}
@@
 
-\section{The Role of FPGAs}
+\subsection{Error Combiner}
+The \ac{ADPLL}s used in a network need to combine the error signals from multiple neighbours to determine what the average difference from its neighbours is, and this necessitates the addition of the Error Combiner.
+In a digital system this can be implemented by a weighted average of the different error signals, with the weight being modifiable at run-time.
+This configurability is what permits the system to implement uni-directional mode and also allows for the weighting applied to certain signals, such as the external reference, to be modified.
+The ease of implementation of a configurable error combiner is one of the main advantages of an \ac{ADPLL} over an analogue system.
+
+\section{The Role of the FPGA}
+A \acl{FPGA} is a type of \ac{IC} that is designed to be configured by a designer after the chip itself has been manufactured.
+An \ac{FPGA} contains a large number of logic elements that can be connected together in order to perform complex logic, written using the same \ac{HDL}s used to design the digital blocks of \ac{ASIC}s.
+They may also implement started modules such as adders, multiplexers and \ac{RAM} as a fundamental element.
+More complex logic is often implemented using multiplexed lookup tables rather than true logic elements.
+High end devices such as the Xilinx Zynq Ultrascale even implement Multi-Processor \ac{SOC}s.
+Compared to an \ac{ASIC} the designer does not have direct control over the layout of the system but rather describes its behaviour, possibly down to the basic logic elements of inverters or other gates.
+Limited control is possible over the placement of the individual modules, but \ac{EDA} tools are responsible for the exact placement of elements.
+As a result, it is not possible to have precise control over the delays experienced as signals propagate through the design.
+To assist with the resolution of any issues \ac{EDA}s provide tools to analyse timing behaviour.
+
+Prototyping on an \ac{FPGA} is a common verification stage for conventional \ac{ASIC} designs as it allows for a hardware validation of any digital circuitry, and the detection of any potential flaws or errors made by the design before the expensive of an \ac{ASIC} implementation.
+In their 2013 \ac{ADPLL} network implementation Zianbetov and Shan used an \ac{FPGA} in order to validate their programming interface, the design of the error processing block, ensure they had eliminated mode locking behaviour and to ensure phase synchronisation was possible \cite{zianbetov2013phd,shan2014phd}.
+However they experienced two main limitations, they were not able to implement the mixed-signal \ac{PFD} and \ac{DCO}, and the maximum frequency of operation possible was orders of magnitude lower than the GHz range of their \ac{ASIC} implementation.
+These issues were circumvented by implementing an alternative \ac{PFD} and \ac{DCO} designs which were driven by the clock distribution network provided by the \ac{FPGA} and every clock frequency scaled by the same amount such that the results of testing would remain indicative.
+One of the main advantages they saw was that the hardware description used for the digital blocks of their \ac{ASIC} could be directly ported over to the \ac{FPGA}.
+
+It is however possible to implement limited mixed-signal circuits on an \ac{FPGA} through the use of primitive logic elements, however, as control over the implementation is restricted to the module level it is not possible to mirror the implementation of a design intended for an \ac{ASIC}.
+As these implementations are analogous to a mixed-signal circuit on an \ac{ASIC}, the verification of theoretical behaviours, as done by Koskin \textit{et all}, in order to test the findings of his PhD thesis \cite{theboys2019}, is made possible without the expense of \ac{ASIC} fabrication.
+An \ac{FPGA} based mixed-signal \ac{ADPLL} network is also seeing use here in \ac{UCD} as a initial prototyping platform for the validation of new modules for use in \ac{ASIC} based \ac{ADPLL} networks.
+
+\ac{FPGA}s have been used in other fields to simulate and experiment with new technologies, although not all of these attempt to implement the mixed-signal circuitry using primitive elements.
+Fernandez-Alvarez \textit{et al} (2016) proposed a method suited to higher end \ac{FPGA}s in which a co-processor on the \ac{FPGA} simulates the mixed-signal circuitry while the digital section of the design is implemented on the \ac{FPGA} itself \cite{fernandez2017hw}.
+While the interfacing between hardware and software remains a challenge they found:
+\begin{quotation}
+	Obtained data are compared to the data obtained by means of using PSIM and ModelSim co-simulation.
+The proposed solution speeds up the evaluation in around one order of magnitude keeping the accuracy.
+The output signal differs in less than 0.6 mV (RMSD).
+\end{quotation}
+
+Mixed-signal circuits were, however, simulated in hardware by \'{O}scar Luc\'{i}a \textit{et al} (2011) on an \ac{FPGA} in order to overcome the excessive time penalty imposed by software based simulations that required the behaviour of both a digital and mixed-signal peripheral and that of a micro controller running code in C to be simulated side by side \cite{lucia2011real}.
+They concluded
+\begin{quotation}
+	...
+the proposed system provides a versatile and fast method to develop ad hoc control architectures, avoiding the need for time-consuming mixed-signal simulations and the risk of damaging the actual power converter implementation.
+
+\end{quotation}
+By carrying out simulations on an \ac{FPGA}, Guanhua Wang \textit{et al} (2013) achieved a 3000 times decrease in run-time when compared to an identical simulation in MATLAB used for the verification of a calibration algorithm for successive approximation analog-to-digital converters \cite{wang2013fast}.
+Many other examples exist of \ac{FPGA}s used for the simulation of mixed-signal or \acl{RF} circuits in literature.
+%TODO Elena had suggestions?

### 2019-04-09 thesis/FPGABasedADPLLs.tex
weak af thursday

\section{Chapter Overview}
The first step in creating an \ac{FPGA} based network of \ac{ADPLL}s is the design of the \ac{ADPLL} itself, which will be addressed in this chapter. The nature of an \ac{FPGA} necessitates a number of compromises in the design of a given block which limits transferability to \ac{ASIC} designs. In this chapter the potential designs for each individual block, or module, investigated will be explained and the case for their selection in an \ac{FPGA} based \ac{ADPLL} made. A number of blocks have implement purely digital circuitry and as such can be transferred in their entirety from an \ac{FPGA} and vice versa. However, those that will be used to emulate mixed-signal circuitry, such as the \ac{DCO} and \ac{PFD}, will be examined in greater detail.
\section{Digitally Controlled Oscillators}
The choices made in the design of the \ac{DCO} have the greatest impact on the effectiveness of the overall platform and which use cases it it suitable for, as the key

### 2019-04-09 thesis/Background.tex

@@
 \end{quotation}
-By carrying out simulations on an \ac{FPGA}, Guanhua Wang \textit{et al} (2013) achieved a 3000 times decrease in run-time when compared to an identical simulation in MATLAB used for the verification of a calibration algorithm for successive approximation analog-to-digital converters \cite{wang2013fast}.
+By carrying out simulations on an \ac{FPGA}, Guanhua Wang \textit{et al} (2013) achieved a 3000 times decrease in run-time when compared to an identical simulation in MATLAB used for the verification of a calibration algorithm for successive approximation analogue-to-digital converters \cite{wang2013fast}.
 Many other examples exist of \ac{FPGA}s used for the simulation of mixed-signal or \acl{RF} circuits in literature.
 %TODO Elena had suggestions?
+
+\section{\ac{ADPLL} Performance Characterisation}
+As already stated, goal of a clock distribution system is to synchronise clocking events in all areas of the chip.
+The effectiveness of this synchronisation is characterised by two main metrics, jitter and skew, which together describe the distribution of clocking events.
+Skew is the average time delay between a clocking event in one area of the chip and that of a reference event.
+It can be easily measured by computing the average value of this delay.
+Jitter has a number of definitions depending on how it is measured, but at its most basic it is the standard deviation of the time delays with respect to the reference clocking edge.
+
+The simplest form of clock performance characterisation is on a \ac{C2C} basis, in which the reference is the signal under test itself.
+Here the standard deviation of the individual periods gives the jitter of the signal, while skew has no meaning with a self reference the mean period can be used to compute the centre frequency over the time interval.
+For a more informative measurement the \ac{TIE} can be computed, which takes into account another signal as the reference.
+\ac{TIE} is calculated by comparing the delay between the signal in question and a reference that is treated as ideal.
+The mean value gives the relative skew between the two signals and once more the standard deviation of the measurements gives the jitter with respect to this reference.
+
+Both above forms of measurement assume a large number of sequential measurements, however, there are other ways that jitter and skew can be calculated \cite{sadsadad}.
+Phase jitter is an important characteristic for communications systems as it can be used to calculate phase noise, important to ensure spurious emissions are within regulations.
+Long term, or accumulated, jitter represents the cumulative effect of jitter on the signal over several cycles.
+Long term jitter in particular affects \acs{RADAR} as it will manifest itself as a Doppler shift in the return signal.
+For an \ac{PLL} network, with an appropriately designed filter, the long term jitter should be zero so long as the reference signal is stable.
+%TODO ask Brian about this one

### 2019-04-10 thesis/Background.tex
4.5 pages down

@@
 
-\section{\ac{ADPLL} Performance Characterisation}
+\section{\acs{ADPLL} Performance Characterisation}
 As already stated, goal of a clock distribution system is to synchronise clocking events in all areas of the chip.
@@
 
-Both above forms of measurement assume a large number of sequential measurements, however, there are other ways that jitter and skew can be calculated \cite{sadsadad}.
+Both above forms of measurement assume a large number of sequential measurements, however, there are other ways that jitter and skew can be calculated \cite{AN007}.
 Phase jitter is an important characteristic for communications systems as it can be used to calculate phase noise, important to ensure spurious emissions are within regulations.

### 2019-04-10 thesis/FPGABasedADPLLs.tex
4.5 pages down

@@
 However, those that will be used to emulate mixed-signal circuitry, such as the \ac{DCO} and \ac{PFD}, will be examined in greater detail.
+
 \section{Digitally Controlled Oscillators}
-The choices made in the design of the \ac{DCO} have the greatest impact on the effectiveness of the overall platform and which use cases it it suitable for, as the key
+The choices made in the design of the \ac{DCO} have the greatest impact on the effectiveness of the overall platform and which use cases the \ac{ADPLL} containing it are suitable for, as the key performance benchmarks are all done using the waveform this block generates.
+This project will address three distinct designs of \ac{ADPLL}s suitable for implementation on an FPGA, two derived from the clocks generated by the \ac{FPGA}s own distribution network and one generated independently of this clock, using a chain of inverters.
+These are not the only ways in which an oscillator could be synthesised on an \ac{FPGA}, however other designs were deemed to be unsuitable for extensible and portable implementations.
+
+A prime example of this is the use of Xilinx proprietary \texttt{IODELAY} blocks to create an oscillator, as detailed in Xilinx Application Note XAPP872 \cite{iodelay}.
+The key idea here is that the bulk of the period is made up by the propagation time through one of the \texttt{IODELAY} blocks, which can be set at implementation time.
+This is combined with a section of an inverter chain, and a multiplexer used modify the length of this segment, the output of which is fed back into the \texttt{IODELAY} block.
+This method was discarded as the number of \texttt{IODELAY} blocks is very limited, so expanding to a larger network would be impossible, and they are all located around the edge of the chip, not suited to the construction of a Cartesian grid.
+
+The main issue with the creation of \ac{DCO}s on an \ac{FPGA} is the inability to create mixed-signal circuits, such as those that would be intended for use on an \ac{ASIC}.
+As such the \ac{FPGA} based oscillator must emulate the behaviour of a mixed-signal circuit in some way.
+
+\section{\acs{FPGA} Driven, Linear Period \acs{DCO}}
+The first design of \ac{DCO} to be examined is of the type used by Zianbetov in his \ac{ADPLL} network test bed and relies on a counter driven by the clock manager on the \ac{FPGA} \cite{zianbetov2013phd}.
+At each event on the \ac{FPGA} provided clock a counter is incremented, overflowing upon increment past the maximum possible value.
+The \ac{MSB} of this counter forms the waveform generated by this oscillator, the period of which is controlled through an adjustable value that is loaded into the counter once overflow is reached, and forms the starting points for the counter.
+The period of oscillation is given by:
+\begin{equation}
+	T_{osc} = \big(2^{width} -~(BIAS+CC)~\big)\times T_{FPGA}
+\end{equation}
+where $T_{FPGA}$ is the clock period of the \ac{FPGA}, $CC$ is the control code input, $BIAS$ centres the oscillator in the middle of the tuning range in the event that the control code is zero and $width$ is the width of the counter used.
+As the only variable here is the control code, period step of this design is $T_{FPGA}$, thereby providing period linearity with respect to the control code.
+This is the key advantage of this design, as most \ac{ASIC} implementations of a \ac{DCO} are also linear in period.
+The other main reason to choose this design is that its \ac{FPGA} clocked nature allows for exact control over the frequency of operation, and the number of tunable parameters make it possible to configure multiple ways to achieve the same frequencies of operation.
+Combined these attributes make it very easy to create an oscillator that emulates the behaviour of a design intended for an \ac{ASIC}, however, at a greatly reduced frequency.
+This restriction on the frequency of operation arises out of the period step size, which in order to obtain a good resolution must be orders of magnitude smaller than the intended period to be generated.
+As the output waveform is taken from the counter's \ac{MSB}, the reload value of the counter must never go beyond $2^{width-1}$, as otherwise the output waveform will become a constant 1.
+As the reload value varies the low time of the \ac{MSB}, if the desired output waveform is a square wave this design will not be suitable.
+
+In being \ac{FPGA} clocked this design has pseudo-deterministic characteristics, with each period step being almost identical across oscillators and over the entire tuning range, unlike an \ac{ASIC} where process variation will impact the layout of a high frequency oscillator.
+The only variation in this design will come, ironically, from jitter or skew in the \ac{FPGA}'s clock distribution network, which as the frequencies will typically be in the low hundreds of MHz is very minor.
+In the case of the Xilinx \acl{Nexys} this is at most 100 picoseconds, or 0.05\% of the period of an intended output clock at 5 MHz.
+To put this value into perspective, on this board the minimum value of $T_{FPGA}$ that could be used to drive this oscillator is 3.87 nanoseconds, 1.935\% of the period.
+
+The resulting \ac{DCO} is best suited to applications that do not seek to gain a better understanding of oscillator performance, but rather those focused on validating the entirely digital blocks in the system, the role in which Zianbetov and Shan used this type of oscillator \cite{zianbetov2013phd,shan2014phd}.
+
+\section{\acs{FPGA} Driven, Linear Frequency \acs{DCO}}
+The second \ac{FPGA} clocked oscillator is similar in most attributes to the above design but eschews period linearity for frequency linearity.
+Again the overflow property of a counter is used with the counter's \ac{MSB} as output of the block, however, this time it forms a square wave.
+Rather than setting the reload value of the counter, instead the increment is adjusted depending on the control code, thus requiring $\frac{2^N}{BIAS+CC}$ increments to overflow.
+Accordingly the frequency of operation is set by:
+\begin{equation}
+	f_{osc} = f_{FPGA}\times\frac{BIAS+CC}{2^{width}}
+\end{equation}
+Here the control code $CC$ and bias are added to the value stored in the accumulator at each event of the \ac{FPGA}, clock until overflow is reached.
+This occurs at $2^{width}$ where, as before, $width$ is the bit width of the counter, thus valuing each control code increment at $\frac{f_{FPGA}}{2^{width}}$ Hz.
+As with the previous design, this oscillator is better suited to frequencies where the output of the \ac{DCO} is orders of magnitude lower than the clock signal driving it, as this ensures that the incremental change due to the control code remains a small fraction of the period.
+
+This design is just as configurable as its linear-in-period counterpart, and well suited to the low frequency emulation of \ac{ASIC} based oscillators that are themselves linear in frequency.
+In sharing the \ac{FPGA} as a clock source again the pseudo-deterministic characteristics return, once more meaning this oscillator is better used for testing, simulating or verifying other blocks in the system.
+
+\section{Inverter Ring \acs{DCO}}

### 2019-04-10 thesis/Background.tex
wednesday halfway

@@
 
-Both above forms of measurement assume a large number of sequential measurements, however, there are other ways that jitter and skew can be calculated \cite{AN007}.
+Both above forms of measurement assume a large number of sequential measurements, however, there are other ways that jitter and skew can be calculated \cite{AN10007}.
 Phase jitter is an important characteristic for communications systems as it can be used to calculate phase noise, important to ensure spurious emissions are within regulations.

### 2019-04-10 thesis/FPGABasedADPLLs.tex
wednesday halfway

@@
 
-\section{\acs{FPGA} Driven, Linear Period \acs{DCO}}
+\subsection{\acs{FPGA} Driven, Linear Period \acs{DCO}}
 The first design of \ac{DCO} to be examined is of the type used by Zianbetov in his \ac{ADPLL} network test bed and relies on a counter driven by the clock manager on the \ac{FPGA} \cite{zianbetov2013phd}.
@@
 
-\section{\acs{FPGA} Driven, Linear Frequency \acs{DCO}}
+\subsection{\acs{FPGA} Driven, Linear Frequency \acs{DCO}}
 The second \ac{FPGA} clocked oscillator is similar in most attributes to the above design but eschews period linearity for frequency linearity.
@@
 
-\section{Inverter Ring \acs{DCO}}
+\subsection{Ring Oscillator}
+The best approximation of the non-idealities of an \ac{ASIC} \ac{DCO} implementation can be obtained using a \ac{RO}, also know as an inverter ring, as the \ac{DCO}.
+In an \ac{ASIC} an \ac{RO} is constructed by connecting a number of NOT gates back in a chain, with the output of the last NOT gate fed input the input of the first.
+As NOT gates are not available on an \ac{FPGA}, they are replaced by inverter primitives.
+The inverter ring is an inherently unstable circuit that leverages the propagation delay through an odd number of inverters to create a waveform that, when viewed at a fixed position, is a square wave.
+\begin{figure}[h]
+	\centering
+	\includegraphics[scale=0.4]{../RO-wiki}
+	\caption{Basic Ring Oscillator \cite{ro_wiki}.}
+	\label{fig:ro_wiki}
+\end{figure}\\
+The output value of the inverter at the end of the chain changes a non-zero amount of time after the input to the first element is set.
+It can be trivially shown that if the number of inverters is odd, this will be the inverse of the value initially applied to the first element.
+This sets in motion an oscillation, the period of which is governed by the number of inverters in the chain.
+The formula to calculate the oscillator's period is simple:
+\begin{equation}
+	T_{osc} = 2\times \tau_{inv} \times N
+\end{equation}
+Here, $\tau_{inv}$ is the propagation time through one inverter, $N$ is the number of inverters in the chain and the multiple of two arises from the necessity to invert twice to produce a square wave.
+
+The main advantage of this type of oscillator stems from its similarity to that used on an \ac{ASIC} which leads to many non-idealities carrying over from the truly mixed-signal implementation.
+Rather than being just a tool to validate that other modules, the performance of this oscillator can be analysed more deeply as well as the performance of networks containing it.
+Koskin \textit{et al} used such a design to test theories proposed regarding the synchronisation of oscillators during his PhD using this type of oscillator implemented on an \ac{FPGA} \cite{koskin2019phd,theboys2019}.
+
+While it is advantageous for the purposes of analysis that this design is affected by variation in temperature, implementation and other non-idealities this brings with it challenges not present in a true mixed-signal design.
+In a design intended for an \ac{ASIC} the design has near total control over the placement of circuit elements, however, the \ac{FPGA} based counterpart control is available over just the area in which the oscillator may lie, and the \ac{EDA} is responsible for the exact placement.
+Unlike the \ac{ASIC} implementation, in which the propagation delay through each element of the ring is well bounded, the lack of control over placement means that the propagation time through each inverter is not well bounded and may vary wildly.
+This makes it time consuming to ensure that the frequency tuning ranges of each oscillator match that desired, and use of this oscillator less straight-forward than that of the previous two designs.
+
+\section{Feedback Divider}
+Many \ac{PLL}s, be they digital or analogue, incorporate a divider in the feedback path so that comparison can be made with a reference operating at a lower frequency.
+This is a feature that is advantageous for \ac{ADPLL} networks, as comparison using this divided clock allows the synchronisation signal sent between modules to be of a frequency orders of magnitude lower than that of the clock distribution network and thus require circuitry that consumes significantly less energy.
+This division can be obtained trivially on an \ac{FPGA} by a number of means, of which this thesis will mention just two.
+The first of these methods allows for integer division, by inverting a signal each time a counter reaches a certain value, $val$.
+When this condition is true the value stored in the counter is also reset to zero, thus implementing division of the generated clock frequency by $val$.
+
+By limiting to division by powers of two, the divider can be simplified with the removal of any comparison logic.
+One flip-flop and an inverter are required per power of two.
+As in Figure \ref{fig:divider_2}, the inverted $\overline{Q}$ output of each flip-flop is inverted and connected to the $D$ input, with the non-inverted $Q$ of the previous flip-flop acting as the clock.
+The first flip-flop in the chain is clocked by the \ac{DCO} output.
+
+\section{Phase Detector}
+
+\section{Error Combiner}
+In a regular \ac{ADPLL} an error combining circuit is not required as there is never more than one reference signal, and as such the input of this block would be equivalent to the output.
+However in an \ac{ADPLL} network multiple reference signals are required, one for each neighbour, thus requiring some method to combine.
+Simply averaging these signals was not found to be viable by Pratt and Nyugen in their original paper for analogue systems, however, the work of Javidan \textit{et al}, showed that in a digital system this method will not result in mode locking, provided start-up is performed uni-directionally \cite{javidan2011all}.
+
+Accordingly any intended error combiner design implementing an average should also be reconfigurable to satisfy the reconfigurability constraint.
+
+\section{Loop Filter}

### 2019-04-10 thesis/FPGABasedADPLLs.tex
4 svgs

@@
 \subsection{\acs{FPGA} Driven, Linear Period \acs{DCO}}
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=0.6\textwidth]{../osc1}
+	\caption{\acs{FPGA} Driven, Linear Period \acs{DCO} \ac{RTL} Diagram.}
+	\label{fig:osc1}
+\end{figure}
 The first design of \ac{DCO} to be examined is of the type used by Zianbetov in his \ac{ADPLL} network test bed and relies on a counter driven by the clock manager on the \ac{FPGA} \cite{zianbetov2013phd}.
@@
 \subsection{\acs{FPGA} Driven, Linear Frequency \acs{DCO}}
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=0.6\textwidth]{../osc2}
+	\caption{\acs{FPGA} Driven, Linear Period \acs{DCO} \ac{RTL} Diagram.}
+	\label{fig:osc2}
+\end{figure}
 The second \ac{FPGA} clocked oscillator is similar in most attributes to the above design but eschews period linearity for frequency linearity.
@@
 	\centering
-	\includegraphics[scale=0.4]{../RO-wiki}
-	\caption{Basic Ring Oscillator \cite{ro_wiki}.}
+	\includegraphics[width=0.4\textwidth]{../RO-wiki}
+	\caption{Basic Ring Oscillator \ac{RTL} Diagram \cite{ro_wiki}.}
 	\label{fig:ro_wiki}
-\end{figure}\\
+\end{figure}
+
 The output value of the inverter at the end of the chain changes a non-zero amount of time after the input to the first element is set.
@@
 This division can be obtained trivially on an \ac{FPGA} by a number of means, of which this thesis will mention just two.
-The first of these methods allows for integer division, by inverting a signal each time a counter reaches a certain value, $val$.
-When this condition is true the value stored in the counter is also reset to zero, thus implementing division of the generated clock frequency by $val$.
-
+The first of these methods allows for integer division, by inverting a signal each time a counter reaches a certain value, $k$.
+When this condition is true the value stored in the counter is also reset to zero, thus implementing division of the generated clock frequency by $2k$.
+\begin{figure}[h]%
+	\centering
+	\subfloat[Divider 1]{{\includegraphics[width=0.6\textwidth]{../divider1} }}\\
+	\subfloat[Divider 2]{{\includegraphics[width=0.6\textwidth]{../divider2} }}%
+	\caption{Divider \ac{RTL} Diagrams.}
+	\label{fig:divs}
+\end{figure}
 By limiting to division by powers of two, the divider can be simplified with the removal of any comparison logic.

### 2019-04-10 thesis/FPGABasedADPLLs.tex
error combiner

--- 2019-04-10 thesis/FPGABasedADPLLs.tex
@@
 Accordingly any intended error combiner design implementing an average should also be reconfigurable to satisfy the reconfigurability constraint.
+Given this is a digital system the most straightforward way to implement this is as a weighted sum, followed by a division.
+The final \ac{ADPLL} network will be laid out in a Cartesian grid so a maximum of four values will possibly be combined, with only the non-corner edge nodes having a number of neighbours that is not a power of two.
+The simplest, and most processing time efficient, method of calculating the weighted sum restricts the weights to powers of two, as this permits their implementation as a left bit shift which is trivially achieved in hardware by a part select.
+Similarly this reasoning can be expanded to the division, where the complicated division by three for a non-corner edge node is instead replaced by division by four.
+This means the division is implemented identically by all error combiners in the network, and implemented trivially by a right shift of two bits.%TODO ask pierre re 3 div 4
 

### 2019-04-11 thesis/Background.tex
more pngs and svgs

@@
 	\includegraphics[scale=0.4]{../eldar_last_30_yrs}
-	\caption{Frequency of the Intel microprocessors over past 30 years \cite{zianbetov2013phd}.}
+	\caption[Frequency of the Intel microprocessors over past 30 years]{Frequency of the Intel microprocessors over past 30 years \cite{zianbetov2013phd}.}
 	\label{fig:eldar_last_30_yrs}
@@
 	\includegraphics[scale=0.6]{../eldar_why_precise_clocking}
-	\caption{Data Flow in a Clocked System \cite{zianbetov2013phd}.}
+	\caption[Data flow in a clocked system]{Data flow in a clocked system \cite{zianbetov2013phd}.}
 	\label{fig:eldar_why_precise_clocking}
@@
 	\includegraphics[scale=0.33]{../trees}
-	\caption{H and Branch Tree Clock Distribution Systems.}
+	\caption[H and branch tree clock distribution systems]{H and branch tree clock distribution systems.}
 	\label{fig:trees}
@@
 	\includegraphics[scale=0.7]{../tex_files/eldar_mesh}
-	\caption{Mesh Clock Distribution System \cite{zianbetov2013phd}.}
+	\caption[Mesh clock distribution system ]{Mesh clock distribution system \cite{zianbetov2013phd}.}
 	\label{fig:mesh}
@@
 	\includegraphics[scale=0.7]{../mizuno1998noise}
-	\caption{Coupled Oscillator Clock Delivery Circuit \cite{mizuno1998noise}.}
+	\caption[Coupled oscillator clock delivery circuit]{Coupled oscillator clock delivery circuit \cite{mizuno1998noise}.}
 	\label{fig:mizuno1998noise}
@@
 	\includegraphics[scale=0.7]{../gutnik2000active}
-	\caption{PLL Network Topology \cite{gutnik2000active}.}
+	\caption[PLL network topology ]{PLL network topology \cite{gutnik2000active}.}
 	\label{fig:gutnik2000active}
@@
 Analogue designs both deliver continuous control signals to the oscillators and have a continuous phase detection capability, unlike a digital system where these actions are carried out with fixed resolution.
-\begin{figure}[h]
+\begin{figure}[!h]
 	\centering
 	\includegraphics[scale=0.4]{../ccirc_2013_arch}
-	\caption{Architecture of the ADPLL network and of a single node \cite{zianbetov2013distributed}.}
+	\caption[Architecture of the ADPLL network and of a single node]{Architecture of the ADPLL network and of a single node \cite{zianbetov2013distributed}.}
 	\label{fig:eldar_node}
@@
 In their paper, they presented a method where initial start-up was performed uni-directionally and, once all nodes are close to alignment, full connectivity could be restored, however this was not viable in an analogue system as reconfigurability was not an option \cite{pratt1995distributed}.
-In creating an entirely digital system, Zianbetov and Shan could exploit reconfigurability and implement a uni-directional start-up and thus avoid the problem of convergence into a mode locked state, without having to design a non-linear phase detector.
+Javidan \textit{et al} (2011) found that this was in fact the case, and stated:
+\begin{quote}
+	[Mode-locking] is solved in a simple and original way, by a dynamic reconfiguration of the network interconnection topology at the starting stage.
+\end{quote}
+In creating an entirely digital system, Zianbetov and Shan could easily reconfigure the network topology and by implement a uni-directional start-up avoid the problem of convergence into a mode locked state, without having to design a non-linear phase detector.
 
@@
 	\includegraphics[scale=0.5]{../tex_files/mulkeen_pll}
-	\caption{Block Diagram of a \acl{PLL}, \textit{Wireless Systems Notes}, B. Mulkeen (2017).}
+	\caption[Block fiagram of a \acl{PLL}]{Block fiagram of a \acl{PLL}, \textit{Wireless Systems Notes}, B. Mulkeen (2017).}
 	\label{fig:mulkeen_pll}
@@
 	\includegraphics[scale=0.275]{../inverter_chain}
-	\caption{Basic Ring/Inverter Chain Oscillator.}
+	\caption[Basic Ring/Inverter Chain Oscillator]{Basic ring/inverter chain oscillator.}
 	\label{fig:my_ring}
@@
 	\includegraphics[scale=0.35]{../shan_bb_pd}
-	\caption{Bang-bang phase/frequency detector architecture. \cite{shan2014phd}.}
+	\caption[Bang-bang phase/frequency detector architecture]{Bang-bang phase/frequency detector architecture \cite{shan2014phd}.}
 	\label{fig:shan_bb_pd}
@@
 	\centering
-	\includegraphics[scale=0.45]{../pi_simple}
-	\caption{Basic PI Controller Architecture.}
+	\includegraphics[width=0.6\textwidth]{../simple_pi}
+	\caption[Basic \ac{PI} controller architecture]{Basic \ac{PI} controller architecture.}
 	\label{fig:my_simple_pi}

### 2019-04-11 thesis/FPGABasedADPLLs.tex
more pngs and svgs

@@
 However, those that will be used to emulate mixed-signal circuitry, such as the \ac{DCO} and \ac{PFD}, will be examined in greater detail.
+\begin{figure}[h] %TODO redo this image
+	\centering
+	\includegraphics[width=\textwidth]{../network_adpll}
+	\caption{\ac{ADPLL} designed for use in a Cartesian grid network.}
+	\label{fig:dist_adpll}
+\end{figure}
 
@@
 	\includegraphics[width=0.6\textwidth]{../osc1}
-	\caption{\acs{FPGA} Driven, Linear Period \acs{DCO} \ac{RTL} Diagram.}
+	\caption{\acs{FPGA} driven, linear period \acs{DCO} \ac{RTL} Diagram.}
 	\label{fig:osc1}
@@
 	\includegraphics[width=0.6\textwidth]{../osc2}
-	\caption{\acs{FPGA} Driven, Linear Period \acs{DCO} \ac{RTL} Diagram.}
+	\caption{\acs{FPGA} driven, linear frequency \acs{DCO} \ac{RTL} Diagram.}
 	\label{fig:osc2}
@@
 	\includegraphics[width=0.4\textwidth]{../RO-wiki}
-	\caption{Basic Ring Oscillator \ac{RTL} Diagram \cite{ro_wiki}.}
+	\caption[Basic ring oscillator \ac{RTL} diagram]{Basic ring oscillator \ac{RTL} diagram \cite{ro_wiki}.}
 	\label{fig:ro_wiki}
@@
 
-\section{Feedback Divider}
-Many \ac{PLL}s, be they digital or analogue, incorporate a divider in the feedback path so that comparison can be made with a reference operating at a lower frequency.
+\section{Frequency Divider}
+Many \ac{PLL}s, be they digital or analogue, incorporate a frequency divider in the feedback path so that comparison can be made with a reference operating at a lower frequency.
 This is a feature that is advantageous for \ac{ADPLL} networks, as comparison using this divided clock allows the synchronisation signal sent between modules to be of a frequency orders of magnitude lower than that of the clock distribution network and thus require circuitry that consumes significantly less energy.
@@
 	\subfloat[Divider 2]{{\includegraphics[width=0.6\textwidth]{../divider2} }}%
-	\caption{Divider \ac{RTL} Diagrams.}
+	\caption{Frequency divider \ac{RTL} diagrams.}
 	\label{fig:divs}
@@
 
-Accordingly any intended error combiner design implementing an average should also be reconfigurable to satisfy the reconfigurability constraint.
+Accordingly, any intended error combiner design implementing an average should also be reconfigurable, so that it satisfies the reconfigurability constraint.
 Given this is a digital system the most straightforward way to implement this is as a weighted sum, followed by a division.

### 2019-04-11 thesis/FPGABasedADPLLs.tex
loop filters

@@
 However, those that will be used to emulate mixed-signal circuitry, such as the \ac{DCO} and \ac{PFD}, will be examined in greater detail.
-\begin{figure}[h] %TODO redo this image
+A \ac{RTL} diagram of the \ac{ADPLL} topology that will be assumed in the following sections is shown in Figure \ref{fig:network_adpll}.
+\begin{figure}[h]
 	\centering
@@
 	\caption{\ac{ADPLL} designed for use in a Cartesian grid network.}
-	\label{fig:dist_adpll}
+	\label{fig:network_adpll}
 \end{figure}
+
+Important to note is that despite a Cartesian grid leading to four neighbours, only two phase detectors are required per oscillator.
+Borrowing matrix indexing, and choosing the oscillator located at index $i,j$, $i,j \neq 1,N$ where $N$ is the number of oscillators in each direction, comparison need only be done with the signals generated by oscillators located at $i-1,j$ and $i,j-1$.
+The comparison with oscillators $i+1,j$ and $i,j+1$ will be carried out in those \ac{ADPLL}s, and the negation of the errors measured used in the \ac{ADPLL} located at $i,j$.
 
@@
 One flip-flop and an inverter are required per power of two.
-As in Figure \ref{fig:divider_2}, the inverted $\overline{Q}$ output of each flip-flop is inverted and connected to the $D$ input, with the non-inverted $Q$ of the previous flip-flop acting as the clock.
+As in Figure \ref{fig:divs} (b), the inverted $\overline{Q}$ output of each flip-flop is inverted and connected to the $D$ input, with the non-inverted $Q$ of the previous flip-flop acting as the clock.
 The first flip-flop in the chain is clocked by the \ac{DCO} output.
@@
 \section{Loop Filter}
+The \acl{LF} in an \ac{ADPLL} can be implemented as a \ac{PI} controller, a circuit that can be implemented in a wide variety of ways, with the exact implementation having little impact on behaviour.
+The first choice to be made is whether the designer wishes to implement the controller as an \ac{FIR} or \ac{IIR} filter.
+Both methods will produce a controller that fundamentally behaves the same way, but will have a starkly different implementation.
+In both cases the filter can be implemented using either fractional or integer representation of the data, depending on the wishes of the designer.
+Carrying out computations using fractional values may be conceptually easier to understand as the magnitude of individual paths or taps will be independent of another, whereas using integer representations allows for a simpler implementation.
+If fractional representation is used, it will be done with fixed-point arithmetic as the time complexity of floating point calculations is too great and the potential benefits are inconsequential.
+
+\subsection{\ac{FIR} \acl{LF}}
+The \ac{FIR} implementation of the \acl{LF} will take the form presented in Figure \ref{fig:fir_pi} (a), and consists series of time delays, or ``taps'', the output of which is multiplied by a coefficient and the result of each multiplier added together to form the filter output.
+Various tools, Matlab for example, can be used to compute the coefficients and number of taps required to obtain a specific filter response.
+Using an \ac{FIR} filter is advantageous as it is always stable due to the lack of a feedback path, however a large number of taps are required to obtain a frequency response akin to that of the \ac{IIR} implementation, thus requiring an equal number of multipliers and $taps-1$ adders.
+Compared to an \ac{IIR}, which requires just a two adders, two multipliers and a flip-flop this has a much greater hardware footprint.
+It is possible to condense the \ac{FIR} multiplication and addition to a single multiplier and adder, depicted in \ref{fig:fir_pi} (b), however this will take as many clock cycles to complete as there are taps.
+This, however, is not viable in this system as the filter must produce its output in a single clock cycle.
+
+The filter can be implemented using fixed-point arithmetic, in which the designer must ensure alignment of the radix in additions and correct radix position in multiplication results.
+Consequently the fractional width of the each multiplication need not be the same, most commonly used when the proportional gain, the output of the 0th tap multiplication, is significantly larger than that of the other taps.
+Integer arithmetic instead assumes all calculations are carried out using integers, and output of the addition stage is divided by a large number to form the module output.
+The integer approach simplifies the task of the designer and reduces the likelihood of incorrect calculations due to misalignment.
+
+\subsection{\ac{IIR} \acl{LF}}
+A \ac{PI} controller can be easily implemented as an \ac{IIR} filter in hardware, as in \Figure \ref{fig:iir_pi}, requiring significantly fewer circuit elements than an \ac{FIR} based design as the long series of taps and multipliers is not required, as instead integration is performed by a single multiplier and an accumulator.
+As before the frequency response of the controller can be calculated using the following equation, where $\alpha$ is the proportional and $\beta$ the integral gain:
+\begin{equation*}
+	H(z) = \alpha + \beta\frac{1}{1-z^{-1}} = \frac{(\alpha + \beta) - \alpha z^{-1}}{1-z^{-1}}
+\end{equation*}
+
+As the scaling factor used in this multiplier is the $k_i$ gain itself, reconfiguration of the integral path of the \acl{LF} can be carried out solely by varying this gain, as opposed to the \ac{FIR} solution where the value of each tap must be recalculated to change the response of the system.
+The combination of reconfigurability and simpler hardware make the \ac{IIR} the more desirable \ac{LF} design.
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=0.6\textwidth]{../simple_pi}
+	\caption[Basic \ac{IIR} \ac{PI} loop filter]{Basic \ac{IIR} \ac{PI} loop filter.}
+	\label{fig:iir_pi}
+\end{figure}
+
+Again both integer and fixed-point methods of \ac{LF} design are possible, with the integer method requiring logic combining the two paths to preserve the relationship between the proportional and integral gains, which from the work of Koskin \textit{et al} is known to be several binary orders of magnitude \cite{koskin2018generation}.
+Failure to preserve a relationship of this order may lead to the filter becoming unstable, as in the upper-left region of Figure \ref{fig:gain_grid}.
+In the fixed-point system the relationship between $k_p$ and $k_i$ is preserved during the combination of proportional and integral paths as both values have their radix point aligned before addition is carried out.
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=0.6\textwidth]{example-image-a}%{../koskin_gain_grid} %TODO
+	\caption[Stability of \acl{LF} gains]{Stability of \acl{LF} gains \cite{koskin2018generation}.}
+	\label{fig:gain_grid}
+\end{figure}

### 2019-04-11 thesis/FPGABasedADPLLs.tex
bang bang

@@
 \section{Phase Detector}
+Three potential \acl{PD} designs will be addressed in this thesis, sign-only detecion with a Bang-Bang Detector, \ac{FPGA} clocked detection using a \ac{FSM} and an emulation of a \ac{TDL} using inverter primitives to create a delay, similarly to the \ac{RO} seen previously.
+Both designs will have to have a representation for both situations where the generated signal leads and lags the reference signal, and for the magnitude of this difference, a problem most obviously solved by using a signed binary representation of the difference.
+Most easily implemented in hardware is twos' complement, as operations can be carried out by the same hardware as unsigned arithmetic.
+The sign convention for phase error assigns the positive values to the situation where the generated signal lags the reference, and vice versa for the leading case.
+Apart from the Bang-Bang Detector, each design features two blocks, one to determine the sign of the error and a second which estimates the magnitude.
+These blocks could theoretically be mixed-and-matched, but as they operate on different paradigms no benefit would be obtained.
+
+\subsection{Bang-Bang Detector}
+A Bang-Bang Detector is a sign-only detector, thus outputting a phase error just a single bit wide.
+This can be accomplished using a single flip-flop, with the reference signal acting as the clock, in this case using the rising edge, a design verified by Predrag in his 2011 thesis \cite{redrag}.
+The generated signal is applied to the $D$ input, thus when a rising edge is seen on the reference the value of the generated signal is held as the $Q$ output until the next rising edge of the reference.
+If the generated signal has gone high prior to the reference, and leading it, the value of $Q$ will be 1, thereby matching the expected sign behaviour.
+Trivially it can be seen that the case where the generated signal lags the reference the $Q$ will show 0.
+
+This design is simple but effective, and can even be used without a loop filter, albeit with a very limited oscillator tuning range.
+This is especially valuable an attribute at an early stage of platform development, as it may be used in order to test that other aspects of the system work as intended.
+The binary resolution of this detector, however, restricts its usefulness as the \ac{ADPLL} cannot react more quickly to larger phase differences.
+
+\subsection{\acs{FPGA} Clocked Phase Detector}
+\subsubsection{State Machine Based Detection}
+The easier to implement of the two designs uses a combination of a state machine to determine
+\subsubsection{Up-Down Counter}
+\subsection{SigNum Detector}
+\subsubsection{Sign Detection}
+\subsubsection{Tapped Delay Line}
 
@@
 If fractional representation is used, it will be done with fixed-point arithmetic as the time complexity of floating point calculations is too great and the potential benefits are inconsequential.
+The \acl{LF} output may be applied directly to the oscillator, or an intermediate calculation done to set a bias point.
 

### 2019-04-12 thesis/FPGABasedADPLLs.tex
31

@@
 Koskin \textit{et al} used such a design to test theories proposed regarding the synchronisation of oscillators during his PhD using this type of oscillator implemented on an \ac{FPGA} \cite{koskin2019phd,theboys2019}.
+Also in this design's favour is the period step size, four times the inverter propagation delay, which was found experimentally to be than in the region of a nanosecond on Xilinx Artix-7 and Zynq \ac{FPGA}s.
 
@@
 Trivially it can be seen that the case where the generated signal lags the reference the $Q$ will show 0.
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=0.4\textwidth]{../BB}
+	\caption[Bang-bang phase detector \ac{RTL} diagram]{Bang-bang phase detector \ac{RTL} diagram.}
+	\label{fig:bang_bang}
+\end{figure}
 
@@
 \subsection{\acs{FPGA} Clocked Phase Detector}
+The easier to implement of the two designs uses a combination of a state machine to determine which edge occurred first, and to control a twos' complement up-down counter the estimates the magnitude of the error.
+
 \subsubsection{State Machine Based Detection}
-The easier to implement of the two designs uses a combination of a state machine to determine
 \subsubsection{Up-Down Counter}
+The up-down counter in this design performs estimates the magnitude of the phase difference between the reference and generated signals.
+The state machine provides control instructions, indicating whether the count should be in the increasing or decreasing directions, and and holds the counter in reset once the measurement interval has been completed.
+The counter counts in twos' complement, so counting in the downward direction can be carried out by adding on the representation of -1, the binary signal containing all 1s.
+The other advantage of counting in two's complement is that count value is already in the correct format for the follow block, the phase detector in a regular \ac{ADPLL} or the error combiner in this instance, therefore the count value at the end of the measurement interval can be stored in a register, which maintains the output value of the phase detector between measurements.
+Twos' complement is an advantageous way to represent signed numbers in hardware as the mathematical operations are identical to their unsigned counterparts, thus allowing use of any adder or carry hardware that may be present on the \ac{FPGA}.
+
+This counter is best suited for use with the aforementioned \ac{FSM} sign detection circuitry, and would see no benefit from a combination with the un-clocked sign detection circuit that will be presented in the following section.
+This is the case as if quantisation to the \ac{FPGA} clock had not already occurred in the sign detection circuit it would occur in this counter, as increments are only possible at \ac{FPGA} clocking events, thereby enforcing quantisation to the \ac{FPGA} clock.
+The coarseness of this measurement, restricted to time delay increments equal to the \ac{FPGA} clock period means that it is an ideal detection method for the clocked oscillators, as the minimum phase difference that can be detected matches exactly with the minimum step by which the oscillator period can be tuned.
+This design can, however, be used with the \acl{RO}, provided the \acl{LF} gains are adjusted to maintain the oscillator to detector step relationship.
+
 \subsection{SigNum Detector}
@@
 This, however, is not viable in this system as the filter must produce its output in a single clock cycle.
+In the case of any coefficient symmetry, addition can be performed prior to multiplication, thus reducing the number of multipliers for each identical coefficient.
+\begin{figure}[h]
+	\centering
+	\subfloat[Typical FIR Filter]{{\includegraphics[width=0.6\textwidth]{../FIR_1} }}\\
+	\subfloat[Single Multiplier FIR Filter]{{\includegraphics[width=0.6\textwidth]{../FIR_2} }}%
+	\caption[\ac{FIR} \ac{PI} loop filter RTL diagrams]{\ac{FIR} \ac{PI} loop filter RTL diagrams.}
+	\label{fig:fir_pi}
+\end{figure}
 
@@
 \subsection{\ac{IIR} \acl{LF}}
-A \ac{PI} controller can be easily implemented as an \ac{IIR} filter in hardware, as in \Figure \ref{fig:iir_pi}, requiring significantly fewer circuit elements than an \ac{FIR} based design as the long series of taps and multipliers is not required, as instead integration is performed by a single multiplier and an accumulator.
+A \ac{PI} controller can be easily implemented as an \ac{IIR} filter in hardware, as in Figure \ref{fig:iir_pi}, requiring significantly fewer circuit elements than an \ac{FIR} based design as the long series of taps and multipliers is not required, as instead integration is performed by a single multiplier and an accumulator.
 As before the frequency response of the controller can be calculated using the following equation, where $\alpha$ is the proportional and $\beta$ the integral gain:
@@
 	\centering
-	\includegraphics[width=0.6\textwidth]{../simple_pi}
+	\includegraphics[width=0.6\textwidth]{../simple_pi.pdf}
 	\caption[Basic \ac{IIR} \ac{PI} loop filter]{Basic \ac{IIR} \ac{PI} loop filter.}
@@
 	\centering
-	\includegraphics[width=0.6\textwidth]{example-image-a}%{../koskin_gain_grid} %TODO
+	\includegraphics[width=0.6\textwidth]{../eugene}%{../koskin_gain_grid} %TODO
 	\caption[Stability of \acl{LF} gains]{Stability of \acl{LF} gains \cite{koskin2018generation}.}

### 2019-04-12 thesis/FPGABasedADPLLs.tex
just TDL left

@@
 \section{Phase Detector}
-Three potential \acl{PD} designs will be addressed in this thesis, sign-only detecion with a Bang-Bang Detector, \ac{FPGA} clocked detection using a \ac{FSM} and an emulation of a \ac{TDL} using inverter primitives to create a delay, similarly to the \ac{RO} seen previously.
+Three potential \acl{PD} designs will be addressed in this thesis, sign-only detection with a Bang-Bang Detector, \ac{FPGA} clocked detection using a \ac{FSM} and an emulation of a \ac{TDL} using inverter primitives to create a delay, similarly to the \ac{RO} seen previously.
 Both designs will have to have a representation for both situations where the generated signal leads and lags the reference signal, and for the magnitude of this difference, a problem most obviously solved by using a signed binary representation of the difference.
-Most easily implemented in hardware is twos' complement, as operations can be carried out by the same hardware as unsigned arithmetic.
+Most easily implemented in hardware is two's complement, as operations can be carried out by the same hardware as unsigned arithmetic.
 The sign convention for phase error assigns the positive values to the situation where the generated signal lags the reference, and vice versa for the leading case.
@@
 \subsection{\acs{FPGA} Clocked Phase Detector}
-The easier to implement of the two designs uses a combination of a state machine to determine which edge occurred first, and to control a twos' complement up-down counter the estimates the magnitude of the error.
+The easier to implement of the two designs uses a combination of a state machine to determine which edge occurred first, and to control a two's complement up-down counter, which acts as the \ac{TDC} in this detector, and estimates the magnitude of the error.
 
 \subsubsection{State Machine Based Detection}
+While sign detection clocked by the \ac{FPGA} can be performed in a variety of ways, a state machine is a straightforward way in which to accomplish this.
+A \acl{FSM} consists of three main components: the states themselves, the next state logic and the output logic.
+In this case the next state logic is the generated and reference signals, and the output logic sets the count instructions for the up-down counter as well as ensuring the phase error output is held constant between measurement periods.
+The \ac{FSM} will have a state representing the case where either the reference or generated signals occurred first, in addition to a wait state used between the end of one measurement interval and the commencement of the next.
+Two types of \ac{FSM} exist, Moore and Mealy Machines, the difference being that Moore Machines base their output only on the current state while Mealy Machines do so on both the current state and the inputs to the system.
+If a Moore Machines is used an extra state is required in order to hold the value of the phase error at the end of the measurement interval.
+
+As this design is clocked by the \ac{FPGA}, the phase error is quantised to the the \ac{FPGA} clock periods, thus setting the resolution of the design.
+While, depending on the oscillator design, the generated clock may be synchronous to the \ac{FPGA}, the reference signal arrives from off-chip and thus will not be synchronised to the measurement clock.
+In order to avoid any metastability events arising from an input signal changing within the set-up and hold times of the registers sampling them, synchroniser circuits are used to ensure the value sampled is either logic high or low at the instance of the measurement.
+As these synchronisers are the first \ac{FPGA} clocked elements in the measurement chain, they also act as the system's quantisers.
+There is no degradation in detection resolution, as the signals would be quantised to the \ac{FPGA} clock anyway.
+
+There are three situations that may occur while the \ac{FSM} is in the waiting state.
+Firstly the case where the reference signal sees a rising edge before the generated clock, and thus the generated clock is said to be lagging the reference.
+As previously mentioned the convention dictates that the phase difference have a positive sign, so the counter is incremented to count in the positive direction.
+Similarly if the generated signal is leading the reference the sign of the error is negative and thus the counter is instructed to count downwards.
+The third potential scenario occurs when both clock signals have a rising edge within the same period of the \ac{FPGA} clock, which due to quantisation will be interpreted as both signals occurring at exactly the same time.
+In this case the counter receives no instruction and a zero value is stored until the next measurement interval.
+Once the \ac{FSM} is in a measurement state, and the counter either incrementing or decrementing, a rising edge occurring on the opposite signal to that which initiated the measurement will end the measurement interval.
+
+%TODO state trans diagram
+
 \subsubsection{Up-Down Counter}
@@
 The state machine provides control instructions, indicating whether the count should be in the increasing or decreasing directions, and and holds the counter in reset once the measurement interval has been completed.
-The counter counts in twos' complement, so counting in the downward direction can be carried out by adding on the representation of -1, the binary signal containing all 1s.
+The counter counts in two's complement, so counting in the downward direction can be carried out by adding on the representation of -1, the binary signal containing all 1s.
 The other advantage of counting in two's complement is that count value is already in the correct format for the follow block, the phase detector in a regular \ac{ADPLL} or the error combiner in this instance, therefore the count value at the end of the measurement interval can be stored in a register, which maintains the output value of the phase detector between measurements.
-Twos' complement is an advantageous way to represent signed numbers in hardware as the mathematical operations are identical to their unsigned counterparts, thus allowing use of any adder or carry hardware that may be present on the \ac{FPGA}.
+Two's complement is an advantageous way to represent signed numbers in hardware as the mathematical operations are identical to their unsigned counterparts, thus allowing use of any adder or carry hardware that may be present on the \ac{FPGA}.
+An important attribute of this counter is the ability to freeze the count at the minimum and maximum count values in order to avoid a situation where the counter overflows, thus at high phase offsets potentially alternating between positive and negative differences which may restrict the locking range of the \ac{ADPLL}.
+Disallowing overflow also aligns the behaviour of this phase detector with that of those implementing delay lines, allowing better interchangeability and comparison, which for a testing platform may be valuable.
 
@@
 \subsection{SigNum Detector}
+The SigNum detector is based on the methodology used in the Bang-Bang \acl{PD} seen previously, but rather than just performing sign detection, the magnitude of the phase difference can also be measured.
+In the case of the \ac{FPGA} clocked design the counter handled the representation of both sign and magnitude, but as the name SigNum (Sign Number) suggested both are handled separately in this design.
+The modified Bang-Bang detector establishes which signal occurred first and generates a control signal that is analogous to the time delay between a rising edge on each input signal.
+A \ac{TDL} then uses this control signal to convert the phase difference between the signals to a digital signal.
+This magnitude is combined with the sign detected by the bang-Bang detector to form a two's complement error signal.
+
+This method of phase detection is advantageous as similarly to the \acl{RO}, the \ac{TDL} can be implemented using inverters, and therefore is not synchronised to the \ac{FPGA} clock and the propagation time through inverters sets the resolution, thus allowing a smaller step than that of the clocked designs.
+The disadvantages, however, are also consistent with the \ac{RO} in that the propagation time will vary significantly depending on the \ac{EDA}'s placement of the circuit.
+This is not as severe a limitation for the \ac{TDL} as the \ac{RO} as while the tuning range is affected by this variation it cannot cause the centre point to vary, whereas the impact of placement on the \ac{RO} caused both variation in range and centre frequency.
+
 \subsubsection{Sign Detection}
+The Bang-Bang detector structure changes significantly in order to provide the control signals required for the \ac{TDC}.
+The \ac{TDC} expects a long pulse of logic high for duration of the measurement interval, going low once the interval has ended.
+The original single flip-flop design maintains a value between rising edges of the reference clock and could not generate the required pulse.
+Instead two flip-flops are used, each clocked on an input signal with their $D$ inputs connected to logic high.
+The $Q$ outputs of these registers are connected to an \acs{XOR} gate which generates a pulse as long as a rising edge has only been seen on one signal.
+Once a second rising edge occurs the \acs{XOR} output will be set to 0, thus terminating the pulse.
+On completion of the measurement interval, the flip-flops must be reset in preparation for the next measurement cycle, and this can be accomplished by ANDing together the $Q$ outputs of each signal, producing a \texttt{clear} signal that will go high when both flip-flops have a logic high output.
+
+To determine the sign itself, which clock edge first had a rising edge must be recorded, and this can be done by setting a \ac{SR} latch if the rising edge of the generated signal occurs first and resetting in the case of the reference clock having the first rising edge.
+Looking at the whole measurement cycle however, it is apparent that all 4 combinations possible of two inputs occur, putting the \ac{SR} latch into which the output is undefined.
+Careful design of the system can avoid problems arising from this.
+Examining the truth table in Figure \ref{fig:sr} (a), the combination to be avoided is the presence of a 1 on both outputs, precisely the situation seen at the end of the measurement interval.
+In order to avoid the undefined state occurring at the end of the measurement interval, both inputs and outputs to the \ac{SR} latch can be inverted using a NOT gate.
+Figure \ref{fig:sr} (b) illustrates how the $\overline{Q}$ output now behaves as $Q$ did previously, however the end of measurement state where both input registers have been triggered now causes the \ac{SR} latch to enter the hold state where it maintains the previous values.
+The sign, $\overline{Q}$ of the \ac{SR}, can then be loaded into a register using the \texttt{clear} signal generated previously, to maintain it until the next measurement completes.
+
+This, however, ignores the potential for metastability which may arise if both signals have a rising edge within an extremely short period of time.
+While it may in other systems be feasible to dismiss this occurrence as a low probability event, the goal of a \ac{PLL} is to reduce the time delay between rising edges to zero which increases the possibility of this occurring, which may lead to indeterminate behaviour in other parts of the feedback network.
+In order to avoid this in an \ac{ASIC}, an arbitration circuit is placed between the \ac{SR} latch and the register storing the sign between measurements, which will default to one of the two signals in this case.
+This is however a mixed signal circuit, such as that proposed by Tierno \textit{et al} (2008) and implemented by Zianbetov and Shan \cite{tierno2008wide,zianbetov2013phd,shan2014phd}, which requires two \ac{PMOS} and two \ac{NMOS} transistors and cannot be implemented on an \ac{FPGA}.
+As the output of the \ac{SR} latch will be indeterminate prior to the detection of the first edge, the impact of both edge detection registers' output is that the \ac{SR} latch attempts to hold this indeterminate value as an output.
+To avoid this unwanted behaviour the arbiter present on an \ac{ASIC} is replaced by a second \ac{SR} latch, connected as in Figure \ref{fig:arbitration}.
+This ensures the value presented to the register storing the sign is always determinate.
+%TODO sr truth table
+%TODO image of simplified detector
+
 \subsubsection{Tapped Delay Line}
+To complement the asynchronous sign detection circuit, an asynchronous \ac{TDC} is also required, as otherwise there would be no benefit to using the above method over the more easily understood \ac{FSM} based design as quantisation to the clock period would happen in the \ac{TDC} rather than the sign detection circuitry.
+The standard method for implementing a \ac{TDC} on an \ac{ASIC} is through a tapped delay line, where a signal is propagated through a chain of fixed length delays, and at the end of the measurement interval it is recorded whether or not the signal has propagated to each point in the chain.
+As with the mixed signal circuitry previously addressed, a \ac{TDL} cannot be exactly replicated on an \ac{FPGA}, but through the use of inverters it is possible to replace the unrealisable fixed delays.
+
+%TODO high level TDL image (using Z-1s)
+%TODO eugene re: why no zero output
 
@@
 	\centering
-	\includegraphics[width=0.6\textwidth]{../eugene}%{../koskin_gain_grid} %TODO
+	\includegraphics[width=0.6\textwidth]{../eugene}
 	\caption[Stability of \acl{LF} gains]{Stability of \acl{LF} gains \cite{koskin2018generation}.}

### 2019-04-12 thesis/FPGABasedADPLLs.tex
tdll

@@
 The standard method for implementing a \ac{TDC} on an \ac{ASIC} is through a tapped delay line, where a signal is propagated through a chain of fixed length delays, and at the end of the measurement interval it is recorded whether or not the signal has propagated to each point in the chain.
-As with the mixed signal circuitry previously addressed, a \ac{TDL} cannot be exactly replicated on an \ac{FPGA}, but through the use of inverters it is possible to replace the unrealisable fixed delays.
+This architecture is illustrated in Figure \ref{fig:tdl}.
+As with the mixed signal circuitry previously addressed, a \ac{TDL} cannot be exactly replicated on an \ac{FPGA}, but through the use of inverters it is possible to replace the unrealisable well bounded delays.
+As with the \ac{RO}, the delay though the inverters is not nearly as well bounded, and significant variation between delays, in both individual \ac{TDL}s and within the same \ac{TDL}, due to the implementation occurs This variation notwithstanding, the inverter pairs are a suitable replacement, and this method of time-to-digital conversion provides both resolution advantages and the presence of implementation based variation over a clocked design.
+These prove particularly beneficial when attempting to verify theoretical results.
 
+Similarly to the sign at the output of the detection circuit, flip-flops store the value of each tap between phase error measurements.
+Prior to storage, the tap values must be converted to two's complement.
+As the signal propagates through the taps, magnitude of the phase error is determined by how many taps the signal has passed through.
+Conceptually, this thermometer coding is first converted to an unsigned integer before conversion to a signed quantity based on the sign of the error.
 %TODO high level TDL image (using Z-1s)

### 2019-04-13 thesis/FPGABasedADPLLs.tex
end of part2

@@
 This makes it time consuming to ensure that the frequency tuning ranges of each oscillator match that desired, and use of this oscillator less straight-forward than that of the previous two designs.
+This is compounded by the simulation tools in \ac{EDA}s intended for \ac{FPGA}s being unable to accurately simulate the time delays, thus negating any benefit of test benches for the alignment process.
 
@@
 In this case the counter receives no instruction and a zero value is stored until the next measurement interval.
-Once the \ac{FSM} is in a measurement state, and the counter either incrementing or decrementing, a rising edge occurring on the opposite signal to that which initiated the measurement will end the measurement interval.
-
-%TODO state trans diagram
+Once the \ac{FSM} is in a measurement state, and the counter either incrementing or decrementing, a rising edge occurring on the opposite signal to that which initiated the measurement will end the interval.
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=0.8\textwidth]{../state_trans_new}
+	\caption[Example State Transition Diagram for a Moore Machine]{Example State Transition Diagram for a Moore Machine.}
+	\label{fig:state_trans}
+\end{figure}
 
@@
 Careful design of the system can avoid problems arising from this.
-Examining the truth table in Figure \ref{fig:sr} (a), the combination to be avoided is the presence of a 1 on both outputs, precisely the situation seen at the end of the measurement interval.
+Examining the truth table in Table \ref{table:sr_tt}, the combination to be avoided is the presence of a 1 on both outputs, precisely the situation seen at the end of the measurement interval.
 In order to avoid the undefined state occurring at the end of the measurement interval, both inputs and outputs to the \ac{SR} latch can be inverted using a NOT gate.
-Figure \ref{fig:sr} (b) illustrates how the $\overline{Q}$ output now behaves as $Q$ did previously, however the end of measurement state where both input registers have been triggered now causes the \ac{SR} latch to enter the hold state where it maintains the previous values.
+Table \ref{table:sr_tt} also illustrates how the $\overline{Q}$ output now behaves as $Q$ did previously, however the end of measurement state where both input registers have been triggered now causes the \ac{SR} latch to enter the hold state where it maintains the previous values.
 The sign, $\overline{Q}$ of the \ac{SR}, can then be loaded into a register using the \texttt{clear} signal generated previously, to maintain it until the next measurement completes.
+\begin{table}[ht]
+	\begin{center}
+		\begin{tabular}{cc|cc}           
+			$S$&$R$&$Q$&$\overline{Q}$\\
+			\hline
+			0&0&$Q_{n-1}$&$\overline{Q}_{n-1}$\T\\
+			0&1&0&1\T\\
+			1&0&1&0\T\\
+			1&1&\multicolumn{2}{c}{undefined}\T\\					
+		\end{tabular}
+		\caption{SR Latch Truth Table}
+		\label{table:sr_tt}
+	\end{center}
+\end{table}
 
@@
 This ensures the value presented to the register storing the sign is always determinate.
-%TODO sr truth table
-%TODO image of simplified detector
+\begin{table}[!ht]
+	\begin{center}
+		\begin{tabular}{l|cc|cc|c|cc|c}           
+			When&$Gen$&$Ref$&$\overline{Gen}/S^a$&$\overline{Ref}/R^a$&$Q^a$&$S^b$&$R^b$&$Sign/Q^b$\\
+			\hline
+			Between Measurements&0&0&1&1&\multicolumn{1}{c|}{undefined}&\multicolumn{2}{c|}{01 or 10}&\multicolumn{1}{c}{0 or 1}\T\\
+			Ref. First&0&1&1&0&1&0&1&0\T\\
+			Gen. First&1&0&0&1&0&1&0&1\T\\	
+			Measurement Complete&1&1&0&0&$Q^a_{n-1}$&$\overline{Q}^a_{n-1}$&$Q^a_{n-1}$&$\overline{Q}^a_{n-1}$\T\\					
+		\end{tabular}
+		\caption{Sign Detector Truth Table}
+		\label{table:sign_tt}
+	\end{center}
+\end{table}
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=0.8\textwidth]{../simple_sign_detection}
+	\caption[Example Sign Detector]{Example Sign Detector.}
+	\label{fig:arbitration}
+\end{figure}
 
@@
 The standard method for implementing a \ac{TDC} on an \ac{ASIC} is through a tapped delay line, where a signal is propagated through a chain of fixed length delays, and at the end of the measurement interval it is recorded whether or not the signal has propagated to each point in the chain.
-This architecture is illustrated in Figure \ref{fig:tdl}.
+This architecture is illustrated in Figure \ref{fig:simple_tdc}.
 As with the mixed signal circuitry previously addressed, a \ac{TDL} cannot be exactly replicated on an \ac{FPGA}, but through the use of inverters it is possible to replace the unrealisable well bounded delays.
@@
 Conceptually, this thermometer coding is first converted to an unsigned integer before conversion to a signed quantity based on the sign of the error.
-%TODO high level TDL image (using Z-1s)
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=0.8\textwidth]{../simple_tdc}
+	\caption[Example \acl{TDL} Structure]{Example \acl{TDL} Structure.}
+	\label{fig:simple_tdc}
+\end{figure}
 %TODO eugene re: why no zero output

### 2019-04-13 thesis/FPGABasedADPLLs.tex
39 pages, onto chapter 4 of 6

@@
 	\subfloat[Divider 2]{{\includegraphics[width=0.6\textwidth]{../divider2} }}%
-	\caption{Frequency divider \ac{RTL} diagrams.}
+	\caption[Frequency Divider \ac{RTL} Diagrams]{Frequency Divider \ac{RTL} Diagrams.}
 	\label{fig:divs}
@@
 		\end{tabular}
-		\caption{SR Latch Truth Table}
+		\caption{SR latch truth table}
 		\label{table:sr_tt}
@@
 	\begin{center}
+		\setlength{\tabcolsep}{.5\tabcolsep}
 		\begin{tabular}{l|cc|cc|c|cc|c}           
@@
 		\end{tabular}
-		\caption{Sign Detector Truth Table}
+		\caption{Sign detector truth table}
 		\label{table:sign_tt}
@@
 
-As the scaling factor used in this multiplier is the $k_i$ gain itself, reconfiguration of the integral path of the \acl{LF} can be carried out solely by varying this gain, as opposed to the \ac{FIR} solution where the value of each tap must be recalculated to change the response of the system.
+As the scaling factor used in this multiplier is the \ac{ki} itself, reconfiguration of the integral path of the \acl{LF} can be carried out solely by varying this gain, as opposed to the \ac{FIR} solution where the value of each tap must be recalculated to change the response of the system.
 The combination of reconfigurability and simpler hardware make the \ac{IIR} the more desirable \ac{LF} design.
@@
 Again both integer and fixed-point methods of \ac{LF} design are possible, with the integer method requiring logic combining the two paths to preserve the relationship between the proportional and integral gains, which from the work of Koskin \textit{et al} is known to be several binary orders of magnitude \cite{koskin2018generation}.
-Failure to preserve a relationship of this order may lead to the filter becoming unstable, as in the upper-left region of Figure \ref{fig:gain_grid}.
-In the fixed-point system the relationship between $k_p$ and $k_i$ is preserved during the combination of proportional and integral paths as both values have their radix point aligned before addition is carried out.
+Failure to preserve a relationship of this order may lead to the \ac{ADPLL} becoming unstable, as in the upper-left region of Figure \ref{fig:gain_grid}.
+In the fixed-point system the relationship between \ac{kp} and \ac{ki} is preserved during the combination of proportional and integral paths as both values have their radix point aligned before addition is carried out.
 \begin{figure}[h]

### 2019-04-13 thesis/NetworkImplemenation.tex
39 pages, onto chapter 4 of 6

\section{Chapter Overview}
This chapter will cover the network implemented in the course of the project, beginning with the individual \ac{ADPLL}'s composition. A number of attributes for each design will be used to analyse and compare their performance. The impact of some minor architectural changes will also be covered, before the implementation of test network will be described. This chapter will also highlight some pitfalls encountered in the process. The \ac{HDL} Verilog was selected for use in this project due to my prior exposure to that language.

\section{\acs{ADPLL} Architectures}
Three different architectures of \ac{ADPLL} were implemented in this project for the purposes of testing, and re-use in the future within the research team in \acl{UCD}. Between each design a number of blocks remain constant, as only a single instance of that block was implemented, as variation its design would have negligible impact on the final system. These blocks were the error combiner and frequency divider. The loop filter was implemented both using fractional and integer arithmetic, but as the exact same calculation was performed in both cases there was no impact on the output of the module.

The three architectures chosen for implementation represent a progression from entirely an \ac{FPGA} clock driven system to an \ac{ADPLL} with all components asynchronous with respect to that clock. Therefore the first \ac{ADPLL} design features a clocked oscillator and phase detector, specifically the linear frequency design mentioned in Chapter \ref{chap:3} and the clocked phase detector using a state machine and a bi-directional counter. Skipping over Design 2, the final design used was asynchronous in totality, featuring the \acl{RO} and SigNum/\ac{TDL} phase detector. Design 2 sat between these two, using the \ac{RO} as its oscillator, but retained the clocked phase detector seen in Design 1.

The extensibility of the end product of this project was to the forefront during the design process. Whilst, apart from frequency, there were no guidelines as to what the system should resemble and a number of attributes could be chosen arbitrarily, a future user of this platform could have entirely different requirements or specifications. As such each block was implemented in such a way that changing the number of bits used for a certain signal, the target frequency or swapping between different blocks could be done with the sole requirement of changing Verilog \texttt{localparam}\footnote{A \texttt{localparam} is a constant that cannot be modified in a module instance statement \cite{hdlworks}}s in the module \footnote{Module is the Verilog term for a number of logic elements grouped to provide a certain functionality} in which change was made, which would then propagate to any sub-modules affected. Starting the design process with this in mind was a major benefit at later stages when frequent modifications were being made to test their impact, or when changing between \ac{ADPLL} designs.

\subsection{Generic Components}
As mentioned in the chapter overview, a number of components were used in all \ac{ADPLL}s implemented, as changes in their design would not impact the overall system.
\subsubsection{Error Combiner}
The error combiner in this design had a number of requirements in order to be suitable for use in both a single \ac{ADPLL} as well as a network operating in both uni- and bi-directional modes, alongside modifications which would change the widths of the error signals requiring averaging. The module was implemented on the basis that the \ac{LF} input width would be identical to the phase detector output width.

The error combiner was designed to take in up to four error signals, all of an identical width which defaults to eight, and four weights to apply in the summation process. A weighted summation is performed by multiplying each error signal by the weight and adding the results together. Regardless of the number of input error signals, a division by four is then performed in order to compute the average error, done by shifting the value right by two. In a \ac{HDL}, unlike C programming or similar, a shift requires no computation and is instead implemented by performing a part-selection. As all error signals are two's complemented signed integers, the error combination process is also carried out using signed values. The code implementing the error combiner can be found in \textsc{ErrorCombiner.v}, attached in Appendix \ref{adx:code} Listing \ref{lst:error_combiner}.

\subsubsection{Frequency Divider}
The frequency divider is implemented using ``Divider 2'' from Chapter \ref{chap:3}, as only divisions ratios of 2, 4 and 8 were selected for use in the testing process. For ease of testing, the divider had outputs representing each of these division ratios simultaneously, which could be multiplexed between at runtime if so desired. A fourth output passed the input through unmodified, allowing for the divider to be removed without regeneration of the \ac{FPGA} configuration.
\begin{figure}[h]%
	\centering
	\includegraphics[width=0.6\textwidth]{../divider2}
	\caption[Frequency divider \ac{RTL} diagram]{Frequency divider \ac{RTL} diagram.}
	\label{fig:divs_iompl}
\end{figure}

\subsubsection{Loop Filter}
Two \acl{LF}s were implemented, both as \ac{IIR} filters. \ac{FIR} was dismissed due to the extra hardware required to implement the calculations within a single clock cycle, as the \ac{LF} is clocked using the oscillator output, and the greater difficulty of gain adjustment compared to an \ac{IIR}. Clocking on using the generated signal ensures that the discrete time integration is only carried out once per phase comparison. A pitfall that may be encountered implementing an \ac{ADPLL} featuring a divider is not clocking the module on the divided clock, which will result in the integration being carried out at the oscillator output frequency.

Each of integer and fixed-point arithmetic were used to implement a filter respectively, however both designs are interchangeable as they perform the integration identically, given the same proportional and integral gains. All testing was carried out using integer arithmetic, as in Figure \ref{fig:integer_lf}, however the interchangeability will be confirmed later in this chapter, in Section \ref{section:minor_variations}. Additionally the \ac{LF} supports variation of both \acs{ki} and \acs{kp} at runtime, however, if a static value is desired the runtime variation may be disabled using a \texttt{parameter}\footnote{Parameters are constant values that may be changed at compile time, or in the module instance statement \cite{hdlworks2}}.
\begin{figure}[h]%
	\centering
	\includegraphics[width=0.8\textwidth]{../integer_lf} 
	\caption[\acl{LF} implemented using integer arithmetic \ac{RTL} diagram]{\acl{LF} implemented using integer arithmetic \ac{RTL} diagram (signal widths using default values).}
	\label{fig:integer_lf}
\end{figure}

Figure \ref{fig:integer_lf} contains an \ac{RTL} diagram describing the implementation. Important to note is the shift applied to the result of the multiplication by \acs{kp} which, as the integration is performed with a larger width to avoid accumulator overflow, preserves the intended relationship between proportional and integral gains.
The Verilog implementation of the \ac{LF} using integer arithmetic can be found in \textsc{LoopFilter.v}, attached in Appendix \ref{adx:code} Listing \ref{lst:loop_filter}.

\subsection{\acs{ADPLL} Design 1}
\subsubsection{Oscillator}
\subsubsection{Phase Detector}

\subsection{\acs{ADPLL} Design 2}
\subsubsection{Oscillator}
\subsubsection{Phase Detector}

\subsection{\acs{ADPLL} Design 3}
\subsubsection{Oscillator}
\subsubsection{Phase Detector}

\section{\acs{ADPLL} Characterisation}

\section{Minor Variations}\label{section:minor_variations}

\section{\acs{ADPLL} Network Implementation}
\subsection{2x2 Network}
\subsection{3x3 Network}

### 2019-04-15 thesis/NetworkImplemenation.tex
44.5

@@
 The \ac{HDL} Verilog was selected for use in this project due to my prior exposure to that language.
+5 MHz was chosen as the target centre frequency for these \ac{ADPLL}s, during initial testing it was discovered that the waveform output via the Digilent Nexys4 development board's \ac{GPIO} ports began to degrade around the 10 MHz mark, preventing accurate analysis.
 
@@
 \subsection{\acs{ADPLL} Design 1}
-\subsubsection{Oscillator}
+\acs{ADPLL} Design 1 is comprised of the aforementioned generic components and \ac{FPGA} clocked implementations of the \ac{PFD} and \ac{DCO}.
+This was the first solution addressed in the course of this project as it is the simplest to both implement and test, as all aspects of the \ac{ADPLL} are driven by the \ac{FPGA} clock.
+Compared to later designs utilising the \ac{RO}, Vivado simulations could be carried out, with accurate timing behaviour, depicting the locking process.
+The \ac{DCO} was chosen to be linear in frequency in order to obtain a square wave output as it was not yet known whether the varying pulse width of a period linear design would cause a problem in modules which, at that stage in the project, had not yet been designed.
+Indeed, a later discarded version of the \ac{FPGA} clocked \ac{PFD} relied on equal low and high times.
+The \ac{PFD} itself was implemented using a Moore type \ac{FSM} driving a up-down counter to convert the time between rising edges to a digital signal.
+
+\subsubsection{\acl{DCO}}
+As the \ac{FPGA} clocked, linear frequency oscillator is based on the use of counters, the first stage in its design was the decisions as to their required widths.
+The equation given previously is used here, with the control code set to zero in order to obtain the centre frequency, and 5 MHz as the target \ac{DCO} frequency:
+\begin{align}
+f_{osc} &= f_{FPGA}\times\frac{BIAS+CC}{2^{width}}\\
+5\times 10^6 &= f_{FPGA}\times\frac{BIAS}{2^{width}}
+\end{align}
+This still leaves three parameters undecided: The \ac{FPGA} clock, the bit width of the counter and the bias point.
+As the \ac{FPGA} clock determines the frequency step of the design, a test counter was implemented initially to determine the maximum frequency at which the timing violations would not occur, at this was determined to be 275 MHz.
+The equation can then be updated, giving the relationship between bias point and counter width:
+\begin{align}
+5\times 10^6 &= f_{FPGA}\times\frac{BIAS}{2^{width}}\\
+\frac{5}{275} &= \frac{BIAS}{2^{width}}
+\end{align}
+The maximum value of the counter must also be large enough such that with a bias point that allows for a reasonable range of frequency steps to be added or removed. 
+This oscillator was originally designed to allow for 64 control codes either side of the bias, which called for a bias point of 65 or greater.
+\begin{align}
+\frac{5}{275} &= \frac{65}{2^{width}}\\
+2^{width} &= \frac{65\cdot275}{5} = 3575 \\
+width &= \left \lceil{\log_2 3575}\right \rceil = 12
+\end{align}
+From the above equation, the smallest counter that can satisfy this constraint is 12 bits wide, however when testing was carried out using a 275 MHz clock timing violations were discovered by Vivado, resulting \ac{FPGA} clock speed reduction to 258 MHz in order to resolve these violations.
+The correct bias point could then be calculated as $80$:
+\begin{align}
+\frac{5}{258} &= \frac{BIAS}{2^{12}}\\
+ BIAS &= \left \lceil{\frac{5\cdot 2^{12}}{258}}\right \rceil = \left \lceil{\frac{20480}{258}}\right \rceil = 80
+\end{align}
+The frequency step can now be computed as all parameters have been chosen:
+\begin{equation}
+f_{step} = \frac{f_{FPGA}}{2^{12}} = 62.988~\si{\kilo\hertz}
+\end{equation}
+
+Figure \ref{fig:osc2_impl} contains an \ac{RTL} diagram of this oscillator's implementation.
+As phase error is a two's complement signed value, its addition to the bias must be carried out using signed arithmetic.
+Provided the bias is greater than the minimum possible phase error, as has been ensured in this case, the result of this addition is positive signed integer.
+Being a positive quantity, additional logic performing a conversion from a signed to an unsigned representation is required.
+\begin{figure}[h]%
+	\centering
+	\includegraphics[width=0.8\textwidth]{../osc2_impl} 
+	\caption[\ac{DCO} \ac{RTL} diagram as implemented]{\ac{DCO} \ac{RTL} diagram as implemented.}
+	\label{fig:osc2_impl}
+\end{figure}
+
 \subsubsection{Phase Detector}
+As previously mentioned the \ac{PFD} is also implemented using the \ac{FPGA} clocked approach, with sign detection carried out using a Moore Machine.
+The state transition diagram given as the example in Chapter \ref{chap:3} is that used to design this sign detector.
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=0.8\textwidth]{../state_trans_new}
+	\caption[Example State Transition Diagram for a Moore Machine]{Example State Transition Diagram for a Moore Machine.}
+	\label{fig:state_trans_reprint}
+\end{figure}
+
+The synchroniser circuits are implemented by a pair of ``D'' flip-flips connected in series, clocked on the \ac{FPGA} clock.
+This synchroniser works by assuming that metastability will only persist for the duration of one \ac{FPGA} clock cycle as depicted in Figure \ref{fig:synchroniser_behav}.
+In the case where only one clock signal may experience metastability, two synchronisers are required in order to maintain symmetry in the phase comparison.
+\begin{figure}[h]
+\centering
+\includegraphics[width=0.4\textwidth]{../synchroniser_behav}
+\caption[Double ``D'' flip-flip synchroniser circuit \ac{RTL} diagram]{Double ``D'' flip-flip synchroniser circuit \ac{RTL} diagram.}
+\label{fig:synchroniser_behav}
+\end{figure}
+%TODO context for above
+
+This \ac{FSM} then controls an Up-Down counter of the same width as the control code, counting in two's complement.
+As the control code has already been defined as a 7 bit wide two's complement integer, the phase detector's output can theoretically lie in the range $[-64,63]\cap\mathbb{Z}$.
+
+As this detector samples each signal at the \ac{FPGA} clock frequency, the phase detector resolution is equal to the \ac{FPGA} clock period at:
+\begin{equation}
+t_{res} = T_{FPGA} = \frac{1}{258\times 10^{6}} = 3.875~\si{\nano\second}
+\end{equation}
+an angular resolution of:
+\begin{equation}
+res = \frac{t_{res}}{T_{osc}} \cdot 360^\circ = \frac{3.875~\text{ns}}{200~\text{ns}} \cdot 360^\circ = 6.975^\circ
+\end{equation}
+
+The counter is implemented using saturation arithmetic to prevent overflow at high phase differences which could potentially harm the ability of the \ac{ADPLL} to obtain a lock.
+In the case of this \ac{PD} the phase error at which overflow may occur can be computed using the bit width of the counter and temporal resolution.
+$error_{max} = 63\cdot3.875~\text{ns} = 244.125~\text{ns}$.
+As this is greater than the period it cannot occur for purely a phase difference, but may occur in the presence of either a frequency divider in the feedback path, thus increasing the period of the signal used for comparison, or of a significant frequency difference between the signals.
+When implementing the clamping logic, the decision was taken to make the detection range symmetrical, thus the output range was reduced by one increment to $[-63,63]\cap\mathbb{Z}$.
+
+%TODO updown counter RTL diagram & discuss
+
+A major pitfall was encountered in this design of phase detector, which lead in some conditions to a form of mode-locking, in which each oscillator would lock $180\si{\degree}$ phase shifted from the oscillators used as a reference.
+This was later discovered to be as a result of the measurement interval termination conditions in the \ac{FSM}, which in addition to those described in Figure \ref{fig:state_trans_reprint}, would terminate the interval if a falling edge was observed on the signal that originally triggered the measurement.
+Figure \ref{fig:uncertainty} will be used to describe the exact circumstances of the error.
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=0.6\textwidth]{../uncertain}
+	\caption[Circumstances of mode locking behaviour]{Circumstances of mode locking behaviour.}
+	\label{fig:uncertainty}
+\end{figure}
+
+The situation would occur when the measurement interval began with a large phase difference between the generated and reference signals and a simultaneous difference in frequency.
+Without a frequency difference the measurement beginning at ``interval start'' would terminate at location ``a'', measuring a large lag, with the following measurement interval starting at location ``d''.
+However in the presence of a frequency differential, the rising edge on the generated signal may occur only at location ``c'', and with the falling edge termination condition causing the interval to terminate before this edge was detected, at point ``b''.
+This lead to a slightly lower magnitude phase lag detected, however this is a minor flaw, and not the cause of the mode locking.
+As the rising edge at point ``c'' occurs after the interval has terminated, it is treated as a fresh measurement interval, terminating at ``d''.
+As in this interval, the generated signal has seen the first rising edge, a large lead will be recorded.
+As each correction is made with a delay of one cycle, in order to avoid timing violations in the feedback path, there is potential for oscillation between lead and lag detection to occur depending on the initial conditions.
+This low probability behaviour was observed by chance, due to coding error, while using a frequency divider in the feedback path as this ensured the lead and lag measurements were of identical magnitude.
 
 \subsection{\acs{ADPLL} Design 2}
-\subsubsection{Oscillator}
+\subsubsection{\acl{DCO}}
 \subsubsection{Phase Detector}
@@
 \subsection{\acs{ADPLL} Design 3}
-\subsubsection{Oscillator}
+\subsubsection{\acl{DCO}}
 \subsubsection{Phase Detector}
@@
 \section{Minor Variations}\label{section:minor_variations}
+%TODO osc side
 

### 2019-04-15 thesis/Conclusion.tex
48.5

\section{Conclusions}
\section{Impact of Research}
\section{Suggested Future Work}

### 2019-04-15 thesis/FPGABasedADPLLs.tex
48.5

@@
 A Bang-Bang Detector is a sign-only detector, thus outputting a phase error just a single bit wide.
-This can be accomplished using a single flip-flop, with the reference signal acting as the clock, in this case using the rising edge, a design verified by Predrag in his 2011 thesis \cite{redrag}.
+This can be accomplished using a single flip-flop, with the reference signal acting as the clock, in this case using the rising edge, a design verified by Predraig in his 2011 thesis \cite{predraig}.
 The generated signal is applied to the $D$ input, thus when a rising edge is seen on the reference the value of the generated signal is held as the $Q$ output until the next rising edge of the reference.

### 2019-04-15 thesis/NetworkImplemenation.tex
48.5

@@
 From the above equation, the smallest counter that can satisfy this constraint is 12 bits wide, however when testing was carried out using a 275 MHz clock timing violations were discovered by Vivado, resulting \ac{FPGA} clock speed reduction to 258 MHz in order to resolve these violations.
-The correct bias point could then be calculated as $80$:
+The correct bias point could then be calculated as $79$:
 \begin{align}
 \frac{5}{258} &= \frac{BIAS}{2^{12}}\\
- BIAS &= \left \lceil{\frac{5\cdot 2^{12}}{258}}\right \rceil = \left \lceil{\frac{20480}{258}}\right \rceil = 80
+ BIAS &= \left \lfloor{\frac{5\cdot 2^{12}}{258}}\right \rceil = \left \lfloor{\frac{20480}{258}}\right \rceil = 79
 \end{align}
@@
 \end{equation}
+At the target frequency of 5 MHz this corresponds to a change in period of:
+\begin{align}
+f_{step0} &= 79 \times f_{step} = 79 \times 62.988~\si{\kilo\hertz} = 4.976~\si{\mega\hertz} \\
+f_{step1} &= 80 \times f_{step} = 80 \times 62.988~\si{\kilo\hertz} = 5.039~\si{\mega\hertz} \\
+T_{step}  &= \frac{1}{f_{step0}} - \frac{1}{f_{step1}} = \frac{1}{4.976\times 10^6} - \frac{1}{5.039\times 10^6} \\
+T_{step}  &= 2.5~\si{\nano\second}
+\end{align}
+From the frequency step, control code range and bias the frequency range of this oscillator can be found:
+\begin{align}
+f_{osc} &= (BIAS+CC)\times f_{step} = (79+CC)\times 62.988~\si{\kilo\hertz}\\
+f_{min} &= (79+CC_{min})\times 62.988~\si{\kilo\hertz} = (79-63)\times 62.988~\si{\kilo\hertz} = 1.008~\si{\mega\hertz}\\
+f_{max} &= (79+CC_{max})\times 62.988~\si{\kilo\hertz} = (79+63)\times 62.988~\si{\kilo\hertz} = 8.944~\si{\mega\hertz}
+\end{align}
 
@@
 \end{figure}
-%TODO context for above
 
@@
 \begin{equation}
-res = \frac{t_{res}}{T_{osc}} \cdot 360^\circ = \frac{3.875~\text{ns}}{200~\text{ns}} \cdot 360^\circ = 6.975^\circ
+res = \frac{t_{res}}{T_{osc}} \cdot 360\si{\degree} = \frac{3.875~\si{\nano\second}}{200~\si{\nano\second}} \cdot 360\si{\degree} = 6.975\si{\degree}
 \end{equation}
@@
 In the case of this \ac{PD} the phase error at which overflow may occur can be computed using the bit width of the counter and temporal resolution.
-$error_{max} = 63\cdot3.875~\text{ns} = 244.125~\text{ns}$.
+$error_{max} = 63\cdot3.875~\si{\nano\second} = 244.125~\si{\nano\second}$.
 As this is greater than the period it cannot occur for purely a phase difference, but may occur in the presence of either a frequency divider in the feedback path, thus increasing the period of the signal used for comparison, or of a significant frequency difference between the signals.
@@
 
-%TODO updown counter RTL diagram & discuss
+Figure \ref{fig:updown_ctr} depicts the implementation of this counter.
+Apart from the aforementioned capping of the measurement range implemented using a pair of comparators and a multiplexer, important to note is the bit width of the increment.
+Despite overflow being disallowed when the number is interpreted as a signed number, the addition of a 7 bit wide -1 to the accumulator value will cause overflow in the adder.
+This, however, is perfectly valid as an implementation of subtraction in two's complement.
+On the right side of the diagram the output register can be seen.
+Clocked using the \ac{FSM}s measurement interval complete signal, this register preserves the phase error until the termination of the following interval.
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=0.6\textwidth]{../bad_waveform}
+	\caption[\ac{RTL} diagram of up-down counter]{\ac{RTL} diagram of up-down counter.}
+	\label{fig:updown_ctr}
+\end{figure}
+%TODO updown counter RTL diagram
 
@@
 \subsection{\acs{ADPLL} Design 2}
+The second \ac{ADPLL} design implemented during the course of this project was somewhat of a stepping stone between fully synchronous and asynchronous to the \ac{FPGA} clock.
+The clocked phase detector is reused from above, albeit with an altered error width, while the \ac{DCO} is replaced by a \acl{RO}.
+While the phase detector was known to work due to testing in \ac{ADPLL} 1, the \ac{DCO} could not be simulated using accurate timings so all testing had to be done in hardware.
+This presented a challenge, as depending on the implementation, the characteristics of the \ac{DCO} could change, most importantly the centre frequency.
+
 \subsubsection{\acl{DCO}}
+\ac{ADPLL} 2 exchanges the \ac{FPGA} clocked oscillator for \ac{RO}, with the aim of obtaining performance more akin to that of an \ac{ASIC} based mixed-signal implementation, due to the variation in period steps both between oscillators and within the same oscillator due to the oscillator layout.
+Also advantageous is the improvement in period resolution obtained using this design, when compared to that of the \ac{FPGA} clocked design.
+Taking the formula given in Chapter \ref{chap:3} and the inverter propagation time used by Vivado in post synthesis/implementation simulations of $315~\si{\pico\second}$, the period step can be computed:
+\begin{align}
+t_{step} &= \text{inverters per step}\times\text{inversions per cycle}\times\text{propagation delay} \\
+t_{step} &= 2\times 2\times 315~\si{\pico\second} = 1.26~\si{\nano\second}
+\end{align}
+Using this period step, the number of inverters required to produce the signal of period $200~\si{\nano\second}$ can easily be computed:
+\begin{equation}
+\text{num\_inverters} = \left \lfloor{ T_{osc 5~\si{\mega\hertz}}}\times \frac{2}{t_{step}}}\right \rceil = \left \lfloor{ 200~\si{\nano\second}\frac{2}{1.26~\si{\nano\second}}}\right \rceil = 317
+\end{equation}
+
+Figure \ref{fig:ro_impl} contains an \ac{RTL} diagram of the \ac{RO} as implemented. %TODO impl & discuss
+
+A control code width of 5 bits was chosen as with two inverters removed per control code, the tunable range would consist of $20\%$ of the total inverter count.
+To achieve this range with a centre of $317$ inverters, the maximum number of inverters in the chain was required to be $317+2\times 2^{5-1} = 349$, dropped to a minimum at $285$.
+The corresponding minimum and maximum frequencies then are:
+\begin{align}
+f_{osc} &= \frac{1}{T_{osc}} = \frac{1}{\text{num. inverters}\times\text{propagation delay}} = \frac{1}{\text{num. inverters}\0.5\timest_{step}} \\
+f_{min} &= \frac{1}{T_{max}} = \frac{1}{349\times 0.5\times1.136~\si{\nano\second}} = 4.354~\si{\mega\hertz} \\
+f_{max} &= \frac{1}{T_{min}} = \frac{1}{285\times 0.5\times1.136~\si{\nano\second}} = 5.332~\si{\mega\hertz}
+\end{align}
+
+The first pitfall with this type of \ac{DCO} comes in the \ac{HDL} stage, with the \ac{EDA} likely to optimise away what it sees as a long chain of inverters carrying out no function.
+This can be avoided by adding an \texttt{attribute} as a prefix to the instantiation of a logic element or net that should not be ``optimised'' away.
+It is important to choose the correct directive, otherwise the compiler may not behave as expected.
+In the case of Vivado, it is important not to confuse the \texttt{KEEP} or \texttt{KEEP\_HIERARCHY} commands with that of \texttt{DONT\_TOUCH}, with former commands only ensuring that the logic elements will be maintained in synthesis, but not in any subsequent stages \cite{synth_ug}.
+
+As mentioned in previous chapters, the absence of direct control over layout can lead to significant variation of the propagation time though inverters.
+This may occur in three ways: Firstly, the layout of individual \ac{RO}s may be significantly different, thus resulting in poor overlap of tuning regions.
+Anecdotally, while implementing a 3x3 network, 7 of the 9 oscillators had centre frequencies within a $100~\si{\kilo\hertz}$ span but two lay more than $500~\si{\kilo\hertz}$ away in opposite directions which prevented the network from locking.
+Secondly, within an \ac{RO} the propagation delay between each inverter may vary, which results in a variable frequency step, possibly changing the locking range.
+These two effects represent an extreme version of the variation due to process or manufacturing seen on \ac{ASIC}s.
+The final variation is possibly the most frustrating, and occurs when between implementations the \ac{EDA}, Vivado in the case of this project, changes the layout of the \ac{RO}.
+This may occur as a result of a direct change, or as a result of seemingly innocuous changes to unrelated modules.
+While the final problem can be avoided, once the performance of the \ac{RO} is satisfactory, by locking down module, the remaining two problems can only be mitigated somewhat.
+
+This is achieved by assigning specific areas of the chip in which that module must lie, although these must be of a size approximately $40\%$ larger than the minimum space required in order to avoid having the router create a complex, delay intensive layout to fit the \ac{RO}.
+To achieve greater consistency, the implementation directive can be modified such that the router will attempt to use the minimum area that does not require complex routing.
+\texttt{congestion_spreadlogic_low} was used for this purpose in this project, although other options may obtain similar results.
+The other options available in Vivado can be viewed in the \textit{Vivado Design Suite User Guide - Implementation} \cite{impl_ug}.
+
+with small tweaks to unrelated modules likely to cause Vivado to rearrange the layout of the \ac{RO}s and thus change the centre frequency of a given \ac{DCO}.
+
 \subsubsection{Phase Detector}
+The \ac{PFD} used in this oscillator is, apart from the bit width of the counter, identical to that used in the previous design.
+The reduction in counter width will have no impact of performance of the \ac{ADPLL} once a lock has been acquired, however, the capture speed will be reduced for large phase differences due to the counter reaching saturation point significantly sooner.
+5 bits corresponds to a counter range of $[-15,15]\cap\mathbb{Z}$, which will saturate at a time delay of $15\times3.875\si{\nano\second} = 58.125~\si{\nano\second}$, a phase lead or lag of $\frac{}{}\times360\si{\degree} = 104.4\si{\degree}$ at $5~\si{\mega\hertz}$.
+This is an adequate range, as saturation will occur well outside of the normal operating range.
+This reduction is most noticeable when a divider is active in the feedback path, however during locked operation a phase shift of $58.125~\si{\nano\second}$ will not occur, so this impact is only felt during acquisition.
 

### 2019-04-15 thesis/TestingAnalysis.tex
48.5

\section{2x2 Network Performance Comparison}
\section{3x3 Network Performance Comparison}

### 2019-04-15 thesis/NetworkImplemenation.tex
diagrams please

@@
 
-with small tweaks to unrelated modules likely to cause Vivado to rearrange the layout of the \ac{RO}s and thus change the centre frequency of a given \ac{DCO}.
+With the hardware description of the \ac{RO} completed, testing of an instance of the \ac{RO} with a constant control code of zero was attempted, and it was discovered that the estimation of the centre frequency based on the propagation delay was off by $100$s of $\si{\kilo\hertz}$.
+The addition of 100 inverters was required to restore a centre of $5~\si{\mega\hertz}$.
+After the \ac{RO} was integrated into \ac{ADPLL} 2 the centre frequency was discovered to have changed once more.
+The minor variation caused by the now variable control code had changed the average propagation delay through the inverters once more, now requiring 373 inverters for $5~\si{\mega\hertz}$.
+
+When, later in the project, additional \ac{RO}s were implemented to form networks, 373 inverters proved a suitable starting point which consistently delievered \ac{DCO}s with a centre frequency within $200~\si{\kilo\hertz}$ of $5~\si{\mega\hertz}$.
+For the purposes of describing a typical tuning range, the average propagation delay of an \ac{RO} consisting of 373 inverters with a centre frequency of $5.06~\si{\mega\hertz}$ will be used.
+To avoid the aforementioned issues with constant control codes, the fixed code was achieved by implementing an \ac{ADPLL} and setting the \ac{PI} filter gains to zero at runtime before performing a reset.
+%TODO possibily problematic in the last two paragraphs
+
+%TODO freqs n stuff
 

### 2019-04-15 thesis/NetworkImplemenation.tex
49

@@
 	\centering
-	\includegraphics[width=0.6\textwidth]{../bad_waveform}
+	\includegraphics[width=0.8\textwidth]{../updown}
 	\caption[\ac{RTL} diagram of up-down counter]{\ac{RTL} diagram of up-down counter.}
@@
 \end{figure}
-%TODO updown counter RTL diagram
 
@@
 \begin{equation}
-\text{num\_inverters} = \left \lfloor{ T_{osc 5~\si{\mega\hertz}}}\times \frac{2}{t_{step}}}\right \rceil = \left \lfloor{ 200~\si{\nano\second}\frac{2}{1.26~\si{\nano\second}}}\right \rceil = 317
+\text{num\_inverters} = \left \lfloor{ T_{osc 5~\si{\mega\hertz}}\times \frac{2}{t_{step}}}\right \rceil = \left \lfloor{ 200~\si{\nano\second}\frac{2}{1.26~\si{\nano\second}}}\right \rceil = 317
 \end{equation}
 
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=0.8\textwidth]{../ro_new}
+	\caption[\acl{RO} RTL diagram]{\acl{RO} RTL diagram.}
+	\label{fig:ro_impl}
+\end{figure}
 Figure \ref{fig:ro_impl} contains an \ac{RTL} diagram of the \ac{RO} as implemented. %TODO impl & discuss
@@
 \begin{align}
-f_{osc} &= \frac{1}{T_{osc}} = \frac{1}{\text{num. inverters}\times\text{propagation delay}} = \frac{1}{\text{num. inverters}\0.5\timest_{step}} \\
+f_{osc} &= \frac{1}{T_{osc}} = \frac{1}{\text{num. inverters}\times 2\times\text{propagation delay}} = \frac{1}{\text{num. inverters}\times 0.5t_{step}} \\
 f_{min} &= \frac{1}{T_{max}} = \frac{1}{349\times 0.5\times1.136~\si{\nano\second}} = 4.354~\si{\mega\hertz} \\
@@
 To achieve greater consistency, the implementation directive can be modified such that the router will attempt to use the minimum area that does not require complex routing.
-\texttt{congestion_spreadlogic_low} was used for this purpose in this project, although other options may obtain similar results.
+\texttt{congestion\_spreadlogic\_low} was used for this purpose in this project, although other options may obtain similar results.
 The other options available in Vivado can be viewed in the \textit{Vivado Design Suite User Guide - Implementation} \cite{impl_ug}.
@@
 The minor variation caused by the now variable control code had changed the average propagation delay through the inverters once more, now requiring 373 inverters for $5~\si{\mega\hertz}$.
+%TODO possibily problematic in the last two paragraphs
 
@@
 To avoid the aforementioned issues with constant control codes, the fixed code was achieved by implementing an \ac{ADPLL} and setting the \ac{PI} filter gains to zero at runtime before performing a reset.
-%TODO possibily problematic in the last two paragraphs
 
@@
 \section{Minor Variations}\label{section:minor_variations}
-%TODO osc side
+%TODO osc size
 

### 2019-04-16 thesis/NetworkImplemenation.tex
50.1 pages

@@
 	\caption[Frequency divider \ac{RTL} diagram]{Frequency divider \ac{RTL} diagram.}
-	\label{fig:divs_iompl}
+	\label{fig:divs_impl}
 \end{figure}
@@
 
+\subsubsection{\acs{ADPLL} 1 Design Summary}
+
 \subsection{\acs{ADPLL} Design 2}
@@
 \end{figure}
-Figure \ref{fig:ro_impl} contains an \ac{RTL} diagram of the \ac{RO} as implemented. %TODO impl & discuss
+Figure \ref{fig:ro_impl} contains an \ac{RTL} diagram of the \ac{RO} as implemented.
+The inverter chain is specified with a maximum length of $n$ and a maximum number of removable inverters of $k$.
+In order to preserve the unstable circuit, by maintaining an odd number of inverters, only multiples of two inverters may be removed at a time.
+The already introduced \ac{DCO} and \ac{PFD} designs all followed the same convention regarding phase error and the corresponding impact on the control code: In the \ac{PFD}, if the generated signal was lagging the reference, meaning it should ``go faster'' in order to ``catch up'', the error given a positive sign.
+Similarly in the \ac{FPGA} clocked \ac{DCO}, a positive change in the magnitude of the control code resulted in a higher frequency.
+In the interests of consistency and extensibility the same behaviour was implemented for this \ac{DCO} also, with the number of inverters in the chain given as:
+\begin{equation}
+\text{num\_inverters} = inv_{max} - 2\times\text{control\_code},~\text{control\_code}<\frac{k}{2}
+\end{equation}
+$k$ is determined by the bit width of the control code which, as in the case of the clocked \ac{DCO}, is also an unsigned number.
+An inverter chain will oscillate freely and cannot be disabled.
+To this end the first inverter in the chain was replaced by a \ac{NAND} gate to grant control over its operation.
 
-A control code width of 5 bits was chosen as with two inverters removed per control code, the tunable range would consist of $20\%$ of the total inverter count.
+The bias point functions differently in this oscillator, as unlike the \ac{FPGA} clocked design it is not required to set the centre of the tuning range, as in the \ac{RO} this lies at $n-\frac{k}{2}$.
+Instead the bias point servers as a conversion between the signed \ac{LF} output, which is in two's complement form, and the unsigned requirement of the control code.
+This is achieved by adding on the midpoint of the unsigned range and storing the result in an unsigned value.
+Using the case where the width of the control code is three: The range of the signed three bit integer is from $-4$ (100) to $3$ (011) which when directly converted to unsigned values become $4$ and $3$, thus destroying the relationship between lead and lag.
+Worse still the signed value of $-1$ (111) becomes $7$.
+Adding on the mid range value solves this problem, with a signed value of $-4$ now mapping to $0$ and $3$ to $7$.
+Comparing this behaviour to response of the \ac{RO} to the control code, it can be trivially seen that the minimum possible period will still correspond to the largest control code prior to conversion and vice versa for the maximum period.
+
+A control code width of 5 bits was chosen, as with two inverters removed per control code, the tunable range would consist of $20\%$ of the total inverter count.
 To achieve this range with a centre of $317$ inverters, the maximum number of inverters in the chain was required to be $317+2\times 2^{5-1} = 349$, dropped to a minimum at $285$.
@@
 f_{osc} &= \frac{1}{T_{osc}} = \frac{1}{\text{num. inverters}\times 2\times\text{propagation delay}} = \frac{1}{\text{num. inverters}\times 0.5t_{step}} \\
-f_{min} &= \frac{1}{T_{max}} = \frac{1}{349\times 0.5\times1.136~\si{\nano\second}} = 4.354~\si{\mega\hertz} \\
-f_{max} &= \frac{1}{T_{min}} = \frac{1}{285\times 0.5\times1.136~\si{\nano\second}} = 5.332~\si{\mega\hertz}
+f_{min} &= \frac{1}{T_{max}} = \frac{1}{349\times 0.5\times1.26~\si{\nano\second}} = 4.548~\si{\mega\hertz} \\
+f_{max} &= \frac{1}{T_{min}} = \frac{1}{285\times 0.5\times1.26~\si{\nano\second}} = 5.569~\si{\mega\hertz}
 \end{align}
@@
 
-When, later in the project, additional \ac{RO}s were implemented to form networks, 373 inverters proved a suitable starting point which consistently delievered \ac{DCO}s with a centre frequency within $200~\si{\kilo\hertz}$ of $5~\si{\mega\hertz}$.
+When, later in the project, additional \ac{RO}s were implemented to form networks, 373 inverters proved a suitable starting point which consistently delivered \ac{DCO}s with a centre frequency within $200~\si{\kilo\hertz}$ of $5~\si{\mega\hertz}$.
 For the purposes of describing a typical tuning range, the average propagation delay of an \ac{RO} consisting of 373 inverters with a centre frequency of $5.06~\si{\mega\hertz}$ will be used.
 To avoid the aforementioned issues with constant control codes, the fixed code was achieved by implementing an \ac{ADPLL} and setting the \ac{PI} filter gains to zero at runtime before performing a reset.
+
+Using this updated maximum number of inverters the frequency range can be recalculated.
+Firstly, the new minimum and maximum numbers of inverters are $309$ and $373$ respectively, with a mid point at $341$.
+The inverter delay and period step are then recomputed based on the centre frequency of $5~\si{\mega\hertz}$ as:
+\begin{align}%TODO 2 x 2 x midpoint justification - in prev. chapter?
+t_{step} = \frac{T_{osc}}{0.5\times\text{inverter midpoint}} = \frac{200\si{\nano\second}}{0.5\times 341} = 1.176~\si{\nano\second}
+\text{propagation delay} &= \frac{t_{step}}{2\times2} = \frac{1.176~\si{\nano\second}}{4} = 293~\si{\pico\second} \\
+\end{align}
+The frequency range is then:
+\begin{align} %TODO inconsisentl calculation approach with previous version
+f_{osc} &= \frac{1}{T_{osc}} = \frac{1}{\text{num. inverters}\times 2\times\text{propagation delay}}
+f_{min} &= \frac{1}{T_{max}} = \frac{1}{373\times 2\times 293~\si{\pico\second}} = 4.571~\si{\mega\hertz} \\
+f_{max} &= \frac{1}{T_{min}} = \frac{1}{309\times 2\times 293~\si{\pico\second}} = 5.518~\si{\mega\hertz}
+\end{align}
 
@@
 
+\subsubsection{\acs{ADPLL} 2 Design Summary}
+
 \subsection{\acs{ADPLL} Design 3}
@@
 \subsubsection{Phase Detector}
+\subsubsection{\acs{ADPLL} 1 Design Summary}
 

### 2019-04-17 thesis/NetworkImplemenation.tex
54

@@
 
+\subsubsection{Simulations}
+%TODO simulations
+As this was the first \ac{ADPLL} design implemented, various Vivado test benches were used to verify the behaviour of each module.
+Simulations with this design were made more efficient due to deterministic timing in all modules, and absence of any inverter based components which required the use of post-implementation simulations in order to avoid zero delay oscillations.
+Deterministic timing means \ac{ADPLL} 1 offers an additional, and significant, advantage over other designs as it can be entirely simulated in Vivado with accurate timings.
+This ability was vital initially, as it revealed the presence of edge situations that had not been accounted for, and through exporting the waveforms to a log file and analysis performed using Matlab comparison made to the measured system behaviour.
+
+\begin{figure}[h]
+	\centering
+	\subfloat[Vivado simulation.]{{\includegraphics[width=0.45\textwidth]{../sim_locking} }}\\
+	\subfloat[Hardware measurement.]{{\includegraphics[width=0.45\textwidth]{../impl_locking} }}%
+	\caption[\ac{ADPLL} 1 Vivado locking simulations]{\ac{ADPLL} 1 Vivado locking simulations.}
+	\label{fig:sim_locking}
+\end{figure}
+In Figure \ref{fig:sim_locking} (a), the startup behaviour of the system can be observed.
+This simulation was performed with a reference at exactly $5~\si{\mega\hertz}$ and an initial deviation of $252~\si{\kilo\hertz}$, and the plot shows the inverse of the \ac{DCO} period against the time since the oscillator was enabled.
+After some initial oscillation the period settles around that of the reference.
+Once locked, cycle-to-cycle jitter was calculated to be $3.491~\si{\nano\second}$ using the standard deviation of the periods, with a worst case cycle-to-cycle change of $9.299~\si{\nano\second}$.
+Lock was visually determined to have occurred after $20.91~\si{\micro\second}$, after which the oscillator frequency began to change in a repetitive pattern.
+later the same test carried in hardware with identical initial conditions, the results of which are shown in Figure \ref{fig:sim_locking} (b).
+This plot is derived from data captured using an Agilent MSO7054A, using a setup that will be further explained in Section \ref{section:measurement_setup}.
+Despite lacking a non ideal reference, less violent initial oscillation, and a lesser until lock could be visually determined, the simulation provided a good insight into the system's behaviour once implemented on an \ac{FPGA}.
+
 \subsubsection{\acs{ADPLL} 1 Design Summary}
+The listed components were brought together in \textsc{NetworkADPLL.v}, attached in Appendix \ref{adx:code} Listing \ref{lst:network_adpll}.
+Important to note is that each \ac{ADPLL} only contains two \acl{PD}s despite during network operation the need for phase comparisons to be made with up to four neighbours.
+The two ``missing'' \ac{PD}s are in fact synthesised as part of the other \ac{ADPLL}s, with which the comparison would be made.
+This allows for a 50\% reduction in the number of \ac{PD}s required, while also ensuring identical error is received by both \ac{ADPLL}s.
+Each \ac{ADPLL} performs comparison with the \ac{DCO} ``above'' and ``left'' of them in the Cartesian grid, and a port in the module outputs the results for use by the other \ac{ADPLL}s involved.
+%TODO juvenile wording
+Table \ref{table:adpll1} contains a brief overview of the bit widths, tuning ranges and other configuration information for this \ac{ADPLL}.
+\begin{table}[!h]
+	\begin{center}
+		\begin{tabular}{|l|r|r|r|}
+			\multicolumn{4}{c}{\ac{DCO} Tuning Range} \T\\
+			\hline
+			\multicolumn{1}{|c|}{-} & \multicolumn{1}{c|}{Minimum} & \multicolumn{1}{c|}{Step} & \multicolumn{1}{c|}{Maximum} \T\\
+			\hline
+			Frequency & $1.008~\si{\mega\hertz}$ & \multicolumn{1}{r|}{$62.988~\si{\kilo\hertz}$} & $8.944~\si{\mega\hertz}$ \T\\
+			\hline
+			Period & $992.0~\si{\nano\second}$ & \multicolumn{1}{r|}{$2.5~\si{\nano\second}$ (at $5~\si{\mega\hertz}$)} & $111.8~\si{\nano\second}$ \T\\
+			\hline
+		\end{tabular}
+		\begin{tabular}{|l|r|l|r|}
+			\multicolumn{4}{c}{Configuration as Implemented} \T\\
+			\hline
+			\ac{DCO} Counter Width & 12 bits & \ac{DCO} Control Width (max) & 12 bits \T\\
+			\hline
+			\ac{DCO} Bias Point & 79 & Runtime Gains & Enabled \T\\
+			\hline
+			Phase Detector Error Width & 7 & Error Sum Weight Width & 4 \T\\
+			\hline
+			\acs{kp} & $2^{-5}\times[0,15]\cap\mathbb{Z}$ & \acs{ki} & $2^{-8}\times[0,15]\cap\mathbb{Z}$ \T\\
+			\hline
+			Detection Resolution & $3.875~\si{\nano\second}$ & Detection Phase Resolution & $6.975\si{\degree}$ (at $5~\si{\mega\hertz}$)\\
+			\hline
+		\end{tabular}
+	\end{center}
+\caption[ADPLL Design 1 Summary]{ADPLL Design 1 Summary.}
+\label{table:adpll1}
+\end{table}
 
@@
 \begin{align}
-t_{step} &= \text{inverters per step}\times\text{inversions per cycle}\times\text{propagation delay} \\
+t_{step} &= \text{inverters per step}\times\text{inversions per period}\times\text{propagation delay} \\
+t_{step} &= \text{inverters per step}\times\text{inversions per period}\times\tau_{inv} \\
 t_{step} &= 2\times 2\times 315~\si{\pico\second} = 1.26~\si{\nano\second}
 \end{align}
-Using this period step, the number of inverters required to produce the signal of period $200~\si{\nano\second}$ can easily be computed:
+Using the propagation delay, $\tau_{inv}$, the number of inverters required to produce the signal of period $200~\si{\nano\second}$ can easily be computed:
 \begin{equation}
-\text{num\_inverters} = \left \lfloor{ T_{osc 5~\si{\mega\hertz}}\times \frac{2}{t_{step}}}\right \rceil = \left \lfloor{ 200~\si{\nano\second}\frac{2}{1.26~\si{\nano\second}}}\right \rceil = 317
+\text{num\_inverters} = \left \lfloor{ T_{osc~@~5~\si{\mega\hertz}}\times \frac{1}{2\times\tau_{inv}}}\right \rceil = \left \lfloor{ \frac{200~\si{\nano\second}}{0.63~\si{\nano\second}}}\right \rceil = 317
 \end{equation}
@@
 \begin{equation}
-\text{num\_inverters} = inv_{max} - 2\times\text{control\_code},~\text{control\_code}<\frac{k}{2}
+\text{num inverters} = n - 2\times CC,~CC<\frac{k}{2}
 \end{equation}
+Where $CC$ is the unsigned integer control code input.
 $k$ is determined by the bit width of the control code which, as in the case of the clocked \ac{DCO}, is also an unsigned number.
 An inverter chain will oscillate freely and cannot be disabled.
-To this end the first inverter in the chain was replaced by a \ac{NAND} gate to grant control over its operation.
+To this end the first inverter in the chain was replaced by a \acs{NAND} gate to grant control over its operation.
 
@@
 \begin{align}
-f_{osc} &= \frac{1}{T_{osc}} = \frac{1}{\text{num. inverters}\times 2\times\text{propagation delay}} = \frac{1}{\text{num. inverters}\times 0.5t_{step}} \\
-f_{min} &= \frac{1}{T_{max}} = \frac{1}{349\times 0.5\times1.26~\si{\nano\second}} = 4.548~\si{\mega\hertz} \\
-f_{max} &= \frac{1}{T_{min}} = \frac{1}{285\times 0.5\times1.26~\si{\nano\second}} = 5.569~\si{\mega\hertz}
+f_{osc} &= \frac{1}{T_{osc}} = \frac{1}{(n-2CC)\times 2\times\tau_{inv}} = \frac{1}{(n-2CC)\times 2\times 315~\si{\pico\second}} \\
+f_{min} &= \frac{1}{T_{max}} = \frac{1}{349\times 2\times 315~\si{\pico\second}} = 4.548~\si{\mega\hertz} \\
+f_{max} &= \frac{1}{T_{min}} = \frac{1}{285\times 2\times 315~\si{\pico\second}} = 5.569~\si{\mega\hertz}
 \end{align}
@@
 \begin{align}%TODO 2 x 2 x midpoint justification - in prev. chapter?
-t_{step} = \frac{T_{osc}}{0.5\times\text{inverter midpoint}} = \frac{200\si{\nano\second}}{0.5\times 341} = 1.176~\si{\nano\second}
-\text{propagation delay} &= \frac{t_{step}}{2\times2} = \frac{1.176~\si{\nano\second}}{4} = 293~\si{\pico\second} \\
+t_{step} &= \frac{T_{osc}}{0.5\times(n-\frac{k}{2})} = \frac{200\si{\nano\second}}{0.5\times 341} = 1.176~\si{\nano\second} \\
+\tau_{inv} &= \frac{t_{step}}{2\times\text{inverters per step}} = \frac{t_{step}}{2\times2} = \frac{1.176~\si{\nano\second}}{4} = 293~\si{\pico\second}
 \end{align}
@@
 \begin{align} %TODO inconsisentl calculation approach with previous version
-f_{osc} &= \frac{1}{T_{osc}} = \frac{1}{\text{num. inverters}\times 2\times\text{propagation delay}}
-f_{min} &= \frac{1}{T_{max}} = \frac{1}{373\times 2\times 293~\si{\pico\second}} = 4.571~\si{\mega\hertz} \\
-f_{max} &= \frac{1}{T_{min}} = \frac{1}{309\times 2\times 293~\si{\pico\second}} = 5.518~\si{\mega\hertz}
+f_{osc} &= \frac{1}{T_{osc}} = \frac{1}{(n-2CC)\times 2\times\tau_{inv}} \\
+f_{min} &= \frac{1}{T_{max}} = \frac{1}{(373-0)\times 2\times 293~\si{\pico\second}} = 4.571~\si{\mega\hertz} \\
+f_{max} &= \frac{1}{T_{min}} = \frac{1}{(373-2\times32)\times 2\times 293~\si{\pico\second}} = 5.518~\si{\mega\hertz}
 \end{align}
-
-%TODO freqs n stuff
 
@@
 \subsubsection{\acs{ADPLL} 2 Design Summary}
+As previously mentioned, \ac{ADPLL} 2 represents a midway point between \ac{FPGA} clocked designs and those which are asynchronous.
+As the source of the asynchronicity is the inverter chain \ac{DCO}, a more realistic jitter would be expected from this design, with there being a significant distribution of period steps.
+Among other aspects, this will have a noticeable impact in steady state conditions, where the degree of fractional synthesis required will vary from oscillator to oscillator depending on the closest distance from the reference period to a period obtainable by the individual oscillator.
+This is a source of jitter that was not present in the clocked \ac{DCO} used in \ac{ADPLL} 1, as all period steps were equal.
+Once again the \ac{DCO} tuning ranges and configuration details for future measurements are listed below, in Table \ref{table:adpll2}.
+
+\begin{table}[!h]%TODO numbers!
+	\begin{center}
+		\begin{tabular}{|l|r|r|r|}
+			\multicolumn{4}{c}{\ac{DCO} Tuning Range} \T\\
+			\hline
+			\multicolumn{1}{|c|}{-} & \multicolumn{1}{c|}{Minimum} & \multicolumn{1}{c|}{Step} & \multicolumn{1}{c|}{Maximum} \T\\
+			\hline
+			Frequency & $4.571~\si{\mega\hertz}$ & $29.180~\si{\kilo\hertz}$ (at $5~\si{\mega\hertz}$) & $5.518~\si{\mega\hertz}$ \T\\
+			\hline
+			Period & $218.2~\si{\nano\second}$ & $1.176~\si{\nano\second}$ & $181.1~\si{\nano\second}$ \T\\
+			\hline
+		\end{tabular}
+		\begin{tabular}{|l|r|l|r|}
+			\multicolumn{4}{c}{Configuration as Implemented} \T\\
+			\hline
+			\ac{RO} num. of inverters (max) & 373 & \ac{RO} Control Width & 5 bits \T\\
+			\hline
+			\ac{RO} Bias Point & 16 & Runtime Gains & Enabled \T\\
+			\hline
+			Phase Detector Error Width & 5 & Error Sum Weight Width & 4 \T\\
+			\hline
+			\acs{kp} & $2^{-7}\times[0,15]\cap\mathbb{Z}$ & \acs{ki} & $2^{-9}\times[0,15]\cap\mathbb{Z}$ \T\\
+			\hline
+			Detection Resolution & $3.875~\si{\nano\second}$ & Detection Phase Resolution & $6.975\si{\degree}$ (at $5~\si{\mega\hertz}$)\\
+			\hline
+		\end{tabular}
+	\end{center}
+	\caption[ADPLL Design 2 Summary]{ADPLL Design 2 Summary.}
+	\label{table:adpll2}
+\end{table}
 
@@
 \subsubsection{Phase Detector}
-\subsubsection{\acs{ADPLL} 1 Design Summary}
+
+\begin{figure}[h]
+	\centering
+	\subfloat[Modified Bang-Bang sign detector.]{{\includegraphics[width=0.8\textwidth]{../new_pdet1} }}\\
+	\subfloat[\acl{TDL}.]{{\includegraphics[width=1.0\textwidth]{../new_pdet2} }}%
+	\caption[\acl{PD} RTL diagrams]{\acl{PD} RTL diagrams.}
+	\label{fig:signum_impl}
+\end{figure}
+
+\section{Measurement Setup}\label{section:measurement_setup}
 

### 2019-04-17 thesis/FPGABasedADPLLs.tex
? pages

@@
 
-\subsection{SigNum Detector}
+\subsection{SigNum Detector}\label{section:signum}
 The SigNum detector is based on the methodology used in the Bang-Bang \acl{PD} seen previously, but rather than just performing sign detection, the magnitude of the phase difference can also be measured.

### 2019-04-17 thesis/NetworkImplemenation.tex
? pages

@@
 The state transition diagram given as the example in Chapter \ref{chap:3} is that used to design this sign detector.
+The hardware description of this block is found in \textsc{PhaseDetector.v}, attached in Appendix \ref{adx:code} Listing \ref{lst:PhaseDetector}.
 \begin{figure}[h]
@@
 This presented a challenge, as depending on the implementation, the characteristics of the \ac{DCO} could change, most importantly the centre frequency.
+The module implementing this \ac{ADPLL} can be found in \textsc{NetworkRingADPLL.v}, attached in Appendix \ref{adx:code} Listing \ref{lst:network_ring_adpll}.
 
@@
 
+The module implementing the \ac{RO} can be found in \textsc{RingOsc.v}, attached in Appendix \ref{adx:code} Listing \ref{lst:ring_osc}.
+
 \subsubsection{Phase Detector}
@@
 Once again the \ac{DCO} tuning ranges and configuration details for future measurements are listed below, in Table \ref{table:adpll2}.
+
+\begin{table}[!h]%TODO redo numbers!
+	\begin{center}
+		\begin{tabular}{|l|r|r|r|}
+			\multicolumn{4}{c}{\ac{DCO} Tuning Range} \T\\
+			\hline
+			\multicolumn{1}{|c|}{-} & \multicolumn{1}{c|}{Minimum} & \multicolumn{1}{c|}{Step} & \multicolumn{1}{c|}{Maximum} \T\\
+			\hline
+			Frequency & $4.571~\si{\mega\hertz}$ & $29.180~\si{\kilo\hertz}$ (at $5~\si{\mega\hertz}$) & $5.518~\si{\mega\hertz}$ \T\\
+			\hline
+			Period & $218.2~\si{\nano\second}$ & $1.176~\si{\nano\second}$ & $181.1~\si{\nano\second}$ \T\\
+			\hline
+		\end{tabular}
+		\begin{tabular}{|l|r|l|r|}
+			\multicolumn{4}{c}{Configuration as Implemented} \T\\
+			\hline
+			\ac{RO} num. of inverters (max) & 373 & \ac{RO} Control Width & 5 bits \T\\
+			\hline
+			\ac{RO} Bias Point & 16 & Runtime Gains & Enabled \T\\
+			\hline
+			Phase Detector Error Width & 5 & Error Sum Weight Width & 4 \T\\
+			\hline
+			\acs{kp} & $2^{-7}\times[0,15]\cap\mathbb{Z}$ & \acs{ki} & $2^{-9}\times[0,15]\cap\mathbb{Z}$ \T\\
+			\hline
+			Detection Resolution & $3.875~\si{\nano\second}$ & Detection Phase Resolution & $6.975\si{\degree}$ (at $5~\si{\mega\hertz}$)\\
+			\hline
+		\end{tabular}
+	\end{center}
+	\caption[\ac{ADPLL} Design 2 Summary]{\ac{ADPLL} Design 2 Summary.}
+	\label{table:adpll2}
+\end{table}
+
+\subsection{\acs{ADPLL} Design 3}
+\acs{ADPLL} Design 3 has no dependence on the clock provided by the \ac{FPGA}'s clock management utility, as both the \ac{RO} and the \ac{TDC} are designed using inverters, while a Bang-Bang detector replaces the \ac{FPGA} clock reliant \ac{FSM} of the two previous designs.
+This introduces a second source of implementation based variation between \ac{ADPLL}s, however, this has significantly lesser impact on the frequency range of the oscillator and no impact on the centre frequencies.
+The lack of an \ac{FPGA} clocked element results means this design will have the closest behaviour to that of an \ac{ASIC} based \ac{ADPLL} implementation, although as the level of variation due to implementation is much larger on an \ac{FPGA}, the impact of non-idealities will be exaggerated.
+%same interface
+In order to allow quick exchange of \ac{PD}s or other modules, the module in which this \ac{ADPLL} is implemented is still retains the \ac{FPGA} clock as an input.
+In fact, this module shares its implementation with that of \ac{ADPLL} 2, as while the \ac{PD} designs are drastically different, the interface is identical.
+The module implementing this \ac{ADPLL} can be found in \textsc{NetworkRingADPLL.v}, attached in Appendix \ref{adx:code} Listing \ref{lst:network_ring_adpll}.
+
+\subsubsection{\acl{DCO}}
+The \ac{DCO} used in \ac{ADPLL} 3 is identical in form to that used in \ac{ADPLL} 2, so the tuning range, period step and all effects of implementation variations carry over to this \ac{ADPLL}.
+The decision to leave the \ac{RO} unmodified is an easy one, as the intended centre frequency is the same and changing the control code width would remove the ability to fix the cells of each \ac{RO}.
+Fixing the cells is beneficial as it will maintain the oscillator layout between implementation ``runs'', and thus allow for a better comparison between designs 1 \& 2.
+
+\subsubsection{Phase Detector}
+Instead of an \ac{FPGA} clocked phase detector, \ac{ADPLL} 3 uses the modified Bang-Bang detector and \ac{TDL} approach to phase comparison seen in Section \ref{section:signum}.
+Sign detection is implemented exactly as described in that section, with the double \ac{SR} latch approach used to determine the sign of the phase error.
+A flip-flop is inserted after the second \ac{SR} latch, clocked on the signal \texttt{done}, which holds the output value and ensures that the sign changes at the same time as the magnitude.
+The previous discussion of this phase detector proposed no implementation of the edge detectors, which are easily implemented using flip-flops clocked on the signal of interest, with a constant $1$ on the $D$ input.
+In this form the edge detector is single use only, as subsequent edges will not cause any change in the flip-flop output.
+This is important as it ensures the measurement interval will only terminate when a rising edge is seen on the signal that did not start the measurement interval, thus avoiding the mode locking seen in the original implementation of the \ac{FPGA} clocked \ac{PD}.
+Once the measurement interval has completed, the \texttt{done} signal is used to reset the edge detectors asynchronously.
+It is possible that an edge may occur within the time taken to reset the counters, and thus the detector incorrectly gauge the start of a measurement interval, however this can only occur in exceptional circumstances in the acquisition period.
+Therefore it is possible to ignore this scenario, as the only impact will be an increased time taken to acquire a lock.
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=0.8\textwidth]{../new_pdet1}
+	\caption[Modified Bang-Bang sign detector RTL diagrams]{Modified Bang-Bang sign detector RTL diagrams.}
+	\label{fig:signdet_impl}
+\end{figure}
+
+A two signal interface controls the \ac{TDC}, a \texttt{count} that enables the \ac{TDC} and a \texttt{done} signal that signals the termination of the interval.
+A single logic gate is required to generate each signal, as ORing the two edge detector outputs will produce logic $1$ when the interval has begun while ANDing these signals will indicate the end of the interval when both edges have occurred.
+The \texttt{done} signal is also used to clock the flip-flop acting as a buffer for the sign.
+The exact implementation of the sign detection circuit can be seen in Figure \ref{fig:signdet_impl}.
+
+As previously mentioned, the time-to-digital conversion is carried out by a tapped delay line.
+As with the \ac{RO}, inverters will be used to provide the delay between each tap, and thus the resolution of the \ac{TDL} and the length of each individual delay is based on the layout chosen by the router in Vivado.
+Due to this, the exact resolution of the phase detector is unknown, although the delay through inverters used in Vivado simulations and the average delay computed based on \ac{RO} centre frequencies can be used to estimate that of the phase detectors.
+Based these figures for the the propagation delay due to an inverter, $\tau_{inv}$ can be estimated to be in the region of $300~\si{\pico\second}$.
+As each tap consists of an inverter pair, the delay through the tap, $\tau_{tap} \approx 600~\si{\pico\second}$.
+Characterisation of the \ac{TDL}, and thus measurement of the delays in each \ac{TDL}, is technically possible, however, this is an time consuming process as many \ac{TDL}s would need to be characterised in order to accurately compute the average delay.
+
+%TODO eugene
+The delay line itself is shown in Figure \ref{fig:tdl_impl}, with each tap implemented by a pair of flip-flops and three inverters.
+For an $N$ bit width phase error detection, $2^{N-1}-1$ taps are required.
+In the case of this \ac{ADPLL}, a 5 bit two's complement error signal is required, thus the range of the signal is $[-16,15]\cap\mathbb{Z}$.
+If the range is made symmetrical, 15 taps are required to fill a signal of this size.
+From the diagram it can be seen that unless edges are detected on each signal at exactly the same instant, it is almost impossible to measure a zero phase difference between reference and generated signals.
+This is an intentional decision, \cite{idkwhattocite}.
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=1.0\textwidth]{../new_pdet2}
+	\caption[Inverter based \ac{TDL} RTL diagrams]{Inverter based \ac{TDL} RTL diagrams.}
+	\label{fig:tdl_impl}
+\end{figure}
+The operation of the \ac{TDL} is as follows: A delayed version of the \texttt{count} signal acts as the clock for flip-flops forming each tap, which are rising edge activated.
+These flip-flops have a constant ``1'' at their $D$ inputs, so as the signal propagates through the delay line, each flip-flop with propagate a logic ``1'' to their corresponding \texttt{error_tap[i]} signal.
+When the measurement interval has completed, the \texttt{count} signal clocks the 15 bit wide \texttt{error_tap} signal into \texttt{error_tap_buff} which act as the output buffers of the \ac{PD}, holding the phase error constant between measurement intervals.
+As with the flip-flops performing edge detection, the flip-flops forming the taps are can only be used once before requiring a reset, which is again carried out using the \texttt[done] signal.
+
+\subsubsection{\acs{ADPLL} 3 Design Summary}
+Table \ref{table:adpll3}.
 
@@
 	\end{center}
-	\caption[ADPLL Design 2 Summary]{ADPLL Design 2 Summary.}
-	\label{table:adpll2}
+	\caption[\ac{ADPLL} Design 3 Summary]{\ac{ADPLL} Design 3 Summary.}
+	\label{table:adpll3}
 \end{table}
 
-\subsection{\acs{ADPLL} Design 3}
-\subsubsection{\acl{DCO}}
-\subsubsection{Phase Detector}
-
+\section{Measurement Setup}\label{section:measurement_setup}
+For the following sections in which the performance of either individual \ac{ADPLL}s or \ac{ADPLL} networks the same measurement setup was used.
+The \ac{FPGA} used was a Xilinx Artix-7 \acl{Nexys} on a Digilent \acs{Nexys} board.
+The \acl{Nexys} features 15850 logic slices, each with 4 lookup tables and 8 flip-flops.
+There are six clock management tiles, 240 dedicated digital signal processing slices and an analog-to-digital converter.
+The maximum operating frequency is $464~\si{\mega\hertz}$ and 210 input/output ports are available for use.
+Further information can be found in the datasheet \cite{a7_datasheet}.
+The \acs{Nexys} is an evaluation/development board for the aforementioned \ac{FPGA}, however directed at the education sector.
+It contains useful peripherals such as 16 switches, 8 7-segment displays and 4 user addressable \ac{PMOD} headers and can be used with the free of charge ``Webpack'' version of Vivado \cite{n4_datasheet}.
 \begin{figure}[h]
 	\centering
-	\subfloat[Modified Bang-Bang sign detector.]{{\includegraphics[width=0.8\textwidth]{../new_pdet1} }}\\
-	\subfloat[\acl{TDL}.]{{\includegraphics[width=1.0\textwidth]{../new_pdet2} }}%
-	\caption[\acl{PD} RTL diagrams]{\acl{PD} RTL diagrams.}
-	\label{fig:signum_impl}
+	\includegraphics[width=0.6\textwidth]{../n4}
+	\caption[\acs{Nexys} development board]{\acs{Nexys} development board.}
+	\label{fig:n4}
 \end{figure}
 
-\section{Measurement Setup}\label{section:measurement_setup}
+Measurements were performed using an Agilent MSO7054A, a four channel, $4~\si[per-mode=symbol]{\giga\sample\per\second}$ oscilloscope.
+Signals were extracted over the regular \ac{PMOD} headers as the board does not provide outputs better suited to the measurement of signals that are available to the user, such as SMA or BNC connectors, and the \ac{PMOD} header for the analog-to-digital converter, which has lower impedance traces, is not available for the designer to re-configure.
+Data was acquired with the oscilloscope in quad channel mode at a sampling rate of $2\si{\giga\sample\per\second}$.
+One of the four channels was consumed by the external reference, leaving three open for signal measurements.
+$200\si{\micro\second}$ long captures were saved to an Agilent specified binary format which could be opened in Matlab for analysis.
+%TODO diagram or picture of some sort
 
 \section{\acs{ADPLL} Characterisation}
+%TODO why each PLL
 

### 2019-04-17 matlab/gain_grid.m

@@ -1,0 +1,79 @@
+for inc = 1:2 %kp loop
+    for inner_inc = 1:15 %ki_loop
+
+        kp = 8+2*(inc-1);
+        ki = inner_inc;
+
+        path3       = "test_conor/tuesday/indep_plls3x3_5m12_kp_" + num2str(kp) + "_ki_" + num2str(ki) + ".bin";
+        disp(path3);
+        fileID      = fopen(path3);
+        A           = fread(fileID, 'double');
+        fclose(fileID);
+        N           = 4; 
+        step        = 2e6;
+        width_vect  = 4e6;
+        ss          = [];
+        for chan=0:N-1 %start at 0 for chan1
+            deb = chan*width_vect+1;
+            ss = [ss A(deb:deb+step-1, 1) A(deb+step:deb+2*step-1, 1)];
+        end
+        clear A
+
+        scope_x_data = ss(:,1);
+        scope_y_data1 = ss(:,2);
+        scope_y_data2 = ss(:,4);
+        scope_y_data3 = ss(:,6);
+        scope_y_data4 = ss(:,8);
+
+        [~,init_cross(1,:),~,~] = pulsewidth(scope_y_data1,scope_x_data, 'Polarity', 'Positive');
+        [~,init_cross(2,:),~,~] = pulsewidth(scope_y_data2,scope_x_data, 'Polarity', 'Positive');
+        [~,init_cross(3,:),~,~] = pulsewidth(scope_y_data3,scope_x_data, 'Polarity', 'Positive');
+        [~,init_cross(4,:),~,~] = pulsewidth(scope_y_data4,scope_x_data, 'Polarity', 'Positive');
+
+        periods(1,:) = getPeriods(init_cross(1,:)); 
+        periods(2,:) = getPeriods(init_cross(2,:));
+        periods(3,:) = getPeriods(init_cross(3,:));
+        periods(4,:) = getPeriods(init_cross(4,:));
+
+        std_period(2,inc,inner_inc) = std(periods(2,:));
+        std_period(3,inc,inner_inc) = std(periods(3,:));
+        std_period(4,inc,inner_inc) = std(periods(4,:));
+        
+        if size(init_cross(2,:),2) == size(init_cross(1,:),2)
[…]

### 2019-04-17 thesis/NetworkImplemenation.tex
58.3

@@
 The operation of the \ac{TDL} is as follows: A delayed version of the \texttt{count} signal acts as the clock for flip-flops forming each tap, which are rising edge activated.
-These flip-flops have a constant ``1'' at their $D$ inputs, so as the signal propagates through the delay line, each flip-flop with propagate a logic ``1'' to their corresponding \texttt{error_tap[i]} signal.
-When the measurement interval has completed, the \texttt{count} signal clocks the 15 bit wide \texttt{error_tap} signal into \texttt{error_tap_buff} which act as the output buffers of the \ac{PD}, holding the phase error constant between measurement intervals.
-As with the flip-flops performing edge detection, the flip-flops forming the taps are can only be used once before requiring a reset, which is again carried out using the \texttt[done] signal.
+These flip-flops have a constant ``1'' at their $D$ inputs, so as the signal propagates through the delay line, each flip-flop with propagate a logic ``1'' to their corresponding \texttt{error\_tap[i]} signal.
+When the measurement interval has completed, the \texttt{count} signal clocks the 15 bit wide \texttt{error\_tap} signal into \texttt{error\_tap\_buff} which act as the output buffers of the \ac{PD}, holding the phase error constant between measurement intervals.
+As with the flip-flops performing edge detection, the flip-flops comprising the taps can only be used once before requiring a reset, which is again carried out using the \texttt{done} signal.
+As flip-flops have a hold time, the duration after the clock edge in which the signal applied at the $D$ input must not change to ensure valid data, an inverter is used to delay the reset to avoid this issue.
+
+The output of the \ac{TDL} itself is not a binary signal, but rather the thermometer coded representation of the number of taps activated in the measurement interval.
+In order to convert to binary, the easiest solution is a switch statement, with an assignment performed based on the temperature code.
+However all \ac{ADPLL} blocks have been designed to be extensible, requiring the lookup to be generated at compile time.
+This is somewhat difficult to implement, and a simpler solution exists for this problem.
+In a \texttt{generate} statement, a variable size loop is used to count the number of nonzero bits as it would be done in C.
+At compile time this loop is converted into lookup tables, similarly to the switch statement.
+
+The now unsigned binary representation of the error magnitude can be combined then with the sign bit to form a two's complement error signal.
+The file \textsc{PhaseDetectorDL.v} in Appendix \ref{adx:code} Listing \ref{lst:phase_detector_dl} contains the Verilog implementation of this module.
+\texttt{DONT\_TOUCH} compiler directives were used here in a number of places to ensure that the inverters and other elements that are key for timing are not removed in the synthesis process, as similarly to \ac{RO}, Vivado sees these as unneeded elements serving no purpose.
 
@@
 Data was acquired with the oscilloscope in quad channel mode at a sampling rate of $2\si{\giga\sample\per\second}$.
-One of the four channels was consumed by the external reference, leaving three open for signal measurements.
+One of the four channels was consumed by the external reference, leaving three free for signal measurements, in most cases the generated output of three \ac{ADPLL}s were connected.
 $200\si{\micro\second}$ long captures were saved to an Agilent specified binary format which could be opened in Matlab for analysis.

### 2019-04-17 thesis/NetworkImplemenation.tex
58.5

@@
 Important to note is the shift applied to the result of the multiplication by \acs{kp} which, as the integration is performed with a larger width to avoid accumulator overflow, preserves the intended relationship between proportional and integral gains.
+The value of $k_p$ and $k_i$ can be computed using the following formulae:
+\begin{gather}
+\text{Proportional gain} = k_p\times 2^{-path\_width\_change\_p} \\
+path\_width\_change\_p = lsh\_width - rsh\_width \\
+lsh\_width = ki\_width - kp\_width  \\
+rsh\_width = error\_width + kp\_width + lsh\_width - control\_code\_width \\
+\text{Integral gain} = k_i\times 2^{-path\_width\_change\_i} \\
+path\_width\_change\_i = - rsh\_width
+\end{gather}
+In the example system shown in Figure \ref{fig:integer_lf} the proportional gain can be computed as $k_p\times 2^{3-8} = \frac{k_p}{32}$ and the integral gain is $k_i\times 2^{-8} = \frac{k_i}{128}$.
 The Verilog implementation of the \ac{LF} using integer arithmetic can be found in \textsc{LoopFilter.v}, attached in Appendix \ref{adx:code} Listing \ref{lst:loop_filter}.
@@
 	\centering
-	\subfloat[Vivado simulation.]{{\includegraphics[width=0.45\textwidth]{../sim_locking} }}\\
+	\subfloat[Vivado simulation.]{{\includegraphics[width=0.45\textwidth]{../sim_locking} }}%
 	\subfloat[Hardware measurement.]{{\includegraphics[width=0.45\textwidth]{../impl_locking} }}%
@@
 			\hline
+			\acs{kp} width & $2^{-5}\times[0,15]\cap\mathbb{Z}$ & \acs{ki} & $2^{-8}\times[0,15]\cap\mathbb{Z}$ \T\\
 			\acs{kp} & $2^{-5}\times[0,15]\cap\mathbb{Z}$ & \acs{ki} & $2^{-8}\times[0,15]\cap\mathbb{Z}$ \T\\
@@
 The frequency range is then:
-\begin{align} %TODO inconsisentl calculation approach with previous version
+\begin{align} 
 f_{osc} &= \frac{1}{T_{osc}} = \frac{1}{(n-2CC)\times 2\times\tau_{inv}} \\
@@
 
-\begin{table}[!h]%TODO redo numbers!
+\begin{table}[!h]
 	\begin{center}
@@
 			\hline
-			\acs{kp} & $2^{-7}\times[0,15]\cap\mathbb{Z}$ & \acs{ki} & $2^{-9}\times[0,15]\cap\mathbb{Z}$ \T\\
+			\acs{kp} width & $4$ & \acs{ki} & $6$ \T\\
+			\hline
+			\acs{kp} & $2^{-4}\times[0,15]\cap\mathbb{Z}$ & \acs{ki} & $2^{-6}\times[0,15]\cap\mathbb{Z}$ \T\\
 			\hline
@@
 \subsubsection{\acs{ADPLL} 3 Design Summary}
-Table \ref{table:adpll3}.
+Compared to the oscillators suggested earlier, \ac{ADPLL} 3 achieves a significantly finer detection resolution, at the expense of significant variation in the detection step.
+This can be viewed as advantageous however, as the \ac{ADPLL} exhibits an exaggerated version of the implementation and manufacturing defect based variability present in an \ac{ASIC}.
+It is this type of \ac{ADPLL} that both the \acs{UCD} and Sorbonne research teams are using to examine potential designs and test theoretical results.
+As before, Table \ref{table:adpll3} presents the important numeric characteristics of this \ac{ADPLL}.
 
-\begin{table}[!h]%TODO numbers!
+\begin{table}[!h]
 	\begin{center}
@@
 			\hline
-			\acs{kp} & $2^{-7}\times[0,15]\cap\mathbb{Z}$ & \acs{ki} & $2^{-9}\times[0,15]\cap\mathbb{Z}$ \T\\
+			\acs{kp} width & $4$ & \acs{ki} & $6$ \T\\
 			\hline
-			Detection Resolution & $3.875~\si{\nano\second}$ & Detection Phase Resolution & $6.975\si{\degree}$ (at $5~\si{\mega\hertz}$)\\
+			\acs{kp} & $2^{-4}\times[0,15]\cap\mathbb{Z}$ & \acs{ki} & $2^{-6}\times[0,15]\cap\mathbb{Z}$ \T\\
+			\hline
+			Detection Resolution & $\approx0.63~\si{\nano\second}$ & Detection Phase Resolution & $1.134\si{\degree}$ (at $5~\si{\mega\hertz}$)\\
 			\hline
@@
 
+% //9 for older tests @1111 = 4@0000, but >> 5 cos of extra size
+% <4 is a no can do //7 for older tests @1111 = 4@0001
+
 \section{Minor Variations}\label{section:minor_variations}

### 2019-04-18 thesis/NetworkImplemenation.tex
per
62

@@
 Signals were extracted over the regular \ac{PMOD} headers as the board does not provide outputs better suited to the measurement of signals that are available to the user, such as SMA or BNC connectors, and the \ac{PMOD} header for the analog-to-digital converter, which has lower impedance traces, is not available for the designer to re-configure.
-Data was acquired with the oscilloscope in quad channel mode at a sampling rate of $2\si{\giga\sample\per\second}$.
+Data was acquired with the oscilloscope in quad channel mode at a sampling rate of $2\si[per-mode=symbol]{\giga\sample\per\second}$.
 One of the four channels was consumed by the external reference, leaving three free for signal measurements, in most cases the generated output of three \ac{ADPLL}s were connected.
@@
 \section{\acs{ADPLL} Characterisation}
-%TODO why each PLL
+Before combining the \ac{ADPLL}s in a network it was important to analyse the designs locked to an ideal reference, in order to establish a baseline performance.
+These tests were carried out with the \ac{ADPLL}s locked to an external reference at $5~\si{\mega\hertz}$.
+As \ac{ADPLL} 2 \& 3 have significant variance in the step sizes, for these oscillators range and step based measurements were obtained by measuring a number of oscillators, and typical values will be shown.
+Jitter was calculated based on the performance on nine individual \ac{ADPLL}s all locked simultaneously to the $5~\si{\mega\hertz}$ external reference.
+While jitter and skew are the key indicators of \ac{ADPLL} network performance, when it comes to the individual \ac{ADPLL}s themselves it is important to check lock and capture ranges also, to check that the implemented designs match with their intended values.
+
+The \label{table:characterisation} contains the results of this analysis.
+The first thing to note is that to the $\si{\kilo\hertz}$ the lock and capture ranges of the \ac{ADPLL}s implementing \ac{RO}s are identical.
+As the lock range would be expected to be slightly wider than the capture range in a design limited by the loop filter, it can be deduced that the limiting factor in the tuning range of the \ac{ADPLL} is the \ac{DCO} frequency range.
+This is confirmed by a cursory comparison against the values in Table \ref{table:adpll2}, where the \ac{RO} range is given as $4.571-5.518~\si{\mega\hertz}$.
+This range of $0.947~si{\mega\hertz}$ is almost identical to those of the \ac{ADPLL} $1$ \& $3$ which confirms the \ac{RO} is the limiting factor.
+In \ac{ADPLL} 1 however, the lock and capture ranges show some difference, with the lock range being slightly wider.
+The lower end of both ranges is identical, as rather than the instant loss of lock seen at all other thresholds, the low end of both lock and capture ranges saw a more gradual degradation, and $1.08~\si{\mega\hertz}$ is the point that the lock was observed to be lost.
+
+\begin{table}[]
+    \begin{centre}
+        \begin{tabular}{lrrr}
+        \multicolumn{1}{c}{}          & \multicolumn{1}{c}{ADPLL Design 1}  & \multicolumn{1}{c}{ADPLL Design 2} & \multicolumn{1}{c}{ADPLL Design 3} \\ \cline{2-4}       
+        Capture Range                 & $1.08-8.61~\si{\mega\hertz}$       & $4.869-5.877~\si{\mega\hertz}$    & $4.377-5.363~\si{\mega\hertz}$       \\
+        Lock Range                    & $1.08-8.635~\si{\mega\hertz}$      & As above                           & As above                            \\
+        Jitter                        & $1.8385~\si{\nano\second}$         & $0.5623~\si{\nano\second}$         & $0.7410~\si{\nano\second}$          \\
+        Skew (Mean \acs{TIE} to Ref.) & $7.7939~\si{\nano\second}$         & $6.2101~\si{\nano\second}$         & $5.8145~\si{\nano\second}$          \\
+        Maximum \acs{TIE}             & $8.7847~\si{\nano\second}$         & $6.4770~\si{\nano\second}$         & $6.4578~\si{\nano\second}$          \\
+        $k_p$                         & $\frac{1}{32}$                     & $\frac{1}{16}$                     & $\frac{1}{16}$                      \\
+        $k_i$                         & $\frac{1}{256}$                    & $\frac{1}{64}$                     & $\frac{1}{64}$  
+        \end{tabular}
+    \end{centre}
+    \caption[\ac{ADPLL} Design Comparison]{\ac{ADPLL} Design Comparison.}
+    \label{table:characterisation}
+\end{table}
+
+Looking then at the jitter and skew data, unsurprisingly the entirely \ac{FPGA} clocked \ac{ADPLL} with its larger period step is has the worst jitter, by a significant amount.
+The maximum \ac{TIE} is also correspondingly worse, which unless significant skew was present in the other \ac{ADPLL} designs is also to be expected.
 
@@
 \section{Minor Variations}\label{section:minor_variations}
+
+\subsection{Impact of Gain Variation}
+The gains chosen for use in the \acl{LF} have significant impact on the performance of the system, as the \ac{LF} governs the response of the \ac{ADPLL} to a given phase error.
+To large a proportional gain will result in overly aggressive overcompensation, thus negatively affecting the jitter.
+Eventually the violence of the jitter will result in a loss of lock.
+However, if the gain is too small, the frequency or period adjustment will not be sufficiently large to cause the \ac{ADPLL} to lock.
+Similarly, appropriate values of $k_i$ are required, and the relationship between proportional and integral gains must satisfy the conditions set out by Koskin \cite{koskin2018generation}.
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=0.6\textwidth]{../fixed_kp.png}
+    \caption[Fixed $k_p$ integral gain sweeps]{\Fixed $k_p$ integral gain sweeps.}
+    \label{fig:gain_sweep}
+\end{figure}
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=0.6\textwidth]{../fixed_kp_zoom.png} 
+    \caption[Fixed $k_p$ integral gain sweeps]{\Fixed $k_p$ integral gain sweeps.}
+    \label{fig:gain_sweep_zoom}
+\end{figure}
+
+Figure \ref{fig:gain_sweep} contains a sweep across the possible range of integral gains for four different proportional gains.
+It can clearly be observed that there are two distinct regions in the graph, with jitter values of less than $1~\si{\nano\second}$ contrasting sharply with the points with jitter in the region of $10~\si{\nano\second}$.
+These distinct sections of the plot represent locked and unlocked operation respectively.
+Zooming in on the low jitter region in Figure \ref{fig:gain_sweep_zoom}, the loss of lock thresholds can be compared with the theoretical relationship of $k_i = 0.5\times k_p$.
+\begin{table}[!h]
+    \begin{center}
+        \begin{tabular}{lrr}
+            \multicolumn{1}{c}{$k_p$} & \multicolumn{1}{c}{$k_i$} & \multicolumn{1}{c}{Ratio} \T\B\\
+            \hlines
+            $\frac{1}{16}$            & $\frac{5}{64}$            & $0.8$                     \T\B\\
+            $\frac{2}{16}$            & $\frac{10}{64}$           & $0.8$                     \T\B\\
+            $\frac{3}{16}$            & $\frac{12}{64}$           & $1$                       \T\B\\
+            $\frac{4}{16}$            & $\frac{14}{64}$           & $1.1428$                  \T\B
+        \end{tabular}
+    \end{center}
+    \vspace{-0.5cm}
+    \caption[Loss of Lock Gains]{Loss of Lock Gains.}
+    \label{table:lolgains}
+\end{table}
+
+Despite the relatively small number of proportional gains, it is clear that the relationship set out by Koskin does not hold in this system, but rather a scaled version of it.
+This difference is attributable to a difference in \ac{LF} structure.
+In his simulations the \ac{LF} design contains a register forming an input buffer, absent in the \ac{LF} used for these tests.
+The presence of this register, clocked on the generated signal, causes the \ac{ADPLL} to react more slowly to changes in the phase difference, thus increasing the jitter.
+The removal of this register causes a change in the relationship between proportional and integral gains, a change that will be investigated further in a later section.
+
+\subsection{Distribution of Period Steps}
+To add context to the jitter numbers produced for each oscillator previously, the variation of period steps in the \ac{RO}, and delays in the \ac{TDL} due to implementation should be examined.
+The best way to visualise this distribution is through a histogram, placing periods of in a certain range into the same bin.
+This can be visualised effectively using Matlab's \texttt{histplot}.
+In order to obtain data, three \ac{ADPLL}s implementing both the \ac{FPGA} clocked and inverter based \ac{DCO}s were locked to an external reference signal at $5~\si{\mega\hertz}$, and the period of each clock cycle measured.
+\begin{figure}[h]%
+    \centering
+    \subfloat[Inverted delay based \ac{RO}.]{{\includegraphics[width=0.5\textwidth]{../conf_paper/distrib_ring} }}%
+    \subfloat[\ac{FPGA} clocked \ac{DCO}.]{{\includegraphics[width=0.5\textwidth]{../conf_paper/distrib_pa} }}%
+    \caption[Example distribution of periods]{Example distribution of periods.}    
+    \label{fig:dists}
+\end{figure}
+Figure \ref{fig:dists} contains the output of \texttt[histplot} for each of the three oscillators superimposed in one plot for each \ac{DCO} type.
+As the \ac{RO} has significantly finer period resolution, the period steps would be overlaid on each-other thereby making the plot difficult to read.
+In order to avoid this behaviour, the \ac{RO}'s mapping of control code to inverters to remove was altered from twice the control code to 16 times the control code.
+This results in a more pronounced separation of period steps while being indicative of the variation due to implementation.
+The plots paint a a picture of the unrealistic nature of the \ac{FPGA} clocked design in comparison to its inverter based counterpart, as each oscillator will have identical intrinsic period steps thus eliminating any jitter that maybe be seen in the system due to misalignment.
+This effects the suitability of the FPGA clocked \ac{DCO} for the improvement of simulation models, or the examination of how a novel block may impact the behaviour of the network.
+If this analysis was carried out for the period steps, similar behaviour would be observed.
+
+%TODO removal of delay
 %TODO osc size
@@
 \section{\acs{ADPLL} Network Implementation}
+%TODO why use each PLL
 \subsection{2x2 Network}

### 2019-04-18 matlab/a1_3.m

@@ -1,8 +1,8 @@
-for inc = 1:15
+for inc = 1:10
     %filename = char("conor_test/settings1_" + num2str(inc) + ".bin");
     %filename = char("scope_41.bin");
     
-    path3       = "test_conor/monday/design2b_" + num2str(inc) + ".bin";
+    path3       = "~/gdrive/measurements/monday2/adpll_1b" + num2str(inc) + ".bin";
     fileID      = fopen(path3);
     A           = fread(fileID, 'double');
     fclose(fileID);
@@ -76,21 +76,35 @@ for inc = 1:15
 %     p2p_4 = max_4-min_4;
 % 
     if length(init_cross2) == length(init_cross1)
-        TIE(2,inc) = max(abs(init_cross2(1:len)-init_cross1));
+        TIE_max(2,inc) = max(abs(init_cross2(1:len)-init_cross1));
     elseif length(init_cross2) > length(init_cross1)
-        TIE(2,inc) = max(abs(init_cross2(2:end)-init_cross1));
+        TIE_max(2,inc) = max(abs(init_cross2(2:end)-init_cross1));
     end
     if length(init_cross3) == length(init_cross1)
-        TIE(3,inc) = max(abs(init_cross3(1:len)-init_cross1));
+        TIE_max(3,inc) = max(abs(init_cross3(1:len)-init_cross1));
     elseif length(init_cross3) > length(init_cross1)
-        TIE(3,inc) = max(abs(init_cross3(2:end)-init_cross1));
-        inc
+        TIE_max(3,inc) = max(abs(init_cross3(2:end)-init_cross1));
+    end
+    if length(init_cross4) == length(init_cross1)
+        TIE_max(4,inc) = max(abs(init_cross4(1:len)-init_cross1));
+    elseif length(init_cross4) > length(init_cross1)
+        TIE_max(4,inc) = max(abs(init_cross4(2:end)-init_cross1));
+    end
+    
+    if length(init_cross2) == length(init_cross1)
+        TIE(2,inc) = max((init_cross2(1:len)-init_cross1));
+    elseif length(init_cross2) > length(init_cross1)
+        TIE(2,inc) = max((init_cross2(2:end)-init_cross1));
+    end
+    if length(init_cross3) == length(init_cross1)
+        TIE(3,inc) = max((init_cross3(1:len)-init_cross1));
[…]

### 2019-04-18 matlab/gain_grid.m

@@ -1,10 +1,10 @@
-for inc = 1:2 %kp loop
+for inc = 1:4 %kp loop
     for inner_inc = 1:15 %ki_loop
 
-        kp = 8+2*(inc-1);
+        kp = inc;
         ki = inner_inc;
 
-        path3       = "test_conor/tuesday/indep_plls3x3_5m12_kp_" + num2str(kp) + "_ki_" + num2str(ki) + ".bin";
+        path3       = "~/gdrive/measurements/wednesday/indep_pll_3x3_5m00_kp_" + num2str(kp) + "_ki_" + num2str(ki) + ".bin";
         disp(path3);
         fileID      = fopen(path3);
         A           = fread(fileID, 'double');
@@ -25,35 +25,52 @@ for inc = 1:2 %kp loop
         scope_y_data3 = ss(:,6);
         scope_y_data4 = ss(:,8);
 
-        [~,init_cross(1,:),~,~] = pulsewidth(scope_y_data1,scope_x_data, 'Polarity', 'Positive');
-        [~,init_cross(2,:),~,~] = pulsewidth(scope_y_data2,scope_x_data, 'Polarity', 'Positive');
-        [~,init_cross(3,:),~,~] = pulsewidth(scope_y_data3,scope_x_data, 'Polarity', 'Positive');
-        [~,init_cross(4,:),~,~] = pulsewidth(scope_y_data4,scope_x_data, 'Polarity', 'Positive');
-
-        periods(1,:) = getPeriods(init_cross(1,:)); 
-        periods(2,:) = getPeriods(init_cross(2,:));
-        periods(3,:) = getPeriods(init_cross(3,:));
-        periods(4,:) = getPeriods(init_cross(4,:));
-
-        std_period(2,inc,inner_inc) = std(periods(2,:));
-        std_period(3,inc,inner_inc) = std(periods(3,:));
-        std_period(4,inc,inner_inc) = std(periods(4,:));
+%         [~,init_cross(1,:),~,~] = pulsewidth(scope_y_data1,scope_x_data, 'Polarity', 'Positive');
+%         [~,init_cross(2,:),~,~] = pulsewidth(scope_y_data2,scope_x_data, 'Polarity', 'Positive');
+%         [~,init_cross(3,:),~,~] = pulsewidth(scope_y_data3,scope_x_data, 'Polarity', 'Positive');
[…]

### 2019-04-18 matlab/test_conor/MATLAB_data_processing_pll_network_Pb.m

@@ -1,936 +1,0 @@
-clear all;
-clc;
-fclose all;
-format long
-
-disp('**********  OPENING FILES   ***********');
-
-file_type = 'binary';       % choice : 'binary', 'csv', 'dat' : Default Value = 'binary'
-       % is there a majority gate measurement in the file? 'on' or 'off' : Default Value = 'off'
-serie_or_append = 'append';  % as the other file, 'serie' or 'append'
-pll_order = [11];           % Default Value = [11]
-% pll_order = [11 12 13 23]; % which pll is opened
-
-for kp=1:1
-    kp
-    for stuff=1:length(pll_order);
-        clear R_EK_2
-        clear R_EK_coef
-        clear R_PB_2
-        clear R_PB_coef 
-        clear ss
-        clear PP
-        clear T_sig
-        stuff
-        if strcmp(file_type,'dat')
-            switch  N
-                case 3
-                    path = './ADPLL_network/measurement/ADPLL_3x1_network/kp=05_ki=1..13_64/';
-                    s1 = importdata(strcat(path, 'sig_div1.dat' ));
-                    s2 = importdata(strcat(path, 'sig_div2.dat'));
-                    s3 = importdata(strcat(path, 'sig_div3.dat'));
-                    s4 = importdata(strcat(path, 'sig_mg.dat'));
-                    s5 = importdata(strcat(path, 'sig_ref.dat'));
-                    ss = [s1 s2 s3 s4 s5];
-                case 5
-                    path = './measure/PLL_x5/serie33/';
-                    s1 = importdata(strcat(path, 'sig_div1.dat' ));
-                    s2 = importdata(strcat(path, 'sig_div2.dat'));
-                    s3 = importdata(strcat(path, 'sig_div3.dat'));
-                    s4 = importdata(strcat(path, 'sig_div4.dat'));
-                    s5 = importdata(strcat(path, 'sig_div5.dat'));
-                    s6 = importdata(strcat(path, 'sig_mg.dat'));
-                    s7 = importdata(strcat(path, 'sig_ref.dat'));
-                    ss = [s1 s2 s3 s4 s5 s6 s7];
-                case 7
-                    path = './measure/vacances7PLL/';
-                    s1 = importdata(strcat(path, 'sig_div1.dat' ));
-                    s2 = importdata(strcat(path, 'sig_div2.dat'));
-                    s3 = importdata(strcat(path, 'sig_div3.dat'));
[…]

### 2019-04-18 matlab/test_conor/fft_Bisiaux.m

@@ -1,15 +1,0 @@
-function [Xfft, Yfft] = fft_Bisiaux(Fs,fft_pts,x_input)
-%FFT_BISIAUX Summary of this function goes here
-%   Detailed explanation goes here
-% Nfft        = pow2(14); %length(v_modul);%   
-% Xfft_moi    = Fs*(0:Nfft/2-1)/Nfft;
-% Yfft        = fft(v_modul, Nfft);
-% Yfft        = Yfft(1:Nfft/2);
-% Yfft        = abs(Yfft)/length(Yfft);
-Nfft        = fft_pts; %length(v_modul);%   
-Xfft        = Fs*(0:Nfft/2-1)/Nfft;
-Yfft        = fft(x_input, Nfft);
-Yfft        = Yfft(1:Nfft/2);
-Yfft        = abs(Yfft)/length(Yfft);
-end
-

### 2019-04-18 matlab/test_conor/hdlcoder_board_customization.m

@@ -1,12 +1,0 @@
-function r = hdlcoder_board_customization
-% Board plugin registration file
-% 1. Any registration file with this name on MATLAB path will be picked up
-% 2. Registration file returns a cell array pointing to the location of
-%    the board plugins
-% 3. Board plugin must be a package folder accessible from MATLAB path,
-%    and contains a board definition file
-r = { ...
-      'ZyboRegistration.plugin_board', ...
-     };
- boardName = 'ZYBO';
-end

### 2019-04-18 matlab/test_conor/importAgilentBin.m

@@ -1,90 +1,0 @@
-function [timeVector, voltageVector] = importAgilentBin(inputFilename, varargin)
-% ImportAgilentBin reads the Agilent Binary Waveform filetype.
-% [timeVector, voltageVector] = importAgilentBin(inputFilename)
-% [timeVector, voltageVector] = importAgilentBin(inputFilename, waveform_index)
-% if waveformIndex is not provided, the first waveform will be read
-% voltageVector may contain two columns [MIN, MAX]
-
-if (~exist(inputFilename))
-    error('inputFilename missing.');
-end
-
-fileId = fopen(inputFilename, 'r');
-
-% read file header
-fileCookie = fread(fileId, 2, 'char');
-fileVersion = fread(fileId, 2, 'char');
-fileSize = fread(fileId, 1, 'int32');
-nWaveforms = fread(fileId, 1, 'int32');
-
-% verify cookie
-fileCookie = char(fileCookie');
-if (~strcmp(fileCookie, 'AG'))
-    fclose(fileId);
-    error('Unrecognized file format.');
-end
-
-% determine which waveform to read
-waveformSelect = 1;
-if (size(varargin) == 1 & varargin{1} <= nWaveforms)
-    waveformSelect = varargin{1};
-end
-
-for waveformIndex = 1:nWaveforms
-    % read waveform header
-    headerSize = fread(fileId, 1, 'int32'); bytesLeft = headerSize - 4;
-    waveformType = fread(fileId, 1, 'int32'); bytesLeft = bytesLeft - 4;
-    nWaveformBuffers = fread(fileId, 1, 'int32'); bytesLeft = bytesLeft - 4;
-    nPoints = fread(fileId, 1, 'int32'); bytesLeft = bytesLeft - 4;
-    count = fread(fileId, 1, 'int32');  bytesLeft = bytesLeft - 4;
-    xDisplayRange = fread(fileId, 1, 'float32');  bytesLeft = bytesLeft - 4;
-    xDisplayOrigin = fread(fileId, 1, 'double');  bytesLeft = bytesLeft - 8;
-    xIncrement = fread(fileId, 1, 'double');  bytesLeft = bytesLeft - 8;
-    xOrigin = fread(fileId, 1, 'double');  bytesLeft = bytesLeft - 8;
-    xUnits = fread(fileId, 1, 'int32');  bytesLeft = bytesLeft - 4;
-    yUnits = fread(fileId, 1, 'int32');  bytesLeft = bytesLeft - 4;
[…]

### 2019-04-18 matlab/test_conor/plugin_board.m

@@ -1,34 +1,0 @@
-function hB = plugin_board()
-% Board definition
-% Construct board object
-hB = hdlcoder.Board;
-hB.BoardName    = 'ZYBO';
-% FPGA device information
-hB.FPGAVendor   = 'Xilinx';
-hB.FPGAFamily   = 'Zynq';
-hB.FPGADevice   = 'xc7z010';
-hB.FPGAPackage  = 'clg400';
-hB.FPGASpeed    = '-1';
-% Tool information
-hB.SupportedTool = {'Xilinx Vivado'};
-% FPGA JTAG chain position
-hB.JTAGChainPosition = 2;
-%% Add interfaces
-% Standard "External Port" interface
-hB.addExternalPortInterface( ...
-    'IOPadConstraint', {'IOSTANDARD = LVCMOS33'});
-% Custom board external I/O interface
-hB.addExternalIOInterface( ...
-  'InterfaceID',    'LEDs General Purpose', ...
-  'InterfaceType',  'OUT', ...
-  'PortName',       'LEDs', ...
-  'PortWidth',      4, ...
-  'FPGAPin',        {'M14', 'M15', 'G14', 'D18'}, ...
-  'IOPadConstraint', {'IOSTANDARD = LVCMOS33'});
-hB.addExternalIOInterface( ...
-   'InterfaceID',    'Push Buttons', ...
-   'InterfaceType',  'IN', ...
-   'PortName',       'PushButtons', ...
-   'PortWidth',      4, ...
-   'FPGAPin',        {'R18', 'P16', 'V16', 'Y16'}, ...
-   'IOPadConstraint', {'IOSTANDARD = LVCMOS33'});

### 2019-04-18 matlab/test_conor/plugin_rd.m

@@ -1,22 +1,0 @@
-function hRD = plugin_rd()
-% Reference design definition
-% Construct reference design object
-hRD = hdlcoder.ReferenceDesign('SynthesisTool', 'Xilinx Vivado');
-hRD.ReferenceDesignName = 'Demo system';
-hRD.BoardName = 'ZYBO';
-% Tool information
-hRD.SupportedToolVersion = {'2017.2','2017.4'};
-% add custom Vivado design
-hRD.addCustomVivadoDesign( ...
-    'CustomBlockDesignTcl', 'design_led.tcl', ...
-    'VivadoBoardPart','digilentinc.com:zybo:part0:1.0');
-%% Add interfaces
-% add clock interface
-hRD.addClockInterface( ...
-  'ClockConnection',   'clk_wiz_0/clk_out1', ...
-  'ResetConnection',   'proc_sys_reset_0/peripheral_aresetn');
-% add AXI4 and AXI4-Lite slave interfaces
-hRD.addAXI4SlaveInterface( ...
-    'InterfaceConnection', 'axi_interconnect_0/M00_AXI', ...
-    'BaseAddress',         '0x40010000', ...
-    'MasterAddressSpace',  'processing_system7_0/Data');

### 2019-04-19 thesis/NetworkImplemenation.tex
fixed skews

@@
 While jitter and skew are the key indicators of \ac{ADPLL} network performance, when it comes to the individual \ac{ADPLL}s themselves it is important to check lock and capture ranges also, to check that the implemented designs match with their intended values.
+The lock and capture ranges are given based on typical values, whereas jitter and skew are the median values from the \ac{ADPLL}s tested.
 
@@
         Jitter                        & $1.8385~\si{\nano\second}$         & $0.5623~\si{\nano\second}$         & $0.7410~\si{\nano\second}$          \\
-        Skew (Mean \acs{TIE} to Ref.) & $7.7939~\si{\nano\second}$         & $6.2101~\si{\nano\second}$         & $5.8145~\si{\nano\second}$          \\
-        Maximum \acs{TIE}             & $8.7847~\si{\nano\second}$         & $6.4770~\si{\nano\second}$         & $6.4578~\si{\nano\second}$          \\
+        Skew (Mean \acs{TIE} to Ref.) & $1.7439~\si{\nano\second}$         & $0.3741~\si{\nano\second}$         & $0.6381~\si{\nano\second}$          \\
+        Maximum \acs{TIE}             & $7.2840~\si{\nano\second}$         & $6.4770~\si{\nano\second}$         & $6.4578~\si{\nano\second}$          \\
         $k_p$                         & $\frac{1}{32}$                     & $\frac{1}{16}$                     & $\frac{1}{16}$                      \\
@@
 The maximum \ac{TIE} is also correspondingly worse, which unless significant skew was present in the other \ac{ADPLL} designs is also to be expected.
+\ac{ADPLL} 2 \& 3 experience low jitter, at less than $1\%$ of the period.
+Somewhat surprisingly there is no apparent benefit in terms of jitter reduction due to the better delay resolution of the inverter based \ac{TDL}.
+This may be down to either the increase in jitter from a second source of variation, or a less favourable fractional-N synthesis due to the distribution of periods.
+
+Despite Vivado being forced to use the same layout for both \ac{ADPLL} 2 \& 3 through the use of fixed cells, the minimum and maximum frequencies are still significantly different, although the size of the range is almost identical.
+This highlights the difficulty of ensuring identical conditions between measurements which is the most significant downside of inverter based modules.
 
@@
 %TODO removal of delay
+%TODO early
 %TODO osc size
@@
 \section{\acs{ADPLL} Network Implementation}
+As the \ac{ADPLL}s were designed with network usage in mind, the \ac{HDL} aspect of their integration into a network is straightforward.
+For \ac{ADPLL} 1, the oscillators need only be wired together, but for Design 2 \& 3 which are reliant on inverters require some care to ensure their lock ranges overlap sufficiently.
+This is a time consuming task, as a number of iterations may be required, even when locking down the cells of lined up \ac{RO}s, as the nets forming the connections may change.
+It is recommended only to carry out alignment when all other aspects of the design are unlikely to change and to save the bitstream generated.
+The projected time requirement in aligning \ac{RO}s meant that the creation of a 4x4 network was abandoned.
+\ac{FPGA} clocked \ac{ADPLL}s are more suited to larger networks as their centre point can be determined with ease.
+
+The network operates using an external reference which is input over a \ac{PMOD} header that is shared with a number of output signals, however it could be configured to use an internal reference generated by either the \ac{FPGA} clock management utility or a \ac{DCO} of the designer's choice.
+
+In order to increase the flexibility of the network, the switches on the \acs{Nexys} are used to change both the proportional and integral gains and the configuration of the network at runtime.
+The reconfigurability aspect of the error combiner allows for the network to operate in either uni- or bi-directional mode, or as a number of individual \ac{PLL}s all connected to the external reference.
+This is particularly advantageous for the inverter based designs, as multiple combinations can be measured without reimplementing, and thus changing the layout of either oscillator or \acl{PD}.
+
 %TODO why use each PLL

### 2019-04-19 matlab/a1_3.m

@@ -1,8 +1,9 @@
-for inc = 1:10
+for inc = 1:5
     %filename = char("conor_test/settings1_" + num2str(inc) + ".bin");
     %filename = char("scope_41.bin");
     
-    path3       = "~/gdrive/measurements/monday2/adpll_1b" + num2str(inc) + ".bin";
+    %path3       = "~/gdrive/measurements/monday2/adpll_1b" + num2str(inc) + ".bin";
+    path3       = "~/gdrive/measurements/monday/design2b_" + num2str(inc) + ".bin";
     fileID      = fopen(path3);
     A           = fread(fileID, 'double');
     fclose(fileID);
@@ -92,19 +93,19 @@ for inc = 1:10
     end
     
     if length(init_cross2) == length(init_cross1)
-        TIE(2,inc) = max((init_cross2(1:len)-init_cross1));
+        TIE(2,inc) = mean((init_cross2(1:len)-init_cross1));
     elseif length(init_cross2) > length(init_cross1)
-        TIE(2,inc) = max((init_cross2(2:end)-init_cross1));
+        TIE(2,inc) = mean((init_cross2(2:end)-init_cross1));
     end
     if length(init_cross3) == length(init_cross1)
-        TIE(3,inc) = max((init_cross3(1:len)-init_cross1));
+        TIE(3,inc) = mean((init_cross3(1:len)-init_cross1));
     elseif length(init_cross3) > length(init_cross1)
-        TIE(3,inc) = max((init_cross3(2:end)-init_cross1));
+        TIE(3,inc) = mean((init_cross3(2:end)-init_cross1));
         end
     if length(init_cross4) == length(init_cross1)
-        TIE(4,inc) = max((init_cross4(1:len)-init_cross1));
+        TIE(4,inc) = mean((init_cross4(1:len)-init_cross1));
     elseif length(init_cross4) > length(init_cross1)
-        TIE(4,inc) = max((init_cross4(2:end)-init_cross1));
+        TIE(4,inc) = mean((init_cross4(2:end)-init_cross1));
     end
 
 

### 2019-04-20 thesis/Background.tex
79git statusgit statusgit statusgit statusgit statusgit statusgit statusgit statusgit statusgit statusgit statusgit status

@@
 \end{figure}
-
 Looking at the design of a given node it is notable that the function carried out by the Error Combiner is akin to a average, therefore as mentioned by Pratt and Nyugen, there is potential to a mode locked equilibrium in which the oscillators are not synchronised.
@@
 
-\section{\acl{ADPLL} Architecture}
+\section{\acs{ADPLL} Architecture}
 As indicated in Figure \ref{fig:mulkeen_pll} the three main building blocks of a conventional \ac{PLL} are the \ac{PD}, \ac{LF} and \ac{VCO}.
@@
 	\includegraphics[scale=0.5]{../tex_files/mulkeen_pll}
-	\caption[Block fiagram of a \acl{PLL}]{Block fiagram of a \acl{PLL}, \textit{Wireless Systems Notes}, B. Mulkeen (2017).}
+	\caption[Block fiagram of a \acl{PLL}]{Block diagram of a \acl{PLL}, \textit{Wireless Systems Notes}, B. Mulkeen (2017).}
 	\label{fig:mulkeen_pll}
@@
 	\centering
-	\includegraphics[scale=0.275]{../inverter_chain}
+	\includegraphics[width=0.6\textwidth]{../inverter_chain}
 	\caption[Basic Ring/Inverter Chain Oscillator]{Basic ring/inverter chain oscillator.}
@@
 	\centering
-	\includegraphics[scale=0.35]{../shan_bb_pd}
+	\includegraphics[width=0.6\textwidth]{../shan_bb_pd}
 	\caption[Bang-bang phase/frequency detector architecture]{Bang-bang phase/frequency detector architecture \cite{shan2014phd}.}
@@
 	\label{fig:my_simple_pi}
-\end{figure}%TODO ask mulkeen if this is IIR
+\end{figure}
 The \acl{LF} in an \ac{ADPLL} can be implemented as a \ac{PI} controller, as only a low-pass filter is required, such as that in Figure \ref{fig:my_simple_pi}.
@@
 \begin{equation*}
-	H(z) = \alpha + \beta\frac{1}{1-z^{-1}} = \frac{(\alpha + \beta) - \alpha z^{-1}}{1-z^{-1}}
+	H(z) = \alpha + \beta\frac{1}{1-z^{-1}} = \frac{(\alpha + \beta) - \alpha z^{-1}}{1-z^{-1}}\vspace{-0.2cm}
 \end{equation*}
@@
 In the same study a range of values was found which would produce low jitter operation of the network.
-As these values are all less than one, the filter must implement fixed point arithmetic in an effort to maintain the simplicity of the clock distribution network, rather than incurring the complexity penalty of floating point calculations.
+As these values are all less than one, the filter must implement fixed point arithmetic in an effort to maintain the simplicity of the clock distribution network, rather than incurring the penalty of floating point calculations.
 
@@
 
+A similar technique was used by Lata \textit{et al} (2013) to implement a \ac{FPGA} clocked oscillator with a $200~\si{\kilo\hertz}$ centre frequency, however it is hard to understand exactly what either the goal was or the specifics of the \ac{DCO} used \cite{lata2013adpll}.
+
 It is however possible to implement limited mixed-signal circuits on an \ac{FPGA} through the use of primitive logic elements, however, as control over the implementation is restricted to the module level it is not possible to mirror the implementation of a design intended for an \ac{ASIC}.
@@
 \ac{FPGA}s have been used in other fields to simulate and experiment with new technologies, although not all of these attempt to implement the mixed-signal circuitry using primitive elements.
-Fernandez-Alvarez \textit{et al} (2016) proposed a method suited to higher end \ac{FPGA}s in which a co-processor on the \ac{FPGA} simulates the mixed-signal circuitry while the digital section of the design is implemented on the \ac{FPGA} itself \cite{fernandez2017hw}.
+Fernandez-Alvarez \textit{et al} (2016) proposed a method for prototyping high-voltage system controllers suited to higher end \ac{FPGA}s in which a co-processor on the \ac{FPGA} simulates the mixed-signal circuitry while the digital section of the design is implemented on the \ac{FPGA} itself \cite{fernandez2017hw}.
 While the interfacing between hardware and software remains a challenge they found:
 \begin{quotation}
-	Obtained data are compared to the data obtained by means of using PSIM and ModelSim co-simulation.
+	Obtained data are compared to the data obtained by means of using ...
+simulation.
 The proposed solution speeds up the evaluation in around one order of magnitude keeping the accuracy.
@@
 the proposed system provides a versatile and fast method to develop ad hoc control architectures, avoiding the need for time-consuming mixed-signal simulations and the risk of damaging the actual power converter implementation.
-
 \end{quotation}
 By carrying out simulations on an \ac{FPGA}, Guanhua Wang \textit{et al} (2013) achieved a 3000 times decrease in run-time when compared to an identical simulation in MATLAB used for the verification of a calibration algorithm for successive approximation analogue-to-digital converters \cite{wang2013fast}.
-Many other examples exist of \ac{FPGA}s used for the simulation of mixed-signal or \acl{RF} circuits in literature.
+Many other examples can be found of \ac{FPGA}s used for the simulation of mixed-signal or \acs{RF} circuits.
 %TODO Elena had suggestions?
@@
 Long term, or accumulated, jitter represents the cumulative effect of jitter on the signal over several cycles.
-Long term jitter in particular affects \acs{RADAR} as it will manifest itself as a Doppler shift in the return signal.
+Long term jitter in particular affects \acs*{RADAR} as it will manifest itself as a Doppler shift in the return signal.
 For an \ac{PLL} network, with an appropriately designed filter, the long term jitter should be zero so long as the reference signal is stable.

### 2019-04-20 thesis/Conclusion.tex
? pages
79git statusgit statusgit statusgit statusgit statusgit statusgit statusgit statusgit statusgit statusgit statusgit status

@@
 \section{Conclusions}
-\section{Impact of Research}
 \section{Suggested Future Work}
+\subsection{\acl{TDL} Characterisation}
+\acreset{TDL}
+As previously mentioned, characterisation of the \ac{TDL} was not carried out due to the time it would have taken to perform the required measurements and tests.
+This characteristion can be accomplished through the use of an off-chip tunable delay.
+Off-chip is unfortunately a requirement, as the only delays available of the order required to test the step sizes are the variable delay of the inverters themselves.
+A resistor \& capacitor delay circuit can be used to perform these test, and the output of the \ac{PFD} analysed on a mixed-signal oscilloscope alongside the delayed and reference waveforms.
+The use of an off-chip delay element necessitates careful construction/cailbration of the measurement environment to eliminate the delay stemming from the signal path through input/output buffers as it leaves the chip.
+
+\subsection{Increased Network Size}
+A 4x4 network was a stretch-goal of this project, depending on the time remaining after once the minumim goal of a 3x3 network had been reached.
+Unfortuantely due to problems with the clocked phase detector which resulted in mode-locking, incorrect \ac{RO} inverter selection logic and the projected time requirement to ensure alignment of each \ac{RO} this was not achieved.
+A 4x4 network would better explain some dynamics, such as the skew growth as the \ac{ADPLL} indices increase.
+However.
+progressing beyond a 3x3 network means that less than $20\%$ of \ac{ADPLL} output signals can be examined at any given point in time.
+As such, it may be useful to develop a more advanced measurement set-up that can either capture more signals simultaneously, or implement a measurement system that makes more efficient use of the available probes.
+
+\subsection{\acs{FPGA} Clocked, Linear Period \acs{DCO}}
+\acreset{FPGA,DCO}
+While this design of \ac{DCO} was proposed earlier in this thesis, it was not implemented in any \ac{ADPLL}s, as it was only realised that the output waveform did not matter, just the location of the rising edges was not made until later in the project, once the pitfall encountered in the clocked \ac{PFD} design was resolved.
+Until this point the non-square wave output had eliminated the design from consideration.
+However, as in an \ac{ASIC} linear period designs are easier to implement, it would be ideal to have the benefit of frequency scaled hardware simulations with a \ac{DCO} that more closely relates the the eventual system.
+
+\subsection{Procedural Network Instantiation}
+This suggestion is directed at making the user experience simpler.
+As implemented, networks based on \ac{ADPLL}s containing \ac{RO}s are implemented in separate files to their \ac{FPGA} clocked \ac{DCO} based counterparts as are the networks of difference sizes.
+At a network size of 2x2 this was not a problem, but at the scale of 3x3 network the modules became somewhat cumbersome to work with, with a large number of interconnects.
+For 3x3 and larger networks it would be expedient to create the network with a \texttt{generate} statement, as this would reduce the number of lines of code and the amount of named interconnects required.
+Instead the user could set a number of configuration parameters and the network would be created accordingly.
+This could even allow for the use of different \ac{ADPLL} types in the same network with ease.
+Unfortunately the tasks of floorplanning and oscillator alignment would still be left to the user to perform.
+
+%TODO cahracterise tdl

### 2019-04-20 thesis/FPGABasedADPLLs.tex
79git statusgit statusgit statusgit statusgit statusgit statusgit statusgit statusgit statusgit statusgit statusgit status

@@
 The only variation in this design will come, ironically, from jitter or skew in the \ac{FPGA}'s clock distribution network, which as the frequencies will typically be in the low hundreds of MHz is very minor.
-In the case of the Xilinx \acl{Nexys} this is at most 100 picoseconds, or 0.05\% of the period of an intended output clock at 5 MHz.
+In the case of the Xilinx \acl*{Nexys} this is at most 100 picoseconds, or 0.05\% of the period of an intended output clock at 5 MHz.
 To put this value into perspective, on this board the minimum value of $T_{FPGA}$ that could be used to drive this oscillator is 3.87 nanoseconds, 1.935\% of the period.
@@
 
-\subsection{\ac{FIR} \acl{LF}}
-The \ac{FIR} implementation of the \acl{LF} will take the form presented in Figure \ref{fig:fir_pi} (a), and consists series of time delays, or ``taps'', the output of which is multiplied by a coefficient and the result of each multiplier added together to form the filter output.
+\subsection{\acs{FIR} \acl{LF}}
+The \acl{FIR} implementation of the \acl{LF} will take the form presented in Figure \ref{fig:fir_pi} (a), and consists series of time delays, or ``taps'', the output of which is multiplied by a coefficient and the result of each multiplier added together to form the filter output.
 Various tools, Matlab for example, can be used to compute the coefficients and number of taps required to obtain a specific filter response.
@@
 
-\subsection{\ac{IIR} \acl{LF}}
-A \ac{PI} controller can be easily implemented as an \ac{IIR} filter in hardware, as in Figure \ref{fig:iir_pi}, requiring significantly fewer circuit elements than an \ac{FIR} based design as the long series of taps and multipliers is not required, as instead integration is performed by a single multiplier and an accumulator.
+\subsection{\acs{IIR} \acl{LF}}
+A \ac{PI} controller can be easily implemented as an \acl{IIR} filter in hardware, as in Figure \ref{fig:iir_pi}, requiring significantly fewer circuit elements than an \ac{FIR} based design as the long series of taps and multipliers is not required, as instead integration is performed by a single multiplier and an accumulator.
 As before the frequency response of the controller can be calculated using the following equation, where $\alpha$ is the proportional and $\beta$ the integral gain:

### 2019-04-20 thesis/NetworkImplemenation.tex
79git statusgit statusgit statusgit statusgit statusgit statusgit statusgit statusgit statusgit statusgit statusgit status

--- 2019-04-17 thesis/NetworkImplemenation.tex
@@
             \hline
-			\acs{kp} width & $2^{-5}\times[0,15]\cap\mathbb{Z}$ & \acs{ki} & $2^{-8}\times[0,15]\cap\mathbb{Z}$ \T\\
+            \acs{kp} width & 5 & \acs{ki} & 8 \T\\
+            \hline
             \acs{kp} & $2^{-5}\times[0,15]\cap\mathbb{Z}$ & \acs{ki} & $2^{-8}\times[0,15]\cap\mathbb{Z}$ \T\\
@@
 The inverter delay and period step are then recomputed based on the centre frequency of $5~\si{\mega\hertz}$ as:
-\begin{align}%TODO 2 x 2 x midpoint justification - in prev. chapter?
+\begin{align}
 t_{step} &= \frac{T_{osc}}{0.5\times(n-\frac{k}{2})} = \frac{200\si{\nano\second}}{0.5\times 341} = 1.176~\si{\nano\second} \\
@@
 The module implementing the \ac{RO} can be found in \textsc{RingOsc.v}, attached in Appendix \ref{adx:code} Listing \ref{lst:ring_osc}.
+In the module instance statement it can be seen that a second generated output is also present.
+This second output is taken from a point in the fixed part of the chain, a significant number of inverters prior to the minimum possible due to control code adjustment.
+This signal is at an identical frequency to that of the generated signal, however will not be immediately affected by a change in the control code, which could in some edge cases change the polarity of the signal.
+This signal is used to drive the \ac{LF} in order to avoid an extraneous addition to the accumulator in the integral path.
+This early signal must be taken from an inverter at an index significantly lower than $n-k$ to prevent this addition causing the problem it has been implemented to avoid.
 
@@
 For the following sections in which the performance of either individual \ac{ADPLL}s or \ac{ADPLL} networks the same measurement setup was used.
-The \ac{FPGA} used was a Xilinx Artix-7 \acl{Nexys} on a Digilent \acs{Nexys} board.
-The \acl{Nexys} features 15850 logic slices, each with 4 lookup tables and 8 flip-flops.
+The \ac{FPGA} used was a Xilinx Artix-7 \acl*{Nexys} on a Digilent \acs*{Nexys} board.
+The \acl*{Nexys} features 15850 logic slices, each with 4 lookup tables and 8 flip-flops.
 There are six clock management tiles, 240 dedicated digital signal processing slices and an analog-to-digital converter.
@@
 Further information can be found in the datasheet \cite{a7_datasheet}.
-The \acs{Nexys} is an evaluation/development board for the aforementioned \ac{FPGA}, however directed at the education sector.
+The \acs*{Nexys} is an evaluation/development board for the aforementioned \ac{FPGA}, however directed at the education sector.
 It contains useful peripherals such as 16 switches, 8 7-segment displays and 4 user addressable \ac{PMOD} headers and can be used with the free of charge ``Webpack'' version of Vivado \cite{n4_datasheet}.
@@
     \includegraphics[width=0.6\textwidth]{../n4}
-	\caption[\acs{Nexys} development board]{\acs{Nexys} development board.}
+    \caption[\acs*{Nexys} development board]{\acs*{Nexys} development board.}
     \label{fig:n4}
@@
 Signals were extracted over the regular \ac{PMOD} headers as the board does not provide outputs better suited to the measurement of signals that are available to the user, such as SMA or BNC connectors, and the \ac{PMOD} header for the analog-to-digital converter, which has lower impedance traces, is not available for the designer to re-configure.
-Data was acquired with the oscilloscope in quad channel mode at a sampling rate of $2\si{\giga\sample\per\second}$.
+Data was acquired with the oscilloscope in quad channel mode at a sampling rate of $2\si[per-mode=symbol]{\giga\sample\per\second}$.
 One of the four channels was consumed by the external reference, leaving three free for signal measurements, in most cases the generated output of three \ac{ADPLL}s were connected.
 $200\si{\micro\second}$ long captures were saved to an Agilent specified binary format which could be opened in Matlab for analysis.
-%TODO diagram or picture of some sort
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=0.6\textwidth]{../measurement_setup}
+	\caption[Measurement setup]{Measurement setup.}
+	\label{fig:setup}
+\end{figure}
 
 \section{\acs{ADPLL} Characterisation}
-%TODO why each PLL
+Before combining the \ac{ADPLL}s in a network it was important to analyse the designs locked to an ideal reference, in order to establish a baseline performance.
+These tests were carried out with the \ac{ADPLL}s locked to an external reference at $5~\si{\mega\hertz}$.
+As \ac{ADPLL} 2 \& 3 have significant variance in the step sizes, for these oscillators range and step based measurements were obtained by measuring a number of oscillators, and typical values will be shown.
+Jitter was calculated based on the performance on nine individual \ac{ADPLL}s all locked simultaneously to the $5~\si{\mega\hertz}$ external reference.
+While jitter and skew are the key indicators of \ac{ADPLL} network performance, when it comes to the individual \ac{ADPLL}s themselves it is important to check lock and capture ranges also, to check that the implemented designs match with their intended values.
+The lock and capture ranges are given based on typical values, whereas jitter and skew are the median values from the \ac{ADPLL}s tested.
+
+The \ref{table:characterisation} contains the results of this analysis.
+The first thing to note is that to the $\si{\kilo\hertz}$ the lock and capture ranges of the \ac{ADPLL}s implementing \ac{RO}s are identical.
+As the lock range would be expected to be slightly wider than the capture range in a design limited by the loop filter, it can be deduced that the limiting factor in the tuning range of the \ac{ADPLL} is the \ac{DCO} frequency range.
+This is confirmed by a cursory comparison against the values in Table \ref{table:adpll2}, where the \ac{RO} range is given as $4.571-5.518~\si{\mega\hertz}$.
+This range of $0.947~\si{\mega\hertz}$ is almost identical to those of the \ac{ADPLL} $1$ \& $3$ which confirms the \ac{RO} is the limiting factor.
+In \ac{ADPLL} 1 however, the lock and capture ranges show some difference, with the lock range being slightly wider.
+The lower end of both ranges is identical, as rather than the instant loss of lock seen at all other thresholds, the low end of both lock and capture ranges saw a more gradual degradation, and $1.08~\si{\mega\hertz}$ is the point that the lock was observed to be lost.
+
+\begin{table}[]
+    \begin{center}
+        \begin{tabular}{lrrr}
+        \multicolumn{1}{c}{}          & \multicolumn{1}{c}{ADPLL Design 1}  & \multicolumn{1}{c}{ADPLL Design 2} & \multicolumn{1}{c}{ADPLL Design 3} \\ \cline{2-4}       
+        Capture Range                 & $1.08-8.61~\si{\mega\hertz}$       & $4.869-5.877~\si{\mega\hertz}$    & $4.377-5.363~\si{\mega\hertz}$       \T\\
+        Lock Range                    & $1.08-8.635~\si{\mega\hertz}$      & As above                           & As above                            \\
+        Jitter                        & $1.8385~\si{\nano\second}$         & $0.5623~\si{\nano\second}$         & $0.7410~\si{\nano\second}$          \\
+        Skew (Mean \acs{TIE} to Ref.) & $1.7439~\si{\nano\second}$         & $0.3741~\si{\nano\second}$         & $0.6381~\si{\nano\second}$          \\
+        Maximum \acs{TIE}             & $7.2840~\si{\nano\second}$         & $6.4770~\si{\nano\second}$         & $6.4578~\si{\nano\second}$          \\
+        $k_p$                         & $\frac{1}{32}$                     & $\frac{1}{16}$                     & $\frac{1}{16}$                      \\
+        $k_i$                         & $\frac{1}{256}$                    & $\frac{1}{64}$                     & $\frac{1}{64}$  
+        \end{tabular}
+    \end{center}
+    \caption[\ac{ADPLL} Design Comparison]{\ac{ADPLL} Design Comparison.}
+    \label{table:characterisation}
+\end{table}
+
+Looking then at the jitter and skew data, unsurprisingly the entirely \ac{FPGA} clocked \ac{ADPLL} with its larger period step is has the worst jitter, by a significant amount.
+The maximum \ac{TIE} is also correspondingly worse, which unless significant skew was present in the other \ac{ADPLL} designs is also to be expected.
+\ac{ADPLL} 2 \& 3 experience low jitter, at less than $1\%$ of the period.
+Somewhat surprisingly there is no apparent benefit in terms of jitter reduction due to the better delay resolution of the inverter based \ac{TDL}.
+This may be down to either the increase in jitter from a second source of variation, or a less favourable fractional-N synthesis due to the distribution of periods.
+
+Despite Vivado being forced to use the same layout for both \ac{ADPLL} 2 \& 3 through the use of fixed cells, the minimum and maximum frequencies are still significantly different, although the size of the range is almost identical.
+This highlights the difficulty of ensuring identical conditions between measurements which is the most significant downside of inverter based modules.
 
@@
 
-\section{Minor Variations}\label{section:minor_variations}
-%TODO osc size
+\section{\acs{ADPLL} Network Implementation}
+As the \ac{ADPLL}s were designed with network usage in mind, the \ac{HDL} aspect of their integration into a network is straightforward.
+For \ac{ADPLL} 1, the oscillators need only be wired together, but for Design 2 \& 3 which are reliant on inverters require some care to ensure their lock ranges overlap sufficiently.
+This is a time consuming task, as a number of iterations may be required, even when locking down the cells of lined up \ac{RO}s, as the nets forming the connections may change.
+It is recommended only to carry out alignment when all other aspects of the design are unlikely to change and to save the bitstream generated.
+The projected time requirement in aligning \ac{RO}s meant that the creation of a 4x4 network was abandoned.
+\ac{FPGA} clocked \ac{ADPLL}s are more suited to larger networks as their centre point can be determined with ease.
 
-\section{\acs{ADPLL} Network Implementation}
-\subsection{2x2 Network}
-\subsection{3x3 Network}
+The network operates using an external reference which is input over a \ac{PMOD} header that is shared with a number of output signals, however it could be configured to use an internal reference generated by either the \ac{FPGA} clock management utility or a \ac{DCO} of the designer's choice.
+
+In order to increase the flexibility of the network, the switches on the \acs*{Nexys} are used to change both the proportional and integral gains and the configuration of the network at runtime.
+The reconfigurability aspect of the error combiner allows for the network to operate in either uni- or bi-directional mode, or as a number of individual \ac{PLL}s all connected to the external reference.
+This is particularly advantageous for the inverter based designs, as multiple combinations can be measured without reimplementing, and thus changing the layout of either oscillator or \acl{PD}.
+
+Two different network sizes were implemented, a 2x2 and a 3x3 grid respectively.
+The 2x2 network was implemented first, and then this design extended to a 3x3 once measurements had been carried out on and any errors made had been corrected.
+In later sections matrix indexing will be used to refer to each oscillator, with \ac{ADPLL}$_{1,1}$ closest to the reference, and higher numbers in either index getting progressively further from the reference, as in Figure \ref{fig:network_mine}.
+A 3x3 network was the minimum goal, with the aforementioned but eventually not implemented 4x4 a stretch-goal if time permitted.
+A 3x3 network set as the primary goal, as it is the minimum size network in which there exists an \ac{ADPLL} with four neighbours.
+As will be explained in Chapter \ref{chap:5}, \ac{ADPLL} Design 1 was only implemented in a 2x2 configuration while both other \ac{ADPLL}s were tested in a 3x3 configuration.
+
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=0.5\textwidth]{../network_annotated}
+	\caption[ADPLL indexing explanation]{ADPLL indexing explanation.}
+	\label{fig:network_mine}
+\end{figure}
+
+%TODO why use each PLL

### 2019-04-20 thesis/TestingAnalysis.tex
79git statusgit statusgit statusgit statusgit statusgit statusgit statusgit statusgit statusgit statusgit statusgit status

@@
+\section{Chapter Overview}
+This chapter contains details of the measurements carried out on networks of various sizes and investigation into the impact of a number of decisions made in the design process.
+Explanations will be given or analysis made of the results, which will be given in the form of of either plots or tables, depending on which allows for more concise presentation of the data.
+Both 2x2 and 3x3 network performance analysis was performed with no input delay to the \ac{LF}, which Section \ref{subsection:lf_delay} will show to have a significant performance impact.
+All tests were carried out, unless otherwise stated, at a centre frequency of $5~\si{\mega\hertz}$.
+
 \section{2x2 Network Performance Comparison}
+In order to investigate the impact of the \ac{ADPLL} network on performance, and conversely the impact of each individual \ac{ADPLL} design on the network, switches were used to alter the configuration of the network without requiring re-implementation and a number of captures made at each setting.
+The three configurations were: each \ac{ADPLL} individually locked to the reference, the network in uni-directional mode and bi-directional mode.
+The centre frequency or gains were not changed between measurements in order to obtain a good comparison.
+For \ac{ADPLL} 1 the gains were: $k_p = \frac{1}{32}$ \& $k_i = \frac{1}{256}$.
+For \ac{ADPLL} 2 and 3 they were: $k_p = \frac{1}{16}$ \& $k_i = \frac{1}{64}$.
+The
+
+There are two main trends worth highlighting here, with the first of these being as the feedback network becomes more complicated a large degree of skew is introduced, with the magnitude becoming larger in the \ac{ADPLL}s further from the reference.
+This would appear to be attributable to a combination of the time taken for a change in the reference to propagate to the \ac{ADPLL} node in question, and the fineness of period and detection resolutions.
+This second observation follows from the significant reduction in the skew growth in Designs 2 \& 3 which have finer resolutions.
+
+The other trend, is the variation in \ac{C2C} jitter across the \ac{ADPLL} designs, with coarseness of the \ac{FPGA} clocked \ac{DCO} becoming somewhat obvious.
+As free \ac{PLL}s, the clocked \ac{DCO} has almost identical \ac{C2C} jitter and maximum \ac{TIE}, implying that only two period steps are used the majority of the time.
+As these steps are somewhat further apart than that of the \ac{RO}, at $3.875~\si{\nano\second}$ apart as opposed to an implementation dependant value in the region of $1.1765~\si{\nano\second}$, a much greater jitter is observed.
+This would confirm the belief that the \ac{FPGA} clocked designs are better suited to operational frequencies where the generated clock period to \ac{FPGA} clock period ratio is significantly lower than the $\frac{3.875~\si{\nano\second}}{200~\si{\nano\second}} = \frac{1}{51.6}$ in this design.
+Comparing the \ac{C2C} jitter of \ac{ADPLL} 1 \& 2, where the main difference is the choice of \ac{PFD}, the approximate $\frac{1.6~\si{\nano\second}}{0.50~\si{\nano\second}} \approx 3.2$ ratio of observed jitters is almost identical to the $\frac{3.875~\si{\nano\second}}{1.1765~\si{\nano\second}} \approx 3.29$ ratio of the period steps at $5~\si{\mega\hertz}$, implying the \ac{PFD} is at fault for this difference.
+
+The other notable difference in \ac{C2C} jitter, is that the use of an inverter based \ac{TDC}, and its accompanying variation due to implementation differences, has the expected result of additional \ac{C2C} jitter, however, this impact is signfiicantly less than the that of changes in the \ac{RO} period step, and the inverter based \ac{TDC} in combination with the \ac{FPGA} clocked \ac{DCO} would still suffer the same problems as it does in \ac{ADPLL} Design 1 with the clocked \ac{PFD}.
+
+%TODO additional comments with new data?
+
+\begin{table}[!ht]
+    \begin{center}
+        \begin{footnotesize}
+            \setlength{\tabcolsep}{.9\tabcolsep}
+            \begin{tabular}{ll|r|r|r|r|r|r|r|r|r|}           
+                \cline{3-11}
+                && \multicolumn{3}{c|}{Jitter Standard Deviation (ns)} & \multicolumn{3}{c|}{Max. Time Interval Error (ns)} & \multicolumn{3}{c|}{Skew (ns)} \T\\
+                \cline{3-11} 
+                &&PLL 11&PLL 12&PLL 22    &PLL 11&PLL 12&PLL 22    &PLL 11&PLL 12&PLL 22\T\\
+                \hline
+                \multicolumn{2}{|l|}{\ac{ADPLL} Design 1}&-&-&-&-&-&-&-&-&-\T\\
+                \multicolumn{2}{|r|}{Free PLLs} &1.6051  &1.6146  &1.6251     &1.6985 &1.5833 &1.2712    &0.2700 &0.5817 &0.2491 \T\\
+                \multicolumn{2}{|r|}{Uni-dir.}  &1.7693  &1.8463  &1.8862     &7.3060 &7.2565 &10.151    &2.2791 &1.2822 &1.6299 \T\\
+                \multicolumn{2}{|r|}{Bi-dir.}   &1.9964  &1.9146  &1.9572     &15.089 &18.283 &19.673    &9.2449 &11.612 &12.052 \T\\
+                \hline
+                \multicolumn{2}{|l|}{\ac{ADPLL} Design 2}&-&-&-&-&-&-&-&-&-\T\\
+                \multicolumn{2}{|r|}{Free PLLs} &0.56222 &0.49224 &0.64124    &5.3039 &7.8626 &5.4634    &&& \T\\
+                \multicolumn{2}{|r|}{Uni-dir.}  &0.55040 &0.51284 &0.62796    &5.5312 &7.5359 &7.9697    &&& \T\\
+                \multicolumn{2}{|r|}{Bi-dir.}   &0.57056 &0.50403 &0.61644    &11.515 &15.943 &17.428    &&& \T\\
+                \hline
+                \multicolumn{2}{|l|}{\ac{ADPLL} Design 3}&-&-&-&-&-&-&-&-&-\T\\
+                \multicolumn{2}{|r|}{Free PLLs} &0.66757 &0.84477 &0.71308    &5.9142 &6.1353 &5.3939    &-0.36997 &-0.20981 &-1.3344  \T\\
+                \multicolumn{2}{|r|}{Uni-dir.}  &0.66934 &0.83894 &0.71271    &5.8779 &7.1746 &6.8914    &-0.36645 &-0.05647 &-0.36178 \T\\
+                \multicolumn{2}{|r|}{Bi-dir.}   &0.68372 &0.83193 &0.68570    &7.5705 &9.6292 &9.7715    & 0.68901 & 1.58720 & 1.5553  \T\\
+                \hline
+                \B                
+            \end{tabular}
+        \end{footnotesize}
+        \caption{2x2 Network performance Comparison.}
+        \label{table:2x2perf}
+    \end{center}
+    \vspace{-0.5cm}
+\end{table}
+%\FloatBarrier
+
 \section{3x3 Network Performance Comparison}
+Subsequent to tests carried out using a 2x2 network, the corresponding tests were carried out using a 3x3 network in order to analyse the impact of a larger network.
+For these tests the gains were left unchanged but, as the 2x2 and 3x3 networks have different floorplans, the \ac{RO}s have changed somewhat.
+This has especially manifested itself in the exceptional individual \ac{PLL} performance of Design 3.
+
+From analysis of the 2x2 network, it was evident that the best performing design was going to be \ac{ADPLL} 3, as both \ac{ADPLL} 2 \& 3 contain the clocked \ac{PFD} which appears to cause significant skew as the \ac{ADPLL} node indices increase and move further from the external reference.
+The increased network size appears to have made the lag between reference edge and the rising edges of the individual generated signals worse, which would be expected as the number of clock cycles required to respond to a change in the phase relationship between reference and the furthest node has increased.
+\ac{ADPLL} 3 suffers least from the greater complication of the network, likely as a result of its period and detector resolution being the finest of the three designs considered.
+
+As \ac{ADPLL} Design 1 suffered significantly more with skew as the node indices increased, so it is no surprise that in the larger network the problem has been amplified, meaning this design is not well suited for use in larger networks at this frequency due to poor resolution.
+This does not, however, impact the main role of this design method, which is the emulation of designs intended for use in \ac{ASIC}s as the frequency of operation will be much lower than the $5~\si{\mega\hertz}$ used in these tests, which will bring with it a finer resolution.
+%TODO fill in this after measurements
+\begin{table}[!ht]
+    \begin{center}
+        \begin{footnotesize}
+            \setlength{\tabcolsep}{.9\tabcolsep}
+            \begin{tabular}{ll|r|r|r|r|r|r|r|r|r|}           
+                \cline{3-11}
+                && \multicolumn{3}{c|}{Jitter Standard Deviation (ns)} & \multicolumn{3}{c|}{Max. Time Interval Error (ns)} & \multicolumn{3}{c|}{Skew (ns)} \T\\
+                \cline{3-11} 
+                &&PLL 11&PLL 12&PLL 22    &PLL 11&PLL 12&PLL 22    &PLL 11&PLL 12&PLL 22\T\\
+                \hline
+                \multicolumn{2}{|l|}{\ac{ADPLL} Design 1}&-&-&-&-&-&-&-&-&-\T\\
+                \multicolumn{2}{|r|}{Free PLLs} &&& &&& &&& \T\\
+                \multicolumn{2}{|r|}{Uni-dir.}  &&& &&& &&& \T\\
+                \multicolumn{2}{|r|}{Bi-dir.}   &&& &&& &&& \T\\
+                \hline
+                \multicolumn{2}{|l|}{\ac{ADPLL} Design 2}&-&-&-&-&-&-&-&-&-\T\\
+                \multicolumn{2}{|r|}{Free PLLs} &0.47167 &0.55012 &0.36251    &5.9649 &6.6965 &5.5010    &3.4045&3.2369&2.3108 \T\\
+                \multicolumn{2}{|r|}{Uni-dir.}  &0.49292 &0.55586 &0.39054    &5.958  &8.9250 &11.622    &3.5052&6.9860&8.874 \T\\
+                \multicolumn{2}{|r|}{Bi-dir.}   &0.43263 &0.48053 &0.33778    &18.867 &35.973 &38.773    &15.193&28.965&30.841 \T\\
+                \hline
+                \multicolumn{2}{|l|}{\ac{ADPLL} Design 3}&-&-&-&-&-&-&-&-&-\T\\
+                \multicolumn{2}{|r|}{Free PLLs} &0.89042 &0.64889 &0.57405    &4.0634 &3.2467 &2.3141    &1.8373&1.1006&0.39334 \T\\
+                \multicolumn{2}{|r|}{Uni-dir.}  &0.93580 &0.93353 &0.71361    &4.2675 &5.8193 &6.4139    &1.8605&2.3172&3.1463  \T\\
+                \multicolumn{2}{|r|}{Bi-dir.}   &0.97717 &0.85131 &0.63863    &5.6439 &8.4856 &9.4467    &2.7129&4.1074&5.1717  \T\\
+                \hline
+                \B                
+            \end{tabular}
+        \end{footnotesize}
+        \caption{3x3 Network performance Comparison.}
+        \label{table:3x3perf}
+    \end{center}
+    \vspace{-0.5cm}
+\end{table}
+%\FloatBarrier
+
+\section{Minor Variations}\label{section:minor_variations}
+A number of minor design decisions will be tested in this section against their alternatives, as will the impact of varying the \ac{LF} gains.
+
+\subsection{Impact of Gain Variation}
+The gains chosen for use in the \acl{LF} have significant impact on the performance of the system, as the \ac{LF} governs the response of the \ac{ADPLL} to a given phase error.
+To large a proportional gain will result in overly aggressive overcompensation, thus negatively affecting the jitter.
+Eventually the violence of the jitter will result in a loss of lock.
+However, if the gain is too small, the frequency or period adjustment will not be sufficiently large to cause the \ac{ADPLL} to lock.
+Similarly, appropriate values of $k_i$ are required, and the relationship between proportional and integral gains must satisfy the conditions set out by Koskin \cite{koskin2018generation}.
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=0.6\textwidth]{../fixed_kp.png}
+    \caption[Fixed $k_p$ integral gain sweeps]{Fixed $k_p$ integral gain sweeps.}
+    \label{fig:gain_sweep}
+\end{figure}
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=0.6\textwidth]{../fixed_kp_zoom.png} 
+    \caption[Fixed $k_p$ integral gain sweeps]{Fixed $k_p$ integral gain sweeps.}
+    \label{fig:gain_sweep_zoom}
+\end{figure}
+
+Figure \ref{fig:gain_sweep} contains a sweep across the possible range of integral gains for four different proportional gains.
+This test was carried out with a 3x3 network of \ac{ADPLL} Design 3, each locked to the external reference at $5~\si{\mega\hertz}$, and the generated waveform from the \ac{ADPLL}s at locations ``11'', ``22'' and  ``33'' were captured.
+It can clearly be observed that there are two distinct regions in the graph, with jitter values of less than $1~\si{\nano\second}$ contrasting sharply with the points with jitter in the region of $10~\si{\nano\second}$.
+These distinct sections of the plot represent locked and unlocked operation respectively.
+Zooming in on the low jitter region in Figure \ref{fig:gain_sweep_zoom}, the loss of lock thresholds can be compared with the theoretical relationship of $k_i = 0.5\times k_p$.
+\begin{table}[!h]
+    \begin{center}
+        \begin{tabular}{lrr}
+            \multicolumn{1}{c}{$k_p$} & \multicolumn{1}{c}{$k_i$} & \multicolumn{1}{c}{Ratio} \T\B\\
+            \hline
+            $\frac{1}{16}$            & $\frac{5}{64}$            & $0.8$                     \T\B\\
+            $\frac{2}{16}$            & $\frac{10}{64}$           & $0.8$                     \T\B\\
+            $\frac{3}{16}$            & $\frac{12}{64}$           & $1$                       \T\B\\
+            $\frac{4}{16}$            & $\frac{14}{64}$           & $1.1428$                  \T\B
+        \end{tabular}
+    \end{center}
+    \vspace{-0.5cm}
+    \caption[Loss of Lock Gains]{Loss of Lock Gains.}
+    \label{table:lolgains}
+\end{table}
+
+Despite the relatively small number of proportional gains, it is clear that the relationship set out by Koskin does not hold in this system, but rather a scaled version of it.
+This difference is attributable to a difference in \ac{LF} structure.
+In his simulations the \ac{LF} design contains a register forming an input buffer, absent in the \ac{LF} used for these tests.
+The presence of this register, clocked on the generated signal, causes the \ac{ADPLL} to react more slowly to changes in the phase difference, thus increasing the jitter.
+The removal of this register causes a change in the relationship between proportional and integral gains, a change that will be investigated further in a later section.
+
+\subsection{Distribution of Period Steps}
+To add context to the jitter numbers produced for each oscillator previously, the variation of period steps in the \ac{RO}, and delays in the \ac{TDL} due to implementation should be examined.
+The best way to visualise this distribution is through a histogram, placing periods of in a certain range into the same bin.
+This can be visualised effectively using Matlab's \texttt{histplot}.
+In order to obtain data, three \ac{ADPLL}s implementing both the \ac{FPGA} clocked and inverter based \ac{DCO}s were locked to an external reference signal at $5~\si{\mega\hertz}$, and the period of each clock cycle measured.
+\begin{figure}[h]%
+    \centering
+    \subfloat[Inverter propagation delay based \ac{RO}.]{{\includegraphics[width=0.5\textwidth]{../conf_paper/distrib_ring} }}%
+    \subfloat[\ac{FPGA} clocked \ac{DCO}.]{{\includegraphics[width=0.5\textwidth]{../conf_paper/distrib_pa} }}%
+    \caption[Example distribution of periods]{Example distribution of periods.}    
+    \label{fig:dists}
+\end{figure}
+Figure \ref{fig:dists} contains the output of \texttt{histfit} for each of the three oscillators superimposed in one plot for each \ac{DCO} type.
+As the \ac{RO} has significantly finer period resolution, the period steps would be overlaid on each-other thereby making the plot difficult to read.
+In order to avoid this behaviour, the \ac{RO}'s mapping of control code to inverters to remove was altered from twice the control code to 16 times the control code.
+This results in a more pronounced separation of period steps while being indicative of the variation due to implementation.
+The plots paint a a picture of the unrealistic nature of the \ac{FPGA} clocked design in comparison to its inverter based counterpart, as each oscillator will have identical intrinsic period steps thus eliminating any jitter that maybe be seen in the system due to misalignment.
+This effects the suitability of the FPGA clocked \ac{DCO} for the improvement of simulation models, or the examination of how a novel block may impact the behaviour of the network.
+If this analysis was carried out for the period steps, similar behaviour would be observed.
+
+\subsection{\ac{FPGA} Clocked \ac{DCO} Width Variation}
+When introducing the design of the \ac{DCO} for \ac{ADPLL} 1, the minimum possible accumulator width was chosen.
+From the equation used to calculate the frequency step size, it may have seemed that by doubling the bias point and increasing the width of the counter to 13 bits a finer resolution could be achieved.
+\begin{align}
+f_{step} &= \frac{f_{FPGA}}{2^{12}} = 62.988~\si{\kilo\hertz} \\
+f_{step} &= \frac{f_{FPGA}}{2^{13}} = 31.494~\si{\kilo\hertz}
+\end{align}
+As Table \ref{table:accum_width} shows, this is not in fact the case.
+Both measurements were carried out using identical loop filter gains, network configurations and reference signal.
+While in uni-directional mode some difference is observed in the maximum \ac{TIE}, but in all other aspects no statistically relevant changes can be seen.
+%TODO explanation for this?
+\begin{table}[!ht]
+    \begin{center}
+        \begin{footnotesize}
+            \setlength{\tabcolsep}{.9\tabcolsep}
+            \begin{tabular}{ll|r|r|r|r|r|r|r|r|r|}           
+                \cline{3-11}
+                && \multicolumn{3}{c|}{Jitter Standard Deviation (ns)} & \multicolumn{3}{c|}{Max. Time Interval Error (ns)} & \multicolumn{3}{c|}{Skew (ns)} \T\\
+                \cline{3-11} 
+                &&PLL 11&PLL 12&PLL 22    &PLL 11&PLL 12&PLL 22    &PLL 11&PLL 12&PLL 22\T\\
+                \hline
+                \multicolumn{2}{|l|}{Accum. Width 12}&-&-&-&-&-&-&-&-&-\T\\
+                \multicolumn{2}{|r|}{Uni-dir.}  &1.7693 &1.8463 &1.8862    &7.3060  &7.2565  &10.151    &2.2791 &1.2822  &1.6299  \T\\
+                \multicolumn{2}{|r|}{Bi-dir.}   &1.9964 &1.9146 &1.9572    &15.0890 &18.2830 &19.673    &9.2449 &11.6120 &12.0520 \T\\
+                \hline
+                \multicolumn{2}{|l|}{Accum. Width 13}&-&-&-&-&-&-&-&-&-\T\\
+                \multicolumn{2}{|r|}{Uni-dir.}  &1.7946 &1.8161 &1.9048    &6.8614  &7.1495 &9.371      &2.2257 &1.6328  &1.3778  \T\\
+                \multicolumn{2}{|r|}{Bi-dir.}   &2.0016 &1.9217 &1.9605    &14.9580 &18.938 &19.510     &9.1678 &12.0000 &11.7690 \T\\
+                \hline
+                \B
+            \end{tabular}
+        \end{footnotesize}
+        \caption{\ac{DCO} Accumulator Width Comparison.}
+        \label{table:accum_width}
+    \end{center}
+    \vspace{-0.5cm}
+\end{table}
+%\FloatBarrier
+
+\subsection{\ac{LF} Input Delay Register}\label{subsection:lf_delay}
+Previously, when comparing the loss of lock gain relationship to that proposed by Koskin \textit{et al}, it was mentioned that the \ac{LF} used in this project does not have an input delay register.
+The lack of agreement between the theory and experimental results was attributed to the input delay register, or lack thereof.
+In order to provide a basis for this claim, the performance of a network in 2x2 configuration, was analysed and the results of this analysis are presented in Table \ref{table:lf_delay}.
+While the \ac{C2C} jitter is statistically similar, there is significant different to be observed in the \ac{TIE}, where an approximately $50\%$ increase can be observed, measurement for measurement.
+As the maximum \ac{TIE} is significantly impacted by the presence of any skew, as skew is the average value of this difference, it is interesting to note that the relationship is in fact the opposite.
+It is in precisely this situation that \ac{C2C} jitter as a measurement is non-ideal.
+\ac{C2C} jitter does not differentiate between the situations in which while the period variation might be the same for multiple clock cycles in the same or opposite directions.
+As the extra cycle of delay due to the input register increases the reaction time of the \ac{ADPLL} to any changes, the delay between clock edges increases.
+\ac{TIE} is, however, affected by this difference which can be clearly seen in the table.
+
+\begin{table}[!ht]
+    \begin{center}
+        \begin{footnotesize}
+            \setlength{\tabcolsep}{.9\tabcolsep}
+            \begin{tabular}{ll|r|r|r|r|r|r|r|r|r|}           
+                \cline{3-11}
+                && \multicolumn{3}{c|}{Jitter Standard Deviation (ns)} & \multicolumn{3}{c|}{Max. Time Interval Error (ns)} & \multicolumn{3}{c|}{Skew (ns)} \T\\
+                \cline{3-11} 
+                &&PLL 11&PLL 12&PLL 22    &PLL 11&PLL 12&PLL 22    &PLL 11&PLL 12&PLL 22\T\\
+                \hline
+                \multicolumn{2}{|l|}{W/ \ac{LF} input reg}&-&-&-&-&-&-&-&-&-\T\\
+                \multicolumn{2}{|r|}{Free PLLs} &0.66757 &0.84477 &0.71308    &5.9142 &6.1353 &5.3939   &-0.36997 &-0.20981 &-1.3344  \T\\
+                \multicolumn{2}{|r|}{Uni-dir.}  &0.66934 &0.83894 &0.71271    &5.8779 &7.1746 &6.8914   &-0.36645 &-0.05647 &-0.36178 \T\\
+                \multicolumn{2}{|r|}{Bi-dir.}   &0.68372 &0.83193 &0.68570    &7.5705 &9.6292 &9.7715   & 0.68901 & 1.58720 & 1.5553  \T\\
+                \hline
+                \multicolumn{2}{|l|}{W/o \ac{LF} input reg}&-&-&-&-&-&-&-&-&-\T\\
+                \multicolumn{2}{|r|}{Free PLLs} &0.67899 &0.65618 &0.61415    &4.2527 &3.6478 &3.8483   &-0.31418 &0.11897 &-0.57619 \T\\
+                \multicolumn{2}{|r|}{Uni-dir.}  &0.68534 &0.69405 &0.63666    &4.3257 &4.2901 &4.6185   &-0.38849 &0.23363 & 0.21156 \T\\
+                \multicolumn{2}{|r|}{Bi-dir.}   &0.68248 &0.70175 &0.61356    &5.3497 &6.2640 &6.8079   & 0.52553 &1.7025  & 1.9499  \T\\
+                \hline
+                \B
+            \end{tabular}
+        \end{footnotesize}
+        \caption{Impact of \ac{LF} Input Delay Register.}
+        \label{table:lf_delay}
+    \end{center}
+    \vspace{-0.5cm}
+\end{table}
+%\FloatBarrier
+
+\subsection{Impact of Frequency Divider}
+The major benefit of an \ac{PLL} network over other forms of coupled oscillator based clock distribution systems, is that the coupling can be done using a signal significantly reduced in frequency, thereby reducing the power consumption of the related hardware.
+It is therefore important to ensure that the insertion of a feedback divider does not introduce any further jitter to the system.
+Table \ref{table:div} presents the results of measurements at a number of divisions levels.
+As re-implemenation was performed between each measurement, there is some variation present.
+However, it is notable that for each \ac{ADPLL} that saw a \ac{C2C} jitter increase, there was another that performed better than at a lower level of division.
+Accordingly, it can be claimed that the insertion of a feedback divider does not impact the \ac{C2C} jitter present.
+Intuitively this makes sense as the control code will change more violently, but proportionally less frequently.
+\begin{table}[!ht]
+    \begin{center}
+        \begin{footnotesize}
+            \setlength{\tabcolsep}{.9\tabcolsep}
+            \begin{tabular}{ll|r|r|r|r|r|r|r|r|r|}           
+                \cline{3-11}
+                && \multicolumn{3}{c|}{Jitter Standard Deviation (ns)} & \multicolumn{3}{c|}{Max. Time Interval Error (ns)} & \multicolumn{3}{c|}{Skew (ns)} \T\\
+                \cline{3-11} 
+                &&PLL 11&PLL 12&PLL 22    &PLL 11&PLL 12&PLL 22    &PLL 11&PLL 12&PLL 22\T\\
+                \hline
+                \multicolumn{2}{|l|}{No Divider}&-&-&-&-&-&-&-&-&-\T\\
+                \multicolumn{2}{|r|}{Free PLLs} &0.66757 &0.84477 &0.71308    &5.9142 &6.1353 &5.3939    &-0.36997 &-0.20981 &-1.3344  \T\\
+                \multicolumn{2}{|r|}{Uni-dir.}  &0.66934 &0.83894 &0.71271    &5.8779 &7.1746 &6.8914    &-0.36645 &-0.05647 &-0.36178 \T\\
+                \multicolumn{2}{|r|}{Bi-dir.}   &0.68372 &0.83193 &0.68570    &7.5705 &9.6292 &9.7715    & 0.68901 & 1.58720 & 1.5553  \T\\
+                \hline
+                \multicolumn{2}{|l|}{Divide by 2}&-&-&-&-&-&-&-&-&-\T\\
+                \multicolumn{2}{|r|}{Free PLLs} &0.68252 &0.51750 &1.1074     &4.9464 &4.0728 &6.7426    &-0.91985 &-1.7024  &-1.84279 \T\\
+                \multicolumn{2}{|r|}{Uni-dir.}  &0.68441 &0.53203 &1.1081     &4.8707 &4.9684 &8.9973    &-0.99034 &-1.0240  &-0.63989 \T\\
+                \multicolumn{2}{|r|}{Bi-dir.}   &0.69524 &0.54953 &1.1033     &6.8180 &7.8760 &12.288    & 0.00337 & 0.69148 & 1.3085  \T\\
+                \hline
+                \multicolumn{2}{|l|}{Divide by 4}&-&-&-&-&-&-&-&-&-\T\\
+                \multicolumn{2}{|r|}{Free PLLs} &0.77564 &0.55934 &0.65770    &10.576 &6.7330 &12.647    &-1.1507  &-1.8670   &-2.3381 \T\\
+                \multicolumn{2}{|r|}{Uni-dir.}  &0.77443 &0.56544 &0.66019    &10.210 &7.4612 &15.166    &-1.2283  &-1.2691   &-1.0948 \T\\
+                \multicolumn{2}{|r|}{Bi-dir.}   &0.77657 &0.57043 &0.66809    &11.408 &10.859 &17.242    &-0.28439 & 0.40697  & 0.8351 \T\\
+                \hline
+                \B
+            \end{tabular}
+        \end{footnotesize}
+        \caption{Network Performance at Different Divider Levels.}
+        \label{table:div}
+    \end{center}
+    \vspace{-0.5cm}
+\end{table}
+%\FloatBarrier
+
+It, however, does have a significant impact on the maximum \ac{TIE}, and notably this occurs without any meaningful change in skew.
+This confirms an increase in jitter due to the greater division levels as the source of the performance degradation.
+By examining how the dividers behave, this can be explained.
+When a divider is inserted into the network, the phase error is computed at a greatly reduced rate.
+This means the rate at which the \ac{LF} is updated also drops, and therefore the control code is changes $n$ times less frequently, where $n$ is the division level.
+The dividers reduce the speed at which the individual \ac{ADPLL}s can react to a phase difference, while locking the resulting control code in place for a greater number of clock cycles as the division level increases.
+This greater duration may cause a control code to be applied for an excessively long number of cycles, resulting in a drift of the rising edge of the divided/generated signal away from the reference, which will manifest as \ac{TIE}.
+These tests were carried out with identical gains for the proportional and integral paths in order to demonstrate the effect, however, an appropriate reduction in these gains can correct this problem by reducing the magnitude of the control code alteration.
+%TODO ask mulkeen
+Figure \ref{fig:todo_div} can confirm this behaviour.
+In Plot (a), as the division level increases, the number of cycles spent at a $200~\si{\nano\second}$ period decreases, which would line up with the overcorrection, thus resulting in the increased \ac{TIE} seen in Plot (b).
+\begin{figure}[h]%
+    \centering
+    \includegraphics[width=0.8\textwidth]{../bad_waveform}
+    \caption[Period distribution at differen divider levels]{Period distribution at differen divider levels.}    
+    \label{fig:todo_div}
+\end{figure}
+
+\section*{\acs{ADPLL} Use Cases}
+As each design was introduced, the most appropriate use cases for that particular \ac{ADPLL} were suggested.
+\ac{ADPLL} networks can be used in a number of roles, from modelling system dynamics or the refinement of behavioural simulation models to the validation, verification and testing of potential \ac{ASIC} based networks (Zianbetov \& Shan) to the design of other hardware for the purposes of network performance improvement or control, however, some architectures of \ac{ADPLL} are more suited to particular roles than others.
+
+As mentioned when \ac{FPGA} clocked designs were introduced, their configurability makes them particularly suited to the validation and verification of designs destined for use in \ac{ASIC}s, as was the case with Zianbetov and Shan \cite{zianbetov2013phd,shan2014phd}.
+In contrast to the \ac{RO} based designs, in which configurability is more more restricted, it is possible to replicate the resolution, centre frequencies and other \ac{ADPLL} characteristics of the design requiring verification at a scaled down frequency.
+This can be carried out at large division ratios, where the \ac{FPGA} clock period is not a restricting factor, such as the $50~\si{kilo\hertz}$ centre frequency used by Zianbetov and Shan.
+In order to reduce the frequency to this extent with inverters over $30$ thousand would be required, introducing the potential for even greater variation in centre frequencies due to implementation.
+The other issue arising out of this number of inverters is the floorspace of the \ac{FPGA} that would be consumed implementing just one \ac{RO}, in the case of the Xilinx \acl{Nexys} and extrapolating based on the resource usage of the \ac{RO}s implemented, a single oscillator would consume half the look-up tables available, eliminating the possibility of a network.
+
+While the \ac{FPGA} based designs feature characteristics making them particularly suitable for use as a validation tool, their performance in a network reduces their suitability for the refinement of software based simulations, or in the analysis of new control or performance enhancement tools.
+Referring back to Table \ref{table:2x2perf}, \ac{C2C} jitter is significantly greater than the inverter based designs, and a large degree of skew is present making this design non-ideal for these purposes.
+In particular, as seen with the increase to a 3x3 network, this skew becomes an issue that limits the network size.
+
+Figure \ref{fig:dists}, which demonstrated the impact of period distribution, highlights a second major unrealistic behaviour present in \ac{FPGA} clocked \ac{ADPLL} designs, as an \ac{ASIC} destined design will not exhibit this identical linear period distribution.
+While \ac{ADPLL} 2 will feature a large degree of variation due to implementation differences, these differences are only present in the \ac{RO} not the \ac{PFD}.
+As testing Designs 2 \& 3 has shown, in particular the results displayed in Table \ref{table:2x2perf}, apart from a larger skew, the jitter performance is quite similar, meaning that the extra complexity of inverter based \ac{PFD} is not required in order to achieve as accurate behaviour as can be achieved in this environment.
+The implementation based variation of the detector step manifests itself as an increase in the jitter across all network modes, as would be expected, compared to the \ac{FPGA} based design.
+
+Accordingly, if the aim of this platform is the enhancement of a model, the validation of theoretical results, or analysing the performance of new control or performance enhancement hardware the \ac{RO} based designs are idea for this purpose.
+Two members of the research team here in \acs{UCD}, Eugene Koskin and Pierre Bisiaux, have been using an \ac{ADPLL} with the same architecture as Design 3 for these purposes.

### 2019-04-21 matlab/a1_3.m
updated images

@@ -1,37 +1,34 @@
-for inc = 1:5
+num_files = 15;
+std_period = zeros(4,num_files);
+TIE_mean = zeros(4,num_files);
+TIE_max = zeros(4,num_files);
+N           = 4; 
+step        = 2e6;
+width_vect  = 4e6;
+div_ratio = 4;
+config = "div" + num2str(div_ratio);
+config = "nodiv";
+
+for inc = 1:num_files
     %filename = char("conor_test/settings1_" + num2str(inc) + ".bin");
     %filename = char("scope_41.bin");
-    
-    %path3       = "~/gdrive/measurements/monday2/adpll_1b" + num2str(inc) + ".bin";
-    path3       = "~/gdrive/measurements/monday/design2b_" + num2str(inc) + ".bin";
+        
+    path3       = "~/gdrive/measurements/monday2/adpll_1b" + num2str(inc) + ".bin";
+%     path3       = "~/gdrive/measurements/tuesday/net3x3_clocked_pd_5m18_" + num2str(inc) + ".bin";
+%     path3       = "~/gdrive/measurements/wednesday/net3x3_5m00_kp_100_ki_1_" + num2str(inc) + ".bin";
+%     path3       = "~/gdrive/measurements/monday/design2b_div" + num2str(div_ratio) + "_" + num2str(inc) + ".bin";    
+%     path3       = "~/gdrive/measurements/monday/design2b_" + num2str(inc) + ".bin";
+
+    disp(path3)
     fileID      = fopen(path3);
     A           = fread(fileID, 'double');
     fclose(fileID);
-    N           = 4; 
-    step        = 2e6;
-    width_vect  = 4e6;
-    ss          = [];
+    ss = [];
     for chan=0:N-1 %start at 0 for chan1
         deb = chan*width_vect+1;
         ss = [ss A(deb:deb+step-1, 1) A(deb+step:deb+2*step-1, 1)];
     end
 
-%     Ni = N-1;
-%     for jkl = 0:Ni
-%          [w,init_cross,final_cross,mid_level] = pulsewidth(ss(:,2+2*jkl),ss(:,1+2*jkl), 'Polarity', 'Positive');
-%          switch jkl
-%              case 0
-%                 PP = init_cross;
-%              otherwise
-%                  if length(PP)>length(init_cross)
-%                     PP = [PP(1:length(init_cross),:) init_cross];
-%                  elseif length(PP)<length(init_cross)
[…]

### 2019-04-21 matlab/div_dist.m

@@ -1,0 +1,39 @@
+periods_div1_st = load('div1_periods.mat');
+periods_div2_st = load('div2_periods.mat');
+periods_div4_st = load('div4_periods.mat');
+
+periods_div1 = periods_div1_st.periods;
+periods_div2 = periods_div2_st.periods;
+periods_div4 = periods_div4_st.periods;
+
+tie_div1_st = load('div1_tie.mat');
+tie_div2_st = load('div2_tie.mat');
+tie_div4_st = load('div4_tie.mat');
+
+tie_div1 = tie_div1_st.tie_save;
+tie_div2 = tie_div2_st.tie_save;
+tie_div4 = tie_div4_st.tie_save;
+
+figure
+hold on;
+h = histfit(periods_div1(2,:)*1e9,67);
+h(1).FaceColor = [1 0 0];
+h = histfit(periods_div2(2,:)*1e9,67);
+h(1).FaceColor = [0 1 0];
+h = histfit(periods_div4(2,:)*1e9,67);
+h(1).FaceColor = [0 0 1];
+title('')
+xlabel('Period (ns)','fontsize',16);
+ylabel('Occurance Rate','fontsize',16);
+yticklabels([])
+ax = gca;
+ax.FontSize = 16;
+alpha(.33)
+legend("div1","div2","div4")
+
+figure
+hold on;
+plot(tie_div1(2,:))
+x = 1:2:(10000-2);
+plot(x,tie_div2(2,:)-200E-9)
+plot(1:4:10000,tie_div4(2,:))

### 2019-04-21 matlab/gain_grid.m

@@ -1,6 +1,6 @@
 for inc = 1:4 %kp loop
     for inner_inc = 1:15 %ki_loop
-
+        tic
         kp = inc;
         ki = inner_inc;
 
@@ -52,9 +52,8 @@ for inc = 1:4 %kp loop
         std_period(2,inc,inner_inc) = std(periods2);
         std_period(3,inc,inner_inc) = std(periods3);
         std_period(4,inc,inner_inc) = std(periods4);
-        
-        
-        mean_jitter = squeeze(mean(std_period,1));
+               
+        toc
         
 %         if size(init_cross(2,:),2) == size(init_cross(1,:),2)
 %             TIE(2,inc,inner_inc) = max(abs(init_cross(2,1:size(init_cross(1,:),2))-init_cross(1,1:size(init_cross(1,:),2))));
@@ -75,15 +74,15 @@ for inc = 1:4 %kp loop
     end
 end
 
-mean_jitter = squeeze(mean(std_period,1));
+std_period_to_disp = std_period(2:4,1:4,1:end);
 
 fig_handle = figure(1);
 axes_handle = subplot(1,1,1);
-scatter([1:15],mean_jitter(1,:)*1E9,'b');
+scatter(axes_handle,reshape(repmat([1:15],[3 1]),1,[]),reshape(squeeze(std_period_to_disp(:,1,:))*1E9,[1 3*15]),'b');
 hold(axes_handle,'on')
-scatter([1:15],mean_jitter(2,:)*1E9,'r');
-scatter([1:15],mean_jitter(3,:)*1E9,'g');
-scatter([1:15],mean_jitter(4,:)*1E9,'k');
+scatter(axes_handle,reshape(repmat([1:15],[3 1]),1,[]),reshape(squeeze(std_period_to_disp(:,2,:))*1E9,[1 3*15]),'r');
+scatter(axes_handle,reshape(repmat([1:15],[3 1]),1,[]),reshape(squeeze(std_period_to_disp(:,3,:))*1E9,[1 3*15]),'g');
+scatter(axes_handle,reshape(repmat([1:15],[3 1]),1,[]),reshape(squeeze(std_period_to_disp(:,4,:))*1E9,[1 3*15]),'k');
[…]

### 2019-04-21 thesis/Background.tex
start of conc

@@
 The main drivers of performance in \acp{SOC} are the number of transistors on a chip, which is correlated with the number of calculations that can be carried out simultaneously, and the frequency at which the device operates, which determines the number of calculations performed per second.
-Moore's Law, based on the famous observation by Gordon Moore in 1965\cite{moore1965cramming}, predicted a doubling in the transistor count of \ac{IC}s per year for the forthcoming decade.
+Moore's Law, based on the famous observation by Gordon Moore in 1965\cite{moore1965cramming}, predicted a doubling in the transistor count of \acl{IC} (\acs{IC}s) per year for the forthcoming decade.
 This behaviour has carried on to this day as a result of the ever decreasing size of transistors, and has only begun to slow down in recent years.
-However, as the number of transistors on a chip has increased roughly following Moore's Law, the increase in clock frequency has not been able to follow a similar linear trajectory, having remained roughly equivalent for the last number of years \cite{ross2008cpu}, indeed the clock speed of the Intel Core family of \ac{CPU}s has not changed since their introduction in 2009 \cite{intelark}.
+However, as the number of transistors on a chip has increased roughly following Moore's Law, the increase in clock frequency has not been able to follow a similar linear trajectory, having remained roughly equivalent for the last number of years \cite{ross2008cpu}, indeed the clock speed of the Intel Core family of \acl{CPU} (\acs{CPU}s) has not changed since their introduction in 2009 \cite{intelark}.
 \begin{figure}[h]
@@
 This plateauing of clock frequency has been caused by high power consumption due to the demands placed by the global distribution of a high frequency clock, often the single biggest consumer of power on the chip \cite{tiwari1998reducing}.
-With the growth of the \ac{IOT} market where low power devices are desirable, with many of the emerging uses of \ac{SOC}s being portable and thus without a permanent power source, high power consumption goes directly against one of the key pillars of the technology.
+With the growth of the \ac{IOT} market where low power devices are desirable, with many of the emerging uses of \acp{SOC} being portable and thus without a permanent power source, high power consumption goes directly against one of the key pillars of the technology.
 This forces many of these devices to use lower performance hardware in order to reduce the power consumption, and increase the battery life, of their devices.
@@
 
-In contrast in a \ac{GALS} clock delivery system the ``local'' areas are not synchronised with other.
+In contrast in \iac{GALS} clock delivery system the ``local'' areas are not synchronised with other.
 This reduces the clocking system's complexity and thus the power consumption and chip area used, at the expense of communication speed between blocks.
 This disadvantage comes from the need to then somehow synchronise the messages being sent from one area to another to avoid the corruption of any messages.
-A \ac{GSLS}, system, however has the advantages of deterministic behaviour and greater rates of communication between clocking areas and, as such, remains a desirable system design.
+\Iac{GSLS}, system, however has the advantages of deterministic behaviour and greater rates of communication between clocking areas and, as such, remains a desirable system design.
 A number of methods which deliver \ac{GSLS} clocking exist at present such as clock trees as well as emerging technologies such as ADPLL networks.
@@
 Correspondingly a positive $\Delta T$ for depicted registers implies a negative $\Delta T$ for $R_f$ and the subsequent register.
-The most common sources of clock error are caused by mismatches which usually stem from production, such as differences in the length of clocking paths, buffer delays or in the parameters of either active or passive components in the clock distribution network, which as the size of components on an \ac{IC} reduces becomes more difficult to avoid.
+The most common sources of clock error are caused by mismatches which usually stem from production, such as differences in the length of clocking paths, buffer delays or in the parameters of either active or passive components in the clock distribution network, which as the size of components on \iac{IC} reduces becomes more difficult to avoid.
 All sources of mismatch will manifest themselves in the clock distribution system as skew between transistors, while the noise in active components or the power supply system will appear as jitter in the clock signal.
@@
 The advantage of this method is the simplicity of the feedback network, requiring just the divided clock output from a single leaf node.
-The \ac{VCO}s then adjusted by the control voltage, $v_c$, which needs delivery to all areas of the chip.
+The \acp{VCO} then adjusted by the control voltage, $v_c$, which needs delivery to all areas of the chip.
 However, this is a regular signal and as such does not suffer from skew or jitter.
 This alleviates the need for a power hungry distribution circuit, while also being more noise-immune than the transmission of a high frequency clock.
-However this design still suffers from clock variation as all \ac{VCO}s are fed the same control voltage, and thus the manufacturing tolerance issues present in conventional designs persists here also.
+However this design still suffers from clock variation as all \acp{VCO} are fed the same control voltage, and thus the manufacturing tolerance issues present in conventional designs persists here also.
 This is acknowledged by the authors:
@@
 	\textit{Unfortunately, as with the conventional ...
-method, distributing the \ac{VCO}s over the entire chip causes the problem that jitter and skew are increased by variations in the fabrication process (static), temperature, and power supply (dynamic)} \cite{mizuno1998noise}.
+method, distributing the \acp{VCO} over the entire chip causes the problem that jitter and skew are increased by variations in the fabrication process (static), temperature, and power supply (dynamic)} \cite{mizuno1998noise}.
 \end{quotation}
@@
 Once again, this negates the requirement for a global distribution structure and the signals used for comparisons need only be sent between neighbouring clocking areas.
-As a \ac{PLL} is being used it is again possible to perform the phase comparisons using a divided version of the generated clock.
+As \iac{PLL} is being used it is again possible to perform the phase comparisons using a divided version of the generated clock.
 This in turn means the hardware transporting the divided clock signal to the phase comparator, has significantly lower requirements placed on it, thus lowering the power consumption due to electrical losses.
-Pratt and Nguyen initially proposed method of clock distribution in their 1995 paper entitled ``\textit{Distributed Synchronous Clocking}'' in which they propose a Cartesian grid of clocking areas, each with their own \ac{PLL}, which has become known as a \ac{PLL} Network \cite{pratt1995distributed}.
+Pratt and Nguyen initially proposed method of clock distribution in their 1995 paper entitled ``\textit{Distributed Synchronous Clocking}'' in which they propose a Cartesian grid of clocking areas, each with their own \ac{PLL}, which has become known as \iac{PLL} Network \cite{pratt1995distributed}.
 In this design any given node is synchronised with its neighbours and one of the corner nodes is additionally synchronised with the reference.
@@
 	\textit{Design and measurements on this chip confirm that generating and synchronizing multiple clocks on chip is feasible.
-Neither the power nor the area overhead of multiple \acs{PLL}s is substantial compared to the cost of distributing the clock by conventional means} \cite{gutnik2000active}.
+Neither the power nor the area overhead of multiple \acsp{PLL} is substantial compared to the cost of distributing the clock by conventional means} \cite{gutnik2000active}.
 \end{quote}
@@
 \section{ADPLL Networks}
-As a \ac{PLL} network is an analog circuit, its integration in a modern \ac{IC} is a barrier to usage, and as such it has not been used in any commercial designs \cite{zianbetov2013distributed}.
+As \iac{PLL} network is an analog circuit, its integration in a modern \ac{IC} is a barrier to usage, and as such it has not been used in any commercial designs \cite{zianbetov2013distributed}.
 An alternative design that is more suitable for current fabrication techniques eschews from using analogue components and instead implements the network of controlled oscillators using only digital circuitry, hence the name All-Digital \ac{PLL}.
@@
 As indicated in Figure \ref{fig:mulkeen_pll} the three main building blocks of a conventional \ac{PLL} are the \ac{PD}, \ac{LF} and \ac{VCO}.
-In an \ac{ADPLL} these blocks are then replaced by their digital counterparts, necessitating quantisation in order to remain physically realisable.
+In \iac{ADPLL} these blocks are then replaced by their digital counterparts, necessitating quantisation in order to remain physically realisable.
 The ``All-Digital'' moniker is a misnomer as the oscillator and \acl{PD} are usually both implemented by mixed signal circuits.
@@
 	\includegraphics[scale=0.5]{../tex_files/mulkeen_pll}
-	\caption[Block fiagram of a \acl{PLL}]{Block diagram of a \acl{PLL}, \textit{Wireless Systems Notes}, B. Mulkeen (2017).}
+	\caption[Block fiagram of \iacl{PLL}]{Block diagram of \iacl{PLL}, \textit{Wireless Systems Notes}, B. Mulkeen (2017).}
 	\label{fig:mulkeen_pll}
@@
 The decisions made in the design of the \ac{DCO}, or \ac{NCO}, determine many of the other \ac{ADPLL} parameters.
-While tuning range and centre frequency, as well as linearity, carry over from the analogue counterpart, a \ac{DCO} also has a frequency step which in combination with the bit width of the control signal determines the range over which the oscillator can be tuned.
+While tuning range and centre frequency, as well as linearity, carry over from the analogue counterpart, \iac{DCO} also has a frequency step which in combination with the bit width of the control signal determines the range over which the oscillator can be tuned.
 Figure \ref{fig:my_ring} illustrates a basic ring oscillator design.
@@
 
-It is also possible to implement an \ac{NCO} by means of a counter, in a manner that will produce either linear period or linear frequency steps.
+It is also possible to implement \iac{NCO} by means of a counter, in a manner that will produce either linear period or linear frequency steps.
 Both methods use the most significant bit of the counter's value to form the output signal.
@@
 \subsection{Digital Phase Detector}
-Once again quantisation impacts the \acl{PD}, as rather than a continuous output the phase detector in an \ac{ADPLL} has a finite number of output values, thus limiting the accuracy of the phase detector.
+Once again quantisation impacts the \acl{PD}, as rather than a continuous output the phase detector in \iac{ADPLL} has a finite number of output values, thus limiting the accuracy of the phase detector.
 A second form of quantisation is also present, as unlike an analog system, a digital phase detector does not provide continuous data in the time domain either, instead relying on sampling.
@@
 As the output only has two levels the resultant word is only 1 bit wide and as such, limits the range over which the output frequency can be controlled.
-More complex designs such as that in Figure \ref{fig:shan_bb_pd}, implemented by Shan, build on this by measuring the time difference between edges of the signals using a \ac{TDC} in his case using a \ac{TDL} \cite{shan2014phd}.
-A \ac{TDL} is constructed by a chain of elements of a fixed delay, and the signal to be timed is applied to this start of this chain.
+More complex designs such as that in Figure \ref{fig:shan_bb_pd}, implemented by Shan, build on this by measuring the time difference between edges of the signals using \iac{TDC} in his case using \iac{TDL} \cite{shan2014phd}.
+\Iac{TDL} is constructed by a chain of elements of a fixed delay, and the signal to be timed is applied to this start of this chain.
 After the timing interval elapsed the values at each point in the chain are examined, and using temperature coding, these are converted to a digital signal, the width of which is the bit width of the \ac{PD}'s error signal.
@@
 \end{figure}
-The \acl{LF} in an \ac{ADPLL} can be implemented as a \ac{PI} controller, as only a low-pass filter is required, such as that in Figure \ref{fig:my_simple_pi}.
-In the case of a node in an \ac{ADPLL} network the input of this filter is a weighted of the phase difference relative to the neighbouring local clocking areas.
+The \acl{LF} in \iac{ADPLL} can be implemented as \iac{PI} controller, as only a low-pass filter is required, such as that in Figure \ref{fig:my_simple_pi}.
+In the case of a node in \iac{ADPLL} network the input of this filter is a weighted of the phase difference relative to the neighbouring local clocking areas.
 In one example topology a digital system the proportional section can be implemented by a simple multiplier, whereas the integral path is constructed by adding the result of a multiplication by the proportional gain to an accumulator.
@@
 \subsection{Error Combiner}
-The \ac{ADPLL}s used in a network need to combine the error signals from multiple neighbours to determine what the average difference from its neighbours is, and this necessitates the addition of the Error Combiner.
+The \acp{ADPLL} used in a network need to combine the error signals from multiple neighbours to determine what the average difference from its neighbours is, and this necessitates the addition of the Error Combiner.
 In a digital system this can be implemented by a weighted average of the different error signals, with the weight being modifiable at run-time.
 This configurability is what permits the system to implement uni-directional mode and also allows for the weighting applied to certain signals, such as the external reference, to be modified.
-The ease of implementation of a configurable error combiner is one of the main advantages of an \ac{ADPLL} over an analogue system.
+The ease of implementation of a configurable error combiner is one of the main advantages of \iac{ADPLL} over an analogue system.
 
 \section{The Role of the FPGA}
-A \acl{FPGA} is a type of \ac{IC} that is designed to be configured by a designer after the chip itself has been manufactured.
-An \ac{FPGA} contains a large number of logic elements that can be connected together in order to perform complex logic, written using the same \ac{HDL}s used to design the digital blocks of \ac{ASIC}s.
+\Iacl{FPGA} is a type of \ac{IC} that is designed to be configured by a designer after the chip itself has been manufactured.
+\Iac{FPGA} contains a large number of logic elements that can be connected together in order to perform complex logic, written using the same \aclp{HDL} (\acsp{HDL}) used to design the digital blocks of \aclp{ASIC} (\acsp{ASIC}).
 They may also implement started modules such as adders, multiplexers and \ac{RAM} as a fundamental element.
@@
 High end devices such as the Xilinx Zynq Ultrascale even implement Multi-Processor \ac{SOC}s.
-Compared to an \ac{ASIC} the designer does not have direct control over the layout of the system but rather describes its behaviour, possibly down to the basic logic elements of inverters or other gates.
+Compared to \iac{ASIC} the designer does not have direct control over the layout of the system but rather describes its behaviour, possibly down to the basic logic elements of inverters or other gates.
 Limited control is possible over the placement of the individual modules, but \ac{EDA} tools are responsible for the exact placement of elements.
@@
 
-Prototyping on an \ac{FPGA} is a common verification stage for conventional \ac{ASIC} designs as it allows for a hardware validation of any digital circuitry, and the detection of any potential flaws or errors made by the design before the expensive of an \ac{ASIC} implementation.
-In their 2013 \ac{ADPLL} network implementation Zianbetov and Shan used an \ac{FPGA} in order to validate their programming interface, the design of the error processing block, ensure they had eliminated mode locking behaviour and to ensure phase synchronisation was possible \cite{zianbetov2013phd,shan2014phd}.
+Prototyping on \iac{FPGA} is a common verification stage for conventional \ac{ASIC} designs as it allows for a hardware validation of any digital circuitry, and the detection of any potential flaws or errors made by the design before the expensive of \iac{ASIC} implementation.
+In their 2013 \ac{ADPLL} network implementation Zianbetov and Shan used \iac{FPGA} in order to validate their programming interface, the design of the error processing block, ensure they had eliminated mode locking behaviour and to ensure phase synchronisation was possible \cite{zianbetov2013phd,shan2014phd}.
 However they experienced two main limitations, they were not able to implement the mixed-signal \ac{PFD} and \ac{DCO}, and the maximum frequency of operation possible was orders of magnitude lower than the GHz range of their \ac{ASIC} implementation.
@@
 
-A similar technique was used by Lata \textit{et al} (2013) to implement a \ac{FPGA} clocked oscillator with a $200~\si{\kilo\hertz}$ centre frequency, however it is hard to understand exactly what either the goal was or the specifics of the \ac{DCO} used \cite{lata2013adpll}.
+A similar technique was used by Lata \textit{et al} (2013) to implement \iac{FPGA} clocked oscillator with a $200~\si{\kilo\hertz}$ centre frequency, however it is hard to understand exactly what either the goal was or the specifics of the \ac{DCO} used \cite{lata2013adpll}.
 
-It is however possible to implement limited mixed-signal circuits on an \ac{FPGA} through the use of primitive logic elements, however, as control over the implementation is restricted to the module level it is not possible to mirror the implementation of a design intended for an \ac{ASIC}.
-As these implementations are analogous to a mixed-signal circuit on an \ac{ASIC}, the verification of theoretical behaviours, as done by Koskin \textit{et all}, in order to test the findings of his PhD thesis \cite{theboys2019}, is made possible without the expense of \ac{ASIC} fabrication.
-An \ac{FPGA} based mixed-signal \ac{ADPLL} network is also seeing use here in \ac{UCD} as a initial prototyping platform for the validation of new modules for use in \ac{ASIC} based \ac{ADPLL} networks.
+It is however possible to implement limited mixed-signal circuits on \iac{FPGA} through the use of primitive logic elements, however, as control over the implementation is restricted to the module level it is not possible to mirror the implementation of a design intended for \iac{ASIC}.
+As these implementations are analogous to a mixed-signal circuit on \iac{ASIC}, the verification of theoretical behaviours, as done by Koskin \textit{et all}, in order to test the findings of his PhD thesis \cite{theboys2019}, is made possible without the expense of \ac{ASIC} fabrication.
+\Iac{FPGA} based mixed-signal \ac{ADPLL} network is also seeing use here in \ac{UCD} as a initial prototyping platform for the validation of new modules for use in \ac{ASIC} based \ac{ADPLL} networks.
 
@@
 
-Mixed-signal circuits were, however, simulated in hardware by \'{O}scar Luc\'{i}a \textit{et al} (2011) on an \ac{FPGA} in order to overcome the excessive time penalty imposed by software based simulations that required the behaviour of both a digital and mixed-signal peripheral and that of a micro controller running code in C to be simulated side by side \cite{lucia2011real}.
+Mixed-signal circuits were, however, simulated in hardware by \'{O}scar Luc\'{i}a \textit{et al} (2011) on \iac{FPGA} in order to overcome the excessive time penalty imposed by software based simulations that required the behaviour of both a digital and mixed-signal peripheral and that of a micro controller running code in C to be simulated side by side \cite{lucia2011real}.
 They concluded
@@
 \end{quotation}
-By carrying out simulations on an \ac{FPGA}, Guanhua Wang \textit{et al} (2013) achieved a 3000 times decrease in run-time when compared to an identical simulation in MATLAB used for the verification of a calibration algorithm for successive approximation analogue-to-digital converters \cite{wang2013fast}.
+By carrying out simulations on \iac{FPGA}, Guanhua Wang \textit{et al} (2013) achieved a 3000 times decrease in run-time when compared to an identical simulation in MATLAB used for the verification of a calibration algorithm for successive approximation analogue-to-digital converters \cite{wang2013fast}.
 Many other examples can be found of \ac{FPGA}s used for the simulation of mixed-signal or \acs{RF} circuits.
@@
 
-The simplest form of clock performance characterisation is on a \ac{C2C} basis, in which the reference is the signal under test itself.
+The simplest form of clock performance characterisation is on \iac{C2C} basis, in which the reference is the signal under test itself.
 Here the standard deviation of the individual periods gives the jitter of the signal, while skew has no meaning with a self reference the mean period can be used to compute the centre frequency over the time interval.
@@
 Long term jitter in particular affects \acs*{RADAR} as it will manifest itself as a Doppler shift in the return signal.
-For an \ac{PLL} network, with an appropriately designed filter, the long term jitter should be zero so long as the reference signal is stable.
+For \iac{PLL} network, with an appropriately designed filter, the long term jitter should be zero so long as the reference signal is stable.
 %TODO ask Brian about this one

### 2019-04-21 thesis/CodeAppendix.tex
start of conc

A selection of files containing the core Verilog modules are included in this appendix. All files used can be located on github at  \href{https://github.com/ConchuOD/fpga_adpll}{\textbf{https://github.com/ConchuOD/fpga\_adpll}}.
\section{Phase Detector}\label{lst:PhaseDetector}
\lstinputlisting[style={verilog-style}]{../design_sources/PhaseDetector.v}
\lstinputlisting[style={verilog-style}]{../design_sources/UpDownCounter.v}

\section{Phase Detector with \acs{TDL}}\label{lst:phase_detector_dl}
\lstinputlisting[style={verilog-style}]{../design_sources/PhaseDetectorDL.v}

\section{Error Combiner}\label{lst:error_combiner}
\lstinputlisting[style={verilog-style}]{../design_sources/ErrorCombiner.v}

\section{Loop Filter}\label{lst:loop_filter}
\lstinputlisting[style={verilog-style}]{../design_sources/LoopFilter.v}

\section{Ring Oscillator}\label{lst:ring_osc}
\lstinputlisting[style={verilog-style}]{../design_sources/RingOsc.v}

\section{\acs{FPGA} Clocked Oscillator}\label{lst:phase_accum}
\lstinputlisting[style={verilog-style}]{../design_sources/PhaseAccum.v}

\section{Network \acs{ADPLL} (\acs{RO})}\label{lst:network_ring_adpll}
\lstinputlisting[style={verilog-style}]{../design_sources/NetworkRingADPLL.v}

\section{Network \acs{ADPLL} (\ac{FPGA} Clocked)}\label{lst:network_adpll}
\lstinputlisting[style={verilog-style}]{../design_sources/NetworkRingADPLL.v}

### 2019-04-21 thesis/Conclusion.tex
start of conc

@@
 \section{Conclusions}
+%TODO dont reset acro
+The objective of this project was to create \iac{FPGA} based platform that could be used to prototype and model \ac{ADPLL} networks in a cost and time effective manner.
+The main effort of the project was to analyse the ways in which the components of an \ac{ADPLL} network could be designed, then select from this pool a number of the most appropriate designs.
+A number of \ac{ADPLL} designs were then created using the selected components in different combinations for the purposes of comparison.
+Once these \ac{ADPLL} designs had been validated, they were implemented in networks of a 2x2 and 3x3 Cartesian grid topology.
+Measurements and performance benchmarks were then carried out on these \ac{ADPLL} networks, and the results used to verify claims made regarding the intended use cases of the individual \ac{ADPLL} designs.
+Finally various small modifications to the \ac{ADPLL} designs were made, and their impact on system performance identified.
+
+%TODO did I manage this?
+
 \section{Suggested Future Work}
@@
 Until this point the non-square wave output had eliminated the design from consideration.
-However, as in an \ac{ASIC} linear period designs are easier to implement, it would be ideal to have the benefit of frequency scaled hardware simulations with a \ac{DCO} that more closely relates the the eventual system.
+However, as in \iac{ASIC} linear period designs are easier to implement, it would be ideal to have the benefit of frequency scaled hardware simulations with \iac{DCO} that more closely relates the the eventual system.
 

### 2019-04-21 thesis/FPGABasedADPLLs.tex
start of conc

@@
 \section{Chapter Overview}
-The first step in creating an \ac{FPGA} based network of \ac{ADPLL}s is the design of the \ac{ADPLL} itself, which will be addressed in this chapter.
-The nature of an \ac{FPGA} necessitates a number of compromises in the design of a given block which limits transferability to \ac{ASIC} designs.
-In this chapter the potential designs for each individual block, or module, investigated will be explained and the case for their selection in an \ac{FPGA} based \ac{ADPLL} made.
-A number of blocks have implement purely digital circuitry and as such can be transferred in their entirety from an \ac{FPGA} and vice versa.
+The first step in creating \iac{FPGA} based network of \acp{ADPLL} is the design of the \ac{ADPLL} itself, which will be addressed in this chapter.
+The nature of \iac{FPGA} necessitates a number of compromises in the design of a given block which limits transferability to \ac{ASIC} designs.
+In this chapter the potential designs for each individual block, or module, investigated will be explained and the case for their selection in \iac{FPGA} based \ac{ADPLL} made.
+A number of blocks have implement purely digital circuitry and as such can be transferred in their entirety from \iac{FPGA} and vice versa.
 However, those that will be used to emulate mixed-signal circuitry, such as the \ac{DCO} and \ac{PFD}, will be examined in greater detail.
-A \ac{RTL} diagram of the \ac{ADPLL} topology that will be assumed in the following sections is shown in Figure \ref{fig:network_adpll}.
+\Iac{RTL} diagram of the \ac{ADPLL} topology that will be assumed in the following sections is shown in Figure \ref{fig:network_adpll}.
 \begin{figure}[h]
@@
 Borrowing matrix indexing, and choosing the oscillator located at index $i,j$, $i,j \neq 1,N$ where $N$ is the number of oscillators in each direction, comparison need only be done with the signals generated by oscillators located at $i-1,j$ and $i,j-1$.
-The comparison with oscillators $i+1,j$ and $i,j+1$ will be carried out in those \ac{ADPLL}s, and the negation of the errors measured used in the \ac{ADPLL} located at $i,j$.
+The comparison with oscillators $i+1,j$ and $i,j+1$ will be carried out in those \acp{ADPLL}, and the negation of the errors measured used in the \ac{ADPLL} located at $i,j$.
 
@@
 The choices made in the design of the \ac{DCO} have the greatest impact on the effectiveness of the overall platform and which use cases the \ac{ADPLL} containing it are suitable for, as the key performance benchmarks are all done using the waveform this block generates.
-This project will address three distinct designs of \ac{ADPLL}s suitable for implementation on an FPGA, two derived from the clocks generated by the \ac{FPGA}s own distribution network and one generated independently of this clock, using a chain of inverters.
-These are not the only ways in which an oscillator could be synthesised on an \ac{FPGA}, however other designs were deemed to be unsuitable for extensible and portable implementations.
+This project will address three distinct designs of \acp{ADPLL} suitable for implementation on an FPGA, two derived from the clocks generated by the \acp{FPGA} own distribution network and one generated independently of this clock, using a chain of inverters.
+These are not the only ways in which an oscillator could be synthesised on \iac{FPGA}, however other designs were deemed to be unsuitable for extensible and portable implementations.
 
@@
 
-The main issue with the creation of \ac{DCO}s on an \ac{FPGA} is the inability to create mixed-signal circuits, such as those that would be intended for use on an \ac{ASIC}.
+The main issue with the creation of \acp{DCO} on \iac{FPGA} is the inability to create mixed-signal circuits, such as those that would be intended for use on \iac{ASIC}.
 As such the \ac{FPGA} based oscillator must emulate the behaviour of a mixed-signal circuit in some way.
@@
 As the only variable here is the control code, period step of this design is $T_{FPGA}$, thereby providing period linearity with respect to the control code.
-This is the key advantage of this design, as most \ac{ASIC} implementations of a \ac{DCO} are also linear in period.
+This is the key advantage of this design, as most \ac{ASIC} implementations of \iac{DCO} are also linear in period.
 The other main reason to choose this design is that its \ac{FPGA} clocked nature allows for exact control over the frequency of operation, and the number of tunable parameters make it possible to configure multiple ways to achieve the same frequencies of operation.
-Combined these attributes make it very easy to create an oscillator that emulates the behaviour of a design intended for an \ac{ASIC}, however, at a greatly reduced frequency.
+Combined these attributes make it very easy to create an oscillator that emulates the behaviour of a design intended for \iac{ASIC}, however, at a greatly reduced frequency.
 This restriction on the frequency of operation arises out of the period step size, which in order to obtain a good resolution must be orders of magnitude smaller than the intended period to be generated.
@@
 
-In being \ac{FPGA} clocked this design has pseudo-deterministic characteristics, with each period step being almost identical across oscillators and over the entire tuning range, unlike an \ac{ASIC} where process variation will impact the layout of a high frequency oscillator.
+In being \ac{FPGA} clocked this design has pseudo-deterministic characteristics, with each period step being almost identical across oscillators and over the entire tuning range, unlike \iac{ASIC} where process variation will impact the layout of a high frequency oscillator.
 The only variation in this design will come, ironically, from jitter or skew in the \ac{FPGA}'s clock distribution network, which as the frequencies will typically be in the low hundreds of MHz is very minor.
@@
 \subsection{Ring Oscillator}
-The best approximation of the non-idealities of an \ac{ASIC} \ac{DCO} implementation can be obtained using a \ac{RO}, also know as an inverter ring, as the \ac{DCO}.
-In an \ac{ASIC} an \ac{RO} is constructed by connecting a number of NOT gates back in a chain, with the output of the last NOT gate fed input the input of the first.
-As NOT gates are not available on an \ac{FPGA}, they are replaced by inverter primitives.
+The best approximation of the non-idealities of \iac{ASIC} \ac{DCO} implementation can be obtained using \iac{RO}, also know as an inverter ring, as the \ac{DCO}.
+In \iac{ASIC} \iac{RO} is constructed by connecting a number of NOT gates back in a chain, with the output of the last NOT gate fed input the input of the first.
+As NOT gates are not available on \iac{FPGA}, they are replaced by inverter primitives.
 The inverter ring is an inherently unstable circuit that leverages the propagation delay through an odd number of inverters to create a waveform that, when viewed at a fixed position, is a square wave.
@@
 
-The main advantage of this type of oscillator stems from its similarity to that used on an \ac{ASIC} which leads to many non-idealities carrying over from the truly mixed-signal implementation.
+The main advantage of this type of oscillator stems from its similarity to that used on \iac{ASIC} which leads to many non-idealities carrying over from the truly mixed-signal implementation.
 Rather than being just a tool to validate that other modules, the performance of this oscillator can be analysed more deeply as well as the performance of networks containing it.
-Koskin \textit{et al} used such a design to test theories proposed regarding the synchronisation of oscillators during his PhD using this type of oscillator implemented on an \ac{FPGA} \cite{koskin2019phd,theboys2019}.
-Also in this design's favour is the period step size, four times the inverter propagation delay, which was found experimentally to be than in the region of a nanosecond on Xilinx Artix-7 and Zynq \ac{FPGA}s.
+Koskin \textit{et al} used such a design to test theories proposed regarding the synchronisation of oscillators during his PhD using this type of oscillator implemented on \iac{FPGA} \cite{koskin2019phd,theboys2019}.
+Also in this design's favour is the period step size, four times the inverter propagation delay, which was found experimentally to be than in the region of a nanosecond on Xilinx Artix-7 and Zynq \acp{FPGA}.
 
 While it is advantageous for the purposes of analysis that this design is affected by variation in temperature, implementation and other non-idealities this brings with it challenges not present in a true mixed-signal design.
-In a design intended for an \ac{ASIC} the design has near total control over the placement of circuit elements, however, the \ac{FPGA} based counterpart control is available over just the area in which the oscillator may lie, and the \ac{EDA} is responsible for the exact placement.
+In a design intended for \iac{ASIC} the design has near total control over the placement of circuit elements, however, the \ac{FPGA} based counterpart control is available over just the area in which the oscillator may lie, and the \ac{EDA} is responsible for the exact placement.
 Unlike the \ac{ASIC} implementation, in which the propagation delay through each element of the ring is well bounded, the lack of control over placement means that the propagation time through each inverter is not well bounded and may vary wildly.
 This makes it time consuming to ensure that the frequency tuning ranges of each oscillator match that desired, and use of this oscillator less straight-forward than that of the previous two designs.
-This is compounded by the simulation tools in \ac{EDA}s intended for \ac{FPGA}s being unable to accurately simulate the time delays, thus negating any benefit of test benches for the alignment process.
+This is compounded by the simulation tools in \ac{EDA} packages intended for \acp{FPGA} being unable to accurately simulate the time delays, thus negating any benefit of test benches for the alignment process.
 
 \section{Frequency Divider}
-Many \ac{PLL}s, be they digital or analogue, incorporate a frequency divider in the feedback path so that comparison can be made with a reference operating at a lower frequency.
+Many \aclp{PLL} (\acsp{PLL}), be they digital or analogue, incorporate a frequency divider in the feedback path so that comparison can be made with a reference operating at a lower frequency.
 This is a feature that is advantageous for \ac{ADPLL} networks, as comparison using this divided clock allows the synchronisation signal sent between modules to be of a frequency orders of magnitude lower than that of the clock distribution network and thus require circuitry that consumes significantly less energy.
-This division can be obtained trivially on an \ac{FPGA} by a number of means, of which this thesis will mention just two.
+This division can be obtained trivially on \iac{FPGA} by a number of means, of which this thesis will mention just two.
 The first of these methods allows for integer division, by inverting a signal each time a counter reaches a certain value, $k$.
@@
 \section{Phase Detector}
-Three potential \acl{PD} designs will be addressed in this thesis, sign-only detection with a Bang-Bang Detector, \ac{FPGA} clocked detection using a \ac{FSM} and an emulation of a \ac{TDL} using inverter primitives to create a delay, similarly to the \ac{RO} seen previously.
+Three potential \acl{PD} designs will be addressed in this thesis, sign-only detection with a Bang-Bang Detector, \ac{FPGA} clocked detection using \iac{FSM} and an emulation of \iac{TDL} using inverter primitives to create a delay, similarly to the \ac{RO} seen previously.
 Both designs will have to have a representation for both situations where the generated signal leads and lags the reference signal, and for the magnitude of this difference, a problem most obviously solved by using a signed binary representation of the difference.
@@
 While sign detection clocked by the \ac{FPGA} can be performed in a variety of ways, a state machine is a straightforward way in which to accomplish this.
-A \acl{FSM} consists of three main components: the states themselves, the next state logic and the output logic.
+\Iacl{FSM} consists of three main components: the states themselves, the next state logic and the output logic.
 In this case the next state logic is the generated and reference signals, and the output logic sets the count instructions for the up-down counter as well as ensuring the phase error output is held constant between measurement periods.
@@
 The modified Bang-Bang detector establishes which signal occurred first and generates a control signal that is analogous to the time delay between a rising edge on each input signal.
-A \ac{TDL} then uses this control signal to convert the phase difference between the signals to a digital signal.
+\Iac{TDL} then uses this control signal to convert the phase difference between the signals to a digital signal.
 This magnitude is combined with the sign detected by the bang-Bang detector to form a two's complement error signal.
@@
 Instead two flip-flops are used, each clocked on an input signal with their $D$ inputs connected to logic high.
-The $Q$ outputs of these registers are connected to an \acs{XOR} gate which generates a pulse as long as a rising edge has only been seen on one signal.
+The $Q$ outputs of these registers are connected to \iacs{XOR} gate which generates a pulse as long as a rising edge has only been seen on one signal.
 Once a second rising edge occurs the \acs{XOR} output will be set to 0, thus terminating the pulse.
@@
 
-To determine the sign itself, which clock edge first had a rising edge must be recorded, and this can be done by setting a \ac{SR} latch if the rising edge of the generated signal occurs first and resetting in the case of the reference clock having the first rising edge.
+To determine the sign itself, which clock edge first had a rising edge must be recorded, and this can be done by setting \iac{SR} latch if the rising edge of the generated signal occurs first and resetting in the case of the reference clock having the first rising edge.
 Looking at the whole measurement cycle however, it is apparent that all 4 combinations possible of two inputs occur, putting the \ac{SR} latch into which the output is undefined.
@@
 This, however, ignores the potential for metastability which may arise if both signals have a rising edge within an extremely short period of time.
-While it may in other systems be feasible to dismiss this occurrence as a low probability event, the goal of a \ac{PLL} is to reduce the time delay between rising edges to zero which increases the possibility of this occurring, which may lead to indeterminate behaviour in other parts of the feedback network.
-In order to avoid this in an \ac{ASIC}, an arbitration circuit is placed between the \ac{SR} latch and the register storing the sign between measurements, which will default to one of the two signals in this case.
-This is however a mixed signal circuit, such as that proposed by Tierno \textit{et al} (2008) and implemented by Zianbetov and Shan \cite{tierno2008wide,zianbetov2013phd,shan2014phd}, which requires two \ac{PMOS} and two \ac{NMOS} transistors and cannot be implemented on an \ac{FPGA}.
+While it may in other systems be feasible to dismiss this occurrence as a low probability event, the goal of \iac{PLL} is to reduce the time delay between rising edges to zero which increases the possibility of this occurring, which may lead to indeterminate behaviour in other parts of the feedback network.
+In order to avoid this in \iac{ASIC}, an arbitration circuit is placed between the \ac{SR} latch and the register storing the sign between measurements, which will default to one of the two signals in this case.
+This is however a mixed signal circuit, such as that proposed by Tierno \textit{et al} (2008) and implemented by Zianbetov and Shan \cite{tierno2008wide,zianbetov2013phd,shan2014phd}, which requires two \ac{PMOS} and two \ac{NMOS} transistors and cannot be implemented on \iac{FPGA}.
 As the output of the \ac{SR} latch will be indeterminate prior to the detection of the first edge, the impact of both edge detection registers' output is that the \ac{SR} latch attempts to hold this indeterminate value as an output.
-To avoid this unwanted behaviour the arbiter present on an \ac{ASIC} is replaced by a second \ac{SR} latch, connected as in Figure \ref{fig:arbitration}.
+To avoid this unwanted behaviour the arbiter present on \iac{ASIC} is replaced by a second \ac{SR} latch, connected as in Figure \ref{fig:arbitration}.
 This ensures the value presented to the register storing the sign is always determinate.
@@
 To complement the asynchronous sign detection circuit, an asynchronous \ac{TDC} is also required, as otherwise there would be no benefit to using the above method over the more easily understood \ac{FSM} based design as quantisation to the clock period would happen in the \ac{TDC} rather than the sign detection circuitry.
-The standard method for implementing a \ac{TDC} on an \ac{ASIC} is through a tapped delay line, where a signal is propagated through a chain of fixed length delays, and at the end of the measurement interval it is recorded whether or not the signal has propagated to each point in the chain.
+The standard method for implementing \iac{TDC} on \iac{ASIC} is through a tapped delay line, where a signal is propagated through a chain of fixed length delays, and at the end of the measurement interval it is recorded whether or not the signal has propagated to each point in the chain.
 This architecture is illustrated in Figure \ref{fig:simple_tdc}.
-As with the mixed signal circuitry previously addressed, a \ac{TDL} cannot be exactly replicated on an \ac{FPGA}, but through the use of inverters it is possible to replace the unrealisable well bounded delays.
-As with the \ac{RO}, the delay though the inverters is not nearly as well bounded, and significant variation between delays, in both individual \ac{TDL}s and within the same \ac{TDL}, due to the implementation occurs This variation notwithstanding, the inverter pairs are a suitable replacement, and this method of time-to-digital conversion provides both resolution advantages and the presence of implementation based variation over a clocked design.
+As with the mixed signal circuitry previously addressed, \iac{TDL} cannot be exactly replicated on \iac{FPGA}, but through the use of inverters it is possible to replace the unrealisable well bounded delays.
+As with the \ac{RO}, the delay though the inverters is not nearly as well bounded, and significant variation between delays, in both individual \acp{TDL} and within the same \ac{TDL}, due to the implementation occurs This variation notwithstanding, the inverter pairs are a suitable replacement, and this method of time-to-digital conversion provides both resolution advantages and the presence of implementation based variation over a clocked design.
 These prove particularly beneficial when attempting to verify theoretical results.
@@
 In a regular \ac{ADPLL} an error combining circuit is not required as there is never more than one reference signal, and as such the input of this block would be equivalent to the output.
-However in an \ac{ADPLL} network multiple reference signals are required, one for each neighbour, thus requiring some method to combine.
+However in \iac{ADPLL} network multiple reference signals are required, one for each neighbour, thus requiring some method to combine.
 Simply averaging these signals was not found to be viable by Pratt and Nyugen in their original paper for analogue systems, however, the work of Javidan \textit{et al}, showed that in a digital system this method will not result in mode locking, provided start-up is performed uni-directionally \cite{javidan2011all}.
@@
 \section{Loop Filter}
-The \acl{LF} in an \ac{ADPLL} can be implemented as a \ac{PI} controller, a circuit that can be implemented in a wide variety of ways, with the exact implementation having little impact on behaviour.
-The first choice to be made is whether the designer wishes to implement the controller as an \ac{FIR} or \ac{IIR} filter.
+The \acl{LF} in \iac{ADPLL} can be implemented as \iac{PI} controller, a circuit that can be implemented in a wide variety of ways, with the exact implementation having little impact on behaviour.
+The first choice to be made is whether the designer wishes to implement the controller as \iac{FIR} or \ac{IIR} filter.
 Both methods will produce a controller that fundamentally behaves the same way, but will have a starkly different implementation.
@@
 Various tools, Matlab for example, can be used to compute the coefficients and number of taps required to obtain a specific filter response.
-Using an \ac{FIR} filter is advantageous as it is always stable due to the lack of a feedback path, however a large number of taps are required to obtain a frequency response akin to that of the \ac{IIR} implementation, thus requiring an equal number of multipliers and $taps-1$ adders.
-Compared to an \ac{IIR}, which requires just a two adders, two multipliers and a flip-flop this has a much greater hardware footprint.
+Using \iac{FIR} filter is advantageous as it is always stable due to the lack of a feedback path, however a large number of taps are required to obtain a frequency response akin to that of the \ac{IIR} implementation, thus requiring an equal number of multipliers and $taps-1$ adders.
+Compared to \iac{IIR}, which requires just a two adders, two multipliers and a flip-flop this has a much greater hardware footprint.
 It is possible to condense the \ac{FIR} multiplication and addition to a single multiplier and adder, depicted in \ref{fig:fir_pi} (b), however this will take as many clock cycles to complete as there are taps.
@@
 \subsection{\acs{IIR} \acl{LF}}
-A \ac{PI} controller can be easily implemented as an \acl{IIR} filter in hardware, as in Figure \ref{fig:iir_pi}, requiring significantly fewer circuit elements than an \ac{FIR} based design as the long series of taps and multipliers is not required, as instead integration is performed by a single multiplier and an accumulator.
+\Iac{PI} controller can be easily implemented as \iacl{IIR} filter in hardware, as in Figure \ref{fig:iir_pi}, requiring significantly fewer circuit elements than \iac{FIR} based design as the long series of taps and multipliers is not required, as instead integration is performed by a single multiplier and an accumulator.
 As before the frequency response of the controller can be calculated using the following equation, where $\alpha$ is the proportional and $\beta$ the integral gain:

### 2019-04-21 thesis/Introduction.tex
79git statusgit statusgit statusgit statusgit statusgit statusgit statusgit statusgit statusgit statusgit statusgit status
start of conc

@@
 This thesis will put forward the \ac{FPGA} as a tool in the design of \ac{ADPLL} networks to bridge the gap between software simulations and implementation in custom silicon by providing a hardware-based simulation, modelling and validation platform.
-While an FPGA lacks the direct control over the schematic and layout that an \ac{ASIC} provides, the hardware nature of this platform enables the analysis of system dynamics that are not easily modelled in simulation and the testing.
+While an FPGA lacks the direct control over the schematic and layout that \iac{ASIC} provides, the hardware nature of this platform enables the analysis of system dynamics that are not easily modelled in simulation and the testing.
+\acp{ADPLL} are an entirely digital implementation of \iac{PLL}, advantageous as this allows for easier integration into the modern high transistor count \acl{IC} (\acs{IC}s) for which \ac{ADPLL} networks are a proposed clock distribution system.
 
@@
 
-An \ac{FPGA} based test platform is ideal for those who wish to examine system dynamics without the time delays, financial burden or expertise required to develop a complete mixed-signal system on an \ac{ASIC} but retain the ability to realistically simulate the behaviour of an \ac{ADPLL} network.
+\Iac{FPGA} based test platform is ideal for those who wish to examine system dynamics without the time delays, financial burden or expertise required to develop a complete mixed-signal system on \iac{ASIC} but retain the ability to realistically simulate the behaviour of \iac{ADPLL} network.
 The result of this project is such a platform, designed to be extensible, with flexibility built into each component/module used.

### 2019-04-21 thesis/NetworkImplemenation.tex
start of conc

@@
 The \ac{HDL} Verilog was selected for use in this project due to my prior exposure to that language.
-5 MHz was chosen as the target centre frequency for these \ac{ADPLL}s, during initial testing it was discovered that the waveform output via the Digilent Nexys4 development board's \ac{GPIO} ports began to degrade around the 10 MHz mark, preventing accurate analysis.
+5 MHz was chosen as the target centre frequency for these \acp{ADPLL}, during initial testing it was discovered that the waveform output via the Digilent Nexys4 development board's \ac{GPIO} ports began to degrade around the 10 MHz mark, preventing accurate analysis.
 
@@
 
-The three architectures chosen for implementation represent a progression from entirely an \ac{FPGA} clock driven system to an \ac{ADPLL} with all components asynchronous with respect to that clock.
+The three architectures chosen for implementation represent a progression from entirely \iac{FPGA} clock driven system to \iac{ADPLL} with all components asynchronous with respect to that clock.
 Therefore the first \ac{ADPLL} design features a clocked oscillator and phase detector, specifically the linear frequency design mentioned in Chapter \ref{chap:3} and the clocked phase detector using a state machine and a bi-directional counter.
@@
 \subsection{Generic Components}
-As mentioned in the chapter overview, a number of components were used in all \ac{ADPLL}s implemented, as changes in their design would not impact the overall system.
+As mentioned in the chapter overview, a number of components were used in all \acp{ADPLL} implemented, as changes in their design would not impact the overall system.
 \subsubsection{Error Combiner}
@@
 Regardless of the number of input error signals, a division by four is then performed in order to compute the average error, done by shifting the value right by two.
-In a \ac{HDL}, unlike C programming or similar, a shift requires no computation and is instead implemented by performing a part-selection.
+In \iac{HDL}, unlike C programming or similar, a shift requires no computation and is instead implemented by performing a part-selection.
 As all error signals are two's complemented signed integers, the error combination process is also carried out using signed values.
@@
 \subsubsection{Loop Filter}
-Two \acl{LF}s were implemented, both as \ac{IIR} filters.
-\ac{FIR} was dismissed due to the extra hardware required to implement the calculations within a single clock cycle, as the \ac{LF} is clocked using the oscillator output, and the greater difficulty of gain adjustment compared to an \ac{IIR}.
+Two \aclp{LF} were implemented, both as \ac{IIR} filters.
+\ac{FIR} was dismissed due to the extra hardware required to implement the calculations within a single clock cycle, as the \ac{LF} is clocked using the oscillator output, and the greater difficulty of gain adjustment compared to \iac{IIR}.
 Clocking on using the generated signal ensures that the discrete time integration is only carried out once per phase comparison.
-A pitfall that may be encountered implementing an \ac{ADPLL} featuring a divider is not clocking the module on the divided clock, which will result in the integration being carried out at the oscillator output frequency.
+A pitfall that may be encountered implementing \iac{ADPLL} featuring a divider is not clocking the module on the divided clock, which will result in the integration being carried out at the oscillator output frequency.
 
@@
 
-Figure \ref{fig:integer_lf} contains an \ac{RTL} diagram describing the implementation.
+Figure \ref{fig:integer_lf} contains \iac{RTL} diagram describing the implementation.
 Important to note is the shift applied to the result of the multiplication by \acs{kp} which, as the integration is performed with a larger width to avoid accumulator overflow, preserves the intended relationship between proportional and integral gains.
@@
 
-Figure \ref{fig:osc2_impl} contains an \ac{RTL} diagram of this oscillator's implementation.
+Figure \ref{fig:osc2_impl} contains \iac{RTL} diagram of this oscillator's implementation.
 As phase error is a two's complement signed value, its addition to the bias must be carried out using signed arithmetic.
@@
 On the right side of the diagram the output register can be seen.
-Clocked using the \ac{FSM}s measurement interval complete signal, this register preserves the phase error until the termination of the following interval.
+Clocked using the \acp{FSM} measurement interval complete signal, this register preserves the phase error until the termination of the following interval.
 \begin{figure}[h]
@@
 This plot is derived from data captured using an Agilent MSO7054A, using a setup that will be further explained in Section \ref{section:measurement_setup}.
-Despite lacking a non ideal reference, less violent initial oscillation, and a lesser until lock could be visually determined, the simulation provided a good insight into the system's behaviour once implemented on an \ac{FPGA}.
+Despite lacking a non ideal reference, less violent initial oscillation, and a lesser until lock could be visually determined, the simulation provided a good insight into the system's behaviour once implemented on \iac{FPGA}.
 
@@
 The listed components were brought together in \textsc{NetworkADPLL.v}, attached in Appendix \ref{adx:code} Listing \ref{lst:network_adpll}.
-Important to note is that each \ac{ADPLL} only contains two \acl{PD}s despite during network operation the need for phase comparisons to be made with up to four neighbours.
-The two ``missing'' \ac{PD}s are in fact synthesised as part of the other \ac{ADPLL}s, with which the comparison would be made.
-This allows for a 50\% reduction in the number of \ac{PD}s required, while also ensuring identical error is received by both \ac{ADPLL}s.
-Each \ac{ADPLL} performs comparison with the \ac{DCO} ``above'' and ``left'' of them in the Cartesian grid, and a port in the module outputs the results for use by the other \ac{ADPLL}s involved.
+Important to note is that each \ac{ADPLL} only contains two \aclp{PD} despite during network operation the need for phase comparisons to be made with up to four neighbours.
+The two ``missing'' \acp{PD} are in fact synthesised as part of the other \acp{ADPLL}, with which the comparison would be made.
+This allows for a 50\% reduction in the number of \acp{PD} required, while also ensuring identical error is received by both \acp{ADPLL}.
+Each \ac{ADPLL} performs comparison with the \ac{DCO} ``above'' and ``left'' of them in the Cartesian grid, and a port in the module outputs the results for use by the other \acp{ADPLL} involved.
 %TODO juvenile wording
@@
 The second \ac{ADPLL} design implemented during the course of this project was somewhat of a stepping stone between fully synchronous and asynchronous to the \ac{FPGA} clock.
-The clocked phase detector is reused from above, albeit with an altered error width, while the \ac{DCO} is replaced by a \acl{RO}.
+The clocked phase detector is reused from above, albeit with an altered error width, while the \ac{DCO} is replaced by \iacl{RO}.
 While the phase detector was known to work due to testing in \ac{ADPLL} 1, the \ac{DCO} could not be simulated using accurate timings so all testing had to be done in hardware.
@@
 \subsubsection{\acl{DCO}}
-\ac{ADPLL} 2 exchanges the \ac{FPGA} clocked oscillator for \ac{RO}, with the aim of obtaining performance more akin to that of an \ac{ASIC} based mixed-signal implementation, due to the variation in period steps both between oscillators and within the same oscillator due to the oscillator layout.
+\ac{ADPLL} 2 exchanges the \ac{FPGA} clocked oscillator for \ac{RO}, with the aim of obtaining performance more akin to that of \iac{ASIC} based mixed-signal implementation, due to the variation in period steps both between oscillators and within the same oscillator due to the oscillator layout.
 Also advantageous is the improvement in period resolution obtained using this design, when compared to that of the \ac{FPGA} clocked design.
@@
 \end{figure}
-Figure \ref{fig:ro_impl} contains an \ac{RTL} diagram of the \ac{RO} as implemented.
+Figure \ref{fig:ro_impl} contains \iac{RTL} diagram of the \ac{RO} as implemented.
 The inverter chain is specified with a maximum length of $n$ and a maximum number of removable inverters of $k$.
@@
 An inverter chain will oscillate freely and cannot be disabled.
-To this end the first inverter in the chain was replaced by a \acs{NAND} gate to grant control over its operation.
+To this end the first inverter in the chain was replaced by \iacs{NAND} gate to grant control over its operation.
 
@@
 As mentioned in previous chapters, the absence of direct control over layout can lead to significant variation of the propagation time though inverters.
-This may occur in three ways: Firstly, the layout of individual \ac{RO}s may be significantly different, thus resulting in poor overlap of tuning regions.
+This may occur in three ways: Firstly, the layout of individual \acp{RO} may be significantly different, thus resulting in poor overlap of tuning regions.
 Anecdotally, while implementing a 3x3 network, 7 of the 9 oscillators had centre frequencies within a $100~\si{\kilo\hertz}$ span but two lay more than $500~\si{\kilo\hertz}$ away in opposite directions which prevented the network from locking.
-Secondly, within an \ac{RO} the propagation delay between each inverter may vary, which results in a variable frequency step, possibly changing the locking range.
-These two effects represent an extreme version of the variation due to process or manufacturing seen on \ac{ASIC}s.
+Secondly, within \iac{RO} the propagation delay between each inverter may vary, which results in a variable frequency step, possibly changing the locking range.
+These two effects represent an extreme version of the variation due to process or manufacturing seen on \acp{ASIC}.
 The final variation is possibly the most frustrating, and occurs when between implementations the \ac{EDA}, Vivado in the case of this project, changes the layout of the \ac{RO}.
@@
 
-When, later in the project, additional \ac{RO}s were implemented to form networks, 373 inverters proved a suitable starting point which consistently delivered \ac{DCO}s with a centre frequency within $200~\si{\kilo\hertz}$ of $5~\si{\mega\hertz}$.
-For the purposes of describing a typical tuning range, the average propagation delay of an \ac{RO} consisting of 373 inverters with a centre frequency of $5.06~\si{\mega\hertz}$ will be used.
-To avoid the aforementioned issues with constant control codes, the fixed code was achieved by implementing an \ac{ADPLL} and setting the \ac{PI} filter gains to zero at runtime before performing a reset.
+When, later in the project, additional \acp{RO} were implemented to form networks, 373 inverters proved a suitable starting point which consistently delivered \acp{DCO} with a centre frequency within $200~\si{\kilo\hertz}$ of $5~\si{\mega\hertz}$.
+For the purposes of describing a typical tuning range, the average propagation delay of \iac{RO} consisting of 373 inverters with a centre frequency of $5.06~\si{\mega\hertz}$ will be used.
+To avoid the aforementioned issues with constant control codes, the fixed code was achieved by implementing \iac{ADPLL} and setting the \ac{PI} filter gains to zero at runtime before performing a reset.
 
@@
 \acs{ADPLL} Design 3 has no dependence on the clock provided by the \ac{FPGA}'s clock management utility, as both the \ac{RO} and the \ac{TDC} are designed using inverters, while a Bang-Bang detector replaces the \ac{FPGA} clock reliant \ac{FSM} of the two previous designs.
-This introduces a second source of implementation based variation between \ac{ADPLL}s, however, this has significantly lesser impact on the frequency range of the oscillator and no impact on the centre frequencies.
-The lack of an \ac{FPGA} clocked element results means this design will have the closest behaviour to that of an \ac{ASIC} based \ac{ADPLL} implementation, although as the level of variation due to implementation is much larger on an \ac{FPGA}, the impact of non-idealities will be exaggerated.
+This introduces a second source of implementation based variation between \acp{ADPLL}, however, this has significantly lesser impact on the frequency range of the oscillator and no impact on the centre frequencies.
+The lack of \iac{FPGA} clocked element results means this design will have the closest behaviour to that of \iac{ASIC} based \ac{ADPLL} implementation, although as the level of variation due to implementation is much larger on \iac{FPGA}, the impact of non-idealities will be exaggerated.
 %same interface
-In order to allow quick exchange of \ac{PD}s or other modules, the module in which this \ac{ADPLL} is implemented is still retains the \ac{FPGA} clock as an input.
+In order to allow quick exchange of \acp{PD} or other modules, the module in which this \ac{ADPLL} is implemented is still retains the \ac{FPGA} clock as an input.
 In fact, this module shares its implementation with that of \ac{ADPLL} 2, as while the \ac{PD} designs are drastically different, the interface is identical.
@@
 \subsubsection{Phase Detector}
-Instead of an \ac{FPGA} clocked phase detector, \ac{ADPLL} 3 uses the modified Bang-Bang detector and \ac{TDL} approach to phase comparison seen in Section \ref{section:signum}.
+Instead of \iac{FPGA} clocked phase detector, \ac{ADPLL} 3 uses the modified Bang-Bang detector and \ac{TDL} approach to phase comparison seen in Section \ref{section:signum}.
 Sign detection is implemented exactly as described in that section, with the double \ac{SR} latch approach used to determine the sign of the phase error.
@@
 As each tap consists of an inverter pair, the delay through the tap, $\tau_{tap} \approx 600~\si{\pico\second}$.
-Characterisation of the \ac{TDL}, and thus measurement of the delays in each \ac{TDL}, is technically possible, however, this is an time consuming process as many \ac{TDL}s would need to be characterised in order to accurately compute the average delay.
+Characterisation of the \ac{TDL}, and thus measurement of the delays in each \ac{TDL}, is technically possible, however, this is an time consuming process as many \acp{TDL} would need to be characterised in order to accurately compute the average delay.
 
@@
 Compared to the oscillators suggested earlier, \ac{ADPLL} 3 achieves a significantly finer detection resolution, at the expense of significant variation in the detection step.
-This can be viewed as advantageous however, as the \ac{ADPLL} exhibits an exaggerated version of the implementation and manufacturing defect based variability present in an \ac{ASIC}.
+This can be viewed as advantageous however, as the \ac{ADPLL} exhibits an exaggerated version of the implementation and manufacturing defect based variability present in \iac{ASIC}.
 It is this type of \ac{ADPLL} that both the \acs{UCD} and Sorbonne research teams are using to examine potential designs and test theoretical results.
@@
 \section{Measurement Setup}\label{section:measurement_setup}
-For the following sections in which the performance of either individual \ac{ADPLL}s or \ac{ADPLL} networks the same measurement setup was used.
+For the following sections in which the performance of either individual \acp{ADPLL} or \ac{ADPLL} networks the same measurement setup was used.
 The \ac{FPGA} used was a Xilinx Artix-7 \acl*{Nexys} on a Digilent \acs*{Nexys} board.
@@
 Data was acquired with the oscilloscope in quad channel mode at a sampling rate of $2\si[per-mode=symbol]{\giga\sample\per\second}$.
-One of the four channels was consumed by the external reference, leaving three free for signal measurements, in most cases the generated output of three \ac{ADPLL}s were connected.
+One of the four channels was consumed by the external reference, leaving three free for signal measurements, in most cases the generated output of three \acp{ADPLL} were connected.
 $200\si{\micro\second}$ long captures were saved to an Agilent specified binary format which could be opened in Matlab for analysis.
@@
 \section{\acs{ADPLL} Characterisation}
-Before combining the \ac{ADPLL}s in a network it was important to analyse the designs locked to an ideal reference, in order to establish a baseline performance.
-These tests were carried out with the \ac{ADPLL}s locked to an external reference at $5~\si{\mega\hertz}$.
+Before combining the \acp{ADPLL} in a network it was important to analyse the designs locked to an ideal reference, in order to establish a baseline performance.
+These tests were carried out with the \acp{ADPLL} locked to an external reference at $5~\si{\mega\hertz}$.
 As \ac{ADPLL} 2 \& 3 have significant variance in the step sizes, for these oscillators range and step based measurements were obtained by measuring a number of oscillators, and typical values will be shown.
-Jitter was calculated based on the performance on nine individual \ac{ADPLL}s all locked simultaneously to the $5~\si{\mega\hertz}$ external reference.
-While jitter and skew are the key indicators of \ac{ADPLL} network performance, when it comes to the individual \ac{ADPLL}s themselves it is important to check lock and capture ranges also, to check that the implemented designs match with their intended values.
-The lock and capture ranges are given based on typical values, whereas jitter and skew are the median values from the \ac{ADPLL}s tested.
+Jitter was calculated based on the performance on nine individual \acp{ADPLL} all locked simultaneously to the $5~\si{\mega\hertz}$ external reference.
+While jitter and skew are the key indicators of \ac{ADPLL} network performance, when it comes to the individual \acp{ADPLL} themselves it is important to check lock and capture ranges also, to check that the implemented designs match with their intended values.
+The lock and capture ranges are given based on typical values, whereas jitter and skew are the median values from the \acp{ADPLL} tested.
 
 The \ref{table:characterisation} contains the results of this analysis.
-The first thing to note is that to the $\si{\kilo\hertz}$ the lock and capture ranges of the \ac{ADPLL}s implementing \ac{RO}s are identical.
+The first thing to note is that to the $\si{\kilo\hertz}$ the lock and capture ranges of the \acp{ADPLL} implementing \acp{RO} are identical.
 As the lock range would be expected to be slightly wider than the capture range in a design limited by the loop filter, it can be deduced that the limiting factor in the tuning range of the \ac{ADPLL} is the \ac{DCO} frequency range.
@@
         Jitter                        & $1.8385~\si{\nano\second}$         & $0.5623~\si{\nano\second}$         & $0.7410~\si{\nano\second}$          \\
-        Skew (Mean \acs{TIE} to Ref.) & $1.7439~\si{\nano\second}$         & $0.3741~\si{\nano\second}$         & $0.6381~\si{\nano\second}$          \\
+        Skew (Me\iacs{TIE} to Ref.) & $1.7439~\si{\nano\second}$         & $0.3741~\si{\nano\second}$         & $0.6381~\si{\nano\second}$          \\
         Maximum \acs{TIE}             & $7.2840~\si{\nano\second}$         & $6.4770~\si{\nano\second}$         & $6.4578~\si{\nano\second}$          \\
@@
 \section{\acs{ADPLL} Network Implementation}
-As the \ac{ADPLL}s were designed with network usage in mind, the \ac{HDL} aspect of their integration into a network is straightforward.
+As the \acp{ADPLL} were designed with network usage in mind, the \ac{HDL} aspect of their integration into a network is straightforward.
 For \ac{ADPLL} 1, the oscillators need only be wired together, but for Design 2 \& 3 which are reliant on inverters require some care to ensure their lock ranges overlap sufficiently.
-This is a time consuming task, as a number of iterations may be required, even when locking down the cells of lined up \ac{RO}s, as the nets forming the connections may change.
+This is a time consuming task, as a number of iterations may be required, even when locking down the cells of lined up \acp{RO}, as the nets forming the connections may change.
 It is recommended only to carry out alignment when all other aspects of the design are unlikely to change and to save the bitstream generated.
-The projected time requirement in aligning \ac{RO}s meant that the creation of a 4x4 network was abandoned.
-\ac{FPGA} clocked \ac{ADPLL}s are more suited to larger networks as their centre point can be determined with ease.
+The projected time requirement in aligning \acp{RO} meant that the creation of a 4x4 network was abandoned.
+\ac{FPGA} clocked \acp{ADPLL} are more suited to larger networks as their centre point can be determined with ease.
 
-The network operates using an external reference which is input over a \ac{PMOD} header that is shared with a number of output signals, however it could be configured to use an internal reference generated by either the \ac{FPGA} clock management utility or a \ac{DCO} of the designer's choice.
+The network operates using an external reference which is input over \iac{PMOD} header that is shared with a number of output signals, however it could be configured to use an internal reference generated by either the \ac{FPGA} clock management utility or \iac{DCO} of the designer's choice.
 
 In order to increase the flexibility of the network, the switches on the \acs*{Nexys} are used to change both the proportional and integral gains and the configuration of the network at runtime.
-The reconfigurability aspect of the error combiner allows for the network to operate in either uni- or bi-directional mode, or as a number of individual \ac{PLL}s all connected to the external reference.
+The reconfigurability aspect of the error combiner allows for the network to operate in either uni- or bi-directional mode, or as a number of individual \aclp{PLL} (\acsp{PLL}) all connected to the external reference.
 This is particularly advantageous for the inverter based designs, as multiple combinations can be measured without reimplementing, and thus changing the layout of either oscillator or \acl{PD}.
@@
 A 3x3 network was the minimum goal, with the aforementioned but eventually not implemented 4x4 a stretch-goal if time permitted.
-A 3x3 network set as the primary goal, as it is the minimum size network in which there exists an \ac{ADPLL} with four neighbours.
-As will be explained in Chapter \ref{chap:5}, \ac{ADPLL} Design 1 was only implemented in a 2x2 configuration while both other \ac{ADPLL}s were tested in a 3x3 configuration.
+A 3x3 network set as the primary goal, as it is the minimum size network in which there exists \iac{ADPLL} with four neighbours.
+As will be explained in Chapter \ref{chap:5}, \ac{ADPLL} Design 1 was only implemented in a 2x2 configuration while both other \acp{ADPLL} were tested in a 3x3 configuration.
 

### 2019-04-21 thesis/TestingAnalysis.tex
start of conc

@@
 
-There are two main trends worth highlighting here, with the first of these being as the feedback network becomes more complicated a large degree of skew is introduced, with the magnitude becoming larger in the \ac{ADPLL}s further from the reference.
+There are two main trends worth highlighting here, with the first of these being as the feedback network becomes more complicated a large degree of skew is introduced, with the magnitude becoming larger in the \acp{ADPLL} further from the reference.
 This would appear to be attributable to a combination of the time taken for a change in the reference to propagate to the \ac{ADPLL} node in question, and the fineness of period and detection resolutions.
@@
 The other trend, is the variation in \ac{C2C} jitter across the \ac{ADPLL} designs, with coarseness of the \ac{FPGA} clocked \ac{DCO} becoming somewhat obvious.
-As free \ac{PLL}s, the clocked \ac{DCO} has almost identical \ac{C2C} jitter and maximum \ac{TIE}, implying that only two period steps are used the majority of the time.
+As free \aclp{PLL} (\acsp{PLL}), the clocked \ac{DCO} has almost identical \ac{C2C} jitter and maximum \ac{TIE}, implying that only two period steps are used the majority of the time.
 As these steps are somewhat further apart than that of the \ac{RO}, at $3.875~\si{\nano\second}$ apart as opposed to an implementation dependant value in the region of $1.1765~\si{\nano\second}$, a much greater jitter is observed.
@@
 Subsequent to tests carried out using a 2x2 network, the corresponding tests were carried out using a 3x3 network in order to analyse the impact of a larger network.
-For these tests the gains were left unchanged but, as the 2x2 and 3x3 networks have different floorplans, the \ac{RO}s have changed somewhat.
+For these tests the gains were left unchanged but, as the 2x2 and 3x3 networks have different floorplans, the \acp{RO} have changed somewhat.
 This has especially manifested itself in the exceptional individual \ac{PLL} performance of Design 3.
@@
 As \ac{ADPLL} Design 1 suffered significantly more with skew as the node indices increased, so it is no surprise that in the larger network the problem has been amplified, meaning this design is not well suited for use in larger networks at this frequency due to poor resolution.
-This does not, however, impact the main role of this design method, which is the emulation of designs intended for use in \ac{ASIC}s as the frequency of operation will be much lower than the $5~\si{\mega\hertz}$ used in these tests, which will bring with it a finer resolution.
+This does not, however, impact the main role of this design method, which is the emulation of designs intended for use in \aclp{ASIC} (\acsp{ASIC}) as the frequency of operation will be much lower than the $5~\si{\mega\hertz}$ used in these tests, which will bring with it a finer resolution.
 %TODO fill in this after measurements
@@
 Figure \ref{fig:gain_sweep} contains a sweep across the possible range of integral gains for four different proportional gains.
-This test was carried out with a 3x3 network of \ac{ADPLL} Design 3, each locked to the external reference at $5~\si{\mega\hertz}$, and the generated waveform from the \ac{ADPLL}s at locations ``11'', ``22'' and  ``33'' were captured.
+This test was carried out with a 3x3 network of \ac{ADPLL} Design 3, each locked to the external reference at $5~\si{\mega\hertz}$, and the generated waveform from the \acp{ADPLL} at locations ``11'', ``22'' and  ``33'' were captured.
 It can clearly be observed that there are two distinct regions in the graph, with jitter values of less than $1~\si{\nano\second}$ contrasting sharply with the points with jitter in the region of $10~\si{\nano\second}$.
@@
 This can be visualised effectively using Matlab's \texttt{histplot}.
-In order to obtain data, three \ac{ADPLL}s implementing both the \ac{FPGA} clocked and inverter based \ac{DCO}s were locked to an external reference signal at $5~\si{\mega\hertz}$, and the period of each clock cycle measured.
+In order to obtain data, three \acp{ADPLL} implementing both the \ac{FPGA} clocked and inverter based \acp{DCO} were locked to an external reference signal at $5~\si{\mega\hertz}$, and the period of each clock cycle measured.
 \begin{figure}[h]%
@@
 \subsection{Impact of Frequency Divider}
-The major benefit of an \ac{PLL} network over other forms of coupled oscillator based clock distribution systems, is that the coupling can be done using a signal significantly reduced in frequency, thereby reducing the power consumption of the related hardware.
+The major benefit of \iac{PLL} network over other forms of coupled oscillator based clock distribution systems, is that the coupling can be done using a signal significantly reduced in frequency, thereby reducing the power consumption of the related hardware.
 It is therefore important to ensure that the insertion of a feedback divider does not introduce any further jitter to the system.
@@
 As re-implemenation was performed between each measurement, there is some variation present.
-However, it is notable that for each \ac{ADPLL} that saw a \ac{C2C} jitter increase, there was another that performed better than at a lower level of division.
+However, it is notable that for each \ac{ADPLL} that saw \iac{C2C} jitter increase, there was another that performed better than at a lower level of division.
 Accordingly, it can be claimed that the insertion of a feedback divider does not impact the \ac{C2C} jitter present.
@@
 This means the rate at which the \ac{LF} is updated also drops, and therefore the control code is changes $n$ times less frequently, where $n$ is the division level.
-The dividers reduce the speed at which the individual \ac{ADPLL}s can react to a phase difference, while locking the resulting control code in place for a greater number of clock cycles as the division level increases.
+The dividers reduce the speed at which the individual \acp{ADPLL} can react to a phase difference, while locking the resulting control code in place for a greater number of clock cycles as the division level increases.
 This greater duration may cause a control code to be applied for an excessively long number of cycles, resulting in a drift of the rising edge of the divided/generated signal away from the reference, which will manifest as \ac{TIE}.
@@
     \centering
-    \includegraphics[width=0.8\textwidth]{../bad_waveform}
+    \includegraphics[width=0.8\textwidth]{../dist2}
     \caption[Period distribution at differen divider levels]{Period distribution at differen divider levels.}    
@@
 
-As mentioned when \ac{FPGA} clocked designs were introduced, their configurability makes them particularly suited to the validation and verification of designs destined for use in \ac{ASIC}s, as was the case with Zianbetov and Shan \cite{zianbetov2013phd,shan2014phd}.
+As mentioned when \ac{FPGA} clocked designs were introduced, their configurability makes them particularly suited to the validation and verification of designs destined for use in \acp{ASIC}, as was the case with Zianbetov and Shan \cite{zianbetov2013phd,shan2014phd}.
 In contrast to the \ac{RO} based designs, in which configurability is more more restricted, it is possible to replicate the resolution, centre frequencies and other \ac{ADPLL} characteristics of the design requiring verification at a scaled down frequency.
@@
 In order to reduce the frequency to this extent with inverters over $30$ thousand would be required, introducing the potential for even greater variation in centre frequencies due to implementation.
-The other issue arising out of this number of inverters is the floorspace of the \ac{FPGA} that would be consumed implementing just one \ac{RO}, in the case of the Xilinx \acl{Nexys} and extrapolating based on the resource usage of the \ac{RO}s implemented, a single oscillator would consume half the look-up tables available, eliminating the possibility of a network.
+The other issue arising out of this number of inverters is the floorspace of the \ac{FPGA} that would be consumed implementing just one \ac{RO}, in the case of the Xilinx \acl{Nexys} and extrapolating based on the resource usage of the \acp{RO} implemented, a single oscillator would consume half the look-up tables available, eliminating the possibility of a network.
 
@@
 
-Figure \ref{fig:dists}, which demonstrated the impact of period distribution, highlights a second major unrealistic behaviour present in \ac{FPGA} clocked \ac{ADPLL} designs, as an \ac{ASIC} destined design will not exhibit this identical linear period distribution.
+Figure \ref{fig:dists}, which demonstrated the impact of period distribution, highlights a second major unrealistic behaviour present in \ac{FPGA} clocked \ac{ADPLL} designs, as \iac{ASIC} destined design will not exhibit this identical linear period distribution.
 While \ac{ADPLL} 2 will feature a large degree of variation due to implementation differences, these differences are only present in the \ac{RO} not the \ac{PFD}.
@@
 Accordingly, if the aim of this platform is the enhancement of a model, the validation of theoretical results, or analysing the performance of new control or performance enhancement hardware the \ac{RO} based designs are idea for this purpose.
-Two members of the research team here in \acs{UCD}, Eugene Koskin and Pierre Bisiaux, have been using an \ac{ADPLL} with the same architecture as Design 3 for these purposes.
+Two members of the research team here in \acs{UCD}, Eugene Koskin and Pierre Bisiaux, have been using \iac{ADPLL} with the same architecture as Design 3 for these purposes.

### 2019-04-21 thesis/Conclusion.tex
draft 1 totally complete

--- 2019-04-20 thesis/Conclusion.tex
@@
 \section{Conclusions}
+%TODO dont reset acro
+The objective of this project was to create \iacs{FPGA} based platform that could be used to prototype and model \acs{ADPLL} networks in a cost and time effective manner.
+The main effort of the project was to analyse the ways in which the components of \iacs{ADPLL} network could be designed, then select from this pool a number of the most appropriate designs.
+A number of \acs{ADPLL} designs were then created using the selected components in different combinations for the purposes of comparison.
+Once these \acs{ADPLL} designs had been validated, they were implemented in networks of a 2x2 and 3x3 Cartesian grid topology.
+Measurements and performance benchmarks were then carried out on these \acs{ADPLL} networks, and the results used to verify claims made regarding the intended use cases of the individual \acs{ADPLL} designs.
+Finally various small modifications to the \acs{ADPLL} designs were made, and their impact on system performance identified.
+
+The project resulted in a platform comprising three \acs{ADPLL} designs suited for different use cases.
+As evidenced by the investigation into the impact of minor variations, this system fulfils the objective that it be suitable for the testing of architectural changes.
+Realistic, though exaggerated, variation due to implementation makes the inverter based designs ideal for the testing of new control schemes or performance enhancing modules.
+Simultaneously, this makes designs based on \acsp{RO}   ideal for the confirmation of either theoretical or simulated results, in a platform that suffers from realistic implementation based variation, and non-ideal effects such as power supply noise that are difficult to simulate well.
+The presence of \acs{FPGA} clocked \acsp{DCO} and \acsp{PFD} makes the emulation of mixed-signal circuits at a scaled down frequency for the purposes of design verification possible.
+
+As the platform is designed with extensibility in mind, rapid changes in the design are possible, with the lengthy and resource consuming tape-out process that would otherwise be required for a test in hardware, cut down to the time taken for the \ac{EDA} tool of choice to generate a configuration file.
+Whereas tape-out may be a costly process, for a figure in the $100$s of Euro, an \acs{FPGA} can be acquired will that bridge the gap between software simulations and implementation in custom silicon, by providing a hardware-based simulation, modelling and validation platform.
+
 \section{Suggested Future Work}
@@
 Unfortunately the tasks of floorplanning and oscillator alignment would still be left to the user to perform.
-
-%TODO cahracterise tdl

### 2019-04-21 thesis/TestingAnalysis.tex
lol locations

@@
 These distinct sections of the plot represent locked and unlocked operation respectively.
-Zooming in on the low jitter region in Figure \ref{fig:gain_sweep_zoom}, the loss of lock thresholds can be compared with the theoretical relationship of $k_i = 0.5\times k_p$.
+Zooming in on the low jitter region in Figure \ref{fig:gain_sweep_zoom}, the approximate loss of lock thresholds can be compared with the theoretical relationship of $k_i = 0.5\times k_p$.
+The approximated loss of lock is determined as gain setting that caused the first \ac{ADPLL} to lose lock.
 \begin{table}[!h]
@@
             $\frac{1}{16}$            & $\frac{5}{64}$            & $0.8$ \T\B\\
-            $\frac{2}{16}$            & $\frac{10}{64}$           & $0.8$                     \T\B\\
-            $\frac{3}{16}$            & $\frac{12}{64}$           & $1$                       \T\B\\
-            $\frac{4}{16}$            & $\frac{14}{64}$           & $1.1428$                  \T\B
+            $\frac{2}{16}$            & $\frac{11}{64}$           & $0.72$ \T\B\\
+            $\frac{3}{16}$            & $\frac{13}{64}$           & $0.92$ \T\B\\
+            $\frac{4}{16}$            & $\frac{15}{64}$           & $1.07$ \T\B
         \end{tabular}

### 2019-04-23 thesis/Abstract.tex
partial corrections

--- 2019-04-05 thesis/Abstract.tex
@@
 Low power, high frequency clock distribution systems will be in ever increasing demand in the near future as the need for high performance digital circuitry grows.
-At these frequencies, however, the conventional clock distribution systems are unable provide a clock signal of adequate quality without compromising on either of these problems.
-Many devices have turned away from using Globally Synchronous clock distribution systems in favour of those that divide the area of the chip into Globally Asynchronous but Locally Synchronous areas.
-However, new technologies seek to enable the use of Globally Synchronous methods at high frequencies, one of which being the use of oscillators coupled in phase, each responsible for delivering the clock to a subregion of the chip.\\
-Each oscillator forms part of a Phase Lock Loop (PLL), and in order to enable the synchronisation of the network each PLL is linked those controlling the adjacent clock regions.
-For a digital system it is expedient to implement these PLLs digitally as an All-Digital Phase Lock Loop (ADPLL) as this provides a number of advantages.
+At these frequencies, however, the conventional clock distribution systems are unable to provide a clock signal of adequate quality without compromising on power consumption.
+Many device designers have turned away from using Globally Synchronous clock distribution systems, in favour of those that divide the area of the chip into Globally Asynchronous but Locally Synchronous areas.
+However, new technologies seek to enable the use of Globally Synchronous methods at high frequencies, one of which is the use of oscillators coupled in phase, each responsible for delivering the clock to a subregion of the chip.\\
+Each oscillator forms part of a Phase Lock Loop (PLL), and to enable the synchronisation of the network, each PLL is linked those controlling the adjacent clock regions.
+For a digital system, it is expedient to implement these PLLs digitally, as an All-Digital Phase Lock Loop (ADPLL), as this provides a number of advantages.
 It has been shown in theory and experimentally that this method can produce the high quality clock desired with low power consumption.\\
-As the production of test chips is expensive and time consuming through simulation and validation of a design is vital, traditionally carried out for mixed signal circuitry using complex behavioural and theoretical models.
-For a ADPLL Network the consistency of a signal both with respect to itself, and to the other clock signals on the chip is a key performance attribute and much of the variation is due to random processes that may be difficult to simulate effectively.\\
+As the production of test chips is expensive and time consuming, through simulation and validation of a design is vital For mixed-signal circuits, his is traditionally carried out using complex behavioural and theoretical models.
+For an ADPLL network, the consistency of a signal, both with respect to itself, and to the other clock signals on the chip, is a key performance attribute and much of the variation is due to random processes that may be difficult to simulate effectively.\\
 This thesis will demonstrate that the Field Programmable Gate Array (FPGA) can be used in order to simulate, model or validate ADPLL network architectures in a cost and time effective manner, as a complement to conventional methods.
-The key benefit is that many system dynamics that will be seen when a network is implemented on an Application Specific Integrated Circuit (ASIC), but be overlooked in software simulations can be examined in the hardware simulation that an FPGA provides.\\
+The key benefit is that many system dynamics that will be seen when a network is implemented on an Application Specific Integrated Circuit (ASIC), but may be overlooked in software simulations, can be examined in the hardware simulation that an FPGA provides.\\
 This thesis implements networks using three designs of ADPLL, each using a different architecture, and highlights the use cases to which each is best suited.
-The performance of each design is then analysed and this compared to the suggested use cases.
+The performance of each design is then analysed and compared to the suggested use cases.
 Additionally the impact of more minor architectural modifications is tested and documented.

### 2019-04-23 thesis/Background.tex
partial corrections

@@
 \section{Brief Overview}
-In a world where the demand for high performance hand-held computing devices continues to grow and the prevalence of ``smart'' devices is increasing, there is unprecedented demand for \ac{SOC} devices to control systems as varied as medical devices and entertainment systems.
+In a world where the demand for high performance hand-held computing devices continues to grow and the prevalence of ``smart'' devices is increasing, there is unprecedented demand for \ac{SOC} devices to control systems as varied as medical devices, watches and entertainment systems.
 As these applications become more and more demanding, with ever increasing amounts of data to process and the expectation that today's devices will outperform those of yesterday, the problem of maintaining the steady gain in performance of \acp{SOC} remains at the fore.
@@
 The main drivers of performance in \acp{SOC} are the number of transistors on a chip, which is correlated with the number of calculations that can be carried out simultaneously, and the frequency at which the device operates, which determines the number of calculations performed per second.
-Moore's Law, based on the famous observation by Gordon Moore in 1965\cite{moore1965cramming}, predicted a doubling in the transistor count of \acl{IC} (\acs{IC}s) per year for the forthcoming decade.
-This behaviour has carried on to this day as a result of the ever decreasing size of transistors, and has only begun to slow down in recent years.
-However, as the number of transistors on a chip has increased roughly following Moore's Law, the increase in clock frequency has not been able to follow a similar linear trajectory, having remained roughly equivalent for the last number of years \cite{ross2008cpu}, indeed the clock speed of the Intel Core family of \acl{CPU} (\acs{CPU}s) has not changed since their introduction in 2009 \cite{intelark}.
+Moore's Law, based on the famous observation by Gordon Moore in 1965\cite{moore1965cramming}, predicted a doubling in the transistor count of \aclp{IC} (\acsp{IC}) per year for the forthcoming decade.
+This behaviour has carried on to this day as a result of the ever decreasing size of transistors, and has only begun to slow down in recent years \cite{ross2008cpu}.
+However, as the number of transistors on a chip has increased roughly following Moore's Law, the increase in clock frequency has not been able to follow a similar linear trajectory, having remained roughly equivalent for the last number of years \cite{ross2008cpu}, indeed the clock speed of the Intel Core family of \acl{CPU} (\acs{CPU}s) has not changed since their introduction in 2009 \cite{intelark}, as shown in Figure \ref{fig:eldar_last_30_yrs}
+.
 \begin{figure}[h]
@@
 \end{figure}\\
-This plateauing of clock frequency has been caused by high power consumption due to the demands placed by the global distribution of a high frequency clock, often the single biggest consumer of power on the chip \cite{tiwari1998reducing}.
-With the growth of the \ac{IOT} market where low power devices are desirable, with many of the emerging uses of \acp{SOC} being portable and thus without a permanent power source, high power consumption goes directly against one of the key pillars of the technology.
-This forces many of these devices to use lower performance hardware in order to reduce the power consumption, and increase the battery life, of their devices.
+This plateauing of clock frequency has been caused by high power consumption, due to the demands placed by the global distribution of a high frequency clock, often the single biggest consumer of power on the chip \cite{tiwari1998reducing}.
+With the growth of the \ac{IOT} market where low power devices are essential, with many of the emerging uses of \acp{SOC} being portable and thus without a permanent power source, high power consumption goes directly against one of the key pillars of the technology.
+This forces many of these devices to use lower performance hardware in order to reduce the power consumption, and increase their battery life,.
 
@@
 
-In contrast in \iac{GALS} clock delivery system the ``local'' areas are not synchronised with other.
+In contrast, in \iac{GALS} clock delivery system the ``local'' areas are not synchronised with each other.
 This reduces the clocking system's complexity and thus the power consumption and chip area used, at the expense of communication speed between blocks.
-This disadvantage comes from the need to then somehow synchronise the messages being sent from one area to another to avoid the corruption of any messages.
-\Iac{GSLS}, system, however has the advantages of deterministic behaviour and greater rates of communication between clocking areas and, as such, remains a desirable system design.
-A number of methods which deliver \ac{GSLS} clocking exist at present such as clock trees as well as emerging technologies such as ADPLL networks.
+This disadvantage comes from the need to then somehow synchronise the messages being sent from one area to another to avoid message corruption.
+\Iac{GSLS}, system, however, has the advantages of deterministic behaviour and greater rates of communication between clocking areas and, as such, remains a desirable system design.
+A number of methods which deliver \ac{GSLS} clocking exist at present, such as clock trees, as well as emerging technologies like the ADPLL network.
 
@@
 A lack of synchronisation between the clock edges will manifest itself as a time difference between the clocking events at both registers, $\Delta T = t^i_k - t^f_k$.
-$\Delta T$ is considered to be ergodic and can be described by an average deviation called skew and random process, normally modelled as a Gaussian random variable.
-If $\Delta T$ is negative this reduces the time available for the intervening combinatorial network thereby, having the same effect as a reduction in clocking frequency.
+$\Delta T$ is considered to be ergodic and can be described by an average deviation called skew and random process called jitter, normally modelled as a Gaussian random variable.
+If $\Delta T$ is negative, this reduces the time available for the intervening combinatorial network, thereby having the same effect as a reduction in clocking frequency.
 Correspondingly a positive $\Delta T$ for depicted registers implies a negative $\Delta T$ for $R_f$ and the subsequent register.
-The most common sources of clock error are caused by mismatches which usually stem from production, such as differences in the length of clocking paths, buffer delays or in the parameters of either active or passive components in the clock distribution network, which as the size of components on \iac{IC} reduces becomes more difficult to avoid.
+The most common sources of clock error are caused by mismatches which usually stem from production, such as differences in the length of clocking paths, buffer delays or in the parameters of either active or passive components in the clock distribution network.
+As the size of components on \iac{IC} reduces, this mismatch becomes more difficult to avoid.
 All sources of mismatch will manifest themselves in the clock distribution system as skew between transistors, while the noise in active components or the power supply system will appear as jitter in the clock signal.
@@
 \section{Traditional Solutions}
-A number of traditional solutions exist which provide \ac{GSLS} clocking systems, using a variety of techniques.
+A number of traditional solutions exist that produce \ac{GSLS} clocking systems, using a variety of techniques.
 The most simple of these implement clock distribution systems that are symmetrical in order to distribute a centrally generated clock signal to all areas of the chip at the same phase.
@@
 
-A mesh clock distribution network is an alternate design where the clock is delivered using a Cartesian grid of distribution lines.
-Compared to a tree type system, the variation in skew seen with a clock mesh is inversely proportional to the density of the grid while the sources of jitter remain identical.
-According to Abdelhadi \textit{et al} (2010) clock meshes ``\textit{achieve low and deterministic skew, low skew variations, and low jitter}'', all desirable characteristics for a clock distribution system.
-However they dissipate more power due to extra capacitive loading, attributable to vast number of lines required to form the grid.
-Similarly mesh distribution networks suffer from potential mismatch in production and alleviation through increasing of the dimensions of interconnects will, as with a tree type system, lead to higher power consumption \cite{abdelhadi2010timing}.
+A mesh clock distribution network, as in Figure \ref{fig:mesh} is an alternate design where the clock is delivered using a Cartesian grid of distribution lines.
+Compared to a tree type system, the variation in skew seen with a clock mesh, is inversely proportional to the density of the grid while the sources of jitter remain identical.
+According to Abdelhadi \textit{et al} (2010) clock meshes ``\textit{achieve low and deterministic skew, low skew variations, and low jitter}''; all desirable characteristics for a clock distribution system.
+However, they dissipate more power due to extra capacitive loading, attributable to the vast number of lines required to form the grid.
+Similarly, mesh distribution networks suffer from potential mismatch in production and alleviation, through increasing of the dimensions of interconnects, will, as with a tree type system, lead to higher power consumption \cite{abdelhadi2010timing}.
 \begin{figure}[h]
@@
 \end{figure}
+
 Alternative designs replace the electrical lines used in the tree networks with waveguides for optical signals, with only the distribution in the local area carried out using regular wires.
 This technique presents many advantages \cite{chen2006chip}: optical clock delivery is immune to the noise sources that affect electrical clock distribution systems, consume less power and do not suffer from the electrical losses present in a regular tree system.
+However, this is a fledgling method, and sufficient research into materials and potential designs has yet to be carried out \ref{zianbetov2013phd,1648640}.
 
@@
 Two main strategies exist to provide skew compensation, each named according to the location of the control mechanism.
-Designs featuring the controller located at the clock source, are known as ``centralised'' methods, and those with multiple controllers in the individual clocking areas known as ``decentralised''.
-Regardless of the controller placement these techniques allow for the tuning of the propagation delay between the centralised clock source and the local clocking areas.
+Designs featuring the controller located at the clock source, are known as ``centralised'' methods, and those with multiple controllers in the individual clocking areas are known as ``decentralised''.
+Regardless of the controller placement, these techniques allow for the tuning of the propagation delay between the centralised clock source and the local clocking areas.
 
-In a centralised skew compensation circuit, the skew across the chip is calculated by the central controller which then manipulates the distribution network in order to deliver a more in-phase clock around the chip.
-This calculation is done by measuring the round trip time from the clock source to both the root of local clock tree, and to the individual ``leaves'' of the tree.
+In a centralised skew compensation circuit, the skew across the chip is calculated by the central controller, which then manipulates the distribution network in order to deliver a more in-phase clock around the chip.
+This calculation is done by measuring the round trip time from the clock source to both the root of the local clock tree, and to the individual ``leaves'' of the tree.
 The controller then has a limited ability to tune the propagation path.
-The downsides of this technique are the resolution of both the measurement and compensation are poor, allowing for the correction of just skew and not of any jitter that may be present in the system, and that the extra circuitry required for both the tunability of the forward path and the two extra return paths contribute to an increased footprint and power consumption.
+The downsides of this technique are that the resolution of both the measurement and compensation are poor, not permitting more than the correction of skew, and not of any jitter that may be present in the system.
+The extra circuitry required for both the tunability of the forward path, and the two extra return paths contributes to the increased footprint and power consumption of the distribution system.
 
-As the name suggest a decentralised skew compensation technique delegates the responsibility of tuning the propagation path to the individual clock regions.
+As the name suggests a decentralised skew compensation technique, delegates the responsibility of tuning the propagation path to the individual clock regions.
 This strategy has the advantage of not requiring the return paths present in a ``centralised'' design.
-Instead comparison is made betweeen the leaves of different clocking areas and on this basis the propagation delay is varied.
+Instead, comparison is made between the leaves of different clocking areas and, on this basis, the propagation delay is varied.
 For example, Yamashita \textit{et al} (2005) designed a system in which each clocking area or ``leaf node'' contains a partial clock tree.
@@
 
-%DOWN TO HERE DONE FOR ACRO
 \section{Multi-oscillator Designs}
-The designs described previously, are all similar in that they have a single central oscillator that provides the clock for all areas of the chip, whereas the following methods attempt to synchronise multiple oscillators, each of which provides the clock for a single clocking area.
-The main advantages of a multi-oscillator design, are that as each clocking area has its clock created locally, there is no degradation in the quality of the signal as it is distributed around the chip and the number of potential noise sources is reduced.
+The designs described previously are similar, in that they have a single central oscillator that provides the clock for all areas of the chip, whereas the following methods attempt to synchronise multiple oscillators, each of which provides the clock for its own clocking area.
+The main advantages of a multi-oscillator design are that as each clocking area has its clock created locally, there is no degradation in the quality of the signal as it is distributed around the chip and the number of potential noise sources is reduced.
 In order to obtain global synchronisation some method of comparison between local clocking areas is required, and how this is done depends on the architecture.
-Regardless of the comparison is made, it is carried out between neighbouring clocking areas and as such the feedback network need not have a large footprint or power overhead.
+Regardless of how the comparison is made, it is carried out between neighbouring clocking areas and, therefore, the feedback network need not have a large footprint or power overhead.
 
@@
 The advantage of this method is the simplicity of the feedback network, requiring just the divided clock output from a single leaf node.
-The \acp{VCO} then adjusted by the control voltage, $v_c$, which needs delivery to all areas of the chip.
-However, this is a regular signal and as such does not suffer from skew or jitter.
+The \acp{VCO} are then adjusted by the control voltage, $v_c$, which needs delivery to all areas of the chip.
+However, this is not a clock signal so does not suffer from skew or jitter.
 This alleviates the need for a power hungry distribution circuit, while also being more noise-immune than the transmission of a high frequency clock.
-However this design still suffers from clock variation as all \acp{VCO} are fed the same control voltage, and thus the manufacturing tolerance issues present in conventional designs persists here also.
+This design does still suffer from clock variation as all \acp{VCO} are fed the same control voltage, and thus the manufacturing tolerance issues present in conventional designs persist here also.
 This is acknowledged by the authors:
@@
 \end{quotation}
-This type of multi-oscillator design is implemented by analogue circuits, and as a result not only are the clock signals, but also the control signals are liable to variation due to noise, fabrication mismatch and power supply dynamics.
+This type of multi-oscillator design is implemented by analogue circuits, and as a result not only the clock signals, but also the control signals are liable to variation due to noise, fabrication mismatch and power supply dynamics.
 
-Another potential multi-oscillator clock distribution system uses the phase relationship between the oscillators driving neighbouring clock areas in order to obtain synchronisation.
+Another potential multi-oscillator clock distribution system uses the phase relationship between the oscillators driving neighbouring clock areas, in order to obtain synchronisation.
 Once again, this negates the requirement for a global distribution structure and the signals used for comparisons need only be sent between neighbouring clocking areas.
-As \iac{PLL} is being used it is again possible to perform the phase comparisons using a divided version of the generated clock.
+As \iac{PLL} is being used, it is again possible to perform the phase comparisons using a divided version of the generated clock.
 This in turn means the hardware transporting the divided clock signal to the phase comparator, has significantly lower requirements placed on it, thus lowering the power consumption due to electrical losses.
-Pratt and Nguyen initially proposed method of clock distribution in their 1995 paper entitled ``\textit{Distributed Synchronous Clocking}'' in which they propose a Cartesian grid of clocking areas, each with their own \ac{PLL}, which has become known as \iac{PLL} Network \cite{pratt1995distributed}.
+Pratt and Nguyen initially addressed this method of clock distribution in their 1995 paper entitled ``\textit{Distributed Synchronous Clocking}'' in which they proposed a Cartesian grid of clocking areas, each with their own \ac{PLL} This distribution method is known as \iac{PLL} Network \cite{pratt1995distributed}.
 In this design any given node is synchronised with its neighbours and one of the corner nodes is additionally synchronised with the reference.
-According to the authors this is ``a simple, effective way to achieve low cost, high quality, low skew clock generation in a synchronous parallel processor''.
-They did, however, note the presence of a phenomenon called ``mode locking'', which is setting of the network into a stable equilibrium where there are non-zero relative phases.
+According to the authors, this is ``a simple, effective way to achieve low cost, high quality, low skew clock generation in a synchronous parallel processor''.
+They did, however, note the presence of a phenomenon called ``mode locking'', which is the settling of the network into a stable equilibrium where there are non-zero relative phases.
 
@@
 This architecture of clock distribution network was then implemented by Gutnik \textit{et al} (2000) who fabricated a 4x4 array of oscillators, operating at a centre frequency of 1.2 MHz.
-The oscillator was implemented as a voltage controlled ``nMOS-loaded differential ring oscillator'', and in order to mode locking the phase detector was implemented as a highly non-linear circuit.
+The oscillator was implemented as a voltage controlled ``nMOS-loaded differential ring oscillator'', and in order to avoid mode locking the phase detector was implemented as a highly non-linear circuit.
 The design was a success and the authors concluded:
@@
 
-The remaining benefits of such a clock distribution system are: As the individual oscillators have their own control signal mismatch between different oscillators is not a factor as they will also receive different control signals.
-Secondly, and unlike the conventional methods, sources of jitter in the system such as power supply dynamics can be accounted for.
-Finally symmetry between the different oscillators is not required, once again attributable to the individual control signals in use.
+The remaining benefits of such a clock distribution system are: as the individual oscillators have their own control signal, mismatch between different oscillators is not a factor, as they will also receive different control signals.
+Secondly, and unlike the conventional methods, sources of jitter in the system, for example power supply dynamics, can be accounted for.
+Finally, symmetry between the different oscillators is not required, again attributable to the individual control signals in use.
+
 \section{ADPLL Networks}
-As \iac{PLL} network is an analog circuit, its integration in a modern \ac{IC} is a barrier to usage, and as such it has not been used in any commercial designs \cite{zianbetov2013distributed}.
-An alternative design that is more suitable for current fabrication techniques eschews from using analogue components and instead implements the network of controlled oscillators using only digital circuitry, hence the name All-Digital \ac{PLL}.
+As \iac{PLL} network is an analogue circuit, its integration in a modern \ac{IC} is a barrier to usage, and, therefore, it has not been used in any commercial designs \cite{zianbetov2013distributed}.
+An alternative design that is more suitable for current fabrication techniques, eschews from using analogue components and, instead, implements the network of controlled oscillators using only digital circuitry, hence the name All-Digital \ac{PLL}.
 A 4x4 \ac{ADPLL} network was designed and prototyped in 65 nm CMOS by Zianbetov and Shan in order to test the suitability of the technique as a clock distributor \cite{zianbetov2013phd,shan2014phd}.
 
-In this design the oscillators are once again laid out in a Cartesian grid, with each node coupled to their neighbours in phase.
-As this is now a digital system the coupling is carried out using digital phase comparators, which attempt to measure the phase difference between two oscillators.
+In this design, the oscillators are once again laid out in a Cartesian grid, with each node phase coupled to their neighbours.
+As this is now a digital system, the coupling is carried out using digital phase comparators, which attempt to measure the phase difference between two oscillators.
 Figure \ref{fig:eldar_node} shows high level detail of the architecture of both the entire clocking system and that of an individual node in the design.
-The digital nature of this architecture brings with it a number of advantages over traditional analogue implementations, as it can benefit from advancements in digital circuit design suites, be reconfigurable and programmable and has a significantly greater immunity to perturbations inherent to its digital nature, as the exact voltage of signals is of no importance \cite{zianbetov2013phd}.
-This last advantage is of particular use in a digital environment, as otherwise there is potential for clock degradation resulting from switching of transistors.
-The drawback of the switch to a digital architecture however is the presence of quantisation.
+The digital nature of this architecture brings with it a number of advantages over traditional analogue implementations, as it can benefit from advancements in digital circuit design suites, be reconfigurable and programmable, and has a significantly greater immunity to perturbations inherent to its digital nature, as the exact voltage of signals is of no importance \cite{zianbetov2013phd}.
+
+This last advantage is particularly useful in a digital environment, as otherwise there is potential for clock degradation resulting from switching of transistors.
+The drawback of the switch to a digital architecture, however, is the presence of quantisation.
 Analogue designs both deliver continuous control signals to the oscillators and have a continuous phase detection capability, unlike a digital system where these actions are carried out with fixed resolution.
@@
 \end{figure}
-Looking at the design of a given node it is notable that the function carried out by the Error Combiner is akin to a average, therefore as mentioned by Pratt and Nyugen, there is potential to a mode locked equilibrium in which the oscillators are not synchronised.
+Looking at the design of a given node, it is notable that the function carried out by the Error Combiner is akin to an average, therefore as mentioned by Pratt and Nyugen, there is potential for a mode locked equilibrium in which the oscillators are not synchronised.
 In their paper, they presented a method where initial start-up was performed uni-directionally and, once all nodes are close to alignment, full connectivity could be restored, however this was not viable in an analogue system as reconfigurability was not an option \cite{pratt1995distributed}.
@@
 \end{quote}
-In creating an entirely digital system, Zianbetov and Shan could easily reconfigure the network topology and by implement a uni-directional start-up avoid the problem of convergence into a mode locked state, without having to design a non-linear phase detector.
+In creating an entirely digital system, Zianbetov and Shan could easily reconfigure the network topology and, by implementing a uni-directional start-up, avoid the problem of convergence into a mode locked state, without having to design a non-linear phase detector.
 
 \section{\acs{ADPLL} Architecture}
-As indicated in Figure \ref{fig:mulkeen_pll} the three main building blocks of a conventional \ac{PLL} are the \ac{PD}, \ac{LF} and \ac{VCO}.
+As indicated in Figure \ref{fig:mulkeen_pll}, the three main building blocks of a conventional \ac{PLL} are the \ac{PD}, \ac{LF} and \ac{VCO}.
 In \iac{ADPLL} these blocks are then replaced by their digital counterparts, necessitating quantisation in order to remain physically realisable.
@@
 	\includegraphics[scale=0.5]{../tex_files/mulkeen_pll}
-	\caption[Block fiagram of \iacl{PLL}]{Block diagram of \iacl{PLL}, \textit{Wireless Systems Notes}, B. Mulkeen (2017).}
+	\caption[Block diagram of \iacl{PLL}]{Block diagram of \iacl{PLL}, \textit{Wireless Systems Notes}, B. Mulkeen (2017).}
 	\label{fig:mulkeen_pll}
@@
 The decisions made in the design of the \ac{DCO}, or \ac{NCO}, determine many of the other \ac{ADPLL} parameters.
-While tuning range and centre frequency, as well as linearity, carry over from the analogue counterpart, \iac{DCO} also has a frequency step which in combination with the bit width of the control signal determines the range over which the oscillator can be tuned.
+While tuning range and centre frequency, as well as linearity, carry over from the analogue counterpart, \iac{DCO} also has a frequency step, which in combination with the bit width of the control signal determines the range over which the oscillator can be tuned.
 Figure \ref{fig:my_ring} illustrates a basic ring oscillator design.
-A ring oscillator is an inherently unstable circuit composed of an odd number of inverters connected in a circle, which allows a signal to propagate infinitely, with the signal at any point in the circuit appearing as a square wave.
-The half-period of this oscillator is the time taken for the signal to propagate once through the chain, $n$ times the propagation delay through one inverter.
-The frequency of operation can then be set by modulating the length of the chain, in steps of two inverters to maintain an odd number, by means of \texttt{f\_select}.
-The main impact of output frequency quantisation is that only frequencies which are integer multiples of the frequency step away from the centre frequency can be easily reproduced, with intermediate values only obtainable in a manner akin to Fractional-N synthesis with the control code toggling back and forth.
+A ring oscillator is an inherently unstable circuit, composed of an odd number of inverters connected in a circle, which allows a signal to propagate infinitely, with the signal at any point in the circuit appearing as a square wave.
+The half-period of this oscillator is the time taken for the signal to propagate once through the chain, $n$ times the propagation delay through one inverter, where $n$ is the number of inverters.
+The frequency of operation can then be set by modulating the length of the chain, in steps of two inverters to maintain an odd number, by means of a control signal.
+The main impact of output frequency quantisation is that only frequencies which are at integer multiples of the frequency step away from the centre frequency can be easily reproduced.
+Intermediate values only obtainable in a manner akin to Fractional-N synthesis, with the control code toggling back and forth.
 This acts as a source of jitter in the system.
@@
 Both methods use the most significant bit of the counter's value to form the output signal.
-Period linearity is achieved by varying the reload value of the counter after overflow depending on the control code, thereby changing the period by a multiple of a fixed step.
-Alternatively frequency linearity can be achieved if the reload value is left constant, but instead the amount added to the counter every clock cycle is changed according to the control code, once again a fixed step size is used.
+Period linearity is achieved by varying the reload value of the counter after overflow, depending on the control code, thereby changing the period by a multiple of a fixed step.
+Alternatively, frequency linearity can be achieved if the reload value is left constant, but instead the amount added to the counter every clock cycle is changed according to the control code, once again a fixed step size is used.
 
 \subsection{Digital Phase Detector}
-Once again quantisation impacts the \acl{PD}, as rather than a continuous output the phase detector in \iac{ADPLL} has a finite number of output values, thus limiting the accuracy of the phase detector.
-A second form of quantisation is also present, as unlike an analog system, a digital phase detector does not provide continuous data in the time domain either, instead relying on sampling.
-At its most basic, a digital phase comparator may only output an indication of which signal is leading, a design known as a Bang-Bang Detector, which can be constructed using a single D Flip Flop with one the generated signal connected to the ``D'' input and the reference signal acting as the clock.
-As the output only has two levels the resultant word is only 1 bit wide and as such, limits the range over which the output frequency can be controlled.
-More complex designs such as that in Figure \ref{fig:shan_bb_pd}, implemented by Shan, build on this by measuring the time difference between edges of the signals using \iac{TDC} in his case using \iac{TDL} \cite{shan2014phd}.
+Once again quantisation impacts the \acl{PD}, as rather than a continuous output, the phase detector in \iac{ADPLL} has a finite number of output values, thus limiting the accuracy of the phase detector.
+A second form of quantisation is also present, as unlike an analog system, a digital phase detector does not provide continuous data in the time domain, instead relying on sampling.
+At its most basic, a digital phase comparator may only output an indication of which signal is leading, a design known as a Bang-Bang Detector.
+This can be constructed using a single D Flip Flop, with the generated signal connected to the ``D'' input and the reference signal acting as the clock.
+As the output only has two levels, the resultant word is only 1 bit wide, limiting the range over which the output frequency can be controlled.
+More complex designs, such as that in Figure \ref{fig:shan_bb_pd} implemented by Shan, build on this by measuring the time difference between edges of the signals using \iac{TDC}, in his case using \iac{TDL} \cite{shan2014phd}.
 \Iac{TDL} is constructed by a chain of elements of a fixed delay, and the signal to be timed is applied to this start of this chain.
-After the timing interval elapsed the values at each point in the chain are examined, and using temperature coding, these are converted to a digital signal, the width of which is the bit width of the \ac{PD}'s error signal.
+After the timing interval elapses the values at each point in the chain are examined, and using temperature coding, these are converted to a digital signal.
+The width of this signal is the bit width of the \ac{PD}'s error signal.
 This mimics a time measurement and allows for the phase difference to be recorded in a non binary manner.
@@
 	\centering
-	\includegraphics[width=0.6\textwidth]{../shan_bb_pd}
+	\includegraphics[width=0.65\textwidth]{../shan_bb_pd}
 	\caption[Bang-bang phase/frequency detector architecture]{Bang-bang phase/frequency detector architecture \cite{shan2014phd}.}
@@
 The \acl{LF} in \iac{ADPLL} can be implemented as \iac{PI} controller, as only a low-pass filter is required, such as that in Figure \ref{fig:my_simple_pi}.
-In the case of a node in \iac{ADPLL} network the input of this filter is a weighted of the phase difference relative to the neighbouring local clocking areas.
-In one example topology a digital system the proportional section can be implemented by a simple multiplier, whereas the integral path is constructed by adding the result of a multiplication by the proportional gain to an accumulator.
-This delayed summation can be easily implemented by an accumulator to which the current value of the multiplication is added each cycle, and as such the system has an infinite impulse response.
+In the case of a node in \iac{ADPLL} network the input of this filter is a weighted average of the phase difference relative to the neighbouring local clocking areas.
+In a digital system, the proportional section can be implemented by a simple multiplier, whereas the integral path is constructed by the delayed addition of error multiplied by the integral gain to the sum of past values.
+This delayed accumulation can be easily implemented by an accumulator to which the current value of the multiplication is added each cycle, and as a result the system has an infinite impulse response.
 The value of these gains determine the response and stability of the \ac{ADPLL} network.
@@
 \end{equation*}
-It has been found by Koskin \textit{et al} (2018) that stable operation can be achieved when the integral gain, $k_i$, is less than half the proportional gain, $k_p$ \cite{koskin2018generation}.
+It has been found by Koskin \textit{et al} (2018), that stable operation can be achieved when the integral gain, $k_i$, is less than half the proportional gain, $k_p$ \cite{koskin2018generation}.
 In the same study a range of values was found which would produce low jitter operation of the network.
-As these values are all less than one, the filter must implement fixed point arithmetic in an effort to maintain the simplicity of the clock distribution network, rather than incurring the penalty of floating point calculations.
+As these values are all less than one, the filter must implement fixed point arithmetic in an effort to maintain the simplicity of the \ac{LF}, rather than incurring the penalty of floating point calculations.
 
 \subsection{Error Combiner}
-The \acp{ADPLL} used in a network need to combine the error signals from multiple neighbours to determine what the average difference from its neighbours is, and this necessitates the addition of the Error Combiner.
-In a digital system this can be implemented by a weighted average of the different error signals, with the weight being modifiable at run-time.
+The \acp{ADPLL} used in a network need to combine the error signals from multiple neighbours to determine what the average difference from its neighbours is, and this necessitates the addition of an Error Combiner.
+In a digital system, this can be implemented by a weighted average of the different error signals, with the weight being modifiable at run-time.
 This configurability is what permits the system to implement uni-directional mode and also allows for the weighting applied to certain signals, such as the external reference, to be modified.
-The ease of implementation of a configurable error combiner is one of the main advantages of \iac{ADPLL} over an analogue system.
+The ease of implementation of a configurable error combiner is one of the key advantages of \iac{ADPLL} over an analogue system.
 
@@
 \Iac{FPGA} contains a large number of logic elements that can be connected together in order to perform complex logic, written using the same \aclp{HDL} (\acsp{HDL}) used to design the digital blocks of \aclp{ASIC} (\acsp{ASIC}).
-They may also implement started modules such as adders, multiplexers and \ac{RAM} as a fundamental element.
+They may also implement standard modules such as adders, multiplexers and \ac{RAM} as a fundamental element.
 More complex logic is often implemented using multiplexed lookup tables rather than true logic elements.
-High end devices such as the Xilinx Zynq Ultrascale even implement Multi-Processor \ac{SOC}s.
+High end devices such as the Xilinx Zynq Ultrascale, may implement Multi-Processor \ac{SOC}s.
 Compared to \iac{ASIC} the designer does not have direct control over the layout of the system but rather describes its behaviour, possibly down to the basic logic elements of inverters or other gates.
@@
 
-Prototyping on \iac{FPGA} is a common verification stage for conventional \ac{ASIC} designs as it allows for a hardware validation of any digital circuitry, and the detection of any potential flaws or errors made by the design before the expensive of \iac{ASIC} implementation.
-In their 2013 \ac{ADPLL} network implementation Zianbetov and Shan used \iac{FPGA} in order to validate their programming interface, the design of the error processing block, ensure they had eliminated mode locking behaviour and to ensure phase synchronisation was possible \cite{zianbetov2013phd,shan2014phd}.
-However they experienced two main limitations, they were not able to implement the mixed-signal \ac{PFD} and \ac{DCO}, and the maximum frequency of operation possible was orders of magnitude lower than the GHz range of their \ac{ASIC} implementation.
-These issues were circumvented by implementing an alternative \ac{PFD} and \ac{DCO} designs which were driven by the clock distribution network provided by the \ac{FPGA} and every clock frequency scaled by the same amount such that the results of testing would remain indicative.
+Prototyping on \iac{FPGA} is a common verification stage for conventional \ac{ASIC} designs, as it allows for a hardware validation of any digital circuitry, and the detection of any potential flaws or errors made by the design before the expense of \iac{ASIC} implementation.
+In their 2013 \ac{ADPLL} network implementation, Zianbetov and Shan used \iac{FPGA} in order to validate their programming interface, the design of the error processing block, ensure they had eliminated mode locking behaviour and to ensure phase synchronisation was possible \cite{zianbetov2013phd,shan2014phd}.
+However they experienced two main limitations: they were not able to implement the mixed-signal \ac{PFD} and \ac{DCO}.
+Secondly, the maximum frequency of operation possible was orders of magnitude lower than the GHz range of their \ac{ASIC} implementation.
+These issues were circumvented by implementing an alternative \ac{PFD} and \ac{DCO} designs which were driven by the clock distribution network provided by the \ac{FPGA}, and frequency scaling each clock by the same amount, such that the results of testing would remain indicative.
 One of the main advantages they saw was that the hardware description used for the digital blocks of their \ac{ASIC} could be directly ported over to the \ac{FPGA}.
@@
 
-It is however possible to implement limited mixed-signal circuits on \iac{FPGA} through the use of primitive logic elements, however, as control over the implementation is restricted to the module level it is not possible to mirror the implementation of a design intended for \iac{ASIC}.
-As these implementations are analogous to a mixed-signal circuit on \iac{ASIC}, the verification of theoretical behaviours, as done by Koskin \textit{et all}, in order to test the findings of his PhD thesis \cite{theboys2019}, is made possible without the expense of \ac{ASIC} fabrication.
+It is possible to implement limited mixed-signal circuits on \iac{FPGA} through the use of primitive logic elements, however, as control over the implementation is restricted to the module level, it is not possible to mirror the implementation of a design intended for \iac{ASIC}.
+As these implementations are analogous to a mixed-signal circuit on \iac{ASIC}, the verification of theoretical behaviours, as done by Koskin \textit{et al}, in order to test the findings of his PhD thesis \cite{theboys2019}, is made possible without the expense of \ac{ASIC} fabrication.
 \Iac{FPGA} based mixed-signal \ac{ADPLL} network is also seeing use here in \ac{UCD} as a initial prototyping platform for the validation of new modules for use in \ac{ASIC} based \ac{ADPLL} networks.
@@
 
-Mixed-signal circuits were, however, simulated in hardware by \'{O}scar Luc\'{i}a \textit{et al} (2011) on \iac{FPGA} in order to overcome the excessive time penalty imposed by software based simulations that required the behaviour of both a digital and mixed-signal peripheral and that of a micro controller running code in C to be simulated side by side \cite{lucia2011real}.
+Mixed-signal circuits were, however, simulated in hardware by \'{O}scar Luc\'{i}a \textit{et al} (2011) on \iac{FPGA}, in order to overcome the excessive time penalty imposed by software based simulations that required the behaviour of both a digital and mixed-signal peripheral and that of a microcontroller running code in C to be simulated side by side \cite{lucia2011real}.
 They concluded
@@
 \section{\acs{ADPLL} Performance Characterisation}
-As already stated, goal of a clock distribution system is to synchronise clocking events in all areas of the chip.
+As already stated, the goal of a clock distribution system is to synchronise clocking events in all areas of the chip.
 The effectiveness of this synchronisation is characterised by two main metrics, jitter and skew, which together describe the distribution of clocking events.
@@
 It can be easily measured by computing the average value of this delay.
-Jitter has a number of definitions depending on how it is measured, but at its most basic it is the standard deviation of the time delays with respect to the reference clocking edge.
+Jitter has a number of definitions depending on how it is measured, but at its most basic, it is the standard deviation of the time delays with respect to the reference clocking edge.
 
-The simplest form of clock performance characterisation is on \iac{C2C} basis, in which the reference is the signal under test itself.
-Here the standard deviation of the individual periods gives the jitter of the signal, while skew has no meaning with a self reference the mean period can be used to compute the centre frequency over the time interval.
+The simplest form of clock performance characterisation is on \iac{C2C} basis, in which the reference is the previous edge of the signal under test itself.
+Here the standard deviation of the individual delays gives the jitter of the signal.
+As the previous edge of the generated signal is used as the reference, and thereby the delay is the period of the signal, the mean value is no longer the skew, but instead can be used to compute the centre frequency over the time interval.
 For a more informative measurement the \ac{TIE} can be computed, which takes into account another signal as the reference.
 \ac{TIE} is calculated by comparing the delay between the signal in question and a reference that is treated as ideal.
-The mean value gives the relative skew between the two signals and once more the standard deviation of the measurements gives the jitter with respect to this reference.
+The mean value gives the relative skew between the two signals and, once more, the standard deviation of the measurements gives the jitter with respect to this reference.
 
 Both above forms of measurement assume a large number of sequential measurements, however, there are other ways that jitter and skew can be calculated \cite{AN10007}.
-Phase jitter is an important characteristic for communications systems as it can be used to calculate phase noise, important to ensure spurious emissions are within regulations.
+Phase jitter is an important characteristic for communications systems as it can be used to calculate phase noise, important to ensure spurious emissions are within legal requirements.
 Long term, or accumulated, jitter represents the cumulative effect of jitter on the signal over several cycles.

### 2019-04-23 thesis/FPGABasedADPLLs.tex
partial corrections

@@
 The first step in creating \iac{FPGA} based network of \acp{ADPLL} is the design of the \ac{ADPLL} itself, which will be addressed in this chapter.
-The nature of \iac{FPGA} necessitates a number of compromises in the design of a given block which limits transferability to \ac{ASIC} designs.
+The nature of \iac{FPGA} necessitates compromises in the design of a certain blocks, limiting direct transfer from \ac{ASIC} designs.
 In this chapter the potential designs for each individual block, or module, investigated will be explained and the case for their selection in \iac{FPGA} based \ac{ADPLL} made.
-A number of blocks have implement purely digital circuitry and as such can be transferred in their entirety from \iac{FPGA} and vice versa.
+A number of blocks have implemented by purely digital circuitry and as such can be transferred in their entirety from \iac{FPGA} and vice versa.
 However, those that will be used to emulate mixed-signal circuitry, such as the \ac{DCO} and \ac{PFD}, will be examined in greater detail.
-\Iac{RTL} diagram of the \ac{ADPLL} topology that will be assumed in the following sections is shown in Figure \ref{fig:network_adpll}.
+\Iac{RTL} diagram of the \ac{ADPLL} topology that will be used in the following sections is shown in Figure \ref{fig:network_adpll}.
 \begin{figure}[h]
@@
 
-Important to note is that despite a Cartesian grid leading to four neighbours, only two phase detectors are required per oscillator.
-Borrowing matrix indexing, and choosing the oscillator located at index $i,j$, $i,j \neq 1,N$ where $N$ is the number of oscillators in each direction, comparison need only be done with the signals generated by oscillators located at $i-1,j$ and $i,j-1$.
-The comparison with oscillators $i+1,j$ and $i,j+1$ will be carried out in those \acp{ADPLL}, and the negation of the errors measured used in the \ac{ADPLL} located at $i,j$.
+It is Important to note that despite a Cartesian grid leading to four neighbours, only two phase detectors are required per oscillator.
+Borrowing matrix indexing, and choosing the oscillator located at index $(i,j)$, $i,j \neq 1,N$ where $N$ is the number of oscillators in each direction, comparison need only be done with the signals generated by oscillators located at $(i-1,j)$ and $(i,j-1)$.
+The comparison with oscillators $(i+1,j)$ and $(i,j+1$) will be carried out in those \acp{ADPLL}, and the negation of the errors measured, used in the \ac{ADPLL} located at $(i,j)$.
 
 \section{Digitally Controlled Oscillators}
-The choices made in the design of the \ac{DCO} have the greatest impact on the effectiveness of the overall platform and which use cases the \ac{ADPLL} containing it are suitable for, as the key performance benchmarks are all done using the waveform this block generates.
-This project will address three distinct designs of \acp{ADPLL} suitable for implementation on an FPGA, two derived from the clocks generated by the \acp{FPGA} own distribution network and one generated independently of this clock, using a chain of inverters.
-These are not the only ways in which an oscillator could be synthesised on \iac{FPGA}, however other designs were deemed to be unsuitable for extensible and portable implementations.
+The choices made in the design of the \ac{DCO} have the greatest impact on the effectiveness of the overall platform and which use cases the \ac{ADPLL} is suitable for, as all performance benchmarks are done using the waveform this block generates.
+This project will address three distinct designs of \acp{DCO} suitable for implementation on an FPGA, two derived from the clocks generated by the \acp{FPGA} own distribution network and one generated independently of this clock, using a chain of inverters.
+These are not the only ways in which an oscillator could be synthesised on \iac{FPGA}, however, other designs were deemed to be unsuitable for extensible and portable implementations.
+%TODO delete if space needed
+A prime example of this, is the use of Xilinx proprietary \texttt{IODELAY} blocks to create an oscillator, as detailed in Xilinx Application Note XAPP872 \cite{iodelay}.
+The key idea here is that the bulk of the period is made up by the propagation time through one of the \texttt{IODELAY} blocks, which can be set at implementation time.
+This is combined with a section of an inverter chain, and a multiplexer is used modify the length of this segment, the output of which is fed back into the \texttt{IODELAY} block.
+This method was discarded as the number of \texttt{IODELAY} blocks is very limited, so expanding to a larger network would be impossible, and as they are all located around the edge of the chip, not suited to the construction of a Cartesian grid.
+%TODO delete if space needed
 
-A prime example of this is the use of Xilinx proprietary \texttt{IODELAY} blocks to create an oscillator, as detailed in Xilinx Application Note XAPP872 \cite{iodelay}.
-The key idea here is that the bulk of the period is made up by the propagation time through one of the \texttt{IODELAY} blocks, which can be set at implementation time.
-This is combined with a section of an inverter chain, and a multiplexer used modify the length of this segment, the output of which is fed back into the \texttt{IODELAY} block.
-This method was discarded as the number of \texttt{IODELAY} blocks is very limited, so expanding to a larger network would be impossible, and they are all located around the edge of the chip, not suited to the construction of a Cartesian grid.
-
-The main issue with the creation of \acp{DCO} on \iac{FPGA} is the inability to create mixed-signal circuits, such as those that would be intended for use on \iac{ASIC}.
-As such the \ac{FPGA} based oscillator must emulate the behaviour of a mixed-signal circuit in some way.
+The main issue with the creation of \acp{DCO} on \iac{FPGA}, is the inability to create mixed-signal circuits, such as those used on \iac{ASIC}.
+Therefore the \ac{FPGA} based oscillator must emulate the behaviour of a mixed-signal circuit in some way.
 
@@
 \end{figure}
-The first design of \ac{DCO} to be examined is of the type used by Zianbetov in his \ac{ADPLL} network test bed and relies on a counter driven by the clock manager on the \ac{FPGA} \cite{zianbetov2013phd}.
-At each event on the \ac{FPGA} provided clock a counter is incremented, overflowing upon increment past the maximum possible value.
-The \ac{MSB} of this counter forms the waveform generated by this oscillator, the period of which is controlled through an adjustable value that is loaded into the counter once overflow is reached, and forms the starting points for the counter.
+The first design of \ac{DCO} to be examined, is of the type used by Zianbetov in his \ac{ADPLL} network test bed and relies on a counter driven by the clock manager on the \ac{FPGA} \cite{zianbetov2013phd}.
+At each event on the \ac{FPGA} provided clock, a counter is incremented, overflowing upon increment past the maximum possible value.
+The \ac{MSB} of this counter forms the waveform generated by this oscillator, the period of which is controlled through an adjustable value that is loaded into the counter once overflow is reached, and forms the starting point for the counter.
 The period of oscillation is given by:
@@
 where $T_{FPGA}$ is the clock period of the \ac{FPGA}, $CC$ is the control code input, $BIAS$ centres the oscillator in the middle of the tuning range in the event that the control code is zero and $width$ is the width of the counter used.
-As the only variable here is the control code, period step of this design is $T_{FPGA}$, thereby providing period linearity with respect to the control code.
+As the only variable here is the control code, the period step of this design is $T_{FPGA}$, thereby providing period linearity with respect to the control code.
 This is the key advantage of this design, as most \ac{ASIC} implementations of \iac{DCO} are also linear in period.
-The other main reason to choose this design is that its \ac{FPGA} clocked nature allows for exact control over the frequency of operation, and the number of tunable parameters make it possible to configure multiple ways to achieve the same frequencies of operation.
-Combined these attributes make it very easy to create an oscillator that emulates the behaviour of a design intended for \iac{ASIC}, however, at a greatly reduced frequency.
-This restriction on the frequency of operation arises out of the period step size, which in order to obtain a good resolution must be orders of magnitude smaller than the intended period to be generated.
+The other main reason to choose this design, is that its \ac{FPGA} clocked nature allows for exact control over the frequency of operation, and the number of tunable parameters make it possible to configure multiple ways to achieve the same frequencies of operation.
+Combined, these attributes make it very easy to create an oscillator that emulates the behaviour of a design intended for \iac{ASIC}, however, at a greatly reduced frequency.
+This restriction on the frequency of operation arises out of the period step size, which in order to obtain a good resolution must be orders of magnitude smaller than the intended period.
 As the output waveform is taken from the counter's \ac{MSB}, the reload value of the counter must never go beyond $2^{width-1}$, as otherwise the output waveform will become a constant 1.
-As the reload value varies the low time of the \ac{MSB}, if the desired output waveform is a square wave this design will not be suitable.
+As the reload value varies the low time of the \ac{MSB}, if the desired output waveform is a square wave, this design will not be suitable.
 
-In being \ac{FPGA} clocked this design has pseudo-deterministic characteristics, with each period step being almost identical across oscillators and over the entire tuning range, unlike \iac{ASIC} where process variation will impact the layout of a high frequency oscillator.
-The only variation in this design will come, ironically, from jitter or skew in the \ac{FPGA}'s clock distribution network, which as the frequencies will typically be in the low hundreds of MHz is very minor.
+Being \ac{FPGA} clocked this design has pseudo-deterministic characteristics, with each period step being almost identical across oscillators and over the entire tuning range, unlike \iac{ASIC} where process variation will impact the layout of a high frequency oscillator.
+The only variation in this design will come, ironically, from jitter or skew in the \ac{FPGA}'s clock distribution network, which as the frequencies will be in the hundreds of MHz is very minor.
 In the case of the Xilinx \acl*{Nexys} this is at most 100 picoseconds, or 0.05\% of the period of an intended output clock at 5 MHz.
-To put this value into perspective, on this board the minimum value of $T_{FPGA}$ that could be used to drive this oscillator is 3.87 nanoseconds, 1.935\% of the period.
+To put this value into perspective, on this board the minimum value of $T_{FPGA}$ that could be used to drive this oscillator is $3.875~\si{\nano\second}$, $1.935\%$ of the period.
 
@@
 The second \ac{FPGA} clocked oscillator is similar in most attributes to the above design but eschews period linearity for frequency linearity.
-Again the overflow property of a counter is used with the counter's \ac{MSB} as output of the block, however, this time it forms a square wave.
+Again the overflow property of a counter is used with the counter's \ac{MSB} as the block's output, however, this time it forms a square wave.
 Rather than setting the reload value of the counter, instead the increment is adjusted depending on the control code, thus requiring $\frac{2^N}{BIAS+CC}$ increments to overflow.
@@
 \end{equation}
-Here the control code $CC$ and bias are added to the value stored in the accumulator at each event of the \ac{FPGA}, clock until overflow is reached.
+Here the control code $CC$ and $BIAS$ are added to the value stored in the accumulator at each event of the \ac{FPGA} clock, until overflow is reached.
 This occurs at $2^{width}$ where, as before, $width$ is the bit width of the counter, thus valuing each control code increment at $\frac{f_{FPGA}}{2^{width}}$ Hz.
@@
 This design is just as configurable as its linear-in-period counterpart, and well suited to the low frequency emulation of \ac{ASIC} based oscillators that are themselves linear in frequency.
-In sharing the \ac{FPGA} as a clock source again the pseudo-deterministic characteristics return, once more meaning this oscillator is better used for testing, simulating or verifying other blocks in the system.
+In sharing the \ac{FPGA} as a clock source, the pseudo-deterministic characteristics return, once more meaning this oscillator is better used for testing, simulating or verifying other blocks in the system.
 
 \subsection{Ring Oscillator}
-The best approximation of the non-idealities of \iac{ASIC} \ac{DCO} implementation can be obtained using \iac{RO}, also know as an inverter ring, as the \ac{DCO}.
-In \iac{ASIC} \iac{RO} is constructed by connecting a number of NOT gates back in a chain, with the output of the last NOT gate fed input the input of the first.
-As NOT gates are not available on \iac{FPGA}, they are replaced by inverter primitives.
-The inverter ring is an inherently unstable circuit that leverages the propagation delay through an odd number of inverters to create a waveform that, when viewed at a fixed position, is a square wave.
+The best approximation of the non-idealities of \iac{ASIC} \ac{DCO} implementation can be obtained using \iac{RO}, also known as an inverter ring, as the \ac{DCO}.
+In \iac{ASIC} \iac{RO} is constructed by connecting a number of inverters back in a chain, with the output of the last NOT gate fed into the input of the first.
+As true inverters are not available on \iac{FPGA}, they are replaced by inverter primitives, constructed from look-up tables.
+The inverter ring is an inherently unstable circuit, that leverages the propagation delay through an odd number of inverters to create a waveform that, when viewed at a fixed position, is a square wave.
 \begin{figure}[h]
@@
 The main advantage of this type of oscillator stems from its similarity to that used on \iac{ASIC} which leads to many non-idealities carrying over from the truly mixed-signal implementation.
-Rather than being just a tool to validate that other modules, the performance of this oscillator can be analysed more deeply as well as the performance of networks containing it.
-Koskin \textit{et al} used such a design to test theories proposed regarding the synchronisation of oscillators during his PhD using this type of oscillator implemented on \iac{FPGA} \cite{koskin2019phd,theboys2019}.
-Also in this design's favour is the period step size, four times the inverter propagation delay, which was found experimentally to be than in the region of a nanosecond on Xilinx Artix-7 and Zynq \acp{FPGA}.
+Rather than being just a tool used to validate other modules, the performance of this oscillator can be analysed more deeply as well as the performance of networks containing it.
+Koskin \textit{et al} tested theories proposed regarding the synchronisation of oscillators during his PhD using this type of oscillator implemented on \iac{FPGA} \cite{koskin2019phd,theboys2019}.
+Also in this design's favour is the period step size, four times the inverter propagation delay, which was found experimentally to be within in the region of a nanosecond on Xilinx Artix-7 and Zynq \acp{FPGA}.
 
-While it is advantageous for the purposes of analysis that this design is affected by variation in temperature, implementation and other non-idealities this brings with it challenges not present in a true mixed-signal design.
-In a design intended for \iac{ASIC} the design has near total control over the placement of circuit elements, however, the \ac{FPGA} based counterpart control is available over just the area in which the oscillator may lie, and the \ac{EDA} is responsible for the exact placement.
+While it is advantageous for the purposes of analysis that this design is affected by variation in temperature, implementation and other non-idealities, this brings with it challenges not present in a true mixed-signal design.
+In a design intended for \iac{ASIC}, the designer has near total control over the placement of circuit elements, however, the \ac{FPGA} based counterpart control the designer may only set the region within the oscillator will lie, and the \acs{EDA} tool is responsible for the exact placement.
 Unlike the \ac{ASIC} implementation, in which the propagation delay through each element of the ring is well bounded, the lack of control over placement means that the propagation time through each inverter is not well bounded and may vary wildly.
-This makes it time consuming to ensure that the frequency tuning ranges of each oscillator match that desired, and use of this oscillator less straight-forward than that of the previous two designs.
-This is compounded by the simulation tools in \ac{EDA} packages intended for \acp{FPGA} being unable to accurately simulate the time delays, thus negating any benefit of test benches for the alignment process.
+This makes it time consuming to ensure that the frequency tuning ranges of each oscillator match that desired, and making use of this oscillator less straight-forward than that of the previous two designs.
+This is compounded by the simulation tools in the \ac{EDA} packages intended for \acp{FPGA} being unable to accurately simulate the time delays, thus negating any benefit of test benches for the alignment process.
 
 \section{Frequency Divider}
-Many \aclp{PLL} (\acsp{PLL}), be they digital or analogue, incorporate a frequency divider in the feedback path so that comparison can be made with a reference operating at a lower frequency.
-This is a feature that is advantageous for \ac{ADPLL} networks, as comparison using this divided clock allows the synchronisation signal sent between modules to be of a frequency orders of magnitude lower than that of the clock distribution network and thus require circuitry that consumes significantly less energy.
-This division can be obtained trivially on \iac{FPGA} by a number of means, of which this thesis will mention just two.
+Many \aclp{PLL} (\acsp{PLL}), be they digital or analogue, incorporate a frequency divider in the feedback path, so that comparisons can be made with a reference operating at a lower frequency.
+This is a feature that is advantageous for \ac{ADPLL} networks, as comparisons using this divided clock allows the synchronisation signal sent between modules to be of a frequency orders of magnitude lower than that of the clock distribution network, and thus requiring circuitry that consumes significantly less energy.
+This division can be obtained trivially on \iac{FPGA} by a number of means; of which this thesis will mention just two.
 The first of these methods allows for integer division, by inverting a signal each time a counter reaches a certain value, $k$.
-When this condition is true the value stored in the counter is also reset to zero, thus implementing division of the generated clock frequency by $2k$.
+When this condition is true the value stored in the counter is also reset to zero, thus dividing the generated clock frequency by $2k$.
 \begin{figure}[h]%
@@
 	\subfloat[Divider 2]{{\includegraphics[width=0.6\textwidth]{../divider2} }}%
-	\caption[Frequency Divider \ac{RTL} Diagrams]{Frequency Divider \ac{RTL} Diagrams.}
+	\caption[Frequency divider \ac{RTL} diagrams]{Frequency divider \ac{RTL} diagrams.}
 	\label{fig:divs}
 \end{figure}
-By limiting to division by powers of two, the divider can be simplified with the removal of any comparison logic.
+
+By limiting division to powers of two, the second option ca simplify the divider by the removal of any comparison logic.
 One flip-flop and an inverter are required per power of two.
@@
 Three potential \acl{PD} designs will be addressed in this thesis, sign-only detection with a Bang-Bang Detector, \ac{FPGA} clocked detection using \iac{FSM} and an emulation of \iac{TDL} using inverter primitives to create a delay, similarly to the \ac{RO} seen previously.
-Both designs will have to have a representation for both situations where the generated signal leads and lags the reference signal, and for the magnitude of this difference, a problem most obviously solved by using a signed binary representation of the difference.
-Most easily implemented in hardware is two's complement, as operations can be carried out by the same hardware as unsigned arithmetic.
+All designs will have to have a representation for both situations where the generated signal leads and lags the reference signal, and for the magnitude of this difference.
+Both can be combined using a signed binary representation of the difference, such as two's complement.
+Two's complement is easily implemented, as operations can be carried out by the same hardware as unsigned arithmetic.
 The sign convention for phase error assigns the positive values to the situation where the generated signal lags the reference, and vice versa for the leading case.
-Apart from the Bang-Bang Detector, each design features two blocks, one to determine the sign of the error and a second which estimates the magnitude.
-These blocks could theoretically be mixed-and-matched, but as they operate on different paradigms no benefit would be obtained.
+Apart from the Bang-Bang Detector, each design features two blocks, one to determine the sign of the error and a second to estimate the magnitude.
+These blocks could theoretically be mixed-and-matched, but as they operate on different paradigms, no benefit would be obtained.
 
 \subsection{Bang-Bang Detector}
-A Bang-Bang Detector is a sign-only detector, thus outputting a phase error just a single bit wide.
+A Bang-Bang Detector is a sign-only detector, thus outputting a phase error a single bit wide.
 This can be accomplished using a single flip-flop, with the reference signal acting as the clock, in this case using the rising edge, a design verified by Predraig in his 2011 thesis \cite{predraig}.
-The generated signal is applied to the $D$ input, thus when a rising edge is seen on the reference the value of the generated signal is held as the $Q$ output until the next rising edge of the reference.
+The generated signal is applied to the $D$ input, thus when a rising edge is seen on the reference, the value of the generated signal is held as the $Q$ output until the next rising edge of the reference.
 If the generated signal has gone high prior to the reference, and leading it, the value of $Q$ will be 1, thereby matching the expected sign behaviour.
-Trivially it can be seen that the case where the generated signal lags the reference the $Q$ will show 0.
+Trivially it can be seen that the case where the generated signal lags the reference the $Q$ will show $0$.
 \begin{figure}[h]
@@
 
-This design is simple but effective, and can even be used without a loop filter, albeit with a very limited oscillator tuning range.
-This is especially valuable an attribute at an early stage of platform development, as it may be used in order to test that other aspects of the system work as intended.
-The binary resolution of this detector, however, restricts its usefulness as the \ac{ADPLL} cannot react more quickly to larger phase differences.
+This design is simple but effective, and can be used without a loop filter, albeit with a very limited oscillator tuning range.
+This is an especially valuable attribute at early stages of platform development, as it may be used in order to test that other aspects of the system work as intended.
+The binary resolution of this detector, however, restricts its usefulness ,as the \ac{ADPLL} cannot react more quickly to larger phase differences.
 
 \subsection{\acs{FPGA} Clocked Phase Detector}
-The easier to implement of the two designs uses a combination of a state machine to determine which edge occurred first, and to control a two's complement up-down counter, which acts as the \ac{TDC} in this detector, and estimates the magnitude of the error.
+The easier to implement of the two designs, uses a combination of a state machine to determine which edge occurred first, and to control a two's complement up-down counter, which acts as the \ac{TDC} in this detector, and estimates the magnitude of the error.
 
@@
 \Iacl{FSM} consists of three main components: the states themselves, the next state logic and the output logic.
-In this case the next state logic is the generated and reference signals, and the output logic sets the count instructions for the up-down counter as well as ensuring the phase error output is held constant between measurement periods.
+In this case the next state logic is the generated and reference signals, and the output logic sets the count instructions for the up-down counter, as well as ensuring the phase error output is held constant between measurement periods.
 The \ac{FSM} will have a state representing the case where either the reference or generated signals occurred first, in addition to a wait state used between the end of one measurement interval and the commencement of the next.
-Two types of \ac{FSM} exist, Moore and Mealy Machines, the difference being that Moore Machines base their output only on the current state while Mealy Machines do so on both the current state and the inputs to the system.
+Two types of \ac{FSM} exist, Moore and Mealy Machines, the difference being that Moore Machines base their output only on the current state, while Mealy Machines do so on both the current state and the inputs to the system.
 If a Moore Machines is used an extra state is required in order to hold the value of the phase error at the end of the measurement interval.
@@
 There are three situations that may occur while the \ac{FSM} is in the waiting state.
-Firstly the case where the reference signal sees a rising edge before the generated clock, and thus the generated clock is said to be lagging the reference.
-As previously mentioned the convention dictates that the phase difference have a positive sign, so the counter is incremented to count in the positive direction.
-Similarly if the generated signal is leading the reference the sign of the error is negative and thus the counter is instructed to count downwards.
+Firstly, the case where the reference signal sees a rising edge before the generated clock, and thus the generated clock is said to be lagging the reference.
+As previously mentioned, the convention dictates that the phase difference have a positive sign, so the counter is incremented to count in the positive direction.
+Secondly, if the generated signal is leading the reference, the sign of the error is negative and thus the counter is instructed to count downwards.
 The third potential scenario occurs when both clock signals have a rising edge within the same period of the \ac{FPGA} clock, which due to quantisation will be interpreted as both signals occurring at exactly the same time.
 In this case the counter receives no instruction and a zero value is stored until the next measurement interval.
-Once the \ac{FSM} is in a measurement state, and the counter either incrementing or decrementing, a rising edge occurring on the opposite signal to that which initiated the measurement will end the interval.
+Once the \ac{FSM} is in a measurement state, and the counter either incrementing or decrementing, a rising edge occurring on the opposite signal to that which initiated the measurement, will end the interval.
 \begin{figure}[h]
@@
 \subsubsection{Up-Down Counter}
-The up-down counter in this design performs estimates the magnitude of the phase difference between the reference and generated signals.
-The state machine provides control instructions, indicating whether the count should be in the increasing or decreasing directions, and and holds the counter in reset once the measurement interval has been completed.
+The up-down counter in this design estimates the magnitude of the phase difference between the reference and generated signals.
+The state machine provides control instructions, indicating whether the count should be in the increasing or decreasing directions, and holds the counter in reset once the measurement interval has been completed.
 The counter counts in two's complement, so counting in the downward direction can be carried out by adding on the representation of -1, the binary signal containing all 1s.
@@
 Two's complement is an advantageous way to represent signed numbers in hardware as the mathematical operations are identical to their unsigned counterparts, thus allowing use of any adder or carry hardware that may be present on the \ac{FPGA}.
-An important attribute of this counter is the ability to freeze the count at the minimum and maximum count values in order to avoid a situation where the counter overflows, thus at high phase offsets potentially alternating between positive and negative differences which may restrict the locking range of the \ac{ADPLL}.
-Disallowing overflow also aligns the behaviour of this phase detector with that of those implementing delay lines, allowing better interchangeability and comparison, which for a testing platform may be valuable.
+An important attribute of this counter is the ability to freeze the count at the minimum and maximum count values, in order to avoid a situation where the counter overflows.
+At high phase offsets this may cause alternation between positive and negative differences, restricting the locking range of the \ac{ADPLL}.
+Disallowing overflow also aligns the behaviour of this phase detector with that of those implementing delay lines, allowing better interchangeability and comparison, which for a testing platform is valuable.
 
@@
 The SigNum detector is based on the methodology used in the Bang-Bang \acl{PD} seen previously, but rather than just performing sign detection, the magnitude of the phase difference can also be measured.
-In the case of the \ac{FPGA} clocked design the counter handled the representation of both sign and magnitude, but as the name SigNum (Sign Number) suggested both are handled separately in this design.
+In the case of the \ac{FPGA} clocked design, the counter handled the representation of both sign and magnitude, but as the name SigNum (Sign Number) suggests, both are handled separately in this design.
 The modified Bang-Bang detector establishes which signal occurred first and generates a control signal that is analogous to the time delay between a rising edge on each input signal.
-\Iac{TDL} then uses this control signal to convert the phase difference between the signals to a digital signal.
-This magnitude is combined with the sign detected by the bang-Bang detector to form a two's complement error signal.
+\Iacf{TDL} then uses this control signal to convert the phase difference between the signals to a digital signal.
+The magnitude is combined with the sign detected by the bang-Bang detector to form a two's complement error signal.
 
-This method of phase detection is advantageous as similarly to the \acl{RO}, the \ac{TDL} can be implemented using inverters, and therefore is not synchronised to the \ac{FPGA} clock and the propagation time through inverters sets the resolution, thus allowing a smaller step than that of the clocked designs.
+This method of phase detection is advantageous, as similarly to the \acl{RO}, the \ac{TDL} can be implemented using inverters, and therefore is not synchronised to the \ac{FPGA} clock.
+The propagation time through inverters once more sets the resolution, thus allowing a smaller step than that of the clocked designs.
 The disadvantages, however, are also consistent with the \ac{RO} in that the propagation time will vary significantly depending on the \ac{EDA}'s placement of the circuit.
-This is not as severe a limitation for the \ac{TDL} as the \ac{RO} as while the tuning range is affected by this variation it cannot cause the centre point to vary, whereas the impact of placement on the \ac{RO} caused both variation in range and centre frequency.
+This is not as severe a limitation for the \ac{TDL} as the \ac{RO} as, while the tuning range is affected by this variation, it cannot cause the centre point to vary, whereas the impact of placement on the \ac{RO} caused both variation in range and centre frequency.
 
@@
 The Bang-Bang detector structure changes significantly in order to provide the control signals required for the \ac{TDC}.
-The \ac{TDC} expects a long pulse of logic high for duration of the measurement interval, going low once the interval has ended.
+The \ac{TDC} expects a long pulse of logic high for the duration of the measurement interval, going low once the interval has ended.
 The original single flip-flop design maintains a value between rising edges of the reference clock and could not generate the required pulse.
@@
 The $Q$ outputs of these registers are connected to \iacs{XOR} gate which generates a pulse as long as a rising edge has only been seen on one signal.
-Once a second rising edge occurs the \acs{XOR} output will be set to 0, thus terminating the pulse.
+Once a second rising edge occurs, the \acs{XOR} output will be set to 0, thus terminating the pulse.
 On completion of the measurement interval, the flip-flops must be reset in preparation for the next measurement cycle, and this can be accomplished by ANDing together the $Q$ outputs of each signal, producing a \texttt{clear} signal that will go high when both flip-flops have a logic high output.
 
-To determine the sign itself, which clock edge first had a rising edge must be recorded, and this can be done by setting \iac{SR} latch if the rising edge of the generated signal occurs first and resetting in the case of the reference clock having the first rising edge.
-Looking at the whole measurement cycle however, it is apparent that all 4 combinations possible of two inputs occur, putting the \ac{SR} latch into which the output is undefined.
-Careful design of the system can avoid problems arising from this.
+To determine the sign itself, the clock edge which first had a rising edge must be recorded, and this can be done by setting \iac{SR} latch if the rising edge of the generated signal occurs first and resetting in the case of the reference clock having the first rising edge.
+Looking at the whole measurement cycle, however, it is apparent that all four combinations possible of two inputs occur, putting the \ac{SR} latch into which the output is undefined.
+This problem can be avoided through careful design.
 Examining the truth table in Table \ref{table:sr_tt}, the combination to be avoided is the presence of a 1 on both outputs, precisely the situation seen at the end of the measurement interval.
 In order to avoid the undefined state occurring at the end of the measurement interval, both inputs and outputs to the \ac{SR} latch can be inverted using a NOT gate.
-Table \ref{table:sr_tt} also illustrates how the $\overline{Q}$ output now behaves as $Q$ did previously, however the end of measurement state where both input registers have been triggered now causes the \ac{SR} latch to enter the hold state where it maintains the previous values.
+Table \ref{table:sr_tt} also illustrates how the $\overline{Q}$ output now behaves as $Q$ did previously, however, the end of measurement state where both input registers have been triggered now causes the \ac{SR} latch to enter the hold state, where it maintains the previous values.
 The sign, $\overline{Q}$ of the \ac{SR}, can then be loaded into a register using the \texttt{clear} signal generated previously, to maintain it until the next measurement completes.
@@
 This, however, ignores the potential for metastability which may arise if both signals have a rising edge within an extremely short period of time.
-While it may in other systems be feasible to dismiss this occurrence as a low probability event, the goal of \iac{PLL} is to reduce the time delay between rising edges to zero which increases the possibility of this occurring, which may lead to indeterminate behaviour in other parts of the feedback network.
+While it may be feasible in other systems to dismiss this occurrence as a low probability event, the goal of \iac{PLL} is to reduce the time delay between rising edges to zero.
+This increases the possibility of a collision occurring, which may lead to indeterminate behaviour in other parts of the feedback network.
 In order to avoid this in \iac{ASIC}, an arbitration circuit is placed between the \ac{SR} latch and the register storing the sign between measurements, which will default to one of the two signals in this case.
-This is however a mixed signal circuit, such as that proposed by Tierno \textit{et al} (2008) and implemented by Zianbetov and Shan \cite{tierno2008wide,zianbetov2013phd,shan2014phd}, which requires two \ac{PMOS} and two \ac{NMOS} transistors and cannot be implemented on \iac{FPGA}.
+This is, however, a mixed signal circuit, such as that proposed by Tierno \textit{et al} (2008) and implemented by Zianbetov and Shan \cite{tierno2008wide,zianbetov2013phd,shan2014phd}, which requires two \ac{PMOS} and two \ac{NMOS} transistors and cannot be implemented on \iac{FPGA}.
 As the output of the \ac{SR} latch will be indeterminate prior to the detection of the first edge, the impact of both edge detection registers' output is that the \ac{SR} latch attempts to hold this indeterminate value as an output.
@@
 \subsubsection{Tapped Delay Line}
-To complement the asynchronous sign detection circuit, an asynchronous \ac{TDC} is also required, as otherwise there would be no benefit to using the above method over the more easily understood \ac{FSM} based design as quantisation to the clock period would happen in the \ac{TDC} rather than the sign detection circuitry.
-The standard method for implementing \iac{TDC} on \iac{ASIC} is through a tapped delay line, where a signal is propagated through a chain of fixed length delays, and at the end of the measurement interval it is recorded whether or not the signal has propagated to each point in the chain.
+To complement the asynchronous sign detection circuit, an asynchronous \ac{TDC} is also required, as otherwise there would be no benefit to using the above method over the more easily understood \ac{FSM} based design, as quantisation to the clock period would happen in the \ac{TDC} rather than the sign detection circuitry.
+The standard method for implementing \iac{TDC} on \iac{ASIC} is through a tapped delay line, where a signal is propagated through a chain of fixed length delays, and at the end of the measurement interval whether or not the signal has propagated to each point in the chain is recorded.
 This architecture is illustrated in Figure \ref{fig:simple_tdc}.
 As with the mixed signal circuitry previously addressed, \iac{TDL} cannot be exactly replicated on \iac{FPGA}, but through the use of inverters it is possible to replace the unrealisable well bounded delays.
-As with the \ac{RO}, the delay though the inverters is not nearly as well bounded, and significant variation between delays, in both individual \acp{TDL} and within the same \ac{TDL}, due to the implementation occurs This variation notwithstanding, the inverter pairs are a suitable replacement, and this method of time-to-digital conversion provides both resolution advantages and the presence of implementation based variation over a clocked design.
-These prove particularly beneficial when attempting to verify theoretical results.
+As with the \ac{RO}, the delay though the inverters is loosely bounded, and significant variation between delays, in both individual \acp{TDL} and within the same \ac{TDL}, due to the implementation exists.
+This variation notwithstanding, the inverter pairs are a suitable replacement, and this method of time-to-digital conversion provides both resolution advantages and the presence of implementation based variation over a clocked design.
+These attributes +prove particularly beneficial when attempting to verify theoretical results.
 
@@
 \section{Error Combiner}
-In a regular \ac{ADPLL} an error combining circuit is not required as there is never more than one reference signal, and as such the input of this block would be equivalent to the output.
-However in \iac{ADPLL} network multiple reference signals are required, one for each neighbour, thus requiring some method to combine.
-Simply averaging these signals was not found to be viable by Pratt and Nyugen in their original paper for analogue systems, however, the work of Javidan \textit{et al}, showed that in a digital system this method will not result in mode locking, provided start-up is performed uni-directionally \cite{javidan2011all}.
+In a regular \ac{ADPLL} an error combining circuit is not required as there is never more than one reference signal, and thus the input of this block would be equivalent to the output.
+However in \iac{ADPLL} network, multiple reference signals are required, one for each neighbour, thus requiring combination.
+Simply averaging these signals was not found to be viable by Pratt and Nyugen in their original paper for analogue systems \cite{pratt1995distributed}.
+However, the work of Javidan \textit{et al}, showed that in a digital system this method will not result in mode locking, provided start-up is performed uni-directionally \cite{javidan2011all}.
 
-Accordingly, any intended error combiner design implementing an average should also be reconfigurable, so that it satisfies the reconfigurability constraint.
-Given this is a digital system the most straightforward way to implement this is as a weighted sum, followed by a division.
-The final \ac{ADPLL} network will be laid out in a Cartesian grid so a maximum of four values will possibly be combined, with only the non-corner edge nodes having a number of neighbours that is not a power of two.
-The simplest, and most processing time efficient, method of calculating the weighted sum restricts the weights to powers of two, as this permits their implementation as a left bit shift which is trivially achieved in hardware by a part select.
+Therefore, any intended error combiner design implementing an average, should also be reconfigurable, so that it satisfies the reconfigurability constraint.
+Given this is a digital system, the most straightforward way to implement this is as a weighted sum, followed by a division.
+The final \ac{ADPLL} network will be laid out in a Cartesian grid, so a maximum of four values will possibly be combined, with only the non-corner edge nodes having a number of neighbours that is not a power of two.
+The simplest, and most processing time efficient, method of calculating the weighted sum restricts the weights to powers of two.
+This permits their implementation as a left bit shift, trivially achieved in hardware by a part select.
 Similarly this reasoning can be expanded to the division, where the complicated division by three for a non-corner edge node is instead replaced by division by four.
-This means the division is implemented identically by all error combiners in the network, and implemented trivially by a right shift of two bits.%TODO ask pierre re 3 div 4
+This means the division is implemented identically by all error combiners in the network, and implemented trivially by a right shift of two bits.
+Appropriate gain choice can negate the error in the averaging.
 
 \section{Loop Filter}
-The \acl{LF} in \iac{ADPLL} can be implemented as \iac{PI} controller, a circuit that can be implemented in a wide variety of ways, with the exact implementation having little impact on behaviour.
+The \acl{LF} in \iac{ADPLL} can be implemented as \iac{PI} controller, a circuit with many valid variations, with the exact implementation having little impact on behaviour.
 The first choice to be made is whether the designer wishes to implement the controller as \iac{FIR} or \ac{IIR} filter.
@@
 In both cases the filter can be implemented using either fractional or integer representation of the data, depending on the wishes of the designer.
-Carrying out computations using fractional values may be conceptually easier to understand as the magnitude of individual paths or taps will be independent of another, whereas using integer representations allows for a simpler implementation.
+Carrying out computations using fractional values may be easier to understand conceptually, as the magnitude of individual paths or taps will be independent of each other, whereas using integer representations allows for a simpler hardware implementation.
 If fractional representation is used, it will be done with fixed-point arithmetic as the time complexity of floating point calculations is too great and the potential benefits are inconsequential.
@@
 \subsection{\acs{FIR} \acl{LF}}
-The \acl{FIR} implementation of the \acl{LF} will take the form presented in Figure \ref{fig:fir_pi} (a), and consists series of time delays, or ``taps'', the output of which is multiplied by a coefficient and the result of each multiplier added together to form the filter output.
+The \acl{FIR} implementation of the \acl{LF} will take the form presented in Figure \ref{fig:fir_pi} (a), and consists series of time delays, or ``taps'', the output of which is multiplied by a coefficient.
+The result of each multiplication is added together to form the filter output.
 Various tools, Matlab for example, can be used to compute the coefficients and number of taps required to obtain a specific filter response.
 Using \iac{FIR} filter is advantageous as it is always stable due to the lack of a feedback path, however a large number of taps are required to obtain a frequency response akin to that of the \ac{IIR} implementation, thus requiring an equal number of multipliers and $taps-1$ adders.
-Compared to \iac{IIR}, which requires just a two adders, two multipliers and a flip-flop this has a much greater hardware footprint.
+Compared to \iac{IIR}, which requires just two adders, two multipliers and a flip-flop this, has a much greater hardware footprint.
 It is possible to condense the \ac{FIR} multiplication and addition to a single multiplier and adder, depicted in \ref{fig:fir_pi} (b), however this will take as many clock cycles to complete as there are taps.
-This, however, is not viable in this system as the filter must produce its output in a single clock cycle.
+This is not viable in an \ac{ADPLL} as the filter must produce its output in a single clock cycle.
 In the case of any coefficient symmetry, addition can be performed prior to multiplication, thus reducing the number of multipliers for each identical coefficient.
@@
 The filter can be implemented using fixed-point arithmetic, in which the designer must ensure alignment of the radix in additions and correct radix position in multiplication results.
-Consequently the fractional width of the each multiplication need not be the same, most commonly used when the proportional gain, the output of the 0th tap multiplication, is significantly larger than that of the other taps.
+Consequently, the fractional width of each multiplication need not be the same, most commonly used when the proportional gain, the output of the 0th tap multiplication, is significantly larger than that of the other taps.
 Integer arithmetic instead assumes all calculations are carried out using integers, and output of the addition stage is divided by a large number to form the module output.
@@
 \subsection{\acs{IIR} \acl{LF}}
-\Iac{PI} controller can be easily implemented as \iacl{IIR} filter in hardware, as in Figure \ref{fig:iir_pi}, requiring significantly fewer circuit elements than \iac{FIR} based design as the long series of taps and multipliers is not required, as instead integration is performed by a single multiplier and an accumulator.
-As before the frequency response of the controller can be calculated using the following equation, where $\alpha$ is the proportional and $\beta$ the integral gain:
+\Iac{PI} controller implemented most easily in hardware as \iacl{IIR} filter, as in Figure \ref{fig:iir_pi}.
+It comprises significantly fewer circuit elements than \iac{FIR} based design as the long series of taps and multipliers is not required.
+Instead integration is performed by a single multiplier and an accumulator.
+As seen before, the frequency response of the controller can be calculated using the following equation, where $\alpha$ is the proportional and $\beta$ the integral gain:
 \begin{equation*}
@@
 As the scaling factor used in this multiplier is the \ac{ki} itself, reconfiguration of the integral path of the \acl{LF} can be carried out solely by varying this gain, as opposed to the \ac{FIR} solution where the value of each tap must be recalculated to change the response of the system.
-The combination of reconfigurability and simpler hardware make the \ac{IIR} the more desirable \ac{LF} design.
+The combination of reconfigurability and simpler hardware, make the \ac{IIR} the more desirable \ac{LF} design.
 \begin{figure}[h]
@@
 
-Again both integer and fixed-point methods of \ac{LF} design are possible, with the integer method requiring logic combining the two paths to preserve the relationship between the proportional and integral gains, which from the work of Koskin \textit{et al} is known to be several binary orders of magnitude \cite{koskin2018generation}.
+Again, both integer and fixed-point methods of \ac{LF} design are possible, with the integer method requiring the logic combining the two paths to preserve the relationship between the proportional and integral gains.
+From the work of Koskin \textit{et al} this known to be several binary orders of magnitude \cite{koskin2018generation}.
 Failure to preserve a relationship of this order may lead to the \ac{ADPLL} becoming unstable, as in the upper-left region of Figure \ref{fig:gain_grid}.
-In the fixed-point system the relationship between \ac{kp} and \ac{ki} is preserved during the combination of proportional and integral paths as both values have their radix point aligned before addition is carried out.
+In the fixed-point system the relationship between \ac{kp} and \ac{ki} is preserved during the combination of proportional and integral paths, as both values are aligned using their radix point, before addition is carried out.
 \begin{figure}[h]

### 2019-04-23 thesis/Introduction.tex
partial corrections

@@
-This thesis will put forward the \ac{FPGA} as a tool in the design of \ac{ADPLL} networks to bridge the gap between software simulations and implementation in custom silicon by providing a hardware-based simulation, modelling and validation platform.
-While an FPGA lacks the direct control over the schematic and layout that \iac{ASIC} provides, the hardware nature of this platform enables the analysis of system dynamics that are not easily modelled in simulation and the testing.
+This thesis will propose the \ac{FPGA} as a tool in the design of \ac{ADPLL} networks, to bridge the gap between software simulations and implementation in custom silicon, by providing a hardware-based simulation, modelling and validation platform.
+While an FPGA lacks the direct control over the schematic and layout that \iac{ASIC} provides, the hardware nature of this platform enables the analysis of system dynamics that are not easily modelled in simulation and testing.
 \acp{ADPLL} are an entirely digital implementation of \iac{PLL}, advantageous as this allows for easier integration into the modern high transistor count \acl{IC} (\acs{IC}s) for which \ac{ADPLL} networks are a proposed clock distribution system.
 
-The goal of this project is to design and implement an extensible platform that can be used by the research team in \ac{UCD} going forward as they seek to understand the behaviour of \ac{ADPLL} networks at a higher level and to serve as a hardware test-bed for proposed new architectures or system components.
-In order to accomplish this goal a number of potential \ac{ADPLL} architectures will be investigated, implemented and tested to ensure they are function correctly.
+The goal of this project is to design and implement an extensible platform that can be used by the research team in \ac{UCD} going forward, as they seek to understand the behaviour of \ac{ADPLL} networks at a higher level and to serve as a hardware test-bed for proposed new architectures or system components.
+In order to accomplish this goal, a number of potential \ac{ADPLL} architectures will be investigated, implemented and tested to ensure they function correctly.
 Individual \acsp{ADPLL}, however, will not give sufficient insight into the behaviour of a network, so once the \ac{ADPLL} designs have been established, each will be implemented as part of a network of increasing sizes.
@@
 
-The design of each block, or component part, used in the network will be discussed, starting with the reasons for their selection and an explanation of the design methodology, along with any major pitfalls encountered their creation.
-The impact on performance caused by modifying the design of these blocks will again be assessed on the basis of measurement results, before comparison is made to both theoretical expectations and their use case.
+The design of each block, or component part, used in the network will be discussed, starting with the reasons for their selection and an explanation of the design methodology, along with any major pitfalls encountered in their creation.
+The impact on performance caused by modifications in the design of these blocks will again be assessed on the basis of measurement results, before comparison is made to both theoretical expectations and their use case.
 
-\Iac{FPGA} based test platform is ideal for those who wish to examine system dynamics without the time delays, financial burden or expertise required to develop a complete mixed-signal system on \iac{ASIC} but retain the ability to realistically simulate the behaviour of \iac{ADPLL} network.
+\Iac{FPGA} based test platform is ideal for those who wish to examine system dynamics without the time delays, financial burden or expertise required to develop a complete mixed-signal system on \iac{ASIC}, yet simultaneously retaining the ability to realistically simulate the behaviour of \iac{ADPLL} network.
 The result of this project is such a platform, designed to be extensible, with flexibility built into each component/module used.

### 2019-04-23 thesis/LayAbstract.tex
partial corrections

@@
 \begin{document}
-With the increasing proliferation of ``smart'' devices the need for low power yet high speed devices has never been greater.
-Each smart device contains a processor to control the device, each containing complex circuitry, requiring extremely exact synchronisation.
+With the increasing proliferation of ``smart'' devices, the need for lower power yet higher speed circuits has never been greater.
+Each smart device contains a processor to control the device, containing complex circuitry, requiring extremely exact synchronisation.
 The task of this synchronisation falls to the ``clock'' signal which must occur at the same instant in all areas of the processor.
-Conventional solutions to this problem cannot satisfy both power and speed requirements simultaneously, so designers have proposed the ADPLL Network which approaches the problem from the other side.
-Rather than generate the clock once and send it around the chip, which consumes a large amount of power, an ADPLL network divides the chip into a grid and generates many signals that each serve a local area, only requiring non time critical control signals to be sent over large distances.\\
-The production of chips to test designs is both expensive and time consuming so designers must ensure that mistakes have not been made, accomplished by simulating the behaviour of the designs using complex behavioural and theoretical models.
-This thesis will discuss the use of Field Programmable Gate Arrays (FPGA) as a hardware testing, simulation and validation platform, to be used prior to test chip production, that is both cost and time effective.
+Conventional solutions to this problem cannot satisfy both power and speed requirements simultaneously, so designers have proposed the ADPLL Network takes a different approach to the problem.
+Rather than generate the clock once and send it around the chip, and action which consumes a large amount of power, an ADPLL network divides the chip into a grid and generates many signals that each serve a local area, thus only requiring non time-critical control signals to be sent over large distances.\\
+The production of chips to test designs is both expensive and time consuming, so designers must ensure that mistakes have not been made.
+This is accomplished by simulating the behaviour of the designs, using complex behavioural and theoretical models.
+This thesis will discuss the use of Field Programmable Gate Arrays (FPGA) as a hardware testing, simulation and validation platform, to be used prior to test chip production, being both cost and time effective.
 The hardware nature of an FPGA enables the analysis of behaviours that would not be possible in software, without the cost and time penalties of a custom chip.
-This is made possible by the ability to reconfigure the chip at will, albeit with comparatively lesser capabilities.\\
+This is made viable by the ability to reconfigure the chip at will, albeit with comparatively lesser capabilities.\\
 This thesis implements networks using three designs of ADPLL, each using a different architecture, and highlights the use cases to which each is best suited.
-The performance of each design is then analysed and this compared to the suggested use cases.
+The performance of each design is then analysed and compared to the suggested use cases.
 Additionally the impact of more minor architectural modifications is tested and documented.

### 2019-04-23 thesis/NetworkImplemenation.tex
partial corrections

@@
 The \ac{HDL} Verilog was selected for use in this project due to my prior exposure to that language.
-5 MHz was chosen as the target centre frequency for these \acp{ADPLL}, during initial testing it was discovered that the waveform output via the Digilent Nexys4 development board's \ac{GPIO} ports began to degrade around the 10 MHz mark, preventing accurate analysis.
+$5~\si{\mega\hertz}$ was chosen as the target centre frequency for these \acp{ADPLL} as during initial testing, it lies well within the $10~\si{\mega\hertz}$ degradation point of the the Digilent Nexys4 board's \ac{PMOD} headers.
 
 \section{\acs{ADPLL} Architectures}
-Three different architectures of \ac{ADPLL} were implemented in this project for the purposes of testing, and re-use in the future within the research team in \acl{UCD}.
-Between each design a number of blocks remain constant, as only a single instance of that block was implemented, as variation its design would have negligible impact on the final system.
+Three different architectures of \ac{ADPLL} were implemented in this project for the purposes of testing and re-use in the future within the research team in \acl{UCD}.
+Between each design a number of blocks remain constant, because only a single instance of that block was implemented as variation in its design would have negligible impact on the final system.
 These blocks were the error combiner and frequency divider.
-The loop filter was implemented both using fractional and integer arithmetic, but as the exact same calculation was performed in both cases there was no impact on the output of the module.
+The loop filter was implemented using both fractional and integer arithmetic, but as the exact same calculation was performed in both cases, there was no impact on the output of the module.
 
-The three architectures chosen for implementation represent a progression from entirely \iac{FPGA} clock driven system to \iac{ADPLL} with all components asynchronous with respect to that clock.
+The three architectures chosen for implementation, represent a progression from an entirely \ac{FPGA} clock driven system to \iac{ADPLL} with all components asynchronous with respect to that clock.
 Therefore the first \ac{ADPLL} design features a clocked oscillator and phase detector, specifically the linear frequency design mentioned in Chapter \ref{chap:3} and the clocked phase detector using a state machine and a bi-directional counter.
@@
 Whilst, apart from frequency, there were no guidelines as to what the system should resemble and a number of attributes could be chosen arbitrarily, a future user of this platform could have entirely different requirements or specifications.
-As such each block was implemented in such a way that changing the number of bits used for a certain signal, the target frequency or swapping between different blocks could be done with the sole requirement of changing Verilog \texttt{localparam}\footnote{A \texttt{localparam} is a constant that cannot be modified in a module instance statement \cite{hdlworks}}s in the module \footnote{Module is the Verilog term for a number of logic elements grouped to provide a certain functionality} in which change was made, which would then propagate to any sub-modules affected.
+As such each block was implemented in such a way that changing the number of bits used for a certain signal, changing the target frequency or swapping between different blocks could be done with the sole requirement of changing Verilog \texttt{localparam}\footnote{A \texttt{localparam} is a constant that cannot be modified in a module instance statement \cite{hdlworks}}s in the module \footnote{Module is the Verilog term for a number of logic elements grouped to provide a certain functionality} in which change was made, which would then propagate to any sub-modules affected.
 Starting the design process with this in mind was a major benefit at later stages when frequent modifications were being made to test their impact, or when changing between \ac{ADPLL} designs.
@@
 \subsubsection{Error Combiner}
-The error combiner in this design had a number of requirements in order to be suitable for use in both a single \ac{ADPLL} as well as a network operating in both uni- and bi-directional modes, alongside modifications which would change the widths of the error signals requiring averaging.
+The error combiner in this design had a number of requirements, in order to be suitable for use in both a single \ac{ADPLL} as well as in a network operating in both uni- and bi-directional modes, alongside modifications which would change the widths of the error signals requiring averaging.
 The module was implemented on the basis that the \ac{LF} input width would be identical to the phase detector output width.
@@
 \subsubsection{Frequency Divider}
-The frequency divider is implemented using ``Divider 2'' from Chapter \ref{chap:3}, as only divisions ratios of 2, 4 and 8 were selected for use in the testing process.
-For ease of testing, the divider had outputs representing each of these division ratios simultaneously, which could be multiplexed between at runtime if so desired.
+The frequency divider is implemented using ``Divider 2'' from Chapter \ref{chap:3}, as only division ratios of 2, 4 and 8 were selected for use in the testing process.
+For ease of testing, the divider had outputs representing each of these division ratios simultaneously, which, at runtime, could be multiplexed between, if so desired.
 A fourth output passed the input through unmodified, allowing for the divider to be removed without regeneration of the \ac{FPGA} configuration.
@@
 \ac{FIR} was dismissed due to the extra hardware required to implement the calculations within a single clock cycle, as the \ac{LF} is clocked using the oscillator output, and the greater difficulty of gain adjustment compared to \iac{IIR}.
-Clocking on using the generated signal ensures that the discrete time integration is only carried out once per phase comparison.
+Clocking on the generated signal ensures that the discrete time integration is only carried out once per phase comparison.
 A pitfall that may be encountered implementing \iac{ADPLL} featuring a divider is not clocking the module on the divided clock, which will result in the integration being carried out at the oscillator output frequency.
 
-Each of integer and fixed-point arithmetic were used to implement a filter respectively, however both designs are interchangeable as they perform the integration identically, given the same proportional and integral gains.
-All testing was carried out using integer arithmetic, as in Figure \ref{fig:integer_lf}, however the interchangeability will be confirmed later in this chapter, in Section \ref{section:minor_variations}.
-Additionally the \ac{LF} supports variation of both \acs{ki} and \acs{kp} at runtime, however, if a static value is desired the runtime variation may be disabled using a \texttt{parameter}\footnote{Parameters are constant values that may be changed at compile time, or in the module instance statement \cite{hdlworks2}}.
+Each of integer and fixed-point arithmetic were used to implement a filter respectively, however, both designs are interchangeable as they perform the integration identically, given the same proportional and integral gains.
+While testing was carried out using both methods, all results in Chapter \ref{chap:5} were obtained using integer arithmetic, as in Figure \ref{fig:integer_lf}.
+Additionally, the \ac{LF} supports variation of both \acs{ki} and \acs{kp} at runtime, however, if a static value is desired the runtime variation may be disabled using a \texttt{parameter}\footnote{Parameters are constant values that may be changed at compile time, or in the module instance statement \cite{hdlworks2}}.
 \begin{figure}[h]%
@@
 Figure \ref{fig:integer_lf} contains \iac{RTL} diagram describing the implementation.
-Important to note is the shift applied to the result of the multiplication by \acs{kp} which, as the integration is performed with a larger width to avoid accumulator overflow, preserves the intended relationship between proportional and integral gains.
+Important to note, is the shift applied to the result of the multiplication by \acs{kp} which, as the integration is performed with a larger width to avoid accumulator overflow, preserves the intended relationship between proportional and integral gains.
 The value of $k_p$ and $k_i$ can be computed using the following formulae:
@@
 This was the first solution addressed in the course of this project as it is the simplest to both implement and test, as all aspects of the \ac{ADPLL} are driven by the \ac{FPGA} clock.
-Compared to later designs utilising the \ac{RO}, Vivado simulations could be carried out, with accurate timing behaviour, depicting the locking process.
-The \ac{DCO} was chosen to be linear in frequency in order to obtain a square wave output as it was not yet known whether the varying pulse width of a period linear design would cause a problem in modules which, at that stage in the project, had not yet been designed.
+Compared to later designs utilising the \ac{RO}, Vivado simulations of the locking process with accurate timing behaviour could be carried out.
+The \ac{DCO} was chosen to be linear in frequency in order to obtain a square wave output, as it was not known at this stage of the project whether the varying pulse width of the \ac{FPGA} clocked, period linear design would cause a problem in modules which, at that stage in the project, had not yet been designed.
 Indeed, a later discarded version of the \ac{FPGA} clocked \ac{PFD} relied on equal low and high times.
@@
 This still leaves three parameters undecided: The \ac{FPGA} clock, the bit width of the counter and the bias point.
-As the \ac{FPGA} clock determines the frequency step of the design, a test counter was implemented initially to determine the maximum frequency at which the timing violations would not occur, at this was determined to be 275 MHz.
+As the \ac{FPGA} clock determines the frequency step of the design, a test counter was implemented initially to determine the maximum frequency at which the timing violations would not occur, which was determined to be 275 MHz.
 The equation can then be updated, giving the relationship between bias point and counter width:
@@
 \end{align}
-The maximum value of the counter must also be large enough such that with a bias point that allows for a reasonable range of frequency steps to be added or removed. 
+The maximum value of the counter must also be large enough, that with a bias point that allows for a reasonable range of frequency steps to be added or removed. 
 This oscillator was originally designed to allow for 64 control codes either side of the bias, which called for a bias point of 65 or greater.
@@
 \end{align}
-From the above equation, the smallest counter that can satisfy this constraint is 12 bits wide, however when testing was carried out using a 275 MHz clock timing violations were discovered by Vivado, resulting \ac{FPGA} clock speed reduction to 258 MHz in order to resolve these violations.
+From the above equation, the smallest counter that can satisfy this constraint is 12 bits wide, however when testing was carried out using a 275 MHz clock, timing violations were discovered by Vivado, resulting in a \ac{FPGA} clock speed reduction to 258 MHz in order to resolve these violations.
 The correct bias point could then be calculated as $79$:
@@
 As phase error is a two's complement signed value, its addition to the bias must be carried out using signed arithmetic.
-Provided the bias is greater than the minimum possible phase error, as has been ensured in this case, the result of this addition is positive signed integer.
+Provided the bias is greater than the minimum possible phase error, as has been ensured in this case, the result of this addition is a positive signed integer.
 Being a positive quantity, additional logic performing a conversion from a signed to an unsigned representation is required.
@@
 
-The counter is implemented using saturation arithmetic to prevent overflow at high phase differences which could potentially harm the ability of the \ac{ADPLL} to obtain a lock.
-In the case of this \ac{PD} the phase error at which overflow may occur can be computed using the bit width of the counter and temporal resolution.
-$error_{max} = 63\cdot3.875~\si{\nano\second} = 244.125~\si{\nano\second}$.
-As this is greater than the period it cannot occur for purely a phase difference, but may occur in the presence of either a frequency divider in the feedback path, thus increasing the period of the signal used for comparison, or of a significant frequency difference between the signals.
+The counter is implemented using saturation arithmetic to prevent overflow at high phase differences, which could potentially harm the ability of the \ac{ADPLL} to obtain a lock.
+In the case of this \ac{PD}, the phase error at which overflow may occur can be computed using the bit width of the counter and temporal resolution.
+\begin{equation}
+error_{max} = 63\cdot3.875~\si{\nano\second} = 244.125~\si{\nano\second} 
+\end{equation}
+As this is greater than the period, it cannot occur for purely a phase difference, but may occur in the presence of either a frequency divider in the feedback path, thus increasing the period of the signal used for comparison, or of a significant frequency difference between the signals.
 When implementing the clamping logic, the decision was taken to make the detection range symmetrical, thus the output range was reduced by one increment to $[-63,63]\cap\mathbb{Z}$.
@@
 Figure \ref{fig:updown_ctr} depicts the implementation of this counter.
-Apart from the aforementioned capping of the measurement range implemented using a pair of comparators and a multiplexer, important to note is the bit width of the increment.
+Apart from the aforementioned capping of the measurement range implemented using a pair of comparators and a multiplexer, it is important to note the bit width of the increment.
 Despite overflow being disallowed when the number is interpreted as a signed number, the addition of a 7 bit wide -1 to the accumulator value will cause overflow in the adder.
@@
     \centering
-    \includegraphics[width=0.6\textwidth]{../uncertain}
+    \includegraphics[width=0.5\textwidth]{../uncertain}
     \caption[Circumstances of mode locking behaviour]{Circumstances of mode locking behaviour.}
@@
 However in the presence of a frequency differential, the rising edge on the generated signal may occur only at location ``c'', and with the falling edge termination condition causing the interval to terminate before this edge was detected, at point ``b''.
-This lead to a slightly lower magnitude phase lag detected, however this is a minor flaw, and not the cause of the mode locking.
+This led to a slightly lower magnitude phase lag detected, however this is a minor flaw, and not the cause of the mode locking.
 As the rising edge at point ``c'' occurs after the interval has terminated, it is treated as a fresh measurement interval, terminating at ``d''.
 As in this interval, the generated signal has seen the first rising edge, a large lead will be recorded.
-As each correction is made with a delay of one cycle, in order to avoid timing violations in the feedback path, there is potential for oscillation between lead and lag detection to occur depending on the initial conditions.
+As each correction is made with a delay of one cycle, in order to avoid timing violations in the feedback path, there is potential for oscillation between lead and lag detection to occur, depending on the initial conditions.
 This low probability behaviour was observed by chance, due to coding error, while using a frequency divider in the feedback path as this ensured the lead and lag measurements were of identical magnitude.
@@
 \subsubsection{Simulations}
-%TODO simulations
 As this was the first \ac{ADPLL} design implemented, various Vivado test benches were used to verify the behaviour of each module.
-Simulations with this design were made more efficient due to deterministic timing in all modules, and absence of any inverter based components which required the use of post-implementation simulations in order to avoid zero delay oscillations.
+Simulations with this design were made more efficient due to deterministic timing in all modules, and the absence of any inverter based components which required the use of post-implementation simulations in order to avoid zero delay oscillations.
 Deterministic timing means \ac{ADPLL} 1 offers an additional, and significant, advantage over other designs as it can be entirely simulated in Vivado with accurate timings.
-This ability was vital initially, as it revealed the presence of edge situations that had not been accounted for, and through exporting the waveforms to a log file and analysis performed using Matlab comparison made to the measured system behaviour.
+This ability was vital initially, as it revealed the presence of edge situations that had not been accounted for and, through exporting the waveforms to a log file and analysis performed using Matlab, comparison made to the measured system behaviour.
 
@@
 Lock was visually determined to have occurred after $20.91~\si{\micro\second}$, after which the oscillator frequency began to change in a repetitive pattern.
-later the same test carried in hardware with identical initial conditions, the results of which are shown in Figure \ref{fig:sim_locking} (b).
+Later the same test was carried in hardware with identical initial conditions; the results of which are shown in Figure \ref{fig:sim_locking} (b).
 This plot is derived from data captured using an Agilent MSO7054A, using a setup that will be further explained in Section \ref{section:measurement_setup}.
@@
 The listed components were brought together in \textsc{NetworkADPLL.v}, attached in Appendix \ref{adx:code} Listing \ref{lst:network_adpll}.
-Important to note is that each \ac{ADPLL} only contains two \aclp{PD} despite during network operation the need for phase comparisons to be made with up to four neighbours.
+It is important to note that each \ac{ADPLL} contains only two \aclp{PD}, despite the fact that during network operation the need for phase comparisons to be made with up to four neighbours.
 The two ``missing'' \acp{PD} are in fact synthesised as part of the other \acp{ADPLL}, with which the comparison would be made.
@@
 Each \ac{ADPLL} performs comparison with the \ac{DCO} ``above'' and ``left'' of them in the Cartesian grid, and a port in the module outputs the results for use by the other \acp{ADPLL} involved.
-%TODO juvenile wording
 Table \ref{table:adpll1} contains a brief overview of the bit widths, tuning ranges and other configuration information for this \ac{ADPLL}.
@@
 \subsection{\acs{ADPLL} Design 2}
-The second \ac{ADPLL} design implemented during the course of this project was somewhat of a stepping stone between fully synchronous and asynchronous to the \ac{FPGA} clock.
+The second \ac{ADPLL} design implemented during the course of this project was sa stepping stone between fully synchronous and asynchronous to the \ac{FPGA} clock.
 The clocked phase detector is reused from above, albeit with an altered error width, while the \ac{DCO} is replaced by \iacl{RO}.
-While the phase detector was known to work due to testing in \ac{ADPLL} 1, the \ac{DCO} could not be simulated using accurate timings so all testing had to be done in hardware.
+While the phase detector was known to work due to testing in \ac{ADPLL} 1, the \ac{DCO} could not be simulated using accurate timings, so all testing had to be done in hardware.
 This presented a challenge, as depending on the implementation, the characteristics of the \ac{DCO} could change, most importantly the centre frequency.
@@
 \ac{ADPLL} 2 exchanges the \ac{FPGA} clocked oscillator for \ac{RO}, with the aim of obtaining performance more akin to that of \iac{ASIC} based mixed-signal implementation, due to the variation in period steps both between oscillators and within the same oscillator due to the oscillator layout.
-Also advantageous is the improvement in period resolution obtained using this design, when compared to that of the \ac{FPGA} clocked design.
+Also advantageous, is the improvement in period resolution obtained using this design, when compared to that of the \ac{FPGA} clocked design.
 Taking the formula given in Chapter \ref{chap:3} and the inverter propagation time used by Vivado in post synthesis/implementation simulations of $315~\si{\pico\second}$, the period step can be computed:
@@
 \end{figure}
-Figure \ref{fig:ro_impl} contains \iac{RTL} diagram of the \ac{RO} as implemented.
+Figure \ref{fig:ro_impl} contains \iac{RTL} diagram of the \ac{RO}, as implemented.
 The inverter chain is specified with a maximum length of $n$ and a maximum number of removable inverters of $k$.
 In order to preserve the unstable circuit, by maintaining an odd number of inverters, only multiples of two inverters may be removed at a time.
-The already introduced \ac{DCO} and \ac{PFD} designs all followed the same convention regarding phase error and the corresponding impact on the control code: In the \ac{PFD}, if the generated signal was lagging the reference, meaning it should ``go faster'' in order to ``catch up'', the error given a positive sign.
-Similarly in the \ac{FPGA} clocked \ac{DCO}, a positive change in the magnitude of the control code resulted in a higher frequency.
+The previously introduced \ac{DCO} and \ac{PFD} designs all followed the same convention regarding phase error and the corresponding impact on the control code: in the \ac{PFD}, if the generated signal is lagging the reference, meaning it should ``go faster'' in order to ``catch up'', the error is given a positive sign.
+Similarly in the \ac{FPGA} clocked \ac{DCO}, a positive change in the magnitude of the control code results in a higher frequency.
 In the interests of consistency and extensibility the same behaviour was implemented for this \ac{DCO} also, with the number of inverters in the chain given as:
@@
 
-The bias point functions differently in this oscillator, as unlike the \ac{FPGA} clocked design it is not required to set the centre of the tuning range, as in the \ac{RO} this lies at $n-\frac{k}{2}$.
-Instead the bias point servers as a conversion between the signed \ac{LF} output, which is in two's complement form, and the unsigned requirement of the control code.
+The bias point functions differently in this oscillator, as unlike the \ac{FPGA} clocked design, it is not required to set the centre of the tuning range, as in the \ac{RO}, this lies at $n-\frac{k}{2}$.
+Instead the bias point serves as a conversion between the two's complement signed \ac{LF} output and the unsigned requirement of the control code.
 This is achieved by adding on the midpoint of the unsigned range and storing the result in an unsigned value.
-Using the case where the width of the control code is three: The range of the signed three bit integer is from $-4$ (100) to $3$ (011) which when directly converted to unsigned values become $4$ and $3$, thus destroying the relationship between lead and lag.
-Worse still the signed value of $-1$ (111) becomes $7$.
-Adding on the mid range value solves this problem, with a signed value of $-4$ now mapping to $0$ and $3$ to $7$.
-Comparing this behaviour to response of the \ac{RO} to the control code, it can be trivially seen that the minimum possible period will still correspond to the largest control code prior to conversion and vice versa for the maximum period.
+Using the case where the width of the control code is three: the range of the signed three bit integer is from $-4$ (100) to $3$ (011) which, when directly converted to unsigned values, become $4$ and $3$, thus destroying the relationship between lead and lag.
+Worse still, the signed value of $-1$ (111) becomes $7$.
+Adding on the mid range value solves this problem, with a signed value of $-4$ now mapping to $0$ and $3$ mapping to $7$.
+Comparing this behaviour to the inverter chain oscillator's response to the control code, it can be trivially seen that the minimum possible period will still correspond to the largest control code prior to conversion and vice versa for the maximum period.
 
@@
 It is important to choose the correct directive, otherwise the compiler may not behave as expected.
-In the case of Vivado, it is important not to confuse the \texttt{KEEP} or \texttt{KEEP\_HIERARCHY} commands with that of \texttt{DONT\_TOUCH}, with former commands only ensuring that the logic elements will be maintained in synthesis, but not in any subsequent stages \cite{synth_ug}.
+In the case of Vivado, it is important not to confuse the \texttt{KEEP} or \texttt{KEEP\_HIERARCHY} commands with that of \texttt{DONT\_TOUCH}, with the former commands only ensuring that the logic elements will be maintained in synthesis, but not in any subsequent stages \cite{synth_ug}.
 
 As mentioned in previous chapters, the absence of direct control over layout can lead to significant variation of the propagation time though inverters.
-This may occur in three ways: Firstly, the layout of individual \acp{RO} may be significantly different, thus resulting in poor overlap of tuning regions.
+This may occur in three ways: firstly, the layout of individual \acp{RO} may be significantly different, thus resulting in poor overlap of tuning regions.
 Anecdotally, while implementing a 3x3 network, 7 of the 9 oscillators had centre frequencies within a $100~\si{\kilo\hertz}$ span but two lay more than $500~\si{\kilo\hertz}$ away in opposite directions which prevented the network from locking.
-Secondly, within \iac{RO} the propagation delay between each inverter may vary, which results in a variable frequency step, possibly changing the locking range.
+Secondly, within \iac{RO} the propagation delay between each inverter may vary, resulting in a variable frequency step, possibly changing the locking range.
 These two effects represent an extreme version of the variation due to process or manufacturing seen on \acp{ASIC}.
-The final variation is possibly the most frustrating, and occurs when between implementations the \ac{EDA}, Vivado in the case of this project, changes the layout of the \ac{RO}.
+The final variation is possibly the most frustrating, and occurs between implementations when the \ac{EDA}, Vivado in the case of this project, changes the layout of the \ac{RO}.
 This may occur as a result of a direct change, or as a result of seemingly innocuous changes to unrelated modules.
@@
 
-With the hardware description of the \ac{RO} completed, testing of an instance of the \ac{RO} with a constant control code of zero was attempted, and it was discovered that the estimation of the centre frequency based on the propagation delay was off by $100$s of $\si{\kilo\hertz}$.
+With the hardware description of the \ac{RO} completed, testing of an instance of the \ac{RO} with a constant control code of zero was attempted, and it was discovered that the estimation of the centre frequency, based on the propagation delaym was off by $100$s of $\si{\kilo\hertz}$.
 The addition of 100 inverters was required to restore a centre of $5~\si{\mega\hertz}$.
-After the \ac{RO} was integrated into \ac{ADPLL} 2 the centre frequency was discovered to have changed once more.
+After the \ac{RO} was integrated into \ac{ADPLL} 2, the centre frequency was discovered to have changed once more.
 The minor variation caused by the now variable control code had changed the average propagation delay through the inverters once more, now requiring 373 inverters for $5~\si{\mega\hertz}$.
@@
 
-When, later in the project, additional \acp{RO} were implemented to form networks, 373 inverters proved a suitable starting point which consistently delivered \acp{DCO} with a centre frequency within $200~\si{\kilo\hertz}$ of $5~\si{\mega\hertz}$.
+When, later in the project, additional \acp{RO} were implemented to form networks, 373 inverters proved a suitable starting point, consistently delivering \acp{DCO} with a centre frequency within $200~\si{\kilo\hertz}$ of $5~\si{\mega\hertz}$.
 For the purposes of describing a typical tuning range, the average propagation delay of \iac{RO} consisting of 373 inverters with a centre frequency of $5.06~\si{\mega\hertz}$ will be used.
-To avoid the aforementioned issues with constant control codes, the fixed code was achieved by implementing \iac{ADPLL} and setting the \ac{PI} filter gains to zero at runtime before performing a reset.
+To avoid the aforementioned issues with constant control codes, the fixed code was achieved by implementing \iac{ADPLL} and setting the \ac{PI} filter gains to zero at runtime, before performing a reset to obtain the mid-point control code.
 
-Using this updated maximum number of inverters the frequency range can be recalculated.
+Using this updated maximum number of inverters, the frequency range can be recalculated.
 Firstly, the new minimum and maximum numbers of inverters are $309$ and $373$ respectively, with a mid point at $341$.
@@
 The module implementing the \ac{RO} can be found in \textsc{RingOsc.v}, attached in Appendix \ref{adx:code} Listing \ref{lst:ring_osc}.
-In the module instance statement it can be seen that a second generated output is also present.
-This second output is taken from a point in the fixed part of the chain, a significant number of inverters prior to the minimum possible due to control code adjustment.
+In the module instance statement, it can be seen that a second generated output is also present.
+This second output is taken from a point in the fixed part of the chain, a significant number of inverters prior to the minimum number of inverters removable by the control code adjustment.
 This signal is at an identical frequency to that of the generated signal, however will not be immediately affected by a change in the control code, which could in some edge cases change the polarity of the signal.
@@
 5 bits corresponds to a counter range of $[-15,15]\cap\mathbb{Z}$, which will saturate at a time delay of $15\times3.875\si{\nano\second} = 58.125~\si{\nano\second}$, a phase lead or lag of $\frac{}{}\times360\si{\degree} = 104.4\si{\degree}$ at $5~\si{\mega\hertz}$.
-This is an adequate range, as saturation will occur well outside of the normal operating range.
-This reduction is most noticeable when a divider is active in the feedback path, however during locked operation a phase shift of $58.125~\si{\nano\second}$ will not occur, so this impact is only felt during acquisition.
+This is an adequate, as saturation will occur well outside of normal operating conditions.
+This reduction is most noticeable when a divider is active in the feedback path, however, even with a divider, during locked operation a phase shift of $58.125~\si{\nano\second}$ will not occur, so this impact is only felt during acquisition.
 
@@
 As previously mentioned, \ac{ADPLL} 2 represents a midway point between \ac{FPGA} clocked designs and those which are asynchronous.
-As the source of the asynchronicity is the inverter chain \ac{DCO}, a more realistic jitter would be expected from this design, with there being a significant distribution of period steps.
-Among other aspects, this will have a noticeable impact in steady state conditions, where the degree of fractional synthesis required will vary from oscillator to oscillator depending on the closest distance from the reference period to a period obtainable by the individual oscillator.
+As the source of the asynchronicity is the inverter chain \ac{DCO}, a more realistic jitter would be expected from this design, stemming from the significant distribution of period steps.
+The biggest impact of this change will be felt in steady state conditions, where the degree of fractional synthesis required will vary from oscillator to oscillator, depending on the closest distance from the reference period to a period obtainable by the individual oscillator.
 This is a source of jitter that was not present in the clocked \ac{DCO} used in \ac{ADPLL} 1, as all period steps were equal.
@@
 \acs{ADPLL} Design 3 has no dependence on the clock provided by the \ac{FPGA}'s clock management utility, as both the \ac{RO} and the \ac{TDC} are designed using inverters, while a Bang-Bang detector replaces the \ac{FPGA} clock reliant \ac{FSM} of the two previous designs.
-This introduces a second source of implementation based variation between \acp{ADPLL}, however, this has significantly lesser impact on the frequency range of the oscillator and no impact on the centre frequencies.
-The lack of \iac{FPGA} clocked element results means this design will have the closest behaviour to that of \iac{ASIC} based \ac{ADPLL} implementation, although as the level of variation due to implementation is much larger on \iac{FPGA}, the impact of non-idealities will be exaggerated.
+This introduces a second source of implementation based variation between \acp{ADPLL}, however, this has significantly less impact on the frequency range of the oscillator and no impact at all on the centre frequencies.
+The lack of \iac{FPGA} clocked element means this design will have the closest behaviour to that of \iac{ASIC} based \ac{ADPLL} implementation, although as the level of variation due to implementation is much larger on \iac{FPGA}, the impact of non-idealities will be exaggerated.
 %same interface
-In order to allow quick exchange of \acp{PD} or other modules, the module in which this \ac{ADPLL} is implemented is still retains the \ac{FPGA} clock as an input.
+In order to allow quick exchange of \acp{PD} or other modules, the module in which this \ac{ADPLL} is implemented still retains the \ac{FPGA} clock as an input.
 In fact, this module shares its implementation with that of \ac{ADPLL} 2, as while the \ac{PD} designs are drastically different, the interface is identical.
+This allows for the changing of \ac{PD} purely by changing the module type in the instance statement.
 The module implementing this \ac{ADPLL} can be found in \textsc{NetworkRingADPLL.v}, attached in Appendix \ref{adx:code} Listing \ref{lst:network_ring_adpll}.
@@
 Once the measurement interval has completed, the \texttt{done} signal is used to reset the edge detectors asynchronously.
-It is possible that an edge may occur within the time taken to reset the counters, and thus the detector incorrectly gauge the start of a measurement interval, however this can only occur in exceptional circumstances in the acquisition period.
-Therefore it is possible to ignore this scenario, as the only impact will be an increased time taken to acquire a lock.
+It is possible that an edge may occur within the time taken to reset the counters, and thus making the detector incorrectly gauge the start of a measurement interval.
+However, this can only occur for exceptionally large phase differences, which can only occur in the acquisition period.
+Therefore, it is possible to ignore this scenario, as the only impact will be an increased time taken to acquire a lock.
 \begin{figure}[h]
@@
 A two signal interface controls the \ac{TDC}, a \texttt{count} that enables the \ac{TDC} and a \texttt{done} signal that signals the termination of the interval.
-A single logic gate is required to generate each signal, as ORing the two edge detector outputs will produce logic $1$ when the interval has begun while ANDing these signals will indicate the end of the interval when both edges have occurred.
+A single logic gate is required to generate each signal, as ORing the two edge detector outputs will produce logic $1$ when the interval has begun, while ANDing these signals will indicate the end of the interval when both edges have occurred.
 The \texttt{done} signal is also used to clock the flip-flop acting as a buffer for the sign.
@@
 As with the \ac{RO}, inverters will be used to provide the delay between each tap, and thus the resolution of the \ac{TDL} and the length of each individual delay is based on the layout chosen by the router in Vivado.
-Due to this, the exact resolution of the phase detector is unknown, although the delay through inverters used in Vivado simulations and the average delay computed based on \ac{RO} centre frequencies can be used to estimate that of the phase detectors.
+Due to this, the exact resolution of the phase detector is unknown, although the delay through inverters used in Vivado simulations, and the average delay computed based on \ac{RO} centre frequencies, can be used to estimate that of the phase detectors.
 Based these figures for the the propagation delay due to an inverter, $\tau_{inv}$ can be estimated to be in the region of $300~\si{\pico\second}$.
 As each tap consists of an inverter pair, the delay through the tap, $\tau_{tap} \approx 600~\si{\pico\second}$.
-Characterisation of the \ac{TDL}, and thus measurement of the delays in each \ac{TDL}, is technically possible, however, this is an time consuming process as many \acp{TDL} would need to be characterised in order to accurately compute the average delay.
+Characterisation of the \ac{TDL}, through measurement of the delays steps in each \ac{TDL}, is technically possible, however, this is an time consuming process as many \acp{TDL} would need to be characterised in order to accurately compute the average delay.
 
@@
 If the range is made symmetrical, 15 taps are required to fill a signal of this size.
-From the diagram it can be seen that unless edges are detected on each signal at exactly the same instant, it is almost impossible to measure a zero phase difference between reference and generated signals.
+From the diagram, it can be seen that unless edges are detected on each signal at exactly the same instant, it is almost impossible to measure a zero phase difference between reference and generated signals.
 This is an intentional decision, \cite{idkwhattocite}.
@@
 The operation of the \ac{TDL} is as follows: A delayed version of the \texttt{count} signal acts as the clock for flip-flops forming each tap, which are rising edge activated.
-These flip-flops have a constant ``1'' at their $D$ inputs, so as the signal propagates through the delay line, each flip-flop with propagate a logic ``1'' to their corresponding \texttt{error\_tap[i]} signal.
-When the measurement interval has completed, the \texttt{count} signal clocks the 15 bit wide \texttt{error\_tap} signal into \texttt{error\_tap\_buff} which act as the output buffers of the \ac{PD}, holding the phase error constant between measurement intervals.
+These flip-flops have a constant ``1'' at their $D$ inputs, so as the signal propagates through the delay line, each flip-flop will assign a logic ``1'' to their corresponding \texttt{error\_tap[i]} signal.
+When the measurement interval has completed, the \texttt{count} signal clocks the 15 bit wide \texttt{error\_tap} signal into \texttt{error\_tap\_buff}.
+These flip-flops act as the output buffers of the \ac{PD}, holding the phase error constant between measurement intervals.
 As with the flip-flops performing edge detection, the flip-flops comprising the taps can only be used once before requiring a reset, which is again carried out using the \texttt{done} signal.
-As flip-flops have a hold time, the duration after the clock edge in which the signal applied at the $D$ input must not change to ensure valid data, an inverter is used to delay the reset to avoid this issue.
+As flip-flops have a hold time, the duration after the clock edge in which the signal applied at the $D$ input must not change to ensure valid data, an inverter is used to delay the reset to avoid a violation.
 
@@
 This is somewhat difficult to implement, and a simpler solution exists for this problem.
-In a \texttt{generate} statement, a variable size loop is used to count the number of nonzero bits as it would be done in C.
+In a \texttt{generate} statement, a variable size loop is used to count the number of non-zero bits as it would be done in C.
 At compile time this loop is converted into lookup tables, similarly to the switch statement.
@@
 Compared to the oscillators suggested earlier, \ac{ADPLL} 3 achieves a significantly finer detection resolution, at the expense of significant variation in the detection step.
-This can be viewed as advantageous however, as the \ac{ADPLL} exhibits an exaggerated version of the implementation and manufacturing defect based variability present in \iac{ASIC}.
+However, depending on the intended use case, this can be viewed as advantageous, as the \ac{ADPLL} exhibits an exaggerated version of the implementation and manufacturing defect based variability present in \iac{ASIC}.
 It is this type of \ac{ADPLL} that both the \acs{UCD} and Sorbonne research teams are using to examine potential designs and test theoretical results.
@@
 \section{Measurement Setup}\label{section:measurement_setup}
-For the following sections in which the performance of either individual \acp{ADPLL} or \ac{ADPLL} networks the same measurement setup was used.
+For the following sections in which the performance of either individual \acp{ADPLL} or \ac{ADPLL} networks was tested, the same measurement setup was used.
 The \ac{FPGA} used was a Xilinx Artix-7 \acl*{Nexys} on a Digilent \acs*{Nexys} board.
@@
 Measurements were performed using an Agilent MSO7054A, a four channel, $4~\si[per-mode=symbol]{\giga\sample\per\second}$ oscilloscope.
-Signals were extracted over the regular \ac{PMOD} headers as the board does not provide outputs better suited to the measurement of signals that are available to the user, such as SMA or BNC connectors, and the \ac{PMOD} header for the analog-to-digital converter, which has lower impedance traces, is not available for the designer to re-configure.
+Signals were extracted over the regular \ac{PMOD} headers, as the board does not provide outputs better suited to the measurement of signals that are available to the user, such as SMA or BNC connectors.
+Unfortunately, the \ac{PMOD} header for the analog-to-digital converter, which has lower impedance traces, is not available for the designer to re-configure.
 Data was acquired with the oscilloscope in quad channel mode at a sampling rate of $2\si[per-mode=symbol]{\giga\sample\per\second}$.
-One of the four channels was consumed by the external reference, leaving three free for signal measurements, in most cases the generated output of three \acp{ADPLL} were connected.
+One of the four channels was consumed by the external reference, leaving three free for signal measurements.
+In most cases the generated output of three \acp{ADPLL} were connected.
 $200\si{\micro\second}$ long captures were saved to an Agilent specified binary format which could be opened in Matlab for analysis.
+Jitter was calculated using the \ac{C2C} method.
 \begin{figure}[h]
@@
 \section{\acs{ADPLL} Characterisation}
-Before combining the \acp{ADPLL} in a network it was important to analyse the designs locked to an ideal reference, in order to establish a baseline performance.
+Before combining the \acp{ADPLL} in a network, it was important to analyse the designs locked to an ideal reference, in order to establish a baseline performance.
 These tests were carried out with the \acp{ADPLL} locked to an external reference at $5~\si{\mega\hertz}$.
-As \ac{ADPLL} 2 \& 3 have significant variance in the step sizes, for these oscillators range and step based measurements were obtained by measuring a number of oscillators, and typical values will be shown.
-Jitter was calculated based on the performance on nine individual \acp{ADPLL} all locked simultaneously to the $5~\si{\mega\hertz}$ external reference.
-While jitter and skew are the key indicators of \ac{ADPLL} network performance, when it comes to the individual \acp{ADPLL} themselves it is important to check lock and capture ranges also, to check that the implemented designs match with their intended values.
+As \ac{ADPLL} 2 \& 3 have significant variance in the step sizes, for these range and step based measurements were obtained by measuring a number of oscillators, and typical values will be shown.
+Jitter was calculated based on the performance of nine individual \acp{ADPLL} all locked simultaneously to the $5~\si{\mega\hertz}$ external reference.
+While jitter and skew are the key indicators of \ac{ADPLL} network performance, when it comes to the individual \acp{ADPLL} designs, it is important to check lock and capture ranges also, to check that the implementions match with their intended values.
 The lock and capture ranges are given based on typical values, whereas jitter and skew are the median values from the \acp{ADPLL} tested.
 
-The \ref{table:characterisation} contains the results of this analysis.
+Table \ref{table:characterisation} contains the results of this analysis.
 The first thing to note is that to the $\si{\kilo\hertz}$ the lock and capture ranges of the \acp{ADPLL} implementing \acp{RO} are identical.
@@
 This is confirmed by a cursory comparison against the values in Table \ref{table:adpll2}, where the \ac{RO} range is given as $4.571-5.518~\si{\mega\hertz}$.
-This range of $0.947~\si{\mega\hertz}$ is almost identical to those of the \ac{ADPLL} $1$ \& $3$ which confirms the \ac{RO} is the limiting factor.
+This range of $0.947~\si{\mega\hertz}$ is almost identical to those of the \ac{ADPLL} $1$ \& $3$, which confirms the \ac{RO} is the limiting factor.
 In \ac{ADPLL} 1 however, the lock and capture ranges show some difference, with the lock range being slightly wider.
-The lower end of both ranges is identical, as rather than the instant loss of lock seen at all other thresholds, the low end of both lock and capture ranges saw a more gradual degradation, and $1.08~\si{\mega\hertz}$ is the point that the lock was observed to be lost.
+The lower end of both ranges is identical, as rather than the instant loss of lock seen at all other thresholds, the low end of both lock and capture ranges saw a more gradual degradation, with $1.08~\si{\mega\hertz}$ the point that the lock was observed to be lost.
 
@@
 
-Looking then at the jitter and skew data, unsurprisingly the entirely \ac{FPGA} clocked \ac{ADPLL} with its larger period step is has the worst jitter, by a significant amount.
+Looking then at the jitter and skew data, unsurprisingly the entirely \ac{FPGA} clocked \ac{ADPLL} with its larger period step has the worst jitter, by a significant amount.
 The maximum \ac{TIE} is also correspondingly worse, which unless significant skew was present in the other \ac{ADPLL} designs is also to be expected.
@@
 Somewhat surprisingly there is no apparent benefit in terms of jitter reduction due to the better delay resolution of the inverter based \ac{TDL}.
-This may be down to either the increase in jitter from a second source of variation, or a less favourable fractional-N synthesis due to the distribution of periods.
+This may be attributable to either the increase in jitter from a second source of variation, or a less favourable fractional-N synthesis due to the distribution of periods.
+Later results, during network tested, will show significant differences in jitter in both \ac{ADPLL} Design 2 \& 3, however, Design 3 has a higher median jitter value implying both the extra variation from the \ac{PD} and the less favourable synthesis play a part.
 
@@
 As the \acp{ADPLL} were designed with network usage in mind, the \ac{HDL} aspect of their integration into a network is straightforward.
-For \ac{ADPLL} 1, the oscillators need only be wired together, but for Design 2 \& 3 which are reliant on inverters require some care to ensure their lock ranges overlap sufficiently.
+For \ac{ADPLL} 1, the oscillators need only be wired together, but for Design 2 \& 3, reliant on inverters, require some care to ensure their lock ranges overlap sufficiently.
 This is a time consuming task, as a number of iterations may be required, even when locking down the cells of lined up \acp{RO}, as the nets forming the connections may change.
-It is recommended only to carry out alignment when all other aspects of the design are unlikely to change and to save the bitstream generated.
-The projected time requirement in aligning \acp{RO} meant that the creation of a 4x4 network was abandoned.
+It is recommended only to carry out alignment when all other aspects of the design are unlikely to change, and to save the bitstream generated.
 \ac{FPGA} clocked \acp{ADPLL} are more suited to larger networks as their centre point can be determined with ease.
 
-The network operates using an external reference which is input over \iac{PMOD} header that is shared with a number of output signals, however it could be configured to use an internal reference generated by either the \ac{FPGA} clock management utility or \iac{DCO} of the designer's choice.
-
-In order to increase the flexibility of the network, the switches on the \acs*{Nexys} are used to change both the proportional and integral gains and the configuration of the network at runtime.
-The reconfigurability aspect of the error combiner allows for the network to operate in either uni- or bi-directional mode, or as a number of individual \aclp{PLL} (\acsp{PLL}) all connected to the external reference.
-This is particularly advantageous for the inverter based designs, as multiple combinations can be measured without reimplementing, and thus changing the layout of either oscillator or \acl{PD}.
+The network operates using an external reference which is input over \iac{PMOD} header, shared with a number of output signals, however it could be configured to use an internal reference generated by either the \ac{FPGA} clock management utility or \iac{DCO} of the designer's choice.
+In order to increase the flexibility of the network, the switches on the \acs*{Nexys} are used to change both the \ac{LF} gains and the configuration of the network at runtime.
+The reconfigurability aspect of the error combiner allows for the network to operate in either uni- or bi-directional mode, or as a number of individual \aclp{PLL} (\acsp{PLL}), all connected to the external reference.
+This is particularly advantageous for the inverter based designs, as multiple combinations can be measured without reimplementing, thus avoiding changing the layout of either oscillator or \acl{PD}.
 
 Two different network sizes were implemented, a 2x2 and a 3x3 grid respectively.
-The 2x2 network was implemented first, and then this design extended to a 3x3 once measurements had been carried out on and any errors made had been corrected.
+The 2x2 network was implemented first, and then this design extended to a 3x3 once measurements had been carried out, and any errors made had been corrected.
 In later sections matrix indexing will be used to refer to each oscillator, with \ac{ADPLL}$_{1,1}$ closest to the reference, and higher numbers in either index getting progressively further from the reference, as in Figure \ref{fig:network_mine}.
-A 3x3 network was the minimum goal, with the aforementioned but eventually not implemented 4x4 a stretch-goal if time permitted.
 A 3x3 network set as the primary goal, as it is the minimum size network in which there exists \iac{ADPLL} with four neighbours.
-As will be explained in Chapter \ref{chap:5}, \ac{ADPLL} Design 1 was only implemented in a 2x2 configuration while both other \acp{ADPLL} were tested in a 3x3 configuration.
-
+A 4x4 network a bonus if time permitted, however, due to the projected time taken to align the 16 \acp{RO} required, this plan was dropped.
+All three \ac{ADPLL} designs were tested in a 2x3 and 3x3 configuration during the course of the project.
 \begin{figure}[h]
@@
 \end{figure}
-
-%TODO why use each PLL

### 2019-04-23 matlab/a1_3.m

@@ -1,3 +1,4 @@
+clearvars
 num_files = 15;
 std_period = zeros(4,num_files);
 TIE_mean = zeros(4,num_files);
@@ -13,7 +14,7 @@ for inc = 1:num_files
     %filename = char("conor_test/settings1_" + num2str(inc) + ".bin");
     %filename = char("scope_41.bin");
         
-    path3       = "~/gdrive/measurements/monday2/adpll_1b" + num2str(inc) + ".bin";
+    path3       = "~/gdrive/measurements/tuesday_last/adpll_des3_2x2_" + num2str(inc) + ".bin";
 %     path3       = "~/gdrive/measurements/tuesday/net3x3_clocked_pd_5m18_" + num2str(inc) + ".bin";
 %     path3       = "~/gdrive/measurements/wednesday/net3x3_5m00_kp_100_ki_1_" + num2str(inc) + ".bin";
 %     path3       = "~/gdrive/measurements/monday/design2b_div" + num2str(div_ratio) + "_" + num2str(inc) + ".bin";    

### 2019-04-24 thesis/Background.tex
DONEgit add -u!

@@
 	\includegraphics[width=0.6\textwidth]{../simple_pi}
-	\caption[Basic \ac{PI} controller architecture]{Basic \ac{PI} controller architecture.}
+	\caption[Basic \acs{PI} controller architecture]{Basic \acs{PI} controller architecture.}
 	\label{fig:my_simple_pi}
 \end{figure}
-The \acl{LF} in \iac{ADPLL} can be implemented as \iac{PI} controller, as only a low-pass filter is required, such as that in Figure \ref{fig:my_simple_pi}.
+The \acl{LF} in \iac{ADPLL} can be implemented as \iacf{PI} controller, as only a low-pass filter is required, such as that in Figure \ref{fig:my_simple_pi}.
 In the case of a node in \iac{ADPLL} network the input of this filter is a weighted average of the phase difference relative to the neighbouring local clocking areas.
@@
 \section{\acs{ADPLL} Performance Characterisation}
-As already stated, the goal of a clock distribution system is to synchronise clocking events in all areas of the chip.
+The goal of a clock distribution system is to synchronise clocking events in all areas of the chip.
 The effectiveness of this synchronisation is characterised by two main metrics, jitter and skew, which together describe the distribution of clocking events.
@@
 As the previous edge of the generated signal is used as the reference, and thereby the delay is the period of the signal, the mean value is no longer the skew, but instead can be used to compute the centre frequency over the time interval.
-For a more informative measurement the \ac{TIE} can be computed, which takes into account another signal as the reference.
-\ac{TIE} is calculated by comparing the delay between the signal in question and a reference that is treated as ideal.
+For a more informative measurement, the \ac{TIE} can be computed, which takes into account another signal as the reference.
+\ac{TIE} compares the delay between the signal in question and a reference that is treated as ideal.
 The mean value gives the relative skew between the two signals and, once more, the standard deviation of the measurements gives the jitter with respect to this reference.
 
-Both above forms of measurement assume a large number of sequential measurements, however, there are other ways that jitter and skew can be calculated \cite{AN10007}.
-Phase jitter is an important characteristic for communications systems as it can be used to calculate phase noise, important to ensure spurious emissions are within legal requirements.
+There are a number of other ways that jitter and skew can be calculated \cite{AN10007}.
+Phase jitter is important for communications systems, for the calcuation of phase noise, used to ensure spurious emissions are within legal requirements.
 Long term, or accumulated, jitter represents the cumulative effect of jitter on the signal over several cycles.
@@
 For \iac{PLL} network, with an appropriately designed filter, the long term jitter should be zero so long as the reference signal is stable.
-%TODO ask Brian about this one

### 2019-04-24 thesis/Conclusion.tex
DONEgit add -u!

--- 2019-04-20 thesis/Conclusion.tex
@@
 \section{Conclusions}
+\acuse{FPGA,ADPLL,RO,TDL,DCO,ASIC,PFD,EDA}
+The objective of this project was to create \iacs{FPGA} based platform that could be used to prototype and model \acs{ADPLL} networks in a cost and time effective manner.
+The main effort of the project was to analyse the ways in which the components of \iacs{ADPLL} network could be designed, and then selecting a number of the most appropriate designs from this pool.
+A number of \acs{ADPLL} designs were then created using the selected components in different combinations for the purposes of comparison.
+Once these \acs{ADPLL} designs had been validated, they were implemented in networks of a 2x2 and 3x3 Cartesian grid topology.
+Measurements and performance benchmarks were then carried out on these \acs{ADPLL} networks, and the results used to verify claims made regarding the intended use cases of the individual \acs{ADPLL} designs.
+Finally, various small modifications to the \acs{ADPLL} designs were made, and their impact on system performance identified.
+
+The project resulted in a platform comprising three \acs{ADPLL} designs suited for different use cases.
+As evidenced by the investigation into the impact of minor variations, this system fulfils the objective that it be suitable for the testing of architectural changes.
+Realistic, though exaggerated, variation due to implementation makes the inverter based designs ideal for the testing of new control schemes or performance enhancing modules.
+Simultaneously, this makes designs based on \acsp{RO} ideal for the confirmation of either theoretical or simulated results, in a platform that suffers from realistic implementation based variation, and non-ideal effects, such as power supply noise, that are difficult to simulate well.
+The presence of \acs{FPGA} clocked \acsp{DCO} and \acsp{PFD} makes the emulation of mixed-signal circuits, at a scaled down frequency, for the purposes of design verification possible.
+
+As the platform is designed with extensibility in mind, rapid changes in the design are possible, with the lengthy and resource consuming tape-out process that would otherwise be required for a test in hardware, cut down to the time taken for the \ac{EDA} tool of choice to generate a configuration file.
+Whereas tape-out may be a costly process, for a figure in the $100$s of Euro, an \acs{FPGA} can be acquired that will bridge the gap between software simulations and implementation in custom silicon, by providing a hardware-based simulation, modelling and validation platform.
+
 \section{Suggested Future Work}
 \subsection{\acl{TDL} Characterisation}
-\acreset{TDL}
 As previously mentioned, characterisation of the \ac{TDL} was not carried out due to the time it would have taken to perform the required measurements and tests.
@@
 Off-chip is unfortunately a requirement, as the only delays available of the order required to test the step sizes are the variable delay of the inverters themselves.
-A resistor \& capacitor delay circuit can be used to perform these test, and the output of the \ac{PFD} analysed on a mixed-signal oscilloscope alongside the delayed and reference waveforms.
+A resistor \& capacitor delay circuit can be used to perform these tests, and the output of the \ac{PFD} analysed on a mixed-signal oscilloscope, alongside the delayed and reference waveforms.
 The use of an off-chip delay element necessitates careful construction/cailbration of the measurement environment to eliminate the delay stemming from the signal path through input/output buffers as it leaves the chip.
@@
 \subsection{Increased Network Size}
-A 4x4 network was a stretch-goal of this project, depending on the time remaining after once the minumim goal of a 3x3 network had been reached.
-Unfortuantely due to problems with the clocked phase detector which resulted in mode-locking, incorrect \ac{RO} inverter selection logic and the projected time requirement to ensure alignment of each \ac{RO} this was not achieved.
+If timed had permitted, 4x4 network was to be implemented as part of this project.
+Unfortuantely due to problems with the clocked phase detector  (which resulted in mode-locking), incorrect \ac{RO} inverter selection logic and the projected time requirement to ensure alignment of 16 \acp{RO} this was not achieved.
 A 4x4 network would better explain some dynamics, such as the skew growth as the \ac{ADPLL} indices increase.
-However.
-progressing beyond a 3x3 network means that less than $20\%$ of \ac{ADPLL} output signals can be examined at any given point in time.
-As such, it may be useful to develop a more advanced measurement set-up that can either capture more signals simultaneously, or implement a measurement system that makes more efficient use of the available probes.
+However, progressing beyond a 3x3 network means that less than $20\%$ of \ac{ADPLL} output signals can be examined at any one point in time.
+Therefore, it may be useful to develop a more advanced measurement set-up, that can either capture more signals simultaneously, or implement a measurement system that makes more efficient use of the available probes.
 
 \subsection{\acs{FPGA} Clocked, Linear Period \acs{DCO}}
-\acreset{FPGA,DCO}
-While this design of \ac{DCO} was proposed earlier in this thesis, it was not implemented in any \ac{ADPLL}s, as it was only realised that the output waveform did not matter, just the location of the rising edges was not made until later in the project, once the pitfall encountered in the clocked \ac{PFD} design was resolved.
-Until this point the non-square wave output had eliminated the design from consideration.
-However, as in an \ac{ASIC} linear period designs are easier to implement, it would be ideal to have the benefit of frequency scaled hardware simulations with a \ac{DCO} that more closely relates the the eventual system.
+While this type of \ac{DCO} was proposed earlier in this thesis, it was not implemented in any \acp{ADPLL}.
+At the time the \acp{ADPLL} were designed, a square wave output was required, something this design cannot accomplish.
+It was not until later in the project, once the pitfall encountered in the clocked \ac{PFD} design was resolved, that a non square wave signal would have been possible.
+Until this point, the output waveform had eliminated the design from consideration.
+However, as in \iac{ASIC}, linear period designs are easier to implement.
+It would be ideal to have the benefit of frequency scaled hardware simulations with \iac{DCO} that more closely relates to the eventual system.
 
@@
 This suggestion is directed at making the user experience simpler.
-As implemented, networks based on \ac{ADPLL}s containing \ac{RO}s are implemented in separate files to their \ac{FPGA} clocked \ac{DCO} based counterparts as are the networks of difference sizes.
-At a network size of 2x2 this was not a problem, but at the scale of 3x3 network the modules became somewhat cumbersome to work with, with a large number of interconnects.
-For 3x3 and larger networks it would be expedient to create the network with a \texttt{generate} statement, as this would reduce the number of lines of code and the amount of named interconnects required.
+As implemented, networks based on \ac{ADPLL}s containing \ac{RO}s are implemented in separate files to their \ac{FPGA} clocked \ac{DCO} based counterparts, as are the networks of difference sizes.
+At a network size of 2x2, this was not a problem, but with a 3x3 network, the modules became somewhat cumbersome to work with.
+This is due to the large number of interconnects.
+For 3x3 and larger networks, it would be expedient to create the network with a \texttt{generate} statement, as this would reduce the number of lines of code and the number of individually named interconnects required.
 Instead the user could set a number of configuration parameters and the network would be created accordingly.
 This could even allow for the use of different \ac{ADPLL} types in the same network with ease.
-Unfortunately the tasks of floorplanning and oscillator alignment would still be left to the user to perform.
-
-%TODO cahracterise tdl
+Unfortunately, the tasks of floorplanning and oscillator alignment would still be left to the user to perform.

### 2019-04-24 thesis/FPGABasedADPLLs.tex
DONEgit add -u!

@@
 \section{Chapter Overview}
-The first step in creating \iac{FPGA} based network of \acp{ADPLL} is the design of the \ac{ADPLL} itself, which will be addressed in this chapter.
+The first step in creating \iac{FPGA} based network of \acfp{ADPLL} is the design of the \ac{ADPLL} itself, which will be addressed in this chapter.
 The nature of \iac{FPGA} necessitates compromises in the design of a certain blocks, limiting direct transfer from \ac{ASIC} designs.
@@
 	\includegraphics[width=\textwidth]{../network_adpll}
-	\caption{\ac{ADPLL} designed for use in a Cartesian grid network.}
+	\caption{\acs{ADPLL} designed for use in a Cartesian grid network.}
 	\label{fig:network_adpll}
@@
 	\subfloat[Single Multiplier FIR Filter]{{\includegraphics[width=0.6\textwidth]{../FIR_2} }}%
-	\caption[\ac{FIR} \ac{PI} loop filter RTL diagrams]{\ac{FIR} \ac{PI} loop filter RTL diagrams.}
+	\caption[\acs{FIR} \acs{PI} loop filter RTL diagrams]{\acs{FIR} \acs{PI} loop filter RTL diagrams.}
 	\label{fig:fir_pi}
@@
 	\includegraphics[width=0.6\textwidth]{../simple_pi.pdf}
-	\caption[Basic \ac{IIR} \ac{PI} loop filter]{Basic \ac{IIR} \ac{PI} loop filter.}
+	\caption[Basic \acs{IIR} \acs{PI} loop filter]{Basic \acs{IIR} \acs{PI} loop filter.}
 	\label{fig:iir_pi}

### 2019-04-24 thesis/TestingAnalysis.tex
partial corrections
citation and tables
DONEgit add -u!

@@
 \section{Chapter Overview}
-This chapter contains details of the measurements carried out on networks of various sizes and investigation into the impact of a number of decisions made in the design process.
-Explanations will be given or analysis made of the results, which will be given in the form of of either plots or tables, depending on which allows for more concise presentation of the data.
-Both 2x2 and 3x3 network performance analysis was performed with no input delay to the \ac{LF}, which Section \ref{subsection:lf_delay} will show to have a significant performance impact.
-All tests were carried out, unless otherwise stated, at a centre frequency of $5~\si{\mega\hertz}$.
+This chapter contains details of the measurements carried out on networks of various sizes in addition to investigation into the impact of a number of decisions made in the design process.
+Explanations will be given, or analysis made, of the results, which will be either in the form of plots or tables, depending on which allows for more concise presentation of the data.
+Both 2x2 and 3x3 network performance analysis was performed with no input delay to the \ac{LF}, which Section \ref{subsection:lf_delay} will show to have had a significant performance impact.
+Unless otherwise stated, all tests were carried out at a centre frequency of $5~\si{\mega\hertz}$.
 
 \section{2x2 Network Performance Comparison}
-In order to investigate the impact of the \ac{ADPLL} network on performance, and conversely the impact of each individual \ac{ADPLL} design on the network, switches were used to alter the configuration of the network without requiring re-implementation and a number of captures made at each setting.
-The three configurations were: each \ac{ADPLL} individually locked to the reference, the network in uni-directional mode and bi-directional mode.
-The centre frequency or gains were not changed between measurements in order to obtain a good comparison.
+In order to investigate the impact of the \ac{ADPLL} network on performance, and conversely the impact of each individual \ac{ADPLL} design on the network, switches were used to alter the configuration of the network without requiring re-implementation.
+A number of captures were made at each setting.
+The three configurations were: each \ac{ADPLL} individually locked to the reference, the network in uni-directional mode and in bi-directional mode.
+In order to obtain a good comparison, the \ac{LF} gains were not changed between measurements.
 For \ac{ADPLL} 1 the gains were: $k_p = \frac{1}{32}$ \& $k_i = \frac{1}{256}$.
 For \ac{ADPLL} 2 and 3 they were: $k_p = \frac{1}{16}$ \& $k_i = \frac{1}{64}$.
-The
-
-There are two main trends worth highlighting here, with the first of these being as the feedback network becomes more complicated a large degree of skew is introduced, with the magnitude becoming larger in the \acp{ADPLL} further from the reference.
-This would appear to be attributable to a combination of the time taken for a change in the reference to propagate to the \ac{ADPLL} node in question, and the fineness of period and detection resolutions.
-This second observation follows from the significant reduction in the skew growth in Designs 2 \& 3 which have finer resolutions.
-
-The other trend, is the variation in \ac{C2C} jitter across the \ac{ADPLL} designs, with coarseness of the \ac{FPGA} clocked \ac{DCO} becoming somewhat obvious.
-As free \aclp{PLL} (\acsp{PLL}), the clocked \ac{DCO} has almost identical \ac{C2C} jitter and maximum \ac{TIE}, implying that only two period steps are used the majority of the time.
-As these steps are somewhat further apart than that of the \ac{RO}, at $3.875~\si{\nano\second}$ apart as opposed to an implementation dependant value in the region of $1.1765~\si{\nano\second}$, a much greater jitter is observed.
-This would confirm the belief that the \ac{FPGA} clocked designs are better suited to operational frequencies where the generated clock period to \ac{FPGA} clock period ratio is significantly lower than the $\frac{3.875~\si{\nano\second}}{200~\si{\nano\second}} = \frac{1}{51.6}$ in this design.
-Comparing the \ac{C2C} jitter of \ac{ADPLL} 1 \& 2, where the main difference is the choice of \ac{PFD}, the approximate $\frac{1.6~\si{\nano\second}}{0.50~\si{\nano\second}} \approx 3.2$ ratio of observed jitters is almost identical to the $\frac{3.875~\si{\nano\second}}{1.1765~\si{\nano\second}} \approx 3.29$ ratio of the period steps at $5~\si{\mega\hertz}$, implying the \ac{PFD} is at fault for this difference.
-
-The other notable difference in \ac{C2C} jitter, is that the use of an inverter based \ac{TDC}, and its accompanying variation due to implementation differences, has the expected result of additional \ac{C2C} jitter, however, this impact is signfiicantly less than the that of changes in the \ac{RO} period step, and the inverter based \ac{TDC} in combination with the \ac{FPGA} clocked \ac{DCO} would still suffer the same problems as it does in \ac{ADPLL} Design 1 with the clocked \ac{PFD}.
-
-%TODO additional comments with new data?
+\ac{ADPLL} requires a smaller gain due to the greater frequency sensitivity.
 
@@
                 \multicolumn{2}{|l|}{\ac{ADPLL} Design 2}&-&-&-&-&-&-&-&-&-\T\\
-                \multicolumn{2}{|r|}{Free PLLs} &0.56222 &0.49224 &0.64124    &5.3039 &7.8626 &5.4634    &&& \T\\
-                \multicolumn{2}{|r|}{Uni-dir.}  &0.55040 &0.51284 &0.62796    &5.5312 &7.5359 &7.9697    &&& \T\\
-                \multicolumn{2}{|r|}{Bi-dir.}   &0.57056 &0.50403 &0.61644    &11.515 &15.943 &17.428    &&& \T\\
+                \multicolumn{2}{|r|}{Free PLLs} &0.66275 &0.62108 &0.63211    &5.7432 &5.4399 &5.2317    &0.00768 &-0.1681 &0.5383  \T\\
+                \multicolumn{2}{|r|}{Uni-dir.}  &0.65990 &0.63995 &0.67278    &5.6926 &6.9422 &8.9609    &0.00663 &-0.3453 &0.76741 \T\\
+                \multicolumn{2}{|r|}{Bi-dir.}   &0.66308 &0.66233 &0.67583    &12.222 &16.719 &19.176    &5.0799  & 7.2752 &8.7174  \T\\
                 \hline
@@
         \end{footnotesize}
-        \caption{2x2 Network performance Comparison.}
+        \caption{2x2 Network Performance Comparison.}
         \label{table:2x2perf}
@@
 \end{table}
-%\FloatBarrier
+There are two main trends worth highlighting here, the first of these being as the feedback network becomes more complicated a large degree of skew is introduced, with the magnitude becoming larger in the \acp{ADPLL} further from the reference.
+This would appear to be attributable to a combination of the time taken for a change in the reference to propagate to the \ac{ADPLL} node in question, and the fineness of period and detection resolutions.
+This second observation follows from the significant reduction in the skew growth in Designs 2 \& 3 which have finer resolutions.
+
+The other trend, is the variation in \ac{C2C} jitter across the \ac{ADPLL} designs, with coarseness of the \ac{FPGA} clocked \ac{DCO} becoming somewhat obvious.
+As free \aclp{PLL} (\acsp{PLL}), the clocked \ac{DCO} has almost identical \ac{C2C} jitter and maximum \ac{TIE}, implying that only two period steps are used the majority of the time.
+As, at $3.875~\si{\nano\second}$, these steps are somewhat further apart than that of the \ac{RO} (an implementation dependant value in the region of $1.1765~\si{\nano\second}$), a much greater jitter is observed.
+This would confirm the belief that the \ac{FPGA} clocked designs are better suited to operational frequencies where the generated clock period is multiple orders of magnitude longer than the \ac{FPGA} clock period.
+In this design, the ratio works out as $\frac{3.875~\si{\nano\second}}{200~\si{\nano\second}} = \frac{1}{51.6}$, which is insufficient.
+Comparing the \ac{C2C} jitter of \ac{ADPLL} 1 \& 2, where the main difference is the choice of \ac{RO}, the approximate $\frac{1.6~\si{\nano\second}}{0.50~\si{\nano\second}} \approx 3.2$ ratio of observed jitters is almost identical to the $\frac{3.875~\si{\nano\second}}{1.1765~\si{\nano\second}} \approx 3.29$ ratio of the period steps at $5~\si{\mega\hertz}$, implying the \ac{PFD} is at fault for this difference.
+
+The other notable difference in \ac{C2C} jitter, is that the use of an inverter based \ac{TDC}, and its accompanying variation due to implementation differences, has the expected result of additional \ac{C2C} jitter.
+However, this impact is significantly less than that of changes in the \ac{RO} period step.
+The combination of inverter based \ac{TDC} in combination with the \ac{FPGA} clocked \ac{DCO} would likely be wasted, as the \ac{DCO} would still suffer the same period distribution issues it does in \ac{ADPLL} Design 1, with the clocked \ac{PFD}.
+
+Overall, the best performer is Design 3, as although the \ac{C2C} jitter is worse that of Design 2, it exhibits the least skew in bi-directional mode.
+This would imply that period and detection resolution are the key attribute guiding network performance.
+\ac{ADPLL} 3's performance confirms it's status as the ideal choice for the testing of new control or performance improvement techniques, or comparison with theoretical models.
 
@@
 Subsequent to tests carried out using a 2x2 network, the corresponding tests were carried out using a 3x3 network in order to analyse the impact of a larger network.
-For these tests the gains were left unchanged but, as the 2x2 and 3x3 networks have different floorplans, the \acp{RO} have changed somewhat.
-This has especially manifested itself in the exceptional individual \ac{PLL} performance of Design 3.
+For these tests the gains were left unchanged but, as the 2x2 and 3x3 networks have different floorplans, the \ac{RO} layouts have changed somewhat.
+This layout change has especially manifested itself in the exceptional individual \ac{PLL} performance of Design 3.
 
-From analysis of the 2x2 network, it was evident that the best performing design was going to be \ac{ADPLL} 3, as both \ac{ADPLL} 2 \& 3 contain the clocked \ac{PFD} which appears to cause significant skew as the \ac{ADPLL} node indices increase and move further from the external reference.
-The increased network size appears to have made the lag between reference edge and the rising edges of the individual generated signals worse, which would be expected as the number of clock cycles required to respond to a change in the phase relationship between reference and the furthest node has increased.
-\ac{ADPLL} 3 suffers least from the greater complication of the network, likely as a result of its period and detector resolution being the finest of the three designs considered.
-
-As \ac{ADPLL} Design 1 suffered significantly more with skew as the node indices increased, so it is no surprise that in the larger network the problem has been amplified, meaning this design is not well suited for use in larger networks at this frequency due to poor resolution.
-This does not, however, impact the main role of this design method, which is the emulation of designs intended for use in \aclp{ASIC} (\acsp{ASIC}) as the frequency of operation will be much lower than the $5~\si{\mega\hertz}$ used in these tests, which will bring with it a finer resolution.
-%TODO fill in this after measurements
 \begin{table}[!ht]
@@
                 \cline{3-11} 
-                &&PLL 11&PLL 12&PLL 22    &PLL 11&PLL 12&PLL 22    &PLL 11&PLL 12&PLL 22\T\\
+                &&PLL 11&PLL 22&PLL 33    &PLL 11&PLL 22&PLL 33    &PLL 11&PLL 22&PLL 33 \T\\
                 \hline
                 \multicolumn{2}{|l|}{\ac{ADPLL} Design 1}&-&-&-&-&-&-&-&-&-\T\\
-                \multicolumn{2}{|r|}{Free PLLs} &&& &&& &&& \T\\
-                \multicolumn{2}{|r|}{Uni-dir.}  &&& &&& &&& \T\\
-                \multicolumn{2}{|r|}{Bi-dir.}   &&& &&& &&& \T\\
+                \multicolumn{2}{|r|}{Free PLLs} &1.8956  &1.8425  &1.7594     &5.7868 &5.8407 &6.2537    &1.8481 &1.5976 &2.3251 \T\\
+                \multicolumn{2}{|r|}{Uni-dir.}  &1.7836  &1.8206  &1.8827     &5.5863 &6.522  &9.4008    &1.7697 &1.0359 &1.5503 \T\\
+                \multicolumn{2}{|r|}{Bi-dir.}   &1.9049  &1.9205  &1.9441     &22.980 &42.771 &47.441    &17.268 &33.739 &35.989 \T\\
                 \hline
@@
         \end{footnotesize}
-        \caption{3x3 Network performance Comparison.}
+        \caption{3x3 Network Performance Comparison.}
         \label{table:3x3perf}
@@
 \end{table}
-%\FloatBarrier
+The results of the 3x3 network testing are largely similar to those from before, with the same observations applying.
+From analysis of the 2x2 network, it was evident that the best performing design was going to be \ac{ADPLL} 3, as both \ac{ADPLL} 2 \& 3 contain the clocked \ac{PFD} which appears to cause significant skew as the \acp{ADPLL} move further from the external reference.
+The increased network size appears to have made the lag between reference edge and the rising edges of the individual generated signals worse, which would be expected as the number of clock cycles required to respond to a change in the phase relationship between reference and the furthest node has increased.
+\ac{ADPLL} 3 suffers least from the greater complication of the network, likely as a result of its period and detector resolution being the finest of the three designs considered.
+
+As \ac{ADPLL} Design 1 suffered significantly more with skew as the node indices increased, it is no surprise that in the larger network the problem has been amplified, meaning this design is not well suited for use in larger networks at this frequency, due to poor resolution.
+This does not, however, impact the main role of this design method, which is the emulation of designs intended for use in \aclp{ASIC} (\acsp{ASIC}) at scaled frequencies.
+As the frequency of operation should be much lower than the $5~\si{\mega\hertz}$ used in these tests, the finer resolution, relative to the generated period, should alleviate this skew.
 
@@
 The gains chosen for use in the \acl{LF} have significant impact on the performance of the system, as the \ac{LF} governs the response of the \ac{ADPLL} to a given phase error.
-To large a proportional gain will result in overly aggressive overcompensation, thus negatively affecting the jitter.
+Too large a proportional gain will result in compensation, thus negatively affecting the jitter.
 Eventually the violence of the jitter will result in a loss of lock.
@@
 Figure \ref{fig:gain_sweep} contains a sweep across the possible range of integral gains for four different proportional gains.
-This test was carried out with a 3x3 network of \ac{ADPLL} Design 3, each locked to the external reference at $5~\si{\mega\hertz}$, and the generated waveform from the \acp{ADPLL} at locations ``11'', ``22'' and  ``33'' were captured.
-It can clearly be observed that there are two distinct regions in the graph, with jitter values of less than $1~\si{\nano\second}$ contrasting sharply with the points with jitter in the region of $10~\si{\nano\second}$.
+This test was carried out with a 3x3 network of \ac{ADPLL} Design 3, each locked to the external reference at $5~\si{\mega\hertz}$.
+Only the generated waveforms from the \acp{ADPLL} at locations ``11'', ``22'' and  ``33'' were captured, alongside the external reference.
+It can be clearly observed that there are two distinct regions in the graph, with jitter values of less than $1~\si{\nano\second}$ contrasting sharply with the section where jitter is in the region of $10~\si{\nano\second}$.
 These distinct sections of the plot represent locked and unlocked operation respectively.
@@
 
-Despite the relatively small number of proportional gains, it is clear that the relationship set out by Koskin does not hold in this system, but rather a scaled version of it.
-This difference is attributable to a difference in \ac{LF} structure.
-In his simulations the \ac{LF} design contains a register forming an input buffer, absent in the \ac{LF} used for these tests.
+Despite the relatively small number of proportional gains, it is clear that the exact relationship set out by Koskin does not hold in this system, but rather a scaled version of it does.
+This difference is attributable to a modification in the structure of the \ac{LF}.
+In his simulations, the \ac{LF} design contains a register forming an input buffer, absent in the \ac{LF} used for these tests.
 The presence of this register, clocked on the generated signal, causes the \ac{ADPLL} to react more slowly to changes in the phase difference, thus increasing the jitter.
-The removal of this register causes a change in the relationship between proportional and integral gains, a change that will be investigated further in a later section.
+This register was removed as it is not required to meet timing constraints in the \acp{ADPLL} used in this project.
+This removal allows the system to respond a cycle faster to a change in the phase relationships, and the impact of this change will be investigated further in a later section.
 
@@
 To add context to the jitter numbers produced for each oscillator previously, the variation of period steps in the \ac{RO}, and delays in the \ac{TDL} due to implementation should be examined.
-The best way to visualise this distribution is through a histogram, placing periods of in a certain range into the same bin.
+The best way to visualise this distribution is through a histogram, placing periods within in a certain range into the same bin.
 This can be visualised effectively using Matlab's \texttt{histplot}.
-In order to obtain data, three \acp{ADPLL} implementing both the \ac{FPGA} clocked and inverter based \acp{DCO} were locked to an external reference signal at $5~\si{\mega\hertz}$, and the period of each clock cycle measured.
+In order to obtain data, three instances of \acp{ADPLL} implementing each of the \ac{FPGA} clocked and inverter based \acp{DCO} were locked to an external reference signal at $5~\si{\mega\hertz}$.
+A capture was taken of each generated signal, the period of each clock cycle measured based on their rising edges.
+
 \begin{figure}[h]%
@@
 \end{figure}
-Figure \ref{fig:dists} contains the output of \texttt{histfit} for each of the three oscillators superimposed in one plot for each \ac{DCO} type.
-As the \ac{RO} has significantly finer period resolution, the period steps would be overlaid on each-other thereby making the plot difficult to read.
-In order to avoid this behaviour, the \ac{RO}'s mapping of control code to inverters to remove was altered from twice the control code to 16 times the control code.
-This results in a more pronounced separation of period steps while being indicative of the variation due to implementation.
-The plots paint a a picture of the unrealistic nature of the \ac{FPGA} clocked design in comparison to its inverter based counterpart, as each oscillator will have identical intrinsic period steps thus eliminating any jitter that maybe be seen in the system due to misalignment.
+Figure \ref{fig:dists} contains the output of \texttt{histfit} for each of the three oscillators, superimposed in one plot for each \ac{DCO} type.
+As the \ac{RO} has significantly finer period resolution, the period steps would be overlaid on each-other, thereby making the plot difficult to read.
+In order to avoid this behaviour, the \ac{RO}'s mapping of control code to inverters was altered, from twice the control code to 16 times the control code.
+This resulted in a more pronounced separation of period steps while being indicative of the variation due to implementation.
+The plots paint a a picture of the unrealistic nature of the \ac{FPGA} clocked design in comparison to its inverter based counterpart, as each \ac{DCO} has identical intrinsic period steps thus eliminating any jitter that may be seen in the system due to misalignment.
 This effects the suitability of the FPGA clocked \ac{DCO} for the improvement of simulation models, or the examination of how a novel block may impact the behaviour of the network.
@@
 f_{step} &= \frac{f_{FPGA}}{2^{13}} = 31.494~\si{\kilo\hertz}
-\end{align}
-As Table \ref{table:accum_width} shows, this is not in fact the case.
-Both measurements were carried out using identical loop filter gains, network configurations and reference signal.
-While in uni-directional mode some difference is observed in the maximum \ac{TIE}, but in all other aspects no statistically relevant changes can be seen.
+\end{align} Both the following measurements were carried out using identical loop filter gains, network configurations and reference signal.
+While in uni-directional mode some minor difference is observed in the maximum \ac{TIE}, but overall no statistically relevant changes were seen.
+
+As Table \ref{table:accum_width} shows, while the resolution may have improved, the jitter has not.
+This occurs as all signals are synchronised to the \ac{FPGA} clock, which has a $3.875~\si{\nano\second}$ period.
+This quantises the output waveform to the \ac{FGPA} clock, and sets the minimum bound on the jitter.
+Accordingly, the most suitable counter width is that which maximises the bit width, and thus achieves the best frequency resolution, with the constraint that, at the operating frequency, the a period resolution is no finer than the \ac{FPGA} clock period.
+In this case, there is no bit width that satisfies both constraints, so the latter is relaxed to get a width of $12$.
 %TODO explanation for this?
@@
 The lack of agreement between the theory and experimental results was attributed to the input delay register, or lack thereof.
-In order to provide a basis for this claim, the performance of a network in 2x2 configuration, was analysed and the results of this analysis are presented in Table \ref{table:lf_delay}.
-While the \ac{C2C} jitter is statistically similar, there is significant different to be observed in the \ac{TIE}, where an approximately $50\%$ increase can be observed, measurement for measurement.
-As the maximum \ac{TIE} is significantly impacted by the presence of any skew, as skew is the average value of this difference, it is interesting to note that the relationship is in fact the opposite.
+In order to provide a basis for this claim, the performance of a network in 2x2 configuration was analysed and the results of this analysis are presented in Table \ref{table:lf_delay}.
+While the \ac{C2C} jitter is statistically similar, there is significant difference to be observed in the \ac{TIE}, where an approximately $50\%$ increase can be observed, measurement for measurement.
+The absolute maximum \ac{TIE} is significantly impacted by the presence of any skew, as skew is the average value of this quantity, it is important to note that the skew has not meaningfully changed.
+As skew has not changed, it can be determined that jitter is the source of the increased \ac{TIE}.
+
 It is in precisely this situation that \ac{C2C} jitter as a measurement is non-ideal.
-\ac{C2C} jitter does not differentiate between the situations in which while the period variation might be the same for multiple clock cycles in the same or opposite directions.
-As the extra cycle of delay due to the input register increases the reaction time of the \ac{ADPLL} to any changes, the delay between clock edges increases.
-\ac{TIE} is, however, affected by this difference which can be clearly seen in the table.
+\ac{C2C} jitter does not differentiate between the situations in which the period variation is the same for multiple clock cycles in the same, or in opposite directions.
+\ac{C2C} jitter is therefore not a good indicator of how far apart the reference and generated signal's rising edges, whereas this is what \ac{TIE} is intended to do.
+As the delay due to the input register increases the reaction time of the \ac{ADPLL} to any changes in the reference, increases in the \ac{TIE} will be countered more slowly, thereby allowing the \ac{TIE} to grow for an extra clock cycle.
+Without a delay this is reacted to more quickly, limiting the maximum deviation.
+Both cases affect the same change in the period, just a cycle apart, so will impact the \ac{C2C} jitter identically.
 
@@
 \subsection{Impact of Frequency Divider}
-The major benefit of \iac{PLL} network over other forms of coupled oscillator based clock distribution systems, is that the coupling can be done using a signal significantly reduced in frequency, thereby reducing the power consumption of the related hardware.
-It is therefore important to ensure that the insertion of a feedback divider does not introduce any further jitter to the system.
-Table \ref{table:div} presents the results of measurements at a number of divisions levels.
+The major benefit of \iac{PLL} network over other forms of coupled oscillator based clock distribution systems is that the coupling can be done using a signal significantly reduced in frequency, thereby reducing the power consumption of the related hardware.
+It is important, therefore, to ensure that the insertion of a feedback divider does not introduce any further jitter to the system.
+Table \ref{table:div} presents the results of measurements at a number of division levels.
 As re-implemenation was performed between each measurement, there is some variation present.
@@
 
-It, however, does have a significant impact on the maximum \ac{TIE}, and notably this occurs without any meaningful change in skew.
+It does have a significant impact on the maximum \ac{TIE}, however, and notably this occurs without any meaningful change in skew.
 This confirms an increase in jitter due to the greater division levels as the source of the performance degradation.
-By examining how the dividers behave, this can be explained.
+This can be explained by examining how the dividers behave.
 When a divider is inserted into the network, the phase error is computed at a greatly reduced rate.
-This means the rate at which the \ac{LF} is updated also drops, and therefore the control code is changes $n$ times less frequently, where $n$ is the division level.
-The dividers reduce the speed at which the individual \acp{ADPLL} can react to a phase difference, while locking the resulting control code in place for a greater number of clock cycles as the division level increases.
+This means the rate at which the \ac{LF} is updated also drops, and therefore the control code changes $n$ times less frequently, where $n$ is the division level.
+The dividers reduce the speed at which the individual \acp{ADPLL} can react to a phase difference, while locking the resulting control code in place for a ever greater number of clock cycles as the division level increases.
 This greater duration may cause a control code to be applied for an excessively long number of cycles, resulting in a drift of the rising edge of the divided/generated signal away from the reference, which will manifest as \ac{TIE}.
@@
 Figure \ref{fig:todo_div} can confirm this behaviour.
-In Plot (a), as the division level increases, the number of cycles spent at a $200~\si{\nano\second}$ period decreases, which would line up with the overcorrection, thus resulting in the increased \ac{TIE} seen in Plot (b).
+In Plot (a), as the division level increases, the number of cycles spent at a $200~\si{\nano\second}$ period decreases.
+This confirms that overcorrection is occurring, thus resulting in the increased \ac{TIE} seen in Plot (b).
 \begin{figure}[h]%
@@
     \includegraphics[width=0.8\textwidth]{../dist2}
-    \caption[Period distribution at differen divider levels]{Period distribution at differen divider levels.}    
+    \caption[Period distribution at different divider levels]{Period distribution at different divider levels.}    
     \label{fig:todo_div}
@@
 As each design was introduced, the most appropriate use cases for that particular \ac{ADPLL} were suggested.
-\ac{ADPLL} networks can be used in a number of roles, from modelling system dynamics or the refinement of behavioural simulation models to the validation, verification and testing of potential \ac{ASIC} based networks (Zianbetov \& Shan) to the design of other hardware for the purposes of network performance improvement or control, however, some architectures of \ac{ADPLL} are more suited to particular roles than others.
+\ac{ADPLL} networks can be used in a number of roles: from modelling system dynamics or the refinement of behavioural simulation models to the validation, verification and testing of potential \ac{ASIC} based networks (Zianbetov \& Shan) or the design of other hardware for the purposes of network performance improvement or control.
+However, some architectures of \ac{ADPLL} are more suited to particular roles than others.
 
-As mentioned when \ac{FPGA} clocked designs were introduced, their configurability makes them particularly suited to the validation and verification of designs destined for use in \acp{ASIC}, as was the case with Zianbetov and Shan \cite{zianbetov2013phd,shan2014phd}.
+As mentioned when \ac{FPGA} clocked designs were introduced, their configurability made them particularly suited to the validation and verification of designs destined for use in \acp{ASIC}, as was the case with Zianbetov and Shan \cite{zianbetov2013phd,shan2014phd}.
 In contrast to the \ac{RO} based designs, in which configurability is more more restricted, it is possible to replicate the resolution, centre frequencies and other \ac{ADPLL} characteristics of the design requiring verification at a scaled down frequency.
-This can be carried out at large division ratios, where the \ac{FPGA} clock period is not a restricting factor, such as the $50~\si{kilo\hertz}$ centre frequency used by Zianbetov and Shan.
-In order to reduce the frequency to this extent with inverters over $30$ thousand would be required, introducing the potential for even greater variation in centre frequencies due to implementation.
-The other issue arising out of this number of inverters is the floorspace of the \ac{FPGA} that would be consumed implementing just one \ac{RO}, in the case of the Xilinx \acl{Nexys} and extrapolating based on the resource usage of the \acp{RO} implemented, a single oscillator would consume half the look-up tables available, eliminating the possibility of a network.
+This can be carried out at large division ratios, where the \ac{FPGA} clock period is not a restricting factor, such as the $50~\si{\kilo\hertz}$ centre frequency used by Zianbetov and Shan.
+In order to reduce the frequency to this extent with inverters, over $30,000$ would be required, introducing the potential for even greater variation in \ac{DCO} centre frequencies due to implementation.
+The other issue arising from this number of inverters, is the floorspace of the \ac{FPGA} that would be consumed implementing just one \ac{RO}.
+In the case of the Xilinx \acl{Nexys}, and extrapolating based on the resource usage of the $373$ inverter \acp{RO} implemented, a single \ac{RO} would consume half the look-up tables available, eliminating the possibility of a network.
 
-While the \ac{FPGA} based designs feature characteristics making them particularly suitable for use as a validation tool, their performance in a network reduces their suitability for the refinement of software based simulations, or in the analysis of new control or performance enhancement tools.
-Referring back to Table \ref{table:2x2perf}, \ac{C2C} jitter is significantly greater than the inverter based designs, and a large degree of skew is present making this design non-ideal for these purposes.
-In particular, as seen with the increase to a 3x3 network, this skew becomes an issue that limits the network size.
+While the \ac{FPGA} based designs feature characteristics which make them particularly suitable for use as a validation tool, their performance in a network reduces their suitability for the refinement of software based simulations, or in the analysis of new control or performance enhancement tools.
+Referring back to Table \ref{table:2x2perf}, not only is \ac{C2C} jitter significantly greater than the inverter based designs, a large degree of skew is also present.
+These combined make \ac{FPGA} clocked designs unsuitable for the aforementioned purposes.
+In particular, as seen with the increase to a 3x3 network, this skew becomes the limiting factor the network size.
 
@@
 While \ac{ADPLL} 2 will feature a large degree of variation due to implementation differences, these differences are only present in the \ac{RO} not the \ac{PFD}.
-As testing Designs 2 \& 3 has shown, in particular the results displayed in Table \ref{table:2x2perf}, apart from a larger skew, the jitter performance is quite similar, meaning that the extra complexity of inverter based \ac{PFD} is not required in order to achieve as accurate behaviour as can be achieved in this environment.
+As testing Designs 2 \& 3 has shown, particularly the results displayed in Table \ref{table:2x2perf}, apart from a larger skew, the jitter performance is quite similar, meaning that the extra complexity of inverter based \ac{PFD} is not required, for smaller networks, in order to achieve as accurate behaviour as can be achieved in this environment.
 The implementation based variation of the detector step manifests itself as an increase in the jitter across all network modes, as would be expected, compared to the \ac{FPGA} based design.
 
-Accordingly, if the aim of this platform is the enhancement of a model, the validation of theoretical results, or analysing the performance of new control or performance enhancement hardware the \ac{RO} based designs are idea for this purpose.
-Two members of the research team here in \acs{UCD}, Eugene Koskin and Pierre Bisiaux, have been using \iac{ADPLL} with the same architecture as Design 3 for these purposes.
+Accordingly, if the aim of this platform is the enhancement of a model, the validation of theoretical results or the analysis of the performance of new control or performance enhancement hardware, the \ac{RO} based designs are idea for this purpose.
+Two members of the research team here in \acs{UCD}, Eugene Koskin and Pierre Bisiaux, have been using \iac{ADPLL} with the same architecture as Design 3, for these very purposes.

### 2019-04-24 thesis/Background.tex
done done

@@
 Alternative designs replace the electrical lines used in the tree networks with waveguides for optical signals, with only the distribution in the local area carried out using regular wires.
-This technique presents many advantages \cite{chen2006chip}: optical clock delivery is immune to the noise sources that affect electrical clock distribution systems, consume less power and do not suffer from the electrical losses present in a regular tree system.
-However, this is a fledgling method, and sufficient research into materials and potential designs has yet to be carried out \ref{zianbetov2013phd,1648640}.
+This technique presents many advantages: optical clock delivery is immune to the noise sources that affect electrical clock distribution systems, consume less power and do not suffer from the electrical losses present in a regular tree system \cite{chen2006chip}.
+However, this is a fledgling method, and sufficient research into materials and potential designs has yet to be carried out \cite{zianbetov2013phd,chen2006chip}.
 

### 2019-04-24 thesis/FPGABasedADPLLs.tex
done done

@@
 However, those that will be used to emulate mixed-signal circuitry, such as the \ac{DCO} and \ac{PFD}, will be examined in greater detail.
-\Iac{RTL} diagram of the \ac{ADPLL} topology that will be used in the following sections is shown in Figure \ref{fig:network_adpll}.
+\Iacf{RTL} diagram of the \ac{ADPLL} topology that will be used in the following sections is shown in Figure \ref{fig:network_adpll}.
 \begin{figure}[h]
@@
 	\includegraphics[width=0.6\textwidth]{../osc1}
-	\caption{\acs{FPGA} driven, linear period \acs{DCO} \ac{RTL} Diagram.}
+	\caption{\acs{FPGA} driven, linear period \acs{DCO} \acs{RTL} Diagram.}
 	\label{fig:osc1}
@@
 		\end{tabular}
-		\caption{SR latch truth table}
+		\caption[SR Latch Truth Table]{SR Latch Truth Table.}
 		\label{table:sr_tt}
@@
 		\end{tabular}
-		\caption{Sign detector truth table}
+		\caption[Sign Detector Truth Table]{Sign Detector Truth Table}
 		\label{table:sign_tt}
@@
 
-Again, both integer and fixed-point methods of \ac{LF} design are possible, with the integer method requiring the logic combining the two paths to preserve the relationship between the proportional and integral gains.
+Again, both integer and fixed-point methods of \ac{LF} design are possible, with the integer method requiring the logic combining the two paths to preserve the relationship between proportional and integral gains.
 From the work of Koskin \textit{et al} this known to be several binary orders of magnitude \cite{koskin2018generation}.
@@
 	\includegraphics[width=0.6\textwidth]{../eugene}
-	\caption[Stability of \acl{LF} gains]{Stability of \acl{LF} gains \cite{koskin2018generation}.}
+	\caption[Oscillator jitter for various \acl{LF} gains]{Oscillator jitter for various \acl{LF} gains \cite{koskin2018generation}.}
 	\label{fig:gain_grid}

### 2019-04-24 thesis/NetworkImplemenation.tex
citation and tables
done done

@@
 As all error signals are two's complemented signed integers, the error combination process is also carried out using signed values.
-The code implementing the error combiner can be found in \textsc{ErrorCombiner.v}, attached in Appendix \ref{adx:code} Listing \ref{lst:error_combiner}.
+The code implementing the error combiner can be found in \textsc{ErrorCombiner.v}, attached in Appendix  \ref{lst:error_combiner}.
 
@@
 In the example system shown in Figure \ref{fig:integer_lf} the proportional gain can be computed as $k_p\times 2^{3-8} = \frac{k_p}{32}$ and the integral gain is $k_i\times 2^{-8} = \frac{k_i}{128}$.
-The Verilog implementation of the \ac{LF} using integer arithmetic can be found in \textsc{LoopFilter.v}, attached in Appendix \ref{adx:code} Listing \ref{lst:loop_filter}.
+The Verilog implementation of the \ac{LF} using integer arithmetic can be found in \textsc{LoopFilter.v}, attached in Appendix \ref{lst:loop_filter}.
 
@@
 The state transition diagram given as the example in Chapter \ref{chap:3} is that used to design this sign detector.
-The hardware description of this block is found in \textsc{PhaseDetector.v}, attached in Appendix \ref{adx:code} Listing \ref{lst:PhaseDetector}.
+The hardware description of this block is found in \textsc{PhaseDetector.v}, attached in Appendix \ref{lst:PhaseDetector}.
 \begin{figure}[h]
@@
 \subsubsection{\acs{ADPLL} 1 Design Summary}
-The listed components were brought together in \textsc{NetworkADPLL.v}, attached in Appendix \ref{adx:code} Listing \ref{lst:network_adpll}.
+The listed components were brought together in \textsc{NetworkADPLL.v}, attached in Appendix \ref{lst:network_adpll}.
 It is important to note that each \ac{ADPLL} contains only two \aclp{PD}, despite the fact that during network operation the need for phase comparisons to be made with up to four neighbours.
@@
 This presented a challenge, as depending on the implementation, the characteristics of the \ac{DCO} could change, most importantly the centre frequency.
-The module implementing this \ac{ADPLL} can be found in \textsc{NetworkRingADPLL.v}, attached in Appendix \ref{adx:code} Listing \ref{lst:network_ring_adpll}.
+The module implementing this \ac{ADPLL} can be found in \textsc{NetworkRingADPLL.v}, attached in Appendix \ref{lst:network_ring_adpll}.
 
@@
 
-The module implementing the \ac{RO} can be found in \textsc{RingOsc.v}, attached in Appendix \ref{adx:code} Listing \ref{lst:ring_osc}.
+The module implementing the \ac{RO} can be found in \textsc{RingOsc.v}, attached in Appendix \ref{lst:ring_osc}.
 In the module instance statement, it can be seen that a second generated output is also present.
@@
     \end{center}
-    \caption[\ac{ADPLL} Design 2 Summary]{\ac{ADPLL} Design 2 Summary.}
+    \caption[\acs{ADPLL} Design 2 Summary]{\acs{ADPLL} Design 2 Summary.}
     \label{table:adpll2}
@@
 This allows for the changing of \ac{PD} purely by changing the module type in the instance statement.
-The module implementing this \ac{ADPLL} can be found in \textsc{NetworkRingADPLL.v}, attached in Appendix \ref{adx:code} Listing \ref{lst:network_ring_adpll}.
+The module implementing this \ac{ADPLL} can be found in \textsc{NetworkRingADPLL.v}, attached in Appendix \ref{lst:network_ring_adpll}.
 
@@
 From the diagram, it can be seen that unless edges are detected on each signal at exactly the same instant, it is almost impossible to measure a zero phase difference between reference and generated signals.
-This is an intentional decision, \cite{idkwhattocite}.
+This is done to maximize the information carried by the digital signal \cite{shan2014phd}.
 \begin{figure}[h]
@@
 The now unsigned binary representation of the error magnitude can be combined then with the sign bit to form a two's complement error signal.
-The file \textsc{PhaseDetectorDL.v} in Appendix \ref{adx:code} Listing \ref{lst:phase_detector_dl} contains the Verilog implementation of this module.
+The file \textsc{PhaseDetectorDL.v} in Appendix \ref{lst:phase_detector_dl} contains the Verilog implementation of this module.
 \texttt{DONT\_TOUCH} compiler directives were used here in a number of places to ensure that the inverters and other elements that are key for timing are not removed in the synthesis process, as similarly to \ac{RO}, Vivado sees these as unneeded elements serving no purpose.

### 2019-04-24 thesis/TestingAnalysis.tex
done done

@@
 As Table \ref{table:accum_width} shows, while the resolution may have improved, the jitter has not.
-This occurs as all signals are synchronised to the \ac{FPGA} clock, which has a $3.875~\si{\nano\second}$ period.
-This quantises the output waveform to the \ac{FGPA} clock, and sets the minimum bound on the jitter.
-Accordingly, the most suitable counter width is that which maximises the bit width, and thus achieves the best frequency resolution, with the constraint that, at the operating frequency, the a period resolution is no finer than the \ac{FPGA} clock period.
+In this design, all signals are synchronised to the \ac{FPGA} clock, which has a $3.875~\si{\nano\second}$ period.
+The output waveform can only change at rising edges of the \ac{FPGA} clock, thus quantising the generated signal's period to increments of $3.875~\si{\nano\second}$.
+If the period step is shorter than the clock period, the extra resolution is lost and the \ac{FPGA} clock sets the minimum bound on the jitter.
+Accordingly, the most suitable counter width is that which maximises the number of bits, and thus achieves the best frequency resolution, with the constraint that, at the operating frequency, the period resolution is no finer than the \ac{FPGA} clock period.
 In this case, there is no bit width that satisfies both constraints, so the latter is relaxed to get a width of $12$.
-%TODO explanation for this?
+
 \begin{table}[!ht]
@@
         \end{footnotesize}
-        \caption{\ac{DCO} Accumulator Width Comparison.}
+        \caption{\acs{DCO} Accumulator Width Comparison.}
         \label{table:accum_width}

### 2019-04-24 arduino/pwm/TimerHelpers.h
Final code commit

@@ -1,207 +1,0 @@
-/*
- Timer Helpers library.
-
-Devised and written by Nick Gammon.
-Date: 21 March 2012
-Version: 1.0
-
-Licence: Released for public use.
- 
-See: http://www.gammon.com.au/forum/?id=11504
- 
- Example:
- 
- // set up Timer 1
- TCNT1 = 0;         // reset counter
- OCR1A =  999;       // compare A register value (1000 * clock speed)
- 
- // Mode 4: CTC, top = OCR1A
- Timer1::setMode (4, Timer1::PRESCALE_1, Timer1::CLEAR_A_ON_COMPARE);
- 
- TIFR1 |= bit (OCF1A);    // clear interrupt flag
- TIMSK1 = bit (OCIE1A);   // interrupt on Compare A Match  
- 
-*/
-
-#ifndef _TimerHelpers_h
-#define _TimerHelpers_h
-
-#if defined(ARDUINO) && ARDUINO >= 100
-  #include "Arduino.h"
-#else
-  #include "WProgram.h"
-#endif
-
-/* ---------------------------------------------------------------
- Timer 0 setup
- --------------------------------------------------------------- */
-
-namespace Timer0 
-{
-  // TCCR0A, TCCR0B
-  const byte Modes [8] [2] =   {
-  { 0,                         0 },            // 0: Normal, top = 0xFF
-  { bit (WGM00),               0 },            // 1: PWM, Phase-correct, top = 0xFF
-  {               bit (WGM01), 0 },            // 2: CTC, top = OCR0A
-  { bit (WGM00) | bit (WGM01), 0 },            // 3: Fast PWM, top = 0xFF
-  { 0,                         bit (WGM02) },  // 4: Reserved
-  { bit (WGM00),               bit (WGM02) },  // 5: PWM, Phase-correct, top = OCR0A
-  {               bit (WGM01), bit (WGM02) },  // 6: Reserved
-  { bit (WGM00) | bit (WGM01), bit (WGM02) },  // 7: Fast PWM, top = OCR0A
-  
-  };  // end of Timer0::Modes
-  
-  // Activation
-  // Note: T0 is pin 6, Arduino port: D4
-  enum { NO_CLOCK, PRESCALE_1, PRESCALE_8, PRESCALE_64, PRESCALE_256, PRESCALE_1024, T0_FALLING, T0_RISING };
-  
[…]

### 2019-04-24 matlab/a10M_quad_plot.m

@@ -1,42 +1,0 @@
-filename = 'c_____1.bin';
-[~,scope_y_data] = importAgilentBin(filename,1);
-[scope_x_data,scope_y_data2] = importAgilentBin(filename,2);
-[w,init_cross,final_cross,mid_level] = pulsewidth(scope_y_data,scope_x_data, 'Polarity', 'Positive');
-[w2,init_cross2,final_cross2,mid_level2] = pulsewidth(scope_y_data2,scope_x_data, 'Polarity', 'Positive');
-
-f_samp = 4E9;
-num_samp = length(scope_y_data);
-t_max = num_samp*(1/f_samp); 
-t=linspace(0,t_max,num_samp);
-
-figure
-hold on; 
-plot(t*1e6-7.95-0.09,scope_y_data+4,'b');
-plot(t*1e6-7.95-0.09,scope_y_data2,'r');
-
-c = scope_y_data2;
-
-filename = 'PA_10M_div4_1.bin';
-[~,scope_y_data] = importAgilentBin(filename,1);
-[scope_x_data,scope_y_data2] = importAgilentBin(filename,2);
-[w,init_cross,final_cross,mid_level] = pulsewidth(scope_y_data,scope_x_data, 'Polarity', 'Positive');
-[w2,init_cross2,final_cross2,mid_level2] = pulsewidth(scope_y_data2,scope_x_data, 'Polarity', 'Positive');
-
-t=linspace(0,t_max/2,num_samp);
-
-move_dist = 190;
-plot((t(move_dist:end)-move_dist*1/4E9)*1e6-7.95,c(move_dist:end)+8,'g'); %
-
-num_samp = length(scope_y_data);
-t_max = num_samp*(1/f_samp); 
-t=linspace(0,t_max/2,num_samp);
-
-move_dist = 178;
-plot((t(move_dist:end)-move_dist*1/4E9)*1e6-7.95,scope_y_data2(move_dist:end)+12,'k');
-
-ylabel('Amplitude')
-xlabel('Time (s)')
-xlim([80 81])
-xlabel("Time (\mu sec)")
-set(gca,'ytick',[])
-legend(["10M div 8", "ref div 8", "10M div 4", "ref div 4"])

### 2019-04-24 matlab/a1_3.m

@@ -1,231 +1,0 @@
-clearvars
-num_files = 15;
-std_period = zeros(4,num_files);
-TIE_mean = zeros(4,num_files);
-TIE_max = zeros(4,num_files);
-N           = 4; 
-step        = 2e6;
-width_vect  = 4e6;
-div_ratio = 4;
-config = "div" + num2str(div_ratio);
-config = "nodiv";
-
-for inc = 1:num_files
-    %filename = char("conor_test/settings1_" + num2str(inc) + ".bin");
-    %filename = char("scope_41.bin");
-        
-    path3       = "~/gdrive/measurements/tuesday_last/adpll_des3_2x2_" + num2str(inc) + ".bin";
-%     path3       = "~/gdrive/measurements/tuesday/net3x3_clocked_pd_5m18_" + num2str(inc) + ".bin";
-%     path3       = "~/gdrive/measurements/wednesday/net3x3_5m00_kp_100_ki_1_" + num2str(inc) + ".bin";
-%     path3       = "~/gdrive/measurements/monday/design2b_div" + num2str(div_ratio) + "_" + num2str(inc) + ".bin";    
-%     path3       = "~/gdrive/measurements/monday/design2b_" + num2str(inc) + ".bin";
-
-    disp(path3)
-    fileID      = fopen(path3);
-    A           = fread(fileID, 'double');
-    fclose(fileID);
-    ss = [];
-    for chan=0:N-1 %start at 0 for chan1
-        deb = chan*width_vect+1;
-        ss = [ss A(deb:deb+step-1, 1) A(deb+step:deb+2*step-1, 1)];
-    end
-
-    scope_x_data = ss(:,1);
-    scope_y_data1 = ss(:,2);
-    scope_y_data2 = ss(:,4);
-    scope_y_data3 = ss(:,6);
-    scope_y_data4 = ss(:,8);
-%     [~,scope_y_data1] = importAgilentBin(filename,1);
-%     [~,scope_y_data2] = importAgilentBin(filename,2);
-%     [~,scope_y_data3] = importAgilentBin(filename,3);
-%     [scope_x_data,scope_y_data4] = importAgilentBin(filename,4);
-    f_samp = 4E9;
-    num_samp = length(scope_y_data1);
-    t_max = num_samp*(1/f_samp); 
-    t=linspace(0,t_max,num_samp);
-
-    [~,init_cross1,~,~] = pulsewidth(scope_y_data1,scope_x_data, 'Polarity', 'Positive');
[…]

### 2019-04-24 matlab/cd___anal.m

@@ -1,31 +1,0 @@
-for inc = 1:10
-    filename = char("new/uhh" + num2str(inc) + ".bin");
-    [scope_x_data,scope_y_data] = importAgilentBin(filename,3);
-
-    f_samp = 4E9;
-    num_samp = length(scope_y_data);
-    t_max = num_samp*(1/f_samp); 
-    t=linspace(0,t_max,num_samp);
-
-    [~,init_cross4,~,~] = pulsewidth(scope_y_data,scope_x_data, 'Polarity', 'Positive');
-
-    periods4 = getPeriods(init_cross4);
-
-    std_period_4 = std(periods4);
-
-    max_4 = max(periods4-mean(periods4));
-    min_4 = min(periods4-mean(periods4));
-    p2p_4 = max_4-min_4;
-
-    clc
-    string = num2str(std_period_4) + ", " + num2str(max_4) + ", " + num2str(min_4) + ", " + num2str(p2p_4)
-end
-
-
-
-% figure
-% hold on;
-% histfit(periods4)
-% title('')
-% xlabel('')
-% ylabel('')

### 2019-04-24 matlab/cheating.m

@@ -1,34 +1,0 @@
-% two_a_plls = [1.6051e-09, 1.5924e-09, 1.6146e-09, 1.6251e-09, 1.619e-09, 1.621e-09,];
-% two_a_uni = [1.9819e-09, 2.0627e-09, 2.2021e-09, 1.9789e-09, 2.0005e-09, 2.0612e-09,];
-% two_a_bi = [2.0303e-09, 2.1097e-09, 2.1253e-09, 1.9946e-09, 2.0264e-09, 2.0212e-09,];
-% 
-% three_a_plls = [6.2499e-07, 7.5919e-07, 7.9049e-07,];
-% three_a_uni = [7.6822e-07, 9.009e-07,  1.1407e-06,];
-% three_a_bi = [7.0393e-07, 7.0201e-07, 1.0241e-06,];
-% 
-% three_b_plls = [4.1561e-09, 4.0044e-10, 4.1883e-09, 3.9882e-10, 4.1628e-09, 3.9888e-10, 4.1648e-09, 3.9396e-10,];
-% three_b_uni = [6.1918e-09, 5.2058e-09, 1.9428e-08, 6.2755e-09, 5.2531e-09, 1.9096e-08, 2.2232e-08,];
-% three_b_bi = [5.5243e-09, 7.1567e-09, 6.9324e-09, 5.4574e-09, 7.2364e-09, 7.4538e-09, 5.3828e-09, 7.1966e-09, 7.4565e-09, 5.8469e-09, 7.1654e-09, 6.9995e-09, 5.8569e-09, 7.1604e-09, 6.9636e-09, ];
-% 
-% mean(two_a_plls)
-% mean(two_a_uni)
-% mean(two_a_bi)
-% mean(three_a_plls)
-% mean(three_a_uni)
-% mean(three_a_bi)
-% mean(three_b_plls)
-% mean(three_b_uni)
-% mean(three_b_bi)
-
-mean([6.9324,
-7.4538,
-7.4565,
-6.9995,
-6.9636,
-
-
-
-
-  
-])
-

### 2019-04-24 matlab/lf.m

@@ -1,27 +1,0 @@
-% loop filter cutoff
-ki = 1;
-ki_frac_bits = 8;
-kp = 1;
-kp_frac_bits = 4;
-div_at_exit = 2;
-target = 5E6;
-loop_div = 4;
-
-b0 = kp/(div_at_exit*2^kp_frac_bits);
-b1 = ki/(div_at_exit*2^ki_frac_bits);
-
-TS = 1/(target/loop_div);
-N = 2*[(2*b0 + TS*b1)  (-2*b0 + TS*b1)];
-D = [1 -1];
-
-PI_tf = tf(N, D, TS);
-
-bode(PI_tf);
-
-%% error code to freq
-pa_freq_sens = 1*b0*(258E6*(1/2^12));
-
-inv_num = 383;
-T_target = 1/target;
-
-ro_freq_sens = 1*b0*(abs(1/(T_target - 2*(T_target/inv_num))-target) + abs(1/(T_target + 2*(T_target/inv_num))-target))/2;

### 2019-04-24 matlab/plot10points.m

@@ -1,5 +1,0 @@
-% plot10points.m 
-% plot and list 10 points
-
-[x,y]=importAgilentBin('points10.bin',1)
-plot(x,y)

### 2019-04-24 matlab/plot_4_waves.m

@@ -1,17 +1,0 @@
-filename = 'scope_37.bin';
-[~,scope_y_data] = importAgilentBin(filename,1);
-[~,scope_y_data2] = importAgilentBin(filename,2);
-[~,scope_y_data3] = importAgilentBin(filename,3);
-[scope_x_data,scope_y_data4] = importAgilentBin(filename,4);
-
-f_samp = 4E9;
-num_samp = length(scope_y_data);
-t_max = num_samp*(1/f_samp); 
-t=linspace(0,t_max,num_samp);
-
-figure
-hold on; 
-plot(t*1e6-7.95-0.09,scope_y_data+15,'b');
-plot(t*1e6-7.95-0.09,scope_y_data2+10,'r');
-plot(t*1e6-7.95-0.09,scope_y_data3+5,'g');
-plot(t*1e6-7.95-0.09,scope_y_data4+0,'k');

### 2019-10-14 conf_paper/paper.tex
supposedly chanaged

@@
 
-\title{FPGA Based Modelling of an ADPLL Network\\}
-
-\author{\IEEEauthorblockN{C. Dooley$^1$, E. Blokhina$^1$, B. Mulkeen$^1$, D. Galayko$^2$}
+\title{FPGA Based Modelling of an ADPLL Network\vspace{-0.3cm}}
+\author{
+	\IEEEauthorblockN{C. Dooley$^1$, E. Blokhina$^1$, B. Mulkeen$^1$, D. Galayko$^2$}
     \IEEEauthorblockA{\textit{$^1$School of Electrical \& Electronic Engineering, University College Dublin}\\
-    \textit{$^2$LIP6 Lab, Sorbonne Université}
-    	\vspace{-0.8cm}
+    \textit{$^2$LIP6 Lab, Sorbonne Universit\'e}
+    	\vspace{-0.7cm}
 	}
 }
+\IEEEoverridecommandlockouts
+\IEEEpubid{\makebox[\columnwidth]{978-1-7281-1201-5/19/\$31.00~\copyright2019 IEEE \hfill} \hspace{\columnsep}\makebox[\columnwidth]{ }}
 \maketitle
+
 \begin{abstract}
     This paper introduces and compares the implementation of a number of FPGA based ADPLL network prototyping architectures.
-Comparison is then made between the expected performance and role of each method as a development tool.
+Networks are then created using three different ADPLL implementations and tests performed on each.
+Based on these test results, comparison is made to both the expected performance and role of each ADPLL design as a development tool.
 \end{abstract}
@@
 The technical motivation underlying this paper is the creation of an Field Programmable Gate Array (FPGA) based prototyping platform for All Digital Phase Lock Loop (ADPLL) networks.
-As the creation of Application Specific Integrated Circuits (ASICs) is an expensive and time consuming process, and mistakes made in the design may require the manufacture of a new chip, it is important to ensure that any errors made in the design have been located and addressed before manufacture..
+As the creation of Application Specific Integrated Circuits (ASICs) is an expensive and time consuming process, and mistakes made in the design may require the manufacture of a new chip, it is important to ensure that any errors made in the design have been located and addressed before manufacture.
 Simulations either at theoretical or gate/transistor levels have global usage in minimising such errors due to the ubiquity of simulators and their ease of use.\\
@@
 Subsequently Javiden et.
-al
-\cite{javidan2011all} proposed the implementation of such a system using All Digital PLLs, which better suited the use cases of the technology and avoided some of the flaws pointed out in the original paper.
+al \cite{javidan2011all} proposed the implementation of such a system using All Digital PLLs, which better suited the use cases of the technology and avoided some of the flaws pointed out in the original paper.
 Zianbetov and Shan followed on from this proposition by implementing such a network on a custom IC \cite{zianbetov2013phd,shan2014phd}.\\
@@
 In combination with the bit width of the control signal, the minimum frequency step limits the range over which the oscillator can be tuned.
-Frequently the output of the DCO is divided down to allow the control blocks in the feedback path that are clocked on this signal, to run at a lower and thus easier frequency.
-\subsection{Loop Filter}
+Frequently the output of the DCO is divided down to allow the control blocks in the feedback path that are clocked on this signal, to run at a lower and thus easier frequency.\vspace{-0.2cm}
+\subsection{ Loop Filter}
 The loop filter is a key aspect of a phase lock loop as without it only the current output of the phase detector could be used to compute the control voltage of the VCO.
-In an ADPLL it performs an identical role however, the digital environment results in the traditional loop filter being replaced by a discrete time Proportional Integral (PI) controller, with the integral path of the controller implemented by an accumulator.
+In an ADPLL it performs an identical role however, the digital environment results in the traditional loop filter being replaced by a discrete time Proportional Integral (PI) controller, with the integral path of the controller implemented by an accumulator.\vspace{-0.2cm}
 \subsection{Error Combiner}
@@
 The ability to change the weights during operation enables ``uni-directional start-up'', which Javiden et.
-al \cite{javidan2011all} found prevented the system from entering unwanted stable equilibriums warned of in Pratt and Nguyen's seminal paper.
+al \cite{javidan2011all} found prevented the system from entering unwanted stable equilibriums warned of in Pratt and Nguyen's seminal paper.\vspace{-0.2cm}
 \subsection{Phase Detector}
@@
         \end{tiny}
-        \caption{}
+        \caption{ADPLL Performance Comparison.}
         \label{table:b_c_perf}
@@
 \begin{figure}[h]%
+	\vspace{-0.8cm}
     \centering
@@
     \caption{Example Distribution of Periods.\label{fig:dists}}    
-    \vspace{-0.5cm}
+    \vspace{-0.4cm}
 \end{figure}
